### 2024-04-25 WGiT-Proposal.qmd
Initial commit

---
title: "Towards caring 15-minute neighbourhoods"
author:
  - name: Anastasia Soukhov 
    email: soukhoa@mcmaster.ca
    affiliations:
        - name: McMaster University
          department: School of Earth, Environment & Society
          city: Hamilton, Canada
    attributes:
        corresponding: true
  - name: Léa Ravensbergen
    email: ravensbl@mcmaster.ca
    affiliations:
        - name: McMaster University
          department: School of Earth, Environment & Society
          city: Hamilton, Canada
  - name: Lucía Mejía Dorantes
    email: lucia.mejia.dorantes@gmail.com
    affiliations:
        - name: Consultant
          city: Karlsruhe, Germany
  - name: Antonio Páez
    email: paezha@mcmaster.ca
    affiliations:
        - name: McMaster University
          department: School of Earth, Environment & Society
          city: Hamilton, Canada
format: 
  #docx
  elsevier-pdf:
    keep-tex: true
    include-in-header:
      text: |
        \usepackage{graphicx}
        \usepackage{unicode-math}
        \usepackage{hyperref}
        \def\tightlist{}
        \usepackage{setspace}
        \doublespacing
        \usepackage{lineno}
        \linenumbers
    journal:
      name: 7th International Conference on Women and Gender in Transportation (WGiT)
      formatting: preprint
      model: 3p
      cite-style: authoryear
keywords: ["Accessibility", "Mobility of Care", "Gender", "15-Minute City", "Care Complete Neighbourhoods"]
bibliography: bibliography.bib
---

The 15-Minute City is a normative conceptualisation gaining ground in urban planning: it frames neighbourhoods as responsive to human needs and environmental sensibilities, where most daily necessities can be reached within a 15-minute walk or bike ride [@Allam2022]. As a related tool, accessibility measures (the ease of reaching opportunities) are increasingly important amongst transport planners aiming to foster just and sustainable cities [@vale2023accessibility]. Both the 15-Minute City and accessibility measures are flexible enough to consider all destination types holistically however, gendered examinations have been lacking in the literature. For instance, accessibility analyses have historically focused on employment-centric and discretionary travel, types of travel more frequent for working-age and higher-income men.

To counter this masculinist bias, this study investigates a way to gender-mainstream the 15-Minute City through a care lens [e.g., @lawWomenTransportNew1999; @uteng2008gendered; @Oestergaard1992; @levy1991towards; @levy2013travel] supported by the Mobility of Care framework (coined by @sanchezdemadariagaMobilityCareIntroducing2013). Mobility of Care emphasizes the importance of travel to unpaid work (care trips) in contrast to the better-studied travel to employment and leisure. While all three trip types (work, care, and leisure) are essential, care trips are often relatively shorter-distance, proximate to residential/work/school, and comprise approximately 30% of adults’ daily trips [@sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory; @mejia2021mobility]; fitting well within the 15-Minute City conceptualisation.

Our study provides an empirical example that maps the 15-Minute City onto the Mobility of Care framework. Specifically, it identifies which areas in Hamilton, Canada are 'caring 15-minute neighbourhoods'. To do so, a database of care destinations is created using secondary data. In this database, care destinations include all places associated with sustaining household tasks needed for the reproduction of life namely: shopping (e.g., groceries), errands (e.g., libraries), health (e.g., dentist), and caring for dependents (e.g., schools). This database is used to estimate the number and mix of care-destinations that can be reached within a 15-minute walk- and cycling- sheds from census centroids. Typologies are generated to illustrate which neighbourhoods can and to what degree facilitate 15-minute access to care.

Results indicate only a few neighbourhoods outside of the downtown core are 'care-complete', i.e., contain a sufficiently high mix of care destinations from all categories and sub-categories. However, some neighbourhoods are almost 'care-complete' and provide 15-minute access to some care categories. Our study frames these neighbourhoods on the continuum of 'caring' and in need of further intervention. The quantitative investigation conducted provides a high-level picture of what neighbourhoods (and their underlying land-use) are connected to transport infrastructure that can support reaching care-destinations.

Taken together, this study provides a theoretical bridge to connect 15-Minute Cities, accessibility analysis and Mobility of Care framework for the purpose of informing policy choice aimed to encourage just and sustainable mobility.

# References

### 2024-04-25 data/data-prep2.qmd
Initial commit

---
title: "data-prep2"
format: html
---
```{r}
rm(list = ls())
```

Let's work with the TTM that we already have ! Only destinations within 15 minutes or less, for walking mode.
```{r library-setup}
library(biscale)
library(cowplot)
library(sf)
library(here)
library(dplyr)
library(tmap)
library(mapview)
library(tmaptools)
library(pander)
library(flextable)
library(ftExtra)
library(scales)
```

```{r spatial-availability-function}
sp_avail_detailed <- function(x, o_id, d_id, pop, opp, r, f, alpha = 1){

  o_id <- rlang::enquo(o_id)
  d_id <- rlang::enquo(d_id)
  pop <- rlang::enquo(pop)
  opp <- rlang::enquo(opp)
  r <- rlang::enquo(r)
  f <- rlang::enquo(f)

  sum_pop <- x |>
    dplyr::distinct(!!o_id,
                    .keep_all = TRUE) |>
    dplyr::mutate(sum_pop = !!r*(!!pop)^alpha) |>
    dplyr::pull(sum_pop) |>
    sum()

  f_p <- dplyr::pull(x, !!r) * dplyr::pull(x, !!pop)^alpha / sum_pop

  sum_impedance <- x |>
    dplyr::group_by(!!d_id) |>
    dplyr::summarize(sum_impedance = sum(!!f))

  x <- x |>
    dplyr::left_join(sum_impedance,
                     by = rlang::as_name(d_id))

  f_c <- dplyr::pull(x, !!f) / x$sum_impedance

  x$f_c <- f_c
  x$f_p <- f_p

  sum_pa <- x |>
    dplyr::group_by(!!d_id) |>
    dplyr::summarize(sum_pa= sum(f_p * f_c))

  x <- x |>
    dplyr::left_join(sum_pa,
                     by = rlang::as_name(d_id))
  x$f_t <- (f_p * f_c) / dplyr::pull(x, sum_pa)

  x |>
    dplyr::mutate(V_ij = !!opp * f_t)
}
```

```{r}
#boundaries for mapping
Community_Boundaries <- st_read("data-raw/Community_Boundaries.shp") |> mutate(COMMUNITY_ = ifelse(COMMUNITY_ == "Hamilton", "Hamilton-Central", COMMUNITY_))|> st_transform(crs=4326)
City_Boundary <- st_read("data-raw/City_Boundary.shp") |> st_transform(crs=4326)
hydro_p_LakeOntario <- st_read("data-raw/hydro_p_LakeOntario.shp") |> st_transform(crs=4326)
ham_bay <- st_read("data-raw/Waterbodies.shp") |> st_transform(crs=4326) |> filter(FEATURE_TY == "Lake")
ham_bay_cropped<-st_crop(ham_bay$geometry, Community_Boundaries$geometry)
hydro_p_LakeOntario_cropped<-st_crop(hydro_p_LakeOntario$geometry, Community_Boundaries$geometry)

load("care_dest.rda")
load("ttm_care_walk.rda")
load("HAM_census_21.rda")
```

```{r}
HAM_census_21 <- HAM_census_21 |>
  rename("LICO_AT_TOTAL" = `v_CA21_1085: Prevalence of low income based on the Low-income cut-offs, after tax (LICO-AT) (%)`) |> 
  mutate(LICO_AT_TOTAL = ifelse(is.na(LICO_AT_TOTAL), 0, LICO_AT_TOTAL)) |>
  rowwise() |> mutate(
         perc_walked = (`v_CA21_7647: Walked`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`, `v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE),
         perc_car = (`v_CA21_7635: Car, truck or van`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`, `v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE),
perc_transit = (`v_CA21_7644: Public transit`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`, `v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE),
 perc_cycle_other = (`v_CA21_7650: Bicycle`+ `v_CA21_7653: Other method`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`, `v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE),
pop_den = Population/`Shape Area`) |>
  select(c("GeoUID", "Population", "LICO_AT_TOTAL", "perc_walked", "perc_car", "perc_transit", "perc_cycle_other", "pop_den"))

HAM_census_21 <- st_join(HAM_census_21, Community_Boundaries|> select("COMMUNITY_"), largest=TRUE)
DA_centroids <- HAM_census_21 |> st_centroid() 
```

```{r reformat care-dest specific categories}
care_dest <- care_dest |> 
  mutate(Care_Category_Specific = case_when(
    Care_Category_Specific  == "Daycare_EarlyON" ~ "Daycare or EarlyON",
    Care_Category_Specific == "Community_Centre" ~ "Community or Recreation Centre",
    Care_Category_Specific == "Daycare_Centre" ~ "Daycare or EarlyON",
    Care_Category_Specific == "Recreation_Centre" ~ "Community or Recreation Centre",
    Care_Category_Specific == "School" ~ "School",
    Care_Category_Specific == "Long-Term_Care_Home" ~ "Long-Term Care or Retirement Home",
    Care_Category_Specific == "Retirement_Home" ~ "Long-Term Care or Retirement Home",
    Care_Category_Specific == "Senior_Centre" ~ "Senior Centre",
    Care_Category_Specific == "Seniors_Active_Living_Centre" ~ "Senior Centre",
    Care_Category_Specific == "Post_Office" ~ "Post Office",
    Care_Category_Specific == "Library" ~ "Library",
    Care_Category_Specific == "Banks" ~ "Bank or ATM",
    Care_Category_Specific == "Grocer_ConvenienceStore" ~ "Convenience Store",
    Care_Category_Specific == "Clinics" ~ "Hospital or Clinic",
    Care_Category_Specific == "Grocer_Retail" ~ "Grocer",
    Care_Category_Specific == "Dentist" ~ "Dentist",
    Care_Category_Specific == "Hosptial" ~ "Hospital or Clinic",
    Care_Category_Specific == "Hospital" ~ "Hospital or Clinic",
    Care_Category_Specific == "Pharmacy" ~ "Pharmacy",
    Care_Category_Specific == "Park" ~ "Park"),
  Care_Category = case_when(
    Care_Category == "Child_Centric" ~ "Dependent-centric",
    Care_Category == "Elder_Centric" ~ "Dependent-centric",
    Care_Category == "Errand_Centric" ~ "Errand-centric",
    Care_Category == "Grocery_Centric" ~ "Grocery-centric",
    Care_Category == "Health_Centric" ~ "Health-centric"))
```

```{r travel-time-15min}
#15 mins or less
ttm_care_walk_15 <- ttm_care_walk |> 
  filter(travel_time_p50 <= 15) |> select(-c("travel_time_p25", "travel_time_p75")) |> 
  left_join(
  care_dest |> st_drop_geometry() |> select(c("ID", "Care_Category", "Care_Category_Specific", "Care_Category_Speciality")) |>
    mutate(ID = as.character(ID)),
  by = c("to_id" = "ID")) |>
  mutate(f = 1)

ttm_care_walk_15 <- ttm_care_walk_15 |>
  left_join(HAM_census_21 |> st_drop_geometry() |> select(c("GeoUID", "Population", "LICO_AT_TOTAL")), by=c("from_id" = "GeoUID")) 
```

```{r}
access_15 <- ttm_care_walk_15 |> group_by(from_id,Care_Category, Care_Category_Specific) |> summarize(copp_isc = n())

test_all <- ttm_care_walk_15 |>
  sp_avail_detailed(o_id = from_id,
                    d_id = to_id, 
                    pop = Population, 
                    opp = f,
                    r = f, 
                    f = f, alpha = 1) |> group_by(from_id, Care_Category, Care_Category_Specific) |> summarize(V_isc = sum(V_ij, na.rm = TRUE))

access_15 <- access_15 |> left_join(test_all, by=c("from_id", "Care_Category", "Care_Category_Specific")) |> ungroup()
rm(test_all)
access_15$V_isc |> sum(na.rm = TRUE) #sums to 1997 opportunities -- out of 2225 total opportunities
```

```{r}
# all speciality categories should be represented, i.e., each ID should have 14 rows -- if that category is empty it should be 0. Let's add those 0s. 
access_15_zeros <- 
  data.frame(from_id = rep(HAM_census_21$GeoUID, length(access_15$Care_Category_Specific |> unique())),
           Care_Category = rep(c("Dependent-centric", "Dependent-centric", "Grocery-centric", "Dependent-centric", 
                                 "Dependent-centric", "Dependent-centric",
                                 "Errand-centric", "Errand-centric", 
                                 "Grocery-centric","Health-centric",
                                 "Health-centric", "Errand-centric",
                                 "Dependent-centric", "Health-centric"), length(HAM_census_21$GeoUID)),

           Care_Category_Specific = rep(c("Park", "Daycare or EarlyON", "Convenience Store", "School",
                                          "Long-Term Care or Retirement Home", "Senior Centre",
                                          "Bank or ATM", "Library", 
                                          "Grocer", "Dentist", 
                                          "Pharmacy", "Post Office",
                                          "Community or Recreation Centre", "Hospital or Clinic"), length(HAM_census_21$GeoUID)),

           copp_isc = rep(0, length(HAM_census_21$GeoUID)*length(c("Park", "Daycare or EarlyON", "Convenience Store", "School",
                                          "Long-Term Care or Retirement Home", "Senior Centre",
                                          "Bank or ATM", "Library", 
                                          "Grocer", "Dentist", 
                                          "Pharmacy", "Post Office",
                                          "Community or Recreation Centre", "Hospital or Clinic"))),

           V_isc = rep(0, length(HAM_census_21$GeoUID)*length(c("Park", "Daycare or EarlyON", "Convenience Store", "School",
                                          "Long-Term Care or Retirement Home", "Senior Centre",
                                          "Bank or ATM", "Library", 
                                          "Grocer", "Dentist", 
                                          "Pharmacy", "Post Office",
                                          "Community or Recreation Centre", "Hospital or Clinic"))))

access_15 <- bind_rows(access_15_zeros, access_15) |>
                         group_by(from_id, Care_Category, Care_Category_Specific) |> 
  summarize(copp_isc = sum(copp_isc),
            V_isc = sum(V_isc))

access_15_bycategory <- access_15 |> group_by(from_id, Care_Category)  |> 
  summarize(copp_ic = sum(copp_isc, na.rm=T),
            V_ic = sum(V_isc, na.rm = T))

access_15 <- access_15 |>
  left_join(HAM_census_21 |> st_drop_geometry() |> select(c("GeoUID", "Population", "LICO_AT_TOTAL", "COMMUNITY_")), 
            by=c("from_id" = "GeoUID")) |> 
  mutate(copp_isc_perpop = copp_isc/Population,
         v_isc = V_isc/Population) |> ungroup()

access_15_bycategory <- access_15_bycategory |>
  left_join(HAM_census_21 |> st_drop_geometry() |> select(c("GeoUID", "Population", "LICO_AT_TOTAL", "COMMUNITY_")), 
            by=c("from_id" = "GeoUID")) |> 
  mutate(copp_ic_perpop = copp_ic/Population,
         v_ic = V_ic/Population) |> ungroup()

rm(access_15_zeros)
```
```{r}
# access_15
access_15 |> select(c(Care_Category, Care_Category_Specific)) |> unique()
# access_15 |> filter(Care_Category == c("Errand-centric", "Grocery-centric","Health-centric"))
```

```{r}
no_dependents <- access_15_bycategory |> filter(Care_Category == "Errand-centric"|
                                                 Care_Category == "Grocery-centric" |
                                                  Care_Category == "Health-centric") |> 
  group_by(from_id) |>
  summarise(Care_Access_Status = case_when(all(copp_ic == 0) ~ "Not caring",
                                 all(copp_ic >= 1 & copp_ic <= 2) ~ "Caring", 
                                 all(copp_ic >= 3) ~ "Very caring"),
            copp_isc = sum(copp_ic), 
            V_isc = sum(V_ic),
            Population = first(Population),
            COMMUNITY_ = first(COMMUNITY_))

no_dependents2 <- access_15 |> filter(Care_Category == "Errand-centric"|
                                                 Care_Category == "Grocery-centric" |
                                                  Care_Category == "Health-centric") |> 
  group_by(from_id) |>
  summarise(Care_Access_Status2 = case_when(all(copp_isc >= 1 & copp_isc < 2) ~ "Completely caring", 
                                 all(copp_isc >= 2) ~ "Very completely caring"))

no_dependents <- no_dependents |> left_join(no_dependents2, by= c("from_id"))

```

```{r}
no_dependents
```

<!-- NOTES FROM LEA CONVO FEB 22:

We changed some category tags. 
 - child and elder centric are now both 'dependent-centric'. These are 'caring for others'. We also consolidated some sub-categories. Daycare and EarlyON are now combined.. Long Term-Care Home and Residential Home are now combined.. Senior Centre and Senior Active Living Centre are combined. Hospital and Clinic are combined. Recreation Centre and COmmunity Centre are now combined.
 - A consideration for the future * maybe convenience stores and grocery stores can be be modulated based on their time -- i.e., if a DA trip to convenience is less than grocery, it is counted, if it is more than grocery it is not. More thought needs to be put into this, for now I did not include this consideration. 
 - We will consider the following 2 (or maybe 3?) faceted (by mode) map with scales:

1.	No-children (oneself and partner, no dependents) city map (15 min walk + 15 min cycle):
-	Errand, grocery, health

-	Scale: 
o	not-caring (does not have 1 of each category), 
o	caring (1 of each category), 
o	very caring (2 or more of each category), 
o	completely caring (1 of each sub-category), 
o	very completely caring (2 of each sub-category)

2.	Children+seniors: -- the truly caring city has all 4 categories. You always preform care but not always care for others. Your status will change. 
-	+elder+children
-	Scale: (same)

3. What neighbourhoods are ‘almost there’ for being completely caring? What are they missing?

And our presentation's angle: a 15 minute city is XYZ, the destinations are contested but typically are 'care' trips as well as discretionary trips (fun stuff). The mobility of care framework: paid, unpaid, and discretionary. All is important. But unpaid travel is under-studied. We are going to explore unpaid trips: they are essential and short-distance trips. A masculine city is one where paid work + discretionary trips are planned for. This analysis presents a tool to conceptual what is and is not a 'caring' or 'caring complete' neighbourhood. presents some visualisations.

APRIL 25 2024
- for simplicity, 15 min walk thresholds as a baseline for 15-min caring neighbourhoods. 15 min cycle thresholds to demonstrate the potential of cyling to get those 'almost there' neighbourhoods to be more caring if cycling infrastructure is implemented. Transit/bike to care isn't so common currently, but it's ane equity argument and shows political potential.

-->
<!-- THINK FROM ABOVE HERE-->

```{r message=FALSE}
all_cats <- access_15 |> group_by(from_id) |>
  summarise(Complete = ifelse(all(c_opp > 0 ), 1, 0),
            c_opp = sum(c_opp), 
            V_ic = sum(V_ic),
            Population = first(Population),
            COMMUNITY_ = first(COMMUNITY_))

all_cats |> group_by(COMMUNITY_) |> 
  summarize(DApop_served_all_categories = sum(Population*Complete, na.rm=T),
            DApopperc_served_all_categories = sum(Population*Complete, na.rm=T)/sum(Population))

one_cat <- access_15 |> group_by(from_id, Care_Category)|>
  summarise(Complete = ifelse(all(c_opp > 0 ), 1, 0),
            c_opp = sum(c_opp), 
            V_ic = sum(V_ic),
            Population = first(Population),
            COMMUNITY_ = first(COMMUNITY_))

one_cat_communities <- one_cat |> group_by(COMMUNITY_, Care_Category) |> 
  summarize(DApop_served_1_category = sum(Population*Complete, na.rm=T),
            DApopperc_served_1_category = sum(Population*Complete, na.rm=T)/sum(Population))

one_cat_communities |> filter(COMMUNITY_ == "Hamilton-Central")
one_cat_communities |> filter(Care_Category == "Grocery_Centric") 
```
```{r}
all_cats <- left_join(HAM_census_21 |> select(c(GeoUID)), all_cats, by=c( "GeoUID" = "from_id"))
```
```{r message=FALSE}
tmap_mode("view")
tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("c_opp",
              palette = "Greens",
              border.alpha = 0.2,
              title = "Number of destinations spatially accessible within 15 mins.",
              legend.is.portrait = FALSE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 20)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12,
            legend.title.size = 0.7,
            legend.text.size = 0.6)
```

```{r message=FALSE}
tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("V_ic",
              palette = "Purples",
              border.alpha = 0,
              title = "Spatially available destinations to mode-using population within 15 mins.",
              legend.is.portrait = FALSE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 20)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12,
            legend.title.size = 0.7,
            legend.text.size = 0.6)

```
```{r}
one_cat <- left_join(HAM_census_21 |> select(c(GeoUID)), one_cat, by=c( "GeoUID" = "from_id"))
```

```{r}
tmap_mode("plot")
tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(one_cat) +
  tm_polygons("Complete",
              border.alpha = 0,
              legend.is.portrait = FALSE,
              n=2,
              labels = c("Not complete", "Complete")) +
  tm_facets(by=c("Care_Category"), nrow=2, showNA = FALSE)+
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 20)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12,
            legend.title.size = 0.7,
            legend.text.size = 0.6)
```

Questions:

- what is a reasonable travel time? Is 15 minute walk indeed fine for our purpose?

- is 1 potential opportunity enough per subcategory?

- Should some subcategories be substitutes? E.g., 1 daycare OR EARLYON centre is enough, not 1 of each.

- Are there ways the categories can overlap to tell a story? Maybe to include google maps images. eg. 
  - communities where children+errand+grocery centric care is easier -- especially for women
  - health_centric+Elder_centric -- people with higher health care needs

- care trips 30% of all trips made. they are essential and have

<!-- ```{r} -->
<!-- test1 <- ttm_care_walk_15 |> filter(Care_Category == "Child_Centric") |> -->
<!--   sp_avail_detailed(o_id = from_id, -->
<!--                     d_id = to_id,  -->
<!--                     pop = Population,  -->
<!--                     opp = f, -->
<!--                     r = f,  -->
<!--                     f = f, alpha = 1)|> -->
<!--   group_by(from_id, Care_Category) |> summarize(V_ic = sum(V_ij, na.rm = TRUE))  -->

<!-- test2 <- ttm_care_walk_15 |> filter(Care_Category == "Grocery_Centric") |> -->
<!--   sp_avail_detailed(o_id = from_id, -->
<!--                     d_id = to_id,  -->
<!--                     pop = Population,  -->
<!--                     opp = f, -->
<!--                     r = f,  -->
<!--                     f = f, alpha = 1)|> -->
<!--   group_by(from_id, Care_Category) |> summarize(V_ic = sum(V_ij, na.rm = TRUE))  -->

<!-- test3 <- ttm_care_walk_15 |> filter(Care_Category == "Elder_Centric") |> -->
<!--   sp_avail_detailed(o_id = from_id, -->
<!--                     d_id = to_id,  -->
<!--                     pop = Population,  -->
<!--                     opp = f, -->
<!--                     r = f,  -->
<!--                     f = f, alpha = 1)|> -->
<!--   group_by(from_id, Care_Category) |> summarize(V_ic = sum(V_ij, na.rm = TRUE))  -->

<!--  test4 <- ttm_care_walk_15 |> filter(Care_Category == "Errand_Centric") |> -->
<!--   sp_avail_detailed(o_id = from_id, -->
<!--                     d_id = to_id,  -->
<!--                     pop = Population,  -->
<!--                     opp = f, -->
<!--                     r = f,  -->
<!--                     f = f, alpha = 1)|> -->
<!--   group_by(from_id, Care_Category) |> summarize(V_ic = sum(V_ij, na.rm = TRUE))  -->

<!-- test5 <- ttm_care_walk_15 |> filter(Care_Category == "Health_Centric") |> -->
<!--   sp_avail_detailed(o_id = from_id, -->
<!--                     d_id = to_id,  -->
<!--                     pop = Population,  -->
<!--                     opp = f, -->
<!--                     r = f,  -->
<!--                     f = f, alpha = 1)|> -->
<!--   group_by(from_id, Care_Category) |> summarize(V_ic = sum(V_ij, na.rm = TRUE))  -->

<!-- test <- rbind(test1, test2, test3, test4, test5) -->

<!-- access_15 <- access_15 |> left_join(test, by=c("from_id", "Care_Category")) |> ungroup() -->

<!-- rm(test, test1, test2, test3, test4, test5) -->
<!-- ``` -->

### 2024-05-16 Manuscript.qmd
Added Lea and Lucia's introduction text. Added preliminary results and methods section. Draft of deliverables 1.

---
title: "Towards caring 15-minute neighbourhoods"
author:
  - name: Anastasia Soukhov 
    email: soukhoa@mcmaster.ca
    affiliations:
        - name: McMaster University
          department: School of Earth, Environment & Society
          city: Hamilton, Canada
    attributes:
        corresponding: true
  - name: Léa Ravensbergen
    email: ravensbl@mcmaster.ca
    affiliations:
        - name: McMaster University
          department: School of Earth, Environment & Society
          city: Hamilton, Canada
  - name: Lucía Mejía Dorantes
    email: lucia.mejia.dorantes@gmail.com
    affiliations:
        - name: Consultant
          city: Karlsruhe, Germany
  - name: Antonio Páez
    email: paezha@mcmaster.ca
    affiliations:
        - name: McMaster University
          department: School of Earth, Environment & Society
          city: Hamilton, Canada
format: 
  docx
  # elsevier-pdf:
  #   keep-tex: true
  #   include-in-header:
  #     text: |
  #       \usepackage{graphicx}
  #       \usepackage{unicode-math}
  #       \usepackage{hyperref}
  #       \def\tightlist{}
  #       \usepackage{setspace}
  #       \doublespacing
  #       \usepackage{lineno}
  #       \linenumbers
  #   journal:
  #     name: 7th International Conference on Women and Gender in Transportation (WGiT)
  #     formatting: preprint
  #     model: 3p
  #     cite-style: authoryear
keywords: ["Accessibility", "Mobility of Care", "Gender", "15-Minute City", "Care Complete Neighbourhoods"]
bibliography: bibliography.bib
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      cache=FALSE)
```

```{r}
#function for creating comments
word_comment <- function(comment, highlight = "") {
  if (isTRUE(knitr:::pandoc_to() == "docx")) {
    paste0('[', comment, ']{.comment-start id="0" author="Anastasia"',
           'date="2024-05-16T15:00:00Z"}', highlight, '[]{.comment-end id="0"}')
  }
}
```

```{r library-setup}
library(biscale)
library(cowplot)
library(sf)
library(here)
library(dplyr)
library(tmap)
library(mapview)
library(tmaptools)
library(pander)
library(flextable)
library(ftExtra)
library(scales)
```

```{r}
load("data/care_dest.rda")
```

```{r reformat care-dest specific categories}
care_dest <- care_dest |> 
  mutate(Care_Category_Specific = case_when(
    Care_Category_Specific  == "Daycare_EarlyON" ~ "Daycare or EarlyON",
    Care_Category_Specific == "Community_Centre" ~ "Community or Recreation Centre",
    Care_Category_Specific == "Daycare_Centre" ~ "Daycare or EarlyON",
    Care_Category_Specific == "Recreation_Centre" ~ "Community or Recreation Centre",
    Care_Category_Specific == "School" ~ "School",
    Care_Category_Specific == "Long-Term_Care_Home" ~ "Long-Term Care or Retirement Home",
    Care_Category_Specific == "Retirement_Home" ~ "Long-Term Care or Retirement Home",
    Care_Category_Specific == "Senior_Centre" ~ "Senior Centre",
    Care_Category_Specific == "Seniors_Active_Living_Centre" ~ "Senior Centre",
    Care_Category_Specific == "Post_Office" ~ "Post Office",
    Care_Category_Specific == "Library" ~ "Library",
    Care_Category_Specific == "Banks" ~ "Bank or ATM",
    Care_Category_Specific == "Grocer_ConvenienceStore" ~ "Convenience Store",
    Care_Category_Specific == "Clinics" ~ "Hospital or Clinic",
    Care_Category_Specific == "Grocer_Retail" ~ "Grocer",
    Care_Category_Specific == "Dentist" ~ "Dentist",
    Care_Category_Specific == "Hosptial" ~ "Hospital or Clinic",
    Care_Category_Specific == "Hospital" ~ "Hospital or Clinic",
    Care_Category_Specific == "Pharmacy" ~ "Pharmacy",
    Care_Category_Specific == "Park" ~ "Park"),
  Care_Category = case_when(
    Care_Category == "Child_Centric" ~ "Dependent-centric",
    Care_Category == "Elder_Centric" ~ "Dependent-centric",
    Care_Category == "Errand_Centric" ~ "Errand-centric",
    Care_Category == "Grocery_Centric" ~ "Grocery-centric",
    Care_Category == "Health_Centric" ~ "Health-centric"))

care_dest <- care_dest |> rename("care_ID" = "ID") |> mutate(care_ID = as.character(care_ID)) 
# care_dest|> st_drop_geometry() |> group_by(Care_Category) |> summarise(count = n())
```

```{r re-weighting points so all destinations are considered spatially equal in access counts, include=FALSE}
Dependent_Centric <- 0.25*2225/1265
Errand_Centric <- 0.25*2225/158
Grocery_Centric <- 0.25*2225/381
Health_Centric <- 0.25*2225/421

#adds up to 2225 opportunities  
0.25*2225/1265*(1265) +  0.25*2225/158*(158) + 0.25*2225/381*(381) + 0.25*2225/421*(421)

#adding the weighted care destination counts -- reweighted such that the 4 categories each make up 25% of the 2225 points.
care_dest <- care_dest |>
          mutate(Dest_count_weighted = case_when(Care_Category == "Dependent-centric" ~ Dependent_Centric,
                                                 Care_Category == "Errand-centric" ~ Errand_Centric,
                                                 Care_Category == "Grocery-centric" ~ Grocery_Centric,
                                                 Care_Category == "Health-centric" ~ Health_Centric))

care_dest$Dest_count_weighted |> sum()
```

# Abstract

The 15-Minute City is a normative conceptualisation gaining ground in urban planning: it frames neighbourhoods as responsive to human needs and environmental sensibilities, where most daily necessities can be reached within a 15-minute walk or bike ride [@Allam2022]. As a related tool, accessibility measures (the ease of reaching opportunities) are increasingly important amongst transport planners aiming to foster just and sustainable cities [@vale2023accessibility]. Both the 15-Minute City and accessibility measures are flexible enough to consider all destination types holistically however, gendered examinations have been lacking in the literature. For instance, accessibility analyses have historically focused on employment-centric and discretionary travel, types of travel more frequent for working-age and higher-income men.

To counter this masculinist bias, this study investigates a way to gender-mainstream the 15-Minute City through a care lens [e.g., @lawWomenTransportNew1999; @cresswell_gendered_2008; @Oestergaard1992; @levy1991towards; @levy2013travel] supported by the Mobility of Care framework (coined by @sanchezdemadariagaMobilityCareIntroducing2013). Mobility of Care emphasizes the importance of travel to unpaid work (care trips) in contrast to the better-studied travel to employment and leisure. While all three trip types (work, care, and leisure) are essential, care trips are often relatively shorter-distance, proximate to residential/work/school, and comprise approximately 30% of adults’ daily trips [@sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory; @mejia2021mobility]; fitting well within the 15-Minute City conceptualisation.

Our study provides an empirical example that maps the 15-Minute City onto the Mobility of Care framework. Specifically, it identifies which areas in Hamilton, Canada are 'caring 15-minute neighbourhoods'. To do so, a database of care destinations is created using secondary data. In this database, care destinations include all places associated with sustaining household tasks needed for the reproduction of life namely: shopping (e.g., groceries), errands (e.g., libraries), health (e.g., dentist), and caring for dependents (e.g., schools). This database is used to estimate the number and mix of care-destinations that can be reached within a 15-minute walk- and cycling- sheds from census centroids. Typologies are generated to illustrate which neighbourhoods can and to what degree facilitate 15-minute access to care.

Results indicate only a few neighbourhoods outside of the downtown core are 'care-complete', i.e., contain a sufficiently high mix of care destinations from all categories and sub-categories. However, some neighbourhoods are almost 'care-complete' and provide 15-minute access to some care categories. Our study frames these neighbourhoods on the continuum of 'caring' and in need of further intervention. The quantitative investigation conducted provides a high-level picture of what neighbourhoods (and their underlying land-use) are connected to transport infrastructure that can support reaching care-destinations.

Taken together, this study provides a theoretical bridge to connect 15-Minute Cities, accessibility analysis and Mobility of Care framework for the purpose of informing policy choice aimed to encourage just and sustainable mobility.

# Introduction

Caring is one of the most fundamental activities in societies, which has allowed humanity to survive and evolve. It involves activities targeted to meet physical, psychological, and emotional needs of others [@ilo_care_2018]. Yet, it is one of the worst devalued, underpaid, and unequal activities worldwide. Traditionally, caring activities have remained on women’s shoulders. Even though the society has evolved, and nowadays women attain higher educational levels and tend to work out of home more than ever before, the bulk of care remains (at any stage of her lifecourse) heavily on her shoulders. The Second Job or the Second Shift remains in all industrial societies, mostly women’s work (Hayden, 1982 (?Citation)) [@hochschild_second_2012].

This unequal share of caring responsibilities leads to gendered differences in among others, career development, differences in contract types and decision-making positions, unbalanced share of women in certain professions, gender pay gap, time poverty, as recognised by different international organisations (European Institute for Gender Equality, 2016 (?Citation)) [@ilo_care_2018]. In fact, already the first feminists in the United States observed that domestic labor was the most relevant basic cause of women’s inequality [@gordon_womans_1976] (Hayden, 1982 (?Citation)).

According to the [@ilo_care_2018], women and girls perform more than three-quarters of the total amount of unpaid care work and two-thirds of care workers are women. The caring economy is part of the women-dominated industries, it is characterised by the evaluative discrimination, which is only one of the dimensions of the gender pay gap [@mejiadorantes_discussing_2019]. Especially in the case of domestic workers, care workers are most of the times women and migrants’ workers, with poor working conditions, and many times associated with informal jobs which imply lack of labour rights, no social and health protection, susceptible of exploitation, one of the poorest working conditions across the care economy as the ILO observes [@ilo_care_2018]. Care workers becomes disrespected, leading to a devaluation of their activity, and consequently remains as one of the less well-paid occupations a person can undertake.

If we turn our attention to the urban planning perspective, Caring is a major human activity that may take form in all urban settings, in all transport modes, all times a day and year, and yet it is not described as neither origin nor destination, and until recently, it was inexistent.

Researchers from various fields have highlighted that western modern industrial societies have imposed a physical separation of the private sphere from the public one, splitting production and reproduction, public and private spaces or simply work and home, concentrating on the economic outcome of being mobile [@cresswell_gendered_2008; @lawWomenTransportNew1999; @levy2013travel; @little_gender_1994; @tronto_toward_1990]. The man with legal rights, paid labour public sphere and woman in the private sphere, dependent, with family duties, personal relations and unpaid labour [@tronto_toward_1990]. In this line, @cresswell_gendered_2008 described how masculinity was related to mobile and active, whereas femininity to static. As an example of this dichotomy, transport planning historically framed non-work/school trips as 'non-mandatory' activities, downgrading these activities in comparison to work.

Women have consistently been found to complete more mobility of care trips than men. In one study, mobility of care comprised 32% of women’s daily trips compared to 28% of men’s. While this gap was significant, it was found to be far greater in lower income households where women complete 20% more care trips than men [@ravensbergen2023exploratory]. This gendered division in mobility of care is perhaps unsurprising as women tend to complete a disproportionate amount of household-serving labour, such as caring for children and other dependents, preparing meals, cleaning, and running errands, when compared to men [@RN4]. This gender gap varies geographically [@RN5]. In Canada, adult women complete 1.9 hours, on average, more of unpaid care than men every day [@RN4]. This labour is essential, and a large component of daily life.

From this research motivation, Sánchez de Madariaga coined the term ‘mobility of care’ in 2013 to refer to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands [@sanchezdemadariagaMobilityCareIntroducing2013]. While decades of research had examined these unique household-serving trips, her seminal work was the first to consider all these trips as one category. In doing so, she showed how mobility of care is a significant proportion of daily travel. Indeed, in her original work, and in the studies that have followed, it appears that approximately 30% of adults’ trips are for care purposes [@sanchezdemadariagaMobilityCareIntroducing2013; @sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory]. Sánchez de Madariaga not only shows how important these mobility of care trips are, but also highlights the ways in which “mobility of care is systematically under-represented in any analysis of urban transport” (p. 37). Transport surveys and tools do not directly capture mobility of care, which re-enforces the idea that these trips are not a significant part of daily mobility.

Cities have simplified individuals as a universal and neutral entity. As a result, public spaces are differently experienced depending on our multiple identities. Indeed, asymmetric in the distribution of time of citizens is deeply anchored with the public and private spaces organization within cities. In this light, we focus on the potential access to destinations of care through an empirical case study of the mid-sized city of Hamilton, Canada. We acknowledge that caring is composed by actors and activities but aim to spotlight the role that the urban environment plays on our daily lives, on easing our participation and on exercising our rights to the city as individuals with multiple identities. Our work hopes to contribute to a fruitful discussion that may trigger urban and transport policy interventions that increase gender-equity in caring activities among all actors.

<!-- We experience our urban environment depending on our personal characteristics, such as socioeconomic level, education, profession, household responsibilities, physical or mental disabilities, and very importantly, by gender. Furthermore, our living environment is also influenced by historical, economic, social, cultural, political, religion and environmental factors. All these issues have an influence on the way we live and experience the city, on how we participate in the community and on our social relations in the public and private space, on how we invest in our socio-economic development, and finally on our well-being.  -->

<!-- Cities have simplified individuals as a universal and neutral entity. As a result, public spaces are differently experienced depending on our multiple identities. Indeed, the distribution of time of citizens is deeply anchored with the public and private spaces organization within cities. In fact, the asymmetries in the distribution of time of individuals are rooted on city spaces, distances to reach them, transport alternatives and different perceived fears, emotions and constraints, along with the roles typically associated with men and women, together with the gendered division of labour [@villagran_urban_2020]. Furthermore, these authors emphasize that it is a global reality that women invest a greater amount of time to care and to unpaid work and that it is socioeconomically stratified. -->
<!-- Many studies worldwide have documented how gender influences mobility patterns. Men commute longer distances and present direct travel patterns, whereas women present chained trips to accomplish on their way many different activities (for example, dropping children off at schools or daycare, run errands or accompany family members to medical services). For these reasons it has been documented that women tend to have shorter but more complicated transport patterns. They also tend to rely more on public transport whereas men use more private transport [@cresswell_gendered_2008; @levy2013travel; @mejia2021mobility; @the_world_bank_handbook_2020]. Sánchez de Madariaga uses the concept “mobility of care” to recognize, measure, make visible, value and account for all the travel associated to caring and unpaid tasks for the household [@sanchezdemadariagaMobilityCareIntroducing2013]. -->

## Feminist 15-minute neighbourhoods

Within the backdrop of urban planning and 'mobility of care', we connect our work to the 15-minute neighbourhood concept.

The spatial structure of cities and regions has been described in different ways and its advantages and disadvantages has been discussed with regard to a wide variety of issues, for example, regarding health, the environment, economic, social and equity impacts. After the traditional monocentric vision of the city with a core Centre Business District (CBD), other alternatives appeared, especially due to the expansion of cities caused by the automobile and motorized transport. The polycentric urban planning emerged based on the idea of sub-centers (or multiple business centers) with their own services and functions, well-connected among other surrounding areas, where the links between land-use and transport were optimised. Economic activities (such as retail or commercial office development) and job centers played an important role in shaping the spatial structure [@schneider_transit_1981; @sweet_are_2017]. This concept may have different interpretations depending on the spatial and geographical scales [@hall_polycentricity_2009].

In the last decades, the need for more sustainable, healthier, and liveable cities has received more attention. Planners and policy makers have proposed different alternatives where residents reach “essential activities” within walkable or cycling distances, reducing the need for using motorised transportation modes. The 15-minutes cities, 20-minutes or x-minutes cities, have been long under the spotlight. This approach has also received the name of Chrono-urbanism or the “cities of proximity” [@logan_xminute_2022; @moreno_introducing_2021].

In the international context, there are different examples that highlight the importance of good access to different services and amenities within walkable distances. Probably one of the first cities to adopt the proximity approach was Portland. The Portland Plan [@city_of_portland_20minute_2010] was adopted by the City Council on April 25, 2012 aiming at fostering an inclusive city development for the following 30 years. The plan is based on the following principles: prosperity, education, health, and equity. The progress report states that equity should be achieved, an Office of Equity and Human Rights was established focusing on equitable service delivery to all residents. Nonetheless the progress report only measures progress on topics related to racial equity or people with disabilities [@portland_government_portland_2017].

In Paris, the adoption of the 15-Minute city put into the spotlight this approach. In Paris, they targeted six social functions that had to be easily accessible at any part of the city: housing, work, health care, groceries, education and leisure [@ville_de_paris_paris_2022]. In 2021, Ottawa also adopted a 15-Minutes approach in the new City of Ottawa Official Plan [@ville_dottawa_quartier_2021].

More recently, @moreno_introducing_2021 considers four dimensions on how the 15-minute city framework: density (in terms of people per sq.km); diversity (including mixed-land use and diversity of people; the temporal and spatial proximity to essential services (residential areas, work, commercial areas, education centres, health facilities and other basic institutions) and digitalisation.

It can be observed that the feminist perspective of the cities of proximity is still underestimated. Only few examples exist that specifically target this issue [@gil_sola_choose_2022; @macintyre_her_2022].

Even if we are on the right direction to facilitate caring activities, develop human-centred urban areas, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities and gender. Similarly, it is necessary to provide alternatives to measure progress on this important issue. On the other hand, an intersectional approach to understand how different personal identities interact in our daily day activities in the cities is necessary. There is still room for improvement.

## Mobility of care and mode-constraints

Given that almost one third of daily trips are for care purposes [@sanchezdemadariagaMobilityCareIntroducing2013], it is important to plan for mobility of care. There are also equity considerations to consider: as discussed, women, and especially low-income women, complete this travel more than men. Planning for mobility of care may also warrant different considerations because mobility of care trip characteristics differ from other trips, notably those of the commute to work. For instance, many of these trips are made through trip-chains, i.e., when a non-work-related stop is made during the commute. Women have been found to do so more than men, which is understood as being due to their need to balance unpaid-household serving care and paid employment (e.g., picking up groceries or children on the way home from work) [@ravensbergen2023exploratory; @scheinerWomenComplexDaily2015].

Further, compared to the trip to work, mobility of care trips are more frequently completed by car and by foot, and less frequently done by transit or bicycle [@ravensbergen2023exploratory]. Evidence for why this modal difference is observed exists. For instance, it is possible that walking is more common for care than for work trips because care trips tend to be short [@ravensbergen2023exploratory]. In other words, if care destinations are closer to the home, it is more likely that these destinations are walkable. Ample literature examines the benefits and barriers to walking care destinations, such as schools (e.g., [@RN7;@RN8;@RN9]) and grocery stores (e.g., [@RN10;@RN11]). Little of this work uses the mobility of care framework. Instead, it tends to consider singular care destinations in research focusing on walkability.

In car dependent cities, cycling commute rates tend to be low. Emerging evidence has found that they are even lower for care trips than for work trips. For instance, in Montréal it is estimated that 2% of work trips completed by adults are done by bicycle, while only 1.1% of care trips are done using this sustainable travel mode [@ravensbergen2023exploratory]. Qualitative research has also found that in car-dependent contexts, many consider it inappropriate and unsafe to conduct escort trips by bicycle [@ravensbergenVelomobilitiesCareLowcycling2020; @sersliRidingAloneTogether2020]. Then again, other studies have uncovered the joys inherent in travelling by bicycle with children [@RN14]. This qualitative evidence has also found that the bicycle is excellent for trip-chaining owning to its flexibility and ease of parking, and that grocery shopping can be done by bicycle, though this can require some additional materials (e.g., paniers) and competencies (e.g., knowing where to put your groceries)[@ravensbergenVelomobilitiesCareLowcycling2020]. Research in the Netherlands, a high-cycling context, has found that parents can and do use their bicycles to complete mobilities of care. Indeed, one study compared women with children’s bicycle behaviours to those without children and found that the having a child did not change the amount of cycling done, but the characteristics of cycling trips: mothers’ trips tended to center around their childrens’ needs [@RN15]. Taken together, it is evident that the bicycle has potential to serve mobility of care: it is flexible, low-cost, and used for these trips in many contexts. Other barriers to this travel mode seems to exist in low-cycling contexts.

The difference in uptake of public transport for work compared to care trips is staggering. In Montréal, it is estimated that 18.7% of trips to work are made by transit. In contrast only 5.8% of care trips are made using transit [@ravensbergen2023exploratory]. While research examining accessibility to care destinations by transit exists (e.g., [@RN16;@RN17;@RN18;@RN19]), little research examines why few people use this sustainable transport mode for mobility of care. Though transit use for mobility of care is low, one study found that is it is significantly higher for low-income women. Indeed, approximately 16% of low-income women in Montreal were found to use transit for care trips (compared to 4% of high income women) [@ravensbergen2023exploratory]. In many cities, public transport is not designed to meet household demands in a convenient way [@RN20;@grantsmithManagingChallengesCombining2016]. Given how low transit use is for care trips amongst the general population, it is possible that many of these low-women using transit are transit-dependent. In other words, they may be using transit for care trips because they do not have access to a car, and not because transit is their preferred travel mode.

It appears that the car is the preferred travel mode for mobility of care. In one study, car use was higher for mobility of care trips than the commute to work (82.3% vs. 74.7%) [@ravensbergen2023exploratory]. Further, the rate of car use is higher amongst trips that are part of a chain than those that are not (85.28% v. 78.09%) [@ravensbergen2023exploratory]. Some studies found that a car is viewed as the easier way to incorporate unpaid mobility of care into one’s daily travel as it provides convenience and mitigates concerns about traffic and stranger danger [@maciejewskaHaveChildrenThus2019, @carverParentalChauffeursWhat2013]. Others, however, have found that women are more likely to complete household-serving travel duties by alternative modes than men due to their lower car access [@RN24;@RN25]. It seems that car use may be preferred for mobility of care trips, but planning mobility of care for the car may leave those who do not have access to a car, many of them low-income women, with difficult mobility of care trips.

The discussion around mobility of care and travel mode may also be relevant to larger spatial trends, including gendered spatial entrapment. Feminist geographers have noted for decades the spatial impact of women’s disproportionate household-serving labour. A relationship has been noted whereby many women’s household responsibilities results in job searches that emphasize job opportunities that are near the home and part-time or with flexible hours to allow for care responsibilities. This leads to a lower probability of securing higher paid, and often further afield, employment [@hansonGenderMobilityNew2010;@lawWomenTransportNew1999;@RN28;@RN29]. This trend has been found to be surprisingly resilient, with improvements only taking places slowly [@RN30]. Evidence has also shown that this spatial entrapment can be more in car dependent contexts, as access to cars impacts employment rates [@RN31], and is lower amongst women, and especially women of colour [@RN32].

Taken together, there appears to be a preference for travel by car for mobility of care trips. This is problematic, however, in at least two major ways. First, cars are less sustainable than other modes of transport. Normalizing car use for care trips does not advance the low-carbon transport transition that most cities are aiming for (at least in principle!). Since care trips tend to be short, they are also perfect candidate to be completed by bus, bicycle, or by foot. That is, if these alternative and sustainable modes are planned with care in mind. Second, there is equity concerns with assuming people are accessing care destinations by car. Many people, and low-income women in particular, do not have access to a car. Some are also experiencing a degree of spatial entrapment. Having safe and convenient access to care destinations using alternatives to the private car is especially important for those who must rely on these options regardless of whether they are well planned.

# Data and Methods

## Case study context
In Canada, women have become more educated and have an achieved an increased participation in the labour market [@government_of_canada_gender_2023]. In fact, the educational level of women has surpassed most men [@government_of_canada_women_2023]. Reports observe that the gender pay gap has narrowed but disparities persist [@drolet_intersectional_2023;@pay_equity_office_gender_2023]. According to the statistics, in Ontario between 1998 to 2022 the gender gap narrowed but still women make 87 cents on average for every dollar made by a man [@pay_equity_office_gender_2023]. All reports observe that it affects women differently, women with disabilities, immigrants, indigenous or racialized suffer especially from the gender pay gap. Similarly to other countries worldwide, the occupational segregation remains and women are segregated into lower paid job classes and there is a dominance of women in minimum wage jobs [@drolet_intersectional_2023]. Furthermore, women are underemployed, and make up most part-time work. In this respect, the lack of childcare alternatives enforces this obligatory part-time work [@equal_pay_coalition_myths_2022]. All researchers and organisations agree that an intersectional approach is of utmost importance to better plan policies to enforce gender equality.

- Next: describe hamilton's land-use context; communities which are urban, rural, suburban. Also describe the residential parcels summary statistics, ~140,000 points, aggregated at the level of Canadian Census DA. See population descriptive plots in @fig-Fig1:
```{r plot-figure1, out.width=600, out.height=600}
#| label: fig-Fig1
#| fig-cap: "The number of residential parcels per DA in 2020 (top) and the rate of population per dwelling per DA (middle) and prevalence of Low-income cut-offs after-tax (LICO-AT) per DA (bottom) retrieved from the 2021 Canadian Census. All scales in quartiles. Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics("data/pop_descriptive_plots.tiff")
```

## Estimated travel times
- Next: describe the estimated 15-minute walk travel times using {r5r}, from all residential parcel points to all care destinations assuming default walking speed and OSM road network for Hamilton. Provide summary statistics.

## Care destination dataset
A spatial dataset of care destinations for Hamilton was compiled; sources include provincial and municipal open data portals [@governmentofontarioOntarioDataCatalogue2023; @cityofhamiltonOpenData2023], Data Axle, a consumer dataset compiled of businesses and companies within Canada [@axledataConsumerData2023] and manual retrieved through Google Maps. Each destination was categorized based on the specific type of care being accessed following the mobility of care research by Sanchez de Madariaga & Zucchini (2019) [@sanchezdemadariagaMeasuringMobilitiesCare2019]. Initial categories include child, elder,  grocery, health, and errand -centric destinations. For the purpose of this study, child- and elder- centric are combined to represent a 'dependent-centric' category. Category sources of data and preparation notes are detailed in @tbl-Tbl1. Their spatial distribution and sub-categories are visualised in @fig-Fig2.

```{r creating-table-content}
dat <- data.frame(matrix(c("Child-centric",
                           "[@opendatahamiltonEducationalInstitutions2022; @governmentofontarioOntarioDataCatalogue2023; @cityofhamiltonChildCareRegistry2023; @opendatahamiltonParksDatabase2022; @opendatahamiltonRecreationCommunityCentres2022]",
                           "Schools, daycares, and community centres, recreation centres, and parks: 1,190 locations are included. After manual review, all locations that typically do not serve children were removed including: Post-Secondary, Adult-Learning Centres, Group Homes, and Foster Care Centres. Further, through examination some Section 23 institutions defined as _“centres for children who cannot attend school to meet the needs of care or treatment, and rehabilitation”_ [@governmentofontarioministryofeducationFundingEducationCommunity2023], were kept due to their innate connection to care.",
                           "Elder-centric",
                           "[@opendatahamiltonRecreationCommunityCentres2022; @ontarioministryofhealthgeohubMinistryHealthService2023]",
                           "Senior centres, long-term care homes, and retirement homes: 75 destinations are identified.",
                           "Grocery-centric",
                           "[@axledataConsumerData2023]",
                           "Grocery stores, namely a place a household could buy groceries ranging from convenience stores to large retail stores: 381 destinations are identified. Data is filtered by Company Name, Suite Number, Address, City, Province, Phone Number and Postal Code. The type was then identified e.g., grocers specialty foods, grocers retail, grocer health food, grocer wholesale, grocer curbside, grocer delicatessen wholesale, grocer convenience. Data was crossreferenced to ensure all included locations were operational and legitimate grocery stores.",
                           "Health-centric",
                           "[@ontarioministryofhealthgeohubMinistryHealthService2023; @hamiltonniagarahaldimandbranthealthlineWalkInMedicalClinics2023]",
                           "Hospitals, pharmacies, clinics, and dentist offices: 421 destinations are identified. Hospitals and pharmacies were retrieved while clinics and dentistry clincs were manually scraped from a healthcare services database and checked via Google Maps to remove non-operational locations and confirm dentistry-orientation. ",
                           "Errand-centric",
                           "Hamilton libraries [@opendatahamiltonLibraries2022], post office locations [@axledataConsumerData2023;@canadapostPostOfficeLocator2023], and datasets of all national bank chains [@bankofmontrealBankMontrealBranch2023; @thehongkongandshanghaibankingcorporationlimitedhsbcHSBCBranchABM2023; @nationalbankNationalBankBranches2023; @royalbankofcanadaRBCBranchABM2023; @scotiabankScotiabankABMBranch2023; @thetorontodominionbankTDBankBranch2023].",
                                                      "Libraries, post offices, and banks: 158 destinations are identified. Post offices are retrieved from a mix of databases, and duplicates are removed. Banks are also derived from Data Axle and then cross-referenced to ensure data quality with a Bank Locator website for all national banking firms."
                           ), ncol = 3, byrow = T))
```

```{r creating-table}
#| tbl-cap: "Details on the preparation and data sources of care destinations."
#| label: tbl-Tbl1
set_flextable_defaults(font.size= 10,
                       font.family = NULL,
                       text.align = 'justify',
                       table.layout = 'fixed')

flextable(dat) %>%
  ftExtra::colformat_md() %>%
  compose(
    i = 1, j = c(1,2,3), part = "header", value = as_paragraph(c("Care category", "Sources", "Data preparation notes"))) %>%
  width(j = 1, width = 1.7, unit = "cm") %>%
  width(j = 2, width = 4.3, unit = "cm") %>%
  width(j = 3, width = 11, unit = "cm")
```

```{r plot-figure2, out.width=600, out.height=600}
#| label: fig-Fig2
#| fig-cap: "The locations of care destinations in Hamilton separated by the author-generated categories of: dependent-, errand-, grocery- and health- centric care categories. Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics("data/care_categories_plot.tiff")
```

For the purpose of this analysis and in absence of city-wide household preferences for care destinations, all locations are re-weighted to make each of the four categories conceptually equivalent i.e., each location within each category $c$ is `r word_comment("Thoughts on this", "weighted to equal")` $W_j^c = \frac{1}{C} \frac{\sum_{c} O_j^c}{O_j^c}$ where $C$ is the total number of categories and $O_j^c$ is a location at destination $j$ within a category $c$.  The total sum of locations is preserved after weighting such that $\sum_{c} \sum_{j} O_j^c = \sum_{c} \sum_{j} W_j^c * O_j^c$. Conceptually, this simplistic re-weight makes destinations in categories with fewer locations more important and destinations in categories with more locations less important. The underlying assumption is the status-quo of locations per category is sufficient, e.g., many daycares (smaller capacities) and few hospitals (larger capacities) sufficiently serves  the populations' care needs. Re-weighting in this way ensures that populations who are near _assumed_ smaller capacity destinations such as daycares are not assigned a disproportionately high accessibility score and vice versa for populations near destinations with larger service capacities like hospitals. In absence of empirical data for amenity weight calibration, this methodological assumption is a limitation.

- Dependent-centric (`r sum(care_dest$Care_Category == "Dependent-centric")` destinations each at `r Dependent_Centric`), 
- Errand-centric (`r sum(care_dest$Care_Category == "Errand-centric")` destinations each at `r Errand_Centric`),
- Grocery-centric (`r sum(care_dest$Care_Category == "Grocery-centric")` destinations each at `r Grocery_Centric`) and,
- Health- centric (`r sum(care_dest$Care_Category == "Health-centric")` destinations each at `r Health_Centric`) .

## Methods
Cumulative opportunity method; 15-minute normative catchment for walk mode (and cycling, travel time matrix already calculated but to be added later) is used to represent the 15-minute neighbourhood concept. This has been done in previous literature such as @vale2023accessibility, @graellsCityCitiesMeasuring2021 and @gaglioneUrbanAccessibility15minute2022.

In this work, the cumulative opportunity score is calculated for each parcel and aggregated by care category at the DA level. The Q3 number of opportunities accessibility in a 15 minute walk per care category in retained for each DA. So in this analysis, $i$ is both the parcel and DA spatial unit. To the author's this is one of the first 15-minute neighbourhood analyse with a focus on care destinations.

**Cumulative opportunity accessibility** takes the following form:
$$
S^c_i=\sum_{j}W_j^{c}\cdot f(c_{ij})
$$
\noindent Where:

- $i$ is a set of parcel point origin locations.
- $j$ is a set of destination locations.
- $c$ is a set of care categories.
- $W^c_j$ is a number of weighted opportunities ($O_j^c$) of category $c$ at $j$.
- $c_{ij}$ is the travel cost between $i$ and $j$.
- $f(\cdot)$ is an impedance function of $c_{ij}$; within the cumulative opportunity approach, it is a binary function that takes the value of 1 if $c_{ij}$ is less than a selected value [@handyMeasuringAccessibilityExploration1997].
- $S_{i}^c$ is the unconstrained accessibility, the sum of weighted opportunities reachable within $f(\cdot)$, at each $i$ for each $c$.

Next, the Q3 (median + standard deviation) of all $\sum_{c}S_i^c$ values for each DA are taken and represented as $S^{c}_{DA}$. $S^{c}_{DA}$ is the upper quartile number of weighted care destinations that can be potentially interacted with within a 15 minute walk by parcels within a DA: an optimistic perspective on how many opportunities can be potentially reached `r word_comment("Thoughts on this choice instead of the median? I do this because the parcels are not population weighted, so 1 parcel = 1 travel time. There can be obviously multiple households in 1 parcel, so using Q3 makes some sense.", ".")`

*Entropy* measure has been widely used to characterise the diversity of land-use mix in early work by @frankLinkingObjectivelyMeasured2005 based on the species diversity Shannon-Wiener index that expresses relative evenness throughout a sequence (this measure among other species diversity indices of the time are reviewed in @whittaker1972evolution). In the realm of urban studies and planning literature, it has since been widely used to characterise the diversity in land-use mix [@ewingTravelBuiltEnvironment2010a], especially in the context of active modes [@luUrbanDensityDiversity2017; @mavoaIdentifyingAppropriateLanduse2018] and suburban sprawl [@randallGISBasedLand2015]. Work that has used diversity indices alongside accessibility analysis are scarce, though examples of its use as parameters within accessibility scores in the case of employment opportunities [@chengMeasuringUrbanJob2013; @daiIncorporatingJobDiversity2018] and more recently to describe the diversity in transit facilities [@yinIncorporatingFacilityDiversity2024] are present. To the authors knowledge, the use of an entropy measure to reflect the diversity in care destination amenity types and in context within a 15-minute neighbourhood, has yet to be done.

In our work, each destination is a specific type $st$ (i.e., `r care_dest$Care_Category_Specific |> unique()`) as well as a care category $c$. When $S_i^c$ is aggregated for each category at the $DA$ it reflects how all opportunities from all four care categories can be reached. The four care categories have meaning to the mobility of care conceptualisation. However, it doesn't explicitly reflect the specific type of destination. In our work, the diversity of the specific type of destination is of interest, we calculate $S^{st}_i$ for each parcel $i$ for each $st$ type following the equation for $S^{st}_i$. This upper quartile is aggregated at the $DA$ to come up with a $S^{st}_{DA}$ `r word_comment("Thoughts on this choice? I think we need to consider both the category type and the specific type in this paper. We justify the categories so that's fair but I still think there's value in presenting the specific type  some how - this is what I think this diversity index is good for so I've presented that as an option. Note though, it'd be much more simple to just focus on calculating the $D^{st}_{DA}$ using the $S_i^c$ and not introducing the $S_i^{st}$ calculation like I've done.. so I've made the analysis more complicated. Also destination weighting is based on the four categories - so that is distorting the diversity in certain ways.",".")`

As such, the entropy measure is used to reflect the diversity of specific type of care destinations. The value ranges from 0 to 1, with 1 representing total evenness in the number of (weighted) specific type $st$ care opportunities can be reached at the aggregated spatial unit $DA$:

$$
D^{st}_{DA} = \frac{-\sum_{c} S_i^{st} / S^{st}_{DA} \times ln(S_i^{st} / S^{st}_{DA} )}{ln(ST)}
$$
\noindent Where:

- $i$ is a set of parcel point origin locations.
- $st$ is a set of specific type care categories. In our work they are `r care_dest$Care_Category_Specific |> unique()`.
- $DA$ is a set of DA spatial units (to aggregate parcel point data).
- $S_{i}^{st}$ is the unconstrained accessibility, the sum of weighted opportunities reachable within $f(\cdot)$, at each $i$ for each $st$.
- $S^{st}_{DA}$ is the aggregated unconstrained accessibility at each $DA$. In our work, it is the upper quartile of $S_i^{st}$ values in a $DA$. 
- $ST$ is the total sum of categories. In our work, this is equal to 14. 
- $D^{st}_{DA}$ is the entropy index.

# Results

```{r plot-figure3, out.width=600, out.height=600}
#| label: fig-Fig3
#| fig-cap: "The cumulative opportunity measure. The number of care destinations that can be reached, per DA, within 15 min walk. Visualised as 'caring' (based on the opportunity reachable number; Caring is approximately the median) and if complete (if all care categories can be reached; this is an arbitrary selection). Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics("data/all_cats_copp_weighted_plot.tiff")
```

```{r plot-figure4, out.width=600, out.height=600}
#| label: fig-Fig4
#| fig-cap: "The diversity measure based on the _specific type_ of destination cumulative opportunity DA measure. Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics("data/all_cats_entropy_copp_weighted_plot.tiff")
```

```{r plot-figure5, out.width=600, out.height=600}
#| label: fig-Fig5
#| fig-cap: "A bivariate map of the cumulative opportuninities map (aggregated per category at DA - fig3) and the diversity measure (fig 4). Amenity diverse = diviersity index above the region median. Caring/Complete = accessibility value above the region median. The Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics("data/all_cats_bivariate_entropy_copp_plot.tiff")
```

BONUS! Per capita spatial availability. This is not discussed in the methods. The spatial availability calculation assumes population is 'population density' and weighted opportunity count. With this map we can start thinking about per capita availability of opportunities. 
```{r plot-figure6, out.width=600, out.height=600}
#| label: fig-Fig6
#| fig-cap: "Per capita spatial availability of care opportunities. Uses weighted care opportunities, population density as competition, and 15 minute binary cut-off. Aggregates parcel level spatial availability at the DA level (by summation) and divides by the DA population. Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics("data/all_cats_vvi_plot.tiff")
```

`r word_comment("If you're interested in how the unweighted results look like for the same maps see the appendix. I've also calculated the cycling travel matrices as well as the 'no dependent' i.e., the same analysis but now only considering 3 care categories: health, errand, and grocery. For the time being though let's focus on this.", " ")`

# References

::: {#refs}
:::

\pagebreak

# Appendix

```{r plot-figure7, out.width=600, out.height=600}
#| label: fig-Fig7
#| fig-cap: "UNWEIGHTED -- The cumulative opportunity measure. The number of care destinations that can be reached, per DA, within 15 min walk. Visualised as 'caring' (based on the opportunity reachable number; Caring is approximately the median) and if complete (if all care categories can be reached; this is an arbitrary selection). Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics("data/all_cats_copp_plot.tiff")
```

```{r plot-figure8, out.width=600, out.height=600}
#| label: fig-Fig8
#| fig-cap: "UNWEIGHTED -- The diversity measure based on the _specific type_ of destination cumulative opportunity DA measure. Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics("data/all_cats_entropy_copp_plot.tiff")
```

### 2024-05-16 data/01-census-and-traveltimes.qmd
Added Lea and Lucia's introduction text. Added preliminary results and methods section. Draft of deliverables 1.

---
title: "01-census-and-traveltimes"
format: html
---

```{r}
rm(list=ls())
```

## Libraries and settings
```{r, message=FALSE}
library(disk.frame)
library(cancensus)
library(sf)
library(leaflet)
library(r5r)
library(dplyr)
library(purrr)

setup_disk.frame()
# increase Java memory
options(java.parameters = "-Xmx6G")
#changing my time zone to Toronto's time zone
Sys.setenv(TZ='EDT')
```

# Census data - for population characteristics (DA, their centroids for robustness checks, and census data)
Pulling data from {cancensus} for the Hamilton Census Subdivision (CSD) and various census variables at the level of the Dissemination Area (DA). We can see that the Hamilton CSD has a population of 569,353 and is identified by the CSD region id of "3525005". It is part of the CMA with ID 35537.
```{r}
list_census_regions('CA21') %>% 
  filter(level == "CSD", name %in% c("Hamilton"))
```

```{r, eval=FALSE}
HAM_census_21 <- get_census(dataset='CA21', 
                            regions=list(CSD="3525005"),
                            vectors=c("v_CA21_11", #total - aged 0-14
                                      "v_CA21_12", #males - aged 0-14
                                      "v_CA21_13", #females - aged 0-14
                                      "v_CA21_68", #total - aged 15-64
                                      "v_CA21_69", #males - aged 15-64
                                      "v_CA21_70", #females - aged 15-64
                                      "v_CA21_251", #total - aged 65 and over
                                      "v_CA21_252", #males - aged 65 and over
                                      "v_CA21_253", #females - aged 65 and over
                                      "v_CA21_1085", #total - prevalence of LICO-AT (%)
                                      "v_CA21_1086", #male - prevalence of LICO-AT (%)
                                      "v_CA21_1087",#female - prevalence of LICO-AT (%)
                                      "v_CA21_7647", #total - walked to work
                                      "v_CA21_7635", #total - drove or was driven to work
                                      "v_CA21_7644", #total - used transit to get to work
                                      "v_CA21_7650", #total - biked to work
                                      "v_CA21_7653" #total - used some other mode to get to work
                                      ), 
                                      level='DA', use_cache = FALSE, geo_format = 'sf', quiet = TRUE)

#NOTE: LICO-AF = "The Low‑income cut‑offs, after tax refer to income thresholds, defined using 1992 expenditure data, below which economic families or persons not in economic families would likely have devoted a larger share of their after‑tax income than average to the necessities of food, shelter and clothing. More specifically, the thresholds represented income levels at which these families or persons were expected to spend 20 percentage points or more of their after‑tax income than average on food, shelter and clothing. "

#NOTE: the 'mode choice' is the count of the TOTAL population aged 15 years and over who is employed in the labour force.
```

Check:
```{r}
leaflet() |>
  addTiles() |>
  addPolygons(data= HAM_census_21, color = "green") 
```

Save DA census data (sf object, polygons) :
```{r, eval=FALSE}
save(HAM_census_21, file="HAM_census_21.rda")
```

# Care destinations
Load the *care destination* database that has been created and reformat it as a sf object:
```{r}
care_dest <- read.csv("data-raw/FINAL_Care_Destinations_2023.csv")

care_dest <- st_as_sf(care_dest, coords = c("LONGITUDE", "LATITUDE"))
```

Check:
```{r}
leaflet() |>
  addTiles() |>
  addMarkers(data= care_dest, popup = ~as.character(Name))
```

Save care destinations (sf object, points):
```{r}
save(care_dest, file="care_dest.rda")
#load( file="care_dest.rda")
```

# Modal-shed 15-minute travel time calculations (not used in the analysis)
Formatting origin (centroids of DAs (from_ids)) and destination (centroids of care destinations) points for the travel time calcs using r5r:
```{r}
load("HAM_census_21.rda")
Community_Boundaries <- read_sf("data-raw/boundaries/Community_Boundaries.shp") |> st_transform(st_crs(HAM_census_21))
City_Boundary <- read_sf("data-raw/boundaries/City_Boundary.shp") |> st_transform(st_crs(HAM_census_21))
```

```{r}
#some DAs are oddly shaped, such that their centroids are outside of their surface. here we create a function if the polygon has a centroid that is inside the polygon, use that, otherwise, find a point within the polygon. (From: https://stackoverflow.com/questions/52522872/r-sf-package-centroid-within-polygon)

st_centroid_within_poly <- function (poly) {

  # check if centroid is in polygon
  centroid <- poly %>% st_centroid() 
  in_poly <- st_within(centroid, poly, sparse = F)[[1]]

  # if it is, return that centroid
  if (in_poly) return(centroid)

  # if not, calculate a point on the surface and return that
  centroid_in_poly <- st_point_on_surface(poly) 
  return(centroid_in_poly)
}
```
```{r}
# orig_sf <- HAM_census_21 |> st_intersection(Community_Boundaries |> filter(COMMUNITY_ == "Dundas")) |>
#   mutate(lon = map_dbl(geometry, ~st_centroid_within_poly(.x)[[1]]),
#          lat = map_dbl(geometry, ~st_centroid_within_poly(.x)[[2]]))

orig_sf <- st_centroid_within_poly(HAM_census_21 |> st_intersection(Community_Boundaries |> filter(COMMUNITY_ == "Dundas")))

orig <- data.frame(id = orig_sf$GeoUID,
                   lon = st_coordinates(orig_sf)[,1],
                   lat = st_coordinates(orig_sf)[,2])

save(orig, file="HAM_census_21_centroids.rda")
```

## Setting up r5r core

```{r set up r5 path, include=FALSE}
# the r5r package requires Java Development Kit version 11, which can be downloaded from https://www.oracle.com/java/technologies/javase-jdk11-downloads.html . See the direction given in the r5r installation here (https://ipeagit.github.io/r5r/articles/r5r.html)
dir.create("./data-raw/tt") #create a folder, if it already exists - this function does nothing

r5_path <- file.path("./data-raw/tt")
list.files(r5_path)
```

```{r download OSM street network, eval=FALSE}
# Specify the URL of the OSM file for Ontario - it takes a few minutes to download (~56MB)
download.file(url = "http://download.geofabrik.de/north-america/canada/ontario-latest.osm.pbf",
              destfile= file.path(r5_path, "Hamilton.osm.pbf"), mode = "wb")
#clip to the the hamilton city limits using OSM Convert .exe (). (NOTE: x = longitude)
osmdata::getbb("Hamilton")
```

```{r download GTFS data, eval=FALSE}
download.file(url = "https://transitfeeds.com/p/hamilton-street-railway/31/latest/download",
              destfile= file.path(r5_path, "HSR_transit.zip"), mode = "wb")
```

```{r build graph, eval = FALSE, cache=FALSE}
#takesa few minutes
r5_HAM <- setup_r5(data_path = r5_path,  verbose = TRUE, elevation = "NONE")
```

## Calculating the isocrones: DA centroids to care destinations

Calculate **walk** 15-minute travel times from centroids of DAs (from_ids).
```{r walk travel time}
library(concaveman) #dundas works. Flamborough, hamilton, ancaster, stoney creek and Glanbrook DO NOT.
iso15_walk <- r5r::isochrone(r5r_core = r5_HAM,
                          origins = orig,
                          mode = "WALK",
                          sample_size = 1,
                          max_trip_duration = 60,
                          cutoffs = c(0,30, 60))

save("iso15_walk", file = "census-cent-ttm/iso15_walk.rda")
```

Check:
```{r}
# Mapping DAs #is weird 35250056
leaflet() |>
  addTiles() |>
  addPolygons(data = iso15_walk |> filter(id == 35250056)) |>
  addPolygons(data= HAM_census_21 |> filter(GeoUID == 35250056), color = "green") |>
  addMarkers(data = orig |> filter(id == 35250056) |> st_as_sf(coords = c("lon", "lat")))
```

```{r}
library(data.table)
library(ggplot2)
street_net <- street_network_to_sf(r5_HAM)
main_roads <- subset(street_net$edges)|>
                     #, street_class %like% 'PRIMARY|SECONDARY') 
                      st_intersection(Community_Boundaries |> filter(COMMUNITY_ == "Dundas"))

colors <- c('#ffe0a5','#ffcb69','#ffa600','#ff7c43','#f95d6a',
            '#d45087','#a05195','#665191','#2f4b7c','#003f5c')

ggplot() +
  geom_sf(data = iso15_walk |> filter(id == 35250056), aes(fill=factor(isochrone)), color = NA, alpha = .7) +
  geom_sf(data = main_roads, color = "black", size=0.01, alpha = 0.2) +
  geom_point(data = orig |> filter(id == 35250056), aes(x=lon, y=lat, color='DA centroids')) +
  scale_fill_manual(values = rev(colors) ) +
  scale_color_manual(values=c('DA centroids'='black')) +
  labs(fill = "Travel time\n(in minutes)", color='') +
  theme_minimal() +
  theme(axis.title = element_blank())
```

# Calculating travel time matrices: from DA centroids to care destinations, all modes
NOTE: in the future, compare these travel times to the population-weighted aggregated travel times. 
```{r}
dest <- care_dest

dest$lon <- st_coordinates(dest)[,1]
dest$lat <- st_coordinates(dest)[,2]

dest <- dest %>% 
  st_drop_geometry() %>%
  transmute(ID = as.character(ID),
            lon,
            lat) %>%
  rename("id" = "ID")
```

Calculate **car** travel times from centroids of DAs (from_ids) to destination points (to_ids). 60 min max travel time. 8am departure on a Tuesday. 30min +/- departure window. 
```{r car travel time, message=FALSE, eval=FALSE}
departure_datetime <- as.POSIXct("08-06-2023 08:00:00 EDT", # 12 GMT is 8am in Toronto
                                 format = "%d-%m-%Y %H:%M:%S")

start.time <- Sys.time()
ttm_care_car <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = orig,
                          destinations = dest,
                          mode = c("CAR"),
                          departure_datetime = departure_datetime,
                          time_window = 30,
                          percentiles = c(25,50,75),
                          max_trip_duration = 30,
                          verbose = TRUE)

end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))

ttm_care_car <- as.data.frame(ttm_care_car)

save("ttm_care_car", file = "census-cent-ttm/ttm_care_car.rda")
```

Calculate **transit** travel times from centroids of DAs (from_ids) to destination points (to_ids). 60 min max travel time, 15 min max walking limit. 8am departure on a Tuesday. 30min +/- departure window.
```{r transit travel time}
departure_datetime <- as.POSIXct("08-06-2023 08:00:00", # 12 GMT is 8am in Toronto (EDT)
                                 format = "%d-%m-%Y %H:%M:%S")

start.time <- Sys.time()
ttm_care_transit <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = orig,
                          destinations = dest,
                          mode = c("WALK", "TRANSIT"),
                          departure_datetime = departure_datetime,
                          time_window = 30,
                          percentiles = c(25,50,75),
                          max_trip_duration = 60,
                          max_walk_time = 15,
                          verbose = TRUE)

end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))

ttm_care_transit <- as.data.frame(ttm_care_transit)

save("ttm_care_transit", file = "census-cent-ttm/ttm_care_transit.rda")
```

Calculate **walk** travel times from centroids of DAs (from_ids) to destination points (to_ids). 60 min max travel time, 15 min max walking limit. 8am departure on a Tuesday. 30min +/- departure window.
```{r walk travel time}
departure_datetime <- as.POSIXct("08-06-2023 08:00:00", # 12 GMT is 8am in Toronto (EDT)
                                 format = "%d-%m-%Y %H:%M:%S")

start.time <- Sys.time()
ttm_care_walk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = orig,
                          destinations = dest,
                          mode = c("WALK"),
                          departure_datetime = departure_datetime,
                          time_window = 30,
                          percentiles = c(25,50,75),
                          max_trip_duration = 60,
                          verbose = TRUE)

end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))

ttm_care_walk <- as.data.frame(ttm_care_walk)

save("ttm_care_walk", file = "census-cent-ttm/ttm_care_walk.rda")
```

Calculate **bike** travel times from centroids of DAs (from_ids) to destination points (to_ids). 60 min max travel time, 15 min max walking limit. 8am departure on a Tuesday. 30min +/- departure window.
```{r bike travel time}
departure_datetime <- as.POSIXct("08-06-2023 08:00:00", # 12 GMT is 8am in Toronto (EDT)
                                 format = "%d-%m-%Y %H:%M:%S")

start.time <- Sys.time()
ttm_care_bike <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = orig,
                          destinations = dest,
                          mode = c("BICYCLE"),
                          departure_datetime = departure_datetime,
                          time_window = 30,
                          percentiles = c(25,50,75),
                          max_trip_duration = 60,
                          verbose = TRUE)

end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))

ttm_care_bike <- as.data.frame(ttm_care_bike)

save("ttm_care_bike", file = "census-cent-ttm/ttm_care_bike.rda")
```

Limitations of travel time calcs:
- does not consider traffic stress (i.e., car, bike, and walk for all percentiles are identical -- the time window does not matter and no traffic is assumed)
- for transit, the time window accounts for departures +/- 30 minutes. This accounts for transit availability variability but also does not accont for traffic stress

# Calculating travel time matrices: from parcel centroids to care destinations (this is used in the analysis!)

```{r}
load( file="care_dest.rda")
dest <- care_dest

dest$lon <- st_coordinates(dest)[,1]
dest$lat <- st_coordinates(dest)[,2]

dest <- dest %>% 
  st_drop_geometry() %>%
  transmute(ID = as.character(ID),
            lon,
            lat) %>%
  rename("id" = "ID")
```

```{r, eval=FALSE}
R_PARCELS_2020 <- st_read("data-raw/parcels/PED_LANDUSE_2020/PED_LANDUSE.shp") |> st_transform(st_crs(HAM_census_21)) |>
  mutate(TYPE = ifelse(LUC1_DESCR == "Residential:Detached House" | 
                         LUC1_DESCR == "Residential:Row/Town House"| 
                         LUC1_DESCR == "Residential:Semi-Detached House" |
                         LUC1_DESCR == "Residential:Apartment (7 or more units)"|
                         LUC1_DESCR == "Residential:Multiplex Dwelling (6 units or less)"|
                         LUC1_DESCR == "Residential:Mobile Home",
                       "Residential", 0)) |>
  filter(TYPE != 0)  #i.e. filter out non-residential

R_PARCELS_2020 <- st_make_valid(R_PARCELS_2020)

R_PARCELS_2020$ID <- 1:nrow(R_PARCELS_2020)

R_PARCELS_CENTS_2020 <- st_centroid(R_PARCELS_2020)
save("R_PARCELS_2020", file = "data-raw/parcels/R_PARCELS_2020.rda")
save("R_PARCELS_CENTS_2020", file = "data-raw/parcels/R_PARCELS_CENTS_2020.rda")
```

```{r}
load( "data-raw/parcels/R_PARCELS_CENTS_2020.rda")

orig_sf <- R_PARCELS_CENTS_2020 |> st_as_sf()
rm(R_PARCELS_CENTS_2020,R_PARCELS_2020)

coords <- st_coordinates(orig_sf)
orig <- data.frame(id = orig_sf$ID,
                   lon = st_coordinates(orig_sf)[,1],
                   lat = st_coordinates(orig_sf)[,2])

rm(orig_sf)
```
```{r}
# due to size, the calculation of travel times needs to be batched. We split the origins into 4 distinct dataframes.

orig_1 <- orig[1:36903,]
orig_2 <- orig[36904:73807,]
orig_3 <- orig[73808:110710,]
orig_4 <- orig[110711:143893,]

#check, make sure sums to 143893
nrow(orig_1)+nrow(orig_2)+nrow(orig_3)+nrow(orig_4)
```

## Walk travel times

Calculate **walk** travel times from centroids of parcels (from_ids) to destination points (to_ids). 30 min max travel time.

```{r walk travel time 1}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(orig_1)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(orig_1,
                          outdir = "parcel-ttm/processing/orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

start.time <- Sys.time()
pb <- txtProgressBar(0, num_chunks, style = 3)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = dest,
                          mode = c("WALK"),
                          departure_datetime = as.POSIXct(strptime("08-06-2023 08:00:00", # 12 GMT is 8am in Toronto (EDT)
                                 format = "%d-%m-%Y %H:%M:%S")),
                          max_trip_duration = 30)

  # export output as disk.frame
  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
                                            nchunks = 1,
                                            outdir = "parcel-ttm/processing/output_ttm",
                                            compress = 50,
                                            overwrite = TRUE),
         add_chunk(output_df, ttm_chunk, chunk_id = i))
  setTxtProgressBar(pb, i)
}
end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))
#"OD matrix calculation took 6.78 minutes..."
ttm_care_walk_parcel1 <- as.data.frame(output_df)

save("ttm_care_walk_parcel1", file = "parcel-ttm/ttm_care_walk_parcel1.rda")
rm("ttm_care_walk_parcel1")
```

```{r walk travel time 2}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(orig_2)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(orig_2,
                          outdir = "parcel-ttm/processing/orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

start.time <- Sys.time()
pb <- txtProgressBar(0, num_chunks, style = 3)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = dest,
                          mode = c("WALK"),
                          departure_datetime = as.POSIXct(strptime("08-06-2023 08:00:00", # 12 GMT is 8am in Toronto (EDT)
                                 format = "%d-%m-%Y %H:%M:%S")),
                          max_trip_duration = 30)

  # export output as disk.frame
  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
                                            nchunks = 1,
                                            outdir = "parcel-ttm/processing/output_ttm",
                                            compress = 50,
                                            overwrite = TRUE),
         add_chunk(output_df, ttm_chunk, chunk_id = i))
  setTxtProgressBar(pb, i)
}
end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))
#"OD matrix calculation took 5.35 minutes..."
ttm_care_walk_parcel2 <- as.data.frame(output_df)

save("ttm_care_walk_parcel2", file = "parcel-ttm/ttm_care_walk_parcel2.rda")
rm("ttm_care_walk_parcel2")
```

```{r walk travel time 3}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(orig_3)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(orig_3,
                          outdir = "parcel-ttm/processing/orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

start.time <- Sys.time()
pb <- txtProgressBar(0, num_chunks, style = 3)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = dest,
                          mode = c("WALK"),
                          departure_datetime = as.POSIXct(strptime("08-06-2023 08:00:00", # 12 GMT is 8am in Toronto (EDT)
                                 format = "%d-%m-%Y %H:%M:%S")),
                          max_trip_duration = 30)

  # export output as disk.frame
  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
                                            nchunks = 1,
                                            outdir = "parcel-ttm/processing/output_ttm",
                                            compress = 50,
                                            overwrite = TRUE),
         add_chunk(output_df, ttm_chunk, chunk_id = i))
  setTxtProgressBar(pb, i)
}
end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))

ttm_care_walk_parcel3 <- as.data.frame(output_df)

save("ttm_care_walk_parcel3", file = "parcel-ttm/ttm_care_walk_parcel3.rda")
rm("ttm_care_walk_parcel3")
```

```{r walk travel time 4}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(orig_4)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(orig_4,
                          outdir = "parcel-ttm/processing/orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

start.time <- Sys.time()
pb <- txtProgressBar(0, num_chunks, style = 3)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = dest,
                          mode = c("WALK"),
                          departure_datetime = as.POSIXct(strptime("08-06-2023 08:00:00", # 12 GMT is 8am in Toronto (EDT)
                                 format = "%d-%m-%Y %H:%M:%S")),
                          max_trip_duration = 30)

  # export output as disk.frame
  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
                                            nchunks = 1,
                                            outdir = "parcel-ttm/processing/output_ttm",
                                            compress = 50,
                                            overwrite = TRUE),
         add_chunk(output_df, ttm_chunk, chunk_id = i))
  setTxtProgressBar(pb, i)
}
end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))
#"OD matrix calculation took 6.23 minutes..."

ttm_care_walk_parcel4 <- as.data.frame(output_df)

save("ttm_care_walk_parcel4", file = "parcel-ttm/ttm_care_walk_parcel4.rda")
rm("ttm_care_walk_parcel4")
```

## Bike travel times

Calculate **bike** travel times from centroids of parcels (from_ids) to destination points (to_ids). 30 min max travel time. 8am departure on a Tuesday.
Calculate **bike** travel times from centroids of parcels (from_ids) to destination points (to_ids). 30 min max travel time. 
```{r bike travel time 1}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(orig_1)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(orig_1,
                          outdir = "parcel-ttm/processing/orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

start.time <- Sys.time()
pb <- txtProgressBar(0, num_chunks, style = 3)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = dest,
                          mode = c("BICYCLE"),
                          departure_datetime = as.POSIXct(strptime("08-06-2023 08:00:00", # 12 GMT is 8am in Toronto (EDT)
                                 format = "%d-%m-%Y %H:%M:%S")),
                          max_trip_duration = 30)

  # export output as disk.frame
  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
                                            nchunks = 1,
                                            outdir = "parcel-ttm/processing/output_ttm",
                                            compress = 50,
                                            overwrite = TRUE),
         add_chunk(output_df, ttm_chunk, chunk_id = i))
  setTxtProgressBar(pb, i)
}
end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))
#"OD matrix calculation took 6.78 minutes..."
ttm_care_bike_parcel1 <- as.data.frame(output_df)

save("ttm_care_bike_parcel1", file = "parcel-ttm/ttm_care_bike_parcel1.rda")
rm("ttm_care_bike_parcel1")

# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(orig_2)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(orig_2,
                          outdir = "parcel-ttm/processing/orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

start.time <- Sys.time()
pb <- txtProgressBar(0, num_chunks, style = 3)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = dest,
                          mode = c("BICYCLE"),
                          departure_datetime = as.POSIXct(strptime("08-06-2023 08:00:00", # 12 GMT is 8am in Toronto (EDT)
                                 format = "%d-%m-%Y %H:%M:%S")),
                          max_trip_duration = 30)

  # export output as disk.frame
  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
                                            nchunks = 1,
                                            outdir = "parcel-ttm/processing/output_ttm",
                                            compress = 50,
                                            overwrite = TRUE),
         add_chunk(output_df, ttm_chunk, chunk_id = i))
  setTxtProgressBar(pb, i)
}
end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))
#"OD matrix calculation took 5.35 minutes..."
ttm_care_bike_parcel2 <- as.data.frame(output_df)

save("ttm_care_bike_parcel2", file = "parcel-ttm/ttm_care_bike_parcel2.rda")
rm("ttm_care_bike_parcel2")

# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(orig_3)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(orig_3,
                          outdir = "parcel-ttm/processing/orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

start.time <- Sys.time()
pb <- txtProgressBar(0, num_chunks, style = 3)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = dest,
                          mode = c("BICYCLE"),
                          departure_datetime = as.POSIXct(strptime("08-06-2023 08:00:00", # 12 GMT is 8am in Toronto (EDT)
                                 format = "%d-%m-%Y %H:%M:%S")),
                          max_trip_duration = 30)

  # export output as disk.frame
  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
                                            nchunks = 1,
                                            outdir = "parcel-ttm/processing/output_ttm",
                                            compress = 50,
                                            overwrite = TRUE),
         add_chunk(output_df, ttm_chunk, chunk_id = i))
  setTxtProgressBar(pb, i)
}
end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))

ttm_care_bike_parcel3 <- as.data.frame(output_df)

save("ttm_care_bike_parcel3", file = "parcel-ttm/ttm_care_bike_parcel3.rda")
rm("ttm_care_bike_parcel3")

# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(orig_4)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(orig_4,
                          outdir = "parcel-ttm/processing/orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

start.time <- Sys.time()
pb <- txtProgressBar(0, num_chunks, style = 3)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = dest,
                          mode = c("BICYCLE"),
                          departure_datetime = as.POSIXct(strptime("08-06-2023 08:00:00", # 12 GMT is 8am in Toronto (EDT)
                                 format = "%d-%m-%Y %H:%M:%S")),
                          max_trip_duration = 30)

  # export output as disk.frame
  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
                                            nchunks = 1,
                                            outdir = "parcel-ttm/processing/output_ttm",
                                            compress = 50,
                                            overwrite = TRUE),
         add_chunk(output_df, ttm_chunk, chunk_id = i))
  setTxtProgressBar(pb, i)
}
end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))

ttm_care_bike_parcel4 <- as.data.frame(output_df)

save("ttm_care_bike_parcel4", file = "parcel-ttm/ttm_care_bike_parcel4.rda")
rm("ttm_care_bike_parcel4")
```

Limitations of travel time calcs:
- does not consider traffic stress (i.e., car, bike, and walk for all percentiles are identical -- the time window does not matter and no traffic is assumed)
- for transit, the time window accounts for departures +/- 30 minutes. This accounts for transit availability variability but also does not accont for traffic stress

INSPECTING:
```{r}
# library(data.table)
# library(ggplot2)
# street_net <- street_network_to_sf(r5_HAM)
# main_roads_all <- street_net$edges |> subset(street_class %like% 'PRIMARY|SECONDARY')
#   
# colors <- c('#ffe0a5','#ffcb69','#ffa600','#ff7c43','#f95d6a',
#             '#d45087','#a05195','#665191','#2f4b7c','#003f5c')
# 
# ggplot() +
#   geom_sf(data = main_roads_all, color = "black", size=0.01, alpha = 0.2) +
#   #geom_point(data = orig, aes(x=lon, y=lat, color='Parcel centroids')) +
#   geom_sf(data = Community_Boundaries, col="yellow", size = 0.01, alpha = 0.2)+
#   theme_minimal() +
#   theme(axis.title = element_blank())
```

### 2024-05-16 data/02-preliminary-analysis.qmd
Added Lea and Lucia's introduction text. Added preliminary results and methods section. Draft of deliverables 1.

---
title: "02-preliminary-analysis"
format: html
---
```{r}
rm(list = ls())
```

Working with the parcel to dest TTM - walk and cycling.

```{r library-setup}
library(biscale)
library(cowplot)
library(sf)
library(here)
library(dplyr)
library(tmap)
library(mapview)
library(tmaptools)
library(pander)
library(flextable)
library(ftExtra)
library(scales)
```

```{r spatial-availability-function}
sp_avail_detailed <- function(x, o_id, d_id, pop, opp, r, f, alpha = 1){

  o_id <- rlang::enquo(o_id)
  d_id <- rlang::enquo(d_id)
  pop <- rlang::enquo(pop)
  opp <- rlang::enquo(opp)
  r <- rlang::enquo(r)
  f <- rlang::enquo(f)

  sum_pop <- x |>
    dplyr::distinct(!!o_id,
                    .keep_all = TRUE) |>
    dplyr::mutate(sum_pop = !!r*(!!pop)^alpha) |>
    dplyr::pull(sum_pop) |>
    sum()

  f_p <- dplyr::pull(x, !!r) * dplyr::pull(x, !!pop)^alpha / sum_pop

  sum_impedance <- x |>
    dplyr::group_by(!!d_id) |>
    dplyr::summarize(sum_impedance = sum(!!f))

  x <- x |>
    dplyr::left_join(sum_impedance,
                     by = rlang::as_name(d_id))

  f_c <- dplyr::pull(x, !!f) / x$sum_impedance

  x$f_c <- f_c
  x$f_p <- f_p

  sum_pa <- x |>
    dplyr::group_by(!!d_id) |>
    dplyr::summarize(sum_pa= sum(f_p * f_c))

  x <- x |>
    dplyr::left_join(sum_pa,
                     by = rlang::as_name(d_id))
  x$f_t <- (f_p * f_c) / dplyr::pull(x, sum_pa)

  x |>
    dplyr::mutate(V_ij = !!opp * f_t)
}
```

```{r}
#boundaries for mapping
Community_Boundaries <- st_read("data-raw/boundaries/Community_Boundaries.shp") |> mutate(COMMUNITY_ = ifelse(COMMUNITY_ == "Hamilton", "Hamilton-Central", COMMUNITY_))|> st_transform(crs=4326)
City_Boundary <- st_read("data-raw/boundaries/City_Boundary.shp") |> st_transform(crs=4326)
hydro_p_LakeOntario <- st_read("data-raw/boundaries/hydro_p_LakeOntario.shp") |> st_transform(crs=4326)
ham_bay <- st_read("data-raw/boundaries/Waterbodies.shp") |> st_transform(crs=4326) |> filter(FEATURE_TY == "Lake")
ham_bay_cropped<-st_crop(ham_bay$geometry, Community_Boundaries$geometry)
hydro_p_LakeOntario_cropped<-st_crop(hydro_p_LakeOntario$geometry, Community_Boundaries$geometry)

load("care_dest.rda")
load("HAM_census_21.rda")

load("parcel-ttm/ttm_care_bike_parcel1.rda")
load("parcel-ttm/ttm_care_bike_parcel2.rda")
load("parcel-ttm/ttm_care_bike_parcel3.rda")
load("parcel-ttm/ttm_care_bike_parcel4.rda")
load("parcel-ttm/ttm_care_walk_parcel1.rda")
load("parcel-ttm/ttm_care_walk_parcel2.rda")
load("parcel-ttm/ttm_care_walk_parcel3.rda")
load("parcel-ttm/ttm_care_walk_parcel4.rda")

ttm_care_bike_parcel <-rbind(ttm_care_bike_parcel1, ttm_care_bike_parcel2,
                             ttm_care_bike_parcel3, ttm_care_bike_parcel4)
rm(ttm_care_bike_parcel1, ttm_care_bike_parcel2, ttm_care_bike_parcel3, ttm_care_bike_parcel4)

ttm_care_walk_parcel <-rbind(ttm_care_walk_parcel1, ttm_care_walk_parcel2,
                             ttm_care_walk_parcel3, ttm_care_walk_parcel4)
rm(ttm_care_walk_parcel1, ttm_care_walk_parcel2, ttm_care_walk_parcel3, ttm_care_walk_parcel4)
```

```{r}
load("data-raw/parcels/R_PARCELS_CENTS_2020.rda") #add the census id to the parcels, then the ttm

#first join it to census. Transfer the GeoUID from the DAs that intersect with the residential land use parcel centroids to the R_PARCELS_CENTS_2020 sf object. (we use 'over' here instead of st_intersection because it is way faster)
R_PARCELS_CENTS_GeoUIDs <- sp::over(as_Spatial(R_PARCELS_CENTS_2020),as_Spatial(HAM_census_21 |> dplyr::select(GeoUID)))

R_PARCELS_CENTS_GeoUIDs_IDs <- cbind(R_PARCELS_CENTS_GeoUIDs, R_PARCELS_CENTS_2020 |> dplyr::select("ID") |> st_drop_geometry()) |> 
  mutate(ID = as.character(ID),
         GeoUID = as.character(GeoUID))

rm(R_PARCELS_CENTS_GeoUIDs)
#Some variables of interest: use "ID" to join "GeoUID", "WARD", "COMMUNITY" "PROP_AREA" and "URBAN_RURA" (either S, U, or R for rural -- to characteris how 'rural' or 'urban' certain areas are by their area)

ttm_care_walk_parcel <- left_join(ttm_care_walk_parcel, R_PARCELS_CENTS_GeoUIDs_IDs, by = c("from_id"="ID"))
ttm_care_bike_parcel <- left_join(ttm_care_bike_parcel, R_PARCELS_CENTS_GeoUIDs_IDs, by = c("from_id"="ID"))
```

```{r}
HAM_census_21 <- HAM_census_21 |>
  rename("LICO_AT_TOTAL" = `v_CA21_1085: Prevalence of low income based on the Low-income cut-offs, after tax (LICO-AT) (%)`) |> 
  mutate(LICO_AT_TOTAL = ifelse(is.na(LICO_AT_TOTAL), 0, LICO_AT_TOTAL)) |>
  rowwise() |> mutate(
         perc_walked = (`v_CA21_7647: Walked`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`, `v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE),
         perc_car = (`v_CA21_7635: Car, truck or van`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`, `v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE),
perc_transit = (`v_CA21_7644: Public transit`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`, `v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE),
 perc_cycle_other = (`v_CA21_7650: Bicycle`+ `v_CA21_7653: Other method`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`, `v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE)) |>
  select(c("GeoUID", "Population", "Households", "Dwellings", "LICO_AT_TOTAL", "perc_walked", "perc_car", "perc_transit", "perc_cycle_other"))

HAM_census_21 <- st_join(HAM_census_21, Community_Boundaries|> select("COMMUNITY_"), largest=TRUE)
HAM_census_21 <- HAM_census_21 |> filter(Population > 0) #drop households with 0 households, only 3 DAs

DA_centroids <- HAM_census_21 |> st_centroid() 
```

```{r reformat care-dest specific categories}
care_dest <- care_dest |> 
  mutate(Care_Category_Specific = case_when(
    Care_Category_Specific  == "Daycare_EarlyON" ~ "Daycare or EarlyON",
    Care_Category_Specific == "Community_Centre" ~ "Community or Recreation Centre",
    Care_Category_Specific == "Daycare_Centre" ~ "Daycare or EarlyON",
    Care_Category_Specific == "Recreation_Centre" ~ "Community or Recreation Centre",
    Care_Category_Specific == "School" ~ "School",
    Care_Category_Specific == "Long-Term_Care_Home" ~ "Long-Term Care or Retirement Home",
    Care_Category_Specific == "Retirement_Home" ~ "Long-Term Care or Retirement Home",
    Care_Category_Specific == "Senior_Centre" ~ "Senior Centre",
    Care_Category_Specific == "Seniors_Active_Living_Centre" ~ "Senior Centre",
    Care_Category_Specific == "Post_Office" ~ "Post Office",
    Care_Category_Specific == "Library" ~ "Library",
    Care_Category_Specific == "Banks" ~ "Bank or ATM",
    Care_Category_Specific == "Grocer_ConvenienceStore" ~ "Convenience Store",
    Care_Category_Specific == "Clinics" ~ "Hospital or Clinic",
    Care_Category_Specific == "Grocer_Retail" ~ "Grocer",
    Care_Category_Specific == "Dentist" ~ "Dentist",
    Care_Category_Specific == "Hosptial" ~ "Hospital or Clinic",
    Care_Category_Specific == "Hospital" ~ "Hospital or Clinic",
    Care_Category_Specific == "Pharmacy" ~ "Pharmacy",
    Care_Category_Specific == "Park" ~ "Park"),
  Care_Category = case_when(
    Care_Category == "Child_Centric" ~ "Dependent-centric",
    Care_Category == "Elder_Centric" ~ "Dependent-centric",
    Care_Category == "Errand_Centric" ~ "Errand-centric",
    Care_Category == "Grocery_Centric" ~ "Grocery-centric",
    Care_Category == "Health_Centric" ~ "Health-centric"))

care_dest <- care_dest |> rename("care_ID" = "ID") |> mutate(care_ID = as.character(care_ID)) 
```
```{r}
care_dest|> st_drop_geometry() |> group_by(Care_Category) |> summarise(count = n())
```

```{r re-weighting points so all destinations are considered spatially equal in access counts, include=FALSE}
Dependent_Centric <- 0.25*2225/1265
Errand_Centric <- 0.25*2225/158
Grocery_Centric <- 0.25*2225/381
Health_Centric <- 0.25*2225/421

#adds up to 2225 opportunities  
0.25*2225/1265*(1265) +  0.25*2225/158*(158) + 0.25*2225/381*(381) + 0.25*2225/421*(421)

#adding the weighted care destination counts -- reweighted such that the 4 categories each make up 25% of the 2225 points.
care_dest <- care_dest |>
          mutate(Dest_count_weighted = case_when(Care_Category == "Dependent-centric" ~ Dependent_Centric,
                                                 Care_Category == "Errand-centric" ~ Errand_Centric,
                                                 Care_Category == "Grocery-centric" ~ Grocery_Centric,
                                                 Care_Category == "Health-centric" ~ Health_Centric))

care_dest$Dest_count_weighted |> sum()
```

```{r travel-time-15min}
#15 mins or less
ttm_care_walk_15 <- ttm_care_walk_parcel |> 
  filter(travel_time_p50 <= 15)|> 
  left_join(
    care_dest |> st_drop_geometry() |> select(c("care_ID", "Care_Category", "Care_Category_Specific", "Care_Category_Speciality", "Dest_count_weighted")),  by = c("to_id" = "care_ID")) |>
  mutate(f = 1)

ttm_care_walk_15 <- ttm_care_walk_15 |>
  left_join(HAM_census_21 |> st_drop_geometry() |> select(c("GeoUID", "Population", "Households", "Dwellings", "LICO_AT_TOTAL")), by=c("GeoUID")) |>
  mutate(pop_dwell = Population/Dwellings)

ttm_care_walk_15 <- ttm_care_walk_15 |> tidyr::drop_na("Population")
```

```{r}
access_15 <- ttm_care_walk_15 |> 
  group_by(from_id, Care_Category, Care_Category_Specific) |> 
  summarize(GeoUID = first(GeoUID),
            copp_isc = n(),
            copp_isc_w = sum(Dest_count_weighted))

test_all <- ttm_care_walk_15 |>
  sp_avail_detailed(o_id = from_id,
                    d_id = to_id, 
                    pop = pop_dwell, #pop assumed at each res. parcel -- it is the pop in the DA divided by the number of dwellings in the DA. 
                    opp = Dest_count_weighted, #instead of just being vector of 1
                    r = f, f = f, alpha = 1) |> 
  group_by(from_id, Care_Category, Care_Category_Specific) |> summarize(V_isc_w = sum(V_ij, na.rm = TRUE))

access_15 <- access_15 |> left_join(test_all, by=c("from_id", "Care_Category", "Care_Category_Specific")) |> ungroup()
rm(test_all)
access_15$V_isc_w |> sum(na.rm = TRUE) #sums to 2206 opportunities -- out of 2225 total opportunities. some opportunities canot be reached wtihin 15 min walk! 
```

```{r}
# all Care_Category_Specific should be represented, i.e., each ID should have 14 rows -- if that category is empty it should be 0. Let's add those 0s. 
access_15_zeros <- 
  data.frame(from_id = rep(R_PARCELS_CENTS_2020$ID, length(access_15$Care_Category_Specific |> unique())),
           Care_Category = rep(c("Dependent-centric", "Dependent-centric", "Grocery-centric", "Dependent-centric", 
                                 "Dependent-centric", "Dependent-centric",
                                 "Errand-centric", "Errand-centric", 
                                 "Grocery-centric","Health-centric",
                                 "Health-centric", "Errand-centric",
                                 "Dependent-centric", "Health-centric"), length(R_PARCELS_CENTS_2020$ID)),

           Care_Category_Specific = rep(c("Park", "Daycare or EarlyON", "Convenience Store", "School",
                                          "Long-Term Care or Retirement Home", "Senior Centre",
                                          "Bank or ATM", "Library", 
                                          "Grocer", "Dentist", 
                                          "Pharmacy", "Post Office",
                                          "Community or Recreation Centre", "Hospital or Clinic"), length(R_PARCELS_CENTS_2020$ID)),
           copp_isc = rep(0, length(R_PARCELS_CENTS_2020$ID)*length(c("Park", "Daycare or EarlyON", "Convenience Store", "School",
                                          "Long-Term Care or Retirement Home", "Senior Centre",
                                          "Bank or ATM", "Library", 
                                          "Grocer", "Dentist", 
                                          "Pharmacy", "Post Office",
                                          "Community or Recreation Centre", "Hospital or Clinic"))),
           copp_isc_w = rep(0, length(R_PARCELS_CENTS_2020$ID)*length(c("Park", "Daycare or EarlyON", "Convenience Store", "School",
                                          "Long-Term Care or Retirement Home", "Senior Centre",
                                          "Bank or ATM", "Library", 
                                          "Grocer", "Dentist", 
                                          "Pharmacy", "Post Office",
                                          "Community or Recreation Centre", "Hospital or Clinic"))),

           V_isc_w = rep(0, length(R_PARCELS_CENTS_2020$ID)*length(c("Park", "Daycare or EarlyON", "Convenience Store", "School",
                                          "Long-Term Care or Retirement Home", "Senior Centre",
                                          "Bank or ATM", "Library", 
                                          "Grocer", "Dentist", 
                                          "Pharmacy", "Post Office",
                                          "Community or Recreation Centre", "Hospital or Clinic")))) |>
  mutate(from_id = as.character(from_id))

access_15_zeros <- left_join(access_15_zeros, R_PARCELS_CENTS_GeoUIDs_IDs, by=c("from_id" = "ID"))

access_15 <- bind_rows(access_15_zeros, access_15) |>
  group_by(from_id, Care_Category, Care_Category_Specific) |> 
  summarize(GeoUID = first(GeoUID),
            copp_isc = sum(copp_isc),
            copp_isc_w = sum(copp_isc_w),
            V_isc_w = sum(V_isc_w))
rm(access_15_zeros)
#check
access_15$V_isc_w |> sum(na.rm=T)
```

Aggregating time!

```{r}
# by care category, for parcels 
access_15_bycategory <- access_15 |> group_by(from_id, Care_Category)  |> 
  summarize(GeoUID = first(GeoUID),
            copp_ic = sum(copp_isc, na.rm=T), #the number of opportunities that can be reached by category from each parcel
            copp_ic_w = sum(copp_isc, na.rm=T), #the number of opportunities that can be reached by category from each parcel
            V_ic_w = sum(V_isc_w, na.rm = T))
#check
access_15_bycategory$V_ic_w |> sum(na.rm=T)
```

```{r}
#by DA and mean:
access_15_DA <- access_15 |>
  left_join(HAM_census_21 |> st_drop_geometry() |> select(c("GeoUID", "Population", "Dwellings", "LICO_AT_TOTAL", "COMMUNITY_")), 
            by=c("GeoUID")) |> 
  group_by(GeoUID) |>
  summarise(copp_iscDA_med = median(copp_isc), #the median number of opportunities (all) that can be reached from each parcel in a DA
            copp_iscDA_sd = sd(copp_isc), #the SD
    copp_iscDA_w_med = median(copp_isc_w), #the median number of opportunities (all) that can be reached from each parcel in a DA
            copp_iscDA_w_sd = sd(copp_isc_w), #the SD
            V_w_iscDA = sum(V_isc_w), #the number of total spatially available opportunities (all) that can be reached from each parcel in a DA
            vv_w_iscDA = sum(V_isc_w)/first(Population), #the per-capita number of total spatially available opportunities (all) that can be reached from each parcel in a DA
            Population = first(Population),
            Dwellings = first(Dwellings),
            LICO_AT_TOTAL = first(LICO_AT_TOTAL),
            COMMUNITY_ = first(COMMUNITY_)) |>
  tidyr::drop_na(Population)

#add the number of parcels per DA 
R_PARCELS_CENTS_count_GeoUID <- R_PARCELS_CENTS_GeoUIDs_IDs |> group_by(GeoUID) |> summarise(parcel_count = n())
access_15_DA <- access_15_DA |> left_join(R_PARCELS_CENTS_count_GeoUID, by="GeoUID")
access_15_DA <- left_join(HAM_census_21 |> select(c(GeoUID)), access_15_DA, by=c("GeoUID"))

# now by category
access_15_bycategory_DA <- access_15 |>
  left_join(HAM_census_21 |> st_drop_geometry() |> select(c("GeoUID", "Population","Dwellings", "LICO_AT_TOTAL", "COMMUNITY_")), 
            by=c("GeoUID")) |>
  group_by(GeoUID, Care_Category) |>
  summarise(copp_ic_med = median(copp_isc), #the median number of opportunities that can be reached by category from each parcel in a DA
            copp_ic_sd = sd(copp_isc), #the SD
            copp_ic_w_med = median(copp_isc_w), #the median number of opportunities that can be reached by category from each parcel in a DA
            copp_ic_w_sd = sd(copp_isc_w), #the SD
            V_ic_w = sum(V_isc_w), #the number of total spatially available opportunities that can be reached by category from each parcel in a DA
            vv_ic_w = sum(V_isc_w)/first(Population), #the per-capita number of total spatially available opportunities that can be reached by category from each parcel in a DA
            Population = first(Population),
            Dwellings = first(Dwellings),
            LICO_AT_TOTAL = first(LICO_AT_TOTAL),
            COMMUNITY_ = first(COMMUNITY_) ) |>
  tidyr::drop_na(Population)

#now by specific category
access_15_byspecificcategory_DA <- access_15 |>
  left_join(HAM_census_21 |> st_drop_geometry() |> select(c("GeoUID", "Population","Dwellings", "LICO_AT_TOTAL", "COMMUNITY_")), 
            by=c("GeoUID")) |>
  group_by(GeoUID, Care_Category, Care_Category_Specific) |>
  summarise(copp_isc_med = median(copp_isc), #the median number of opportunities that can be reached by specific category from each parcel in a DA
            copp_isc_sd = sd(copp_isc), #the SD
            copp_isc_w_med = median(copp_isc_w), #the median number of opportunities that can be reached by specific category from each parcel in a DA
            copp_isc_w_sd = sd(copp_isc_w), #the SD
            V_isc_w = sum(V_isc_w), #the number of total spatially available opportunities that can be reached by specific category from each parcel in a DA
            vv_isc = sum(V_isc_w)/first(Population), #the per-capita number of total spatially available opportunities that can be reached by specific category from each parcel in a DA
            Population = first(Population),
            Dwellings = first(Dwellings),
            LICO_AT_TOTAL = first(LICO_AT_TOTAL),
            COMMUNITY_ = first(COMMUNITY_) ) |>
  tidyr::drop_na(Population)

#checks
# access_15_DA |> summary()
access_15_bycategory_DA |> summary()
access_15_DA$V_w_iscDA |> sum(na.rm=T)
access_15_bycategory_DA$V_ic_w |> sum(na.rm=T)
access_15_byspecificcategory_DA$V_isc_w |> sum(na.rm=T)

#RECALL, categories are either 
# - "Dependent-centric" and include "Community or Recreation Centre", "Daycare or EarlyON", "Long-Term Care or Retirement Home", "Park", "School", "Senior Centre"
# - "Errand-centric" and include: "Bank or ATM", "Library", "Post Office"
# - "Grocery-centric" and include: "Convenience Store", "Grocer"
# - "Health-centric" and include: "Dentist", "Hospital or Clinic", "Pharmacy"
```
```{r}
access_15_byspecificcategory_DA |> group_by(Care_Category_Specific) |> summarise(V_isc_w = sum(V_isc_w))
care_dest |> st_drop_geometry() |> group_by(Care_Category_Specific) |> 
  summarise(Dest_count_weighted = sum(Dest_count_weighted),
            count = n())

```

```{r}
#calculating entropy, here I do it by the 14 specific categories. This is a way to consider their diversity within and between categories. Because they have been weighted before calc. the cumulative opp counts, we can proceed without anymore adjustments.
step1 <- access_15_byspecificcategory_DA  |> group_by(GeoUID) |> 
  summarise(OPPi = sum(copp_isc_med, na.rm=T) + sum(copp_isc_sd, na.rm=T),
            OPPi_w = sum(copp_isc_w_med, na.rm=T) + sum(copp_isc_w_sd, na.rm=T),
            Vi_w = sum(V_isc_w))

step2 <- access_15_byspecificcategory_DA |> left_join(step1, by="GeoUID")

step2 <- step2 |> mutate(top1_copp = (copp_isc_med+copp_isc_sd) / OPPi * log((copp_isc_med+copp_isc_sd) / OPPi),
             top1_copp = ifelse(is.nan(top1_copp), 0, top1_copp),
             top1_coppw = (copp_isc_w_med+copp_isc_w_sd) / OPPi_w * log((copp_isc_w_med+copp_isc_w_sd) / OPPi_w),
             top1_coppw = ifelse(is.nan(top1_coppw), 0, top1_coppw),
             top1_Vw= (V_isc_w) / Vi_w * log((V_isc_w) / Vi_w),
             top1_Vw = ifelse(is.nan(top1_Vw), 0, top1_Vw),)  |> 
  group_by(GeoUID) |>
  summarise(entropy_copp_isc_medsd = sum(top1_copp) / -log(14),
            entropy_copp_isc_medsd_w = sum(top1_coppw) / -log(14),
            entropy_V_isc_w = sum(top1_Vw) / -log(14))

#join this to access_15_bycategory_DA

access_15_bycategory_DA <- access_15_bycategory_DA |> left_join(step2, by="GeoUID")
```

```{r}
#the number of opportunities from each categories
all_cats <- access_15_bycategory_DA |> 
  mutate(copp_all = sum(copp_ic_med, na.rm=T) + sum(copp_ic_sd, na.rm=T),
         copp_all_w = sum(copp_ic_w_med, na.rm=T) + sum(copp_ic_w_sd, na.rm=T)) |>
  group_by(GeoUID) |>
         summarise(summed_copp_all = sum(copp_all),
                   summed_copp_all_w = sum(copp_all_w),
                   summed_Vic_w = sum(V_ic_w, na.rm=T),
                   Population = mean(Population),
                   Care_Access_Status1 = case_when(summed_copp_all < 4 ~ "[0-04 des.) Not caring",
                                 summed_copp_all >= 4 & summed_copp_all < 25 ~ "[04-25 des.) Caring", 
                                 summed_copp_all >= 25 ~ "[25+ des.) Very caring"),
                   Care_Access_Status1_w = case_when(summed_copp_all_w < 4 ~ "[0-04 des.) Not caring",
                                 summed_copp_all_w >= 4 & summed_copp_all_w < 25 ~ "[04-25 des.) Caring", 
                                 summed_copp_all_w >= 25 ~ "[25+ des.) Very caring"),

                   Care_Access_Status2 = case_when(all(copp_all >= 2 & copp_all < 4) ~ "- complete (2 to 4 options per cat.)", 
                                                   all(copp_all >= 4) ~ "- very complete (4 or more options per cat.)"),
                   Care_Access_Status2_w = case_when(all(copp_all_w >= 2 & copp_all_w < 4) ~ "- complete (2 to 4 options per cat.)", 
                                                   all(copp_all_w >= 4) ~ "- very complete (4 or more options per cat.)"),
                   entropy_copp_isc_medsd = first(entropy_copp_isc_medsd),
                   entropy_copp_isc_medsd_w = first(entropy_copp_isc_medsd_w),
                   entropy_V_isc_w = first(entropy_V_isc_w)) |>
  mutate(Care_Access_Status2 = ifelse(is.na(Care_Access_Status2), "", Care_Access_Status2),
         Care_Access_Status2_w = ifelse(is.na(Care_Access_Status2_w), "", Care_Access_Status2_w),
         Care_Access_Status = paste(Care_Access_Status1,Care_Access_Status2, sep=" "),
         Care_Access_Status_w = paste(Care_Access_Status1_w,Care_Access_Status2_w, sep=" "),
         vvi = summed_Vic_w/Population)

all_cats$summed_Vic_w |> sum() #only 2206  opps can be reached out of the total 2225, so that's what the numbers are based on.
care_dest|> nrow()

all_cats <- left_join(HAM_census_21 |> select(c(GeoUID)), all_cats, by=c("GeoUID"))
```

```{r message=FALSE}
#tmap_mode("plot")
all_cats_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("Care_Access_Status",
              border.alpha = 0.2,
              palette = c("darkred", "orangered", "orange", "yellow", "forestgreen", "lightgreen"),
              title = "Number of care destinations (all) spatially accessible within 15 min. walk",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12
            # legend.title.size = 0.7,
            # legend.text.size = 0.6
            )

tmap_save(all_cats_copp_plot, file="all_cats_copp_plot.tiff")

all_cats_copp_weighted_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("Care_Access_Status_w",
              border.alpha = 0.2,
              palette = c("darkred", "orangered", "orange", "yellow", "forestgreen", "lightgreen"),
              title = "Number of care destinations (all - weighted) spatially accessible within 15 min. walk",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12
            # legend.title.size = 0.7,
            # legend.text.size = 0.6
            )
tmap_save(all_cats_copp_weighted_plot, file="all_cats_copp_weighted_plot.tiff")
```

```{r}
all_cats_entropy_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("entropy_copp_isc_medsd",
              border.alpha = 0.2,
              palette = "Reds",
              title = "Diversity of (all) care destinations spatially accessible within 15 min. walk",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12)

tmap_save(all_cats_entropy_copp_plot, file="all_cats_entropy_copp_plot.tiff")

all_cats_entropy_copp_weighted_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("entropy_copp_isc_medsd_w",
              border.alpha = 0.2,
              palette = "Reds",
              title = "Diversity of (all - weighted) care destinations spatially accessible within 15 min. walk",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12)

tmap_save(all_cats_entropy_copp_weighted_plot, file="all_cats_entropy_copp_weighted_plot.tiff")
```

```{r}
all_cats_vvi_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("vvi",
              border.alpha = 0.2,
              palette = c("orangered", "forestgreen"),
              breaks = c(0, 0.003804, 0.022174),
              labels = c("Below Median", "Median and above"),
              title = "Per capita spatially available care destinations (all - weighted) within 15 min. walk",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12
            )

tmap_save(all_cats_vvi_plot, file="all_cats_vvi_plot.tiff")
```

```{r}
all_cats |> group_by(Care_Access_Status_w) |> summarise(min = min(summed_copp_all_w),
                                                        median = median(summed_copp_all_w),
                                                        max = max(summed_copp_all_w))
all_cats |> group_by(Care_Access_Status) |> summarise(min = min(summed_copp_all),
                                                        median = median(summed_copp_all),
                                                        max = max(summed_copp_all))
all_cats$entropy_copp_isc_medsd|> summary()
all_cats$entropy_copp_isc_medsd_w |> summary()
all_cats$Care_Access_Status_w |> unique()
```

```{r}
#adding entropy labels
all_cats <- all_cats |> 
  mutate(bivariate_entropy_w = 
           case_when( ((Care_Access_Status_w == "[04-25 des.) Caring - complete (2 to 4 options per cat.)"  |
                       Care_Access_Status_w =="[04-25 des.) Caring - very complete (4 or more options per cat.)" |
                         Care_Access_Status_w =="[25+ des.) Very caring - very complete (4 or more options per cat.)") &
                         entropy_copp_isc_medsd_w >= 0.7154) ~ "Caring, complete and amenity diverse",  #the median diversity,

                     ((Care_Access_Status_w == "[04-25 des.) Caring - complete (2 to 4 options per cat.)"  |
                       Care_Access_Status_w =="[04-25 des.) Caring - very complete (4 or more options per cat.)" |
                         Care_Access_Status_w =="[25+ des.) Very caring - very complete (4 or more options per cat.)") &
                        entropy_copp_isc_medsd_w < 0.7154) ~ "Caring, complete but not amenity diverse",

                     ((Care_Access_Status_w == "[0-04 des.) Not caring " |
                       Care_Access_Status_w == "[04-25 des.) Caring ")  &
                        entropy_copp_isc_medsd_w >= 0.7154) ~ "Not caring or not complete, but amenity diverse",

                     ((Care_Access_Status_w == "[0-04 des.) Not caring "  |
                       Care_Access_Status_w == "[04-25 des.) Caring ")  &
                        entropy_copp_isc_medsd_w < 0.7154) ~ "Not caring, complete, or amenity diverse"),

         bivariate_entropy = 
           case_when( ((Care_Access_Status == "[04-25 des.) Caring - complete (2 to 4 options per cat.)"  |
                       Care_Access_Status =="[04-25 des.) Caring - very complete (4 or more options per cat.)" |
                         Care_Access_Status =="[25+ des.) Very caring - very complete (4 or more options per cat.)") &
                         entropy_copp_isc_medsd >= 0.7418) ~ "Caring, complete and amenity diverse",  #the median diversity,

                     ((Care_Access_Status == "[04-25 des.) Caring - complete (2 to 4 options per cat.)"  |
                       Care_Access_Status =="[04-25 des.) Caring - very complete (4 or more options per cat.)" |
                         Care_Access_Status =="[25+ des.) Very caring - very complete (4 or more options per cat.)") &
                        entropy_copp_isc_medsd < 0.7418) ~ "Caring, complete but not amenity diverse",

                     ((Care_Access_Status == "[0-04 des.) Not caring " |
                       Care_Access_Status == "[04-25 des.) Caring ")  &
                        entropy_copp_isc_medsd >= 0.7418) ~ "Not caring or not complete, but amenity diverse",

                     ((Care_Access_Status == "[0-04 des.) Not caring "  |
                       Care_Access_Status == "[04-25 des.) Caring ")  &
                        entropy_copp_isc_medsd < 0.7418) ~ "Not caring, complete, or amenity diverse"))

```

```{r}
all_cats_bivariate_entropy_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("bivariate_entropy_w",
              border.alpha = 0.2,
              palette = c("yellowgreen","yellow","blue", "#FFFFDF"),
              title = "Diversity of (all - weighted) care destinations spatially accessible within 15 min. walk",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12)

tmap_save(all_cats_bivariate_entropy_copp_plot, file="all_cats_bivariate_entropy_copp_plot.tiff")
```

```{r}
all_cats_entropy_copp_weighted_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("entropy_copp_isc_medsd_w",
              border.alpha = 0.2,
              palette = "Reds", style="quantile", n=4,
              title = "Diversity of (all - weighted) care destinations spatially accessible within 15 min. walk",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12)

tmap_save(all_cats_entropy_copp_weighted_plot, file="all_cats_entropy_copp_weighted_plot.tiff")
```

```{r}
care_categories_plot <- 
 tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(Community_Boundaries, bbox=Community_Boundaries) + tm_polygons(col = "white", border.col = "black", lwd=1)+
  tm_shape(care_dest) + tm_dots("Care_Category_Specific", palette = RColorBrewer::brewer.pal(n = 4, name = 'Dark2'), size=0.025, title = "")+
  tm_facets("Care_Category", nrow = 2, free.scales= TRUE)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 20)) +
  tm_compass(position = c("left", "top"), size=0.9) +
  tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5, legend.bg.color = "white", legend.bg.alpha = 0.4,
            panel.labels = c("1,265 Dependent-centric destinations", 
                             "158 Errand-centric destinations", 
                             "381 Grocery-centric destinations", 
                             "421 Health-centric destinations"),
            panel.label.size = 0.8) 
care_categories_plot

tmap_save(care_categories_plot, filename ="care_categories_plot.tiff")
```
```{r}
parcels_DA_plot <- 
 tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA) +
  tm_polygons("parcel_count",
              border.alpha = 0.2,
              palette = "Reds",style = "quantile", n=4,
              title = "Count of residential parcels",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5,legend.title.size = 0.7,  legend.bg.color = "white", legend.bg.alpha = 0.4)

poprate_DA_plot <- 
 tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA |> mutate(popden = Population/Dwellings)) +
  tm_polygons("popden",
              border.alpha = 0.2,
              palette = "Purples", style = "quantile", n=4,
              title = "Population per dwelling",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5,legend.title.size = 0.7,  legend.bg.color = "white", legend.bg.alpha = 0.4)

LICOAT_DA_plot <- 
 tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA) +
  tm_polygons("LICO_AT_TOTAL",
              border.alpha = 0.2,
              palette = "Oranges", style = "quantile", n=4,
              title = "Prevalance of household LICO-AT",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5, legend.title.size = 0.7, legend.bg.color = "white", legend.bg.alpha = 0.4)

pop_descriptive_plots <- tmap_arrange(parcels_DA_plot, poprate_DA_plot, LICOAT_DA_plot, nrow=3)
tmap_save(pop_descriptive_plots, file="pop_descriptive_plots.tiff", asp = 1/0.6)
```

<!-- let's exclude this for now and focus on ALL categories -->
<!-- ```{r} -->
<!-- #the number of opportunities from each 'non dependent' categories -->
<!-- no_dependents <- access_15_bycategory_DA |> filter(Care_Category == "Errand-centric"| -->
<!--                                                       Care_Category == "Grocery-centric" | -->
<!--                                                       Care_Category == "Health-centric") |> -->
<!--   mutate(copp_nodeps = sum(copp_ic_med, na.rm=T) + sum(copp_ic_sd, na.rm=T)) |> -->
<!--   group_by(GeoUID) |> -->
<!--          summarise(summed_copp_nodeps = sum(copp_nodeps), -->
<!--                    summed_Vic = sum(V_ic, na.rm=T), -->
<!--                    Population = mean(Population), -->
<!--                    Care_Access_Status1 = case_when(summed_copp_nodeps < 3 ~ "[0-03 des.) Not caring", -->
<!--                                  summed_copp_nodeps >= 3 & summed_copp_nodeps <= 10 ~ "[03-10 des.) Caring", -->
<!--                                  summed_copp_nodeps > 10 ~ "[10+ des.) Very caring"), -->

<!--                    Care_Access_Status2 = case_when(all(copp_nodeps >= 2 & copp_nodeps <= 4) ~ "- complete (2 to 4 options per cat.)", -->
<!--                                                    all(copp_nodeps > 4) ~ "- very complete (4 or more options per cat.)")) |> -->
<!--   mutate(Care_Access_Status2 = ifelse(is.na(Care_Access_Status2), "", Care_Access_Status2), -->
<!--          Care_Access_Status = paste(Care_Access_Status1,Care_Access_Status2, sep=" "), -->
<!--          vvi = summed_Vic/Population) -->

<!-- no_dependents$summed_Vic |> sum() #only 952 of the no-dependent opps can be reached out of the 960, so that's what the numbers are based on. -->
<!-- care_dest |> filter(Care_Category == "Errand-centric"| -->
<!--                       Care_Category == "Grocery-centric" | -->
<!--                       Care_Category == "Health-centric") |> nrow() -->

<!-- no_dependents <- left_join(HAM_census_21 |> select(c(GeoUID)), no_dependents, by=c("GeoUID")) -->
<!-- ``` -->

<!-- ```{r message=FALSE} -->
<!-- #tmap_mode("plot") -->
<!-- no_dependents_copp_plot<-tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) + -->
<!--   tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+ -->
<!--   tm_shape(no_dependents) + -->
<!--   tm_polygons("Care_Access_Status", -->
<!--               border.alpha = 0.2, -->
<!--               palette = c("darkred", "orangered", "orange", "yellow", "forestgreen", "lightgreen"), -->
<!--               title = "Number of care destinations (no dependents) spatially accessible within 15 min. walk", -->
<!--               legend.is.portrait = TRUE) + -->
<!--   tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+ -->
<!--   tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) + -->
<!--   tm_compass(position = c("right", "top"), size=1.0)+ -->
<!--   tm_layout(bg.color = "grey", -->
<!--             panel.label.bg.color = "white", -->
<!--             panel.label.color = "black", -->
<!--             legend.outside = TRUE, -->
<!--             legend.position = c("center", "bottom"), -->
<!--             legend.outside.position = "bottom", -->
<!--             legend.outside.size = 0.2, -->
<!--             legend.width = 12 -->
<!--             # legend.title.size = 0.7, -->
<!--             # legend.text.size = 0.6 -->
<!--             ) -->
<!-- tmap_save(no_dependents_copp_plot, file="no_dependents_copp_plot.tiff") -->
<!-- ``` -->

<!-- ```{r, message=FALSE} -->
<!-- no_dependents_vvi_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) + -->
<!--   tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+ -->
<!--   tm_shape(no_dependents) + -->
<!--   tm_polygons("vvi", -->
<!--               border.alpha = 0.2, -->
<!--               palette = c("orangered", "forestgreen"), -->
<!--               breaks = c(0, 0.0013, 0.0148680), -->
<!--               labels = c("Below Median", "Median and above"), -->
<!--               title = "Per capita spatially available care destinations (no dependents) within 15 min. walk", -->
<!--               legend.is.portrait = TRUE) + -->
<!--   tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+ -->
<!--   tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) + -->
<!--   tm_compass(position = c("right", "top"), size=1.0)+ -->
<!--   tm_layout(bg.color = "grey", -->
<!--             panel.label.bg.color = "white", -->
<!--             panel.label.color = "black", -->
<!--             legend.outside = TRUE, -->
<!--             legend.position = c("center", "bottom"), -->
<!--             legend.outside.position = "bottom", -->
<!--             legend.outside.size = 0.2, -->
<!--             legend.width = 12 -->
<!--             # legend.title.size = 0.7, -->
<!--             # legend.text.size = 0.6 -->
<!--             ) -->

<!-- tmap_save(no_dependents_vvi_plot, file="no_dependents_vvi_plot.tiff") -->
<!-- ``` -->

<!-- NOTES FROM LEA CONVO FEB 22:

We changed some category tags. 
 - child and elder centric are now both 'dependent-centric'. These are 'caring for others'. We also consolidated some sub-categories. Daycare and EarlyON are now combined.. Long Term-Care Home and Residential Home are now combined.. Senior Centre and Senior Active Living Centre are combined. Hospital and Clinic are combined. Recreation Centre and COmmunity Centre are now combined.
 - A consideration for the future * maybe convenience stores and grocery stores can be be modulated based on their time -- i.e., if a DA trip to convenience is less than grocery, it is counted, if it is more than grocery it is not. More thought needs to be put into this, for now I did not include this consideration. 
 - We will consider the following 2 (or maybe 3?) faceted (by mode) map with scales:

1.	No-children (oneself and partner, no dependents) city map (15 min walk + 15 min cycle):
-	Errand, grocery, health

-	Scale: 
o	not-caring (does not have 1 of each category), 
o	caring (1 of each category), 
o	very caring (2 or more of each category), 
o	completely caring (1 of each sub-category), 
o	very completely caring (2 of each sub-category)

2.	Children+seniors: -- the truly caring city has all 4 categories. You always preform care but not always care for others. Your status will change. 
-	+elder+children
-	Scale: (same)

3. What neighbourhoods are ‘almost there’ for being completely caring? What are they missing?

And our presentation's angle: a 15 minute city is XYZ, the destinations are contested but typically are 'care' trips as well as discretionary trips (fun stuff). The mobility of care framework: paid, unpaid, and discretionary. All is important. But unpaid travel is under-studied. We are going to explore unpaid trips: they are essential and short-distance trips. A masculine city is one where paid work + discretionary trips are planned for. This analysis presents a tool to conceptual what is and is not a 'caring' or 'caring complete' neighbourhood. presents some visualisations.

APRIL 25 2024
- for simplicity, 15 min walk thresholds as a baseline for 15-min caring neighbourhoods. 15 min cycle thresholds to demonstrate the potential of cyling to get those 'almost there' neighbourhoods to be more caring if cycling infrastructure is implemented. Transit/bike to care isn't so common currently, but it's ane equity argument and shows political potential.

-->

```{r}
tmap_mode("plot")
tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(one_cat) +
  tm_polygons("Complete",
              border.alpha = 0,
              legend.is.portrait = FALSE,
              n=2,
              labels = c("Not complete", "Complete")) +
  tm_facets(by=c("Care_Category"), nrow=2, showNA = FALSE)+
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 20)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12,
            legend.title.size = 0.7,
            legend.text.size = 0.6)
```

Questions:

- what is a reasonable travel time? Is 15 minute walk indeed fine for our purpose?

- is 1 potential opportunity enough per subcategory?

- Should some subcategories be substitutes? E.g., 1 daycare OR EARLYON centre is enough, not 1 of each.

- Are there ways the categories can overlap to tell a story? Maybe to include google maps images. eg. 
  - communities where children+errand+grocery centric care is easier -- especially for women
  - health_centric+Elder_centric -- people with higher health care needs

- care trips 30% of all trips made. they are essential and have

<!-- ```{r} -->
<!-- test1 <- ttm_care_walk_15 |> filter(Care_Category == "Child_Centric") |> -->
<!--   sp_avail_detailed(o_id = from_id, -->
<!--                     d_id = to_id,  -->
<!--                     pop = Population,  -->
<!--                     opp = f, -->
<!--                     r = f,  -->
<!--                     f = f, alpha = 1)|> -->
<!--   group_by(from_id, Care_Category) |> summarize(V_ic = sum(V_ij, na.rm = TRUE))  -->

<!-- test2 <- ttm_care_walk_15 |> filter(Care_Category == "Grocery_Centric") |> -->
<!--   sp_avail_detailed(o_id = from_id, -->
<!--                     d_id = to_id,  -->
<!--                     pop = Population,  -->
<!--                     opp = f, -->
<!--                     r = f,  -->
<!--                     f = f, alpha = 1)|> -->
<!--   group_by(from_id, Care_Category) |> summarize(V_ic = sum(V_ij, na.rm = TRUE))  -->

<!-- test3 <- ttm_care_walk_15 |> filter(Care_Category == "Elder_Centric") |> -->
<!--   sp_avail_detailed(o_id = from_id, -->
<!--                     d_id = to_id,  -->
<!--                     pop = Population,  -->
<!--                     opp = f, -->
<!--                     r = f,  -->
<!--                     f = f, alpha = 1)|> -->
<!--   group_by(from_id, Care_Category) |> summarize(V_ic = sum(V_ij, na.rm = TRUE))  -->

<!--  test4 <- ttm_care_walk_15 |> filter(Care_Category == "Errand_Centric") |> -->
<!--   sp_avail_detailed(o_id = from_id, -->
<!--                     d_id = to_id,  -->
<!--                     pop = Population,  -->
<!--                     opp = f, -->
<!--                     r = f,  -->
<!--                     f = f, alpha = 1)|> -->
<!--   group_by(from_id, Care_Category) |> summarize(V_ic = sum(V_ij, na.rm = TRUE))  -->

<!-- test5 <- ttm_care_walk_15 |> filter(Care_Category == "Health_Centric") |> -->
<!--   sp_avail_detailed(o_id = from_id, -->
<!--                     d_id = to_id,  -->
<!--                     pop = Population,  -->
<!--                     opp = f, -->
<!--                     r = f,  -->
<!--                     f = f, alpha = 1)|> -->
<!--   group_by(from_id, Care_Category) |> summarize(V_ic = sum(V_ij, na.rm = TRUE))  -->

<!-- test <- rbind(test1, test2, test3, test4, test5) -->

<!-- access_15 <- access_15 |> left_join(test, by=c("from_id", "Care_Category")) |> ungroup() -->

<!-- rm(test, test1, test2, test3, test4, test5) -->
<!-- ``` -->

### 2024-05-16 WGiT/WGiT-Proposal.tex
Tidying up the repo

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  authoryear,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{graphicx}
\usepackage{unicode-math}
\usepackage{hyperref}
\def\tightlist{}
\usepackage{setspace}
\doublespacing
\usepackage{lineno}
\linenumbers
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{7th International Conference on Women and Gender in Transportation (WGiT)}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Towards caring 15-minute neighbourhoods},
  pdfauthor={Anastasia Soukhov; Léa Ravensbergen; Lucía Mejía Dorantes; Antonio Páez},
  pdfkeywords={Accessibility, Mobility of Care, Gender, 15-Minute
City, Care Complete Neighbourhoods},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Towards caring 15-minute neighbourhoods}
\author[1]{Anastasia Soukhov%
\corref{cor1}%
}
 \ead{soukhoa@mcmaster.ca} 
\author[1]{Léa Ravensbergen%
%
}
 \ead{ravensbl@mcmaster.ca} 
\author[2]{Lucía Mejía Dorantes%
%
}
 \ead{lucia.mejia.dorantes@gmail.com} 
\author[1]{Antonio Páez%
%
}
 \ead{paezha@mcmaster.ca}

\affiliation[1]{organization={McMaster University, School of Earth,
Environment \& Society},city={Hamilton, Canada},postcodesep={}}
\affiliation[2]{organization={Consultant},city={Karlsruhe,
Germany},postcodesep={}}

\cortext[cor1]{Corresponding author}

\begin{keyword}
    Accessibility \sep Mobility of Care \sep Gender \sep 15-Minute
City \sep 
    Care Complete Neighbourhoods
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, enhanced, sharp corners, breakable, borderline west={3pt}{0pt}{shadecolor}, frame hidden]}{\end{tcolorbox}}\fi

The 15-Minute City is a normative conceptualisation gaining ground in
urban planning: it frames neighbourhoods as responsive to human needs
and environmental sensibilities, where most daily necessities can be
reached within a 15-minute walk or bike ride \citep{Allam2022}. As a
related tool, accessibility measures (the ease of reaching
opportunities) are increasingly important amongst transport planners
aiming to foster just and sustainable cities
\citep{vale2023accessibility}. Both the 15-Minute City and accessibility
measures are flexible enough to consider all destination types
holistically however, gendered examinations have been lacking in the
literature. For instance, accessibility analyses have historically
focused on employment-centric and discretionary travel, types of travel
more frequent for working-age and higher-income men.

To counter this masculinist bias, this study investigates a way to
gender-mainstream the 15-Minute City through a care lens
\citep[e.g.,][]{lawWomenTransportNew1999, uteng2008gendered, Oestergaard1992, levy1991towards, levy2013travel}
supported by the Mobility of Care framework (coined by
\citet{sanchezdemadariagaMobilityCareIntroducing2013}). Mobility of Care
emphasizes the importance of travel to unpaid work (care trips) in
contrast to the better-studied travel to employment and leisure. While
all three trip types (work, care, and leisure) are essential, care trips
are often relatively shorter-distance, proximate to
residential/work/school, and comprise approximately 30\% of adults'
daily trips
\citep{sanchezdemadariagaMeasuringMobilitiesCare2019, ravensbergen2023exploratory, mejia2021mobility};
fitting well within the 15-Minute City conceptualisation.

Our study provides an empirical example that maps the 15-Minute City
onto the Mobility of Care framework. Specifically, it identifies which
areas in Hamilton, Canada are `caring 15-minute neighbourhoods'. To do
so, a database of care destinations is created using secondary data. In
this database, care destinations include all places associated with
sustaining household tasks needed for the reproduction of life namely:
shopping (e.g., groceries), errands (e.g., libraries), health (e.g.,
dentist), and caring for dependents (e.g., schools). This database is
used to estimate the number and mix of care-destinations that can be
reached within a 15-minute walk- and cycling- sheds from census
centroids. Typologies are generated to illustrate which neighbourhoods
can and to what degree facilitate 15-minute access to care.

Results indicate only a few neighbourhoods outside of the downtown core
are `care-complete', i.e., contain a sufficiently high mix of care
destinations from all categories and sub-categories. However, some
neighbourhoods are almost `care-complete' and provide 15-minute access
to some care categories. Our study frames these neighbourhoods on the
continuum of `caring' and in need of further intervention. The
quantitative investigation conducted provides a high-level picture of
what neighbourhoods (and their underlying land-use) are connected to
transport infrastructure that can support reaching care-destinations.

Taken together, this study provides a theoretical bridge to connect
15-Minute Cities, accessibility analysis and Mobility of Care framework
for the purpose of informing policy choice aimed to encourage just and
sustainable mobility.

\renewcommand\refname{References}
  \bibliography{bibliography.bib}

\end{document}

### 2024-05-16 paper/Manuscript.qmd
Tidying up the repo

--- 2024-05-16 Manuscript.qmd
@@
 ---
-```{r setup, include=FALSE}
+```{r}
+#| label: setup-knitr-options
+#| include: false
+
 knitr::opts_chunk$set(echo = FALSE,
@@
                       cache=FALSE)
+
+```
+```{r }
+#| label: set-up-copy-bibliography
+#| include: false
+
+# WARNING: DO NOT MODIFY THE BIBLIOGRAPHY ON THIS FOLDER, WORK WITH THE ONE IN FOLDER ./bibliography/
+file.copy(glue::glue(here::here(), "/bibliography/bibliography.bib"), 
+          getwd(),
+          overwrite = TRUE)
 ```
@@
 ```{r}
-#function for creating comments
+#| label: function for creating comments
+
 word_comment <- function(comment, highlight = "") {
@@
 library(cowplot)
-library(sf)
-library(here)
 library(dplyr)
-library(tmap)
-library(mapview)
-library(tmaptools)
-library(pander)
 library(flextable)
 library(ftExtra)
+library(glue)
+library(here)
+library(mapview)
+library(pander)
 library(scales)
+library(sf)
+library(tmap)
+library(tmaptools)
 ```
@@
 ```{r}
-load("data/care_dest.rda")
+load(glue::glue(here::here(), "/data/care_dest.rda"))
 ```
@@
 Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics("data/pop_descriptive_plots.tiff")
+knitr::include_graphics(glue::glue(here::here(), "/data/pop_descriptive_plots.tiff"))
 ```
@@
 Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics("data/care_categories_plot.tiff")
+knitr::include_graphics(glue::glue(here::here(), "/data/care_categories_plot.tiff"))
 ```
@@
 Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics("data/all_cats_copp_weighted_plot.tiff")
+knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_copp_weighted_plot.tiff"))
 ```
@@
 Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics("data/all_cats_entropy_copp_weighted_plot.tiff")
+knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_entropy_copp_weighted_plot.tiff"))
 ```
@@
 The Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics("data/all_cats_bivariate_entropy_copp_plot.tiff")
+knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_bivariate_entropy_copp_plot.tiff"))
 ```
@@
 Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics("data/all_cats_vvi_plot.tiff")
+knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_vvi_plot.tiff"))
 ```
@@
 Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics("data/all_cats_copp_plot.tiff")
+knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_copp_plot.tiff"))
 ```
@@
 Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics("data/all_cats_entropy_copp_plot.tiff")
+knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_entropy_copp_plot.tiff"))
 ```

### 2024-05-24 data/data-raw/SOM-share/Session 10 - Self Organizing Maps III.Rmd
added SOM notebooks and edits to the manuscript.qmd (from this morning's team meeting regarding deliverables 2)

---
title: "Artificial Neural Networks - Session 10: Self Organizing Maps III"
---

# Introduction

Previously we used a toy example to illustrate the application of self-organizing maps. In this session, we will now use a real-world example to develop a customer segmentation system.

This example uses data from Ireland used by Shane Lynn in this [blogpost](https://www.shanelynn.ie/self-organising-maps-for-customer-segmentation-using-r/). The data are sourced from the Irish Census of 2011 and includes the Dublin area. For this session, we will continue working with the package `SOMbrero`.

As usual, we will begin by clearing the workspace:
```{r}
rm(list = ls())
```

And also loading the packages used for this session:
```{r message = FALSE}
library(AppliedPredictiveModeling) # Functions and Data Sets for 'Applied Predictive Modeling'
library(caret) # Classification and Regression Training
library(formattable) # Create 'Formattable' Data Structures
library(kableExtra) # Construct Complex Table with 'kable' and Pipe Syntax
library(knitr) # A General-Purpose Package for Dynamic Report Generation in R
library(plotly) # Create Interactive Web Graphics via 'plotly.js'
library(rpart) # Recursive Partitioning and Regression Trees
library(rpart.plot) # Plot 'rpart' Models: An Enhanced Version of 'plot.rpart'
library(sf) # Simple Features for R
library(spdep) # Spatial Dependence: Weighting Schemes, Statistics
library(SOMbrero) # SOM Bound to Realize Euclidean and Relational Outputs
library(tidyverse) # Easily Install and Load the 'Tidyverse'
library(tmap) # Thematic Maps
```

# Load and preprocess example data

```{r}
data_raw <- read.csv("data/Selected Themes.csv")
```

Use the following function to preprocess data, for instance to convert from totals to percentages, averages, or rates, as appropriate:
```{r}
convertCSOdata <- function(data, idcol="GEOGDESC"){
  results <- data.frame(id = data[[idcol]])

  # extract weighted average age.
  age_data <- data[,74:108]
  mean_ages <- c(0:19, 22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87)
  results$avr_age <- apply(age_data, MARGIN=1, function(x){
    #want this function to calculate the average weighted age
    # weighted mean age = E(num_people in age bin * average age of bin) / total_people
    age <- sum(x[1:length(mean_ages)] * mean_ages) / x[length(x)]
  })

  # average household size
  household_data <- data[, 331:339]
  mean_household_size <- c(1:8)
  results$avr_household_size <- apply(household_data, 1, function(x){
    size <- sum(x[1:length(mean_household_size)] * mean_household_size) / x[length(x)]
  })

  # Average education level
  education_data <- data[,622:634]
  mean_education_level <- c(0:10) # the 11th column of data is "not stated" answers - to be removed
  results$avr_education_level <- apply(education_data, 1, function(x){
    # need to remove the not-stated answers from the total - thus measuring only average on only people who answered questions
    education <- sum(x[1:length(mean_education_level)] * mean_education_level) / (x[length(x)] - x[length(x)-1]) 
  })

  # Average number of cars per household
  car_data <- data[754:758]
  num_cars <- c(0:4) # the number of cars per household for each column of temp data
  results$avr_num_cars <- apply(car_data, 1, function(x){
    # need to remove the not-stated answers from the total - thus measuring only average on only people who answered questions  
    cars <- sum(x * num_cars) / (sum(x))
  })

  # Average health reported
  health_data <- data[,696:702]
  mean_health <- c(5:1) # the 6th column of data is "not stated" answers - to be removed
  results$avr_health <- apply(health_data, 1, function(x){
    # need to remove the not-stated answers from the total - thus measuring only average on only people who answered questions
    health <- sum(x[1:length(mean_health)] * mean_health) / (x[length(x)] - x[length(x)-1]) 
  })

  # Percentage rented accomodation = Rented from private landlord / (total - not_stated)
  rent_data <- data[, c("T6_3_RPLH", "T6_3_NSH", "T6_3_TH")]
  results$rented_percent <- (rent_data[, 1] / (rent_data[, 3] - rent_data[,2])) * 100

  # Average unemployment = (looking_for_first_job + lost_job) / (total - disability)
  employment_data <- data[, c("T8_1_LFFJT", "T8_1_ULGUPJT", "T8_1_UTWSDT", "T8_1_TT")]
  results$unemployment_percent <- ((rowSums(employment_data[, 1:2])) / (employment_data[, 4] - employment_data[,3])) * 100

  # Average internet penetration = (broadband + other) / (total - not_stated)
  internet_data <- data[, c("T15_3_B", "T15_3_OTH", "T15_3_NS", "T15_3_T")]
  results$internet_percent <- ((rowSums(internet_data[, 1:2])) / (internet_data[, 4] - internet_data[,3])) * 100

  # Single Percent
  # Married Percent
  # Divorced Percent
  # Widowed Percent
  marital_data <- data[, c("T1_2SGLT", "T1_2MART", "T1_2SEPT", "T1_2DIVT", "T1_2WIDT", "T1_2T")]
  marital_percents <- data.frame(t(apply(marital_data, 1, function(x) {x[1:5]/x[6]})) * 100)
  names(marital_percents) <- c("single_percent", "married_percent", "separated_percent", "divorced_percent", "widow_percent")  
  results <- cbind(results, marital_percents)

  return(results)
}
```

Using the function defined above process the data and print a summary:
```{r}
data <- convertCSOdata(data_raw, idcol = "GEOGDESC")
data <- rename(data, 
               SMALL_AREA = id)
summary(data)
```

Geospatial data:
```{r}
ireland_map <- st_read("data/Census2011_Small_Areas_generalised20m.shp")

# Join the tabular data
ireland_map <- ireland_map |>
  left_join(data,
            by = "SMALL_AREA") |>
  st_as_sf()

ggplot() +
  geom_sf(data = ireland_map)
```

Extract only the counties corresponding to the Dublin area:
```{r}
ireland_map <- ireland_map |>
  filter(COUNTYNAME %in% c("Fingal", "Dublin City", "South Dublin", "Dn Laoghaire-Rathdown"))
```

This is the map of the small areas in the Dublin area: 
```{r}
ggplot() +
  geom_sf(data = ireland_map)
```

It might be useful to visualize some of these variables. The following code creates a `tmap` object:, essentially a choropleth map of the variable specified:
```{r fig.height= 8}
map <- tm_shape(ireland_map) +
  tm_polygons("internet_percent")
```

The map can be displayed statically, or dynamically, depending on the `tmap_mode` selected. For static maps use "plot" and for dynamic maps use "view":
```{r plot-map, message=FALSE}
tmap_mode("view")
map
```

# Train SOM

## Simple example with only two variables

To illustrate SOMs in two dimensions, we will use only two of the variables in the dataframe, `avr_age` and `avr_education_level`:
```{r}
df <- ireland_map |>
  st_drop_geometry() |>
  select(avr_age,
             avr_education_level)
```

```{r}
ggplot(data = df, 
       aes(x = avr_age, 
           y = avr_education_level)) + 
  geom_point()
```

Given these two attributes, a SOM can be trained using the function `trainSOM` from `SOMbrero`. Only one argument is needed, the input data `x.data`. Other arguments can be assigned by default. In what follows the dimension of the map is selected in such a way that the number of cells is $5\sqrt n$, and I request that 100 intermediate steps be saved. 
```{r cache = TRUE}
set.seed(593)
# set size of SOM
nrows <- 17
ncols <- 17
# initialize
#initSOM(maxit = 5000)
# run the SOM algorithm 
som1 <- trainSOM(x.data = df, 
                 dimension=c(nrows, ncols), 
                 nb.save = 25, radius.type = "letremy", maxit = 5000, verbose = TRUE)
```

Remember that the `somRes` object includes the following items:

- clustering: the results of the clustering algorithm, which gives each observation a number that corresponds to their best matching unit (BMU).

- prototypes: the final weights of the cells.

- energy: the final energy.

- backup: a list that includes the prototypes (the weights) at intermediate steps, the steps at which backups were recorded, intermediate clustering results, etc.

Since backups were requested, it is possible to plot the evolution of the training algorithm. For instance, the following plot shows the energy:
```{r}
plot(som1, what = "energy")
```

Which can be visualized by means of a _hitmap_:
```{r}
plot(som1, what = "obs", type = "hitmap")
```

```{r}
plot(som1, what = "prototypes", type = "mds")
```

```{r}
som.grid <- cell2nb(nrow = nrows, ncol = ncols, type = "rook")
```

```{r}
som.grid.final <- data.frame(x1 = numeric(length(unlist(som.grid))))
count <- 0
for(i in 1:(nrows * ncols)){
  for(j in 1:length(som.grid[[i]])){
    count <- count + 1
    som.grid.final$x1[count] <- som1$prototypes[i,1]
    som.grid.final$y1[count] <- som1$prototypes[i,2]
    som.grid.final$x2[count] <- som1$prototypes[som.grid[[i]][j],1]
    som.grid.final$y2[count] <- som1$prototypes[som.grid[[i]][j],2]
  }
}  
```

```{r}
ggplot(data = data.frame(node = 1:ncols * nrows, som1$prototypes), 
       aes(x = avr_age, y = avr_education_level)) +
  geom_segment(data = som.grid.final, 
               aes(x = x1, y = y1, xend = x2, yend = y2)) +
  geom_point(aes(color = node), size = 4) +
  geom_point(shape = 1, size = 4) +
  scale_color_distiller(palette = "RdYlBu")
```

## More attributes

Lets now try this but with the full array of numeric stats:
```{r}
df <- ireland_map |>
  st_drop_geometry() |>
  select(avr_age:widow_percent)
```

A rule of thumb for selecting the size of the SOM is $5\sqrt n$ neurons. Since the size of the sample is `r nrow(data)`, this means a square with `r round(sqrt(5 * sqrt(nrow(data))))`:
```{r}
set.seed(593)
# run the SOM algorithm 
nrows <- 17
ncols <- 17
som2 <- trainSOM(x.data = df, dimension=c(nrows, ncols), radius.type = "letremy", verbose = TRUE)
```

This gives the summary for the map:
```{r}
summary(som2)
```

The results of ANOVA indicate that all variables contribute to discrimination in the clustering.

The following shows the number of cases assigned to each node:
```{r}
table(som2$clustering)
```

Which can be visualized by means of a _hitmap_:
```{r}
plot(som2, what = "obs", type = "hitmap")
```

The hitmap does not reveal empty cells, that is, neurons that did not classify a single observation.

# tools for exploring the results

Interpretation of clusters typically involves descriptive statistics of the clusters. `SOMbrero` implements a number of graphical tools to support interpretation.

The first example is a heatmap of the variables in clusters:
```{r fig.width= 10}
plot(som2, what="obs", type="color", variable=1, print.title=TRUE, 
     main="avg_age")
```

# Create super clusters

The clusters obtained from the SOM are perhaps too numerous to be of practical use. It is then convenient to further group them to obtain super clusters. This is easily done in `SOMbrero` by means of the function `superClass`, which takes the results of `trainSOM` and applies a hierarchical clustering algorithm to reduce the number of classes.

In this example, I select $k=8$ for the number of super clusters:
```{r}
som2.sc <- superClass(som2, k=8)
summary(som2.sc)
```
It can be seen that all variables retain their significance. The frequency table, on the other hand, shows how many of the original clusters have been assigned to each of the super clusters.

The hierarchical clustering results can be visualized as follows:
```{r fig.width=12}
plot(som2.sc, plot.var=FALSE)
```

And also plotted in the grid.
```{r}
plot(som2.sc, type="grid", plot.legend=TRUE)
```

# Profiling the results

Once that clusters have been obtained, it is interesting to profile the results. By profiling the results, we can derive insights as to the (hopefully) distinctive characteristics of the clusters. For instance, is Cluster 1 mostly composed of young people? People with internet access?

Explore the results. First append the results of the cluster and super cluster analysis to dataframe:
```{r}
ireland_map$cluster.som2 <- som2$clustering
ireland_map$scluster.som2 <- som2.sc$cluster[som2$clustering]
```

Select four of the features with the greatest discriminatory power, and do a feature plot (from package `caret`):
```{r fig.width=16}
transparentTheme(trans = .4)
featurePlot(x = data.frame(ireland_map$avr_education_level, 
                           ireland_map$avr_num_cars, 
                           ireland_map$rented_percent, 
                           ireland_map$internet_percent), 
            y = as.factor(ireland_map$scluster.som2), 
            plot = "pairs",
            ## Add a key at the top
            auto.key = list(columns = 8))
```

Descriptive statistics of superclusters.

First group data by super cluster:
```{r}
super.clusters <- ireland_map |>
  group_by(scluster.som2)
```

Summarize profiles by super cluster:
```{r}
profiles <- summarize(super.clusters, 
                      Age = mean(avr_age), 
                      `HH Size` = mean(avr_household_size),
                      Cars = mean(avr_num_cars),
                      Health = mean(avr_health),
                      `Pct Rent` = mean(rented_percent),
                      `Pct Unemployed` = mean(unemployment_percent),
                      `Pct Internet` = mean(internet_percent),
                      Single = mean(single_percent),
                      Married = mean(married_percent),
                      Separated = mean(separated_percent),
                      Divorced = mean(divorced_percent),
                      Widow = mean(widow_percent))
profiles <- rename(profiles, `Super Cluster` = scluster.som2)
```

Format table and print:
```{r}
profiles %>%
  format.data.frame(digits = 3) %>%
  mutate(Age = cell_spec(Age, "html", color = ifelse(Age < mean(data$avr_age), "red", "blue")),
         `HH Size` = cell_spec(`HH Size`, "html", color = ifelse(`HH Size` < mean(data$avr_household_size), "red", "blue")),
         Cars = cell_spec(Cars, "html", color = ifelse(Cars < mean(data$avr_num_cars), "red", "blue")),
         Health = cell_spec(Health, "html", color = ifelse(Health < mean(data$avr_health), "red", "blue")),
         `Pct Rent` = cell_spec(`Pct Rent`, "html", color = ifelse(`Pct Rent` < mean(data$unemployment_percent), "red", "blue")),
         `Pct Unemployed` = color_bar("lightpink")(`Pct Unemployed`),
         `Pct Internet` = cell_spec(`Pct Internet`, "html", color = ifelse(`Pct Internet` < mean(data$internet_percent), "red", "blue")),
         Single = cell_spec(Single, "html", color = ifelse(Single < mean(data$single_percent), "red", "blue")),
         Married = cell_spec(Married, "html", color = ifelse(Married < mean(data$married_percent), "red", "blue")),
         Separated = cell_spec(Separated, "html", color = ifelse(Separated < mean(data$separated_percent), "red", "blue")),
         Divorced = cell_spec(Divorced, "html", color = ifelse(Divorced < mean(data$divorced_percent), "red", "blue")),
         Widow = cell_spec(Widow, "html", color = ifelse(Widow < mean(data$widow_percent), "red", "blue"))) %>%
  kable("html", escape = F, digits = 2) %>%
  column_spec(1:13, width = 16) %>%
  kable_styling("striped", full_width = F)
```

Map the superclusters:
```{r}
map2 <- tm_shape(ireland_map) +
  tm_polygons("scluster.som2")
map2
```

# Assessing the quality of the classifier

A thorny issue in unsupervised learning is how to assess the quality of a clustering algorithm. Say that you train a few different classifiers, either using different methods or with different combinations of variables. How do you decide which to use?

One possibility is to analyze the results of the clustering algorithm by means of a classification technique. This is illustrated here by means of a decision tree. A decision tree is a supervised learning technique, which in this case can use the results of the clustering algorithm as labels, and the input variables as features. If the clustering algorithm performed well, class membership must be relatively homogeneous. Since a decision tree

Select data for classification:
```{r}
#df <- select(data, -c(SMALL_AREA, cluster.som2))
df <- ireland_map |>
  st_drop_geometry() |>
  transmute(`Super Cluster` = as.factor(scluster.som2),
                Age = avr_age,
                `HH Size` = avr_household_size,
                Cars = avr_num_cars,
                Health = avr_health,
                `Pct Rent` = rented_percent,
                `Pct Unemployed` = unemployment_percent,
                `Pct Internet` = internet_percent,
                Single = single_percent,
                Married = married_percent,
                Separated = separated_percent,
                Divorced = divorced_percent,
                Widow = widow_percent)
```

Train a decision tree using the data:
```{r}
classmod.som2 <- rpart(`Super Cluster` ~., data = df)
```

Plot results of the decision tree:
```{r fig.width= 12}
rpart.plot(classmod.som2, cex = 0.75, box.palette = 0, type = 3)
```

## Train a SOM using only the variables identified by the Decision Tree

Lets now try this but with the full array of numeric stats:
```{r}
df <- ireland_map |>
  st_drop_geometry() |>
    transmute(Age = avr_age,
                Cars = avr_num_cars,
                Health = avr_health,
                `Pct Rent` = rented_percent,
                `Pct Unemployed` = unemployment_percent,
                `Pct Internet` = internet_percent,
                Separated = separated_percent)
```

```{r}
set.seed(593)
# run the SOM algorithm 
nrows <- 17
ncols <- 17
som3 <- trainSOM(x.data = df, dimension=c(nrows, ncols), radius.type = "letremy", verbose = TRUE)
```

This gives the summary for the map:
```{r}
summary(som3)
```

The clusters obtained from the SOM are perhaps too numerous to be of practical use. It is then convenient to further group them to obtain super clusters. This is easily done in `SOMbrero` by means of the function `superClass`, which takes the results of `trainSOM` and applies a hierarchical clustering algorithm to reduce the number of classes.

In this example, I select $k=8$ for the number of super clusters:
```{r}
som3.sc <- superClass(som3, k=8)
summary(som3.sc)
```
It can be seen that all variables retain their significance. The frequency table, on the other hand, shows how many of the original clusters have been assigned to each of the super clusters.

The clustering results can be visualized as follows
```{r}
plot(som3.sc, type="grid", plot.legend=TRUE)
```

Explore the results. First append the results of the cluster and super cluster analysis to dataframe:
```{r}
ireland_map$cluster.som3 <- som3$clustering
ireland_map$scluster.som3 <- som3.sc$cluster[som3$clustering]
```

```{r}
#df <- select(data, -c(SMALL_AREA, cluster.som2))
df <- ireland_map |>
  st_drop_geometry() |>
  transmute(`Super Cluster` = as.factor(scluster.som2),
                Age = avr_age,
                Cars = avr_num_cars,
                Health = avr_health,
                `Pct Rent` = rented_percent,
                `Pct Unemployed` = unemployment_percent,
                `Pct Internet` = internet_percent,
                Separated = separated_percent)
```

Train a decision tree using the data:
```{r}
classmod.som3 <- rpart(`Super Cluster` ~., data = df)
```

Plot results of the decision tree:
```{r fig.width= 12}
rpart.plot(classmod.som3, cex = 0.75, box.palette = 0, type = 3)
```

# Alternative table format

Format table and print:
```{r}
profiles %>%
  format.data.frame(digits = 3) %>%
  mutate(Age = cell_spec(Age, "html", color = ifelse(Age < mean(data$avr_age), "red", "blue")),
         `HH Size` = cell_spec(`HH Size`, "html", color = ifelse(`HH Size` < mean(data$avr_household_size), "red", "blue")),
         Cars = cell_spec(Cars, "html", color = ifelse(Cars < mean(data$avr_num_cars), "red", "blue")),
         Health = cell_spec(Health, "html", color = ifelse(Health < mean(data$avr_health), "red", "blue")),
         `Pct Rent` = cell_spec(`Pct Rent`, "html", color = ifelse(`Pct Rent` < mean(data$unemployment_percent), "red", "blue")),
         `Pct Unemployed` = cell_spec(`Pct Unemployed`, "html", color = ifelse(`Pct Unemployed` < mean(data$unemployment_percent), "red", "blue")),
         `Pct Internet` = cell_spec(`Pct Internet`, "html", color = ifelse(`Pct Internet` < mean(data$internet_percent), "red", "blue")),
         Single = cell_spec(Single, "html", color = ifelse(Single < mean(data$single_percent), "red", "blue")),
         Married = cell_spec(Married, "html", color = ifelse(Married < mean(data$married_percent), "red", "blue")),
         Separated = cell_spec(Separated, "html", color = ifelse(Separated < mean(data$separated_percent), "red", "blue")),
         Divorced = cell_spec(Divorced, "html", color = ifelse(Divorced < mean(data$divorced_percent), "red", "blue")),
         Widow = cell_spec(Widow, "html", color = ifelse(Widow < mean(data$widow_percent), "red", "blue"))) %>%
  kable("html", escape = F, digits = 2) %>%
  kable_styling("striped", full_width = F)
```

### 2024-05-24 data/data-raw/SOM-share/Session 8 - Self Organizing Maps.Rmd
added SOM notebooks and edits to the manuscript.qmd (from this morning's team meeting regarding deliverables 2)

---
title: "Artificial Neural Networks - Session 8: Self Organizing Maps"
---

# Introduction

In previous sessions we have considered artificial neural networks with a few different network architectures. The architecture of the network is one of the most important aspects of developing a model based on the neural network approach. For instance, _multilayer perceptron networks_ (MPN) consist of one input layer, one or more hidden layers, and one output layer. _Convolutional neural networks_ use a different architecture that is suitable for multi-channel, multi-dimensional data, such as images.

A common aspect of these networks is the way they learn.

If we recall the architecture of networks of this type, they are in general terms as seen in Figure 1.

![Figure 1. Schematic feedforward network](figures/Figure 1.jpg)

Figure 1 is a schematic only, since the input layer in a convolutional neural network will not be a dense layer, but a data convolution. Nonetheless, at the end of the network, the output is always an estimate of a known value, irrespective of whether the inputs are images (in the case of convolutional neural networks) or tabular data (in the case of multiperceptron networks). For instance, in handwritting recognition, the output will be a probability that the input is of a certain class (say, a digit between 0 and 9). In spatial interpolation it will be an estimated value of an observation.

As seen in Figure 1, the network is trained on the data $x$ with the objective of producing an estimated value of $y$, say $\hat{y}$, that will minimize a loss function.

Training of these networks involves feeding data forward across the network, and then backpropagating the error to adjust the parameters of the model. In this way, we have worked with neural networks in a _supervised learning_ context.

There are classes of problems, on the other hand, that are not amenable to supervised learning approaches. Classification, for instance, is an important problem where there are no "outcomes" that can be used to train a model.

You might be already familiar with a number of classifications approaches when the "labels" or "outcomes" of the observations are not known, but want to be retrieved from the data. These include clustering algorithms such as k-means, k-modes, hierarchical clustering, etc. The idea behind these algorithms is to arrange inputs according to their similarity, and are therefore also considered dimensional reduction techniques: by reducing multiple dimensions to only one or two it becomes possible to visualy inspect relationships in complex datasets.

![Figure 2. Dimensional Reduction: data points in higher dimensions are reduced to a 2 dimensional representation](figures/Figure 2.jpg)

The objective of this seminar is to introduce an artificial neural network architecture used for dimension reduction and classification in an unsupervised learning context, the Self Organizing Map.

# The Self Organizing Map

The concept of Self Organizing Maps is due to T. Kohonen circa 1981. The architecture of the network is either a string of neurons for one dimensional maps, or a tesselation of neurons for two dimensional maps (see Figure 3).

![Figure 3. One and two dimensional maps](figures/Figure 3.jpg)

In theory, higher dimensional maps are possible, although they defeat the purpose of providing a visually convenient representation of similiraties among the inputs.

Each of the cells in a map, like other neural networks, contains a set of parameters or _weights_. At first glance, the Self Organizing Map may resemble a convolutional neural network. This is only a superficial similarity.

The inputs for a Self Organizing Map are tabular data of the form:
$$
\textbf{x}_i=[x_{i1}, x_{i2},...,x_{ik}],
$$
where $\textbf{x}$ is a vector of size $1\times k$.

Similar to feeding the data in batches to a feedback neural network, inputs are passed on to a Self Organizing Map in sequential fashion, in principle with a batch size of one, although larger batches are possible too [Kohonen, 1998](https://www.sciencedirect.com/science/article/pii/S0925231298000307).

Each input is fed to every neuron in the network. Each neuron has as many weights as there are variables in the input vector, in other words, neuron $n$ has weights:
$$
\textbf{w}_n=[w_{n1}, w_{n2},...,w_{nk}].
$$

Self Organizing Maps learn based on a competition protocol. Each time an input is fed to the network, the neurons in the network compete for it using a discrimination function. A natural candidate for a discriminant function is the Euclidean distance between the neuron and the input:
$$
d_{in} = \sqrt{(x_{i1}-w_{n1})^2+(x_{i1}-w_{n1})^2+...+(x_{i1}-w_{n1})^2}
$$

The neuron that is most similar to the input (i.e., the one with the smallest distance to the input) becomes the winner, and it adjusts its parameters (weights) to further increase its similartiy to the input.

Here I want to illustrate the mechanics for creating SOMs.

# A simple one dimensional map and initialization

Lets begin by clearing the workspace.
```{r}
rm(list = ls())
```

Next, we load the packages that are used in this document:
```{r}
library(tidyverse) # Easily Install and Load the 'Tidyverse'
library(plotly) # Create Interactive Web Graphics via 'plotly.js'
```

And then we proceed to create a sample data set:
```{r}
no <- 75
y <- 4 * runif(no, min = 0, max = 0.2)
z <- 1 - y + runif(no, min = 0, max = 0.1)
df <- data.frame(y = y, z = z, map = NA)
```

This dataset has seventy-five observations of $k=2$ stochastic variables `y` and `z`.

The first step in working with a Self Organizing Map is to create a feature map.

As mentioned above, the map can be one-, two-, or higher-dimensional, although in practice two-dimensional maps are the most common. Although there is little practical advantage to creating a one-dimensional map, in this case it makes visualization of the key mechanics of a Self Organizing Map easier, so we will create a simple one-dimensional feature map.

A useful way to think about a Self Organizing Map is as a maleable string (in one dimension) or sheet (in two dimensions). The initial map does not have an even surface, but is instead "bumpy". The bumps are created by a random initialization of the weights.

A quantitative guideline to select the number of cells N in a map as a function of the number of observations $n$:
$$
N=5\sqrt{n}
$$

This means that larger samples require bigger maps, as seen below:
```{r}
f <- ggplot(data.frame(x = c(0, 1000)), aes(x))
size <- function(x) {5 * sqrt(x)}
f + stat_function(fun = size)
```

A large number of neurons enhances the ability of the map to adapt to the input; however, it also increases the probability of "blank" neurons, that is, neurons that do not classify any of inputs. A small number of blank neurons is to be expected, but a large number of blank neurons is indicative of inefficient size and may make classification problematic.

Here, we create and initialize a map with $N = 35$ neurons.

Since each input consists of $k=2$ variables the neurons are initialized with the same number of weights, so that attached to each neuron there are values `w1` and `w2`. These values are initialized to a small random number:
```{r}
set.seed(19)
SOM <- data.frame(x = c(1:35), w1 = runif(35, min = 0.1, max = 0.9), w2 = runif(35, min = 0.1, max = 0.9), t = factor(1))
```

We can now visualize the initial network. To plot the neurons, we utilize their position in the network (their index $n$) for the x-axis, and a summary of their weights, essentially their distance from an arbitrary point (say zero):
```{r}
ggplot(data = SOM, aes(x = x, y = sqrt(w1^2 + w2^2))) +
  geom_line() +
  geom_point(aes(color = sqrt(w1^2 + w2^2)), shape = 16, size = 10) +
  geom_point(color = "black", shape = 1, size = 10)  +
  scale_color_distiller(palette = "Spectral")
```

As seen above, the map is a string of neurons with initial random values for weights.

# Feeding inputs and competition

The inputs are fed sequentially to the network. This can be done by drawing an observation at random and without replacement from the dataset, or if the data are not sorted in any obvious way, by feeding observation $i=1$, followed by $i=2$, and so on until all observations available for training have been used.

For this example, we will feed the observations in order from 1 to 75.

To feed the first input, we select the first observation in the dataset and use the weights to calculate the discriminant function:
```{r}
count <- 0
count <- count + 1
D <- sqrt((SOM$w1 - df$y[count])^2 + (SOM$w2 - df$z[count])^2)
```

Notice that `D` is simply the Euclidean distance between the observation and the weights of the neuron. The winning neuron is the one closest to the input, presently:
```{r}
paste("The node closest to this input observation is", which.min(D), "with a distance D of", format(min(D), digits = 3))
df$map[count] <- which.min(D)
```

# Updating the weights and topological dependency of neurons

Once that the winning neuron has been identified, the weights can be updated.

A possible updating mechanis is as folows.

Calculating a new set of weights (say, the weights at $t+1$) by adjusting the current weights (say, the weights at $t$) with the difference between the current weights and the input:
$$
\textbf{w}_{n}(t+1)=\textbf{w}_{n}(t)+(\textbf{x}_i(t)-\textbf{w}_{n}(t))
$$

By reducing their distance, the neuron and the input become _more similar_.

The above mechanism applies to the winning neuron. Unlike neurons in a feedforward network that do not interact cotemporaneously (a neuron receives information from neurons in a preceding layer, and sends information to neurons in a subsequent layer), neurons in a Self Organizing Map display a property of _topological dependency_.

What this means is that neurons in the map can influence each other. This is illustrated in Figure 4.

![Figure 4. Updating and topological dependency](figures/Figure 4.jpg)

When the winning neuron is updated, other neurons in the map are updated too, but perhaps at a rate that decreases with increasing distance from the winning neuron. Thinking again of the map as a maleable surface, a neuron that adapts to more closely resemble its input will "pull" with it its neighbors, just like what would happen if you pressed or pulled a point on a rubber sheet.

The updating mechanism needs to be modified to capture this effect:
$$
\textbf{w}_{n}(t+1)=\textbf{w}_{n}(t)+h(\rho)(\textbf{x}_i(t)-\textbf{w}_{n}(t)).
$$

The updating mechanism is applied to every neuron on the map at time t, but the magnitude of the adjustment $(\textbf{x}_i(t)-\textbf{w}_{n}(t))$ is a function $h$ of the distance $\rho$ between the winning neuron and the neuron being updated.

Function $h()$ needs to be a monotonically decreasing function of distance $\rho$ to ensure a dampening effect on the adjustments to the weights. A candidate function is the negative exponential, for example:
$$
h(\rho)=exp(-\frac{\rho}{2})
$$

To implement the updating procedure in the example we need to calculate the distance of all neurons to the winning neuron (i.e., $\rho$):
```{r}
rho <- sqrt((which.min(D) - seq(1, 35))^2)
rho
```

The vector of distances that we calculated above defines the neighborhood of the winning neuron:
```{r}
ggplot(data = data.frame(x = c(1:35), rho = rho), aes(x = x, y = exp(-rho^2/2))) +
  geom_line() +
  ylab("h(rho)")
```

We can see from the plot above that the winning neuron will be fully adjusted by $(\textbf{x}_i(t)-\textbf{w}_{n}(t))$, whereas neurons at a distance of $\rho = 1$ unit away will be adjusted by only `r format(exp(-1/2), digits = 3)` of the difference between the weights and the inputs, and so on.

We now proceed to update the weights:
```{r}
w1 <- subset(SOM, t == as.character(count))$w1 + exp(-rho^2/2)*(y[count] - subset(SOM, t == as.character(count))$w1)
w2 <- subset(SOM, t == as.character(count))$w2 + exp(-rho^2/2)*(z[count] - subset(SOM, t == as.character(count))$w2)
SOM <- rbind(SOM, data.frame(x = c(1:35), w1 = w1, w2 = w2, t = factor(count + 1)))
```

If we plot again the map, we can see how it has updated:
```{r}
ggplot(data = subset(SOM, t == as.character(count + 1)), aes(x = x, y = sqrt(w1^2 + w2^2))) +
  geom_line() +
  geom_point(aes(color = sqrt(w1^2 + w2^2)), shape = 16, size = 10) +
  geom_point(color = "black", shape = 1, size = 10)  +
  scale_color_distiller(palette = "Spectral")
```

# Competition and updating

The training process continues by feeding the _next_ observation in the dataset to the network:
```{r}
count <- count + 1
D <- sqrt((subset(SOM, t == as.character(count))$w1 - df$y[count])^2 + (subset(SOM, t == as.character(count))$w2 - df$z[count])^2)
paste("The node most closely resembling this observation is", which.min(D), "with a distance D of", format(min(D), digits = 3))
df$map[count] <- which.min(D)
```

We again calculate the distance of all neurons to the winning neuron:
```{r}
rho <- sqrt((which.min(D) - seq(1, 35))^2)
rho
```

And update the weights:
```{r}
w1 <- subset(SOM, t == as.character(count))$w1 + exp(-rho^2/2)*(y[count] - subset(SOM, t == as.character(count))$w1)
w2 <- subset(SOM, t == as.character(count))$w2 + exp(-rho^2/2)*(z[count] - subset(SOM, t == as.character(count))$w2)
SOM <- rbind(SOM, data.frame(x = c(1:35), w1 = w1, w2 = w2, t = factor(count + 1)))
ggplot(data = subset(SOM, t == as.character(count + 1)), aes(x = x, y = sqrt(w1^2 + w2^2))) +
  geom_line() +
  geom_point(aes(color = sqrt(w1^2 + w2^2)), shape = 16, size = 10) +
  geom_point(color = "black", shape = 1, size = 10)  +
  scale_color_distiller(palette = "Spectral")
```

# Completing the training process

It is straightforward to loop the rest of the training procedure:
```{r}
for(count in 3:nrow(df)){
  # Pick the _next_ observation in the dataset and compare to the weights
  D <- sqrt((subset(SOM, t == as.character(count))$w1 - df$y[count])^2 + (subset(SOM, t == as.character(count))$w2 - df$z[count])^2)

  # Assign observation to node on map  
  df$map[count] <- which.min(D)

  # Calculate distance of all neurons to the winning neuron (call this $\rho$)
  rho <- sqrt((which.min(D) - seq(1, 35))^2)

  # Update the weights
  w1 <- subset(SOM, t == as.character(count))$w1 + exp(-rho^2/2)*(y[count] - subset(SOM, t == as.character(count))$w1)
  w2 <- subset(SOM, t == as.character(count))$w2 + exp(-rho^2/2)*(z[count] - subset(SOM, t == as.character(count))$w2)
  SOM <- rbind(SOM, data.frame(x = c(1:35), w1 = w1, w2 = w2, t = factor(count + 1)))
}
```

We can inspect the process by creating a `plotly` animation (more about `plotly [here](https://plot.ly/r/getting-started/)):
```{r}
p <- ggplot(data = SOM, aes(x = x, y = sqrt(w1^2 + w2^2))) +
  geom_line(aes(frame = t)) +
  geom_point(aes(color = sqrt(w1^2 + w2^2), frame = t), shape = 16, size = 5) +
  scale_color_distiller(palette = "Spectral")
p <- ggplotly(p)
p
```

The animation shows how the map has "self organized" in response to the inputs. If we compare initial map to final map:
```{r}
ggplot() +
  geom_line(data = subset(SOM, t == as.character(no)), aes(x = x, y = sqrt(w1^2 + w2^2))) +
  geom_point(data = subset(SOM, t == as.character(no)), aes(x = x, y = sqrt(w1^2 + w2^2), color = sqrt(w1^2 + w2^2)), shape = 16, size = 10) +
  geom_point(data = subset(SOM, t == as.character(no)), aes(x = x, y = sqrt(w1^2 + w2^2)), color = "black", shape = 1, size = 10) +
  geom_line(data = subset(SOM, t == as.character(1)), aes(x = x, y = sqrt(w1^2 + w2^2)), alpha = 0.3) +
  geom_point(data = subset(SOM, t == as.character(1)), aes(x = x, y = sqrt(w1^2 + w2^2)), color = "black", shape = 1, size = 10, alpha = 0.3) +
  scale_color_distiller(palette = "Spectral")
```

So the map has organized, and the observations have been classified, in the sense that they have been assigned to a neuron as part of the training process. Clearly, more similar observations will have been assigned to neighboring neurons (since they will themselves be more similar among them),  and possibly the same neuron.

Since in this example the inputs consisted only of two variables, it is possible to visualize them directly. The following plot of the inputs gives a sense of how well this procedure classified the observations:
```{r}
ggplot(data = df, aes(x = y, y = z)) + 
  geom_point(aes(color = sqrt(y^2 + z^2)), size = 5) +
  geom_point(size = 5, shape = 1) + 
  scale_color_distiller(palette = "Spectral")
  theme(legend.position = "bottom")
```

# Time-dependent parameters

The updating mechanism introduced above allows the map to remain equally malleable throughout the whole training exercise. A modified version of the updating mechanism introduces time-dependent parameters that reduce the malleability of the map as the training exercise advances.

This version of the updating mechanism is as follows:
$$
\textbf{w}_{n}(t+1)=\textbf{w}_{n}(t)+\alpha(t)h(\rho t)(\textbf{x}_i(t)-\textbf{w}_{n}(t))
$$

In the equation above, the topological control $h()$ becomes a function of, in addition of distance to the winning neurone, the training period $t$. For example:
$$
h(\rho t)=exp\Big(-\frac{\rho}{2\sigma(t)}\Big),
$$
where $\sigma(t)$ is also a monotonically decreasing function, such as:
$$
\sigma(t)=exp\Big(-\frac{t}{n}\Big)
$$
so the function grows smaller as t increases. The effect of this parameter is to make the size of the neighborbood smaller over time:
```{r}
f <- ggplot(data.frame(x = c(-10, 10)), aes(x))
h <- function(x) {exp(-x^2/50)}
f + stat_function(fun = h)
```

The second modification to the updating mechanism is the introduction of a _learning rate_ function $\alpha\$. This function is also time-dependent, and could be as follows:
$$
\alpha(t)=\alpha_0\Big(1-\frac{t}{n}\Big)
$$
As $t$ increases, the learning rate declines, in effect making later adjustments to the weights smaller.

Let try again training a map but now with time-dependent dampening parameters.

We'll first reset the data:
```{r}
set.seed(19)
SOM2 <- data.frame(x = c(1:35), w1 = runif(35, min = 0.1, max = 0.9), w2 = runif(35, min = 0.1, max = 0.9), t = factor(1))
```

We also recreate sample dataset:
```{r}
no <- 75
y <- 4 * runif(no, min = 0, max = 0.2)
z <- 1 - y + runif(no, min = 0, max = 0.1)
df2 <- data.frame(y = y, z = z, map = NA)
```

Here we train a new map with time-dependent parameters:
```{r}
for(count in 1:no){
  # Pick the _next_ observation in the dataset and compare to the weights
  D <- sqrt((subset(SOM2, t == as.character(count))$w1 - df2$y[count])^2 + (subset(SOM2, t == as.character(count))$w2 - df2$z[count])^2)

  # Assign observation to node on map  
  df2$map[count] <- which.min(D)

  # Calculate distance of all neurons to the winning neuron (call this $\rho$)
  rho <- sqrt((which.min(D) - seq(1, 35))^2)

  # Update the weights
  w1 <- subset(SOM2, t == as.character(count))$w1 + 0.9 * (1 - count/no) * exp(-rho^2/(2 * exp(-count/no)))*(df2$y[count] - subset(SOM2, t == as.character(count))$w1)
  w2 <- subset(SOM2, t == as.character(count))$w2 + 0.9 * (1 - count/no) * exp(-rho^2/(2 * exp(-count/no)))*(df2$z[count] - subset(SOM2, t == as.character(count))$w2)
  SOM2 <- rbind(SOM2, data.frame(x = c(1:35), w1 = w1, w2 = w2, t = factor(count + 1)))
}
```

Lets inspect the training procedure:
```{r}
p <- ggplot(data = SOM2, aes(x = x, y = sqrt(w1^2 + w2^2))) +
  geom_line(aes(frame = t)) +
  geom_point(aes(color = sqrt(w1^2 + w2^2), frame = t), shape = 16, size = 5) +
  scale_color_distiller(palette = "Spectral")
p <- ggplotly(p)
p
```

Compare the initial maps to the final maps.
```{r}
SOMs <- rbind(SOM, SOM2)
SOMs$som <- as.factor(rep(c("SOM1", "SOM2"), each = 2660))
ggplot(SOMs) +
  geom_line(data = subset(SOMs, t == as.character(no)), aes(x = x, y = sqrt(w1^2 + w2^2))) +
  geom_point(data = subset(SOMs, t == as.character(no)), aes(x = x, y = sqrt(w1^2 + w2^2), color = sqrt(w1^2 + w2^2)), shape = 16, size = 10) +
  geom_point(data = subset(SOMs, t == as.character(no)), aes(x = x, y = sqrt(w1^2 + w2^2)), color = "black", shape = 1, size = 10) +
  geom_line(data = subset(SOMs, t == as.character(1)), aes(x = x, y = sqrt(w1^2 + w2^2)), alpha = 0.3) +
  geom_point(data = subset(SOMs, t == as.character(1)), aes(x = x, y = sqrt(w1^2 + w2^2)), color = "black", shape = 1, size = 10, alpha = 0.3) +
  scale_color_distiller(palette = "Spectral") +
  facet_grid(.~ som) +
  theme(legend.position = "bottom")
```

What difference do you notice? What is a possible advantage of introducing time-dampening parameters in the training algorithm?

With this we have covered the basic mechanics of Self Organizing Maps. In the following sessions we will expand this to two dimensional maps and we will see practical applications.

### 2024-05-24 data/data-raw/SOM-share/Session 11 - Self Organizaing Maps IV.Rmd

---
title: "Artificial Neural Networks - Session 11: Self Organizing Maps IV: Analysis of Contingency Tables"
---

# Setup

Clear the workspace:
```{r clear-workspace}
rm(list = ls())
```

Load the libraries used in this notebook:
```{r load-libraries, include = FALSE}
library(readxl) # Read Excel Files
library(tidyverse) # Easily Install and Load the 'Tidyverse'
library(rgdal) # Bindings for the 'Geospatial' Data Abstraction Library
library(spdep) # Spatial Dependence: Weighting Schemes, Statistics
library(rgeos) # Interface to Geometry Engine - Open Source ('GEOS')
library(dplyr) # A Grammar of Data Manipulation
library(tmap) # Thematic Maps
library(SOMbrero) # SOM Bound to Realize Euclidean and Relational Outputs
library(rpart) # Recursive Partitioning and Regression Trees
library(rpart.plot) # Plot 'rpart' Models: An Enhanced Version of 'plot.rpart'
library(kableExtra) # Construct Complex Table with 'kable' and Pipe Syntax
```

Load geography and data files for this session:
```{r load-ridings}
ridings <- st_read("data/Ridings.shp")
```

# Introduction

Self Organizing Maps are often used to produce clusters when numeric data are available. Numeric data is often presented in the form of a table, with each record corresponding to a unit of analysis, and each column is an attribute. An example of this is a table showing, for instance, socio-economic and demographic attributes for zones.

Consider a selection of attributes for political jurisdictions in Ontario, which are available from the following file:
```{r read-census-data}
data <- read.csv("data/Ridings-Census Subdivisions-Overlay-Data.csv")
ridings$ED_NAME_EN <- data$ED_NAME_EN
```

The table below is an example of numeric information in a table 
```{r numeric-table-example}

kable(data |>
        select(ED_NAME_EN, Population, `Private.Dwellings`, `Avg.Median.Age`), 
      "html", 
      caption = "Example of table with numeric data",
      col.names = c("Jurisdiction", "Population", "Private Dwellings", "Avg Median Age"), 
      digits = 0) %>%
  kable_styling()
```

Each jurisdiction is described by a set of attributes, which could be, and often are, used to produce clustering/segmentation systems, as seen in the preceding session.

Other times, data is available that does not necessarily conform to the format of a table as shown above. Sometimes, for example, we have tables that describe relationships between variables, typically in the form of frequencies of occurrence. Such tables are then cross-tabulations or contingency tables.

An example of a table of this type is, for example, the results of Ontario's provincial election of 2018, which are available from the following file:
```{r load-election-data}
results.raw <- read.csv("data/Unofficial Results.csv")
results.raw <- rename(results.raw, ED_NAME_EN = Riding)
results.raw$ED_NAME_EN <- ridings$ED_NAME_EN
```

The table below is an example of a contingency table: 
```{r contingency-table}
kable(results.raw[,c(1,14:18)], "html", caption = "Contingency table: votes for each party by riding",
      col.names = c("Riding", "Votes Liberal", "Votes PC", "Votes NDP", "Votes Green", "Votes Other"), digits = 0) %>%
  kable_styling()
```

What is a key difference between a contingency table, and the other commonly used data tables?

Notice how in a contingency table, both the rows AND the columns can be considered units of analysis. In the table above, the rows are for "Jurisdictions", whereas the columns are for "Parties". The contents of the table are "Votes".

This is not the same for all data tables. The columns of the table with socio-economic and demographic information above do not necessarily represent an underlying variable, and in fact more commonly are used to collect multivariate information.

Whereas the sum of all values in a contingency table adds up to some universe (for instance total number of votes cast in the election), the sum of all values in a data table is not necessarily a meaningful quantity, since the variables are likely to be in different units! (e.g., number of people, number of dwellings, years).

Contingency tables then collect information about bivariate relationships.

When a contingency table is small, relationships are relatively easy to spot at a glance. Consider the following toy example of apartment occupancy by proximity to transit:
```{r toy-example}
df <- data.frame(Occupation = c(rep("Occupied", 240), rep("Vacant", 10), rep("Occupied", 180), rep("Vacant", 20)),
                     Distance = c(rep("Within Walking Distance", 240), rep("Within Walking Distance", 10), rep("Not Within Walking Distance", 180), rep("Not Within Walking Distance", 20)))
kable(addmargins(xtabs(~ Occupation + Distance, df)), "html") %>% kable_styling()
```

A relevant question is whether occupancy is related to proximity to transit: is "Occupied" more closely related to "Not Within Walking Distance" than to "Within Walking Distance"? In this case, the answer appears to be yes: apartments are more frequently occupied when they are _within_ walking distance of transit (occupancy rate within walking distance is 240/250 = 0.96, whereas beyond walking distance is 180/200 = 0.90).

There are statistics that can be implemented to test whether any variations in the frequency of events are significantly different from what could be expected purely by chance. For instance, a $\chi^2$ test of independence is as follows:
```{r chi2-toy-example}
chisq.test(xtabs(~ Occupation + Distance, df))
```

The test tells us that occupancy rates are significantly higher/lower based on proximity to transit, than what would be expected by chance.

Making sense of a larger contingency table is considerably more challenging based on visual inspection. Lets return to the results of the election in Ontario.

First, we will create a contingency table:
```{r create-contingency-table}
#Copy to create contingency table
results.ct <- results.raw
#Create row names for table
results.ct <- results.ct %>% rename(Riding = ED_NAME_EN) %>% select(Votes_Liberal, Votes_PC, Votes_NDP, Votes_Green, Votes_Other)
results.ct <- round(results.ct)
row.names(results.ct) <- results.raw$ED_NAME_EN
```

A test of independence indicates that the votes are not allocated at random in each riding (if that were the case, the results of the election would have been quite different!):
```{r chi2-votes}
chisq.test(results.ct)
```

However, this test does not help to answer the question of which ridings are more closely associated with each party (check the contingency table again).

The results of the election already give a natural way to cluster ridings: we can group ridings based on the party that won. See the contingency table, now coded by winner of the riding (with Liberals in red, PC in blue, NDP in orange, and Greens in green):
```{r contingency-table-with-winners}
results.raw %>%
  mutate(Votes_Liberal = cell_spec(round(Votes_Liberal), 
                              color = "white",
                              align = "c",
                              background = factor(Winner, c("Liberal", "PC", "NDP", "Green", "Other"), 
                                        c("red", "blue", "orange", "green", "white"))),
         Votes_PC = cell_spec(round(Votes_PC), 
                              color = "white",
                              align = "c",
                              background = factor(Winner, c("Liberal", "PC", "NDP", "Green", "Other"), 
                                        c("red", "blue", "orange", "green", "white"))),
         Votes_NDP = cell_spec(round(Votes_NDP), 
                              color = "white",
                              align = "c",
                              background = factor(Winner, c("Liberal", "PC", "NDP", "Green", "Other"), 
                                        c("red", "blue", "orange", "green", "white"))),
         Votes_Green = cell_spec(round(Votes_Green), 
                              color = "white",
                              align = "c",
                              background = factor(Winner, c("Liberal", "PC", "NDP", "Green", "Other"), 
                                        c("red", "blue", "orange", "green", "white"))),
         Votes_Other = cell_spec(round(Votes_Other), 
                              color = "white",
                              align = "c",
                              background = factor(Winner, c("Liberal", "PC", "NDP", "Green", "Other"), 
                                        c("red", "blue", "orange", "green", "white"))))%>%
  select(ED_NAME_EN, Votes_Liberal, Votes_PC, Votes_NDP, Votes_Green, Votes_Other) %>%
  kable(escape = F) %>%
  kable_styling()
```

The summary of ridings won by each party is:
```{r}
table(results.raw$Winner)
```

Clustering the ridings based on the winners of the election, however, potentially misses some important nuances. For example, Brampton West was won by the PC, but it is much closer to the NDP than Bruce-Grey-Owen Sound, also won by the PC.

A mechanism to simultaneously cluster the ridings _and_ the parties could help to understand some of these nuances. A canonical approach to do this is Correspondence Analysis, a statistical method closely related to Principal Component Analysis. Self Organizing Maps can also be used to this end [see paper by Cottrell et al. 1993](Analysing a contingency table with Kohonen maps.pdf). This session illustrates this application of SOMs as implemented in the `SOMbrero` package.

# Train SOM

The `SOMbrero` package can be used to train a SOM on a contingency table, by setting the argument `type` to "korresp".

A measure of the quality of a SOM is its topographic error, which measures the quality of the projection to the data. The topographic error varies between zero (perfect projection) to one (poor projection).

In this example, we can try different map configurations (i.e., the dimensions of the map) to try to improve the quality of the projection.

The first map uses the default 8 by 8 parameters:
```{r train-som-8by8}
set.seed(2343)
#Try different dimensions to reduce topographic error
korresp.som <- trainSOM(x.data = results.ct, dimension = c(8, 8),
                        type = "korresp", scaling = "chi2", nb.save = 10,
                        radius.type = "letremy")
summary(korresp.som)
```

Try several dimensions:
```{r train-som-mbyn}
set.seed(2343)
#Try different dimensions to reduce topographic error
korresp.som <- trainSOM(x.data = results.ct, dimension = c(4, 7),
                        type = "korresp", scaling = "chi2", nb.save = 10,
                        radius.type = "letremy")
summary(korresp.som)
```

After some experimentation, a map of size 4-by-5 seems to give the best projection quality (notice as well that it also has the lowest final energy):
```{r train-som-final}
set.seed(2343)
#Try different dimensions to reduce topographic error
korresp.som <- trainSOM(x.data = results.ct, dimension = c(4, 5),
                        type = "korresp", scaling = "chi2", nb.save = 10,
                        radius.type = "letremy")
summary(korresp.som)
```

Since we saved intermediate steps when training we can check the progress of the algorithm:
```{r energy-plot}
plot(korresp.som, what="energy")
```

The results of the clustering are as follows, with cell 1 clustering four ridings, cell 2 nine ridings, and so on. There are twenty cells in total (4 times 5) and only cell 8 is empty:
```{r}
table(korresp.som$clustering)
```

Note that there are 124 ridings and 5 types of vote totals: Liberal, PC, NDP, Green, Other. Therefore, a total of 129 units are clustered. As shown next, 
```{r clusters-table}
kable(korresp.som$clustering, "html", caption = "Results of SOM clustering algorithm (unit and cluster number)") %>% 
  kable_styling()
```

Notice how Liberal was clustered at cell 17, PC at cell 20, NDP at 1, Green at 4, and Other at 3. In this way, the ridings that are closest to Liberal are Eglinton and Orleans, also in cell 17:
```{r cluster-17}
kable(korresp.som$clustering[korresp.som$clustering == 17], "html", caption = "Results of SOM clustering algorithm (unit and cluster number)") %>% 
  kable_styling()
```

A hitmap can help to visualize the distribution of units to cells:

```{r hitmap}
plot(korresp.som, what="obs", type="hitmap", print.title=TRUE)
```

The hitmap can be complemented with line plots of the columns (recall that each column corresponds to a party):
```{r line-plot}
plot(korresp.som, what="prototypes", type="lines", view="c", print.title=TRUE)
```

The peak in Cluster 1 corresponds to the third column, i.e., NDP, whereas the peak in Cluster 20 corresponds to the second column, i.e., PC. The algorithm has clustered the main parties in opposite corners of the map, with Other right next to Green in cell 3.

Notice that ridings clustered in cells 2, 5, and 6 tend to be similar to those that align closely with NDP, whereas ridings in cells 12, 15, 16, and 19 are more similar to those closely aligned with PC. Cluster 17, associated with Liberal, on the other hand has an immediate neighbor with several ridings that did not prefer strongly any party (cluster 14). Cluster 14 is the largest cluster, and is comprised of 16 ridings.

# Obtain superclusters

A way to refine the insights derived from the previous analysis is to collect clusters into superclusters. The choice of the number of superclusters is up to the analyst, and depends on the tradeoff between conciseness and loss of information. For the example, I chose $k=6$, so that the superclusters can capture at least the five major parties, and possibly a catchall class:
```{r obtain-superclusters}
korresp.sc <- superClass(korresp.som, k=6)
summary(korresp.sc)
```

The frequency table above shows that supercluster 1 is comprised of four SOM cells, supercluster 2 of three and so on, as also seen in the dendogram below:
```{r dendogram}
plot(korresp.sc, plot.var=FALSE)
```

Lets repeat the lines plot with the superclusters:
```{r}
plot(korresp.sc, type="hitmap", print.title=TRUE, view="c")
```

The superclusters can be characterized as follows: SC1: NDP, SC2: Green, SC3: Leans NDP, SC4: Leans PC, SC5: PC, SC6: Liberal. Change the labels:
```{r relabel-clusters}
korresp.sc$cluster <- factor(korresp.sc$cluster,labels=c("NDP","Green","Leans NDP", "Leans PC", "PC", "Liberal"))
```

To explore these results we first append the results of the cluster and supercluster analysis to dataframe:
```{r append-clusters-results}
clusters.df <- data.frame(ED_NAME_EN = names(korresp.som$clustering))
clusters.df$cluster.som <- korresp.som$clustering
clusters.df$scluster.som <- korresp.sc$cluster[korresp.som$clustering]
```

Append to dataframe with results from election:
```{r}
results <- left_join(results.raw, clusters.df, by = "ED_NAME_EN")
```

Here we "tidy" the table to facilitate facetting during visualization:
```{r tidy-results-table}
results.t <- gather(results, key = "Party", "Votes", 14:18)
```

And relevel the variable `Party`:
```{r}
results.t$Party <- factor(results.t$Party)
results.t$Party <- factor(results.t$Party, levels(results.t$Party)[c(5, 2, 4, 3, 1)])
```

Define color palettes for visualization:
```{r color-palettes}
pal1 <- c("orange", "green", "gold", "dodgerblue", "blue", "red")
pal2 <- c("red", "green", "orange", "blue")
pal3 <- c( "blue", "red", "white", "orange", "green")
```

The following boxplot shows the distribution of votes per riding for each of the superclusters:
```{r}
ggplot(data = results.t, aes(x = Party, y = Votes, fill = Party)) + 
  geom_boxplot() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  scale_fill_manual(values = pal3) +
  facet_wrap(~ scluster.som)
```

With the exception of Liberal, it can be seen that superclusters affiliated with the major parties are distinctly in favor of the respective parties. The ridings in the "Leans PC" supercluster, on the other hand, have on average a vote advantage for PC, compared to the ridings in "Leans NDP", where the NDP does not have a clear advantage.

Compare the results of the clustering procedure to the results of the election:
```{r}
kable(table(results$Winner, results$scluster.som), "html", caption = "Ridings Won by Party and SOM Supercluster") %>%
  kable_styling() %>%
  add_header_above(c(" " = 1, "SOM Superclusters" = 6))
```

Three PC-leaning ridings were won by the NDP, six NDP-leaning ridings were won by PC and two more by the Liberals.

Lets now check the geography of the election. First, append the results to the `SpatialPolygonsDataFrame` for mapping:
```{r join-election-to-geography}
ridings <- ridings |>
  left_join(results, 
            by = "ED_NAME_EN")
```

Convert the variable `Winner` to a factor:
```{r convert-outcome-to-factor}
ridings$Winner <- factor(ridings$Winner)
ridings$Winner <- factor(ridings$Winner, levels(ridings$Winner)[c(1, 3, 2, 4)])
```

```{r}
tmap_mode("view")
map1 <- tm_shape(ridings) + tm_polygons("scluster.som", palette = pal1)
map2 <- tm_shape(ridings) + tm_polygons("Winner", palette = pal2)
tmap_arrange(map1, map2)
```

```{r}
results %>%
  mutate(Votes_Liberal = cell_spec(round(Votes_Liberal), 
                              color = "white",
                              align = "c",
                              background = factor(scluster.som, c("NDP", "Green", "Leans NDP", "Leans PC", "PC", "Liberal"), 
                                        c("orange", "green", "gold", "dodgerblue", "blue", "red"))),
         Votes_PC = cell_spec(round(Votes_PC), 
                              color = "white",
                              align = "c",
                              background = factor(scluster.som, c("NDP", "Green", "Leans NDP", "Leans PC", "PC", "Liberal"), 
                                        c("orange", "green", "gold", "dodgerblue", "blue", "red"))),
         Votes_NDP = cell_spec(round(Votes_NDP), 
                              color = "white",
                              align = "c",
                              background = factor(scluster.som, c("NDP", "Green", "Leans NDP", "Leans PC", "PC", "Liberal"), 
                                        c("orange", "green", "gold", "dodgerblue", "blue", "red"))),
         Votes_Green = cell_spec(round(Votes_Green), 
                              color = "white",
                              align = "c",
                              background = factor(scluster.som, c("NDP", "Green", "Leans NDP", "Leans PC", "PC", "Liberal"), 
                                        c("orange", "green", "gold", "dodgerblue", "blue", "red"))),
         Votes_Other = cell_spec(round(Votes_Other), 
                              color = "white",
                              align = "c",
                              background = factor(scluster.som, c("NDP", "Green", "Leans NDP", "Leans PC", "PC", "Liberal"), 
                                        c("orange", "green", "gold", "dodgerblue", "blue", "red")))) %>%
  select(ED_NAME_EN, Votes_Liberal, Votes_PC, Votes_NDP, Votes_Green, Votes_Other) %>%
  kable(escape = F) %>%
  kable_styling()
```

#Profile ridings

Collect the data to profile:
```{r select-census-variables}
data <- transmute(data, ED_NAME_EN, 
                  Population, `Population.Density` = `Avg.Pop.Density`,
                  `Median.Age` = `Avg.Median.Age`,
                  `Median.Income` = `Avg.Median.income`, `Male.Median.Income` = `Avg.Male_Median.income`, `Female.Median.Income` = `Avg.Female_Median.income`,
                  `Pct.Government.Transfer.Payments` = `Avg.Government.transfer.payments..`,
                  `Median Commuting Duration` = `Avg.Median.commuting.duration`,
                  `Avg Income Taxes as Pct Income` = `Avg.Income.taxes.as...of.total.income`,
                  `Pct Population in Low Income` = `Avg.Population.in.low.income..`,
                  `Population.25.64`,
                  `X15._No.certificate.diploma.degree`,
                  `X15._HS.diploma.equivalent`,
                  `X15._Post.HS.diploma`,
                  `X15._Apprentice.trades.certificate`,
                  `X15._College.certificate.diploma`,
                  `X15._University..bachelor.s`,
                  `X15._University.bachelor.s.`,
                  `X15._University_Bachelor.s.degree`,
                  `X15._University_Above.bachelor.s`
                  )
```

Transform Census Variables:
```{r transform-census-variables}
data.model <- mutate(data,
                     `Prop no certificate/diploma/degree` = `X15._No.certificate.diploma.degree`/`Population.25.64`,
                     `Prop HS diploma/equivalent` = `X15._HS.diploma.equivalent`/`Population.25.64`,
                     `Prop Post HS diploma` = `X15._Post.HS.diploma`/`Population.25.64`,
                     `Prop apprentice/trades certificate` = `X15._Apprentice.trades.certificate`/`Population.25.64`,
                     `Prop college certificate/diploma` = `X15._College.certificate.diploma`/`Population.25.64`,
                     `Prop university bachelor` = `X15._University..bachelor.s`/`Population.25.64`,
                     `Prop postgraduate` = `X15._University_Bachelor.s.degree`/`Population.25.64`)
```

Select variables for modelling:
```{r}
data.model <- select(data.model,-starts_with("X15"), -`Population.25.64`)
```

Join results of superclusters:
```{r}
data.model$`Super Cluster` <- results$scluster.som
#data.model$Winner <- results$Winner
data.model <- select(data.model, -c(Population, ED_NAME_EN))
```

Join to census geography:
```{r join-census-to-geography}
#Join results dataframe to spatial polygons:
#ridings@data <- left_join(ridings@data, data.model, by = "ED_NAME_EN")
```

Train a decision tree using the data:
```{r}
classmod.som <- rpart(`Super Cluster` ~., data = data.model)
#summary(classmod.som)
```

Plot results of the decision tree:
```{r fig.width= 12}
rpart.plot(classmod.som, cex = 0.75, box.palette = 0, type = 3)
```

### 2024-05-24 data/data-raw/SOM-share/Session 9 - Self Organizing Maps II.Rmd

---
title: "Artificial Neural Networks - Session 9: Self Organizing Maps II"
---

# Introduction

In the preceding session self-organizing maps were introduced. As seen there, self-organizing maps are a form of neural network that operates on some distinctive principles, importantly:

1) Learning happens in a competitive way
2) The network has topological relationships

In this way, a lattice of neurons compete for training samples, and in the process learn from those samples. The outcome of the learning process is that training samples are grouped (or classified), in a self-organized map.

The example in the preceding seminar consisted of a simple 1D map. In this seminar, two-dimensional maps will be introduced, along with the resources to train SOMs in R.

There are several packages that implement SOMs in R, including `kohonen`, `som`, and `SOMbrero`. Each package offers slightly different functions, some of which we will explore. For instance, `kohonen` implements unsupervised SOMs, supervised SOMs with two parallel maps, SOMs with multiple parallel maps. The package `SOMbrero` (which stands for _SOM Bound to Realize Euclidean and Relational Outputs_) can be used to train SOMs using numerical data, contingency tables, and disimilarity matrices.

```{r}
rm(list = ls())
```

Lets begin by loading the packages used for this:
```{r message = FALSE}
library(tidyverse) # Easily Install and Load the 'Tidyverse'
library(SOMbrero) # SOM Bound to Realize Euclidean and Relational Outputs
library(plotly) # Create Interactive Web Graphics via 'plotly.js'
library(spdep) # Spatial Dependence: Weighting Schemes, Statistics
```

The dataset for the example is a collection of pokemon characters, compiled from several sources, as described [here](https://www.kaggle.com/shwetp/pokemon-visualizations-unsupervised-learning/data)
```{r message = FALSE}
pokemon <- read_csv("data/pokemon_gen.csv")
```

The dataframe contains the following pieces of information for n = `r nrow(pokemon)` pokemon:
```{r}
summary(pokemon)
```

Note: Pokemon are sometimes cute.

![Figure 1. Pikachu](figures/Pikachu.png)

Pokemon are described by their so-called _stats_, including hit points (which determine how much damage they can take), Attack (how much damage they cause), Defense (influences how much damage they receive when attacked), Special Attack and Special Defense (similar to Attack and Defense but when using special abilities), and Speed (which determines their order in battle: higher speed monsters attack first).

These are all numerical attributes.

In addition, there is the legendary status of the monster. Legendary monsters are rare and typically very powerful. The generation, on the other hand, is a sequence for release of the games, and successive generations modify previous monsters.

Note that generation should be a factor, so lets change it:
```{r}
pokemon$Generation <- as.factor(pokemon$Generation)
```

To illustrate SOMs in two dimensions, we will use only two of the _stats_ in the dataframe, `Defense` and `Speed`:
```{r}
df <- select(pokemon, Defense, Speed)
```

Since these are two numerical attributes, we can visualize them using a scattetplot:
```{r}
ggplot(data = df, aes(x = Defense, y = Speed)) + 
  geom_point() +
  coord_fixed()
```

Given these two attributes, a SOM can be trained using the function `trainSOM` from `SOMbrero`. Only one argument is needed, the input data `x.data`. Other arguments can be assigned by default. In what follows the dimension of the map is given as 5-by-5, and I request that 100 intermediate steps be saved. 
```{r}
set.seed(593)
# run the SOM algorithm 
nrows <- 5
ncols <- 5
poke.som <- trainSOM(x.data=df, dimension=c(nrows, ncols), nb.save = 100, radius.type = "letremy", verbose = TRUE)
```

The value of the function (output) is an object of class somRes:
```{r}
class(poke.som)
```

The summary reports some attributes of the map:
```{r}
summary(poke.som)
```

In this implementation of SOM, the energy is a measure of deformation of the map, when conceptualized as an elastic membrane. The _topographic error_, on the other hand, is the proportion of all data vectors for which first and second best matching units are not adjacent. This measures topology preservation.

In addition, the summary of the SOM includes the results of ANOVA, using the the results of the clustering as the factor. This helps to assess the relevance of each variable to discriminate between clusters. In the present example, the results of ANOVA indicate that the means of the variables for the clusters are significantly different.

The `somRes` object includes the following items:

- clustering: the results of the clustering algorithm, which gives each observation a number that corresponds to their best matching unit (BMU).

- prototypes: the final weights of the cells.

- energy: the final energy.

- backup: a list that includes the prototypes (the weights) at intermediate steps, the steps at which backups were recorded, intermediate clustering results, etc.

Since backups were requested, it is possible to plot the evolution of the training algorithm. For instance, the following plot shows the energy:
```{r}
plot(poke.som, what = "energy")
```

It is also possible to visualize the evolution of the map. This necessitates that we define the topology of the lattice:
```{r}
poke.grid <- cell2nb(nrow = nrows, ncol = ncols, type = "rook")
```

The above gives the list of neighbors for neurons in the lattice.

The prototypes from the backups are extracted for each of the backups and flattened for plotting. The topology obtained above can be used to obtain the edges, by giving the start and endpoints by neighbors:
```{r}
som.grid <- data.frame(x1 = numeric(length(poke.som$backup$prototypes) * length(unlist(poke.grid))))
count <- 0
for(p in 1:length(poke.som$backup$prototypes)){
  for(i in 1:(nrows * ncols)){
    for(j in 1:length(poke.grid[[i]])){
      count <- count + 1
      som.grid$x1[count] <- poke.som$backup$prototypes[[p]][i,1]
      som.grid$y1[count] <- poke.som$backup$prototypes[[p]][i,2]
      som.grid$x2[count] <- poke.som$backup$prototypes[[p]][poke.grid[[i]][j],1]
      som.grid$y2[count] <- poke.som$backup$prototypes[[p]][poke.grid[[i]][j],2]
      som.grid$backup[count] <- p
    }
  }  
}
```

And these are the points:
```{r}
som.points <- do.call(rbind.data.frame, poke.som$backup$prototypes)
rownames(som.points) <- c()
som.points$backup <- rep(1:100, each = 25)
```

Now the evolution of the map can be plotted as follows:
```{r warning = FALSE, message = FALSE}
p <- ggplot(data = som.points, 
            aes(x = Defense, y = Speed, frame = backup)) +
  geom_segment(data = som.grid, 
               aes(x = x1, y = y1, xend = x2, yend = y2, frame = backup)) +
  geom_point(aes(frame = backup), size = 2) +
  coord_equal()

p <- ggplotly(p)
p
```

The final map is:
```{r}
som.grid.final <- data.frame(x1 = numeric(length(unlist(poke.grid))))
count <- 0
for(i in 1:(nrows * ncols)){
    for(j in 1:length(poke.grid[[i]])){
      count <- count + 1
      som.grid.final$x1[count] <- poke.som$prototypes[i,1]
      som.grid.final$y1[count] <- poke.som$prototypes[i,2]
      som.grid.final$x2[count] <- poke.som$prototypes[poke.grid[[i]][j],1]
      som.grid.final$y2[count] <- poke.som$prototypes[poke.grid[[i]][j],2]
    }
  }  
```

```{r}
ggplot(data = data.frame(node = factor(1:25), poke.som$prototypes), 
       aes(x = Defense, y = Speed)) +
  geom_segment(data = som.grid.final, 
               aes(x = x1, y = y1, xend = x2, yend = y2)) +
  geom_point(aes(color = node), size = 4) +
  coord_equal()
```

```{r}
ggplot(data = data.frame(node = factor(1:25), poke.som$prototypes), 
       aes(x = Defense, y = Speed)) +
    geom_point(data = data.frame(node = factor(poke.som$clustering), df),
       aes(x = Defense, y = Speed, color = node)) +
  geom_segment(data = som.grid.final, 
               aes(x = x1, y = y1, xend = x2, yend = y2),
               size = 1) +
  geom_point(aes(color = node), size = 4) +
  geom_point(shape = 1, size = 4) +
  coord_equal()
```

Is this a "good" result?

Lets change the parameters:
```{r}
set.seed(593)
# run the SOM algorithm 
nrows <- 5
ncols <- 7
poke.som2 <- trainSOM(x.data=df, dimension=c(nrows, ncols), radius.type = "letremy", verbose = TRUE)
```

```{r}
summary(poke.som2)
```

Now the final map is:

```{r}
poke.grid2 <- cell2nb(nrow = nrows, ncol = ncols, type = "rook")

som.grid.final2 <- data.frame(x1 = numeric(length(unlist(poke.grid2))))
count <- 0
for(i in 1:(nrows * ncols)){
    for(j in 1:length(poke.grid2[[i]])){
      count <- count + 1
      som.grid.final2$x1[count] <- poke.som2$prototypes[i,1]
      som.grid.final2$y1[count] <- poke.som2$prototypes[i,2]
      som.grid.final2$x2[count] <- poke.som2$prototypes[poke.grid2[[i]][j],1]
      som.grid.final2$y2[count] <- poke.som2$prototypes[poke.grid2[[i]][j],2]
    }
  }  
```

```{r}
ggplot(data = data.frame(node = factor(1:(nrows * ncols)),
                         poke.som2$prototypes), 
       aes(x = Defense, y = Speed)) +
    geom_point(data = data.frame(node = factor(poke.som2$clustering), df),
       aes(x = Defense, y = Speed, color = node)) +
  geom_segment(data = som.grid.final2, 
               aes(x = x1, y = y1, xend = x2, yend = y2),
               size = 1) +
  geom_point(aes(color = node), size = 4) +
  geom_point(shape = 1, size = 4) +
  coord_equal()
```

Lets now try this but with the full array of numeric stats:
```{r}
df <- select(pokemon, HP, Attack, Defense, SpAtk = `Sp. Atk`, SpDef = `Sp. Def`, Speed)
```

A rule of thumb for selecting the size of the SOM is $5\sqrt n$ neurons. Since the size of the sample is `r nrow(pokemon)`, this means a square with `r round(sqrt(5 * sqrt(nrow(pokemon))))`:
```{r}
set.seed(593)
# run the SOM algorithm 
nrows <- 12
ncols <- 12
poke.som3 <- trainSOM(x.data=df, dimension=c(nrows, ncols), radius.type = "letremy", verbose = TRUE)
```

This gives the summary for the map:
```{r}
summary(poke.som3)
```

The results of ANOVA indicate that all variables contribute to discrimination in the clustering.

The following shows the number of cases assigned to each node:
```{r}
table(poke.som3$clustering)
```

Which can be visualized by means of a _hitmap_:
```{r}
plot(poke.som3, what = "obs", type = "hitmap")
```

The hitmap reveals two empty cells, that is, neurons that did not classify a single observation.

Change the size of the map:
```{r}
set.seed(593)
# run the SOM algorithm 
nrows <- 8
ncols <- 8
poke.som3 <- trainSOM(x.data=df, dimension=c(nrows, ncols), radius.type = "letremy", verbose = TRUE)
```

```{r}
summary(poke.som3)
```

Plotting the _hitmap_:
```{r}
plot(poke.som3, what = "obs", type = "hitmap")
```

# Interpretation of clusters

Interpretation of clusters typically involves descriptive statistics of the clusters. `SOMbrero` implements a number of graphical tools to support interpretation.

The first example is a heatmap of the variables in clusters:
```{r}
par(mfrow=c(3,2))
plot(poke.som3, what="obs", type="color", variable=1, print.title=TRUE, 
     main="HP")
plot(poke.som3, what="obs", type="color", variable=2, print.title=TRUE, 
     main="Attack")
plot(poke.som3, what="obs", type="color", variable=3, print.title=TRUE, 
     main="Defense")
plot(poke.som3, what="obs", type="color", variable=4, print.title=TRUE, 
     main="Special Attack")
plot(poke.som3, what="obs", type="color", variable=3, print.title=TRUE, 
     main="Special Defense")
plot(poke.som3, what="obs", type="color", variable=4, print.title=TRUE, 
     main="Speed")
```

The lower left corner of the map concentrates monsters with high speed, high special attack, and high special attack. the upper left corner concentrates high defense and high special defense. High hit points are towards the center of the map.

Two additional way of visualizing the characteristics of the clusters are as follows.

This creates barplots:
```{r}
plot(poke.som, what="obs", type="barplot", print.title=TRUE)
```

And this creates mean lines:
```{r}
plot(poke.som3, what="obs", type="meanline", key.loc=c(-2,8), mar=c(0,2,2,0))
```

And of course, we would like to know how much dispersion there is within the clusters, something that can be explored by means of boxplots:
```{r}
plot(poke.som3, what="obs", type="boxplot")
```

Seeing how some of the stats indicate puny monsters, I am curious to see if this is due to their status as legendary monsters. A plot of an additional variable is as follows:
```{r}
plot(poke.som3, what="add", type="pie", variable=pokemon$Legendary)
```

```{r}
poke.sc <- superClass(poke.som3, k=5)
summary(poke.sc)
```

This is a dendogram with the super-clusters:
```{r}
plot(poke.sc, plot.var=FALSE)
```

And the map with the superclusters:

```{r}
plot(poke.sc, type="grid", plot.legend=TRUE)
```

### 2024-05-24 paper/Manuscript.qmd

@@
 ---
+
+<!--
+
+•	Establishing definitive research questions and the work’s contributions. Us deciding on this together will help me sculpt from what’s there
+  •	Thoughts on methods? Can we improve presentation?
+  •	Thoughts on results? Can we improve presentation? What are insights that are significant to you? What do you want to know more about?
+o	Thoughts on how I weighted care destinations
+o	Thoughts on aggregation by ‘care category’ vs ‘specific type’ of destination
+o	Thoughts on additional dimensions – i.e., the inclusion of results that only consider ‘grocery, errand, and health centric’ destinations (for people who have no dependents).
+For the time being, I left this consideration out and just demo _all_ destinations.
+Also, what about LICO-AT discussion?
+  •	Critical thoughts on how to tie walking and cycling modes together.
+For now, I did not include cycling results, though I have the cycling travel time matrices.
+We should use the cycling results perhaps to demo how the range of care access can be extended for certain neighbourhoods.
+A potential ‘what-if’ scenario?
+  •	Next steps
+
+-->
+
 ```{r}
@@
 
+RESEARCH OBJECTIVES:
+
+1)	15 minute cities stem from conventional methods that ignore mobility of care perspectives. There is an applied way of doing this, 
+2)	Use of the innovative destination dataset
+3)	Mid-size typical Canadian city, how frequently do we find these caring neighbourhoods (potential - accessibility). 
+4)	Who enjoys them and who does not? Clustering methodology. 
+a.	Types of profiles will emerge – i.e. with LICO > X + Prevalence of children >Y + walking accessibility score > U + diversity score > W = “caring neighbourhood”
+
 # Data and Methods
@@
 - $W^c_j$ is a number of weighted opportunities ($O_j^c$) of category $c$ at $j$.
-- $c_{ij}$ is the travel cost between $i$ and $j$.
+- $c_{ij}$ is the travel cost between $i$ and $j$. <!--AS NOTE: TAKE THIS AWAY, have it be non-weighted-->
 - $f(\cdot)$ is an impedance function of $c_{ij}$; within the cumulative opportunity approach, it is a binary function that takes the value of 1 if $c_{ij}$ is less than a selected value [@handyMeasuringAccessibilityExploration1997].
@@
 Next, the Q3 (median + standard deviation) of all $\sum_{c}S_i^c$ values for each DA are taken and represented as $S^{c}_{DA}$.
-$S^{c}_{DA}$ is the upper quartile number of weighted care destinations that can be potentially interacted with within a 15 minute walk by parcels within a DA: an optimistic perspective on how many opportunities can be potentially reached `r word_comment("Thoughts on this choice instead of the median?
-I do this because the parcels are not population weighted, so 1 parcel = 1 travel time.
-There can be obviously multiple households in 1 parcel, so using Q3 makes some sense.", ".")`
+$S^{c}_{DA}$ is the upper quartile number of weighted care destinations that can be potentially interacted with within a 15 minute walk by parcels within a DA: an optimistic perspective on how many opportunities can be potentially reached.
 
@@
 In our work, the diversity of the specific type of destination is of interest, we calculate $S^{st}_i$ for each parcel $i$ for each $st$ type following the equation for $S^{st}_i$.
-This upper quartile is aggregated at the $DA$ to come up with a $S^{st}_{DA}$ `r word_comment("Thoughts on this choice?
-I think we need to consider both the category type and the specific type in this paper.
-We justify the categories so that's fair but I still think there's value in presenting the specific type  some how - this is what I think this diversity index is good for so I've presented that as an option.
-Note though, it'd be much more simple to just focus on calculating the $D^{st}_{DA}$ using the $S_i^c$ and not introducing the $S_i^{st}$ calculation like I've done..
-so I've made the analysis more complicated.
-Also destination weighting is based on the four categories - so that is distorting the diversity in certain ways.",".")`
+This upper quartile is aggregated at the $DA$ to come up with a $S^{st}_{DA}$.
+<!--AS NOTE: remove 'st' and just have diveristy be based on 'c'.
+-->
 
@@
 # Results
+<!--Initial thoughts on results:
 
+Initial results seem optimistic. With SOM , method will classify neighbourhoods into these clusters we then tag/name.
+
+A part of this - spatial knowledge of people doesn't match the dataset. i.e. all stores are not equal, they have qualities that make them less/more accessibly.
+
+But we can speak in potentials. They facilities are there and they could potentially be improved. -->
 ```{r plot-figure3, out.width=600, out.height=600}

### 2024-05-24 data/02-preliminary-analysis.qmd
added 'access_15.rds', the parcel level cumulative opp count per parcel and category

@@
 ```{r re-weighting points so all destinations are considered spatially equal in access counts, include=FALSE}
-Dependent_Centric <- 0.25*2225/1265
-Errand_Centric <- 0.25*2225/158
-Grocery_Centric <- 0.25*2225/381
-Health_Centric <- 0.25*2225/421
-
-#adds up to 2225 opportunities  
-0.25*2225/1265*(1265) +  0.25*2225/158*(158) + 0.25*2225/381*(381) + 0.25*2225/421*(421)
-
-#adding the weighted care destination counts -- reweighted such that the 4 categories each make up 25% of the 2225 points.
-care_dest <- care_dest |>
-          mutate(Dest_count_weighted = case_when(Care_Category == "Dependent-centric" ~ Dependent_Centric,
-                                                 Care_Category == "Errand-centric" ~ Errand_Centric,
-                                                 Care_Category == "Grocery-centric" ~ Grocery_Centric,
-                                                 Care_Category == "Health-centric" ~ Health_Centric))
-
-care_dest$Dest_count_weighted |> sum()
+# Dependent_Centric <- 0.25*2225/1265
+# Errand_Centric <- 0.25*2225/158
+# Grocery_Centric <- 0.25*2225/381
+# Health_Centric <- 0.25*2225/421
+# 
+# #adds up to 2225 opportunities  
+# 0.25*2225/1265*(1265) +  0.25*2225/158*(158) + 0.25*2225/381*(381) + 0.25*2225/421*(421)
+# 
+# #adding the weighted care destination counts -- reweighted such that the 4 categories each make up 25% of the 2225 points.
+# care_dest <- care_dest |>
+#           mutate(Dest_count_weighted = case_when(Care_Category == "Dependent-centric" ~ Dependent_Centric,
+#                                                  Care_Category == "Errand-centric" ~ Errand_Centric,
+#                                                  Care_Category == "Grocery-centric" ~ Grocery_Centric,
+#                                                  Care_Category == "Health-centric" ~ Health_Centric))
+# 
+# care_dest$Dest_count_weighted |> sum()
 ```
@@
   left_join(
-    care_dest |> st_drop_geometry() |> select(c("care_ID", "Care_Category", "Care_Category_Specific", "Care_Category_Speciality", "Dest_count_weighted")),  by = c("to_id" = "care_ID")) |>
+    care_dest |> st_drop_geometry() |> select(c("care_ID", "Care_Category", "Care_Category_Specific", "Care_Category_Speciality")),  by = c("to_id" = "care_ID")) |>
   mutate(f = 1)
@@
   summarize(GeoUID = first(GeoUID),
-            copp_isc = n(),
-            copp_isc_w = sum(Dest_count_weighted))
+            copp_isc = n())
 
@@
                     pop = pop_dwell, #pop assumed at each res. parcel -- it is the pop in the DA divided by the number of dwellings in the DA. 
-                    opp = Dest_count_weighted, #instead of just being vector of 1
+                    opp = f, 
                     r = f, f = f, alpha = 1) |> 
@@
             copp_isc = sum(copp_isc),
-            copp_isc_w = sum(copp_isc_w),
             V_isc_w = sum(V_isc_w))
@@
 Aggregating time!
+```{r}
+#ID = care destination ID (in care_dest), GeoUID is id of the DA, from_id is parcel ID
+saveRDS(access_15, "access_15.rds")
+```
 
@@
 #join this to access_15_bycategory_DA
-
 access_15_bycategory_DA <- access_15_bycategory_DA |> left_join(step2, by="GeoUID")
 ```
-
+COMEBACKHERE
 ```{r}

### 2024-05-25 data/02-preliminary-analysis.qmd
Experiments profiling the neighborhoods based on their accessibility characteristics

@@
 ```
+
 ```{r}
@@
 #ID = care destination ID (in care_dest), GeoUID is id of the DA, from_id is parcel ID
-saveRDS(access_15, "access_15.rds")
+save(access_15, file = "access_15.rda", compress = "xz")
 ```

### 2024-05-25 data/03-clustering-and-profiling.qmd
Experiments profiling the neighborhoods based on their accessibility characteristics

---
title: "Clustering"
author: "Antonio Paez & Anastasia Soukhov"
format: html
---

## Quarto

Lets begin by loading the packages used for this:
```{r message = FALSE}
library(AppliedPredictiveModeling) # Functions and Data Sets for 'Applied Predictive Modeling'
library(caret) # Classification and Regression Training
library(ggparty)
library(ggpubr)
library(party)
library(plotly) # Create Interactive Web Graphics via 'plotly.js'
library(rpart) # Recursive Partitioning and Regression Trees
library(rpart.plot) # Plot 'rpart' Models: An Enhanced Version of 'plot.rpart'
library(skimr)
#library(spdep) # Spatial Dependence: Weighting Schemes, Statistics
library(SOMbrero) # SOM Bound to Realize Euclidean and Relational Outputs
library(tidyverse) # Easily Install and Load the 'Tidyverse'

# library(formattable) # Create 'Formattable' Data Structures
# library(kableExtra) # Construct Complex Table with 'kable' and Pipe Syntax
# library(knitr) # A General-Purpose Package for Dynamic Report Generation in R
# library(plotly) # Create Interactive Web Graphics via 'plotly.js'
# library(sf) # Simple Features for R
# library(spdep) # Spatial Dependence: Weighting Schemes, Statistics
# library(SOMbrero) # SOM Bound to Realize Euclidean and Relational Outputs
# library(tidyverse) # Easily Install and Load the 'Tidyverse'
# library(tmap) # Thematic Maps
```

Read od table:
```{r}
load("access_15.rda")
```

Aggregate by Care Category:
```{r}
access_15_aggregated <- access_15 |>
  group_by(from_id,
           Care_Category) |>
  summarize(copp_isc = sum(copp_isc),
            .groups = "drop")
```

Calculate total Care Category opportunities by parcel:
```{r}
access_15_aggregated_total <- access_15_aggregated |>
  group_by(from_id) |>
  summarize(tot_copp_isc = sum(copp_isc),
            .groups = "drop")
```

Notice that some parcels have access to zero opportunities:
```{r}
summary(access_15_aggregated_total)
```

How many parcels have zero accessibility (i.e., total number of opportunities is zero):
```{r}
access_15_aggregated_total |>
  filter(tot_copp_isc == 0) |>
  nrow()
```

Join total opportunities to table:
```{r}
access_15_aggregated <- access_15_aggregated |>
  left_join(access_15_aggregated_total,
            by = "from_id")
```

Calculate proportion of opportunities by class (use a zero for the parcels that have accessibility zero):
```{r}
access_15_aggregated <- access_15_aggregated |>
  mutate(prop_copp_isc = ifelse(tot_copp_isc > 0, copp_isc/tot_copp_isc, 0))  
```

Check that the proportions are correct (all should add up to one):
```{r}
access_15_aggregated |>
  group_by(from_id) |>
  summarize(prop = sum(prop_copp_isc)) |>
  filter(round(prop) != 1)
```

Normalize number of opportunities to min-max range:
```{r}
access_15_aggregated <- access_15_aggregated |>
  mutate(copp_isc_norm = ifelse(tot_copp_isc != 0, (copp_isc - min(tot_copp_isc))/(max(tot_copp_isc) - min(tot_copp_isc)), 0))
```

Calculate diversity index:
```{r}
access_15_agg_div <- access_15_aggregated |>
  group_by(from_id) |>
  summarize(div = -sum(prop_copp_isc * log(prop_copp_isc), na.rm = TRUE)/log(4))
```

Pivot od table wider so that each Care Category is a column then retrieve the proportion of opportunities by Care Category:
```{r}
access_15_prop_wide <- access_15_aggregated |>
  select(from_id, Care_Category, prop_copp_isc) |>
  pivot_wider(names_from = "Care_Category", 
              values_from = "prop_copp_isc") |>
  rename_with(-from_id, 
              .fn = ~paste0("prop_",
                            str_replace(.x, "-","_")))
```

Pivot od table wider so that each Care Category is a column and retrieve the normalized number of opportunities:
```{r}
access_15_norm_wide <- access_15_aggregated |>
  select(from_id, Care_Category, copp_isc_norm) |>
  pivot_wider(names_from = "Care_Category", 
              values_from = "copp_isc_norm") |>
  rename_with(-from_id, 
              .fn = ~paste0("norm_", str_replace(.x, "-","_")))
```

Pivot od table wider so that each Care Category is a column and retrieve the total number of opportunities by Category:
```{r}
access_15_total_wide <- access_15_aggregated |>
  select(from_id, Care_Category, copp_isc) |>
  pivot_wider(names_from = "Care_Category", 
              values_from = "copp_isc") |>
  rename_with(-from_id, 
              .fn = ~paste0("total_", str_replace(.x, "-","_")))
```

Summary of these tables:
```{r}
summary(access_15_total_wide)
summary(access_15_norm_wide)
summary(access_15_prop_wide)
summary(access_15_agg_div)
```
Join the tables:
```{r}
access_15_wide <- access_15_total_wide |>
  left_join(access_15_norm_wide,
            by = "from_id") |>
  left_join(access_15_prop_wide,
            by = "from_id") |>
  left_join(access_15_agg_div,
            by = "from_id")
```

Summary of this table:
```{r}
summary(access_15_wide)
```

Now this table can be used for the clustering analysis.

Given these normalized attributes, a SOM can be trained using the function `trainSOM` from `SOMbrero`. Only one argument is needed, the input data `x.data`. Other arguments can be assigned by default. In what follows the dimension of the map is given as 5-by-5, and I request that 100 intermediate steps be saved. 
```{r}
set.seed(593)
# run the SOM algorithm 
nrows <- 5
ncols <- 5
access.som <- trainSOM(x.data = access_15_wide |>
                       # Remove the non-normalized accessibility scores
                       select(-c(from_id, starts_with("total_"))),
                     dimension=c(nrows, ncols),
                     nb.save = 100,
                     radius.type = "letremy",
                     verbose = TRUE)
```

The value of the function (output) is an object of class somRes:
```{r}
class(access.som)
```

The summary reports some attributes of the map:
```{r}
summary(access.som)
```

In this implementation of SOM, the energy is a measure of deformation of the map, when conceptualized as an elastic membrane. The _topographic error_, on the other hand, is the proportion of all data vectors for which first and second best matching units are not adjacent. This measures topology preservation.

In addition, the summary of the SOM includes the results of ANOVA, using the the results of the clustering as the factor. This helps to assess the relevance of each variable to discriminate between clusters.

The `somRes` object includes the following items:

- clustering: the results of the clustering algorithm, which gives each observation a number that corresponds to their best matching unit (BMU).

- prototypes: the final weights of the cells.

- energy: the final energy.

- backup: a list that includes the prototypes (the weights) at intermediate steps, the steps at which backups were recorded, intermediate clustering results, etc.

Since backups were requested, it is possible to plot the evolution of the training algorithm. For instance, the following plot shows the energy:
```{r}
plot(access.som, 
     what = "energy")
```

The following shows the number of cases assigned to each node:
```{r}
table(access.som$clustering)
```
Which can be visualized by means of a _hitmap_:
```{r}
plot(access.som,
     what = "obs",
     type = "hitmap")
```

# Interpretation of clusters

Interpretation of clusters typically involves descriptive statistics of the clusters. `SOMbrero` implements a number of graphical tools to support interpretation.

The first example is a heatmap of the variables in clusters:
```{r}
par(mfrow=c(3,2))
plot(access.som, 
     what = "obs", 
     type = "color", 
     variable = 1, 
     show.names=TRUE, 
     main="norm_Dependent_centric")
plot(access.som, 
     what = "obs", 
     type = "color", 
     variable = 2, 
     show.names = TRUE, 
     main = "norm_Errand_centric")
plot(access.som, 
     what = "obs", 
     type = "color", 
     variable = 3,
     show.names = TRUE, 
     main="norm_Grocery_centric")
plot(access.som, 
     what="obs", 
     type="color", 
     variable=4, 
     show.names = TRUE, 
     main="norm_Health_centric")
plot(access.som, 
     what="obs", 
     type="color",
     variable = 9, 
     show.names = TRUE, 
     main="Diversity")
```

The lower left corner of the map concentrates monsters with high speed, high special attack, and high special attack. the upper left corner concentrates high defense and high special defense. High hit points are towards the center of the map.

Two additional way of visualizing the characteristics of the clusters are as follows.

This creates barplots:
```{r}
plot(access.som, 
     what = "obs", 
     type="barplot", 
     show.names=TRUE)
```

And this creates mean lines:
```{r}
plot(access.som, 
     what="obs", 
     type="meanline", 
     key.loc=c(-2,8), 
     mar=c(0,2,2,0))
```

And of course, we would like to know how much dispersion there is within the clusters, something that can be explored by means of boxplots:
```{r}
plot(access.som, 
     what="obs", 
     type="boxplot")
```

Super clusters:
```{r}
access.sc <- superClass(access.som,
                        k = 6)
summary(access.sc)
```

This is a dendogram with the super-clusters:
```{r}
plot(access.sc, 
     plot.var = TRUE)
```

And the map with the superclusters:
```{r}
plot(access.sc, 
     type = "grid", 
     plot.legend = TRUE)
```

## Profile the superclusters

Once that clusters have been obtained, it is interesting to profile the results. By profiling the results, we can derive insights as to the (hopefully) distinctive characteristics of the clusters. For instance, is Cluster 1 mostly composed of young people? People with internet access?

Explore the results. First append the results of the cluster and super cluster analysis to dataframe:
```{r}
access_15_wide$cluster <- access.som$clustering
access_15_wide$scluster <- access.sc$cluster[access.som$clustering]
```

Select four of the features with the greatest discriminatory power, and do a feature plot (from package `caret`):
```{r fig.width=16}
# transparentTheme(trans = .4)
# featurePlot(x = data.frame(access_15_wide$div, 
#                            access_15_wide$Dependent_centric_norm, 
#                            access_15_wide$Errand_centric_norm,
#                            access_15_wide$Grocery_centric_norm,
#                            access_15_wide$Health_centric_norm), 
#             y = as.factor(access_15_wide$scluster), 
#             plot = "pairs",
#             ## Add a key at the top
#             auto.key = list(columns = 8))
```

Descriptive statistics of superclusters.

```{r}
access_15_wide |>
  select(scluster, starts_with("total_"), div) |>
  group_by(scluster) |>
  skim()
```

## Profiling the super clusters

One possibility is to analyze the results of the clustering algorithm by means of a classification technique. This is illustrated here by means of a decision tree. A decision tree is a supervised learning technique, which in this case can use the results of the clustering algorithm as labels, and the input variables as features. If the clustering algorithm performed well, class membership must be relatively homogeneous. Since a decision tree

Select data for classification:
```{r}
df <- access_15_wide |>
  select(-c(cluster,
            from_id, 
            starts_with("prop"), 
            starts_with("norm"))) |>
  rename(Dependent = total_Dependent_centric,
         Grocery = total_Grocery_centric,
         Errand = total_Errand_centric,
         Health = total_Health_centric) |>
  mutate(scluster = factor(scluster),
         div = round(div, digits = 3))
```

Train a decision tree using the data:
```{r}
classmod.som <- rpart(scluster ~ ., 
                      data = df)
```

Visualize the decision tree:
```{r}
rpart.plot(classmod.som, cex = 0.5, type = 0, varlen = 8)
```

This does not look too pretty. Try ggparty, but first convert the rpart object to a party object:
```{r}
classmod.sompar <- as.party(classmod.som)
```

Autoplot as ggplot:
```{r}
autoplot(classmod.sompar)
```

Frequency of parcels in the super clusters (look at the data of the first node):
```{r}
freq_sc <- ggplot(classmod.sompar[1]$data) +
  geom_bar(aes(x = "", 
               fill = scluster)) +
  scale_fill_discrete(name = "super cluster") +
  theme_void()

freq_sc
```

Extract the legend:
```{r}
leg <- ggpubr::get_legend(freq_sc,
                          position = "bottom")

leg <- as_ggplot(leg)

leg
```

Make a prettier plot:
```{r}
party_plot <- ggparty(classmod.sompar) +
  geom_edge() +
  geom_edge_label(size = 1.5) +
  geom_node_splitvar(size = 2) +
  # pass list to gglist containing all ggplot components we want to plot for each
  # (default: terminal) node
  geom_node_plot(gglist = list(geom_bar(aes(x = "", fill = scluster),
                                        position = position_fill()),
                               xlab(""),
                               ylab(""),
                               theme_void(),
                               theme(axis.text = element_blank())))
```

Plot the tree:
```{r}
party_plot
```

### 2024-05-25 data/03-clustering-and-profiling.qmd
Completed initial experiments developing accessibility profiles and demographic profiles of parcels.

@@
 library(caret) # Classification and Regression Training
-library(ggparty)
-library(ggpubr)
-library(party)
+library(cowplot) # Streamlined Plot Theme and Plot Annotations for 'ggplot2'
+library(ggparty) # 'ggplot' Visualizations for the 'partykit' Package
+library(party) # A Laboratory for Recursive Partytioning
 library(plotly) # Create Interactive Web Graphics via 'plotly.js'
@@
 library(rpart.plot) # Plot 'rpart' Models: An Enhanced Version of 'plot.rpart'
-library(skimr)
-#library(spdep) # Spatial Dependence: Weighting Schemes, Statistics
+library(skimr) # Compact and Flexible Summaries of Data
+library(sf) # Simple Features for R
 library(SOMbrero) # SOM Bound to Realize Euclidean and Relational Outputs
 library(tidyverse) # Easily Install and Load the 'Tidyverse'
-
-# library(formattable) # Create 'Formattable' Data Structures
-# library(kableExtra) # Construct Complex Table with 'kable' and Pipe Syntax
-# library(knitr) # A General-Purpose Package for Dynamic Report Generation in R
-# library(plotly) # Create Interactive Web Graphics via 'plotly.js'
-# library(sf) # Simple Features for R
-# library(spdep) # Spatial Dependence: Weighting Schemes, Statistics
-# library(SOMbrero) # SOM Bound to Realize Euclidean and Relational Outputs
-# library(tidyverse) # Easily Install and Load the 'Tidyverse'
-# library(tmap) # Thematic Maps
 ```
@@
 ```{r}
-set.seed(593)
+set.seed(5935)
 # run the SOM algorithm 
@@
 
-Frequency of parcels in the super clusters (look at the data of the first node):
+Frequency of parcels in the super clusters:
 ```{r}
-freq_sc <- ggplot(classmod.sompar[1]$data) +
+freq_sc <- ggplot(access_15_wide |> 
+         mutate(scluster = factor(scluster))) +
   geom_bar(aes(x = "", 
+               group = scluster, 
                fill = scluster)) +
-  scale_fill_discrete(name = "super cluster") +
-  theme_void()
+  scale_fill_discrete(name = "super cluster") #+
+  #theme_void() +
+  #guides(fill = guide_legend(nrow = 1)) +
+  #theme(legend.position = "bottom")
 
@@
 ```{r}
-leg <- ggpubr::get_legend(freq_sc,
-                          position = "bottom")
+legd <- cowplot::get_legend(freq_sc)
 
-leg <- as_ggplot(leg)
-
-leg
+legd
 ```
@@
   geom_edge() +
-  geom_edge_label(size = 1.5) +
-  geom_node_splitvar(size = 2) +
+  geom_edge_label(size = 1.8) +
+  geom_node_label(aes(label = splitvar),
+                  id = "inner", 
+                  size = 2, nudge_y = -0.022) +
   # pass list to gglist containing all ggplot components we want to plot for each
@@
 ```
+
+```{r}
+library(cowplot) # Streamlined Plot Theme and Plot Annotations for 'ggplot2' # Streamlined Plot Theme and Plot Annotations for 'ggplot2'
+
+plot_grid(party_plot, legd, nrow = 1, rel_widths = c(1, 0.2))
+
+```
+
+Cluster 6 is high in suckitude: low in diversity, with 4.6 Dependent Centric destinations on average, and typically no other destinations: zero Errand centric destinations, and only about 3.5% of parcels have access to Grocery-centric destinations.
+```{r}
+access_15_wide |>
+  select(scluster, starts_with("total_"), div) |>
+  rename_with(.fn = ~str_replace(.x, "total_", "")) |>
+  filter(scluster == 6) |>
+  skim()
+```
+
+## Demographic profile of clusters
+
+```{r}
+load("HAM_census_21.rda")
+```
+
+```{r}
+access_15_wide <- access_15_wide |>
+  left_join(access_15 |>
+              ungroup() |>
+              distinct(from_id, 
+                       .keep_all = TRUE) |>
+              select(from_id, 
+                     GeoUID),
+            by = "from_id") |>
+  left_join(HAM_census_21 |>
+              st_drop_geometry(),
+            by = "GeoUID")
+```
+
+Select data for classification:
+```{r}
+df <- access_15_wide |>
+  transmute(scluster = factor(scluster),
+            LICO = `v_CA21_1085: Prevalence of low income based on the Low-income cut-offs, after tax (LICO-AT) (%)`)
+```
+
+Train a decision tree using the data:
+```{r}
+classdemo.som <- rpart(scluster ~ ., 
+                      data = df)
+```
+
+Visualize the decision tree:
+```{r}
+rpart.plot(classdemo.som, cex = 0.5, type = 0, varlen = 8)
+```
+
+convert the rpart object to a party object:
+```{r}
+classdemo.sompar <- as.party(classdemo.som)
+```
+
+Make a prettier plot:
+```{r}
+party_plot <- ggparty(classdemo.sompar) +
+  geom_edge() +
+  geom_edge_label(size = 1.8) +
+  geom_node_label(aes(label = splitvar),
+                  id = "inner", 
+                  size = 2, nudge_y = -0.022) +
+  # pass list to gglist containing all ggplot components we want to plot for each
+  # (default: terminal) node
+  geom_node_plot(gglist = list(geom_bar(aes(x = "", fill = scluster),
+                                        position = position_fill()),
+                               xlab(""),
+                               ylab(""),
+                               theme_void(),
+                               theme(axis.text = element_blank())))
+```
+
+Plot the tree:
+```{r}
+party_plot
+```
+
+```{r}
+library(cowplot) # Streamlined Plot Theme and Plot Annotations for 'ggplot2' # Streamlined Plot Theme and Plot Annotations for 'ggplot2'
+
+plot_grid(party_plot, legd, nrow = 1, rel_widths = c(1, 0.2))
+
+```
+
+Notice that super cluster 3 () makes most of the node with the highest prevalence of poverty (LICO > 10.9%). What is the accessibility profile of this cluster?
+
+It is high diversity, with multiple Grocery with multiple Grocery- and Dependant-centric opportunities, and at least some Errand-centric opportunities:
+```{r}
+access_15_wide |>
+  select(scluster, starts_with("total_"), div) |>
+  rename_with(.fn = ~str_replace(.x, "total_", "")) |>
+  filter(scluster == 3) |>
+  skim()
+```
+
+It is clear that LICO alone is not sufficient to sharply profile the clusters. Ask Anastasia to choose other variables.

### 2024-05-27 paper/Manuscript.qmd
removed 'weighted' results from manuscript

@@
 # care_dest|> st_drop_geometry() |> group_by(Care_Category) |> summarise(count = n())
-```
-
-```{r re-weighting points so all destinations are considered spatially equal in access counts, include=FALSE}
-Dependent_Centric <- 0.25*2225/1265
-Errand_Centric <- 0.25*2225/158
-Grocery_Centric <- 0.25*2225/381
-Health_Centric <- 0.25*2225/421
-
-#adds up to 2225 opportunities  
-0.25*2225/1265*(1265) +  0.25*2225/158*(158) + 0.25*2225/381*(381) + 0.25*2225/421*(421)
-
-#adding the weighted care destination counts -- reweighted such that the 4 categories each make up 25% of the 2225 points.
-care_dest <- care_dest |>
-          mutate(Dest_count_weighted = case_when(Care_Category == "Dependent-centric" ~ Dependent_Centric,
-                                                 Care_Category == "Errand-centric" ~ Errand_Centric,
-                                                 Care_Category == "Grocery-centric" ~ Grocery_Centric,
-                                                 Care_Category == "Health-centric" ~ Health_Centric))
-
-care_dest$Dest_count_weighted |> sum()
 ```
@@
 
-For the purpose of this analysis and in absence of city-wide household preferences for care destinations, all locations are re-weighted to make each of the four categories conceptually equivalent i.e., each location within each category $c$ is `r word_comment("Thoughts on this", "weighted to equal")` $W_j^c = \frac{1}{C} \frac{\sum_{c} O_j^c}{O_j^c}$ where $C$ is the total number of categories and $O_j^c$ is a location at destination $j$ within a category $c$.
-The total sum of locations is preserved after weighting such that $\sum_{c} \sum_{j} O_j^c = \sum_{c} \sum_{j} W_j^c * O_j^c$.
-Conceptually, this simplistic re-weight makes destinations in categories with fewer locations more important and destinations in categories with more locations less important.
-The underlying assumption is the status-quo of locations per category is sufficient, e.g., many daycares (smaller capacities) and few hospitals (larger capacities) sufficiently serves  the populations' care needs.
-Re-weighting in this way ensures that populations who are near _assumed_ smaller capacity destinations such as daycares are not assigned a disproportionately high accessibility score and vice versa for populations near destinations with larger service capacities like hospitals.
-In absence of empirical data for amenity weight calibration, this methodological assumption is a limitation.
-
-- Dependent-centric (`r sum(care_dest$Care_Category == "Dependent-centric")` destinations each at `r Dependent_Centric`), 
-- Errand-centric (`r sum(care_dest$Care_Category == "Errand-centric")` destinations each at `r Errand_Centric`),
-- Grocery-centric (`r sum(care_dest$Care_Category == "Grocery-centric")` destinations each at `r Grocery_Centric`) and,
-- Health- centric (`r sum(care_dest$Care_Category == "Health-centric")` destinations each at `r Health_Centric`) .
-
 ## Methods
@@
 - $c$ is a set of care categories.
-- $W^c_j$ is a number of weighted opportunities ($O_j^c$) of category $c$ at $j$.
-- $c_{ij}$ is the travel cost between $i$ and $j$. <!--AS NOTE: TAKE THIS AWAY, have it be non-weighted-->
+- $O^c_j$ is a number of opportunities of category $c$ at $j$.
+- $c_{ij}$ is the travel cost between $i$ and $j$.
 - $f(\cdot)$ is an impedance function of $c_{ij}$; within the cumulative opportunity approach, it is a binary function that takes the value of 1 if $c_{ij}$ is less than a selected value [@handyMeasuringAccessibilityExploration1997].
@@
 This upper quartile is aggregated at the $DA$ to come up with a $S^{st}_{DA}$.
-<!--AS NOTE: remove 'st' and just have diveristy be based on 'c'.
+<!--AS NOTE: potentially remove 'st' and just have diveristy be based on 'c'.
+For now, let's keep as is..
 -->
@@
 - $DA$ is a set of DA spatial units (to aggregate parcel point data).
-- $S_{i}^{st}$ is the unconstrained accessibility, the sum of weighted opportunities reachable within $f(\cdot)$, at each $i$ for each $st$.
+- $S_{i}^{st}$ is the unconstrained accessibility, the sum of opportunities reachable within $f(\cdot)$, at each $i$ for each $st$.
 - $S^{st}_{DA}$ is the aggregated unconstrained accessibility at each $DA$. In our work, it is the upper quartile of $S_i^{st}$ values in a $DA$. 
@@
 <!--Initial thoughts on results:
-
-Initial results seem optimistic. With SOM , method will classify neighbourhoods into these clusters we then tag/name.
-
-A part of this - spatial knowledge of people doesn't match the dataset. i.e. all stores are not equal, they have qualities that make them less/more accessibly.
-
-But we can speak in potentials. They facilities are there and they could potentially be improved. -->
+-Initial results seem optimistic. With SOM , method will classify neighbourhoods into these clusters we then tag/name. 
+- A part of this - spatial knowledge of people doesn't match the dataset. i.e. all stores are not equal, they have qualities that make them less/more accessibly. 
+- But we can speak in potentials. They facilities are there and they could potentially be improved. -->
 ```{r plot-figure3, out.width=600, out.height=600}
@@
 Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_copp_weighted_plot.tiff"))
+knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_copp_plot.tiff"))
 ```
@@
 Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_entropy_copp_weighted_plot.tiff"))
+knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_entropy_copp_plot.tiff"))
 ```

### 2024-05-27 data/02-preliminary-analysis.qmd
part 2 removed weighted calculations in 'preliminary-analysis'

@@
                     r = f, f = f, alpha = 1) |> 
-  group_by(from_id, Care_Category, Care_Category_Specific) |> summarize(V_isc_w = sum(V_ij, na.rm = TRUE))
+  group_by(from_id, Care_Category, Care_Category_Specific) |> summarize(V_isc = sum(V_ij, na.rm = TRUE))
 
@@
 rm(test_all)
-access_15$V_isc_w |> sum(na.rm = TRUE) #sums to 2206 opportunities -- out of 2225 total opportunities. some opportunities canot be reached wtihin 15 min walk! 
+access_15$V_isc |> sum(na.rm = TRUE) #sums to 2206 opportunities -- out of 2225 total opportunities. some opportunities canot be reached wtihin 15 min walk! 
 ```
@@
 
-           V_isc_w = rep(0, length(R_PARCELS_CENTS_2020$ID)*length(c("Park", "Daycare or EarlyON", "Convenience Store", "School",
+           V_isc = rep(0, length(R_PARCELS_CENTS_2020$ID)*length(c("Park", "Daycare or EarlyON", "Convenience Store", "School",
                                           "Long-Term Care or Retirement Home", "Senior Centre",
@@
             copp_isc = sum(copp_isc),
-            V_isc_w = sum(V_isc_w))
+            V_isc = sum(V_isc))
 rm(access_15_zeros)
 #check
-access_15$V_isc_w |> sum(na.rm=T)
+access_15$V_isc |> sum(na.rm=T)
 ```
@@
             copp_ic = sum(copp_isc, na.rm=T), #the number of opportunities that can be reached by category from each parcel
-            copp_ic_w = sum(copp_isc, na.rm=T), #the number of opportunities that can be reached by category from each parcel
-            V_ic_w = sum(V_isc_w, na.rm = T))
+            V_ic = sum(V_isc, na.rm = T))
 #check
-access_15_bycategory$V_ic_w |> sum(na.rm=T)
+access_15_bycategory$V_ic |> sum(na.rm=T)
 ```
@@
             copp_iscDA_w_sd = sd(copp_isc_w), #the SD
-            V_w_iscDA = sum(V_isc_w), #the number of total spatially available opportunities (all) that can be reached from each parcel in a DA
-            vv_w_iscDA = sum(V_isc_w)/first(Population), #the per-capita number of total spatially available opportunities (all) that can be reached from each parcel in a DA
+            V_iscDA = sum(V_isc), #the number of total spatially available opportunities (all) that can be reached from each parcel in a DA
+            vv_iscDA = sum(V_isc)/first(Population), #the per-capita number of total spatially available opportunities (all) that can be reached from each parcel in a DA
             Population = first(Population),
@@
             copp_ic_sd = sd(copp_isc), #the SD
-            copp_ic_w_med = median(copp_isc_w), #the median number of opportunities that can be reached by category from each parcel in a DA
-            copp_ic_w_sd = sd(copp_isc_w), #the SD
-            V_ic_w = sum(V_isc_w), #the number of total spatially available opportunities that can be reached by category from each parcel in a DA
-            vv_ic_w = sum(V_isc_w)/first(Population), #the per-capita number of total spatially available opportunities that can be reached by category from each parcel in a DA
+            V_ic = sum(V_isc), #the number of total spatially available opportunities that can be reached by category from each parcel in a DA
+            vV_ic = sum(V_isc)/first(Population), #the per-capita number of total spatially available opportunities that can be reached by category from each parcel in a DA
             Population = first(Population),
@@
   summarise(copp_isc_med = median(copp_isc), #the median number of opportunities that can be reached by specific category from each parcel in a DA
-            copp_isc_sd = sd(copp_isc), #the SD
-            copp_isc_w_med = median(copp_isc_w), #the median number of opportunities that can be reached by specific category from each parcel in a DA
-            copp_isc_w_sd = sd(copp_isc_w), #the SD
-            V_isc_w = sum(V_isc_w), #the number of total spatially available opportunities that can be reached by specific category from each parcel in a DA
-            vv_isc = sum(V_isc_w)/first(Population), #the per-capita number of total spatially available opportunities that can be reached by specific category from each parcel in a DA
+            copp_isc_sd = sd(copp_isc), 
+            V_isc = sum(V_isc), #the number of total spatially available opportunities that can be reached by specific category from each parcel in a DA
+            vv_isc = sum(V_isc)/first(Population), #the per-capita number of total spatially available opportunities that can be reached by specific category from each parcel in a DA
             Population = first(Population),
@@
 access_15_bycategory_DA |> summary()
-access_15_DA$V_w_iscDA |> sum(na.rm=T)
-access_15_bycategory_DA$V_ic_w |> sum(na.rm=T)
-access_15_byspecificcategory_DA$V_isc_w |> sum(na.rm=T)
+access_15_DA$V_iscDA |> sum(na.rm=T)
+access_15_bycategory_DA$V_ic |> sum(na.rm=T)
+access_15_byspecificcategory_DA$V_isc |> sum(na.rm=T)
 
@@
 ```{r}
-access_15_byspecificcategory_DA |> group_by(Care_Category_Specific) |> summarise(V_isc_w = sum(V_isc_w))
+access_15_byspecificcategory_DA |> group_by(Care_Category_Specific) |> summarise(V_isc = sum(V_isc))
 care_dest |> st_drop_geometry() |> group_by(Care_Category_Specific) |> 
-  summarise(Dest_count_weighted = sum(Dest_count_weighted),
-            count = n())
+  summarise(count = n())
 
@@
   summarise(OPPi = sum(copp_isc_med, na.rm=T) + sum(copp_isc_sd, na.rm=T),
-            OPPi_w = sum(copp_isc_w_med, na.rm=T) + sum(copp_isc_w_sd, na.rm=T),
-            Vi_w = sum(V_isc_w))
+            Vi = sum(V_isc))
 
@@
              top1_copp = ifelse(is.nan(top1_copp), 0, top1_copp),
-             top1_coppw = (copp_isc_w_med+copp_isc_w_sd) / OPPi_w * log((copp_isc_w_med+copp_isc_w_sd) / OPPi_w),
-             top1_coppw = ifelse(is.nan(top1_coppw), 0, top1_coppw),
-             top1_Vw= (V_isc_w) / Vi_w * log((V_isc_w) / Vi_w),
-             top1_Vw = ifelse(is.nan(top1_Vw), 0, top1_Vw),)  |> 
+             top1_V= (V_isc) / Vi * log((V_isc) / Vi),
+             top1_V = ifelse(is.nan(top1_V), 0, top1_V),)  |> 
   group_by(GeoUID) |>
   summarise(entropy_copp_isc_medsd = sum(top1_copp) / -log(14),
-            entropy_copp_isc_medsd_w = sum(top1_coppw) / -log(14),
-            entropy_V_isc_w = sum(top1_Vw) / -log(14))
+            entropy_V_isc = sum(top1_V) / -log(14))
 
@@
 ```
-COMEBACKHERE
+
 ```{r}
@@
 all_cats <- access_15_bycategory_DA |> 
-  mutate(copp_all = sum(copp_ic_med, na.rm=T) + sum(copp_ic_sd, na.rm=T),
-         copp_all_w = sum(copp_ic_w_med, na.rm=T) + sum(copp_ic_w_sd, na.rm=T)) |>
+  mutate(copp_all = sum(copp_ic_med, na.rm=T) + sum(copp_ic_sd, na.rm=T)) |>
   group_by(GeoUID) |>
          summarise(summed_copp_all = sum(copp_all),
-                   summed_copp_all_w = sum(copp_all_w),
-                   summed_Vic_w = sum(V_ic_w, na.rm=T),
+                   summed_Vic = sum(V_ic, na.rm=T),
                    Population = mean(Population),
@@
                                  summed_copp_all >= 25 ~ "[25+ des.) Very caring"),
-                   Care_Access_Status1_w = case_when(summed_copp_all_w < 4 ~ "[0-04 des.) Not caring",
-                                 summed_copp_all_w >= 4 & summed_copp_all_w < 25 ~ "[04-25 des.) Caring", 
-                                 summed_copp_all_w >= 25 ~ "[25+ des.) Very caring"),
-
                    Care_Access_Status2 = case_when(all(copp_all >= 2 & copp_all < 4) ~ "- complete (2 to 4 options per cat.)", 
                                                    all(copp_all >= 4) ~ "- very complete (4 or more options per cat.)"),
-                   Care_Access_Status2_w = case_when(all(copp_all_w >= 2 & copp_all_w < 4) ~ "- complete (2 to 4 options per cat.)", 
-                                                   all(copp_all_w >= 4) ~ "- very complete (4 or more options per cat.)"),
                    entropy_copp_isc_medsd = first(entropy_copp_isc_medsd),
-                   entropy_copp_isc_medsd_w = first(entropy_copp_isc_medsd_w),
-                   entropy_V_isc_w = first(entropy_V_isc_w)) |>
+                   entropy_V_isc = first(entropy_V_isc)) |>
   mutate(Care_Access_Status2 = ifelse(is.na(Care_Access_Status2), "", Care_Access_Status2),
-         Care_Access_Status2_w = ifelse(is.na(Care_Access_Status2_w), "", Care_Access_Status2_w),
          Care_Access_Status = paste(Care_Access_Status1,Care_Access_Status2, sep=" "),
-         Care_Access_Status_w = paste(Care_Access_Status1_w,Care_Access_Status2_w, sep=" "),
-         vvi = summed_Vic_w/Population)
+         vvi = summed_Vic/Population)
 
-all_cats$summed_Vic_w |> sum() #only 2206  opps can be reached out of the total 2225, so that's what the numbers are based on.
+all_cats$summed_Vic |> sum() #only 2206  opps can be reached out of the total 2225, so that's what the numbers are based on.
 care_dest|> nrow()
@@
 tmap_save(all_cats_copp_plot, file="all_cats_copp_plot.tiff")
-
-all_cats_copp_weighted_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
-  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
-  tm_shape(all_cats) +
-  tm_polygons("Care_Access_Status_w",
-              border.alpha = 0.2,
-              palette = c("darkred", "orangered", "orange", "yellow", "forestgreen", "lightgreen"),
-              title = "Number of care destinations (all - weighted) spatially accessible within 15 min. walk",
-              legend.is.portrait = TRUE) +
-  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
-  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
-  tm_compass(position = c("right", "top"), size=1.0)+
-  tm_layout(bg.color = "grey",
-            panel.label.bg.color = "white",
-            panel.label.color = "black",
-            legend.outside = TRUE,
-            legend.position = c("center", "bottom"),
-            legend.outside.position = "bottom",
-            legend.outside.size = 0.2,
-            legend.width = 12
-            # legend.title.size = 0.7,
-            # legend.text.size = 0.6
-            )
-tmap_save(all_cats_copp_weighted_plot, file="all_cats_copp_weighted_plot.tiff")
 ```
@@
 tmap_save(all_cats_entropy_copp_plot, file="all_cats_entropy_copp_plot.tiff")
-
-all_cats_entropy_copp_weighted_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
-  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
-  tm_shape(all_cats) +
-  tm_polygons("entropy_copp_isc_medsd_w",
-              border.alpha = 0.2,
-              palette = "Reds",
-              title = "Diversity of (all - weighted) care destinations spatially accessible within 15 min. walk",
-              legend.is.portrait = TRUE) +
-  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
-  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
-  tm_compass(position = c("right", "top"), size=1.0)+
-  tm_layout(bg.color = "grey",
-            panel.label.bg.color = "white",
-            panel.label.color = "black",
-            legend.outside = TRUE,
-            legend.position = c("center", "bottom"),
-            legend.outside.position = "bottom",
-            legend.outside.size = 0.2,
-            legend.width = 12)
-
-tmap_save(all_cats_entropy_copp_weighted_plot, file="all_cats_entropy_copp_weighted_plot.tiff")
 ```
@@
 ```{r}
-all_cats |> group_by(Care_Access_Status_w) |> summarise(min = min(summed_copp_all_w),
-                                                        median = median(summed_copp_all_w),
-                                                        max = max(summed_copp_all_w))
 all_cats |> group_by(Care_Access_Status) |> summarise(min = min(summed_copp_all),
@@
 all_cats$entropy_copp_isc_medsd|> summary()
-all_cats$entropy_copp_isc_medsd_w |> summary()
-all_cats$Care_Access_Status_w |> unique()
+all_cats$Care_Access_Status |> unique()
 ```
@@
 all_cats <- all_cats |> 
-  mutate(bivariate_entropy_w = 
-           case_when( ((Care_Access_Status_w == "[04-25 des.) Caring - complete (2 to 4 options per cat.)"  |
-                       Care_Access_Status_w =="[04-25 des.) Caring - very complete (4 or more options per cat.)" |
-                         Care_Access_Status_w =="[25+ des.) Very caring - very complete (4 or more options per cat.)") &
-                         entropy_copp_isc_medsd_w >= 0.7154) ~ "Caring, complete and amenity diverse",  #the median diversity,
-
-                     ((Care_Access_Status_w == "[04-25 des.) Caring - complete (2 to 4 options per cat.)"  |
-                       Care_Access_Status_w =="[04-25 des.) Caring - very complete (4 or more options per cat.)" |
-                         Care_Access_Status_w =="[25+ des.) Very caring - very complete (4 or more options per cat.)") &
-                        entropy_copp_isc_medsd_w < 0.7154) ~ "Caring, complete but not amenity diverse",
-
-                     ((Care_Access_Status_w == "[0-04 des.) Not caring " |
-                       Care_Access_Status_w == "[04-25 des.) Caring ")  &
-                        entropy_copp_isc_medsd_w >= 0.7154) ~ "Not caring or not complete, but amenity diverse",
-
-                     ((Care_Access_Status_w == "[0-04 des.) Not caring "  |
-                       Care_Access_Status_w == "[04-25 des.) Caring ")  &
-                        entropy_copp_isc_medsd_w < 0.7154) ~ "Not caring, complete, or amenity diverse"),
-
-         bivariate_entropy = 
+  mutate(bivariate_entropy = 
            case_when( ((Care_Access_Status == "[04-25 des.) Caring - complete (2 to 4 options per cat.)"  |
@@
   tm_shape(all_cats) +
-  tm_polygons("bivariate_entropy_w",
+  tm_polygons("bivariate_entropy",
               border.alpha = 0.2,
               palette = c("yellowgreen","yellow","blue", "#FFFFDF"),
-              title = "Diversity of (all - weighted) care destinations spatially accessible within 15 min. walk",
+              title = "Diversity of (all) care destinations spatially accessible within 15 min. walk",
               legend.is.portrait = TRUE) +
@@
 tmap_save(all_cats_bivariate_entropy_copp_plot, file="all_cats_bivariate_entropy_copp_plot.tiff")
-```
-
-```{r}
-all_cats_entropy_copp_weighted_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
-  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
-  tm_shape(all_cats) +
-  tm_polygons("entropy_copp_isc_medsd_w",
-              border.alpha = 0.2,
-              palette = "Reds", style="quantile", n=4,
-              title = "Diversity of (all - weighted) care destinations spatially accessible within 15 min. walk",
-              legend.is.portrait = TRUE) +
-  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
-  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
-  tm_compass(position = c("right", "top"), size=1.0)+
-  tm_layout(bg.color = "grey",
-            panel.label.bg.color = "white",
-            panel.label.color = "black",
-            legend.outside = TRUE,
-            legend.position = c("center", "bottom"),
-            legend.outside.position = "bottom",
-            legend.outside.size = 0.2,
-            legend.width = 12)
-
-tmap_save(all_cats_entropy_copp_weighted_plot, file="all_cats_entropy_copp_weighted_plot.tiff")
 ```
@@
 
-Questions:
+<!-- Questions: -->
 
-- what is a reasonable travel time? Is 15 minute walk indeed fine for our purpose?
+<!-- - what is a reasonable travel time? Is 15 minute walk indeed fine for our purpose? -->
 
-- is 1 potential opportunity enough per subcategory?
+<!-- - is 1 potential opportunity enough per subcategory?  -->
 
-- Should some subcategories be substitutes? E.g., 1 daycare OR EARLYON centre is enough, not 1 of each.
+<!-- - Should some subcategories be substitutes? E.g., 1 daycare OR EARLYON centre is enough, not 1 of each.  -->
 
-- Are there ways the categories can overlap to tell a story? Maybe to include google maps images. eg. 
-  - communities where children+errand+grocery centric care is easier -- especially for women
-  - health_centric+Elder_centric -- people with higher health care needs
+<!-- - Are there ways the categories can overlap to tell a story? Maybe to include google maps images. eg.  -->
+<!--   - communities where children+errand+grocery centric care is easier -- especially for women -->
+<!--   - health_centric+Elder_centric -- people with higher health care needs -->
 
-- care trips 30% of all trips made. they are essential and have
+<!-- - care trips 30% of all trips made. they are essential and have -->
 

### 2024-05-27 paper/Manuscript.qmd
part 2 removed weighted calculations in 'preliminary-analysis'

@@
 
+Self organizing map (**SOM**) is used to profile the types of neighbourhoods by their magnitude in 15-minute walking accessibility to caring opportunities, diversity in these opportunities and socio-economic characteristics.
+SOM is a type of artificial neural network (ANN) trained using unsupervised learning that produces a two-dimensional representation of the input space.
+It is a dimensionality reduction method as it applies competitive learning by way of a neighbourhood function thereby preserving the topological properties of the input space and producing combinations of the input space that we intuitively profile.
+
 # Results
@@
 
-BONUS!
-Per capita spatial availability.
-This is not discussed in the methods.
-The spatial availability calculation assumes population is 'population density' and weighted opportunity count.
-With this map we can start thinking about per capita availability of opportunities.
-```{r plot-figure6, out.width=600, out.height=600}
-#| label: fig-Fig6
-#| fig-cap: "Per capita spatial availability of care opportunities.
-Uses weighted care opportunities, population density as competition, and 15 minute binary cut-off.
-Aggregates parcel level spatial availability at the DA level (by summation) and divides by the DA population.
-Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_vvi_plot.tiff"))
-```
-
-`r word_comment("If you're interested in how the unweighted results look like for the same maps see the appendix.
-I've also calculated the cycling travel matrices as well as the 'no dependent' i.e., the same analysis but now only considering 3 care categories: health, errand, and grocery.
-For the time being though let's focus on this.", " ")`
-
 # References
@@
 
-\pagebreak
-
-# Appendix
-
-```{r plot-figure7, out.width=600, out.height=600}
-#| label: fig-Fig7
-#| fig-cap: "UNWEIGHTED -- The cumulative opportunity measure.
-The number of care destinations that can be reached, per DA, within 15 min walk.
-Visualised as 'caring' (based on the opportunity reachable number; Caring is approximately the median) and if complete (if all care categories can be reached; this is an arbitrary selection).
-Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_copp_plot.tiff"))
-```
-
-```{r plot-figure8, out.width=600, out.height=600}
-#| label: fig-Fig8
-#| fig-cap: "UNWEIGHTED -- The diversity measure based on the _specific type_ of destination cumulative opportunity DA measure.
-Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_entropy_copp_plot.tiff"))
-```
+<!-- `r word_comment("If you're interested in how the unweighted results look like for the same maps see the appendix.
+I've also calculated the cycling travel matrices as well as the 'no dependent' i.e., the same analysis but now only considering 3 care categories: health, errand, and grocery.
+For the time being though let's focus on this.", " ")` -->

### 2024-05-28 data/01-census-and-traveltimes.qmd
Added more variables to "HAM_census_21.rda" using 01-census-and-traveltimes.qmd. Played around in 03-clustering-and-profile.qmd

@@
                             vectors=c("v_CA21_11", #total - aged 0-14
-                                      "v_CA21_12", #males - aged 0-14
-                                      "v_CA21_13", #females - aged 0-14
                                       "v_CA21_68", #total - aged 15-64
-                                      "v_CA21_69", #males - aged 15-64
-                                      "v_CA21_70", #females - aged 15-64
                                       "v_CA21_251", #total - aged 65 and over
-                                      "v_CA21_252", #males - aged 65 and over
-                                      "v_CA21_253", #females - aged 65 and over
-                                      "v_CA21_1085", #total - prevalence of LICO-AT (%)
-                                      "v_CA21_1086", #male - prevalence of LICO-AT (%)
-                                      "v_CA21_1087",#female - prevalence of LICO-AT (%)
-                                      "v_CA21_7647", #total - walked to work
-                                      "v_CA21_7635", #total - drove or was driven to work
-                                      "v_CA21_7644", #total - used transit to get to work
-                                      "v_CA21_7650", #total - biked to work
-                                      "v_CA21_7653" #total - used some other mode to get to work
+
+                                       "v_CA21_498", #Average number of children in census families with children
+                                      "v_CA21_500", # No. of couple families
+                                      "v_CA21_507", # No. of (non-couple) one-parent families
+                                      "v_CA21_1085", #Prevalence of LICO-AT (%)
+
+                                      "v_CA21_1103", #No. of pop in priv households in the Adjusted after-tax family income decile group in the BOTTOM HALF of distribution
+                                      "v_CA21_1121", #No. of pop in priv households in the Adjusted after-tax family income decile group in the TOP HALF of distribution
+
+                                      "v_CA21_1142", #Gini index on adjusted household after-tax income
+                                      "v_CA21_907", #Median after-tax income of household in 2020 ($)
+
+                                      "v_CA21_945", # no. of households making under $5,000 (AT income)
+                                      "v_CA21_946", # no. of households making $5000-$9999 (AT income)
+                                      "v_CA21_947", # no. of households making $10000-$14999 (AT income)
+                                      "v_CA21_948", # no. of households making $15000-$19999 (AT income)
+                                      "v_CA21_949", # no. of households making $20000-$24999 (AT income)
+                                      "v_CA21_950", # no. of households making $25000-$29999 (AT income)
+                                      "v_CA21_951", # no. of households making $30000-$34999 (AT income)
+                                      "v_CA21_952", # no. of households making $35000-$39999 (AT income)
+                                      "v_CA21_953", # no. of households making $40000-$44999 (AT income)
+                                      "v_CA21_954", # no. of households making $45000-$49999 (AT income)
+                                      "v_CA21_955", # no. of households making $50000-$59999 (AT income)
+                                      "v_CA21_956", # no. of households making $60000-$69999 (AT income)
+                                      "v_CA21_957", # no. of households making $70000-$79999 (AT income)
+                                      "v_CA21_958", # no. of households making $80000-$89999 (AT income)
+                                      "v_CA21_959", # no. of households making $90000-$99999 (AT income)
+                                      "v_CA21_960", # no. of households making over $100000 (AT income)
+
+                                      "v_CA21_6498",# No. of pop. who are employed (in the labour force)
+                                      "v_CA21_6501",#  No. of pop. who are not employed (in the labour force)
+                                      "v_CA21_6504",#  No. of pop. who are not in the labour force
+
+                                      "v_CA21_1144",# TOTAL - knowledge of official languages for the total population excluding institutional residents
+                                       "v_CA21_1156",# No.
+of pop population with no knowledge of official languages (either English or French), NOTE: _1144 - 1156 = the No.
+of pop who have knowledge of at least 1 off.
+lang.
+
+                                      "v_CA21_4875",# No. of pop who are visible minority
+                                      "v_CA21_4914",# No. of pop who are are NOT visible minority
+
+                                      "v_CA21_5868",#  No. of pop w/ No certificate, diploma or degree for the population aged 25 to 64 years in private households
+                                      "v_CA21_5871",#  No. of pop (age 25 to 64 year) w/ Highest degree is a High (secondary) school diploma or equivalency certificate
+                                      "v_CA21_5874",# No. of pop (age 25 to 64 year) w/ Highest degree is Postsecondary certificate, diploma or degree (bachelor or higher)
+
+                                      "v_CA21_7647", # % total - walked to work
+                                      "v_CA21_7635", # % total - drove or was driven to work
+                                      "v_CA21_7644", # % total - used transit to get to work
+                                      "v_CA21_7650", # % total - biked to work
+                                      "v_CA21_7653" # % total - used some other mode to get to work
                                       ), 

### 2024-05-28 data/02-access-and-prelim-plots.qmd
Added more variables to "HAM_census_21.rda" using 01-census-and-traveltimes.qmd. Played around in 03-clustering-and-profile.qmd

--- 2024-05-27 data/02-preliminary-analysis.qmd
@@
 library(scales)
+library(here)
 ```
@@
   rowwise() |> mutate(
-         perc_walked = (`v_CA21_7647: Walked`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`, `v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE),
+         perc_walked = (`v_CA21_7647: Walked`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`,`v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE),
          perc_car = (`v_CA21_7635: Car, truck or van`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`, `v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE),
@@
 #ID = care destination ID (in care_dest), GeoUID is id of the DA, from_id is parcel ID
-save(access_15, file = "access_15.rda", compress = "xz")
+save(access_15, file= file.path(here("data/access_15.rda")), compress = "xz")
 ```
@@
             copp_iscDA_sd = sd(copp_isc), #the SD
-    copp_iscDA_w_med = median(copp_isc_w), #the median number of opportunities (all) that can be reached from each parcel in a DA
-            copp_iscDA_w_sd = sd(copp_isc_w), #the SD
             V_iscDA = sum(V_isc), #the number of total spatially available opportunities (all) that can be reached from each parcel in a DA
@@
 ```{r}
-#calculating entropy, here I do it by the 14 specific categories.
-This is a way to consider their diversity within and between categories.
-Because they have been weighted before calc.
-the cumulative opp counts, we can proceed without anymore adjustments.
-step1 <- access_15_byspecificcategory_DA  |> group_by(GeoUID) |> 
-  summarise(OPPi = sum(copp_isc_med, na.rm=T) + sum(copp_isc_sd, na.rm=T),
+#calculating entropy, here I do it by the 4 categories for each parcel and then aggregate to DA.
+Another way is by the 14 specific categories.
+For now - keep it simple and just focus on the 4 categories.
+This assumes that all specific types are the SAME within categories.
+step1 <- access_15  |> group_by(from_id) |> 
+  summarise(OPPi = sum(copp_isc, na.rm=T),
             Vi = sum(V_isc))
 
-step2 <- access_15_byspecificcategory_DA |> left_join(step1, by="GeoUID")
+step2 <- access_15 |> left_join(step1, by="from_id")
 
-step2 <- step2 |> mutate(top1_copp = (copp_isc_med+copp_isc_sd) / OPPi * log((copp_isc_med+copp_isc_sd) / OPPi),
+step2 <- step2 |> mutate(
+  top1_copp = (copp_isc) / OPPi * log((copp_isc) / OPPi),
   top1_copp = ifelse(is.nan(top1_copp), 0, top1_copp),
+
   top1_V= (V_isc) / Vi * log((V_isc) / Vi),
-             top1_V = ifelse(is.nan(top1_V), 0, top1_V),)  |> 
+  top1_V = ifelse(is.nan(top1_V), 0, top1_V),
+
+  entropy_copp_isc = sum(top1_copp) / -log(4),
+  entropy_V_isc = sum(top1_V) / -log(4))  |> 
   group_by(GeoUID) |>
-  summarise(entropy_copp_isc_medsd = sum(top1_copp) / -log(14),
-            entropy_V_isc = sum(top1_V) / -log(14))
+  summarise(entropy_copp_isc = mean(entropy_copp_isc, na.rm=T) + sd(entropy_copp_isc, na.rm=T),
+            entropy_V_isc = mean(entropy_V_isc, na.rm=T) + sd(entropy_V_isc, na.rm=T))
 
@@
                                                    all(copp_all >= 4) ~ "- very complete (4 or more options per cat.)"),
-                   entropy_copp_isc_medsd = first(entropy_copp_isc_medsd),
+                   entropy_copp_isc = first(entropy_copp_isc),
                    entropy_V_isc = first(entropy_V_isc)) |>
@@
               palette = c("darkred", "orangered", "orange", "yellow", "forestgreen", "lightgreen"),
-              title = "Number of care destinations (all) spatially accessible within 15 min. walk",
+              title = "Number of care destinations spatially accessible within 15 min. walk",
               legend.is.portrait = TRUE) +
@@
 
-tmap_save(all_cats_copp_plot, file="all_cats_copp_plot.tiff")
+tmap_save(all_cats_copp_plot, file= file.path(here("figures/all_cats_copp_plot.tiff")), dpi=600)
 ```
@@
   tm_shape(all_cats) +
-  tm_polygons("entropy_copp_isc_medsd",
+  tm_polygons("entropy_copp_isc",
               border.alpha = 0.2,
               palette = "Reds",
-              title = "Diversity of (all) care destinations spatially accessible within 15 min. walk",
+              title = "Diversity of care destinations spatially accessible within 15 min. walk",
               legend.is.portrait = TRUE) +
@@
 
-tmap_save(all_cats_entropy_copp_plot, file="all_cats_entropy_copp_plot.tiff")
+tmap_save(all_cats_entropy_copp_plot, file= file.path(here("figures/all_cats_entropy_copp_plot.tiff")), dpi=600)
 ```
@@
               labels = c("Below Median", "Median and above"),
-              title = "Per capita spatially available care destinations (all - weighted) within 15 min. walk",
+              title = "Per capita spatially available care destinations within 15 min. walk",
               legend.is.portrait = TRUE) +
@@
 
-tmap_save(all_cats_vvi_plot, file="all_cats_vvi_plot.tiff")
+tmap_save(all_cats_vvi_plot, file= file.path(here("figures/all_cats_vvi_plot.tiff")), dpi=600)
 ```
@@
                          Care_Access_Status =="[25+ des.) Very caring - very complete (4 or more options per cat.)") &
-                         entropy_copp_isc_medsd >= 0.7418) ~ "Caring, complete and amenity diverse",  #the median diversity,
+                         entropy_copp_isc >= 0.7418) ~ "Caring, complete and amenity diverse",  #the median diversity,
 
@@
                          Care_Access_Status =="[25+ des.) Very caring - very complete (4 or more options per cat.)") &
-                        entropy_copp_isc_medsd < 0.7418) ~ "Caring, complete but not amenity diverse",
+                        entropy_copp_isc < 0.7418) ~ "Caring, complete but not amenity diverse",
 
@@
                        Care_Access_Status == "[04-25 des.) Caring ")  &
-                        entropy_copp_isc_medsd >= 0.7418) ~ "Not caring or not complete, but amenity diverse",
+                        entropy_copp_isc >= 0.7418) ~ "Not caring or not complete, but amenity diverse",
 
@@
                        Care_Access_Status == "[04-25 des.) Caring ")  &
-                        entropy_copp_isc_medsd < 0.7418) ~ "Not caring, complete, or amenity diverse"))
+                        entropy_copp_isc < 0.7418) ~ "Not caring, complete, or amenity diverse"))
 
@@
               palette = c("yellowgreen","yellow","blue", "#FFFFDF"),
-              title = "Diversity of (all) care destinations spatially accessible within 15 min. walk",
+              title = "Diversity of care destinations spatially accessible within 15 min. walk",
               legend.is.portrait = TRUE) +
@@
 
-tmap_save(all_cats_bivariate_entropy_copp_plot, file="all_cats_bivariate_entropy_copp_plot.tiff")
+tmap_save(all_cats_bivariate_entropy_copp_plot, 
+          file= file.path(here("figures/all_cats_bivariate_entropy_copp_plot.tiff")),
+          dpi=600)
 ```
@@
             panel.label.size = 0.8)
-care_categories_plot
 
-tmap_save(care_categories_plot, filename ="care_categories_plot.tiff")
+tmap_save(care_categories_plot, 
+          file= file.path(here("figures/care_categories_plot.tiff")),
+          dpi=600)
 ```
@@
 pop_descriptive_plots <- tmap_arrange(parcels_DA_plot, poprate_DA_plot, LICOAT_DA_plot, nrow=3)
-tmap_save(pop_descriptive_plots, file="pop_descriptive_plots.tiff", asp = 1/0.6)
+
+tmap_save(pop_descriptive_plots, 
+          file= file.path(here("figures/pop_descriptive_plots.tiff")) , asp = 1/0.6,
+          dpi=600)
 ```

### 2024-05-28 data/03-clustering-and-profiling.qmd
Added more variables to "HAM_census_21.rda" using 01-census-and-traveltimes.qmd. Played around in 03-clustering-and-profile.qmd

@@
   transmute(scluster = factor(scluster),
-            LICO = `v_CA21_1085: Prevalence of low income based on the Low-income cut-offs, after tax (LICO-AT) (%)`)
+            LICO = `v_CA21_1085: Prevalence of low income based on the Low-income cut-offs, after tax (LICO-AT) (%)`,
+            Unemployed = `v_CA21_6501: Unemployed`/(`v_CA21_6498: Employed` + `v_CA21_6501: Unemployed` + `v_CA21_6504: Not in the labour force`),
+            Income_gini = `v_CA21_1142: Gini index on adjusted household after-tax income`,
+            Vis_minority = `v_CA21_4875: Total visible minority population`/ (`v_CA21_4914: Not a visible minority`+ `v_CA21_4875: Total visible minority population`),
+            Age_0to14_and_Age_65plus =( `v_CA21_11: 0 to 14 years` + `v_CA21_251: 65 years and over`) / (`v_CA21_11: 0 to 14 years` + `v_CA21_68: 15 to 64 years` + `v_CA21_251: 65 years and over`),
+            One_parent_families = `v_CA21_507: Total one-parent families` / (`v_CA21_507: Total one-parent families` + `v_CA21_500: Total couple families`) )
+
+# # examining correlations
+# cor(df |> select(-c("scluster")), use = "complete.obs")
 ```

### 2024-05-28 paper/Manuscript.qmd
Added more variables to "HAM_census_21.rda" using 01-census-and-traveltimes.qmd. Played around in 03-clustering-and-profile.qmd

@@
 Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics(glue::glue(here::here(), "/data/pop_descriptive_plots.tiff"))
+knitr::include_graphics(glue::glue(here::here(), "/figures/pop_descriptive_plots.tiff"))
 ```
@@
 Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics(glue::glue(here::here(), "/data/care_categories_plot.tiff"))
+knitr::include_graphics(glue::glue(here::here(), "/figures/care_categories_plot.tiff"))
 ```
@@
 Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_copp_plot.tiff"))
+knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_copp_plot.tiff"))
 ```
@@
 Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_entropy_copp_plot.tiff"))
+knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_entropy_copp_plot.tiff"))
 ```
@@
 The Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_bivariate_entropy_copp_plot.tiff"))
+knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_bivariate_entropy_copp_plot.tiff"))
 ```

### 2024-06-03 data/01-census-and-traveltimes.qmd
Draft using SOM method

@@
                                       "v_CA21_7650", # % total - biked to work
-                                      "v_CA21_7653" # % total - used some other mode to get to work
+                                      "v_CA21_7653", # % total - used some other mode to get to work
+
+                                      "v_CA21_4306", # % of owner households with a mortgage,
+                                      "v_CA21_4307", # % of owner households spending 30% or more of its income on shelter costs,
+                                      "v_CA21_4308", # % of owner households in core housing need
+                                      "v_CA21_4314", # % of tenant households in subsidized housing
+                                      "v_CA21_4315", # % of tenant households spending 30% or more of its income on shelter costs,
+                                      "v_CA21_4316" # % in core housing need
+
                                       ), 

### 2024-06-03 data/02-access-and-prelim-plots.qmd
Draft using SOM method

@@
 ```{r}
-rm(list = ls())
+# rm(list = ls())
 ```
 
-Working with the parcel to dest TTM - walk and cycling.
-
+Working with the parcel to dest TTM - walk mode.
 ```{r library-setup}
@@
 
-load("parcel-ttm/ttm_care_bike_parcel1.rda")
-load("parcel-ttm/ttm_care_bike_parcel2.rda")
-load("parcel-ttm/ttm_care_bike_parcel3.rda")
-load("parcel-ttm/ttm_care_bike_parcel4.rda")
+# load("parcel-ttm/ttm_care_bike_parcel1.rda")
+# load("parcel-ttm/ttm_care_bike_parcel2.rda")
+# load("parcel-ttm/ttm_care_bike_parcel3.rda")
+# load("parcel-ttm/ttm_care_bike_parcel4.rda")
 load("parcel-ttm/ttm_care_walk_parcel1.rda")
@@
 
-ttm_care_bike_parcel <-rbind(ttm_care_bike_parcel1, ttm_care_bike_parcel2,
-                             ttm_care_bike_parcel3, ttm_care_bike_parcel4)
-rm(ttm_care_bike_parcel1, ttm_care_bike_parcel2, ttm_care_bike_parcel3, ttm_care_bike_parcel4)
+# ttm_care_bike_parcel <-rbind(ttm_care_bike_parcel1, ttm_care_bike_parcel2,
+#                              ttm_care_bike_parcel3, ttm_care_bike_parcel4)
+# rm(ttm_care_bike_parcel1, ttm_care_bike_parcel2, ttm_care_bike_parcel3, ttm_care_bike_parcel4)
 
@@
 ttm_care_walk_parcel <- left_join(ttm_care_walk_parcel, R_PARCELS_CENTS_GeoUIDs_IDs, by = c("from_id"="ID"))
-ttm_care_bike_parcel <- left_join(ttm_care_bike_parcel, R_PARCELS_CENTS_GeoUIDs_IDs, by = c("from_id"="ID"))
+#ttm_care_bike_parcel <- left_join(ttm_care_bike_parcel, R_PARCELS_CENTS_GeoUIDs_IDs, by = c("from_id"="ID"))
 ```
@@
 
-```{r}
+```{r, eval=FALSE}
 # all Care_Category_Specific should be represented, i.e., each ID should have 14 rows -- if that category is empty it should be 0. Let's add those 0s. 
@@
 Aggregating time!
-```{r}
+```{r, eval=FALSE}
 #ID = care destination ID (in care_dest), GeoUID is id of the DA, from_id is parcel ID
 save(access_15, file= file.path(here("data/access_15.rda")), compress = "xz")
+```
+```{r}
+load(file= file.path(here("data/access_15.rda")))
 ```
@@
 #by DA and mean:
-access_15_DA <- access_15 |>
+access_15_DA_2 <- access_15 |>
   left_join(HAM_census_21 |> st_drop_geometry() |> select(c("GeoUID", "Population", "Dwellings", "LICO_AT_TOTAL", "COMMUNITY_")), 
@@
 R_PARCELS_CENTS_count_GeoUID <- R_PARCELS_CENTS_GeoUIDs_IDs |> group_by(GeoUID) |> summarise(parcel_count = n())
-access_15_DA <- access_15_DA |> left_join(R_PARCELS_CENTS_count_GeoUID, by="GeoUID")
-access_15_DA <- left_join(HAM_census_21 |> select(c(GeoUID)), access_15_DA, by=c("GeoUID"))
+access_15_DA_2 <- access_15_DA_2 |> left_join(R_PARCELS_CENTS_count_GeoUID, by="GeoUID")
+access_15_DA_2 <- left_join(HAM_census_21 |> select(c(GeoUID)), access_15_DA_2, by=c("GeoUID"))
 
@@
 #checks
-# access_15_DA |> summary()
+# access_15_DA_2 |> summary()
 access_15_bycategory_DA |> summary()
-access_15_DA$V_iscDA |> sum(na.rm=T)
+access_15_DA_2$V_iscDA |> sum(na.rm=T)
 access_15_bycategory_DA$V_ic |> sum(na.rm=T)
@@
 ```
-
-```{r message=FALSE}
+<!--
+SUPERCEEDED BY THE PLOTS IN 03-CLUSTERING..qmd
+```{r message=FALSE, eval=FALSE}
 #tmap_mode("plot")
@@
 
-```{r}
+```{r, eval=FALSE}
 all_cats_entropy_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
@@
 ```
+--->
 
-```{r}
+```{r, eval=FALSE}
 all_cats_vvi_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
@@
 
-```{r}
+```{r, eval=FALSE}
 all_cats_bivariate_entropy_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
@@
 tmap_save(all_cats_bivariate_entropy_copp_plot, 
-          file= file.path(here("figures/all_cats_bivariate_entropy_copp_plot.tiff")),
-          dpi=600)
+          file= file.path(here("figures/all_cats_bivariate_entropy_copp_plot.png")),
+          dpi=300)
 ```
 
-```{r}
+```{r, eval=FALSE}
 care_categories_plot <- 
@@
 tmap_save(care_categories_plot, 
-          file= file.path(here("figures/care_categories_plot.tiff")),
-          dpi=600)
+          file= file.path(here("figures/care_categories_plot.png")),
+          dpi=300)
 ```
+
 ```{r}
-parcels_DA_plot <- 
- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
-  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
-  tm_shape(access_15_DA) +
-  tm_polygons("parcel_count",
-              border.alpha = 0.2,
-              palette = "Reds",style = "quantile", n=4,
-              title = "Count of residential parcels",
-              legend.is.portrait = TRUE) +
-  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
-  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
-  tm_compass(position = c("left", "top"), size=1.0)+
-   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5,legend.title.size = 0.7,  legend.bg.color = "white", legend.bg.alpha = 0.4)
+## SUPERCEDED
 
-poprate_DA_plot <- 
- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
-  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
-  tm_shape(access_15_DA |> mutate(popden = Population/Dwellings)) +
-  tm_polygons("popden",
-              border.alpha = 0.2,
-              palette = "Purples", style = "quantile", n=4,
-              title = "Population per dwelling",
-              legend.is.portrait = TRUE) +
-  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
-  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
-  tm_compass(position = c("left", "top"), size=1.0)+
-   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5,legend.title.size = 0.7,  legend.bg.color = "white", legend.bg.alpha = 0.4)
-
-LICOAT_DA_plot <- 
- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
-  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
-  tm_shape(access_15_DA) +
-  tm_polygons("LICO_AT_TOTAL",
-              border.alpha = 0.2,
-              palette = "Oranges", style = "quantile", n=4,
-              title = "Prevalance of household LICO-AT",
-              legend.is.portrait = TRUE) +
-  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
-  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
-  tm_compass(position = c("left", "top"), size=1.0)+
-   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5, legend.title.size = 0.7, legend.bg.color = "white", legend.bg.alpha = 0.4)
-
-pop_descriptive_plots <- tmap_arrange(parcels_DA_plot, poprate_DA_plot, LICOAT_DA_plot, nrow=3)
-
-tmap_save(pop_descriptive_plots, 
-          file= file.path(here("figures/pop_descriptive_plots.tiff")) , asp = 1/0.6,
-          dpi=600)
+# parcels_DA_plot <- 
+#  tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+#   tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
+#   tm_shape(access_15_DA_2) +
+#   tm_polygons("parcel_count",
+#               border.alpha = 0.2,
+#               palette = "Reds",style = "quantile", n=4,
+#               title = "Count of residential parcels",
+#               legend.is.portrait = TRUE) +
+#   tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
+#   tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
+#   tm_compass(position = c("left", "top"), size=1.0)+
+#    tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5,legend.title.size = 0.7,  legend.bg.color = "white", legend.bg.alpha = 0.4)
+# 
+# poprate_DA_plot <- 
+#  tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+#   tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
+#   tm_shape(access_15_DA_2 |> mutate(popden = Population/Dwellings)) +
+#   tm_polygons("popden",
+#               border.alpha = 0.2,
+#               palette = "Purples", style = "quantile", n=4,
+#               title = "Population per dwelling",
+#               legend.is.portrait = TRUE) +
+#   tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
+#   tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
+#   tm_compass(position = c("left", "top"), size=1.0)+
+#    tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5,legend.title.size = 0.7,  legend.bg.color = "white", legend.bg.alpha = 0.4)
+# 
+# LICOAT_DA_plot <- 
+#  tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+#   tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
+#   tm_shape(access_15_DA_2) +
+#   tm_polygons("LICO_AT_TOTAL",
+#               border.alpha = 0.2,
+#               palette = "Oranges", style = "quantile", n=4,
+#               title = "Prevalance of household LICO-AT",
+#               legend.is.portrait = TRUE) +
+#   tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
+#   tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
+#   tm_compass(position = c("left", "top"), size=1.0)+
+#    tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5, legend.title.size = 0.7, legend.bg.color = "white", legend.bg.alpha = 0.4)
+# 
+# pop_descriptive_plots <- tmap_arrange(parcels_DA_plot, poprate_DA_plot, nrow=2)
+# 
+# tmap_save(pop_descriptive_plots, 
+#           file= file.path(here("figures/pop_descriptive_plots.tiff")) , asp = 1/0.6,
+#           dpi=600)
 ```

### 2024-06-03 data/03-clustering-and-profiling.qmd
Draft using SOM method

--- 2024-05-25 data/03-clustering-and-profiling.qmd
@@
 ---
-title: "Clustering"
+title: "Clustering, profiling and plots"
 author: "Antonio Paez & Anastasia Soukhov"
@@
 
-## Quarto
+## Setting up table
 
@@
 library(rpart.plot) # Plot 'rpart' Models: An Enhanced Version of 'plot.rpart'
-library(skimr)
-#library(spdep) # Spatial Dependence: Weighting Schemes, Statistics
+library(skimr) # Compact and Flexible Summaries of Data
+library(sf) # Simple Features for R
 library(SOMbrero) # SOM Bound to Realize Euclidean and Relational Outputs
 library(tidyverse) # Easily Install and Load the 'Tidyverse'
-
-# library(formattable) # Create 'Formattable' Data Structures
-# library(kableExtra) # Construct Complex Table with 'kable' and Pipe Syntax
-# library(knitr) # A General-Purpose Package for Dynamic Report Generation in R
-# library(plotly) # Create Interactive Web Graphics via 'plotly.js'
-# library(sf) # Simple Features for R
-# library(spdep) # Spatial Dependence: Weighting Schemes, Statistics
-# library(SOMbrero) # SOM Bound to Realize Euclidean and Relational Outputs
-# library(tidyverse) # Easily Install and Load the 'Tidyverse'
-# library(tmap) # Thematic Maps
+library(tmap)
+library(here)
+library(parsnip)
 ```
 
-Read od table:
+Read od table and the census:
 ```{r}
 load("access_15.rda")
+load("HAM_census_21.rda")
 ```
 
-Aggregate by Care Category:
+Loading boundaries for mapping:
+```{r}
+Community_Boundaries <- st_read("data-raw/boundaries/Community_Boundaries.shp") |> mutate(COMMUNITY_ = ifelse(COMMUNITY_ == "Hamilton", "Hamilton-Central", COMMUNITY_))|> st_transform(crs=4326)
+City_Boundary <- st_read("data-raw/boundaries/City_Boundary.shp") |> st_transform(crs=4326)
+hydro_p_LakeOntario <- st_read("data-raw/boundaries/hydro_p_LakeOntario.shp") |> st_transform(crs=4326)
+ham_bay <- st_read("data-raw/boundaries/Waterbodies.shp") |> st_transform(crs=4326) |> filter(FEATURE_TY == "Lake")
+ham_bay_cropped<-st_crop(ham_bay$geometry, Community_Boundaries$geometry)
+hydro_p_LakeOntario_cropped<-st_crop(hydro_p_LakeOntario$geometry, Community_Boundaries$geometry)
+```
+
+Aggregate by Care Category (dependent, errand, health, and grocery centric):
 ```{r}
@@
            Care_Category) |>
-  summarize(copp_isc = sum(copp_isc),
-            .groups = "drop")
+  summarize(GeoUID = first(GeoUID),
+            copp_isc = sum(copp_isc),
+            .groups = "drop") |> 
+  left_join(HAM_census_21 |> # and dssociate with census data and drop parcels in DAs with 0 population:
+              st_drop_geometry(),
+            by = "GeoUID")
+
+access_15_aggregated <- access_15_aggregated |> filter(Population >= 0 | !is.na(Population))
 ```
@@
 
-Now this table can be used for the clustering analysis.
+## Plotting population descriptions, DA
+
+For plotting, create a spatial DA sf object that summarizes these inputs:
+```{r}
+access_15_DA <- left_join(access_15_wide, 
+                          access_15_aggregated |> group_by(from_id) |>
+  summarise(GeoUID = first(GeoUID),
+            tot_copp_isc = first(tot_copp_isc),
+            Population = first(Population),
+            LICO = first(`v_CA21_1085: Prevalence of low income based on the Low-income cut-offs, after tax (LICO-AT) (%)`),
+            Income_bottom_n = first( `v_CA21_1103: In bottom half of the distribution`),
+            Income_top_n = first(`v_CA21_1121: In top half of the distribution`),
+            Age_0to14_n = first(`v_CA21_11: 0 to 14 years`),
+            Age_15to64_n = first(`v_CA21_68: 15 to 64 years`),   
+            Age_65plus_n = first(`v_CA21_251: 65 years and over`),
+            One_parent_families_n = first(`v_CA21_507: Total one-parent families`),
+            Couple_families_n = first(`v_CA21_500: Total couple families`)),
+            by=c("from_id")) |>
+  group_by(GeoUID)|>
+  summarise(Population = first(Population),
+            parcel_count = n(),
+            LICO = first(LICO),
+            Income_bottom_n = first(Income_bottom_n),
+            Income_top_n = first(Income_top_n),
+            Age_0to14_n = first(Age_0to14_n),
+            Age_15to64_n = first(Age_15to64_n),   
+            Age_65plus_n = first(Age_65plus_n),
+            One_parent_families_n = first(One_parent_families_n),
+            Couple_families_n = first(Couple_families_n),
+
+            med_tot_copp_isc = median(tot_copp_isc),
+            med_div = median(div),
+            med_dep_copp = median(total_Dependent_centric),
+            med_err_copp = median(total_Errand_centric),
+            med_groc_copp = median(total_Grocery_centric),
+            med_health_copp = median(total_Health_centric)) |>
+  left_join(HAM_census_21 |> select(c(GeoUID)), by=c("GeoUID")) |> st_as_sf() |> st_make_valid()
+
+access_15_DA <- access_15_DA |> mutate(Populationparcel = Population/parcel_count,
+                                       Income_bottom = Income_bottom_n / (Income_bottom_n + Income_top_n),
+                                       Age_0to14 = Age_0to14_n / (Age_0to14_n + Age_15to64_n + Age_65plus_n),
+                                       One_parent_families = One_parent_families_n / (One_parent_families_n + Couple_families_n))
+
+access_15_DA$parcel_count |> sum() #a check
+```
+
+```{r, eval=FALSE}
+parcels_DA_plot <-
+ tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
+  tm_shape(access_15_DA) +
+  tm_polygons("parcel_count",
+              border.alpha = 0.2,
+              palette = "Reds",style = "quantile", n=4,
+              title = "Count of residential parcels",
+              legend.is.portrait = TRUE) +
+  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
+  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
+  tm_compass(position = c("left", "top"), size=1.0)+
+   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5,legend.title.size = 0.7,  legend.bg.color = "white", legend.bg.alpha = 0.4)
+
+pop_DA_plot <-
+ tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
+  tm_shape(access_15_DA) +
+  tm_polygons("Population",
+              border.alpha = 0.2,
+              palette = "Greens", style = "quantile", n=4,
+              title = "Population",
+              legend.is.portrait = TRUE) +
+  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
+  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
+  tm_compass(position = c("left", "top"), size=1.0)+
+   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5,legend.title.size = 0.7,  legend.bg.color = "white", legend.bg.alpha = 0.4)
+
+poprate_DA_plot <-
+ tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
+  tm_shape(access_15_DA) +
+  tm_polygons("Populationparcel",
+              border.alpha = 0.2,
+              palette = "Purples", style = "quantile", n=4,
+              title = "Population per residential parcel rate",
+              legend.is.portrait = TRUE) +
+  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
+  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
+  tm_compass(position = c("left", "top"), size=1.0)+
+   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5,legend.title.size = 0.7,  legend.bg.color = "white", legend.bg.alpha = 0.4)
+
+pop_descriptive_plots_1 <- tmap_arrange(parcels_DA_plot, pop_DA_plot, poprate_DA_plot, nrow=3)
+tmap_save(pop_descriptive_plots_1,
+          file= file.path(here("figures/pop_descriptive_plots_1.png")) , asp = 1/0.6,
+          dpi=300)
+```
+
+```{r, eval=FALSE}
+INCOME_B_DA_plot <-
+ tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
+  tm_shape(access_15_DA) +
+  tm_polygons("Income_bottom",
+              border.alpha = 0.2,
+              palette = "Greens", style = "quantile", n=4,
+              title = "% hld below AT median income",
+              legend.is.portrait = TRUE) +
+  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
+  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
+  tm_compass(position = c("left", "top"), size=1.0)+
+   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5, legend.title.size = 0.7, legend.bg.color = "white", legend.bg.alpha = 0.4)
+
+AGE014_DA_plot <-
+ tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
+  tm_shape(access_15_DA) +
+  tm_polygons("Age_0to14",
+              border.alpha = 0.2,
+              palette = "Oranges", style = "quantile", n=4,
+              title = "% population aged 0 to 14",
+              legend.is.portrait = TRUE) +
+  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
+  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
+  tm_compass(position = c("left", "top"), size=1.0)+
+   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5, legend.title.size = 0.7, legend.bg.color = "white", legend.bg.alpha = 0.4)
+
+ONE_PARENT_DA_plot <-
+ tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
+  tm_shape(access_15_DA) +
+  tm_polygons("One_parent_families",
+              border.alpha = 0.2,
+              palette = "Reds", style = "quantile", n=4,
+              title = "% one parent hlds",
+              legend.is.portrait = TRUE) +
+  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
+  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
+  tm_compass(position = c("left", "top"), size=1.0)+
+   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5, legend.title.size = 0.7, legend.bg.color = "white", legend.bg.alpha = 0.4)
+
+LICOAT_DA_plot <-
+ tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
+  tm_shape(access_15_DA) +
+  tm_polygons("LICO",
+              border.alpha = 0.2,
+              palette = "Purples", style = "quantile", n=4,
+              title = "Prevalance of household LICO-AT",
+              legend.is.portrait = TRUE) +
+  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
+  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
+  tm_compass(position = c("left", "top"), size=1.0)+
+   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5, legend.title.size = 0.7, legend.bg.color = "white", legend.bg.alpha = 0.4)
+
+pop_descriptive_plots_2 <- tmap_arrange(INCOME_B_DA_plot,AGE014_DA_plot, ONE_PARENT_DA_plot, LICOAT_DA_plot, ncol=2)
+
+tmap_save(pop_descriptive_plots_2,
+          file= file.path(here("figures/pop_descriptive_plots_2.png")) , asp = 1/1,
+          dpi=300)
+```
+
+## Plotting accessibility and diversity index, DA
+
+```{r, eval=FALSE}
+all_cats_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
+  tm_shape(access_15_DA) +
+  tm_polygons("med_tot_copp_isc",
+              border.alpha = 0.2,
+              palette = c("darkred", "orangered", "forestgreen", "lightgreen"),
+              style = "quantile", n=4,
+              #breaks = c(0,5,14,27,108),
+              title = "The number of care destinations within a 15 min. walk for a median parcel",
+              label = c("[0 to 5) Not caring",
+                        "[5 to 14) Somewhat caring",
+                        "[14 to 27) Caring",
+                        "[27 to 108] Very caring"),  #these are for quantiles, the clusters p50 seem to roughly follows these.
+              legend.is.portrait = TRUE) +
+  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
+  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
+  tm_compass(position = c("right", "top"), size=1.0)+
+  tm_layout(bg.color = "grey",
+            panel.label.bg.color = "white",
+            panel.label.color = "black",
+            legend.outside = TRUE,
+            legend.position = c("center", "bottom"),
+            legend.outside.position = "bottom",
+            legend.outside.size = 0.2,
+            legend.width = 12
+            # legend.title.size = 0.7,
+            # legend.text.size = 0.6
+            )
+tmap_save(all_cats_copp_plot, file= file.path(here("figures/all_cats_copp_plot.png")), dpi=300)
+```
+
+```{r, eval=FALSE}
+dep_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
+  tm_shape(access_15_DA) +
+  tm_polygons("med_dep_copp",
+              border.alpha = 0.2,
+              style = "quantile", n=4,
+              palette = "Oranges",
+              title = "Dependent-centric",) +
+  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
+  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
+  tm_compass(position = c("left", "top"), size=1.0)+
+  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4,
+            bg.color = "grey",
+            legend.position = c("right","top"))
+
+health_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
+  tm_shape(access_15_DA) +
+  tm_polygons("med_health_copp",
+              border.alpha = 0.2,
+              style = "quantile", n=4,
+              palette = "Purples",
+              title = "Health-centric",) +
+  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
+  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
+  tm_compass(position = c("left", "top"), size=1.0)+
+  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4,
+            bg.color = "grey",
+            legend.position = c("right","top"))
+
+groc_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
+  tm_shape(access_15_DA) +
+  tm_polygons("med_groc_copp",
+              border.alpha = 0.2,
+              style = "quantile", n=4,
+              palette = "Greens",
+              title = "Grocery-centric",) +
+  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
+  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
+  tm_compass(position = c("left", "top"), size=1.0)+
+  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4,
+            bg.color = "grey",
+            legend.position = c("right","top"))
+
+err_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
+  tm_shape(access_15_DA) +
+  tm_polygons("med_err_copp",
+              border.alpha = 0.2,
+              style = "quantile", n=4,
+              palette = "Reds",
+              title = "Errand-centric",) +
+  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
+  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
+  tm_compass(position = c("left", "top"), size=1.0)+
+  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4,
+            bg.color = "grey",
+            legend.position = c("right","top"))
+
+percats_copp_plot <- tmap_arrange(dep_copp_plot, err_copp_plot, groc_copp_plot, health_copp_plot, nrow=2)
+
+tmap_save(percats_copp_plot, 
+          file= file.path(here("figures/percats_copp_plot.png")),
+          dpi=300)
+```
+
+```{r}
+# very caring (mean 4 to 18.6; total p50 46) - very complete (div=0.9; p50 0.90) (Cluster 2)
+# caring (mean 2.1 to 9; total p50 18) -  complete (div=0.81; p50 0.82) (Cluster 1)
+# somewhat caring (0.2 to 6; total p50 13) - complete (div=0.71; p50 0.74) (Cluster 4)
+# somewhat caring (0.2 to 6; total p50 7) - not complete (div=0.36; p50 0.40) (Cluster 3)
+# not caring (0 to 4.7; total p50 3) - not complete (div=0.02; p50 0) (Cluster 5)
+```
+
+```{r, eval=FALSE}
+all_cats_entropy_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
+  tm_shape(access_15_DA) +
+  tm_polygons("med_div",
+              border.alpha = 0.2,
+              style = "quantile", n=3, #are the quantiles, but are also similar to the clustered profiles
+              label = c("[0 to 0.49) Not complete", "[0.49 to 0.85) Complete", "[0.85 to 1] Very complete"),
+              palette = "Greys",
+              title = "Diversity of care destinations within a 15-minute walk for a median parcel",
+              legend.is.portrait = TRUE) +
+  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
+  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
+  tm_compass(position = c("right", "top"), size=1.0)+
+  tm_layout(bg.color = "grey",
+            panel.label.bg.color = "white",
+            panel.label.color = "black",
+            legend.outside = TRUE,
+            legend.position = c("center", "bottom"),
+            legend.outside.position = "bottom",
+            legend.outside.size = 0.2,
+            legend.width = 12)
+
+tmap_save(all_cats_entropy_copp_plot, file= file.path(here("figures/all_cats_entropy_copp_plot.png")), dpi=300)
+```
+
+## Creating clusters (from parcels)
+
+Now access_15_wide can be used for the clustering analysis.
 
@@
 
-The lower left corner of the map concentrates monsters with high speed, high special attack, and high special attack.
-the upper left corner concentrates high defense and high special defense.
-High hit points are towards the center of the map.
-
+The lower right corner of the map concentrates monsters has low levels of all variables.
+The high concentration of all variables is found in the upper left corner.
 Two additional way of visualizing the characteristics of the clusters are as follows.
@@
 access.sc <- superClass(access.som,
-                        k = 6)
+                        k = 5)
 summary(access.sc)
 ```
-
-This is a dendogram with the super-clusters:
-```{r}
+The combined dendogram and grid-map of the super-clusters. Turned off the variance explained plot:
+```{r, eval=FALSE}
+png(file= file.path(here("figures/access_sc_plot_1.png")), width= 2404, height= 1600, units="px",  res=300)
 plot(access.sc, 
-     plot.var = TRUE)
+     type= c("dendrogram"),
+     plot.var = FALSE,
+     show.names = TRUE,
+     main = " ",
+     ylab="Dissimilarity Index",
+     xlab= " ")
+dev.off()
 ```
 
-And the map with the superclusters:
-```{r}
+```{r, eval=FALSE}
+png(file= file.path(here("figures/access_sc_plot_2.png")), width= 2404, height= 1600, units="px",  res=300)
 plot(access.sc,
      type = "grid", 
+     plot.var = FALSE,
+     main = " ",
      plot.legend = TRUE)
+dev.off()
 ```
@@
   select(-c(cluster,
-            from_id, 
             starts_with("prop"), 
@@
          div = round(div, digits = 3))
+
+df <- df |> 
+  mutate(Total_copp = Dependent + Grocery + Errand + Health)
 ```
@@
 ```{r}
-classmod.som <- rpart(scluster ~ ., 
-                      data = df)
+classmod.som <- rpart(data = df,
+                      formula = scluster ~ (Dependent + Grocery + Errand + Health + div))
+
+# Predict the labels and add them to the dataset
+predicted_labels <- predict(classmod.som, df) |> as.data.frame()
+
+df$SC1 <- predicted_labels$`1`
+df$SC2 <- predicted_labels$`2`
+df$SC3 <- predicted_labels$`3`
+df$SC4 <- predicted_labels$`4`
+df$SC5 <- predicted_labels$`5`
+
+# df <- df |> mutate(max_label_val = pmax(predicted_labels$`1`, predicted_labels$`2`, predicted_labels$`3`, predicted_labels$`4`, predicted_labels$`5`),
+#                    max_label = case_when(max_label_val == predicted_labels$`1` ~ "SC1",
+#                                          max_label_val == predicted_labels$`2` ~ "SC2",
+#                                          max_label_val == predicted_labels$`3` ~ "SC3",
+#                                          max_label_val == predicted_labels$`4` ~ "SC4",
+#                                          max_label_val == predicted_labels$`5` ~ "SC5"))
+
+df_DA <- df |> left_join(access_15 |> dplyr::select(c("from_id","GeoUID")), by = "from_id") |> 
+  group_by(GeoUID) |> 
+  summarise(
+    SC1 = median(SC1),
+    SC2 = median(SC2),
+    SC3 = median(SC3),
+    SC4 = median(SC4),
+    SC5 = median(SC5),
+    max_label_val = pmax(SC1, SC2, SC3, SC4, SC5),
+            med_tot_copp_isc = median(Total_copp),
+            med_div = median(div),
+            med_dep_copp = median(Dependent),
+            med_err_copp = median(Errand),
+            med_groc_copp = median(Grocery),
+            med_health_copp = median(Health)) |>
+  mutate(max_label = case_when(max_label_val == SC1 ~ "b. Caring - complete (SC1)",
+                               max_label_val == SC2 ~ "a. Very caring - very complete (SC2)",
+                               max_label_val == SC3 ~ "d. Somewhat caring - not complete (SC3)",
+                               max_label_val == SC4 ~ "c. Somewhat caring - complete (SC4)",
+                               max_label_val == SC5 ~ "d. not caring - not complete (SC5)")) |>
+  left_join(HAM_census_21 |> select(c(GeoUID)), by=c("GeoUID")) |> st_as_sf() |> st_make_valid() 
+```
+```{r, eval=FALSE}
+total_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
+  tm_shape(df_DA) +
+  tm_polygons("med_tot_copp_isc",
+              border.alpha = 0.2,
+              style = "quantile", n=4,
+              palette = "Purples",
+              title = "Total access to care destinations",) +
+  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
+  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
+  tm_compass(position = c("left", "top"), size=1.0)+
+  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4,
+            bg.color = "grey",
+            legend.position = c("right","top"))
+
+SClabel_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
+  tm_shape(df_DA) +
+  tm_polygons("max_label",
+              border.alpha = 0.2,
+              palette = c("darkolivegreen","red", "dodgerblue", "darkorchid","darkgreen"),
+              labels = (c("Very caring - very complete (SC2)","Caring - Complete (SC1)",
+              "Somewhat caring - complete (SC4)","Somewhat caring - somewhat complete (SC5)", 
+              "Somewhat caring - somewhat complete (SC3)")),
+              title = "Prevelant supercluster") +
+  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
+  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
+  tm_compass(position = c("left", "top"), size=1.0)+
+  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4,
+            bg.color = "grey",
+            legend.position = c("right","top"))
+SClabel_copp_plot
+SCLab_and_total_copp_plot <- tmap_arrange(SClabel_copp_plot, total_copp_plot, nrow=2)
+
+tmap_save(SCLab_and_total_copp_plot, 
+          file= file.path(here("figures/SCLab_and_total_copp_plot.png")),
+          dpi=300)
 ```
@@
 
-Frequency of parcels in the super clusters (look at the data of the first node):
+Frequency of parcels in the super clusters:
 ```{r}
-freq_sc <- ggplot(classmod.sompar[1]$data) +
+freq_sc <- ggplot(access_15_wide |> 
+         mutate(scluster = factor(scluster))) +
   geom_bar(aes(x = "", 
+               group = scluster, 
                fill = scluster)) +
-  scale_fill_discrete(name = "super cluster") +
-  theme_void()
+  scale_fill_discrete(name = "Supercluster") #+
+  #theme_void() +
+  #guides(fill = guide_legend(nrow = 1)) +
+  #theme(legend.position = "bottom")
 
@@
   geom_edge() +
-  geom_edge_label(size = 1.5) +
-  geom_node_splitvar(size = 2) +
+  geom_edge_label(size = 1.6) +
+  geom_node_label(aes(label = splitvar),
+                  id = "inner", 
+                  size = 2, nudge_y = -0.022) +
   # pass list to gglist containing all ggplot components we want to plot for each
@@
                                theme(axis.text = element_blank())))
+party_plot
 ```
@@
 ```{r}
-party_plot
+png(file= file.path(here("figures/decision_tree_profiles_sc.png")),width= 2404, height= 1600, units="px",  res=300)
+plot_grid(party_plot, legd, nrow = 1, rel_widths = c(1, 0.2))
+dev.off()
 ```
+
+Cluster 6 is high in suckitude: low in diversity, with 4.6 Dependent Centric destinations on average, and typically no other destinations: zero Errand centric destinations, and only about 3.5% of parcels have access to Grocery-centric destinations.
+```{r}
+access_15_wide |>
+  select(scluster, starts_with("total_"), div) |>
+  rename_with(.fn = ~str_replace(.x, "total_", "")) |>
+  filter(scluster == 1) |>
+  skim()
+```
+
+What I've deduced from these superclusters:
+
+very caring (mean 4 to 18.6; total p50 46) - very complete (div=0.9; p50 0.90) (Cluster 2)
+caring (mean 2.1 to 9; total p50 18) - complete (div=0.81; p50 0.82) (Cluster 1)
+somewhat caring (0.2 to 6; total p50 13) - complete (div=0.71; p50 0.74) (Cluster 4)
+somewhat caring (0.2 to 6; total p50 7) - not complete (div=0.36; p50 0.40) (Cluster 3)
+not caring (0 to 4.7; total p50 3) - not complete (div=0.02; p50 0) (Cluster 5)
+
+## Demographic profile of clusters
+
+Select data for classification:
+```{r}
+df <- access_15_wide |>
+  transmute(scluster = factor(scluster),
+            Income_bottom = `v_CA21_1103: In bottom half of the distribution`/(`v_CA21_1103: In bottom half of the distribution` + `v_CA21_1121: In top half of the distribution`),
+            LICO = `v_CA21_1085: Prevalence of low income based on the Low-income cut-offs, after tax (LICO-AT) (%)`,
+            Avg_numb_of_kids = `v_CA21_498: Average number of children in census families with children`,
+            Not_in_labforce = `v_CA21_6504: Not in the labour force`/(`v_CA21_6498: Employed` + `v_CA21_6501: Unemployed` + `v_CA21_6504: Not in the labour force`),
+            Income_gini = `v_CA21_1142: Gini index on adjusted household after-tax income`,
+            Vis_minority = `v_CA21_4875: Total visible minority population`/ (`v_CA21_4914: Not a visible minority`+ `v_CA21_4875: Total visible minority population`),
+            Age_0to14 =(`v_CA21_11: 0 to 14 years`) / (`v_CA21_11: 0 to 14 years` + `v_CA21_68: 15 to 64 years` + `v_CA21_251: 65 years and over`),
+            One_parent_families = `v_CA21_507: Total one-parent families` / (`v_CA21_507: Total one-parent families` + `v_CA21_500: Total couple families`),
+            walk_towork =  `v_CA21_7647: Walked` / (`v_CA21_7647: Walked` + `v_CA21_7635: Car, truck or van` + `v_CA21_7644: Public transit` + `v_CA21_7650: Bicycle` + `v_CA21_7653: Other method` ),
+            Owner_hld_CHN = `v_CA21_4308: % in core housing need (57)`,
+            Hld_ownership = `v_CA21_4306: % of owner households with a mortgage (58)`,
+            Tenant_subsi_hld = `v_CA21_4314: % of tenant households in subsidized housing (61)`,
+            Tenant_hld_CHH =`v_CA21_4316: % in core housing need (57)`,
+            Population,
+            Low_or_no_deg = (`v_CA21_5868: No certificate, diploma or degree` + `v_CA21_5871: High (secondary) school diploma or equivalency certificate`) / (`v_CA21_5868: No certificate, diploma or degree` + `v_CA21_5871: High (secondary) school diploma or equivalency certificate` + `v_CA21_5874: Postsecondary certificate, diploma or degree`)
+            )
+
+# examining correlations
+cor(df |> select(-c("scluster")), use = "complete.obs")
+
+#Train a decision tree using the data:
+classdemo.som <- rpart(data = df,
+                      formula = scluster ~ (Income_bottom + Age_0to14 + One_parent_families + Vis_minority + Not_in_labforce+ Income_gini+walk_towork+Owner_hld_CHN+Hld_ownership+Tenant_subsi_hld+Tenant_hld_CHH),
+                      weights = Population
+                      )
+# # Visualize the decision tree:
+# rpart.plot(classdemo.som, cex = 0.5, type = 0, varlen = 8)
+```
+
+```{r}
+# convert the rpart object to a party object:
+classdemo.sompar <- as.party(classdemo.som)
+
+#Make a prettier plot:
+party_plot <- ggparty(classdemo.sompar) +
+  geom_edge() +
+  #geom_edge_label(size = 1.8) +
+  geom_edge_label(mapping = aes(label = paste(substr(breaks_label, start = 1, stop = 15))), size = 3.5) +
+  geom_node_splitvar(id = "inner", size = 4) +
+  #geom_edge_label(mapping = aes(label = prettyNum(breaks_label, digits = 1)), size = 1.8) +
+
+  # # pass list to gglist containing all ggplot components we want to plot for each
+  # # (default: terminal) node
+  geom_node_plot(gglist = list(geom_bar(aes(x = "", fill = scluster),
+                                        position = position_fill()),
+                               xlab(""),
+                               ylab(""),
+                               theme_void(),
+                               theme(axis.text = element_blank())))
+legd <- cowplot::get_legend(freq_sc)
+```
+
+```{r}
+#plot and save
+png(file= file.path(here("figures/access_sc_profiles.png")),width= 2404, height= 1600, units="px",  res=300)
+plot_grid(party_plot, legd, nrow = 1, rel_widths = c(1, 0.2))
+dev.off()
+```
+
+This is interesting!
+So for parcels with less than 45% of bottom-income people (richer) -- parcels with higher proportion of 1 parent households (13% and above) are more likely to have bad care/incomplete access (c3) than _really_ bad care/very incomplete access (c5).
+
+For parcels with higher than 45% of bottom-income (poorer), those with high a high proportion of children (age 0 to 14) are likely to have bad care/incomplete acces (c3).
+But those with a lower proportion of children - especially 9.8% LICO and above are going to have _super high_ care/complete (c1) or under 9.8% LICO will have _high_ care/complete (c2)
+For less than 63%, it's a different story. If between 45% and 63%, they are described by cluster 1 (caring and complete).

### 2024-06-03 paper/Manuscript.qmd
Draft using SOM method

--- 2024-05-27 paper/Manuscript.qmd
@@
 library(cowplot)
+library(magick)
 library(dplyr)
@@
 # Introduction
+
+`r word_comment("We should focus this discussion on: defining mobility of care and connecting it to the 15-minute neighbourhoods, the importance of walk-accessibility, and socio-demographic/economic characteristics associated with this overlap (esp.
+single parent households, areas with a high proportion of children, low-income prevalence).", " ")`
 
@@
 
-1)	15 minute cities stem from conventional methods that ignore mobility of care perspectives. There is an applied way of doing this, 
+1)	15 minute cities stem from conventional methods that ignore mobility of care perspectives. There is an applied way of doing this.
 2)	Use of the innovative destination dataset
-3)	Mid-size typical Canadian city, how frequently do we find these caring neighbourhoods (potential - accessibility). 
-4)	Who enjoys them and who does not? Clustering methodology. 
-a.	Types of profiles will emerge – i.e. with LICO > X + Prevalence of children >Y + walking accessibility score > U + diversity score > W = “caring neighbourhood”
+3)	Mid-size typical Canadian city case study. How frequently do we find these caring neighbourhoods (potential - accessibility)? Clustering methodology to determine this.
+4)	Who enjoys them and who does not? Clustering methodology to determine these profiles.
+  - i.e. with LICO > X + Prevalence of children >Y + walking accessibility score > U + diversity score > W = “caring neighbourhood”
 
@@
 
-- Next: describe hamilton's land-use context; communities which are urban, rural, suburban.
+- Next: we describe Hamilton's land-use context; communities which are urban, rural, suburban.
 Also describe the residential parcels summary statistics, ~140,000 points, aggregated at the level of Canadian Census DA.
-See population descriptive plots in @fig-Fig1:
+See parcel, population and population per parcel plots in @fig-Fig1:
 ```{r plot-figure1, out.width=600, out.height=600}
 #| label: fig-Fig1
-#| fig-cap: "The number of residential parcels per DA in 2020 (top) and the rate of population per dwelling per DA (middle) and prevalence of Low-income cut-offs after-tax (LICO-AT) per DA (bottom) retrieved from the 2021 Canadian Census.
+#| fig-cap: "The number of residential parcels per DA in 2020 (top), the population (middle) retrieved from the 2021 Canadian Census, and the rate of population per parcel per DA (bottom).
 All scales in quartiles.
 Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics(glue::glue(here::here(), "/data/pop_descriptive_plots.tiff"))
+knitr::include_graphics(glue::glue(here::here(), "/figures/pop_descriptive_plots_1.png"))
+```
+Some of the socio-economic and demographic variables that characterise accessibility and the diversity of accessibility to care destinations are pictured in @fig-Fig2:
+
+```{r plot-figure2, out.width=600, out.height=600}
+#| label: fig-Fig2
+#| fig-cap: "Socio-economic and demographic variables that characterise accessibility to care destinations retrieved from the 2021 Canadian Census.
+All scales in quartiles.
+Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
+knitr::include_graphics(glue::glue(here::here(), "/figures/pop_descriptive_plots_2.png"))
 ```
@@
 ## Estimated travel times
-- Next: describe the estimated 15-minute walk travel times using {r5r}, from all residential parcel points to all care destinations assuming default walking speed and OSM road network for Hamilton.
+
+Here we will describe the estimated 15-minute walk travel times using {r5r}, from all residential parcel points to all care destinations assuming default walking speed (3.6km/hr) and OSM road network for Hamilton.
 Provide summary statistics.
@@
 ## Care destination dataset
+
 A spatial dataset of care destinations for Hamilton was compiled; sources include provincial and municipal open data portals [@governmentofontarioOntarioDataCatalogue2023; @cityofhamiltonOpenData2023], Data Axle, a consumer dataset compiled of businesses and companies within Canada [@axledataConsumerData2023] and manual retrieved through Google Maps.
-Each destination was categorized based on the specific type of care being accessed following the mobility of care research by Sanchez de Madariaga & Zucchini (2019) [@sanchezdemadariagaMeasuringMobilitiesCare2019].
+Each destination was categorized based on the specific type of care being accessed following the mobility of care research by @sanchezdemadariagaMeasuringMobilitiesCare2019.
 Initial categories include child, elder,  grocery, health, and errand -centric destinations.
@@
 Category sources of data and preparation notes are detailed in @tbl-Tbl1.
-Their spatial distribution and sub-categories are visualised in @fig-Fig2.
+Their spatial distribution and sub-categories are visualised in @fig-Fig3.
 
@@
 
-```{r plot-figure2, out.width=600, out.height=600}
-#| label: fig-Fig2
+```{r plot-figure3, out.width=600, out.height=600}
+#| label: fig-Fig3
 #| fig-cap: "The locations of care destinations in Hamilton separated by the author-generated categories of: dependent-, errand-, grocery- and health- centric care categories.
 Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics(glue::glue(here::here(), "/data/care_categories_plot.tiff"))
+knitr::include_graphics(glue::glue(here::here(), "/figures/care_categories_plot.png"))
 ```
@@
 ## Methods
-Cumulative opportunity method; 15-minute normative catchment for walk mode (and cycling, travel time matrix already calculated but to be added later) is used to represent the 15-minute neighbourhood concept.
+Cumulative opportunity method; 15-minute normative catchment for walk mode is used to represent the 15-minute neighbourhood concept.
 This has been done in previous literature such as @vale2023accessibility, @graellsCityCitiesMeasuring2021 and @gaglioneUrbanAccessibility15minute2022.
 
-In this work, the cumulative opportunity score is calculated for each parcel and aggregated by care category at the DA level.
-The Q3 number of opportunities accessibility in a 15 minute walk per care category in retained for each DA.
-So in this analysis, $i$ is both the parcel and DA spatial unit.
-To the author's this is one of the first 15-minute neighbourhood analyse with a focus on care destinations.
-
-**Cumulative opportunity accessibility** takes the following form:
+In this work, the cumulative opportunity score is calculated for each parcel $i$ and aggregated by care category $c$ and takes following mathetical form:
 $$
@@
 
-Next, the Q3 (median + standard deviation) of all $\sum_{c}S_i^c$ values for each DA are taken and represented as $S^{c}_{DA}$.
-$S^{c}_{DA}$ is the upper quartile number of weighted care destinations that can be potentially interacted with within a 15 minute walk by parcels within a DA: an optimistic perspective on how many opportunities can be potentially reached.
-
 *Entropy* measure has been widely used to characterise the diversity of land-use mix in early work by @frankLinkingObjectivelyMeasured2005 based on the species diversity Shannon-Wiener index that expresses relative evenness throughout a sequence (this measure among other species diversity indices of the time are reviewed in @whittaker1972evolution).
@@
 
-In our work, each destination is a specific type $st$ (i.e., `r care_dest$Care_Category_Specific |> unique()`) as well as a care category $c$.
-When $S_i^c$ is aggregated for each category at the $DA$ it reflects how all opportunities from all four care categories can be reached.
-The four care categories have meaning to the mobility of care conceptualisation.
-However, it doesn't explicitly reflect the specific type of destination.
-In our work, the diversity of the specific type of destination is of interest, we calculate $S^{st}_i$ for each parcel $i$ for each $st$ type following the equation for $S^{st}_i$.
-This upper quartile is aggregated at the $DA$ to come up with a $S^{st}_{DA}$.
-<!--AS NOTE: potentially remove 'st' and just have diveristy be based on 'c'.
-For now, let's keep as is..
--->
-
-As such, the entropy measure is used to reflect the diversity of specific type of care destinations.
-The value ranges from 0 to 1, with 1 representing total evenness in the number of (weighted) specific type $st$ care opportunities can be reached at the aggregated spatial unit $DA$:
+As such, the entropy measure is used to reflect the diversity of care destinations categories.
+The value ranges from 0 to 1, with 1 representing total evenness in the number of care opportunities in each category that can be reached.
+$i$ is a parcel, $c$ is the category of care destinations, $n_c$ is the count of care destination categories (in this work 4), $S_i^c$ is the spatial accessibility of care destinations within a 15-minute walk for a parcel, and $D_i$ is the diversity.
+As $D_i$ represents evenness in categories, a parcel $i$ can be assigned a diversity of 1 if it has $S_i^c$=1 of each category or if it has $S_i^c$=100 of each category.
+Contrariwise, it can be assigned a low score if it has low but different scores for each category as well as if it has high but different scores for each category.
 
 $$
-D^{st}_{DA} = \frac{-\sum_{c} S_i^{st} / S^{st}_{DA} \times ln(S_i^{st} / S^{st}_{DA} )}{ln(ST)}
+D_{i} = \frac{-\sum_{c} S_i^{c} / \sum_{c}S^{c}_{i} \times ln(S_i^{c} / \sum_{c}S^{c}_{i} )}{ln(n_c)}
 $$
-\noindent Where:
 
-- $i$ is a set of parcel point origin locations.
-- $st$ is a set of specific type care categories. In our work they are `r care_dest$Care_Category_Specific |> unique()`.
-- $DA$ is a set of DA spatial units (to aggregate parcel point data).
-- $S_{i}^{st}$ is the unconstrained accessibility, the sum of opportunities reachable within $f(\cdot)$, at each $i$ for each $st$.
-- $S^{st}_{DA}$ is the aggregated unconstrained accessibility at each $DA$. In our work, it is the upper quartile of $S_i^{st}$ values in a $DA$. 
-- $ST$ is the total sum of categories. In our work, this is equal to 14. 
-- $D^{st}_{DA}$ is the entropy index.
-
-Self organizing map (**SOM**) is used to profile the types of neighbourhoods by their magnitude in 15-minute walking accessibility to caring opportunities, diversity in these opportunities and socio-economic characteristics.
+Self organizing map (**SOM**) is used to cluster the types of neighbourhoods by their magnitude in 15-minute walking accessibility to caring opportunities and diversity in these opportunities and profile these clusters based on their socio-economic characteristics.
 SOM is a type of artificial neural network (ANN) trained using unsupervised learning that produces a two-dimensional representation of the input space.
 It is a dimensionality reduction method as it applies competitive learning by way of a neighbourhood function thereby preserving the topological properties of the input space and producing combinations of the input space that we intuitively profile.
+It has been used in transportation related research, e.g., @victorianoTimeSpaceMoney2020.
+
+The SOM procedure is broken into 2 steps: first, five superclusters of the four $S_i^c$ scores and the diversity score $D_i$ are identified.
+The number of superclusters strike a balance between variance explained and supercluster interpretation in what we define as ranging from "not caring" to "very caring" and "not complete" to "complete".
+Second, the superclusters are profiled using a variety of socio-economic and demographic indicators that the mobility of care literature finds significant (@fig-Fig2).
 
 # Results
-<!--Initial thoughts on results:
--Initial results seem optimistic. With SOM , method will classify neighbourhoods into these clusters we then tag/name. 
-- A part of this - spatial knowledge of people doesn't match the dataset. i.e. all stores are not equal, they have qualities that make them less/more accessibly. 
-- But we can speak in potentials. They facilities are there and they could potentially be improved. -->
-```{r plot-figure3, out.width=600, out.height=600}
-#| label: fig-Fig3
-#| fig-cap: "The cumulative opportunity measure.
-The number of care destinations that can be reached, per DA, within 15 min walk.
-Visualised as 'caring' (based on the opportunity reachable number; Caring is approximately the median) and if complete (if all care categories can be reached; this is an arbitrary selection).
+
+Accessibility per category @fig-Fig4:
+```{r plot-figure4, out.width=600, out.height=600}
+#| label: fig-Fig4
+#| fig-cap: "The number of care destinations that can be reached within 15 min walk per care category for a median parcel in each DA.
+Scales are in quartiles.
+These are four of five inputs into the superclusters.
 Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_copp_plot.tiff"))
+knitr::include_graphics(glue::glue(here::here(), "/figures/percats_copp_plot.png"))
 ```
 
-```{r plot-figure4, out.width=600, out.height=600}
-#| label: fig-Fig4
-#| fig-cap: "The diversity measure based on the _specific type_ of destination cumulative opportunity DA measure.
+Diversity @fig-Fig5:
+```{r plot-figure5, out.width=600, out.height=600}
+#| label: fig-Fig5
+#| fig-cap: "The diversity measure based on the proportion of care category spatial accessibility (figure 4).
+Scale is in tertiles.
+The fifth input into the superclusters.
 Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_entropy_copp_plot.tiff"))
+knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_entropy_copp_plot.png")) 
 ```
 
-```{r plot-figure5, out.width=600, out.height=600}
-#| label: fig-Fig5
-#| fig-cap: "A bivariate map of the cumulative opportuninities map (aggregated per category at DA - fig3) and the diversity measure (fig 4).
-Amenity diverse = diviersity index above the region median.
-Caring/Complete = accessibility value above the region median.
-The Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics(glue::glue(here::here(), "/data/all_cats_bivariate_entropy_copp_plot.tiff"))
+Superclusters based on accessibility per category and diversity (the two above figures). 5 clusters are identified, see their dendrogram (right) and grid map (left) in @fig-Fig6 :
+
+```{r plot-figure6-create, eval=FALSE}
+plot1 <- ggdraw() + draw_image(glue::glue(here::here(), "/figures/access_sc_plot_2.png"))
+plot2 <- ggdraw() + draw_image(glue::glue(here::here(), "/figures/access_sc_plot_1.png"))
+plot3 <- plot_grid(plot1, plot2)
+
+ggsave(glue::glue(here::here(), "/figures/access_sc_plot_combo.png"), dpi=300)
 ```
+
+```{r plot-figure6, out.width=800, out.height=800}
+#| label: fig-Fig6
+#| fig-cap: "The 5 resulting superclusters from the SOM output (left) and the associated dendrogram structure (right)."
+#| 
+knitr::include_graphics(glue::glue(here::here(), "/figures/access_sc_plot_combo.png")) 
+```
+
+The constitution of the superclusters in shown in the decision tree diagram in @fig-Fig7.
+Decision trees area a supervised learning technique, which can use the results of the clustering algorithm as labels, and the input variables as features.
+The clustering algorithm performed well as class membership is relatively homogeneous as the majority of bar plots consist of a supermajority of one type of supercluster class membership, i.e., supercluster 5 are parcels that have an Errand-centric access of less than 0.5 opportunities (bottom quartile), a diversity score of less than 0.277 (bottom tertile), and any Dependent-centric access score of greater than 0.5 (and can take on any value of the remaining input variables).
+This technique provides a data-driven approach to supercluster formation which we then attribute meaning to.
+
+```{r plot-figure7, out.width=600, out.height=600}
+#| label: fig-Fig7
+#| fig-cap: "The decision tree profiles of the SOM output."
+knitr::include_graphics(glue::glue(here::here(), "/figures/decision_tree_profiles_sc.png"))
+```
+
+The dendrogram demonstrating the supercluster class membership of each parcel in @fig-Fig7 is summarized, with our interpretation, as follows:
+- Very caring (mean category accessibility between 4 to 18.6 and total accessibility median of 46) - very complete (diversity index of mean 0.9 and median 0.90) (SCluster 2)
+- Caring (mean category accessibility between 2.1 to 9 and total accessibility median of 18) -  complete (div index of mean 0.81 and median 0.82) (SCluster 1)
+- Somewhat caring (mean category accessibility between 0.2 to 6 and total accessibility median of 13) - complete div index of mean 0.71 and median 0.74) (SCluster 4)
+- Somewhat caring (mean category accessibility between 0.2 to 6 and total accessibility median 7) - not complete (div index of mean 0.36 and median 0.40) (SCluster 3)
+- Not caring (mean category accessibility between 0 to 4.7 and total accessibility median of 3) - not complete (div index of mean 0.02 and median 0) (SCluster 5)
+
+To further visualize supercluster membership, we spatially demonstrate which supercluster best describes each parcel in @fig-Fig8 (top plot): the median parcel's supercluster membership per DA is tabulated and the majority membership is represented.
+Visually, the super clusters closely follow the quartile split of the accessibility to care destinations (all categories combined) and tertile split of the diversity scores (visualised in @fig-Fig5).
+
+```{r plot-figure8, out.width=600, out.height=600}
+#| label: fig-Fig8
+#| fig-cap: "The median parcel supercluster membership (top).
+The parcel median total accessibility (bottom).
+The parcel m Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
+knitr::include_graphics(glue::glue(here::here(), "/figures/SCLab_and_total_copp_plot.png"))
+```
+To enhance the meaning of the superclusters beyond a spatial conceptualisation of "caring" and "complete", was _who?_ resides there.
+We implement another decision tree in which the population-weighted socio-demographic characteristics of the parcels (retrieved from the 2021 Canadian Census) are input variables and the resulting output are 'socio-domographic superclusters' of the caring/complete superclusters.
+The input variables included: % households below median household after-tax income, LICO prevalence, average number of children per household, % not in the labour force, Gini index on adjusted household after-tax income, % visible minority, % population aged 0 to 14, % single parent household, % who walk to work (relative to bike, care/truck/van, public transit and 'other'), % of owner household in core housing need (i.e., inadequate housing structure or paying higher than 30% income on housing), % of owner households with a mortgage, % of tenant households in subsidized housing, % of tenant households in core housing need, % no certification, diploma or degree (relative to population with high school, college, or university certificate, diploma or degree).
+
+Of all the variables, those that proved to be significant in profiling the five superclusters were only four variables: % households below median household after-tax income (bottom income), % population aged 0 to 14, % single parent household, and LICO.
+See @fig-Fig9 that provides the splits.
+Unfortunately, the supercluster membership is not as homogeneous as the caring/complete supercluster dendrogram (@fig-Fig7).
+However, we can interpret the socio-demographic superclusters (five that result) as labels to tell a narrative of who may reside in what caring/complete supercluster.
+
+```{r plot-figure9, out.width=600, out.height=300}
+#| label: fig-Fig9
+#| fig-cap: "The dendrogram of the supercluster demographic profiles.."
+knitr::include_graphics(glue::glue(here::here(), "/figures/access_sc_profiles.png"))
+```
+(NOTE: for reference, see the old bivariate map in the appendix for comparison)
+
+In @fig-Fig9, discussing from right-to-left, parcels in DAs with 45% of the population below the median or lower (richer) are often classified with poor accessibility: either "Somewhat caring and not complete" SC3 or "Not caring and not complete" SC5, where parcels within DAs with a higher proportion of single parent households (13% and above) are most often SC3 and below are most often SC5.
+For parcels with higher than 45% of bottom-income (poorer), those with a high proportion of children (age 0 to 14) are also likely to be within the "Somewhat caring and not complete" SC3 - just like parcels in richer and more-single-parent DA counterparts.
+
+These findings related to potential child-care access are notable: single parents provide more care and hence may be more likely to choose, if they are able, more accessible neighbourhoods than the rest of the population in the same household income grouping.
+In a Canadian and U.S.
+context, households with children may locate to catchments with 'better' schools, often more suburban with lower potential access scores overall.
+
+Interestingly, parcels in DAs with lower than 13% children population demonstrate a better accessibility to care destination story.
+Parcels within DAs that constitute more than 45% population below median household income (poorer) - especially above 9.8% LICO (low income cut-off prevalence - likely to be the poorest) are often within "Very caring and very complete" SC1.
+Parcels located in DAs with the same lower median income and child population but have _lower_ than 9.8% LICO (not the poorest) also have good accessibility, as they are most often in "Caring and complete" SC1 as well as being in most often in the "Somewhat caring and complete" SC4 relative to other parcels.
+
+Our findings illustrate differences in patterns of 15-minute walk accessibility to care destinations most prominently profiled by income, low-income prevalence, children and single-parent status.
 
@@
 For the time being though let's focus on this.", " ")` -->
+
+# Appendix
+
+```{r plot-figure10, out.width=600, out.height=600}
+#| label: fig-Fig10
+#| fig-cap: "A bivariate map of the cumulative opportuninities map (aggregated per category at DA) and the diversity measure.
+Amenity diverse = diviersity index above the region median.
+Caring/Complete = accessibility value above the region median.
+The Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
+knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_bivariate_entropy_copp_plot.png"))
+```

### 2024-06-07 data/03-clustering-and-profiling.qmd
Small change (reproducibility) to the clustering-and-profiling.qmd

@@
 ```{r}
+
 df <- access_15_wide |>
@@
 ```{r}
-df <- access_15_wide |>
+
+df <- access_15_wide |> left_join(access_15_aggregated, by = "from_id") |> 
   transmute(scluster = factor(scluster),
@@
             Population,
-            Low_or_no_deg = (`v_CA21_5868: No certificate, diploma or degree` + `v_CA21_5871: High (secondary) school diploma or equivalency certificate`) / (`v_CA21_5868: No certificate, diploma or degree` + `v_CA21_5871: High (secondary) school diploma or equivalency certificate` + `v_CA21_5874: Postsecondary certificate, diploma or degree`)
-            )
+            Low_or_no_deg = (`v_CA21_5868: No certificate, diploma or degree` + `v_CA21_5871: High (secondary) school diploma or equivalency certificate`) / (`v_CA21_5868: No certificate, diploma or degree` + `v_CA21_5871: High (secondary) school diploma or equivalency certificate` + `v_CA21_5874: Postsecondary certificate, diploma or degree`))
 
@@
 classdemo.som <- rpart(data = df,
-                      formula = scluster ~ (Income_bottom + Age_0to14 + One_parent_families + Vis_minority + Not_in_labforce+ Income_gini+walk_towork+Owner_hld_CHN+Hld_ownership+Tenant_subsi_hld+Tenant_hld_CHH),
+                      formula = scluster ~ (Income_bottom + Age_0to14 + One_parent_families + Vis_minority + Not_in_labforce+ Income_gini+walk_towork+Owner_hld_CHN+Hld_ownership+Tenant_subsi_hld+Tenant_hld_CHH+LICO),
                       weights = Population

### 2024-06-16 data/01-census-and-traveltimes.qmd
Incorporated Lea and Lucia's comments

@@
 save(HAM_census_21, file="HAM_census_21.rda")
+```
+
+```{r}
+load( file="HAM_census_21.rda")
 ```
@@
 
+## Transit travel times
+
+Calculate **Transit** travel times from centroids of parcels (from_ids) to destination points (to_ids). 30 min max travel time.
+
+```{r Transit travel time 1}
+# set up batching according to how many origin rows to process at one time
+chunksize = 50 
+num_chunks = ceiling(nrow(orig_1)/chunksize)
+
+# create origin-destination pairs
+origins_chunks <- as.disk.frame(orig_1,
+                          outdir = "parcel-ttm/processing/orig",
+                          nchunks = num_chunks,
+                          overwrite = TRUE)
+
+start.time <- Sys.time()
+pb <- txtProgressBar(0, num_chunks, style = 3)
+
+for (i in 1:num_chunks){
+  Orig_chunk <- get_chunk(origins_chunks, i)
+  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
+                          origins = Orig_chunk,
+                          destinations = dest,
+                          mode = c("TRANSIT"),
+                          departure_datetime = as.POSIXct(strptime("08-06-2023 08:00:00", # 12 GMT is 8am in Toronto (EDT)
+                                 format = "%d-%m-%Y %H:%M:%S")),
+                          max_trip_duration = 30)
+
+  # export output as disk.frame
+  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
+                                            nchunks = 1,
+                                            outdir = "parcel-ttm/processing/output_ttm",
+                                            compress = 50,
+                                            overwrite = TRUE),
+         add_chunk(output_df, ttm_chunk, chunk_id = i))
+  setTxtProgressBar(pb, i)
+}
+end.time <- Sys.time()
+print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))
+#"OD matrix calculation took 6.78 minutes..."
+ttm_care_TRANSIT_parcel1 <- as.data.frame(output_df)
+
+save("ttm_care_TRANSIT_parcel1", file = "parcel-ttm/ttm_care_TRANSIT_parcel1.rda")
+rm("ttm_care_TRANSIT_parcel1")
+```
+
+```{r TRANSIT travel time 2}
+# set up batching according to how many origin rows to process at one time
+chunksize = 50 
+num_chunks = ceiling(nrow(orig_2)/chunksize)
+
+# create origin-destination pairs
+origins_chunks <- as.disk.frame(orig_2,
+                          outdir = "parcel-ttm/processing/orig",
+                          nchunks = num_chunks,
+                          overwrite = TRUE)
+
+start.time <- Sys.time()
+pb <- txtProgressBar(0, num_chunks, style = 3)
+
+for (i in 1:num_chunks){
+  Orig_chunk <- get_chunk(origins_chunks, i)
+  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
+                          origins = Orig_chunk,
+                          destinations = dest,
+                          mode = c("TRANSIT"),
+                          departure_datetime = as.POSIXct(strptime("08-06-2023 08:00:00", # 12 GMT is 8am in Toronto (EDT)
+                                 format = "%d-%m-%Y %H:%M:%S")),
+                          max_trip_duration = 30)
+
+  # export output as disk.frame
+  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
+                                            nchunks = 1,
+                                            outdir = "parcel-ttm/processing/output_ttm",
+                                            compress = 50,
+                                            overwrite = TRUE),
+         add_chunk(output_df, ttm_chunk, chunk_id = i))
+  setTxtProgressBar(pb, i)
+}
+end.time <- Sys.time()
+print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))
+#"OD matrix calculation took 5.35 minutes..."
+ttm_care_TRANSIT_parcel2 <- as.data.frame(output_df)
+
+save("ttm_care_TRANSIT_parcel2", file = "parcel-ttm/ttm_care_TRANSIT_parcel2.rda")
+rm("ttm_care_TRANSIT_parcel2")
+```
+
+```{r TRANSIT travel time 3}
+# set up batching according to how many origin rows to process at one time
+chunksize = 50 
+num_chunks = ceiling(nrow(orig_3)/chunksize)
+
+# create origin-destination pairs
+origins_chunks <- as.disk.frame(orig_3,
+                          outdir = "parcel-ttm/processing/orig",
+                          nchunks = num_chunks,
+                          overwrite = TRUE)
+
+start.time <- Sys.time()
+pb <- txtProgressBar(0, num_chunks, style = 3)
+
+for (i in 1:num_chunks){
+  Orig_chunk <- get_chunk(origins_chunks, i)
+  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
+                          origins = Orig_chunk,
+                          destinations = dest,
+                          mode = c("TRANSIT"),
+                          departure_datetime = as.POSIXct(strptime("08-06-2023 08:00:00", # 12 GMT is 8am in Toronto (EDT)
+                                 format = "%d-%m-%Y %H:%M:%S")),
+                          max_trip_duration = 30)
+
+  # export output as disk.frame
+  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
+                                            nchunks = 1,
+                                            outdir = "parcel-ttm/processing/output_ttm",
+                                            compress = 50,
+                                            overwrite = TRUE),
+         add_chunk(output_df, ttm_chunk, chunk_id = i))
+  setTxtProgressBar(pb, i)
+}
+end.time <- Sys.time()
+print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))
+
+ttm_care_TRANSIT_parcel3 <- as.data.frame(output_df)
+
+save("ttm_care_TRANSIT_parcel3", file = "parcel-ttm/ttm_care_TRANSIT_parcel3.rda")
+rm("ttm_care_TRANSIT_parcel3")
+```
+
+```{r TRANSIT travel time 4}
+# set up batching according to how many origin rows to process at one time
+chunksize = 50 
+num_chunks = ceiling(nrow(orig_4)/chunksize)
+
+# create origin-destination pairs
+origins_chunks <- as.disk.frame(orig_4,
+                          outdir = "parcel-ttm/processing/orig",
+                          nchunks = num_chunks,
+                          overwrite = TRUE)
+
+start.time <- Sys.time()
+pb <- txtProgressBar(0, num_chunks, style = 3)
+
+for (i in 1:num_chunks){
+  Orig_chunk <- get_chunk(origins_chunks, i)
+  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
+                          origins = Orig_chunk,
+                          destinations = dest,
+                          mode = c("TRANSIT"),
+                          departure_datetime = as.POSIXct(strptime("08-06-2023 08:00:00", # 12 GMT is 8am in Toronto (EDT)
+                                 format = "%d-%m-%Y %H:%M:%S")),
+                          max_trip_duration = 30)
+
+  # export output as disk.frame
+  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
+                                            nchunks = 1,
+                                            outdir = "parcel-ttm/processing/output_ttm",
+                                            compress = 50,
+                                            overwrite = TRUE),
+         add_chunk(output_df, ttm_chunk, chunk_id = i))
+  setTxtProgressBar(pb, i)
+}
+end.time <- Sys.time()
+print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))
+#"OD matrix calculation took 6.23 minutes..."
+
+ttm_care_TRANSIT_parcel4 <- as.data.frame(output_df)
+
+save("ttm_care_TRANSIT_parcel4", file = "parcel-ttm/ttm_care_TRANSIT_parcel4.rda")
+rm("ttm_care_TRANSIT_parcel4")
+```
+
 Limitations of travel time calcs:

### 2024-06-16 data/03-clustering-and-profiling.qmd
Incorporated Lea and Lucia's comments

@@
               border.alpha = 0.2,
-              palette = c("darkolivegreen","red", "dodgerblue", "darkorchid","darkgreen"),
+              palette = c("#44ce1b","#bbdb44", "#f7e379", "#f2a134","#e51f1f"),
               labels = (c("Very caring - very complete (SC2)","Caring - Complete (SC1)",

### 2024-06-16 paper/Manuscript.qmd
Incorporated Lea and Lucia's comments

@@
 
-`r word_comment("We should focus this discussion on: defining mobility of care and connecting it to the 15-minute neighbourhoods, the importance of walk-accessibility, and socio-demographic/economic characteristics associated with this overlap (esp.
-single parent households, areas with a high proportion of children, low-income prevalence).", " ")`
+Caring is one of the most fundamental activities in societies, which has allowed humanity to survive and evolve.
+It involves activities targeted to meet the physical, psychological, and emotional needs of others [@ilo_care_2018].
+Yet, it is one of the most devalued, underpaid, and unequal activities worldwide.
+Traditionally, caring activities have remained on women’s shoulders.
+Though women today attain higher educational levels and tend to work out of home more than ever before, the bulk of care remains (at any life stage) heavily on women's shoulders.
+The 'Second Shift' remains mostly women’s work  in all industrial societies [@hayden_grand_2000; @hochschild_second_2012].
+According to the @ilo_care_2018, women and girls perform more than three-quarters of the total amount of unpaid care.
+This gender gap varies geographically [@RN5].
+In Canada, adult women complete 1.9 hours, on average, more of unpaid care than men every day [@RN4].
 
-Caring is one of the most fundamental activities in societies, which has allowed humanity to survive and evolve.
-It involves activities targeted to meet physical, psychological, and emotional needs of others [@ilo_care_2018].
-Yet, it is one of the worst devalued, underpaid, and unequal activities worldwide.
-Traditionally, caring activities have remained on women’s shoulders.
-Even though the society has evolved, and nowadays women attain higher educational levels and tend to work out of home more than ever before, the bulk of care remains (at any stage of her lifecourse) heavily on her shoulders.
-The Second Job or the Second Shift remains in all industrial societies, mostly women’s work (Hayden, 1982 (?Citation)) [@hochschild_second_2012].
+This unequal share of caring responsibilities leads to gendered differences in: career development, profession selection, contract type, pay gap, and time poverty as recognized by different international organisations [@EIGE2016; @ilo_care_2018].
+In fact, even early feminists in the United States observed that domestic labor was the most relevant basic cause of women’s inequality [@gordon_womans_1976; @hayden_grand_2000].
 
-This unequal share of caring responsibilities leads to gendered differences in among others, career development, differences in contract types and decision-making positions, unbalanced share of women in certain professions, gender pay gap, time poverty, as recognised by different international organisations (European Institute for Gender Equality, 2016 (?Citation)) [@ilo_care_2018].
-In fact, already the first feminists in the United States observed that domestic labor was the most relevant basic cause of women’s inequality [@gordon_womans_1976] (Hayden, 1982 (?Citation)).
-
-According to the [@ilo_care_2018], women and girls perform more than three-quarters of the total amount of unpaid care work and two-thirds of care workers are women.
-The caring economy is part of the women-dominated industries, it is characterised by the evaluative discrimination, which is only one of the dimensions of the gender pay gap [@mejiadorantes_discussing_2019].
-Especially in the case of domestic workers, care workers are most of the times women and migrants’ workers, with poor working conditions, and many times associated with informal jobs which imply lack of labour rights, no social and health protection, susceptible of exploitation, one of the poorest working conditions across the care economy as the ILO observes [@ilo_care_2018].
-Care workers becomes disrespected, leading to a devaluation of their activity, and consequently remains as one of the less well-paid occupations a person can undertake.
-
-If we turn our attention to the urban planning perspective, Caring is a major human activity that may take form in all urban settings, in all transport modes, all times a day and year, and yet it is not described as neither origin nor destination, and until recently, it was inexistent.
-
-Researchers from various fields have highlighted that western modern industrial societies have imposed a physical separation of the private sphere from the public one, splitting production and reproduction, public and private spaces or simply work and home, concentrating on the economic outcome of being mobile [@cresswell_gendered_2008; @lawWomenTransportNew1999; @levy2013travel; @little_gender_1994; @tronto_toward_1990].
+Care also has significant planning considerations.
+Indeed, researchers from various fields have highlighted that western modern industrial societies have imposed a physical separation of the private sphere from the public one, splitting production and reproduction, public and private spaces or simply work and home, concentrating on the economic outcome of being mobile [@cresswell_gendered_2008; @lawWomenTransportNew1999; @levy2013travel; @little_gender_1994; @tronto_toward_1990].
 The man with legal rights, paid labour public sphere and woman in the private sphere, dependent, with family duties, personal relations and unpaid labour [@tronto_toward_1990].
@@
 As an example of this dichotomy, transport planning historically framed non-work/school trips as 'non-mandatory' activities, downgrading these activities in comparison to work.
+In more recent years, planners have begun to resist this spatial separation of the public and private sphere to promote for sustainable cities.
+Proximity planning has become more normalized, at popular example of which is the 15-minute city, in which residents can reach all necessary destinations within a 15-minute walk.
+The 15-minute city concept, however, has yet to directly focus on care.
+
+While much care labour takes place in the private, and feminised sphere, completing care activities requires one to be mobile.
+In other words: caring for oneself and others requires trips to, amongst other places, grocery stores, day cares, elder care, and health care destinations.
 
@@
 
-From this research motivation, Sánchez de Madariaga coined the term ‘mobility of care’ in 2013 to refer to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands [@sanchezdemadariagaMobilityCareIntroducing2013].
+From this research motivation, Sánchez de Madariaga coined the term ‘mobility of care’ in 2013 to refer to this travel or all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands [@sanchezdemadariagaMobilityCareIntroducing2013].
 While decades of research had examined these unique household-serving trips, her seminal work was the first to consider all these trips as one category.
@@
 Indeed, in her original work, and in the studies that have followed, it appears that approximately 30% of adults’ trips are for care purposes [@sanchezdemadariagaMobilityCareIntroducing2013; @sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory].
+Perhaps unsurprisingly given the gendered division of care work discussed above, women have consistently been found to complete more mobility of care trips than men.
+In one study, mobility of care comprised 32% of women’s daily trips compared to 28% of men’s.
+While this gap was significant, it was found to be far greater in lower income households where women complete 20% more care trips than men [@ravensbergen2023exploratory].
 Sánchez de Madariaga not only shows how important these mobility of care trips are, but also highlights the ways in which “mobility of care is systematically under-represented in any analysis of urban transport” (p.
@@
 Indeed, asymmetric in the distribution of time of citizens is deeply anchored with the public and private spaces organization within cities.
-In this light, we focus on the potential access to destinations of care through an empirical case study of the mid-sized city of Hamilton, Canada.
+In this light, we bring together the literatures on 15-minute cities and mobility of care by focusing on the potential access to destinations of care through an empirical case study of the mid-sized city of Hamilton, Canada.
 We acknowledge that caring is composed by actors and activities but aim to spotlight the role that the urban environment plays on our daily lives, on easing our participation and on exercising our rights to the city as individuals with multiple identities.
@@
 
-<!-- We experience our urban environment depending on our personal characteristics, such as socioeconomic level, education, profession, household responsibilities, physical or mental disabilities, and very importantly, by gender.
-Furthermore, our living environment is also influenced by historical, economic, social, cultural, political, religion and environmental factors.
-All these issues have an influence on the way we live and experience the city, on how we participate in the community and on our social relations in the public and private space, on how we invest in our socio-economic development, and finally on our well-being.
--->
+## Feminist 15-minute neighbourhoods
 
-<!-- Cities have simplified individuals as a universal and neutral entity.
-As a result, public spaces are differently experienced depending on our multiple identities.
-Indeed, the distribution of time of citizens is deeply anchored with the public and private spaces organization within cities.
-In fact, the asymmetries in the distribution of time of individuals are rooted on city spaces, distances to reach them, transport alternatives and different perceived fears, emotions and constraints, along with the roles typically associated with men and women, together with the gendered division of labour [@villagran_urban_2020].
-Furthermore, these authors emphasize that it is a global reality that women invest a greater amount of time to care and to unpaid work and that it is socioeconomically stratified.
--->
-<!-- Many studies worldwide have documented how gender influences mobility patterns.
-Men commute longer distances and present direct travel patterns, whereas women present chained trips to accomplish on their way many different activities (for example, dropping children off at schools or daycare, run errands or accompany family members to medical services).
-For these reasons it has been documented that women tend to have shorter but more complicated transport patterns.
-They also tend to rely more on public transport whereas men use more private transport [@cresswell_gendered_2008; @levy2013travel; @mejia2021mobility; @the_world_bank_handbook_2020].
-Sánchez de Madariaga uses the concept “mobility of care” to recognize, measure, make visible, value and account for all the travel associated to caring and unpaid tasks for the household [@sanchezdemadariagaMobilityCareIntroducing2013].
--->
-
-## Feminist 15-minute neighbourhoods
+`r word_comment("Good 15 minute city example, but what about feminist 15 minute cities?
+It's novel but movements exists such as https://x.com/brenttoderian/status/1800371515877384527?s=48&t=QI0WSvV_FDb82ZGeLdJ9lQ", " ")`
 
@@
 In Paris, the adoption of the 15-Minute city put into the spotlight this approach.
-In Paris, they targeted six social functions that had to be easily accessible at any part of the city: housing, work, health care, groceries, education and leisure [@ville_de_paris_paris_2022].
+They targeted six social functions that had to be easily accessible at any part of the city: housing, work, health care, groceries, education and leisure [@ville_de_paris_paris_2022].
 In 2021, Ottawa also adopted a 15-Minutes approach in the new City of Ottawa Official Plan [@ville_dottawa_quartier_2021].
-
 More recently, @moreno_introducing_2021 considers four dimensions on how the 15-minute city framework: density (in terms of people per sq.km); diversity (including mixed-land use and diversity of people; the temporal and spatial proximity to essential services (residential areas, work, commercial areas, education centres, health facilities and other basic institutions) and digitalisation.
-
-It can be observed that the feminist perspective of the cities of proximity is still underestimated.
+In this respect, the feminist perspective of the cities of proximity is still underestimated.
 Only few examples exist that specifically target this issue [@gil_sola_choose_2022; @macintyre_her_2022].
-
-Even if we are on the right direction to facilitate caring activities, develop human-centred urban areas, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities and gender.
-Similarly, it is necessary to provide alternatives to measure progress on this important issue.
-On the other hand, an intersectional approach to understand how different personal identities interact in our daily day activities in the cities is necessary.
-There is still room for improvement.
+Given the impact of caring activities on everyday life, the chrono-urbanism should specifically target the analysis of caring activities.
 
@@
 Given that almost one third of daily trips are for care purposes [@sanchezdemadariagaMobilityCareIntroducing2013], it is important to plan for mobility of care.
-There are also equity considerations to consider: as discussed, women, and especially low-income women, complete this travel more than men.
+There are also equity considerations to consider: as discussed, women, and especially low-income women, are particularly in charge of these duties and therefore, complete this travel more than men.
 Planning for mobility of care may also warrant different considerations because mobility of care trip characteristics differ from other trips, notably those of the commute to work.
@@
 74.7%) [@ravensbergen2023exploratory].
-Further, the rate of car use is higher amongst trips that are part of a chain than those that are not (85.28% v.
+Further, the rate of car use is higher amongst trips that are part of a chain than those that are not (85.28% vs.
 78.09%) [@ravensbergen2023exploratory].
@@
 
-RESEARCH OBJECTIVES:
+The objectives of the paper are as follows:
 
-1)	15 minute cities stem from conventional methods that ignore mobility of care perspectives. There is an applied way of doing this.
-2)	Use of the innovative destination dataset
-3)	Mid-size typical Canadian city case study. How frequently do we find these caring neighbourhoods (potential - accessibility)? Clustering methodology to determine this.
-4)	Who enjoys them and who does not? Clustering methodology to determine these profiles.
-  - i.e. with LICO > X + Prevalence of children >Y + walking accessibility score > U + diversity score > W = “caring neighbourhood”
+1)	City planning is moving to a people-centered approach focusing on accessibility to basic needs.
+However, chrono-urbanism stems from conventional methods that overlooks mobility of care perspectives.
+This paper presents a 15-minute cities methodological approach to assess mobility of care.
+2)	We develop this approach using a mid-size typical Canadian city and a novel care destination dataset as case study
+3)	Furthermore, the potential accessibility to and diversity of care destinations in the case study is analysed.
+We identify to what degree are neighbourhoods caring and/or complete.
+To do so, we make use of self organizing maps (SOM), an unsupervised machine learning data-dimensionality reduction technique to identify caring and complete neighbourhoods.
+4)	We then apply the same SOM methodology to determine profiles of population groups to understand who enjoys these neighbourhoods and who does not.
 
@@
 
-- Next: we describe Hamilton's land-use context; communities which are urban, rural, suburban.
-Also describe the residential parcels summary statistics, ~140,000 points, aggregated at the level of Canadian Census DA.
-See parcel, population and population per parcel plots in @fig-Fig1:
+This study focuses on Hamilton, Ontario, a mid-sized city on the shore of Lake Ontario.
+Hamilton has a heterogenous land-use, with a populated and dense urban core, surrounded by suburban development, which is itself surrounded by rural communities made possible through a greenbelt.
+In this work, we analyse the residential parcel centroids, approximately ~140,000 locations.
+We aggregated the points at the level of Canadian Census Dissemination Area (DA) along with the population and population per parcel plots in @fig-Fig1.
+
 ```{r plot-figure1, out.width=600, out.height=600}
@@
 ```
-Some of the socio-economic and demographic variables that characterise accessibility and the diversity of accessibility to care destinations are pictured in @fig-Fig2:
+
+Hamilton also exhibits spatial disparities in social and economic indicators @fig-Fig2.
+The densely populated inner city is characterised by lower average incomes, and a higher prevalence of households living under the LICO.
+The suburban parts of the city tend to have a greater proportion of children and a lower proportion of one parent households.
 
@@
 Here we will describe the estimated 15-minute walk travel times using {r5r}, from all residential parcel points to all care destinations assuming default walking speed (3.6km/hr) and OSM road network for Hamilton.
-Provide summary statistics.
+**Provide summary statistics.**
 
@@
 The type was then identified e.g., grocers specialty foods, grocers retail, grocer health food, grocer wholesale, grocer curbside, grocer delicatessen wholesale, grocer convenience.
-Data was crossreferenced to ensure all included locations were operational and legitimate grocery stores.",
+Data was cross-referenced to ensure all included locations were operational and legitimate grocery stores.",
                            "Health-centric",
@@
                            "Hospitals, pharmacies, clinics, and dentist offices: 421 destinations are identified.
-Hospitals and pharmacies were retrieved while clinics and dentistry clincs were manually scraped from a healthcare services database and checked via Google Maps to remove non-operational locations and confirm dentistry-orientation.
+Hospitals and pharmacies were retrieved while clinics and dentistry clinics were manually scraped from a healthcare services database and checked via Google Maps to remove non-operational locations and confirm dentistry-orientation.
 ",
@@
 ## Methods
-Cumulative opportunity method; 15-minute normative catchment for walk mode is used to represent the 15-minute neighbourhood concept.
+A 15-minute normative walk catchment is used to represent the 15-minute neighbourhood concept.
 This has been done in previous literature such as @vale2023accessibility, @graellsCityCitiesMeasuring2021 and @gaglioneUrbanAccessibility15minute2022.
-
-In this work, the cumulative opportunity score is calculated for each parcel $i$ and aggregated by care category $c$ and takes following mathetical form:
+In this work, the cumulative opportunity score is calculated for each parcel $i$ and aggregated by care category $c$ and takes the following mathematical form:
 $$
-S^c_i=\sum_{j}W_j^{c}\cdot f(c_{ij})
+S^c_i=\sum_{j}O_j^{c}\cdot f(c_{ij})
 $$
@@
 *Entropy* measure has been widely used to characterise the diversity of land-use mix in early work by @frankLinkingObjectivelyMeasured2005 based on the species diversity Shannon-Wiener index that expresses relative evenness throughout a sequence (this measure among other species diversity indices of the time are reviewed in @whittaker1972evolution).
-In the realm of urban studies and planning literature, it has since been widely used to characterise the diversity in land-use mix [@ewingTravelBuiltEnvironment2010a], especially in the context of active modes [@luUrbanDensityDiversity2017; @mavoaIdentifyingAppropriateLanduse2018] and suburban sprawl [@randallGISBasedLand2015].
+It has also been used in social sciences to comprehend the occurrence of social events in a structured manner specially with regard to life trajectories [@ritschardSequenceAnalysisRelated2018].
+
+In the realm of urban studies and planning literature the malleable concept of entropy has been used to characterise the diversity in land-use mix [@ewingTravelBuiltEnvironment2010a], especially in the context of active modes [@luUrbanDensityDiversity2017; @mavoaIdentifyingAppropriateLanduse2018] and suburban sprawl [@randallGISBasedLand2015].
+As well, it has been used to understand mobility behaviour [@mcbride2020exploration; @montero2023applying; @montero2023role].
 Work that has used diversity indices alongside accessibility analysis are scarce, though examples of its use as parameters within accessibility scores in the case of employment opportunities [@chengMeasuringUrbanJob2013; @daiIncorporatingJobDiversity2018] and more recently to describe the diversity in transit facilities [@yinIncorporatingFacilityDiversity2024] are present.
@@
 #| label: fig-Fig5
-#| fig-cap: "The diversity measure based on the proportion of care category spatial accessibility (figure 4).
+#| fig-cap: "The diversity measure (div) based on the proportion of care category spatial accessibility (figure 4).
 Scale is in tertiles.
@@
 
-```{r plot-figure7, out.width=600, out.height=600}
+```{r plot-figure7, out.width=600, out.height=300}
 #| label: fig-Fig7
@@
 context, households with children may locate to catchments with 'better' schools, often more suburban with lower potential access scores overall.
+Furthermore, statistics show that in the last years, the number of single parents families in Canada has increased.
+In 2022, there were about 1.84 million single parents compared to about 1.56 million in 2010 and 1,3 million in 2001 (https://www.statista.com/statistics/443342/single-parent-families-in-canada/ & https://www.justice.gc.ca/eng/rp-pr/jr/jr12/p5a.html#ftn4).
+Furthermore, Ontario is the Canadian province where most of the lone parent families live (749,769 families, https://www.statista.com/statistics/446667/lone-parent-families-in-canada-by-province/).
+Most lone-families are headed by female-lone parent.
+(AS TO ADD IN HAMILTON SPECIFIC NUMBERS)
 
@@
 Our findings illustrate differences in patterns of 15-minute walk accessibility to care destinations most prominently profiled by income, low-income prevalence, children and single-parent status.
+
+#Conclusions and Discussion
+
+There seems to be a spatial mismatch where dwelling availability and affordability play an important role?
+(NOTE: LMD has have questions about the results) Are there social housing programs for lower income households?
+
+Even if we are on the right direction to develop human-centred urban areas by facilitating caring activities, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities, given the multiple impacts it has on everyday mobility, the unequal distribution of tasks among the population and its related socioeconomic consequences.
+Therefore, it is necessary to provide alternatives to measure progress on this important issue.
+City planning should respond to these issues to promote equity.
+On the other hand, an intersectional approach to understand how different personal identities interact in our daily day activities in the cities is necessary.
 

### 2024-06-21 data/data-raw/SOM-share/Session 11 - Self Organizaing Maps IV.Rmd
Started my edits, pushing back to Anastasia to incorporate the discussion of today's meeting with Lucia and Lea.

@@
 library(tidyverse) # Easily Install and Load the 'Tidyverse'
-library(rgdal) # Bindings for the 'Geospatial' Data Abstraction Library
+#library(rgdal) # Bindings for the 'Geospatial' Data Abstraction Library
 library(spdep) # Spatial Dependence: Weighting Schemes, Statistics
-library(rgeos) # Interface to Geometry Engine - Open Source ('GEOS')
+#library(rgeos) # Interface to Geometry Engine - Open Source ('GEOS')
 library(dplyr) # A Grammar of Data Manipulation

### 2024-06-21 paper/Manuscript.qmd
Started my edits, pushing back to Anastasia to incorporate the discussion of today's meeting with Lucia and Lea.

@@
   #     cite-style: authoryear
-keywords: ["Accessibility", "Mobility of Care", "Gender", "15-Minute City", "Care Complete Neighbourhoods"]
+keywords: ["Accessibility", "Mobility of Care", "Gender", "15-minute neighbourhood", "Care Complete Neighbourhoods"]
 bibliography: bibliography.bib
@@
 
-The 15-Minute City is a normative conceptualisation gaining ground in urban planning: it frames neighbourhoods as responsive to human needs and environmental sensibilities, where most daily necessities can be reached within a 15-minute walk or bike ride [@Allam2022].
-As a related tool, accessibility measures (the ease of reaching opportunities) are increasingly important amongst transport planners aiming to foster just and sustainable cities [@vale2023accessibility].
-Both the 15-Minute City and accessibility measures are flexible enough to consider all destination types holistically however, gendered examinations have been lacking in the literature.
-For instance, accessibility analyses have historically focused on employment-centric and discretionary travel, types of travel more frequent for working-age and higher-income men.
+The 15-minute neighbourhood is a concept in planning that promotes urban spaces that are responsive to human needs and environmental sensibilities.
+These neighbuorhoods aim to accomplish this by ensuring that most daily necessities are not further away than a 15 minutes active trip [@Allam2022].
+In practice, 15-minute neighborhoods are a form of normative accessibility, [@paez_measuring_2012], and therefore related to accessibility analysis, an increasingly important tool that planners use to foster just and sustainable cities [@vale2023accessibility; @handyAccessibilityIdeaWhose2020].
+Both the 15-minute neighborhood and accessibility measures are flexible enough to consider all destination types holistically, however, gendered examinations have been lacking in the literature.
+For instance, accessibility analyses have historically been employment-centric even when adopting a gendered perspective [@paez_jobs_2024].
 
-To counter this masculinist bias, this study investigates a way to gender-mainstream the 15-Minute City through a care lens [e.g., @lawWomenTransportNew1999; @cresswell_gendered_2008; @Oestergaard1992; @levy1991towards; @levy2013travel] supported by the Mobility of Care framework (coined by @sanchezdemadariagaMobilityCareIntroducing2013).
-Mobility of Care emphasizes the importance of travel to unpaid work (care trips) in contrast to the better-studied travel to employment and leisure.
-While all three trip types (work, care, and leisure) are essential, care trips are often relatively shorter-distance, proximate to residential/work/school, and comprise approximately 30% of adults’ daily trips [@sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory; @mejia2021mobility]; fitting well within the 15-Minute City conceptualisation.
+To address this gender bias, this study investigates the 15-minute neighbourhood through the lens of Mobility of Care [@@sanchezdemadariagaMobilityCareIntroducing2013].
+This framework emphasizes the importance of travel related to unpaid work (care trips) in contrast to the better-studied travel to employment and leisure [@lawWomenTransportNew1999; @cresswell_gendered_2008; @Oestergaard1992; @levy1991towards; @levy2013travel].
+While these trip types (work, care, and leisure) are essential, care trips are often relatively shorter-distance, proximate to residential/work/school, and comprise approximately 30% of adults’ daily trips [@sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory; @mejia2021mobility], thus fitting neatly within the concept of 15-minute neighbourhoods.
 
-Our study provides an empirical example that maps the 15-Minute City onto the Mobility of Care framework.
-Specifically, it identifies which areas in Hamilton, Canada are 'caring 15-minute neighbourhoods'.
-To do so, a database of care destinations is created using secondary data.
+Our study casts the 15-minute neighbourhood onto the Mobility of Care framework.
+Empirically, we identify which areas in Hamilton, Canada, are 'caring 15-minute neighbourhoods'.
+To do so, a database of care destinations was compiled.
 In this database, care destinations include all places associated with sustaining household tasks needed for the reproduction of life namely: shopping (e.g., groceries), errands (e.g., libraries), health (e.g., dentist), and caring for dependents (e.g., schools).
-This database is used to estimate the number and mix of care-destinations that can be reached within a 15-minute walk- and cycling- sheds from census centroids.
-Typologies are generated to illustrate which neighbourhoods can and to what degree facilitate 15-minute access to care.
+This database is used to estimate the number and mix of care-destinations that can be reached within a 15-minute walk-sheds from census centroids.
+We generate typologies to illustrate which neighbourhoods can and to what degree facilitate 15-minute access to care.
 
-Results indicate only a few neighbourhoods outside of the downtown core are 'care-complete', i.e., contain a sufficiently high mix of care destinations from all categories and sub-categories.
-However, some neighbourhoods are almost 'care-complete' and provide 15-minute access to some care categories.
-Our study frames these neighbourhoods on the continuum of 'caring' and in need of further intervention.
-The quantitative investigation conducted provides a high-level picture of what neighbourhoods (and their underlying land-use) are connected to transport infrastructure that can support reaching care-destinations.
-
-Taken together, this study provides a theoretical bridge to connect 15-Minute Cities, accessibility analysis and Mobility of Care framework for the purpose of informing policy choice aimed to encourage just and sustainable mobility.
+Our results indicate which 15-minute neighbourhoods are almost 'care-complete', that is, contain a sufficiently high mix of care destinations.
+Very few of these are found outside of the downtown core.
+Our study frames these neighbourhoods on a continuum of 'caring' and helps identify areas for policy intervention.
+Taken together, this study provides a theoretical bridge to connect and Mobility of Care, 15-minute neighbourhoods, and accessibility analysis.
+This should help to inform policy choices aimed at encouraging more just and sustainable mobility.
 
@@
 
-Caring is one of the most fundamental activities in societies, which has allowed humanity to survive and evolve.
-It involves activities targeted to meet the physical, psychological, and emotional needs of others [@ilo_care_2018].
-Yet, it is one of the most devalued, underpaid, and unequal activities worldwide.
-Traditionally, caring activities have remained on women’s shoulders.
-Though women today attain higher educational levels and tend to work out of home more than ever before, the bulk of care remains (at any life stage) heavily on women's shoulders.
-The 'Second Shift' remains mostly women’s work  in all industrial societies [@hayden_grand_2000; @hochschild_second_2012].
-According to the @ilo_care_2018, women and girls perform more than three-quarters of the total amount of unpaid care.
-This gender gap varies geographically [@RN5].
+Caring is one of the most fundamental activities in societies, since it consists of numerous activities that allow them to thrive.
+Caring activities are those that strive to meet the physical, psychological, and emotional needs of others [@ilo_care_2018].
+Yet, it is one of the most unequal, undervalued, and even devalued activities worldwide.
+Traditionally, caring activities have been borne on women’s shoulders.
+Though women today attain higher educational levels and tend to work out of home more than ever before, the 'Second Shift' remains mostly women’s work  in all industrial societies [@hayden_grand_2000; @hochschild_second_2012].
+According to @ilo_care_2018, women and girls perform more than three-quarters of the total amount of unpaid care, a gender gap that varies geographically [@RN5].
 In Canada, adult women complete 1.9 hours, on average, more of unpaid care than men every day [@RN4].
 
-This unequal share of caring responsibilities leads to gendered differences in: career development, profession selection, contract type, pay gap, and time poverty as recognized by different international organisations [@EIGE2016; @ilo_care_2018].
-In fact, even early feminists in the United States observed that domestic labor was the most relevant basic cause of women’s inequality [@gordon_womans_1976; @hayden_grand_2000].
+This unequal share of caring responsibilities leads to multifaceted gendered differences: in career development, profession selection, contract type, pay gap, and time poverty, as recognized by various international organisations [@EIGE2016; @ilo_care_2018].
+In fact, early feminists in the United States had already observed that domestic labor was the single most relevant cause of women’s inequality in society [@gordon_womans_1976; @hayden_grand_2000].
 
-Care also has significant planning considerations.
+Care also has important implications for spatial and transportation planning.
 Indeed, researchers from various fields have highlighted that western modern industrial societies have imposed a physical separation of the private sphere from the public one, splitting production and reproduction, public and private spaces or simply work and home, concentrating on the economic outcome of being mobile [@cresswell_gendered_2008; @lawWomenTransportNew1999; @levy2013travel; @little_gender_1994; @tronto_toward_1990].
@@
 In more recent years, planners have begun to resist this spatial separation of the public and private sphere to promote for sustainable cities.
-Proximity planning has become more normalized, at popular example of which is the 15-minute city, in which residents can reach all necessary destinations within a 15-minute walk.
-The 15-minute city concept, however, has yet to directly focus on care.
+Proximity planning has become more normalized, at popular example of which is the 15-minute neighbourhood, in which residents can reach all necessary destinations within a 15-minute walk.
+The 15-minute neighbourhood concept, however, has yet to directly focus on care.
 
@@
 
-In Paris, the adoption of the 15-Minute city put into the spotlight this approach.
+In Paris, the adoption of the 15-minute neighbourhood put into the spotlight this approach.
 They targeted six social functions that had to be easily accessible at any part of the city: housing, work, health care, groceries, education and leisure [@ville_de_paris_paris_2022].
 In 2021, Ottawa also adopted a 15-Minutes approach in the new City of Ottawa Official Plan [@ville_dottawa_quartier_2021].
-More recently, @moreno_introducing_2021 considers four dimensions on how the 15-minute city framework: density (in terms of people per sq.km); diversity (including mixed-land use and diversity of people; the temporal and spatial proximity to essential services (residential areas, work, commercial areas, education centres, health facilities and other basic institutions) and digitalisation.
+More recently, @moreno_introducing_2021 considers four dimensions on how the 15-minute neighbourhood framework: density (in terms of people per sq.km); diversity (including mixed-land use and diversity of people; the temporal and spatial proximity to essential services (residential areas, work, commercial areas, education centres, health facilities and other basic institutions) and digitalisation.
 In this respect, the feminist perspective of the cities of proximity is still underestimated.

### 2024-07-06 data/01-census-and-traveltimes.qmd
current draft version - added some comments from discussion 2 weeks ago with the team.

@@
 
-## Transit travel times
+##Transit travel times
 
 Calculate **Transit** travel times from centroids of parcels (from_ids) to destination points (to_ids). 30 min max travel time.
+
+```{r}
+load("parcel-ttm/ttm_care_TRANSIT_parcel1_1.rda")
+load("parcel-ttm/ttm_care_TRANSIT_parcel1.rda")
+
+ttm_care_TRANSIT_parcel1|> summary()
+ttm_care_TRANSIT_parcel1_1 |> summary()
+```
 
@@
                           destinations = dest,
-                          mode = c("TRANSIT"),
+                          mode = c("TRANSIT","WALK"),
+                          max_walk_time = 15,
                           departure_datetime = as.POSIXct(strptime("08-06-2023 08:00:00", # 12 GMT is 8am in Toronto (EDT)
@@
 
-save("ttm_care_TRANSIT_parcel1", file = "parcel-ttm/ttm_care_TRANSIT_parcel1.rda")
-rm("ttm_care_TRANSIT_parcel1")
+save("ttm_care_TRANSIT_parcel1", file = "parcel-ttm/ttm_care_TRANSIT_parcel1_1.rda")
+rm("ttm_care_TRANSIT_parcel1_1")
 ```

### 2024-07-06 paper/Manuscript.qmd
current draft version - added some comments from discussion 2 weeks ago with the team.

@@
 These neighbuorhoods aim to accomplish this by ensuring that most daily necessities are not further away than a 15 minutes active trip [@Allam2022].
-In practice, 15-minute neighborhoods are a form of normative accessibility, [@paez_measuring_2012], and therefore related to accessibility analysis, an increasingly important tool that planners use to foster just and sustainable cities [@vale2023accessibility; @handyAccessibilityIdeaWhose2020].
+In practice, 15-minute neighborhoods are a form of normative accessibility [@paez_measuring_2012], and therefore related to accessibility analysis, an increasingly important tool that planners use to foster just and sustainable cities [@vale2023accessibility; @handyAccessibilityIdeaWhose2020].
 Both the 15-minute neighborhood and accessibility measures are flexible enough to consider all destination types holistically, however, gendered examinations have been lacking in the literature.
@@
 
-To address this gender bias, this study investigates the 15-minute neighbourhood through the lens of Mobility of Care [@@sanchezdemadariagaMobilityCareIntroducing2013].
+To address this gender bias, this study investigates the 15-minute neighbourhood through the lens of Mobility of Care [@sanchezdemadariagaMobilityCareIntroducing2013].
 This framework emphasizes the importance of travel related to unpaid work (care trips) in contrast to the better-studied travel to employment and leisure [@lawWomenTransportNew1999; @cresswell_gendered_2008; @Oestergaard1992; @levy1991towards; @levy2013travel].
@@
 Yet, it is one of the most unequal, undervalued, and even devalued activities worldwide.
-Traditionally, caring activities have been borne on women’s shoulders.
-Though women today attain higher educational levels and tend to work out of home more than ever before, the 'Second Shift' remains mostly women’s work  in all industrial societies [@hayden_grand_2000; @hochschild_second_2012].
+Traditionally, caring activities have been borne on women’s shoulders [@hayden_grand_2000; @hochschild_second_2012].
 According to @ilo_care_2018, women and girls perform more than three-quarters of the total amount of unpaid care, a gender gap that varies geographically [@RN5].
-In Canada, adult women complete 1.9 hours, on average, more of unpaid care than men every day [@RN4].
 
@@
 Proximity planning has become more normalized, at popular example of which is the 15-minute neighbourhood, in which residents can reach all necessary destinations within a 15-minute walk.
-The 15-minute neighbourhood concept, however, has yet to directly focus on care.
+The 15-minute neighbourhood concept, however, has yet to foreground care.
 
-While much care labour takes place in the private, and feminised sphere, completing care activities requires one to be mobile.
+While much care labour takes place in the private and feminised sphere, completing care activities requires one to be mobile.
 In other words: caring for oneself and others requires trips to, amongst other places, grocery stores, day cares, elder care, and health care destinations.
-
-Women have consistently been found to complete more mobility of care trips than men.
-In one study, mobility of care comprised 32% of women’s daily trips compared to 28% of men’s.
-While this gap was significant, it was found to be far greater in lower income households where women complete 20% more care trips than men [@ravensbergen2023exploratory].
-This gendered division in mobility of care is perhaps unsurprising as women tend to complete a disproportionate amount of household-serving labour, such as caring for children and other dependents, preparing meals, cleaning, and running errands, when compared to men [@RN4].
-This gender gap varies geographically [@RN5].
-In Canada, adult women complete 1.9 hours, on average, more of unpaid care than men every day [@RN4].
-This labour is essential, and a large component of daily life.
 
@@
 
-`r word_comment("Good 15 minute city example, but what about feminist 15 minute cities?
-It's novel but movements exists such as https://x.com/brenttoderian/status/1800371515877384527?s=48&t=QI0WSvV_FDb82ZGeLdJ9lQ", " ")`
+`r word_comment("TO ADD: some good feminist 15 minute cities.
+It's novel but movements exists in the Canadian context i.e., https://x.com/brenttoderian/status/1800371515877384527?s=48&t=QI0WSvV_FDb82ZGeLdJ9lQ.
+Also discuss Barcelona and Bogota (care blocks)", " ")`
 
@@
 Planners and policy makers have proposed different alternatives where residents reach “essential activities” within walkable or cycling distances, reducing the need for using motorised transportation modes.
-The 15-minutes cities, 20-minutes or x-minutes cities, have been long under the spotlight.
+The 15-minutes cities or "x"-minutes cities, have been long under the spotlight.
 This approach has also received the name of Chrono-urbanism or the “cities of proximity” [@logan_xminute_2022; @moreno_introducing_2021].
@@
 Probably one of the first cities to adopt the proximity approach was Portland.
-The Portland Plan [@city_of_portland_20minute_2010] was adopted by the City Council on April 25, 2012 aiming at fostering an inclusive city development for the following 30 years.
-The plan is based on the following principles: prosperity, education, health, and equity.
+The Portland Plan [@city_of_portland_20minute_2010] was adopted by the City Council on April 25, 2012 aiming at fostering an inclusive city development based on prosperity, education, health, and equity for the following 30 years.
 The progress report states that equity should be achieved, an Office of Equity and Human Rights was established focusing on equitable service delivery to all residents.
 Nonetheless the progress report only measures progress on topics related to racial equity or people with disabilities [@portland_government_portland_2017].
+`r word_comment("NOTE TO SELF: And so what?
+Is this a research gap?", " ")`
 
@@
 Given that almost one third of daily trips are for care purposes [@sanchezdemadariagaMobilityCareIntroducing2013], it is important to plan for mobility of care.
-There are also equity considerations to consider: as discussed, women, and especially low-income women, are particularly in charge of these duties and therefore, complete this travel more than men.
+There are also equity considerations to consider; women, and especially low-income women, are particularly in charge of these duties and therefore, complete this travel more than men.
 Planning for mobility of care may also warrant different considerations because mobility of care trip characteristics differ from other trips, notably those of the commute to work.
@@
 
-Further, compared to the trip to work, mobility of care trips are more frequently completed by car and by foot, and less frequently done by transit or bicycle [@ravensbergen2023exploratory].
-Evidence for why this modal difference is observed exists.
-For instance, it is possible that walking is more common for care than for work trips because care trips tend to be short [@ravensbergen2023exploratory].
-In other words, if care destinations are closer to the home, it is more likely that these destinations are walkable.
+Further, compared to the trip to work, mobility of care trips are more frequently completed by foot, and less frequently done by transit or bicycle [@ravensbergen2023exploratory].
+Further, evidence suggests that walking is more common for care than for work trips because care trips tend to be short [@ravensbergen2023exploratory]; if care destinations are closer to the home, it is more likely that these destinations are walkable.
 Ample literature examines the benefits and barriers to walking care destinations, such as schools (e.g., [@RN7;@RN8;@RN9]) and grocery stores (e.g., [@RN10;@RN11]).
-Little of this work uses the mobility of care framework.
-Instead, it tends to consider singular care destinations in research focusing on walkability.
+However, little of this work uses the mobility of care framework.
+Instead, they tend to consider singular care destinations in research focusing on walkability.
 
+`r word_comment("TO ADD: Replace the following three paragrapgs with a more elaborated discussion on walking as the primary mode for 15-min cities and how it overlaps with mobility of care. ", ".")`
 In car dependent cities, cycling commute rates tend to be low.
@@
 ## Case study context
+
 In Canada, women have become more educated and have an achieved an increased participation in the labour market [@government_of_canada_gender_2023].
@@
 Here we will describe the estimated 15-minute walk travel times using {r5r}, from all residential parcel points to all care destinations assuming default walking speed (3.6km/hr) and OSM road network for Hamilton.
-**Provide summary statistics.**
+`r word_comment("TO ADD: add summary statistics", " ")`
 
@@
 
-Accessibility per category @fig-Fig4:
+Accessibility per category @fig-Fig4.
+The concentration of opportunities for all category types is concentrated in Hamilton Central - with the highest concentration being in the lower city and pockets of the high concentration on the mountain `r word_comment("(TO ADD: The escarpment line)", " ")`.
+Grocery-centric opportunities are most concentrated, followed by dependent-centric and health-centric.
+Errand-centric is the most sprawled.
+
 ```{r plot-figure4, out.width=600, out.height=600}
@@
 ```
-
-Diversity @fig-Fig5:
+Diversity @fig-Fig5. Concentrated in the downtown core.
 ```{r plot-figure5, out.width=600, out.height=600}
@@
 
-Superclusters based on accessibility per category and diversity (the two above figures). 5 clusters are identified, see their dendrogram (right) and grid map (left) in @fig-Fig6 :
+Superclusters based on accessibility per category and diversity (the two above figures).
+5 clusters are identified, see their dendrogram (right) and grid map (left) in @fig-Fig6 `r word_comment("change colours of SC to match the following plots", ":")`.
 
@@
 To further visualize supercluster membership, we spatially demonstrate which supercluster best describes each parcel in @fig-Fig8 (top plot): the median parcel's supercluster membership per DA is tabulated and the majority membership is represented.
-Visually, the super clusters closely follow the quartile split of the accessibility to care destinations (all categories combined) and tertile split of the diversity scores (visualised in @fig-Fig5).
+Visually, the super clusters closely follow the quartile split of the accessibility to care destinations (all categories combined) and tertile split of the diversity scores (visualised in @fig-Fig5) `r word_comment("AN IDEA DISCUSSED: Overall, in @fig-Fig8 there are many pockets of good/great care accessibility+diversity of opportunities Hamilton.
+These are objectively caring neighbourhoods.
+However, people may subjectively disagree...
+", ".")`.
 
@@
 
-#Conclusions and Discussion
+ `r word_comment("IDEAS DISCUSSED: regarding @fig-Fig9 - 1) We know that highest accessibility & complete is downtown.
+But people there have fewer children + lowest income are downtown.
+On the horizon as rent increases,  the most precarious may be pushed out and replaced by people who can pay higher rents.
+2) The lowest accessibility & most incomplete are suburban -- they are likely higher income.
+Interesting to note: one-parent families are more likely to be time poor -- and leaving in low access/incomplete suburbs puts more at risk.
+We know a think or 2 about single parents and jobs -- access to jobs was high where single parents were not.
+(Catherine, farber Paez paper).
+3) Pockets of the suburbs have high access to care.
+The suburbs do something -- they have care amenities.
+", " ")`
+  `r word_comment("IDEAS DISCUSSED: regarding @fig-Fig9 - 2) The lowest accessibility & most incomplete are suburban -- they are likely higher income.
+Interesting to note: one-parent families are more likely to be time poor -- and leaving in low access/incomplete suburbs puts more at risk.
+We know a think or 2 about single parents and jobs -- access to jobs was high where single parents were not.
+(Catherine, farber Paez paper).
+", ":")`.
 
-There seems to be a spatial mismatch where dwelling availability and affordability play an important role?
-(NOTE: LMD has have questions about the results) Are there social housing programs for lower income households?
+ `r word_comment("IDEAS DISCUSSED: regarding @fig-Fig9 - 3) Pockets of the suburbs have high access to care. The suburbs do something -- they have care amenities. ", ":")`.
+
+# Conclusions and Discussion
+
+There seems to be a spatial mismatch where dwelling availability and affordability play an important role? Are there social housing programs for lower income households?
 

### 2024-09-03 WGiT/WGiT-pres.Rmd
Draft WGiT Presentation and edits to manuscript

---
title: "Towards caring 15-minute neighbourhoods"
subtitle: "Proximity Planning Congress 2024"
author: "Anastasia Soukhov, Léa Ravensbergen, Lucía Mejía Dorantes, Antonio Páez"
date: "09-10-2024"
output:
  xaringan::moon_reader:
    seal: false
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current% / 10" 
      highlightStyle: github
      highlightLines: true
      ratio: '16:9'
      countIncrementalSlides: false
      beforeInit: "macros.js"
---
```{r xaringan-themer, include=FALSE, warning=FALSE}
#this themer package allows us to change the .css themes through this chunk. Easier than changing the .css directly.
library(xaringanthemer)
style_mono_accent(
  base_color = "#7A003C", #maroon
  link_color = "#5E6A71", #grey
  text_bold_color = "#FDBF57", #gold
  text_slide_number_color = "#FDBF57", #gold
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

```{r load-packages, include=FALSE, cache=FALSE}
options(htmltools.dir.version = FALSE)
library(RefManageR)

library(dplyr)
library(sf)
library(ggplot2)
library(tidyverse)
library(units)
library(patchwork)
library(ggspatial)
library(ggforce)
# library(geomtextpath)
library(knitr)
library(kableExtra)
library(here)
#install.packages("emo")
library(emo)
options(scipen = 999)
```

```{r bib-setup, include=FALSE, cache=FALSE}
BibOptions(check.entries = FALSE, bib.style = "authoryear", style = "markdown",
           dashed = TRUE, max.names = 1)
bib <- ReadBib(paste0(here(), "/bibliography/bibliography.bib"))
```

```{r sp_avail-function,include=FALSE}
#defining the spatial availability function, this prints all the intermediates
sp_avail_detailed <- function(x, o_id, d_id, pop, opp, r, f, alpha = 1){

  o_id <- rlang::enquo(o_id)
  d_id <- rlang::enquo(d_id)
  pop <- rlang::enquo(pop)
  opp <- rlang::enquo(opp)
  r <- rlang::enquo(r)
  f <- rlang::enquo(f)

  sum_pop <- x %>%
    dplyr::distinct(!!o_id,
                    .keep_all = TRUE) %>%
    dplyr::mutate(sum_pop = !!r*(!!pop)^alpha) %>%
    dplyr::pull(sum_pop) %>%
    sum()

  f_p <- dplyr::pull(x, !!r) * dplyr::pull(x, !!pop)^alpha / sum_pop

  sum_impedance <- x %>%
    dplyr::group_by(!!d_id) %>%
    dplyr::summarize(sum_impedance = sum(!!f))

  x <- x %>%
    dplyr::left_join(sum_impedance,
                     by = rlang::as_name(d_id))

  f_c <- dplyr::pull(x, !!f) / x$sum_impedance

  x$f_c <- f_c
  x$f_p <- f_p

  sum_pa <- x %>%
    dplyr::group_by(!!d_id) %>%
    dplyr::summarize(sum_pa= sum(f_p * f_c))

  x <- x %>%
    dplyr::left_join(sum_pa,
                     by = rlang::as_name(d_id))
  f_t <- (f_p * f_c) / dplyr::pull(x, sum_pa)

  x %>%
    dplyr::mutate(V_ij = !!opp * !!f_t)
}
```

class: center

# Towards caring 15-minute neighbourhoods

```{r, echo=FALSE, out.width='30%', fig.align='center'}
knitr::include_graphics(here("figures/mobility-of-care.png"))
```

#### Anastasia Soukhov (McMaster University)
#### Léa Ravensbergen (McMaster University), Lucía Mejía Dorantes (Karlsruhe, Germany), Antonio Páez (McMaster University)

#### _7th TRB International Conference on Women and Gender in Transportation: 10 September, 2024_

---

### The 15-Minute City

- a normative conceptualisation gaining ground in urban planning: reaching daily necessities/activities within 15\* minutes by walk, bike\*, and transit\*. `r Citep(bib, c("Allam2022","moreno_introducing_2021"))`

- Practical challenges:
  - Non-specific: 15-minutes? By which travel modes?
  - **For whom? To what destination?**

```{r, echo=FALSE, out.width='50%', fig.align='center'}
knitr::include_graphics(here("figures/15-m-city.jpg"))
```
_Shutterstock; Getty | Aerial views of Glasgow, Paris and Melbourne. Retrieved from Financial Times article "Welcome to the 15 Minute City" by N. Whittle in 2020_

???

Let's apply a gendered lens. The research objective: answering **For whom? To what destination?** with Mobility of Care

---

### The 15-Minute City and Mobility of Care

- Mobility of Care is _also_ a normative conceptualisation; coined in `r Citep(bib, c("sanchezdemadariagaMobilityCareIntroducing2013"))`, emphasizes all travel needed to sustain yourself and the household, i.e., all non-paid work related travel.

- Mobility of Care trips are often: under 15 minutes, proximate to home, and more frequently conducted by non-car modes (than work trips). Pair well with the 15-Minute City conceptualisation

- **Guiding research aim**: a method to identify what neighbourhoods are, are somewhat, and are not 15-minute _caring_ neighbourhoods

```{r, echo=FALSE, out.width='50%', fig.align='center'}
knitr::include_graphics(here("figures/graphs-trips-by-purpose.webp"))
```
_image from `r Citep(bib, c("sanchezdemadariagaMeasuringMobilitiesofCare2019"))`_

???

Mobility of Care... concerns everyone But as it stands, concerns women most.

- This type of travel has been traditionally understudied: 1) travel from home-to-work is often the focus of transportation planning, 2) travel surveys to inform this planning often obscure and other mobility of care travel.

---
### Case study: Hamilton, Canada

Research question: what neighbourhoods are, are somewhat, and are not 15-minute _caring_ neighbourhoods?

```{r, echo=FALSE, out.width='50%', fig.align='center'}
knitr::include_graphics(here("figures/pop_descriptive_plot_onlypop.png"))
```

---
### Case study: Hamilton, Canada

Research question: what neighbourhoods are, are somewhat, and are not 15-minute _caring_ neighbourhoods?

```{r, echo=FALSE, out.width='50%', fig.align='center'}
knitr::include_graphics(here("figures/care_categories_plot.png"))
```

---

### Quantifying 15-minute caring neighbourhoods

.pull-left[

#### A) Spatial accessibility of care destinations

How _many_ care destinations can be reached within a 15 minute walk? The _potential of opportunities for interaction_ `r Citep(bib, "hansenHowAccessibilityShapes1959")`.

  - Dependent-centric care destinations
  - Errand-centric care destinations
  - Grocery-centric care destinations
  - Health-centric care destinations

$S_i = \sum_j O_j f(c_{ij})$

]

.pull-right[

#### B) The diversity of categories

What is the _mix_ of care destination types within a 15 minute walk? The entropy measure:

$D_{i} = \frac{-\sum_{c} S_i^{c} / \sum_{c}S^{c}_{i} \times ln(S_i^{c} / \sum_{c}S^{c}_{i} )}{ln(n_c)}$

]

- $i$ is a set of parcel point origin locations; $j$ is a set of destination locations.
- $O^_j$ is a number of opportunities at $j$.
- $c_{ij}$ is the travel cost between $i$ and $j$.
- $f(\cdot)$ is an impedance function of $c_{ij}$; within the cumulative opportunity approach, it is a binary function that takes the value of 1 if $c_{ij}$ is less than a selected value

???

---

### Quantifying 15-minute caring neighbourhoods

.pull-left[

#### A) Spatial accessibility of care destinations

```{r, echo=FALSE, out.width='90%',fig.align='center'}
knitr::include_graphics(glue::glue(here::here(), "/figures/percats_copp_plot.png"))
```

]

.pull-right[

#### B) The diversity of categories

```{r, echo=FALSE, out.width='90%',fig.align='center'}
knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_entropy_copp_plot.png"))
```

]

???

### Typology and Profiles of 15-minute caring neighbourhoods: clustering via Self Organizing Maps

- To combine the Spatial Accessibility measures and the Diversity measure, and to profile the neighbourhoods, SOM is implemented.

- SOM is a pattern recognition method which reduces the dimensionality of the data into clusters without relying on any assumptions about the data structure.

The SOM procedure:
1 ) five superclusters of the four $S_i^c$ scores and the diversity score $D_i$ are identified. The number of superclusters strike a balance between variance explained and supercluster interpretation in what we define as ranging from "not caring" to "very caring" and "not complete" to "complete".

2) The superclusters are profiled using a variety of socio-economic and demographic indicators that the mobility of care literature finds significant.

---

### Typologies of 15-minute caring neighbourhoods

```{r, echo=FALSE, out.width='55%',fig.align='center'}
knitr::include_graphics(glue::glue(here::here(), "/figures/SClabel_copp_plot.png"))
```

???

very caring (mean 4 to 18.6; total p50 46) - very complete (div=0.9; p50 0.90) (Cluster 2)
caring (mean 2.1 to 9; total p50 18) - complete (div=0.81; p50 0.82) (Cluster 1)
somewhat caring (0.2 to 6; total p50 13) - complete (div=0.71; p50 0.74) (Cluster 4)
somewhat caring (0.2 to 6; total p50 7) - not complete (div=0.36; p50 0.40) (Cluster 3)
not caring (0 to 4.7; total p50 3) - not complete (div=0.02; p50 0) (Cluster 5)

---

### Profiles of 15-minute caring neighbourhoods typologies

```{r, echo=FALSE, out.width='60%',fig.align='center'}
knitr::include_graphics(glue::glue(here::here(), "/figures/access_sc_profiles_renamed.png"))
```

- Lower income households tend to reside in $\color{Green}{\text{very completely caring}}$ or $\color{GreenYellow}{\text{completely caring}}$ neighbourhoods; 
- Those with a `r emo::ji("up_arrow")` proportion of children tend to reside in $\color{orange}{\text{somewhat caring - not complete}}$ neighbourhoods;
- Higher income households in DAs with a `r emo::ji("down_arrow")` of single-parent households tend to reside in $\color{red}{\text{not caring - not complete}}$ neighbourhoods.

???

All variables that are tested:  Income_bottom + Age_0to14 + One_parent_families + Vis_minority + Not_in_labforce+ Income_gini+ walk_towork+ Owner_hld_CHN+ Hld_ownership+ Tenant_subsi_hld+ Tenant_hld_CHH+ LICO

Of a list of socio-economic variables, ones that characterise the 5 typologies the most are: 1) % in bottom half of the household income distribution, 2) children population (0 to 14) %, 3) single-parent household %, and 4) Low Income Cutoff (LICO) %

Who has resides in $\color{Green}{\text{very completely caring}$ or $\color{GreenYellow}{\text{completely caring}$ neighbourhoods?  (maybe people who cannot afford cars???)
- DAs that have lower median household incomes and a smaller children population (lower than 24% of the population are children 14 and under).
- and DAs which have especially high LICO are often in $\color{Green}{\text{very completely caring}$ neighbourhoods.

Who more often resides in $\color{BurntOrange}{\text{somewhat caring - not complete}$ neighbourhoods?  
- Of bottom income DAs, DAs with a high proportion of children population (higher than 24% of the population are children 14 and under).

Who more often resides in $\color{BrickRed}{\text{not caring - not complete}$ neighbourhoods?  
- DAs of higher income DAs and with lower proportion of single-parent households (13% or lower)

---

### Conclusions: classifying and profiling caring 15-minute neighbourhoods

.pull-left[

#### What we did:

- 15 minute-city is bridged with "Mobility of Care": 15-minute caring neighbourhoods.
- A data-driven pattern recognition method (Self-organizing maps) to create **15-minute caring neighbourhood** typologies using **accessibility** and **diversity** indicators is presented.
- Using Self-Organizing Maps again; each parcel tagged with a neighbourhood typology was profiled based on all relevant census variables.
]

.pull-right[

#### What we found and future considerations:

- **Gentrification concerns**: populations residing in completely caring neighbourhoods are often lower-income and vice-versa.

- **Concerning caring for children and children's self-care**: Lower-income DAs with a `r emo::ji("up_arrow")` proportion of children, and DAs with `r emo::ji("up_arrow")` proportion single-parent households have medium complete care access, why?

]

???

### 2024-09-03 data/02-access-and-prelim-plots.qmd
Draft WGiT Presentation and edits to manuscript

@@
   mutate(Care_Category_Specific = case_when(
-    Care_Category_Specific  == "Daycare_EarlyON" ~ "Daycare or EarlyON",
-    Care_Category_Specific == "Community_Centre" ~ "Community or Recreation Centre",
-    Care_Category_Specific == "Daycare_Centre" ~ "Daycare or EarlyON",
-    Care_Category_Specific == "Recreation_Centre" ~ "Community or Recreation Centre",
+    Care_Category_Specific  == "Daycare_EarlyON" ~ "Daycare | EarlyON",
+    Care_Category_Specific == "Community_Centre" ~ "Community | Rec. Centre",
+    Care_Category_Specific == "Daycare_Centre" ~ "Daycare | EarlyON",
+    Care_Category_Specific == "Recreation_Centre" ~ "Community | Rec. Centre",
     Care_Category_Specific == "School" ~ "School",
-    Care_Category_Specific == "Long-Term_Care_Home" ~ "Long-Term Care or Retirement Home",
-    Care_Category_Specific == "Retirement_Home" ~ "Long-Term Care or Retirement Home",
+    Care_Category_Specific == "Long-Term_Care_Home" ~ "LTC | Ret. Home",
+    Care_Category_Specific == "Retirement_Home" ~ "LTC | Ret. Home",
     Care_Category_Specific == "Senior_Centre" ~ "Senior Centre",
@@
 rm(test_all)
-access_15$V_isc |> sum(na.rm = TRUE) #sums to 2206 opportunities -- out of 2225 total opportunities. some opportunities canot be reached wtihin 15 min walk! 
+access_15$V_isc |> sum(na.rm = TRUE) #sums to 2206 opportunities -- out of 2225 total opportunities. some opportunities cannot be reached within 15 min walk! 
 ```
@@
 
-           Care_Category_Specific = rep(c("Park", "Daycare or EarlyON", "Convenience Store", "School",
-                                          "Long-Term Care or Retirement Home", "Senior Centre",
+           Care_Category_Specific = rep(c("Park", "Daycare | EarlyON", "Convenience Store", "School",
+                                          "LTC | Ret. Home", "Senior Centre",
                                           "Bank or ATM", "Library", 
@@
                                           "Pharmacy", "Post Office",
-                                          "Community or Recreation Centre", "Hospital or Clinic"), length(R_PARCELS_CENTS_2020$ID)),
-           copp_isc = rep(0, length(R_PARCELS_CENTS_2020$ID)*length(c("Park", "Daycare or EarlyON", "Convenience Store", "School",
-                                          "Long-Term Care or Retirement Home", "Senior Centre",
+                                          "Community | Rec. Centre", "Hospital or Clinic"), length(R_PARCELS_CENTS_2020$ID)),
+           copp_isc = rep(0, length(R_PARCELS_CENTS_2020$ID)*length(c("Park", "Daycare | EarlyON", "Convenience Store", "School",
+                                          "LTC | Ret. Home", "Senior Centre",
                                           "Bank or ATM", "Library", 
@@
                                           "Pharmacy", "Post Office",
-                                          "Community or Recreation Centre", "Hospital or Clinic"))),
-           copp_isc_w = rep(0, length(R_PARCELS_CENTS_2020$ID)*length(c("Park", "Daycare or EarlyON", "Convenience Store", "School",
-                                          "Long-Term Care or Retirement Home", "Senior Centre",
+                                          "Community | Rec. Centre", "Hospital or Clinic"))),
+           copp_isc_w = rep(0, length(R_PARCELS_CENTS_2020$ID)*length(c("Park", "Daycare | EarlyON", "Convenience Store", "School",
+                                          "LTC | Ret. Home", "Senior Centre",
                                           "Bank or ATM", "Library", 
@@
                                           "Pharmacy", "Post Office",
-                                          "Community or Recreation Centre", "Hospital or Clinic"))),
+                                          "Community | Rec. Centre", "Hospital or Clinic"))),
 
-           V_isc = rep(0, length(R_PARCELS_CENTS_2020$ID)*length(c("Park", "Daycare or EarlyON", "Convenience Store", "School",
-                                          "Long-Term Care or Retirement Home", "Senior Centre",
+           V_isc = rep(0, length(R_PARCELS_CENTS_2020$ID)*length(c("Park", "Daycare | EarlyON", "Convenience Store", "School",
+                                          "LTC | Ret. Home", "Senior Centre",
                                           "Bank or ATM", "Library", 
@@
                                           "Pharmacy", "Post Office",
-                                          "Community or Recreation Centre", "Hospital or Clinic")))) |>
+                                          "Community | Rec. Centre", "Hospital or Clinic")))) |>
   mutate(from_id = as.character(from_id))
@@
 #RECALL, categories are either 
-# - "Dependent-centric" and include "Community or Recreation Centre", "Daycare or EarlyON", "Long-Term Care or Retirement Home", "Park", "School", "Senior Centre"
+# - "Dependent-centric" and include "Community | Rec. Centre", "Daycare | EarlyON", "LTC | Ret. Home", "Park", "School", "Senior Centre"
 # - "Errand-centric" and include: "Bank or ATM", "Library", "Post Office"
@@
   tm_compass(position = c("left", "top"), size=0.9) +
-  tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5, legend.bg.color = "white", legend.bg.alpha = 0.4,
+  tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 1, legend.bg.color = "white", legend.bg.alpha = 0.4,
             panel.labels = c("1,265 Dependent-centric destinations", 
@@
                              "421 Health-centric destinations"),
-            panel.label.size = 0.8)
+            panel.label.size = 1.1)
 

### 2024-09-03 data/03-clustering-and-profiling.qmd
Draft WGiT Presentation and edits to manuscript

@@
 
+pop_DA_plot_2 <-
+ tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
+  tm_shape(access_15_DA) +
+  tm_polygons("Population",
+              border.alpha = 0.2,
+              palette = "Greens", style = "quantile", n=4,
+              title = "Population",
+              legend.is.portrait = TRUE) +
+  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
+  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
+  tm_compass(position = c("left", "top"))+
+   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.bg.color = "white", legend.bg.alpha = 0.4)
+
 pop_descriptive_plots_1 <- tmap_arrange(parcels_DA_plot, pop_DA_plot, poprate_DA_plot, nrow=3)
@@
           file= file.path(here("figures/pop_descriptive_plots_1.png")) , asp = 1/0.6,
+          dpi=300)
+
+tmap_save(pop_DA_plot_2,
+          file= file.path(here("figures/pop_descriptive_plot_onlypop.png")),
           dpi=300)
@@
               palette = c("#44ce1b","#bbdb44", "#f7e379", "#f2a134","#e51f1f"),
-              labels = (c("Very caring - very complete (SC2)","Caring - Complete (SC1)",
-              "Somewhat caring - complete (SC4)","Somewhat caring - somewhat complete (SC5)", 
-              "Somewhat caring - somewhat complete (SC3)")),
-              title = "Prevelant supercluster") +
+              labels = (c("Very caring - very complete",
+                          "Caring - Complete",
+              "Somewhat caring - complete",
+              "Somewhat caring - somewhat complete", 
+              "Somewhat caring - somewhat complete")),
+              title = "Neighbourhood typology") +
   tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
@@
           file= file.path(here("figures/SCLab_and_total_copp_plot.png")),
+          dpi=300)
+
+tmap_save(SClabel_copp_plot, 
+          file= file.path(here("figures/SClabel_copp_plot.png")),
           dpi=300)
@@
 freq_sc
+
+#with new names and colours
+freq_sc_renamed <- ggplot(access_15_wide |> 
+         mutate(scluster = factor(scluster))) +
+  geom_bar(aes(x = "", 
+               group = scluster, 
+               fill = scluster)) +
+  scale_fill_manual(name = "Neighbourhood typologies",
+                    values = c("#44ce1b","#bbdb44", "#f7e379", "#f2a134","#e51f1f"),
+                    labels = c("Very caring - very complete","Caring - complete","Somewhat caring - not complete","Somewhat caring - complete","Not caring - not complete"))
 ```
@@
 
-#Train a decision tree using the data:
+#Train a decision tree using the data (renamed!):
 classdemo.som <- rpart(data = df,
@@
                                         position = position_fill()),
-                               xlab(""),
-                               ylab(""),
+                               #xlab(""),
+                               #ylab(""),
                                theme_void(),
+                               scale_fill_manual(values = c("#bbdb44","#44ce1b",  "#f2a134", "#f7e379","#e51f1f")),
                                theme(axis.text = element_blank())))
-legd <- cowplot::get_legend(freq_sc)
+
+legd_renamed <- cowplot::get_legend(freq_sc_renamed)
+party_plot
 ```
 
-```{r}
+```{r, eval=FALSE}
 #plot and save
-png(file= file.path(here("figures/access_sc_profiles.png")),width= 2404, height= 1600, units="px",  res=300)
-plot_grid(party_plot, legd, nrow = 1, rel_widths = c(1, 0.2))
+# NOTE: edited this in paint.. be careful saving over
+png(file= file.path(here("figures/access_sc_profiles_renamed.png")),width= 2600, height= 1600, units="px",  res=300)
+plot_grid(party_plot, legd_renamed, nrow = 1, rel_widths = c(1, 0.3))
 dev.off()

### 2024-09-03 paper/Manuscript.qmd
Draft WGiT Presentation and edits to manuscript

@@
 ---
-
-<!--
-
-•	Establishing definitive research questions and the work’s contributions. Us deciding on this together will help me sculpt from what’s there
-  •	Thoughts on methods? Can we improve presentation?
-  •	Thoughts on results? Can we improve presentation? What are insights that are significant to you? What do you want to know more about?
-o	Thoughts on how I weighted care destinations
-o	Thoughts on aggregation by ‘care category’ vs ‘specific type’ of destination
-o	Thoughts on additional dimensions – i.e., the inclusion of results that only consider ‘grocery, errand, and health centric’ destinations (for people who have no dependents).
-For the time being, I left this consideration out and just demo _all_ destinations.
-Also, what about LICO-AT discussion?
-  •	Critical thoughts on how to tie walking and cycling modes together.
-For now, I did not include cycling results, though I have the cycling travel time matrices.
-We should use the cycling results perhaps to demo how the range of care access can be extended for certain neighbourhoods.
-A potential ‘what-if’ scenario?
-  •	Next steps
-
--->
 
@@
 
-The 15-minute neighbourhood is a concept in planning that promotes urban spaces that are responsive to human needs and environmental sensibilities.
-These neighbuorhoods aim to accomplish this by ensuring that most daily necessities are not further away than a 15 minutes active trip [@Allam2022].
-In practice, 15-minute neighborhoods are a form of normative accessibility [@paez_measuring_2012], and therefore related to accessibility analysis, an increasingly important tool that planners use to foster just and sustainable cities [@vale2023accessibility; @handyAccessibilityIdeaWhose2020].
-Both the 15-minute neighborhood and accessibility measures are flexible enough to consider all destination types holistically, however, gendered examinations have been lacking in the literature.
-For instance, accessibility analyses have historically been employment-centric even when adopting a gendered perspective [@paez_jobs_2024].
+The "15-Minute Neighbourhood" concept has been adopted by leaders around the world as a way to promote human-scale cities.
+The concept emphasizes a transport and land-use mix that supports short-trips to daily necessities by foot or bicycle from ones' residence.
+This study adopts the 15-minute neighbourhood concept through the lens of "Mobility of Care", a framework that foregrounds travel to care destinations, travel done predominately by women, in contrast to the better-studied travel to employment and leisure.
+While the 15-minute neighbourhood concept is flexible enough to consider all destination types, gendered examinations are lacking in the literature and no research to date has focused explicitly on care destinations.
+In this study, we address these gaps by identifying which areas in Hamilton, Canada are ‘caring 15-minute neighbourhoods'.
+To do so, a database of care destinations was compiled and used to estimate both the number (cumulative opportunities indicator) and diversity of mix (entropy measure) of destinations reachable within a 15-minute walk from all residential parcels.
+Parcel-level results were classified into 'caring' (quantity of reachable opportunities) and 'complete' (diversity of care destinations) neighbourhood typologies of varying degrees using the Self-Organizing Maps (SOM) machine learning technique.
+Inequalities in the relationship between the identified caring typologies and social-economic indicators were then explored using SOM.
+Our results suggest the majority of 'caring and complete' neighbourhoods are within the urban core where a higher proportion of lower-income and single-parent households reside.
+Further, certain neighbourhoods in suburban and rural communities are 'caring and somewhat complete', showing the potential to become 'complete' through possible urban policy intervention.
+Taken together, this study offers a theoretical bridge to connect Mobility of Care and the 15-minute neighbourhoods concept, a methodology to assess neighbourhoods on a 'caring and complete' continuum, and discussion of relevant gender mainstreaming policy interventions.
 
-To address this gender bias, this study investigates the 15-minute neighbourhood through the lens of Mobility of Care [@sanchezdemadariagaMobilityCareIntroducing2013].
-This framework emphasizes the importance of travel related to unpaid work (care trips) in contrast to the better-studied travel to employment and leisure [@lawWomenTransportNew1999; @cresswell_gendered_2008; @Oestergaard1992; @levy1991towards; @levy2013travel].
-While these trip types (work, care, and leisure) are essential, care trips are often relatively shorter-distance, proximate to residential/work/school, and comprise approximately 30% of adults’ daily trips [@sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory; @mejia2021mobility], thus fitting neatly within the concept of 15-minute neighbourhoods.
-
-Our study casts the 15-minute neighbourhood onto the Mobility of Care framework.
-Empirically, we identify which areas in Hamilton, Canada, are 'caring 15-minute neighbourhoods'.
-To do so, a database of care destinations was compiled.
-In this database, care destinations include all places associated with sustaining household tasks needed for the reproduction of life namely: shopping (e.g., groceries), errands (e.g., libraries), health (e.g., dentist), and caring for dependents (e.g., schools).
-This database is used to estimate the number and mix of care-destinations that can be reached within a 15-minute walk-sheds from census centroids.
-We generate typologies to illustrate which neighbourhoods can and to what degree facilitate 15-minute access to care.
-
-Our results indicate which 15-minute neighbourhoods are almost 'care-complete', that is, contain a sufficiently high mix of care destinations.
-Very few of these are found outside of the downtown core.
-Our study frames these neighbourhoods on a continuum of 'caring' and helps identify areas for policy intervention.
-Taken together, this study provides a theoretical bridge to connect and Mobility of Care, 15-minute neighbourhoods, and accessibility analysis.
-This should help to inform policy choices aimed at encouraging more just and sustainable mobility.
+Keywords: Accessibility; 15-Minute Neighborhoods; Chrono-Urbanism; Cities of Proximity; Mobility of Care; Inequality; Gender Gap;
 
@@
 
-Caring is one of the most fundamental activities in societies, since it consists of numerous activities that allow them to thrive.
-Caring activities are those that strive to meet the physical, psychological, and emotional needs of others [@ilo_care_2018].
-Yet, it is one of the most unequal, undervalued, and even devalued activities worldwide.
-Traditionally, caring activities have been borne on women’s shoulders [@hayden_grand_2000; @hochschild_second_2012].
+The "15-Minute City" has been adopted by leaders as a way to promote human-scaled cities.
+As defined in @moreno_introducing_2021, the 15-Minute City is a urban planning model based in "chrono-urbanism", a theory that emphasizes the positive impact on quality of urban life when urban space becomes multi-rythmic [related to chronotopy @mulicekUrbanRhythmsChronotopic2015].
+Multi-rhythmic urban space in opposition to industrial Fordism and single-use zoning [@mulicekUrbanRhythmsChronotopic2015; @morenoVilleQuartHeure2016]; the segregation of urban time and space for individual uses and mobility.
+The 15-Minute City is one that accommodates the ability to reach essential destinations such as shops, work, and healthcare within a walkable radius.
+It would allow individuals to reclaim time spent on car mobility and give way to sustainable modes; prompting urban spaces that are responsive to human needs and environmental sensibilities [@Allam2022].
+The 15-Minute City generated much interest in planning practice and public discourse, especially after Paris mayor Anne Hidalgo proposed a 15-minute city plan in her 2020 re-election campaign.
+Though the term 15-Minute City is relatively new, the core of the concept is not.
+
+Literature has drawn parallels from the 15-Minute City to Perry's debated Neighbourhood Unit Concept (NUC) from the 1930s [@kissfazekasCircleParadigms15minute2022]; a 'socio-spatial' normative scheme widely adopted by government agencies (and the real-estate community) in the Western world after the Second World War [@talenSocialSciencePlanned2017; @solow1969concept].
+Pairing well with the objectives of planning agencies at the time, Perry's NUC would allow for efficient mass-building of cellular units that prioritized the perceived functional needs of women and children: each unit providing proximate access to an elementary school and sufficent supporting community facilities, shopping, parks and housing [@talenSocialSciencePlanned2017; @brodyNeighbourhoodUnitConcept2013].
+The NUC primarily prioritized local service provision, though Perry had confidence in good design's contribution to 'neighbourhood spirit' [@hall2014cities].
+From the NUC's conception until the end of the 1960s, planner's attempts to prescribe social meaning to the neighbourhood's physical form had been criticized to near extinction [@talenSocialSciencePlanned2017].
+By social scientists, the overestimation on the impact of built environment on social life was critiqued.
+By other planners, the inability to agree on the specificity of the neighbourhood (i.e., population sized, and the type, quality, and quantity of amenity) as well as how neighbourhood units are connected to other neighbourhoods and the rest of the region.
+<!-- failing to combine work, home and cultre was a dilbirate outcome of state-sponsored capitalism X.
+As stated in [@talenSocialSciencePlanned2017],-->  In response to these criticisms, neighbourhood planning proponents have redefined deterministic terminology: prescribed physical forms is 'enabling' rather than engendering, and the provision of social contact 'opportunities' instead of determined social behaviour.
+The redefined bottom-up concept to community and neighbourhood planning has been adopted by New Urbanist from which the 15-Minute City stems.
+The 15-Minute City is created through bottom-up proximity-based processes of community planning [@kissfazekasCircleParadigms15minute2022].
+
+<!-- These critiques were supported by case studies of planned neighbourhoods as...
+lead to social segregation of classes, destination functions, and automobile use.
+From it's onset, it was an efficient way to develop housing after the Second World War.
+The emphasis of proximity to some 'daily essentials' but disconnection from specialised activites (e.g., work) paved the way for car-based subrbia.
+by capital of cities.
+Can the neighbourhood form might be seen as a vehicle for capitalist exploitation, gentrification and displacement???
+The place of neighbourhood planning is actively being debated: can it's social crtiques be addressed in the new form as a "15-Minute City"?
+This study seeks to contribute to this decades old debate.
+-->
+
+However, critiques of social scientists and planners from decades past still persist.
+Can physical form be planned to enable an improved quality of life?
+This line of reasoning has occupied urban planning research in recent years: what are the impacts of built environment on residence's outcomes.
+A popular tool to analyse this relationship has often been accessibility analysis: proximity to X amenity results in X outcomes for a neighboourhood.
+Accessibility is a way of quantifying the ease of opportunity interaction from some point in space, this indicator is becoming increasingly important for planners in the pursuit of just and sustainable city planning [@vale2023accessibility; @handyAccessibilityIdeaWhose2020].
+As a amenity-provision concept of neighbourhood planning that seeks to enable an improvement on urban quality of life, the 15-Minute City lends well to analysis using accessibility tools.
+
+Before discussing 'what outcome', we first ask 'what amenity' as the focus of this study.
+Often, accessibility analysis studies have focused on a single amenity type (X), often employment destinations and often focused on a scale beyond the neighbourhood.
+However, due to 15-Minute City's emphasis on proximate-level service provision, recent academic work on the topic has employed accessibility analysis to account for multiple 'local' amenities types, but seldom included employment.
+We argue that the inclusion of employment opportunities into the 15-Minute City conceptualisation is critical to it's definition.
+Why?
+We echo critics made by planners and social scientistics in decades past: neighbourhoods that did not mix work with service and culture within the neighbourhood, paved the way for car-based suburbia: cars allowed for the connection of neighbourhoods with their 'lower-order' amenities to areas with 'high-order' destinations like work (X).
+The lower- and higher- order framing of these destinations supported the continued segregation of urban life across many dimensions (X).
+
+Combing all amenity types into the 15-Minute City concept is also a matter of gender equality.
+There is a branch of research that encompases this critique: Mobility of Care.
+This framework encourages the explicit consideration of destination types that serve unpaid labour (care destinations) into transport planning (as opposed to only considering destinations with paid work).
+Not considering these care destinations have travel time (and hence quality of life) implications on those who do care activites, commonly women and especially disadvantaged women.
+The consideration of _what_ amenity type is a gender issue.
+
+From this position, this study aims to holistically investigate the ease of interacting with proximate amenities (accessibility) from the perspective of the 15-Minute City and the Mobility of Care framework.
+As mentioned, accessibility analysis have typically focused on employment (X) and recent 15-Minute City flavoured analysis focus on 'lower-order' amentities.
+We focus on a case study of the mid-sized City of Hamilton, Canada, and aim to classify areas by what degree they may be considered a '15-Minute Neighbourhood" based on the relative quantity of amenity access and diversity of access.
+We offer a novel methodology of holistically considering all proximate-amenities, offer empirical findings, and contribute theoretically to the ongoing debate on the definition of a 15-Minute City by foregrounding care destinations.
+
+--------------
+
+Caring activities meet the physical, psychological, and emotional needs of others, and are one of the most essential and fundamental activities in society [@ilo_care_2018].
+Yet, they are the most unequal, undervalued, and even devalued activities worldwide.
+Traditionally, caring activities have been borne on women’s shoulders [@haydenGrandDomesticRevolution1982; @hochschild_second_2012].
 According to @ilo_care_2018, women and girls perform more than three-quarters of the total amount of unpaid care, a gender gap that varies geographically [@RN5].
@@
 This unequal share of caring responsibilities leads to multifaceted gendered differences: in career development, profession selection, contract type, pay gap, and time poverty, as recognized by various international organisations [@EIGE2016; @ilo_care_2018].
-In fact, early feminists in the United States had already observed that domestic labor was the single most relevant cause of women’s inequality in society [@gordon_womans_1976; @hayden_grand_2000].
+In fact, early feminists in the United States had already observed that domestic labor was the single most relevant cause of women’s inequality in society [@gordon_womans_1976; @haydenGrandDomesticRevolution1982].
 
@@
 Indeed, researchers from various fields have highlighted that western modern industrial societies have imposed a physical separation of the private sphere from the public one, splitting production and reproduction, public and private spaces or simply work and home, concentrating on the economic outcome of being mobile [@cresswell_gendered_2008; @lawWomenTransportNew1999; @levy2013travel; @little_gender_1994; @tronto_toward_1990].
-The man with legal rights, paid labour public sphere and woman in the private sphere, dependent, with family duties, personal relations and unpaid labour [@tronto_toward_1990].
-In this line, @cresswell_gendered_2008 described how masculinity was related to mobile and active, whereas femininity to static.
-As an example of this dichotomy, transport planning historically framed non-work/school trips as 'non-mandatory' activities, downgrading these activities in comparison to work.
-In more recent years, planners have begun to resist this spatial separation of the public and private sphere to promote for sustainable cities.
-Proximity planning has become more normalized, at popular example of which is the 15-minute neighbourhood, in which residents can reach all necessary destinations within a 15-minute walk.
+Men tend to be associated with legal rights, paid labour, and the public sphere while woman are associated with the private sphere, dependents, family duties, personal relations and unpaid labour [@tronto_toward_1990].
+In this line, @cresswell_gendered_2008 described how masculinity has been constructed as mobile and active, whereas femininity is static.
+As an example of this dichotomy, transport planning has historically framed non-work/school trips as 'non-mandatory' activities, downgrading these activities in comparison to mandatory trips like work and school (XXX).
+In more recent years, planners have begun to resist the spatial separation of the public and private sphere to promote for sustainable cities.
+Proximity planning has become more normalized, a popular example is the 15-minute neighbourhood i.e., neighbourhoods where residents can reach all necessary destinations within a 15-minute active transport trip.
 The 15-minute neighbourhood concept, however, has yet to foreground care.
@@
 In other words: caring for oneself and others requires trips to, amongst other places, grocery stores, day cares, elder care, and health care destinations.
-
-From this research motivation, Sánchez de Madariaga coined the term ‘mobility of care’ in 2013 to refer to this travel or all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands [@sanchezdemadariagaMobilityCareIntroducing2013].
+From this research motivation, Sánchez de Madariaga coined the term ‘mobility of care’ in 2013 to refer to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands [@sanchezdemadariagaMobilityCareIntroducing2013].
 While decades of research had examined these unique household-serving trips, her seminal work was the first to consider all these trips as one category.
@@
 It's novel but movements exists in the Canadian context i.e., https://x.com/brenttoderian/status/1800371515877384527?s=48&t=QI0WSvV_FDb82ZGeLdJ9lQ.
-Also discuss Barcelona and Bogota (care blocks)", " ")`
+Also discuss Barcelona and Bogota (care blocks) non academic sources: https://oecd-opsi.org/innovations/bogota-care-blocks/ AND https://www.npr.org/sections/goatsandsoda/2023/10/16/1205218707/how-bogota-cares-for-its-family-caregivers-from-dance-classes-to-job-training )", " ")`
 
-Within the backdrop of urban planning and 'mobility of care', we connect our work to the 15-minute neighbourhood concept.
-
-The spatial structure of cities and regions has been described in different ways and its advantages and disadvantages has been discussed with regard to a wide variety of issues, for example, regarding health, the environment, economic, social and equity impacts.
+<!-- The spatial structure of cities and regions has been described in different ways and its advantages and disadvantages has been discussed with regard to a wide variety of issues, for example, regarding health, the environment, economic, social and equity impacts.
 After the traditional monocentric vision of the city with a core Centre Business District (CBD), other alternatives appeared, especially due to the expansion of cities caused by the automobile and motorized transport.
@@
 This concept may have different interpretations depending on the spatial and geographical scales [@hall_polycentricity_2009].
+-->
 
@@
 The Portland Plan [@city_of_portland_20minute_2010] was adopted by the City Council on April 25, 2012 aiming at fostering an inclusive city development based on prosperity, education, health, and equity for the following 30 years.
-The progress report states that equity should be achieved, an Office of Equity and Human Rights was established focusing on equitable service delivery to all residents.
-Nonetheless the progress report only measures progress on topics related to racial equity or people with disabilities [@portland_government_portland_2017].
+The progress report describes the tasks of the established Office of Equity and Human Rights was established focusing on equitable service delivery to all residents.
+In the first example, equity is constrained to topics related to racial equity or people with disabilities [@portland_government_portland_2017].
 `r word_comment("NOTE TO SELF: And so what?
@@
 
+Later, other examples have highlighted the benefits of mobility of proximity without specifically targeting mobility of care, where a universal and neutral entity approach persists.
 In Paris, the adoption of the 15-minute neighbourhood put into the spotlight this approach.
@@
 Given the impact of caring activities on everyday life, the chrono-urbanism should specifically target the analysis of caring activities.
+
+Within the backdrop of urban planning and ‘mobility of care’, we connect our work to the 15-minute neighbourhood concept.
+
+Both the 15-minute neighborhood and accessibility measures are flexible enough to consider all destination types holistically, however, gendered examinations have been lacking in the literature.
+For instance, accessibility analyses have historically been employment-centric even when adopting a gendered perspective [@paez_jobs_2013].
+Some 15-minute city analyses focus on care destinations, but they seem to not put care at the forefront, or to consider but one care destination (e.g., [@hosford15minuteCityReach2022] on grocery stores).
+Though few 15-minute neighbourhood analyses are explicitly feminist, their potential to build feminist communities have been noted outside of academia.
+For instance, Women in Urbanism have noted how women may benefit more than men from this 15-minute approach to planning.
+For instance, women tend to walk and use transit more than men, as well as to make more frequent, but shorter trips -and often as part of a chain (i.e., with stops along the way).
+These trip characteristics are all well served in 15-minute frameworks.
+15-minute neighbourhoods also promote walkability, which results in more “eyes on the streets” and its associated safety, an important enabler of mobility for women and girls.
+Finally, because women tend to earn less money than men and cars are a large financial burden, planning paradigms that reduce car dependency, such as 15 minute cities, may on average benefit women more than men [@Bertolo2023].
 
@@
 It seems that car use may be preferred for mobility of care trips, but planning mobility of care for the car may leave those who do not have access to a car, many of them low-income women, with difficult mobility of care trips.
+Studies also observe disparities regarding gender mobility and type of activities, depending on the area of residence; women living in the city center, where more transport systems and activities are available, use more sustainable travel modes than the ones living in the outskirts, which tend to rely more on cars to carry out daily activities [@marquet2014walking; @mejia2021mobility; @montero2023role; @montero2023applying].
+Outer areas are not as dense as in the downtown, and are basically car-oriented developments.
+Thus, studies observe that in the outer areas gender equity is confronted to sustainable mobility.
 
@@
 
-Taken together, there appears to be a preference for travel by car for mobility of care trips.
+Taken together, there appears to be a preference for travel by car for mobility of care trips in many geographical contexts.
 This is problematic, however, in at least two major ways.
@@
 
+Taken together, city planning is moving to a people-centered approach focusing on accessibility to basic needs.
+However, chrono-urbanism stems from conventional methods that overlooks mobility of care perspectives.
+To counter this, this paper presents a 15-minute cities methodological approach to assess mobility of care.
+We develop this approach using a mid-size typical Canadian city and a novel care destination dataset as case study.
 The objectives of the paper are as follows:
 
-1)	City planning is moving to a people-centered approach focusing on accessibility to basic needs.
-However, chrono-urbanism stems from conventional methods that overlooks mobility of care perspectives.
-This paper presents a 15-minute cities methodological approach to assess mobility of care.
-2)	We develop this approach using a mid-size typical Canadian city and a novel care destination dataset as case study
-3)	Furthermore, the potential accessibility to and diversity of care destinations in the case study is analysed.
+1)	To examine the potential accessibility to, and diversity of, care destinations in the case study.
 We identify to what degree are neighbourhoods caring and/or complete.
 To do so, we make use of self organizing maps (SOM), an unsupervised machine learning data-dimensionality reduction technique to identify caring and complete neighbourhoods.
-4)	We then apply the same SOM methodology to determine profiles of population groups to understand who enjoys these neighbourhoods and who does not.
+2)	We then apply the same SOM methodology to determine profiles of population groups to understand who enjoys these neighbourhoods and who does not.
 
-# Data and Methods
+# Data
 
@@
 This study focuses on Hamilton, Ontario, a mid-sized city on the shore of Lake Ontario.
-Hamilton has a heterogenous land-use, with a populated and dense urban core, surrounded by suburban development, which is itself surrounded by rural communities made possible through a greenbelt.
+Hamilton has a heterogenous land-use, with a populated and dense urban core, surrounded by suburban development, which is itself surrounded by rural communities.
+The Niagara Escarpment runs through Hamilton, and results in a city with two key elevations: the dense lower city that contains the downtown core and the elevated suburban development referred to as ‘the Mountain’ (ADD THIS TO FIGURES).
 In this work, we analyse the residential parcel centroids, approximately ~140,000 locations.
@@
 Hamilton also exhibits spatial disparities in social and economic indicators @fig-Fig2.
-The densely populated inner city is characterised by lower average incomes, and a higher prevalence of households living under the LICO.
+The densely populated inner city is characterised by lower average incomes, and a higher prevalence of households living under the low-income cuttoff thresholds (LICO).
 The suburban parts of the city tend to have a greater proportion of children and a lower proportion of one parent households.
@@
 ## Care destination dataset
-
 A spatial dataset of care destinations for Hamilton was compiled; sources include provincial and municipal open data portals [@governmentofontarioOntarioDataCatalogue2023; @cityofhamiltonOpenData2023], Data Axle, a consumer dataset compiled of businesses and companies within Canada [@axledataConsumerData2023] and manual retrieved through Google Maps.
@@
 ```{r creating-table-content}
-dat <- data.frame(matrix(c("Child-centric",
-                           "[@opendatahamiltonEducationalInstitutions2022; @governmentofontarioOntarioDataCatalogue2023; @cityofhamiltonChildCareRegistry2023; @opendatahamiltonParksDatabase2022; @opendatahamiltonRecreationCommunityCentres2022]",
-                           "Schools, daycares, and community centres, recreation centres, and parks: 1,190 locations are included.
+dat <- data.frame(matrix(c("Depedent-centric",
+                           "[@opendatahamiltonEducationalInstitutions2022; @governmentofontarioOntarioDataCatalogue2023; @cityofhamiltonChildCareRegistry2023; @opendatahamiltonParksDatabase2022; @opendatahamiltonRecreationCommunityCentres2022; @ontarioministryofhealthgeohubMinistryHealthService2023]",
+                           "Schools, daycares, and community centres, recreation centres, parks, senior centres, long-term care homes, and retirement homes: 1,265 locations are included.
 After manual review, all locations that typically do not serve children were removed including: Post-Secondary, Adult-Learning Centres, Group Homes, and Foster Care Centres.
 Further, through examination some Section 23 institutions defined as _“centres for children who cannot attend school to meet the needs of care or treatment, and rehabilitation”_ [@governmentofontarioministryofeducationFundingEducationCommunity2023], were kept due to their innate connection to care.",
-                           "Elder-centric",
-                           "[@opendatahamiltonRecreationCommunityCentres2022; @ontarioministryofhealthgeohubMinistryHealthService2023]",
-                           "Senior centres, long-term care homes, and retirement homes: 75 destinations are identified.",
                            "Grocery-centric",
@@
 
-## Methods
+# Methods
+
+## Accessibility and entropy
 A 15-minute normative walk catchment is used to represent the 15-minute neighbourhood concept.
@@
 
-Self organizing map (**SOM**) is used to cluster the types of neighbourhoods by their magnitude in 15-minute walking accessibility to caring opportunities and diversity in these opportunities and profile these clusters based on their socio-economic characteristics.
+## Self-organizing maps
+
+Self organizing map (SOM) is used to cluster the types of neighbourhoods by their magnitude in 15-minute walking accessibility to caring opportunities and diversity in these opportunities and profile these clusters based on their socio-economic characteristics.
 SOM is a type of artificial neural network (ANN) trained using unsupervised learning that produces a two-dimensional representation of the input space.
 It is a dimensionality reduction method as it applies competitive learning by way of a neighbourhood function thereby preserving the topological properties of the input space and producing combinations of the input space that we intuitively profile.
-It has been used in transportation related research, e.g., @victorianoTimeSpaceMoney2020.
+It has been used widely in XX but scarcely in transportation related research [@victorianoTimeSpaceMoney2020]
 
@@
 
+## Decision trees
+
+Decision trees area a supervised learning technique, which can use the results of the clustering algorithm as labels, and the input variables as features.
+
 # Results
 
-Accessibility per category @fig-Fig4.
+Accessibility results per care category are presented in @fig-Fig4.
 The concentration of opportunities for all category types is concentrated in Hamilton Central - with the highest concentration being in the lower city and pockets of the high concentration on the mountain `r word_comment("(TO ADD: The escarpment line)", " ")`.
@@
 ```
-Diversity @fig-Fig5. Concentrated in the downtown core.
+
+The diversity of care destinations accessible is presented in @fig-Fig5.
+Here, it appears that neighbourhoods with a diversity of care destinations accessible are concentrated in the downtown core.
+Of note, some suburban areas also exhibit a high diversity of care access, these areas _have the potential_ to be retrofitted/ designed for access care (ADD MORE).
+
 ```{r plot-figure5, out.width=600, out.height=600}
@@
 
-The constitution of the superclusters in shown in the decision tree diagram in @fig-Fig7.
+The constitution of the superclusters is shown in the decision tree diagram in @fig-Fig7.
 Decision trees area a supervised learning technique, which can use the results of the clustering algorithm as labels, and the input variables as features.
@@
 Visually, the super clusters closely follow the quartile split of the accessibility to care destinations (all categories combined) and tertile split of the diversity scores (visualised in @fig-Fig5) `r word_comment("AN IDEA DISCUSSED: Overall, in @fig-Fig8 there are many pockets of good/great care accessibility+diversity of opportunities Hamilton.
-These are objectively caring neighbourhoods.
+These are objectively caring neighbourhoods (BY MEASURE).
 However, people may subjectively disagree...
-", ".")`.
+Maybe they don't feel safe, that is not captured.
+the limitations of objective measures ", ".")`.
 
+FIX THIS FIGURE'S LEGEND!!!!!!!
 ```{r plot-figure8, out.width=600, out.height=600}
@@
 
-In @fig-Fig9, discussing from right-to-left, parcels in DAs with 45% of the population below the median or lower (richer) are often classified with poor accessibility: either "Somewhat caring and not complete" SC3 or "Not caring and not complete" SC5, where parcels within DAs with a higher proportion of single parent households (13% and above) are most often SC3 and below are most often SC5.
+<!--EXPLAIN THE FOLLOWING PARAGRAPH BETTER-->
+In @fig-Fig9 from right-to-left, parcels in DAs with 45% of the population below the median or lower (richer) are often classified with poor accessibility: either "Somewhat caring and not complete" SC3 or "Not caring and not complete" SC5.
+Parcels within DAs with a higher proportion of single parent households (13% and above) are most often SC3: "Somewhat caring, not complete" and below are most often SC5: "Not caring, not complete".
 For parcels with higher than 45% of bottom-income (poorer), those with a high proportion of children (age 0 to 14) are also likely to be within the "Somewhat caring and not complete" SC3 - just like parcels in richer and more-single-parent DA counterparts.
@@
 Interestingly, parcels in DAs with lower than 13% children population demonstrate a better accessibility to care destination story.
+<!--PERHAPS A MISMATCH ON WHO NEEDS / WHO GETS CARE ACCESSIBLE DESTINATIONS-->
+
 Parcels within DAs that constitute more than 45% population below median household income (poorer) - especially above 9.8% LICO (low income cut-off prevalence - likely to be the poorest) are often within "Very caring and very complete" SC1.
+<!-- -maybe people who can’t afford a car choose to live here?
+-->
+
 Parcels located in DAs with the same lower median income and child population but have _lower_ than 9.8% LICO (not the poorest) also have good accessibility, as they are most often in "Caring and complete" SC1 as well as being in most often in the "Somewhat caring and complete" SC4 relative to other parcels.
@@
 On the horizon as rent increases,  the most precarious may be pushed out and replaced by people who can pay higher rents.
-2) The lowest accessibility & most incomplete are suburban -- they are likely higher income.
-Interesting to note: one-parent families are more likely to be time poor -- and leaving in low access/incomplete suburbs puts more at risk.
-We know a think or 2 about single parents and jobs -- access to jobs was high where single parents were not.
-(Catherine, farber Paez paper).
-3) Pockets of the suburbs have high access to care.
-The suburbs do something -- they have care amenities.
 ", " ")`
+
   `r word_comment("IDEAS DISCUSSED: regarding @fig-Fig9 - 2) The lowest accessibility & most incomplete are suburban -- they are likely higher income.
@@
 
+  <!--Yes: low access and more incomplete = suburban (Figure 8) AND higher income (Figure 9).
+Are higher income folks living in low access and low caring places, but don’t care because they can afford to drive to get to these places?
+This has implications for sustainability and car dependence
+
+  And also: poorer one parent families with children are also in these not caring/ complete places.
+They are who may feel it: as they likely are more time poor and have lower access to cars.
+This could impact their wellbeing / access to care (which is essential) -->
+
  `r word_comment("IDEAS DISCUSSED: regarding @fig-Fig9 - 3) Pockets of the suburbs have high access to care. The suburbs do something -- they have care amenities. ", ":")`.
 
-# Conclusions and Discussion
+<!--Potentially suburban have greater access to care compared to jobs.
+I mean people live there for a reason I guess!
+And they are often seen as ‘child friendly’.
+If these suburban areas have high access to care by foot, are they going there by foot?
+What can be done to get them to walk to care destinations?
+Or are these examples of suburban that have been designed to be walkable (which should be emulated)?
+-->
+
+# Discussion
 
@@
 On the other hand, an intersectional approach to understand how different personal identities interact in our daily day activities in the cities is necessary.
+
+The process of gentrification and displacement are trends also occurring in Hamilton.
+On the horizon of rising rents, there is a risk of displacement among more vulnerable households, who may be replaced by individuals with higher incomes who can afford different transport modes, with more flexible schedules and are more likely to use ICT arrangements (including telework, e-shopping, and other online services).
+While, lower income households may be particularly vulnerable to locating in areas with worse accessibility to care alternatives, which could have negative consequences for these households and for the society as a whole.
+Even if we are on the right direction to develop human-centred urban areas by facilitating caring activities, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities, given the multiple impacts it has on everyday mobility, the unequal distribution of tasks among the population and its related socioeconomic consequences.
+Therefore, it is necessary to provide alternatives to measure progress on this important issue. 
+This work leaves many doors open for further research. For example, 
+•	the role that gender plays in these households. 
+•	Job arrangements per type of household.
+Some studies observe that working from home arrangements are more likely available for higher education attainment households (likely to have higher income), and families with children, but also living in neighbourhoods with good accessibility,
+•	Finally, an intersectional approach to understand how our personal identities interact in our access to essential activities and caring services.
+City planning should respond to these issues to promote equity.
+On the other hand, an intersectional approach to understand how different personal identities interact in our daily day activities in the cities is necessary.
+
+LUCIA'S:
+As in other case studies, essential opportunities and services tend to concentrate on Hamilton central.
+Therefore, it is more likely to carry out daily activities, including caring ones, in this area using sustainable modes such as walking or biking.
+
+However, this study highlights that there seems to be a spatial mismatch between dwelling availability, family profile and eventually (housing and mobility) affordability.
+At neighbourhood level the concentration of opportunities for all category types is in Hamilton Central, but households with children may not live there.
+
+Richer families may concentrate in either neighborhoods classified with poor accessibility: either “Somewhat caring and not complete” or “Not caring and not complete.
+However, these families are likely to have different mobility alternatives available.
+The literature shows that higher income households have a lower mobility share compared to lower income households.
+Furthermore, they are more likely to be assisted by different service arrangements (i.e.
+e-shopping, service deliveries, taxis and MaaS services, working from home arrangements,,…).
+
+Our results show a different story for low-income households.
+If they can afford it, these households will prefer to live in very caring and very complete neighbourhoods, which will ease their accessibility to daily activities and essential services and therefore, avoiding social exclusion.
+In this respect, one-parent families are more likely to be time poor.
+For them, living in low access/incomplete suburbs increases their chancesof  falling into social exclusion.
+Furthermore, school closures or any other type of relocation of essential services puts more pressure on their shoulders.
+
+Measuring caring neighbourhoods is of utmost importance to promote equity in city planning.
+Our results are of use for social housing programs for lower income households.
+Housing is a human right which may not be disentangle from mobility needs.
+Otherwise, lower income household risk of social exclusion, especially when accessibility to essential services is limited.
+
+Similarly, this tool has great potential to support city planning, by analyzing the population groups benefited or harmed by relocation, closure or opening of different types of essential services.
+
+The process of gentrification and displacement are trends also occurring in Hamilton.
+On the horizon of rising rents, there is a risk of displacement among more vulnerable households, who may be replaced by individuals with higher incomes who can afford different transport modes, with more flexible schedules and are more likely to use ICT arrangements (including telework, e-shopping, and other online services).
+While, lower income households may be particularly vulnerable to locating in areas with worse accessibility to care alternatives, which could have negative consequences for these households and for the society as a whole.
+
+Even if we are on the right direction to develop human-centred urban areas by facilitating caring activities, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities, given the multiple impacts it has on everyday mobility, the unequal distribution of tasks among the population and its related socioeconomic consequences.
+
+ANATASIA's
+- harkens back to what role a planned neighbourhood has. Is it about serivice provision (the what), social homogenity (the who), local power and democracy? []
+
+# Conclusion
+
+This study makes three types of contributions: empirical, methodological and theoretical.
+At the empirical level, areas of the mid-sized City of Hamilton have been typified by their degree of '15-Minute Caring Neighbourhood' potential.
+Methodologically, we applied the longstanding accessibility measure of cumulative opportunities and entropy to classify spatial areas based on how many destinations could be reached in a 15-minute walk and the diversity of destination type.
+These values were then clustered using a novel machine learning approach, SOM, to generate meaningful typologies for discussion and further comparision with socio-economic characteristic of the area.
+We find...
+
+Theoretically, our work puts forth an explicitly caring 15-Minute Neighbourhod conceptualisation, bridging the Mobility of Care and the "15-Minute City" concept.
 

### 2024-09-18 WGiT/WGiT-pres.Rmd
WGiT presentation

@@
     nature:
-      slideNumberFormat: "%current% / 10" 
+      slideNumberFormat: "%current% / 11" 
       highlightStyle: github
@@
 - $i$ is a set of parcel point origin locations; $j$ is a set of destination locations.
-- $O^_j$ is a number of opportunities at $j$.
+- $O_j$ is a number of opportunities at $j$.
 - $c_{ij}$ is the travel cost between $i$ and $j$.
@@
 
-### Typologies of 15-minute caring neighbourhoods
+### "Self Organinzing Maps" typologies: 15-min caring neighbourhoods
 
@@
 
-- Lower income households tend to reside in $\color{Green}{\text{very completely caring}}$ or $\color{GreenYellow}{\text{completely caring}}$ neighbourhoods; 
+- Lower income households tend to reside in $\color{Forestgreen}{\text{very completely caring}}$ or $\color{GreenYellow}{\text{completely caring}}$ neighbourhoods; 
 - Those with a `r emo::ji("up_arrow")` proportion of children tend to reside in $\color{orange}{\text{somewhat caring - not complete}}$ neighbourhoods;
@@
 
-Who has resides in $\color{Green}{\text{very completely caring}$ or $\color{GreenYellow}{\text{completely caring}$ neighbourhoods?  (maybe people who cannot afford cars???)
+Who has resides in $\color{Forestgreen}{\text{very completely caring}}$ or $\color{GreenYellow}{\text{completely caring}}$ neighbourhoods?  (maybe people who cannot afford cars???)
 - DAs that have lower median household incomes and a smaller children population (lower than 24% of the population are children 14 and under).
-- and DAs which have especially high LICO are often in $\color{Green}{\text{very completely caring}$ neighbourhoods.
+- and DAs which have especially high LICO are often in $\color{Forestgreen}{\text{very completely caring}}$ neighbourhoods.
 
-Who more often resides in $\color{BurntOrange}{\text{somewhat caring - not complete}$ neighbourhoods?  
+Who more often resides in $\color{BurntOrange}{\text{somewhat caring - not complete}}$ neighbourhoods?  
 - Of bottom income DAs, DAs with a high proportion of children population (higher than 24% of the population are children 14 and under).
 
-Who more often resides in $\color{BrickRed}{\text{not caring - not complete}$ neighbourhoods?  
+Who more often resides in $\color{BrickRed}{\text{not caring - not complete}}$ neighbourhoods?  
 - DAs of higher income DAs and with lower proportion of single-parent households (13% or lower)
@@
 
-### Conclusions: classifying and profiling caring 15-minute neighbourhoods
+### Conclusions
 
@@
 
-#### What we did:
+#### Implications of findings and future considerations:
 
-- 15 minute-city is bridged with "Mobility of Care": 15-minute caring neighbourhoods.
-- A data-driven pattern recognition method (Self-organizing maps) to create **15-minute caring neighbourhood** typologies using **accessibility** and **diversity** indicators is presented.
-- Using Self-Organizing Maps again; each parcel tagged with a neighbourhood typology was profiled based on all relevant census variables.
+- **Gentrification concerns**: populations residing in $\color{Forestgreen}{\text{very completely caring}}$  neighbourhoods are often lower-income and those in $\color{red}{\text{not caring - not complete}}$ neighbourhoods are better-off.
+
+- **Potential to improve caring for children and children's self-maintenance**: 
+  - Why do lower-income DAs with a `r emo::ji("up_arrow")` proportion of children, and DAs with `r emo::ji("up_arrow")` proportion single-parent households have (medium) $\color{orange}{\text{somewhat caring - not complete}}$ access?
+  - What barriers are they facing?
 ]
@@
 
-#### What we found and future considerations:
+#### Recap:
 
-- **Gentrification concerns**: populations residing in completely caring neighbourhoods are often lower-income and vice-versa.
-
-- **Concerning caring for children and children's self-care**: Lower-income DAs with a `r emo::ji("up_arrow")` proportion of children, and DAs with `r emo::ji("up_arrow")` proportion single-parent households have medium complete care access, why?
+- **15 minute-city** is bridged with **Mobility of Care**: 15-minute caring neighbourhoods.
+- A data-driven pattern recognition method (i.e., self-organizing maps) is used to create **15-minute caring neighbourhood typologies** using **accessibility** and **diversity** indicators.
+- Used Self-Organizing Maps again to **profile** these typologies based on relevant **census variables**.
+- Discussed implications: concerns about gentrification and mobility of care children's trips
 
@@
 
+---
+class: center, middle, inverse
+
+## Thank you for listening! `r emo::ji("smile")`
+
+### Any questions? Remarks?
+
+-
+
+#### Anastasia Soukhov, **soukhoa@mcmaster.ca** 
+#### Léa Ravensbergen, Lucía Mejía Dorantes, Antonio Páez
+
 ???

### 2024-09-18 editorial correspondance/Highlights.qmd
WGiT presentation

---
title: "Highlights"
format: html
---

Highlights:

- 15-minute neighbourhoods (FMN) have (re)-emerged as a popular concept in urban-planning city-planning; emphasizing the access to urban amenities by foot or bicycle. However, historically, neighbourhood planning processes have been top-down; FMC is linked to the Neighbourhood Unit Concept of the 1920s, largely critized for supporting segregated neighbourhoods, leaving a void in neighbourhood planning concepts that were filled by auto-centric suburban sprawl. Regardless, fundamental concepts within FMC remain popular: people identify with neighbourhoods, see cities through that lense. These ideas have in part re-emerged in the popularity of 15-minute neighbourhoods.

- But how do we uplift the positive, popular, aspects of FMN without opening it up to the critics of the NUC? In this work, we turn to feminist geographies and the Mobility of Care conceptualisation and apply it to FMC. We argue, conventional planning practice is masculinist. This can threaten the FMC concept by shaping how 'important' amenities are conceptualised thus favouring work- and leisure- centered amenities in cities that are increasingly capital-driven. Though FMN is fundamentally based in bottom-up collaborative planning processes, we exist in the context of planning pracice which is largely masculinist.

- FMN are for people - people do unpaid labour, those people are often women and intersectionally women-of-colour and low-income. However, this work should be for all people. Cities should be designed with this conceptualisation in mind. This is to combate the growing capitalisation of cities.

- Our paper offers this nuanced understanding of FNC -- calling for them to center Care Destinations as important amenity types with the explicit conceptualisation that their destination type matters especially for women (low-income and of colour) who interact with them most, but also for all people. Their explicit positioning also can indirectly combat finacialisation of cities and co-opting of the FMC concept to one that centers capital-generating destination types like work and prohibitive leisure. Not to say those destination types should or cannot co-exist with Care destinations -- they should -- but markets are already incentivizing them. The market does not speak for non-economically geberating destination types, so we must explicitly call for them and support them within the FNC framework. 
- 
- 
-

### 2024-09-27 data/03-clustering-and-profiling.qmd
Still working on the draft...

@@
 Normalize number of opportunities to min-max range:
+*MAYBE FIX THIS?? See 04*
 ```{r}

### 2024-09-27 data/04-bycategoriy-clustering.qmd
Still working on the draft...

---
title: "Clustering by category and plots"
author: "Antonio Paez & Anastasia Soukhov"
format: html
---

# Setting up table

Lets begin by loading the packages used for this:
```{r message = FALSE}
library(AppliedPredictiveModeling) # Functions and Data Sets for 'Applied Predictive Modeling'
library(caret) # Classification and Regression Training
library(cowplot) # Streamlined Plot Theme and Plot Annotations for 'ggplot2'
library(ggparty) # 'ggplot' Visualizations for the 'partykit' Package
library(party) # A Laboratory for Recursive Partytioning
library(plotly) # Create Interactive Web Graphics via 'plotly.js'
library(rpart) # Recursive Partitioning and Regression Trees
library(rpart.plot) # Plot 'rpart' Models: An Enhanced Version of 'plot.rpart'
library(skimr) # Compact and Flexible Summaries of Data
library(sf) # Simple Features for R
library(SOMbrero) # SOM Bound to Realize Euclidean and Relational Outputs
library(tidyverse) # Easily Install and Load the 'Tidyverse'
library(tmap)
library(here)
library(parsnip)
```

Read od table and the census:
```{r}
load("access_15.rda")
load("HAM_census_21.rda")
```

# Care Category superclusters

Filter by Care Category: first dependent, then  errand, health, and grocery centric:

## Dependent centric supercluster
```{r}
access_15_dep <- access_15 |>  filter(Care_Category == "Dependent-centric") |> 
  left_join(HAM_census_21 |> select(c(GeoUID, Population)) |> # and dssociate with census data and drop parcels in DAs with 0 population:
              st_drop_geometry(),
            by = "GeoUID")

access_15_dep <- access_15_dep |> filter(Population >= 0 | !is.na(Population))
```

Calculate total "Dependent" Specific Care Category opportunities by parcel, proportion, and normalized number of opportunities to min-max range:
```{r}
access_15_dep_total <- access_15_dep |>
  group_by(from_id) |>
  summarize(tot_copp_isc = sum(copp_isc),
            .groups = "drop")

#Join total opportunities to table:
access_15_dep <- access_15_dep |>
  left_join(access_15_dep_total,
            by = "from_id")

# proportion of opportunities by class (use a zero for the parcels that have accessibility zero):
access_15_dep <- access_15_dep |>
  mutate(prop_copp_isc = ifelse(tot_copp_isc > 0, copp_isc/tot_copp_isc, 0))

# Normalize number of opportunities to min-max range:
max_tot <- max(access_15_dep$copp_isc)
min_tot <- min(access_15_dep$copp_isc)

max_tot_2 <- max(access_15_dep$tot_copp_isc) #max number of category opportunities reached (e.g., for dep, includes all parks, rec centres, schools etc. reached)
min_tot_2 <- min(access_15_dep$tot_copp_isc)

access_15_dep <- access_15_dep |>
  mutate(copp_isc_norm = ifelse(tot_copp_isc != 0,
                                (copp_isc - min_tot)/(max_tot - min_tot),0),
         tot_copp_isc_norm = ifelse(tot_copp_isc != 0,
                                (tot_copp_isc - min_tot_2)/(max_tot_2 - min_tot_2),0))
```

Calculate diversity index:
```{r}
no_specific_care_categories <- access_15_dep$Care_Category_Specific |> unique() |> length()

access_15_dep_div <- access_15_dep |>
  group_by(from_id) |>
  summarize(div_dep = -sum(prop_copp_isc * log(prop_copp_isc), na.rm = TRUE)/log(no_specific_care_categories),
            .groups = "drop") #
```

Pivot od table wider so that each parcel has a column for the proportion of opportunities by each Specific Care Category:
```{r}
access_15_dep_prop_wide <- access_15_dep |>
  select(from_id, Care_Category_Specific, prop_copp_isc) |>
  pivot_wider(names_from = "Care_Category_Specific", 
              values_from = "prop_copp_isc") |>
  rename_with(-from_id, 
              .fn = ~paste0("prop_",
                            str_replace(.x, "-","_"))) 
```

Pivot od table wider so that each parcel has a column for the normalized number of opportunities by each Specific Care Category:
```{r}
access_15_dep_norm_wide <- access_15_dep |>
  select(from_id, Care_Category_Specific, copp_isc_norm,tot_copp_isc_norm) |>
  pivot_wider(names_from = "Care_Category_Specific", 
              values_from = "copp_isc_norm") |>
  rename_with(-from_id, 
              .fn = ~paste0("norm_", str_replace(.x, "-","_")))
```

Pivot od table wider so that each parcel has a column for the total number of opportunities by each Specific Care Category:
```{r}
access_15_dep_total_wide <- access_15_dep |>
  select(from_id, Care_Category_Specific, copp_isc, tot_copp_isc) |>
  pivot_wider(names_from = "Care_Category_Specific", 
              values_from = "copp_isc") |>
  rename_with(-from_id, 
              .fn = ~paste0("total_", str_replace(.x, "-","_")))
```

Join the tables:
```{r}
access_15_dep_wide <- access_15_dep_total_wide |>
  left_join(access_15_dep_norm_wide,
            by = "from_id") |>
  left_join(access_15_dep_prop_wide,
            by = "from_id") |>
  left_join(access_15_dep_div,
            by = "from_id") |>
  rename("total_tot_copp_isc_dep" = "total_tot_copp_isc",
         "norm_tot_copp_isc_norm_dep" = "norm_tot_copp_isc_norm")

input.dep <- access_15_dep_wide |> ungroup() |> select(-c(from_id, 
                                                          #starts_with("total_"), 
                                                          total_Care_Category, norm_Care_Category, prop_Care_Category)) 
  # Remove the non-normalized and proportional accessibility scores, as well as the Care Category character names. We are only clustering on the normalized accessibility scores (how many of this type of opportunity can you reach, normalized for the region) and the diversity (the evenness of type of opportunities that can be reached).

#rm(access_15_dep, access_15_dep_div, access_15_dep_norm_wide, access_15_dep_prop_wide, access_15_dep_total, access_15_dep_total_wide, access_15_dep_wide)
```

Now let's repeat for the other three categories.

## Errand-centric

```{r}
access_15_err <- access_15 |>  filter(Care_Category == "Errand-centric") |> 
  left_join(HAM_census_21 |> select(c(GeoUID, Population)) |> # and dssociate with census data and drop parcels in DAs with 0 population:
              st_drop_geometry(),
            by = "GeoUID")

access_15_err <- access_15_err |> filter(Population >= 0 | !is.na(Population))
```

Calculate total err Specific Care Category opportunities by parcel, proportion, and normalized number of opportunities to min-max range:
```{r}
access_15_err_total <- access_15_err |>
  group_by(from_id) |>
  summarize(tot_copp_isc = sum(copp_isc),
            .groups = "drop")

#Join total opportunities to table:
access_15_err <- access_15_err |>
  left_join(access_15_err_total,
            by = "from_id")

# proportion of opportunities by class (use a zero for the parcels that have accessibility zero):
access_15_err <- access_15_err |>
  mutate(prop_copp_isc = ifelse(tot_copp_isc > 0, copp_isc/tot_copp_isc, 0))

# Normalize number of opportunities to min-max range:
max_tot <- max(access_15_err$copp_isc)
min_tot <- min(access_15_err$copp_isc)

max_tot_2 <- max(access_15_err$tot_copp_isc) #max number of category opportunities reached (e.g., for dep, includes all parks, rec centres, schools etc. reached)
min_tot_2 <- min(access_15_err$tot_copp_isc)

access_15_err <- access_15_err |>
  mutate(copp_isc_norm = ifelse(tot_copp_isc != 0,
                                (copp_isc - min_tot)/(max_tot - min_tot),0),
         tot_copp_isc_norm = ifelse(tot_copp_isc != 0,
                                (tot_copp_isc - min_tot_2)/(max_tot_2 - min_tot_2),0))
```

Calculate diversity index:
```{r}
no_specific_care_categories <- access_15_err$Care_Category_Specific |> unique() |> length()

access_15_err_div <- access_15_err |>
  group_by(from_id) |>
  summarize(div_err = -sum(prop_copp_isc * log(prop_copp_isc), na.rm = TRUE)/log(no_specific_care_categories),
            .groups = "drop") #
```

Pivot od table wider so that each parcel has a column for the proportion of opportunities by each Specific Care Category:
```{r}
access_15_err_prop_wide <- access_15_err |>
  select(from_id, Care_Category_Specific, prop_copp_isc) |>
  pivot_wider(names_from = "Care_Category_Specific", 
              values_from = "prop_copp_isc") |>
  rename_with(-from_id, 
              .fn = ~paste0("prop_",
                            str_replace(.x, "-","_")))
```

Pivot od table wider so that each parcel has a column for the normalized number of opportunities by each Specific Care Category:
```{r}
access_15_err_norm_wide <- access_15_err |>
  select(from_id, Care_Category_Specific, copp_isc_norm, tot_copp_isc_norm) |>
  pivot_wider(names_from = "Care_Category_Specific", 
              values_from = "copp_isc_norm") |>
  rename_with(-from_id, 
              .fn = ~paste0("norm_", str_replace(.x, "-","_")))
```

Pivot od table wider so that each parcel has a column for the total number of opportunities by each Specific Care Category:
```{r}
access_15_err_total_wide <- access_15_err |>
  select(from_id, Care_Category_Specific, copp_isc, tot_copp_isc) |>
  pivot_wider(names_from = "Care_Category_Specific", 
              values_from = "copp_isc") |>
  rename_with(-from_id, 
              .fn = ~paste0("total_", str_replace(.x, "-","_")))
```

Join the tables:
```{r}
access_15_err_wide <- access_15_err_total_wide |>
  left_join(access_15_err_norm_wide,
            by = "from_id") |>
  left_join(access_15_err_prop_wide,
            by = "from_id") |>
  left_join(access_15_err_div,
            by = "from_id")|>
  rename("total_tot_copp_isc_err" = "total_tot_copp_isc",
         "norm_tot_copp_isc_norm_err" = "norm_tot_copp_isc_norm")

input.err <- access_15_err_wide |> ungroup() |> select(-c(from_id, 
                                                          #starts_with("total_"), 
                                                          total_Care_Category, norm_Care_Category, prop_Care_Category)) # Remove the non-normalized and proportional accessibility scores, as well as the Care Category character names. We are only clustering on the normalized accessibility scores (how many of this type of opportunity can you reach, normalized for the region) and the diversity (the evenness of type of opportunities that can be reached).

#rm(access_15_err, access_15_err_div, access_15_err_norm_wide, access_15_err_prop_wide, access_15_err_total, access_15_err_total_wide, access_15_err_wide)
```

## Health-centric

```{r}
access_15_health <- access_15 |>  filter(Care_Category == "Health-centric") |> 
  left_join(HAM_census_21 |> select(c(GeoUID, Population)) |> # and associate with census data and drop parcels in DAs with 0 population:
              st_drop_geometry(),
            by = "GeoUID")

access_15_health <- access_15_health |> filter(Population >= 0 | !is.na(Population))
```

Calculate total health Specific Care Category opportunities by parcel, proportion, and normalized number of opportunities to min-max range:
```{r}
access_15_health_total <- access_15_health |>
  group_by(from_id) |>
  summarize(tot_copp_isc = sum(copp_isc),
            .groups = "drop")

#Join total opportunities to table:
access_15_health <- access_15_health |>
  left_join(access_15_health_total,
            by = "from_id")

# proportion of opportunities by class (use a zero for the parcels that have accessibility zero):
access_15_health <- access_15_health |>
  mutate(prop_copp_isc = ifelse(tot_copp_isc > 0, copp_isc/tot_copp_isc, 0))

# Normalize number of opportunities to min-max range:
max_tot <- max(access_15_health$copp_isc)
min_tot <- min(access_15_health$copp_isc)

max_tot_2 <- max(access_15_health$tot_copp_isc) #max number of category opportunities reached (e.g., for dep, includes all parks, rec centres, schools etc. reached)
min_tot_2 <- min(access_15_health$tot_copp_isc)

access_15_health <- access_15_health |>
  mutate(copp_isc_norm = ifelse(tot_copp_isc != 0,
                                (copp_isc - min_tot)/(max_tot - min_tot),0),
         tot_copp_isc_norm = ifelse(tot_copp_isc != 0,
                                (tot_copp_isc - min_tot_2)/(max_tot_2 - min_tot_2),0))
```

Calculate diversity index:
```{r}
no_specific_care_categories <- access_15_health$Care_Category_Specific |> unique() |> length()

access_15_health_div <- access_15_health |>
  group_by(from_id) |>
  summarize(div_health = -sum(prop_copp_isc * log(prop_copp_isc), na.rm = TRUE)/log(no_specific_care_categories),
            .groups = "drop") #
```

Pivot od table wider so that each parcel has a column for the proportion of opportunities by each Specific Care Category:
```{r}
access_15_health_prop_wide <- access_15_health |>
  select(from_id, Care_Category_Specific, prop_copp_isc) |>
  pivot_wider(names_from = "Care_Category_Specific", 
              values_from = "prop_copp_isc") |>
  rename_with(-from_id, 
              .fn = ~paste0("prop_",
                            str_replace(.x, "-","_")))
```

Pivot od table wider so that each parcel has a column for the normalized number of opportunities by each Specific Care Category:
```{r}
access_15_health_norm_wide <- access_15_health |>
  select(from_id, Care_Category_Specific, copp_isc_norm, tot_copp_isc_norm) |>
  pivot_wider(names_from = "Care_Category_Specific", 
              values_from = "copp_isc_norm") |>
  rename_with(-from_id, 
              .fn = ~paste0("norm_", str_replace(.x, "-","_")))
```

Pivot od table wider so that each parcel has a column for the total number of opportunities by each Specific Care Category:
```{r}
access_15_health_total_wide <- access_15_health |>
  select(from_id, Care_Category_Specific, copp_isc, tot_copp_isc) |>
  pivot_wider(names_from = "Care_Category_Specific", 
              values_from = "copp_isc") |>
  rename_with(-from_id, 
              .fn = ~paste0("total_", str_replace(.x, "-","_")))
```

Join the tables:
```{r}
access_15_health_wide <- access_15_health_total_wide |>
  left_join(access_15_health_norm_wide,
            by = "from_id") |>
  left_join(access_15_health_prop_wide,
            by = "from_id") |>
  left_join(access_15_health_div,
            by = "from_id") |>
  rename("total_tot_copp_isc_health" = "total_tot_copp_isc",
         "norm_tot_copp_isc_norm_health" = "norm_tot_copp_isc_norm")

input.health <- access_15_health_wide |> ungroup() |> select(-c(from_id, 
                                                                #starts_with("total_"), 
                                                                total_Care_Category, norm_Care_Category, prop_Care_Category)) # Remove the non-normalized and proportional accessibility scores, as well as the Care Category character names. We are only clustering on the normalized accessibility scores (how many of this type of opportunity can you reach, normalized for the region) and the diversity (the evenness of type of opportunities that can be reached).

#rm(access_15_health, access_15_health_div, access_15_health_norm_wide, access_15_health_prop_wide, access_15_health_total, access_15_health_total_wide, access_15_health_wide)
```

## Grocery-centric

```{r}
access_15_groc <- access_15 |>  filter(Care_Category == "Grocery-centric") |> 
  left_join(HAM_census_21 |> select(c(GeoUID, Population)) |> # and associate with census data and drop parcels in DAs with 0 population:
              st_drop_geometry(),
            by = "GeoUID")

access_15_groc <- access_15_groc |> filter(Population >= 0 | !is.na(Population))
```

Calculate total groc Specific Care Category opportunities by parcel, proportion, and normalized number of opportunities to min-max range:
```{r}
access_15_groc_total <- access_15_groc |>
  group_by(from_id) |>
  summarize(tot_copp_isc = sum(copp_isc),
            .groups = "drop")

#Join total opportunities to table:
access_15_groc <- access_15_groc |>
  left_join(access_15_groc_total,
            by = "from_id")

# proportion of opportunities by class (use a zero for the parcels that have accessibility zero):
access_15_groc <- access_15_groc |>
  mutate(prop_copp_isc = ifelse(tot_copp_isc > 0, copp_isc/tot_copp_isc, 0))

# Normalize number of opportunities to min-max range:
max_tot <- max(access_15_groc$copp_isc)
min_tot <- min(access_15_groc$copp_isc)

max_tot_2 <- max(access_15_groc$tot_copp_isc) #max number of category opportunities reached (e.g., for dep, includes all parks, rec centres, schools etc. reached)
min_tot_2 <- min(access_15_groc$tot_copp_isc)

access_15_groc <- access_15_groc |>
  mutate(copp_isc_norm = ifelse(tot_copp_isc != 0,
                                (copp_isc - min_tot)/(max_tot - min_tot),0),
         tot_copp_isc_norm = ifelse(tot_copp_isc != 0,
                                (tot_copp_isc - min_tot_2)/(max_tot_2 - min_tot_2),0))
```

Calculate diversity index:
```{r}
no_specific_care_categories <- access_15_groc$Care_Category_Specific |> unique() |> length()

access_15_groc_div <- access_15_groc |>
  group_by(from_id) |>
  summarize(div_groc = -sum(prop_copp_isc * log(prop_copp_isc), na.rm = TRUE)/log(no_specific_care_categories),
            .groups = "drop") #
```

Pivot od table wider so that each parcel has a column for the proportion of opportunities by each Specific Care Category:
```{r}
access_15_groc_prop_wide <- access_15_groc |>
  select(from_id, Care_Category_Specific, prop_copp_isc) |>
  pivot_wider(names_from = "Care_Category_Specific", 
              values_from = "prop_copp_isc") |>
  rename_with(-from_id, 
              .fn = ~paste0("prop_",
                            str_replace(.x, "-","_")))
```

Pivot od table wider so that each parcel has a column for the normalized number of opportunities by each Specific Care Category:
```{r}
access_15_groc_norm_wide <- access_15_groc |>
  select(from_id, Care_Category_Specific, copp_isc_norm, tot_copp_isc_norm) |>
  pivot_wider(names_from = "Care_Category_Specific", 
              values_from = "copp_isc_norm") |>
  rename_with(-from_id, 
              .fn = ~paste0("norm_", str_replace(.x, "-","_")))
```

Pivot od table wider so that each parcel has a column for the total number of opportunities by each Specific Care Category:
```{r}
access_15_groc_total_wide <- access_15_groc |>
  select(from_id, Care_Category_Specific, copp_isc, tot_copp_isc) |>
  pivot_wider(names_from = "Care_Category_Specific", 
              values_from = "copp_isc") |>
  rename_with(-from_id, 
              .fn = ~paste0("total_", str_replace(.x, "-","_")))
```

Join the tables:
```{r}
access_15_groc_wide <- access_15_groc_total_wide |>
  left_join(access_15_groc_norm_wide,
            by = "from_id") |>
  left_join(access_15_groc_prop_wide,
            by = "from_id") |>
  left_join(access_15_groc_div,
            by = "from_id") |>
  rename("total_tot_copp_isc_groc" = "total_tot_copp_isc",
         "norm_tot_copp_isc_norm_groc" = "norm_tot_copp_isc_norm")

input.groc <- access_15_groc_wide |> ungroup() |> select(-c( #from_id, 
  #starts_with("total_"), 
  total_Care_Category, 
  norm_Care_Category, prop_Care_Category)) # Remove the non-normalized and proportional accessibility scores, as well as the Care Category character names. We are only clustering on the normalized accessibility scores (how many of this type of opportunity can you reach, normalized for the region) and the diversity (the evenness of type of opportunities that can be reached).

#rm(access_15_groc, access_15_groc_div, access_15_groc_norm_wide, access_15_groc_prop_wide, access_15_groc_total, access_15_groc_total_wide, access_15_groc_wide)
```

# Creating clusters (from parcels)

Now let's create the input data! To be used for the clustering analysis.
```{r}
input.data <- cbind(input.dep,input.err,input.groc,input.health)

input.data <- input.data |>
  mutate(
    prop_dep = ifelse(
    (total_tot_copp_isc_dep + total_tot_copp_isc_err + total_tot_copp_isc_groc + total_tot_copp_isc_health) > 0,
    total_tot_copp_isc_dep / (total_tot_copp_isc_dep + total_tot_copp_isc_err + total_tot_copp_isc_groc + total_tot_copp_isc_health), 0),

    prop_groc = ifelse(
    (total_tot_copp_isc_dep + total_tot_copp_isc_err + total_tot_copp_isc_groc + total_tot_copp_isc_health) > 0,
    total_tot_copp_isc_groc / (total_tot_copp_isc_dep + total_tot_copp_isc_err + total_tot_copp_isc_groc + total_tot_copp_isc_health), 0),

    prop_err = ifelse(
    (total_tot_copp_isc_dep + total_tot_copp_isc_err + total_tot_copp_isc_groc + total_tot_copp_isc_health) > 0,
    total_tot_copp_isc_err / (total_tot_copp_isc_dep + total_tot_copp_isc_err + total_tot_copp_isc_groc + total_tot_copp_isc_health), 0),

    prop_health = ifelse(
    (total_tot_copp_isc_dep + total_tot_copp_isc_err + total_tot_copp_isc_groc + total_tot_copp_isc_health) > 0,
    total_tot_copp_isc_health / (total_tot_copp_isc_dep + total_tot_copp_isc_err + total_tot_copp_isc_groc + total_tot_copp_isc_health), 0),

    div_allcats = -(prop_dep*log(prop_dep) + prop_groc*log(prop_groc) +
                     prop_err*log(prop_err) + prop_health*log(prop_health))/log(4),
    div_allcats = ifelse(is.na(div_allcats) | is.nan(div_allcats), 0, div_allcats)

         )

input.data <- input.data |> select(-c( from_id, 
                                       starts_with("total_"), 
                                       starts_with("prop_"), 
                                       "div_err","div_groc","div_health","div_dep", 
                                       "norm_tot_copp_isc_norm_dep", "norm_tot_copp_isc_norm_err", "norm_tot_copp_isc_norm_groc", "norm_tot_copp_isc_norm_health")) 
```

Given these normalized attributes, a SOM can be trained using the function `trainSOM` from `SOMbrero`. Only one argument is needed, the input data `x.data`. Other arguments can be assigned by default. In what follows the dimension of the map is given as 5-by-5, and I request that 100 intermediate steps be saved. 
```{r}
set.seed(2024)

# run the SOM algorithm, saving 10 steps
# access.som <- trainSOM(x.data = input.data,
#                        nb.save=10, 
#                        #maxit=2000, 
#                        scaling="none", 
#                        radius.type="letremy", 
#                        topo="hexagonal",
#                        dimension=c(43,43), #notably, I tested grid sizes larger then this -- https://stackoverflow.com/questions/19163214/kohonen-self-organizing-maps-determining-the-number-of-neurons-and-grid-size
#                       verbose = TRUE)
# 
# 
# plot(access.som, #make sure very few nodes have no observations
#      what = "obs",
#      type = "hitmap")
# 
# table(access.som$clustering) #make sure very few nodes have no observations
# 
# save(access.som, file = "access_som.RData")
# # 43*43 = 1849
# # 5*sqrt(143882) = 1896.589

access.som_small <- trainSOM(x.data = input.data,
                             #topo="hexagonal",
                             #dimension=c(5,5),
                             verbose = TRUE)
```

Plot super clusters and variance not explained:
```{r, eval=FALSE}
access.sc <- superClass(access.som_small,
                        k = 7)
# summary(access.sc)

png(file= file.path(here("figures/all_access_sc_plot_small.png")), width= 2404, height= 1600, units="px", res=300)

plot(access.sc, 
     type= c("dendrogram"),
     #col = c("#FC8D59","#D73027","#FFFFBF", "#FEE08B","#91CF60","#D9EF8B", "#1A9850"),
     plot.var = TRUE,
     show.names = TRUE,
     main = " ",
     ylab="Dissimilarity index",
     xlab= " "
     )
dev.off()

# Not caring - not complete (C2)
# Somewhat caring - not complete (C1)
# Somewhat caring - complete (C4)
# Caring - somewhat complete (C3)
# Caring - very complete (C6)
# Very caring - complete (C5)
# Very caring - very complete (C7)
```
```{r}
#par(mfrow=c(3,2))
plot(access.som_small, 
     what = "obs", 
     type = "color", 
     variable = 14, 
     show.names=TRUE, 
     main="norm_Dependent_centric")
plot(access.som_small, 
     what = "obs", 
     type = "color", 
     variable = 17, 
     show.names = TRUE, 
     main = "norm_Errand_centric")
# plot(access.som_small, 
#      what = "obs", 
#      type = "color", 
#      variable = 3,
#      show.names = TRUE, 
#      main="norm_Grocery_centric")
# plot(access.som_small, 
#      what="obs", 
#      type="color", 
#      variable=4, 
#      show.names = TRUE, 
#      main="norm_Health_centric")
# plot(access.som_small, 
#      what="obs", 
#      type="color",
#      variable = 9, 
#      show.names = TRUE, 
#      main="Diversity")

access.som_small$data |> colnames()
```

## Exploring results

First append the results of the cluster and super cluster analysis to dataframe:
```{r}
access_15_wide <- cbind(
  access_15_groc_wide |> ungroup() |> select(-c(starts_with("prop_"),
                                                total_Care_Category, 
                                                norm_Care_Category,
                                                prop_Care_Category)),
  access_15_dep_wide |> ungroup() |> select(-c(from_id,
                                                starts_with("prop_"),
                                                total_Care_Category, 
                                                norm_Care_Category,
                                                prop_Care_Category)),
  access_15_health_wide |> ungroup() |> select(-c(from_id,
                                                starts_with("prop_"),
                                                total_Care_Category, 
                                                norm_Care_Category,
                                                prop_Care_Category)),
  access_15_err_wide |> ungroup() |> select(-c(from_id,
                                                starts_with("prop_"),
                                                total_Care_Category, 
                                                norm_Care_Category,
                                                prop_Care_Category)),
  data.frame(div_allcats = input.data$div_allcats))

access_15_wide$cluster <- access.som_small$clustering
access_15_wide$scluster <- access.sc$cluster[access.som_small$clustering]
#save(access_15_wide, file = "access_15_wide.RData")
```

```{r}
access_15_wide |> select(total_tot_copp_isc_groc,total_tot_copp_isc_dep,total_tot_copp_isc_health,total_tot_copp_isc_err,div_allcats) |> summary()
access_15_wide |> group_by(scluster) |> summarise(total_tot_copp_isc_groc = median(total_tot_copp_isc_groc),
                                                  total_tot_copp_isc_dep = median(total_tot_copp_isc_dep),
                                                  total_tot_copp_isc_health = median(total_tot_copp_isc_health),
                                                  total_tot_copp_isc_err = median(total_tot_copp_isc_err),
                                                  div_allcats = median(div_allcats),
                                                  pc = n()/143882)
```

How many parcels are in each cluster? with their %

The low diversities - low accessibility
C7 - the WORST - below Q3 in everything

Mixed diversities - Higher accessibility for some stuff
C4 - above median dependent-centric - but low diversity and low everything else (THESE AREAS HAVE POTENTIAL! the need more GROC, HEALTH, and ERRs) 
C5 - like medium - about median everything with slightly lower HIGH diversity
C3 - exceptionally high dependent-centric, and high everything else but low diversity

High diversities and high accessibility
C6 - medium - about median everything with the highest diversity
C2 - better than medium - slightly above median everything with high diversity
C1 - the best - above Q3 everything

THE LABELS:
Not caring - not complete (C7)
Somewhat caring - not complete (C4)
Somewhat caring - complete (C5)
Caring - somewhat complete (C3)
Caring - very complete (C6)
Very caring - complete (C2)
Very caring - very complete (C1)

## Profiling the super clusters

One possibility is to analyze the results of the SOM algorithm by means of a classification technique. This is illustrated here by means of a decision tree. A decision tree is a supervised learning technique, which in this case can use the results of the clustering algorithm as labels, and the input variables as features. If the clustering algorithm performed well, class membership must be relatively homogeneous.

Select data for classification:
```{r}
df <- access_15_wide |> ungroup() |>
  select(-c(cluster,
            starts_with("prop"))) |>
  rename(Convience = "total_Convenience Store",
         Grocer ="total_Grocer",
         Rec = "total_Community or Recreation Centre",
         Daycare = "total_Daycare or EarlyON",
         LTC = "total_Long_Term Care or Retirement Home",
         Park = "total_Park",
         School = "total_School",
         Senior = "total_Senior Centre",
         Dentist = "total_Dentist",
         Hos_clinc = "total_Hospital or Clinic",
         Pharmacy = "total_Pharmacy",
         Bank_ATM = "total_Bank or ATM",
         Library = "total_Library",
         PostOff = "total_Post Office",

         norm_Convience = "norm_Convenience Store",
         norm_Grocer ="norm_Grocer",
         norm_Rec = "norm_Community or Recreation Centre",
         norm_Daycare = "norm_Daycare or EarlyON",
         norm_LTC = "norm_Long_Term Care or Retirement Home",
         norm_Park = "norm_Park",
         norm_School = "norm_School",
         norm_Senior = "norm_Senior Centre",
         norm_Dentist = "norm_Dentist",
         norm_Hos_clinc = "norm_Hospital or Clinic",
         norm_Pharmacy = "norm_Pharmacy",
         norm_Bank_ATM = "norm_Bank or ATM",
         norm_Library = "norm_Library",
         norm_PostOff = "norm_Post Office") |>
  mutate(scluster = factor(scluster),
         div_groc = round(div_groc, digits = 3),
         div_dep = round(div_dep, digits = 3),
         div_health = round(div_health, digits = 3),
         div_err = round(div_err, digits = 3),
         div_allcats = round(div_allcats, digits = 3))

df <- df |> 
  mutate(Total_copp_groc = Convience +Grocer,
         Total_copp_dep = Rec + Daycare + LTC + Park + School + Senior,
         Total_copp_health = Hos_clinc+Dentist,
         Total_copp_err = PostOff +Library+Bank_ATM + Pharmacy)
```

Train a decision tree using the input SOM input variables:
```{r}
# classmod.som <- rpart(data = df,
#                       formula = scluster ~ (Convience + Grocer + div_groc +
#                                               Rec + Daycare + LTC + Park + School + Senior + div_dep +
#                                               Dentist + Hos_clinc + div_health +
#                                               PostOff +Library+Bank_ATM + Pharmacy + div_err))
classmod.som <- rpart(data = df,
                      formula = scluster ~ (norm_tot_copp_isc_norm_groc + norm_Convience + norm_Grocer + #div_allcats +
                                              norm_tot_copp_isc_norm_dep + norm_Rec + norm_Daycare + norm_LTC + norm_Park + norm_School + norm_Senior + #div_dep +
                                              norm_tot_copp_isc_norm_health + norm_Pharmacy + norm_Dentist + norm_Hos_clinc + #div_health +
                                              norm_tot_copp_isc_norm_err + norm_PostOff + norm_Library+ norm_Bank_ATM + div_allcats#+ div_err
                                            ))

# Predict the labels and add them to the dataset
predicted_labels <- predict(classmod.som, df) |> as.data.frame()

df$SC1 <- predicted_labels$`1`
df$SC2 <- predicted_labels$`2`
df$SC3 <- predicted_labels$`3`
df$SC4 <- predicted_labels$`4`
df$SC5 <- predicted_labels$`5`
df$SC6 <- predicted_labels$`6`
df$SC7 <- predicted_labels$`7`
# df$SC8 <- predicted_labels$`8`
# df$SC9 <- predicted_labels$`9`

```

Trying ggparty to visualise, but first convert the rpart object to a party object:
```{r}
classmod.sompar <- as.party(classmod.som)

#Autoplot as ggplot:
autoplot(classmod.sompar)
```

Frequency of parcels in the super clusters:
```{r}
freq_sc <- ggplot(access_15_wide |> 
         mutate(scluster = factor(scluster))) +
  geom_bar(aes(x = "", 
               group = scluster, 
               fill = scluster)) +
  scale_fill_discrete(name = "Supercluster") #+
  #theme_void() +
  #guides(fill = guide_legend(nrow = 1)) +
  #theme(legend.position = "bottom")

#with new names and colours
freq_sc_renamed <- ggplot(access_15_wide |> 
         mutate(scluster = factor(scluster))) +
  geom_bar(aes(x = "", 
               group = scluster, 
               fill = scluster)) +
  scale_fill_manual(name = "Neighbourhood typologies",
                    values = c( "#1A9850","#91CF60", "#D9EF8B", "#FFFFBF","#FEE08B","#FC8D59","#D73027"),
                    labels = c("Not caring - not complete (C2)",
                               "Somewhat caring - not complete (C1)",
                               "Somewhat caring - complete (C4)",
                               "Caring - somewhat complete (C3)",
                               "Caring - very complete (C6)",
                               "Very caring - complete (C5)",
                               "Very caring - very complete (C7)"))

```

Extract the legend:
```{r}
legd <- cowplot::get_legend(freq_sc)

legd
```

Make a prettier plot:
```{r}
party_plot <- ggparty(classmod.sompar) +
  geom_edge() +
  geom_edge_label(size = 1.6) +
  geom_node_label(aes(label = splitvar),
                  id = "inner", 
                  size = 2, nudge_y = -0.022) +
  # pass list to gglist containing all ggplot components we want to plot for each
  # (default: terminal) node
  geom_node_plot(gglist = list(geom_bar(aes(x = "", fill = scluster),
                                        position = position_fill()),
                               xlab(""),
                               ylab(""),
                               theme_void()
                               #theme(axis.text = element_blank())
                               ))
party_plot
```

Plot the tree:
```{r}
png(file= file.path(here("figures/decision_tree_profiles_sc.png")),width= 2404, height= 1600, units="px",  res=300)
plot_grid(party_plot, legd, nrow = 1, rel_widths = c(1, 0.2))
dev.off()
```

## Maps
OKAY! SO LET"S MAKE SOME VISUAL PLOTS:

```{r}
df_DA <- df |> left_join(access_15 |> dplyr::select(c("from_id","GeoUID")), by = "from_id") |> 
  group_by(GeoUID) |> 
  summarise(
    SC1 = median(SC1),
    SC2 = median(SC2),
    SC3 = median(SC3),
    SC4 = median(SC4),
    SC5 = median(SC5),
    SC6 = median(SC6),
    SC7 = median(SC7),
    parcel_count = n_distinct(from_id),
    max_label_val = pmax(SC1, SC2, SC3, SC4, SC5, SC6, SC7),
    div_allcats = median(div_allcats),
    Total_copp_groc = median(Total_copp_groc),
    Total_copp_dep = median(Total_copp_dep),
    Total_copp_health = median(Total_copp_health),
    Total_copp_err = median(Total_copp_err)) |>
  mutate(max_label = case_when(max_label_val == SC1 ~ "a. Very caring - very complete (SC1)",
                               max_label_val == SC2 ~ "b. Very caring - mostly complete (SC2)",
                               max_label_val == SC3 ~ "d. Somewhat caring - mostly complete (SC3)",
                               max_label_val == SC4 ~ "c. Very caring - mostly complete (SC4)",
                               max_label_val == SC5 ~ "j. not caring - somewhat complete (SC5)",
                               max_label_val == SC6 ~ "e. Somewhat caring - mostly complete (SC6)",
                               max_label_val == SC7 ~ "f. not caring - not complete (SC7)"))

                    labels = c("Not caring - not complete (C2)",
                               "Somewhat caring - not complete (C1)",
                               "Somewhat caring - complete (C4)",
                               "Caring - somewhat complete (C3)",
                               "Caring - very complete (C6)",
                               "Very caring - complete (C5)",
                               "Very caring - very complete (C7)"))

df_DA <- df_DA |> left_join(HAM_census_21, by=c("GeoUID")) |>
  rename( 
    LICO =`v_CA21_1085: Prevalence of low income based on the Low-income cut-offs, after tax (LICO-AT) (%)`,
            Income_bottom_n = `v_CA21_1103: In bottom half of the distribution`,
            Income_top_n =`v_CA21_1121: In top half of the distribution`,
            Age_0to14_n = `v_CA21_11: 0 to 14 years`,
            Age_15to64_n = `v_CA21_68: 15 to 64 years`,   
            Age_65plus_n = `v_CA21_251: 65 years and over`,
            One_parent_families_n = `v_CA21_507: Total one-parent families`,
            Couple_families_n = `v_CA21_500: Total couple families`) |>
  mutate(Populationparcel = Population/parcel_count,
         Income_bottom = Income_bottom_n / (Income_bottom_n + Income_top_n),
         Age_0to14 = Age_0to14_n / (Age_0to14_n + Age_15to64_n + Age_65plus_n),
         One_parent_families = One_parent_families_n / (One_parent_families_n + Couple_families_n)) |>
  st_as_sf() |> st_make_valid()

df_DA$parcel_count |> sum() #a check
```
Loading boundaries for mapping:
```{r}
Community_Boundaries <- st_read("data-raw/boundaries/Community_Boundaries.shp") |> mutate(COMMUNITY_ = ifelse(COMMUNITY_ == "Hamilton", "Hamilton-Central", COMMUNITY_))|> st_transform(crs=4326)
City_Boundary <- st_read("data-raw/boundaries/City_Boundary.shp") |> st_transform(crs=4326)
hydro_p_LakeOntario <- st_read("data-raw/boundaries/hydro_p_LakeOntario.shp") |> st_transform(crs=4326)
ham_bay <- st_read("data-raw/boundaries/Waterbodies.shp") |> st_transform(crs=4326) |> filter(FEATURE_TY == "Lake")
ham_bay_cropped<-st_crop(ham_bay$geometry, Community_Boundaries$geometry)
hydro_p_LakeOntario_cropped<-st_crop(hydro_p_LakeOntario$geometry, Community_Boundaries$geometry)
```

```{r, eval=FALSE}
dep_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(Community_Boundaries) + tm_polygons(col= "white", border.alpha=0)+
  tm_shape(df_DA |> filter(Total_copp_dep >0)) +
  tm_polygons("Total_copp_dep",
              border.alpha = 0,
              n=10,
              palette = "Oranges",
              title = "Dependent-\ncentric") +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4, 
            legend.title.size = 0.8,
            legend.text.size = 0.6,
            bg.color = "grey",
            legend.position = c("right","top"))

health_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
    tm_shape(Community_Boundaries) + tm_polygons(col= "white", border.alpha=0)+
  tm_shape(df_DA|> filter(Total_copp_health >0)) +
  tm_polygons("Total_copp_health",
              border.alpha = 0,
              n=10,
              palette = "Purples",
              title = "Health-\ncentric",) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4,
            legend.title.size = 0.8,
            legend.text.size = 0.6,
            bg.color = "grey",
            legend.position = c("right","top"))

groc_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
    tm_shape(Community_Boundaries) + tm_polygons(col= "white", border.alpha=0)+
  tm_shape(df_DA|> filter(Total_copp_groc >0)) +
  tm_polygons("Total_copp_groc",
              border.alpha = 0,
              n=10,
              palette = "Greens",
              title = "Grocery-\ncentric",) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4,
            legend.title.size = 0.8,
            legend.text.size = 0.6,
            bg.color = "grey",
            legend.position = c("right","top"))

err_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
    tm_shape(Community_Boundaries) + tm_polygons(col= "white", border.alpha=0)+
  tm_shape(df_DA|> filter(Total_copp_err >0)) +
  tm_polygons("Total_copp_err",
              border.alpha = 0,
              n=10,
              palette = "Reds",
              title = "Errand-\ncentric",) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4,
            legend.title.size = 0.8,
            legend.text.size = 0.6,
            bg.color = "grey",
            legend.position = c("right","top"))

percats_copp_plot <- tmap_arrange(dep_copp_plot, err_copp_plot, groc_copp_plot, health_copp_plot, nrow=2)

tmap_save(percats_copp_plot, 
          file= file.path(here("figures/percats_copp_plot.png")),
          dpi=300)
```

```{r, eval=FALSE}
all_cats_entropy_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
      tm_shape(Community_Boundaries) + tm_polygons(col= "white", border.alpha=0)+
  tm_shape(df_DA) +
  tm_polygons("div_allcats",
              border.alpha = 0.1,
              n=10, 
              #label = c("[0 to 0.49) Not complete", "[0.49 to 0.85) Complete", "[0.85 to 1] Very complete"),
              palette = "Greys",
              title = "Diversity") +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
    tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4,
            legend.title.size = 0.8,
            legend.text.size = 0.6,
            bg.color = "grey",
            legend.position = c("right","top"))

tmap_save(all_cats_entropy_copp_plot, file= file.path(here("figures/all_cats_entropy_copp_plot.png")), dpi=300)
```

```{r, eval=FALSE}
total_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(df_DA) +
  tm_polygons("Total_copp_groc",
              border.alpha = 0.2,
              #style = "quantile", n=4,
              palette = "Purples",
              title = "Total access to care destinations",) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4,
            bg.color = "grey",
            legend.position = c("right","top"))

total_copp_plot
```

```{r, eval=FALSE}
SClabel_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(df_DA) +
  tm_polygons("max_label",
              border.alpha = 0.2,
              # palette = c("#44ce1b","#bbdb44", "#f7e379", "#f2a134","#e51f1f"),
              # labels = (c("Very caring - very complete",
              #             "Caring - Complete",
              # "Somewhat caring - complete",
              # "Somewhat caring - somewhat complete", 
              # "Somewhat caring - somewhat complete")),
              title = "Neighbourhood typology") +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4,
            bg.color = "grey",
            legend.position = c("right","top"))
SClabel_copp_plot

SCLab_and_total_copp_plot <- tmap_arrange(SClabel_copp_plot, total_copp_plot, nrow=2)
```

```{r, eval=FALSE}
SCLab_and_total_copp_plot
```

```{r, eval=FALSE}
tmap_save(SCLab_and_total_copp_plot, 
          file= file.path(here("figures/SCLab_and_total_copp_plot.png")),
          dpi=300)

tmap_save(SClabel_copp_plot, 
          file= file.path(here("figures/SClabel_copp_plot.png")),
          dpi=300)
```

## Demographic profile of clusters

Select data for classification:
```{r}

df <- access_15_wide |> left_join(access_15_aggregated, by = "from_id") |> 
  transmute(scluster = factor(scluster),
            Income_bottom = `v_CA21_1103: In bottom half of the distribution`/(`v_CA21_1103: In bottom half of the distribution` + `v_CA21_1121: In top half of the distribution`),
            LICO = `v_CA21_1085: Prevalence of low income based on the Low-income cut-offs, after tax (LICO-AT) (%)`,
            Avg_numb_of_kids = `v_CA21_498: Average number of children in census families with children`,
            Not_in_labforce = `v_CA21_6504: Not in the labour force`/(`v_CA21_6498: Employed` + `v_CA21_6501: Unemployed` + `v_CA21_6504: Not in the labour force`),
            Income_gini = `v_CA21_1142: Gini index on adjusted household after-tax income`,
            Vis_minority = `v_CA21_4875: Total visible minority population`/ (`v_CA21_4914: Not a visible minority`+ `v_CA21_4875: Total visible minority population`),
            Age_0to14 =(`v_CA21_11: 0 to 14 years`) / (`v_CA21_11: 0 to 14 years` + `v_CA21_68: 15 to 64 years` + `v_CA21_251: 65 years and over`),
            One_parent_families = `v_CA21_507: Total one-parent families` / (`v_CA21_507: Total one-parent families` + `v_CA21_500: Total couple families`),
            walk_towork =  `v_CA21_7647: Walked` / (`v_CA21_7647: Walked` + `v_CA21_7635: Car, truck or van` + `v_CA21_7644: Public transit` + `v_CA21_7650: Bicycle` + `v_CA21_7653: Other method` ),
            Owner_hld_CHN = `v_CA21_4308: % in core housing need (57)`,
            Hld_ownership = `v_CA21_4306: % of owner households with a mortgage (58)`,
            Tenant_subsi_hld = `v_CA21_4314: % of tenant households in subsidized housing (61)`,
            Tenant_hld_CHH =`v_CA21_4316: % in core housing need (57)`,
            Population,
            Low_or_no_deg = (`v_CA21_5868: No certificate, diploma or degree` + `v_CA21_5871: High (secondary) school diploma or equivalency certificate`) / (`v_CA21_5868: No certificate, diploma or degree` + `v_CA21_5871: High (secondary) school diploma or equivalency certificate` + `v_CA21_5874: Postsecondary certificate, diploma or degree`))

# examining correlations
cor(df |> select(-c("scluster")), use = "complete.obs")

#Train a decision tree using the data (renamed!):
classdemo.som <- rpart(data = df,
                      formula = scluster ~ (Income_bottom + Age_0to14 + One_parent_families + Vis_minority + Not_in_labforce+ Income_gini+walk_towork+Owner_hld_CHN+Hld_ownership+Tenant_subsi_hld+Tenant_hld_CHH+LICO),
                      weights = Population
                      )
# # Visualize the decision tree:
# rpart.plot(classdemo.som, cex = 0.5, type = 0, varlen = 8)
```

```{r}
# convert the rpart object to a party object:
classdemo.sompar <- as.party(classdemo.som)

#Make a prettier plot:
party_plot <- ggparty(classdemo.sompar) +
  geom_edge() +
  #geom_edge_label(size = 1.8) +
  geom_edge_label(mapping = aes(label = paste(substr(breaks_label, start = 1, stop = 15))), size = 3.5) +
  geom_node_splitvar(id = "inner", size = 4) +
  #geom_edge_label(mapping = aes(label = prettyNum(breaks_label, digits = 1)), size = 1.8) +

  # # pass list to gglist containing all ggplot components we want to plot for each
  # # (default: terminal) node
  geom_node_plot(gglist = list(geom_bar(aes(x = "", fill = scluster),
                                        position = position_fill()),
                               #xlab(""),
                               #ylab(""),
                               theme_void(),
                               scale_fill_manual(values = c("#bbdb44","#44ce1b",  "#f2a134", "#f7e379","#e51f1f")),
                               theme(axis.text = element_blank())))

legd_renamed <- cowplot::get_legend(freq_sc_renamed)
party_plot
```

```{r, eval=FALSE}
#plot and save
# NOTE: edited this in paint.. be careful saving over
png(file= file.path(here("figures/access_sc_profiles_renamed.png")),width= 2600, height= 1600, units="px",  res=300)
plot_grid(party_plot, legd_renamed, nrow = 1, rel_widths = c(1, 0.3))
dev.off()
```

This is interesting! So for parcels with less than 45% of bottom-income people (richer) -- parcels with higher proportion of 1 parent households (13% and above) are more likely to have bad care/incomplete access (c3) than _really_ bad care/very incomplete access (c5).

For parcels with higher than 45% of bottom-income (poorer), those with high a high proportion of children (age 0 to 14) are likely to have bad care/incomplete acces (c3). But those with a lower proportion of children - especially 9.8% LICO and above are going to have _super high_ care/complete (c1) or under 9.8% LICO will have _high_ care/complete (c2)
For less than 63%, it's a different story. If between 45% and 63%, they are described by cluster 1 (caring and complete).

### 2024-09-27 paper/Manuscript.qmd
Still working on the draft...

@@
 ---
-title: "Towards caring 15-minute neighbourhoods"
+title: "Identifying 15-minute neighbourhoods with complete caring potential"
 author:
@@
 library(tmaptools)
+library(gt)
 ```
@@
 load(glue::glue(here::here(), "/data/care_dest.rda"))
+load(glue::glue(here::here(), "/data/access_15_wide.RData"))
 ```
@@
 This study adopts the 15-minute neighbourhood concept through the lens of "Mobility of Care", a framework that foregrounds travel to care destinations, travel done predominately by women, in contrast to the better-studied travel to employment and leisure.
-While the 15-minute neighbourhood concept is flexible enough to consider all destination types, gendered examinations are lacking in the literature and no research to date has focused explicitly on care destinations.
-In this study, we address these gaps by identifying which areas in Hamilton, Canada are ‘caring 15-minute neighbourhoods'.
+While the 15-minute neighbourhood concept is flexible enough to consider all destination types, gendered examinations are relatively lacking in the literature and no research to date has focused explicitly on care destinations.
+In this study, we address these gaps by identifying which areas in Hamilton, Canada have the potential to be ‘caring 15-minute neighbourhoods'.
 To do so, a database of care destinations was compiled and used to estimate both the number (cumulative opportunities indicator) and diversity of mix (entropy measure) of destinations reachable within a 15-minute walk from all residential parcels.
-Parcel-level results were classified into 'caring' (quantity of reachable opportunities) and 'complete' (diversity of care destinations) neighbourhood typologies of varying degrees using the Self-Organizing Maps (SOM) machine learning technique.
-Inequalities in the relationship between the identified caring typologies and social-economic indicators were then explored using SOM.
+Parcel-level results were classified into 'caring' (quantity of reachable opportunities) and 'complete' (diversity of care destinations) neighbourhood typologies of varying degrees using machine learning techniques (Self-Organizing Maps (SOM) and hierarchical clustering).
+Inequalities in the relationship between the identified caring typologies and social-economic indicators of who currently reside in the neighbourhoods were then explored.
 Our results suggest the majority of 'caring and complete' neighbourhoods are within the urban core where a higher proportion of lower-income and single-parent households reside.
 Further, certain neighbourhoods in suburban and rural communities are 'caring and somewhat complete', showing the potential to become 'complete' through possible urban policy intervention.
-Taken together, this study offers a theoretical bridge to connect Mobility of Care and the 15-minute neighbourhoods concept, a methodology to assess neighbourhoods on a 'caring and complete' continuum, and discussion of relevant gender mainstreaming policy interventions.
+Taken together, this study offers a theoretical bridge to connect Mobility of Care and the 15-minute neighbourhoods concept, a data-driven classification and clustering methodology to assess neighbourhoods on a 'caring and complete' continuum, and discussion of relevant gender mainstreaming policy interventions.
 
@@
 
-The "15-Minute City" has been adopted by leaders as a way to promote human-scaled cities.
-As defined in @moreno_introducing_2021, the 15-Minute City is a urban planning model based in "chrono-urbanism", a theory that emphasizes the positive impact on quality of urban life when urban space becomes multi-rythmic [related to chronotopy @mulicekUrbanRhythmsChronotopic2015].
-Multi-rhythmic urban space in opposition to industrial Fordism and single-use zoning [@mulicekUrbanRhythmsChronotopic2015; @morenoVilleQuartHeure2016]; the segregation of urban time and space for individual uses and mobility.
-The 15-Minute City is one that accommodates the ability to reach essential destinations such as shops, work, and healthcare within a walkable radius.
+The "15-Minute City" has been recently adopted by leaders as a way to promote human-scaled cities (e.g., Paris mayor Anne Hidalgo proposed a 15-minute city plan in her 2020 re-election campaign, 15-minute walkable neighbourhood guidance in Shanghai, China [@weng15minuteWalkableNeighborhoods2019]), As defined in @moreno_introducing_2021, the 15-Minute City is a urban planning model based in chrono-urbanism, a theory that emphasizes the positive impact on quality of urban life when urban space becomes multi-rythmic, in opposition to industrial Fordism, single-use zoning, and the segregation of urban time and space for individual uses and mobility [@mulicekUrbanRhythmsChronotopic2015; @morenoVilleQuartHeure2016].
+The 15-Minute City is one that accommodates the ability to reach essential destinations within a walkable, bikeable or transit supportive radius.
 It would allow individuals to reclaim time spent on car mobility and give way to sustainable modes; prompting urban spaces that are responsive to human needs and environmental sensibilities [@Allam2022].
-The 15-Minute City generated much interest in planning practice and public discourse, especially after Paris mayor Anne Hidalgo proposed a 15-minute city plan in her 2020 re-election campaign.
-Though the term 15-Minute City is relatively new, the core of the concept is not.
 
-Literature has drawn parallels from the 15-Minute City to Perry's debated Neighbourhood Unit Concept (NUC) from the 1930s [@kissfazekasCircleParadigms15minute2022]; a 'socio-spatial' normative scheme widely adopted by government agencies (and the real-estate community) in the Western world after the Second World War [@talenSocialSciencePlanned2017; @solow1969concept].
+Though the term 15-Minute City is relatively new, neighbourhood planning practice aiming to foster community and local travel to amenities is not.
+As recent examples, planned neighbourhoods dominated post-WWII built urban form: the Neighbourhood Unit Concept in the western world [@brodyNeighbourhoodUnitConcept2013] and a parallel concept of the “mikrorayons” neighbourhoods in the Soviet Union  [@kissfazekasCircleParadigms15minute2022].
+These planning forms have also been extensively critiqued: as summarized in @talenSocialSciencePlanned2017, urban planners struggled to determine suitably universal design aspects such as neighborhood size, the type and quality of amenities, and local and regional connectivity.
+Social scientists also emphasized that top-down neighbourhood planning approaches overestimated the built environment's influence on social life [@talenSocialSciencePlanned2017].
+Neighbourhood planning approaches such as the 15-Minute City concept are being re-embraced, but from bottom-up perspective.
+However how to equitably prescribe urban form that 'enables' contact with social opportunities (instead of 'engendering it' from the top-down) is still under debate.
+For instance, questions such as _what_ are destinations that matter, by what mode and travel time threshold, and _for whom_ remain under discussion among proximity-based planners [@silvaProximitycentredAccessibilityConceptual2023; @silvaRegionalAccessibilityUndermining2022].
+
+To add clarity to these questions surrounding the 15-Minute City concept, in this work we aim to build a theoretical bridge to "Mobility of Care" as introduced in @sanchezdemadariagaMobilityCareIntroducing2013.
+Mobility of Care is especially relevant to the 15-Minute City, as it is a purpose of travel that is necessary for all and often occurs at the local level.
+Namely, Mobility of Care refers to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands.
+While decades of research have examined types of household-serving trips (e.g., shopping trips, escorting to school), @sanchezdemadariagaMobilityCareIntroducing2013 was the first to consider them all as one category.
+In doing so, Mobility of Care is shown to be a significant proportion of daily travel, characterizing the trip purpose of approximately 30% of adults’ trips [@sanchezdemadariagaMobilityCareIntroducing2013; @sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory].
+Notably, care trips take on tremendous gendered and socio-economic importance, as women, and especially those from lower income households, complete the majority of care trips [@ravensbergen2023exploratory].
+Despite its importance for all and in advancing gender equality, the examination of mobility of care and access to care destinations have been under-examined relative to employment destinations.
+This is especially pertinent in accessibility research that has largely focused on employment () and retail ().
+
+We aim to theoretically bridge Mobility of Care with the 15-Minute City concept to clarify questions surrounding _what destinations within a 15 minute walkable radius matter_.
+We hypothesis that _care_ destinations matter, they matter for all people but especially for women and lower-income women preforming care trips.
+In this vein, this work pairs this hypothesis with an exploratory clustering methodology to identify which neighbourhoods have the potential to support 15-minute access to care destinations in an empirical case study of Hamilton, Ontario, typical mid-sized Canadian city.
+More concretely, the research objectives are two-fold: after a review of the 15-Minute City concept and associated measurement tools, this work first introduces the Mobility of Care conceptualisation to re-imagine local amenities importance within the 15-Minute City concept, and second to apply a data-driven classification and clustering machine learning methodology to support the bottom-up examination of _15-minute caring neighbourhoods_  through an empirical examination of Hamilton.
+
+# Review of neighbourhood planning literature
+
+## From the 15-Minute City to the NUC
+
+In the last decades, the need for more sustainable, healthier, and liveable cities have received more attention.
+Planners and decision-makers have proposed alternatives to mobility-based neighbourhood planning to one centered on proximity to urban functions [@pozoukidou15MinuteCityDecomposing2021].
+In this context, the 15-Minute City is now in the public spotlight [@logan_xminute_2022; @moreno_introducing_2021].
+In @moreno_introducing_2021 where the 15-Minute City concept was introduced, four universal dimensions are outlined: density (in terms of people per km$^2$), diversity (including mixed-land use and diversity of people), the temporal and spatial proximity to essential services, <!-- (residential areas, work, commercial areas, education centres, health facilities and other basic institutions)--> and digitalisation.
+The framework presented has been critised within academic and planning circles [@guzmanProximityEnoughCritical2024].
+Namely, critiques have centered on the concept's inadequecy in addressing pre-existing structural forces and individual charctertes that drive inequalities in who does and who could benefit more from such a paradigm [@dimarino15minuteCityConcept2023; @willberg15minuteCityAll2023].
+A 15-Minute City for all people is a utopic ideal, however how will existing inequalities in how people move through and benefit from the city be addressed.
+_Who is in the most need of the 15-Minute City before we all get there?_.
+Without directed answers to these context-sensitive questions makes this increasingly popular concept susceptible to becoming a place-branding slogan [@pozoukidou15MinuteCityDecomposing2021].
+
+Reflecting this flexible framework; cities that have adopted 15-Minute City concepts have done so using a diverse range of definitions and tools along with similar indistinct universal approaches.
+An early adopter was Portland, US in the Portland Plan [@city_of_portland_20minute_2010] of April 2012.
+It aims to foster an inclusive city development based on prosperity, education, health, and equity for a 30 year horizon.
+It promotes neighbourhood self-suffiency and connectivity to city centers and centres of employment.
+The progress report describes a high-level focus on equitable service delivery to all residents with equity concerning topics related to racial equity and people with disabilities [@portland_government_portland_2017]; similarly adopting an "all populations" approach.
+Later, other cities also adopted proximity-based goals into their plans, taking similar neutral entity approaches.
+For instance, in the plan that later catapulted the 15-Minute City into public discourse: Paris mayor Anne Hidalgo proposed a 15-minute city plan in her 2020 re-election campaign that foregrounded six social functions of the city that should be easily accessible from any location, namely: housing, work, health care, groceries, education and leisure [@ville_de_paris_paris_2022].
+These ideas inspired language in the agendas of other cities in the Western world such as Ottawa, in the Canadian context, who also adopted a 15-Minutes approach in their recent Official Plan [@ville_dottawa_quartier_2021].
+From the review of worldwide practice literature in @teixeiraClassifying15minuteCities2024, overall the 15-Minute City concept is in early phases of implementation around the world and the diverse range of definitions, strategies, and instruments present a significant knowledge and implementation gap: the 15-Minute City seems nice but _how do we get there and whom will benefit_?
+
+The past can shed light on cautionary tales.
+While the 15-Minute City concept is new, neighbourhood planning to improve society outcomes is not: in fact the literature has drawn parallels from the 15-Minute City to Clarence A.
+Perry's Neighbourhood Unit Concept (NUC) from the 1930s [@kissfazekasCircleParadigms15minute2022].
+The NUC is a socio-spatial normative scheme widely adopted by government agencies (and the real-estate community) in the Western world after the Second World War [@talenSocialSciencePlanned2017; @solow1969concept].
 Pairing well with the objectives of planning agencies at the time, Perry's NUC would allow for efficient mass-building of cellular units that prioritized the perceived functional needs of women and children: each unit providing proximate access to an elementary school and sufficent supporting community facilities, shopping, parks and housing [@talenSocialSciencePlanned2017; @brodyNeighbourhoodUnitConcept2013].
@@
 By other planners, the inability to agree on the specificity of the neighbourhood (i.e., population sized, and the type, quality, and quantity of amenity) as well as how neighbourhood units are connected to other neighbourhoods and the rest of the region.
-<!-- failing to combine work, home and cultre was a dilbirate outcome of state-sponsored capitalism X.
-As stated in [@talenSocialSciencePlanned2017],-->  In response to these criticisms, neighbourhood planning proponents have redefined deterministic terminology: prescribed physical forms is 'enabling' rather than engendering, and the provision of social contact 'opportunities' instead of determined social behaviour.
-The redefined bottom-up concept to community and neighbourhood planning has been adopted by New Urbanist from which the 15-Minute City stems.
-The 15-Minute City is created through bottom-up proximity-based processes of community planning [@kissfazekasCircleParadigms15minute2022].
+In response to these criticisms, neighbourhood planning proponents have redefined deterministic terminology: prescribed physical forms as 'enabling' rather than engendering, and the provision of social contact 'opportunities' instead of determined social behaviour.
+The re-defined bottom-up approach to community and neighbourhood planning has been adopted by New Urbanist from which the 15-Minute City stems [@kissfazekasCircleParadigms15minute2022].
+But with _what bottom-up processes?_; small scale experiments are being tested () with preliminary small-scacle results.
+These processes are only in early stages so critiques of social scientists and planners from decades past still persist.
 
-<!-- These critiques were supported by case studies of planned neighbourhoods as...
-lead to social segregation of classes, destination functions, and automobile use.
-From it's onset, it was an efficient way to develop housing after the Second World War.
-The emphasis of proximity to some 'daily essentials' but disconnection from specialised activites (e.g., work) paved the way for car-based subrbia.
-by capital of cities.
-Can the neighbourhood form might be seen as a vehicle for capitalist exploitation, gentrification and displacement???
-The place of neighbourhood planning is actively being debated: can it's social crtiques be addressed in the new form as a "15-Minute City"?
-This study seeks to contribute to this decades old debate.
--->
+_How can physical form be planned to enable an improved quality of life, and for whom?_  has also occupied urban planning research in recent years, exploring what are the impacts of built environment on residence's outcomes.
+For instance, the examination of low-income households access to transportation and their likelihood of gaining employment [@blumenbergDriveWorkRelationship2017; @bastiaanssenDoesBetterJob2022] or the relationship between children's access to public transit and participation in after-school activities [@palmRolePublicTransit2020].
+Relatedly, a new wave of researchers and practitioners focused on local and context-specific relationships in the proximity to destinations have also emerged [@silvaProximitycentredAccessibilityConceptual2023; @silvaRegionalAccessibilityUndermining2022].
+As reviewed  in the city plans that have adopted 15-Minute City approaches, the NUC's critisms, and recent proximity-based examinations, the question of _how to enable improved quality of life through urban built environment_ has shown to be highly context-senstive with the need for further investigation.
 
-However, critiques of social scientists and planners from decades past still persist.
-Can physical form be planned to enable an improved quality of life?
-This line of reasoning has occupied urban planning research in recent years: what are the impacts of built environment on residence's outcomes.
-A popular tool to analyse this relationship has often been accessibility analysis: proximity to X amenity results in X outcomes for a neighboourhood.
-Accessibility is a way of quantifying the ease of opportunity interaction from some point in space, this indicator is becoming increasingly important for planners in the pursuit of just and sustainable city planning [@vale2023accessibility; @handyAccessibilityIdeaWhose2020].
-As a amenity-provision concept of neighbourhood planning that seeks to enable an improvement on urban quality of life, the 15-Minute City lends well to analysis using accessibility tools.
+## Tools: accessibility methods, diversity measures and typology-classification
 
-Before discussing 'what outcome', we first ask 'what amenity' as the focus of this study.
-Often, accessibility analysis studies have focused on a single amenity type (X), often employment destinations and often focused on a scale beyond the neighbourhood.
-However, due to 15-Minute City's emphasis on proximate-level service provision, recent academic work on the topic has employed accessibility analysis to account for multiple 'local' amenities types, but seldom included employment.
-We argue that the inclusion of employment opportunities into the 15-Minute City conceptualisation is critical to it's definition.
-Why?
-We echo critics made by planners and social scientistics in decades past: neighbourhoods that did not mix work with service and culture within the neighbourhood, paved the way for car-based suburbia: cars allowed for the connection of neighbourhoods with their 'lower-order' amenities to areas with 'high-order' destinations like work (X).
-The lower- and higher- order framing of these destinations supported the continued segregation of urban life across many dimensions (X).
+In examining how to enable improved quality of life through urban built environment, an increasingly popular tool have been accessibility measures.
+They are a way to quantify the ease of reaching destinations from a given point in space and have been used to examine urban areas through just and sustainable city planning agendas [@vale2023accessibility; @handyAccessibilityIdeaWhose2020].
+As the 15-Minute City is an amenity-provision neighbourhood planning concept which aims to enable urban environment that improves life quality, its analysis lends well to the use of accessibility tools [@guzmanProximityEnoughCritical2024].
+Foreseeably, recent works have applied accessibility measures to investigate the 15-Minute City across various geographic scopes.
+For instance, in Naples, Italy [@gaglioneUrbanAccessibility15minute2022], Barcelona, Spain [@graellsCityCitiesMeasuring2021], Vancouver, Canada [@hosford15minuteCityReach2022], and in urban areas across Europe [@vale2023accessibility].
+The "cumulative opportunity" measure has been applied to many 15-Minute City accessibility examinations.
+This form of accessibility quantifies how many destinations can be reached from a point in space within a given travel time threshold, pairing well with examinations that have a normative travel time threshold, like the 15-Minute City.
+Furthermore, the cumulative opportunity measure is widely appreciated for its intuitive computation and is popular among practitioners [@handyAccessibilityIdeaWhose2020; @handyMeasuringAccessibilityExploration1997; @chengInvestigatingWalkingAccessibility2019].
+However, reflecting the diversity of the accessibility literature, accessibility measures other than the cumulative opportunity have also been applied [@guzmanProximityEnoughCritical2024].
 
-Combing all amenity types into the 15-Minute City concept is also a matter of gender equality.
-There is a branch of research that encompases this critique: Mobility of Care.
-This framework encourages the explicit consideration of destination types that serve unpaid labour (care destinations) into transport planning (as opposed to only considering destinations with paid work).
-Not considering these care destinations have travel time (and hence quality of life) implications on those who do care activites, commonly women and especially disadvantaged women.
-The consideration of _what_ amenity type is a gender issue.
+Another method which compliments the assessment of the 15-Minute City as it relates to the diversity of opportunity types, is entropy.
+The entropy measure has been widely used to characterise the diversity of land-use mix in early work by @frankLinkingObjectivelyMeasured2005 based on the species diversity Shannon-Wiener index that expresses relative evenness throughout a sequence (this measure among other species diversity indices of the time are reviewed in @whittaker1972evolution).
+It has also been used in social sciences to comprehend the occurrence of social events in a structured manner specially with regard to life trajectories [@ritschardSequenceAnalysisRelated2018].In the realm of urban studies and planning literature the malleable concept of entropy has been used to characterise the diversity in land-use mix [@ewingTravelBuiltEnvironment2010a], especially in the context of active modes [@luUrbanDensityDiversity2017; @mavoaIdentifyingAppropriateLanduse2018] and suburban sprawl [@randallGISBasedLand2015].
+As well, it has been used to understand mobility behaviour [@mcbride2020exploration; @montero2023applying; @montero2023role].
+Work that has used diversity indices alongside accessibility analysis are scarce, though examples of its use as parameters within accessibility scores in the case of employment opportunities [@chengMeasuringUrbanJob2013; @daiIncorporatingJobDiversity2018] and more recently to describe the diversity in transit facilities [@yinIncorporatingFacilityDiversity2024] are present.
+To the authors knowledge, the use of an entropy measure to reflect the diversity in care destination amenity types and in context within a 15-minute neighbourhood, has yet to be done.
+However, entropy measures thematically compliments the 15-Minute City.
 
-From this position, this study aims to holistically investigate the ease of interacting with proximate amenities (accessibility) from the perspective of the 15-Minute City and the Mobility of Care framework.
-As mentioned, accessibility analysis have typically focused on employment (X) and recent 15-Minute City flavoured analysis focus on 'lower-order' amentities.
-We focus on a case study of the mid-sized City of Hamilton, Canada, and aim to classify areas by what degree they may be considered a '15-Minute Neighbourhood" based on the relative quantity of amenity access and diversity of access.
-We offer a novel methodology of holistically considering all proximate-amenities, offer empirical findings, and contribute theoretically to the ongoing debate on the definition of a 15-Minute City by foregrounding care destinations.
+`r word_comment(Maybe Antonio to add? I need to fill this out a bit", " ")`
+In addition, a useful tool has been classification algorithms in the identification of spatial typologies and in profiling who may reside there.
+Those too have K means...
+Hierarchical...
+and SOM [@delmelleDifferentiatingPathwaysNeighborhood2017]...
+regionlization techniques such as spatially constrained multivariate clustering [@burkeGeospatialAnalysisFramework2022] Use of tree diagrams!
 
---------------
+SOM is a type of artificial neural network (ANN) trained using unsupervised learning that produces a two-dimensional representation of the input space.
+It applies competitive learning by way of a neighbourhood function thereby preserving the topological properties of the input space and producing combinations of the input space that we can then intuitively profile.
+It has been used widely in XX but scarcely in transportation related research [@victorianoTimeSpaceMoney2020]
 
+All-in-all, within the context of examining the 15-Minute City there has been a wide range of tools applied in the literature that appear appriorpiate.
+However, the question of _15-minute cities for whom_ requires further examination.
+In this way, this work connects the 15-Minute City to another normative conceptualisation: Mobility of Care.
+
+## Mobility of Care and feminist 15-minute neighbourhoods
+
+As the 15-Minute City is in early stages of conceptualisation; an alternative way to conceive of _what_ urban functions matter could be through a feminist lens.
+Instead of focusing on _all_ destinations, it may be useful to examine destinations involved in caring activities.
 Caring activities meet the physical, psychological, and emotional needs of others, and are one of the most essential and fundamental activities in society [@ilo_care_2018].
@@
 According to @ilo_care_2018, women and girls perform more than three-quarters of the total amount of unpaid care, a gender gap that varies geographically [@RN5].
+This unequal share of caring responsibilities leads to multifaceted gendered differences: in career development, profession selection, contract type, pay gap, and time poverty, as recognized by various international organisations [@EIGE2016; @ilo_care_2018].
+In terms of spatial and transportation planning, almost one third of daily trips are for care purposes [@sanchezdemadariagaMobilityCareIntroducing2013; @sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory].
+From this research motivation, Sánchez de Madariaga coined the term Mobility of Care in 2013 to refer to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands [@sanchezdemadariagaMobilityCareIntroducing2013].
+While decades of research had examined these unique household-serving trips, her seminal work was the first to consider all these trips as one category and demonstrate how mobility of care is a significant proportion of daily travel.
 
-This unequal share of caring responsibilities leads to multifaceted gendered differences: in career development, profession selection, contract type, pay gap, and time poverty, as recognized by various international organisations [@EIGE2016; @ilo_care_2018].
-In fact, early feminists in the United States had already observed that domestic labor was the single most relevant cause of women’s inequality in society [@gordon_womans_1976; @haydenGrandDomesticRevolution1982].
-
-Care also has important implications for spatial and transportation planning.
-Indeed, researchers from various fields have highlighted that western modern industrial societies have imposed a physical separation of the private sphere from the public one, splitting production and reproduction, public and private spaces or simply work and home, concentrating on the economic outcome of being mobile [@cresswell_gendered_2008; @lawWomenTransportNew1999; @levy2013travel; @little_gender_1994; @tronto_toward_1990].
-Men tend to be associated with legal rights, paid labour, and the public sphere while woman are associated with the private sphere, dependents, family duties, personal relations and unpaid labour [@tronto_toward_1990].
-In this line, @cresswell_gendered_2008 described how masculinity has been constructed as mobile and active, whereas femininity is static.
-As an example of this dichotomy, transport planning has historically framed non-work/school trips as 'non-mandatory' activities, downgrading these activities in comparison to mandatory trips like work and school (XXX).
-In more recent years, planners have begun to resist the spatial separation of the public and private sphere to promote for sustainable cities.
-Proximity planning has become more normalized, a popular example is the 15-minute neighbourhood i.e., neighbourhoods where residents can reach all necessary destinations within a 15-minute active transport trip.
-The 15-minute neighbourhood concept, however, has yet to foreground care.
-
-While much care labour takes place in the private and feminised sphere, completing care activities requires one to be mobile.
-In other words: caring for oneself and others requires trips to, amongst other places, grocery stores, day cares, elder care, and health care destinations.
-From this research motivation, Sánchez de Madariaga coined the term ‘mobility of care’ in 2013 to refer to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands [@sanchezdemadariagaMobilityCareIntroducing2013].
-While decades of research had examined these unique household-serving trips, her seminal work was the first to consider all these trips as one category.
-In doing so, she showed how mobility of care is a significant proportion of daily travel.
-Indeed, in her original work, and in the studies that have followed, it appears that approximately 30% of adults’ trips are for care purposes [@sanchezdemadariagaMobilityCareIntroducing2013; @sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory].
+The adoption of the Mobility of Care concept within the 15-Minute City also explicitly incorporates equity considerations.
 Perhaps unsurprisingly given the gendered division of care work discussed above, women have consistently been found to complete more mobility of care trips than men.
@@
 Transport surveys and tools do not directly capture mobility of care, which re-enforces the idea that these trips are not a significant part of daily mobility.
-
-Cities have simplified individuals as a universal and neutral entity.
-As a result, public spaces are differently experienced depending on our multiple identities.
-Indeed, asymmetric in the distribution of time of citizens is deeply anchored with the public and private spaces organization within cities.
-In this light, we bring together the literatures on 15-minute cities and mobility of care by focusing on the potential access to destinations of care through an empirical case study of the mid-sized city of Hamilton, Canada.
-We acknowledge that caring is composed by actors and activities but aim to spotlight the role that the urban environment plays on our daily lives, on easing our participation and on exercising our rights to the city as individuals with multiple identities.
-Our work hopes to contribute to a fruitful discussion that may trigger urban and transport policy interventions that increase gender-equity in caring activities among all actors.
-
-## Feminist 15-minute neighbourhoods
-
-`r word_comment("TO ADD: some good feminist 15 minute cities.
-It's novel but movements exists in the Canadian context i.e., https://x.com/brenttoderian/status/1800371515877384527?s=48&t=QI0WSvV_FDb82ZGeLdJ9lQ.
-Also discuss Barcelona and Bogota (care blocks) non academic sources: https://oecd-opsi.org/innovations/bogota-care-blocks/ AND https://www.npr.org/sections/goatsandsoda/2023/10/16/1205218707/how-bogota-cares-for-its-family-caregivers-from-dance-classes-to-job-training )", " ")`
-
-<!-- The spatial structure of cities and regions has been described in different ways and its advantages and disadvantages has been discussed with regard to a wide variety of issues, for example, regarding health, the environment, economic, social and equity impacts.
-After the traditional monocentric vision of the city with a core Centre Business District (CBD), other alternatives appeared, especially due to the expansion of cities caused by the automobile and motorized transport.
-The polycentric urban planning emerged based on the idea of sub-centers (or multiple business centers) with their own services and functions, well-connected among other surrounding areas, where the links between land-use and transport were optimised.
-Economic activities (such as retail or commercial office development) and job centers played an important role in shaping the spatial structure [@schneider_transit_1981; @sweet_are_2017].
-This concept may have different interpretations depending on the spatial and geographical scales [@hall_polycentricity_2009].
--->
-
-In the last decades, the need for more sustainable, healthier, and liveable cities has received more attention.
-Planners and policy makers have proposed different alternatives where residents reach “essential activities” within walkable or cycling distances, reducing the need for using motorised transportation modes.
-The 15-minutes cities or "x"-minutes cities, have been long under the spotlight.
-This approach has also received the name of Chrono-urbanism or the “cities of proximity” [@logan_xminute_2022; @moreno_introducing_2021].
-
-In the international context, there are different examples that highlight the importance of good access to different services and amenities within walkable distances.
-Probably one of the first cities to adopt the proximity approach was Portland.
-The Portland Plan [@city_of_portland_20minute_2010] was adopted by the City Council on April 25, 2012 aiming at fostering an inclusive city development based on prosperity, education, health, and equity for the following 30 years.
-The progress report describes the tasks of the established Office of Equity and Human Rights was established focusing on equitable service delivery to all residents.
-In the first example, equity is constrained to topics related to racial equity or people with disabilities [@portland_government_portland_2017].
-`r word_comment("NOTE TO SELF: And so what?
-Is this a research gap?", " ")`
-
-Later, other examples have highlighted the benefits of mobility of proximity without specifically targeting mobility of care, where a universal and neutral entity approach persists.
-In Paris, the adoption of the 15-minute neighbourhood put into the spotlight this approach.
-They targeted six social functions that had to be easily accessible at any part of the city: housing, work, health care, groceries, education and leisure [@ville_de_paris_paris_2022].
-In 2021, Ottawa also adopted a 15-Minutes approach in the new City of Ottawa Official Plan [@ville_dottawa_quartier_2021].
-More recently, @moreno_introducing_2021 considers four dimensions on how the 15-minute neighbourhood framework: density (in terms of people per sq.km); diversity (including mixed-land use and diversity of people; the temporal and spatial proximity to essential services (residential areas, work, commercial areas, education centres, health facilities and other basic institutions) and digitalisation.
-In this respect, the feminist perspective of the cities of proximity is still underestimated.
-Only few examples exist that specifically target this issue [@gil_sola_choose_2022; @macintyre_her_2022].
+In this respect, the feminist perspective of the cities of proximity is still underestimated with only few examples on the topic [@gil_sola_choose_2022; @macintyre_her_2022].
 Given the impact of caring activities on everyday life, the chrono-urbanism should specifically target the analysis of caring activities.
 
-Within the backdrop of urban planning and ‘mobility of care’, we connect our work to the 15-minute neighbourhood concept.
-
-Both the 15-minute neighborhood and accessibility measures are flexible enough to consider all destination types holistically, however, gendered examinations have been lacking in the literature.
-For instance, accessibility analyses have historically been employment-centric even when adopting a gendered perspective [@paez_jobs_2013].
-Some 15-minute city analyses focus on care destinations, but they seem to not put care at the forefront, or to consider but one care destination (e.g., [@hosford15minuteCityReach2022] on grocery stores).
-Though few 15-minute neighbourhood analyses are explicitly feminist, their potential to build feminist communities have been noted outside of academia.
-For instance, Women in Urbanism have noted how women may benefit more than men from this 15-minute approach to planning.
-For instance, women tend to walk and use transit more than men, as well as to make more frequent, but shorter trips -and often as part of a chain (i.e., with stops along the way).
-These trip characteristics are all well served in 15-minute frameworks.
-15-minute neighbourhoods also promote walkability, which results in more “eyes on the streets” and its associated safety, an important enabler of mobility for women and girls.
-Finally, because women tend to earn less money than men and cars are a large financial burden, planning paradigms that reduce car dependency, such as 15 minute cities, may on average benefit women more than men [@Bertolo2023].
-
-## Mobility of care and mode-constraints
-
-Given that almost one third of daily trips are for care purposes [@sanchezdemadariagaMobilityCareIntroducing2013], it is important to plan for mobility of care.
-There are also equity considerations to consider; women, and especially low-income women, are particularly in charge of these duties and therefore, complete this travel more than men.
-Planning for mobility of care may also warrant different considerations because mobility of care trip characteristics differ from other trips, notably those of the commute to work.
+Mobility of Care also explicitly prioritizes shorter-distance and local trips, those that could be done by sustainable modes if sufficently supported, making the perspective highly compatible with the 15-Minute City paradigm.Compared to the trip to work, mobility of care trips are more frequently completed by foot, and less frequently done by transit or bicycle [@ravensbergen2023exploratory].
+Further, evidence suggests that walking is more common for care than for work trips because care trips tend to be short [@ravensbergen2023exploratory]; if care destinations are closer to the home, it is more likely that these destinations are walkable.
+Ample literature examines the benefits and barriers to walking care destinations, such as schools (e.g., [@RN7;@RN8;@RN9]) and grocery stores (e.g., [@RN10;@RN11]).
+However, little of this work uses the mobility of care framework.
+Instead, they tend to consider singular care destinations in research focusing on walkability.Furthermore, mobility of care trip characteristics differ from other trips, notably those of the commute to work.
 For instance, many of these trips are made through trip-chains, i.e., when a non-work-related stop is made during the commute.
@@
 
-Further, compared to the trip to work, mobility of care trips are more frequently completed by foot, and less frequently done by transit or bicycle [@ravensbergen2023exploratory].
-Further, evidence suggests that walking is more common for care than for work trips because care trips tend to be short [@ravensbergen2023exploratory]; if care destinations are closer to the home, it is more likely that these destinations are walkable.
-Ample literature examines the benefits and barriers to walking care destinations, such as schools (e.g., [@RN7;@RN8;@RN9]) and grocery stores (e.g., [@RN10;@RN11]).
-However, little of this work uses the mobility of care framework.
-Instead, they tend to consider singular care destinations in research focusing on walkability.
-
-`r word_comment("TO ADD: Replace the following three paragrapgs with a more elaborated discussion on walking as the primary mode for 15-min cities and how it overlaps with mobility of care. ", ".")`
-In car dependent cities, cycling commute rates tend to be low.
-Emerging evidence has found that they are even lower for care trips than for work trips.
-For instance, in Montréal it is estimated that 2% of work trips completed by adults are done by bicycle, while only 1.1% of care trips are done using this sustainable travel mode [@ravensbergen2023exploratory].
-Qualitative research has also found that in car-dependent contexts, many consider it inappropriate and unsafe to conduct escort trips by bicycle [@ravensbergenVelomobilitiesCareLowcycling2020; @sersliRidingAloneTogether2020].
-Then again, other studies have uncovered the joys inherent in travelling by bicycle with children [@RN14].
-This qualitative evidence has also found that the bicycle is excellent for trip-chaining owning to its flexibility and ease of parking, and that grocery shopping can be done by bicycle, though this can require some additional materials (e.g., paniers) and competencies (e.g., knowing where to put your groceries)[@ravensbergenVelomobilitiesCareLowcycling2020].
-Research in the Netherlands, a high-cycling context, has found that parents can and do use their bicycles to complete mobilities of care.
-Indeed, one study compared women with children’s bicycle behaviours to those without children and found that the having a child did not change the amount of cycling done, but the characteristics of cycling trips: mothers’ trips tended to center around their childrens’ needs [@RN15].
-Taken together, it is evident that the bicycle has potential to serve mobility of care: it is flexible, low-cost, and used for these trips in many contexts.
-Other barriers to this travel mode seems to exist in low-cycling contexts.
-
-The difference in uptake of public transport for work compared to care trips is staggering.
-In Montréal, it is estimated that 18.7% of trips to work are made by transit.
-In contrast only 5.8% of care trips are made using transit [@ravensbergen2023exploratory].
-While research examining accessibility to care destinations by transit exists (e.g., [@RN16;@RN17;@RN18;@RN19]), little research examines why few people use this sustainable transport mode for mobility of care.
-Though transit use for mobility of care is low, one study found that is it is significantly higher for low-income women.
-Indeed, approximately 16% of low-income women in Montreal were found to use transit for care trips (compared to 4% of high income women) [@ravensbergen2023exploratory].
-In many cities, public transport is not designed to meet household demands in a convenient way [@RN20;@grantsmithManagingChallengesCombining2016].
-Given how low transit use is for care trips amongst the general population, it is possible that many of these low-women using transit are transit-dependent.
-In other words, they may be using transit for care trips because they do not have access to a car, and not because transit is their preferred travel mode.
-
-It appears that the car is the preferred travel mode for mobility of care.
-In one study, car use was higher for mobility of care trips than the commute to work (82.3% vs.
-74.7%) [@ravensbergen2023exploratory].
-Further, the rate of car use is higher amongst trips that are part of a chain than those that are not (85.28% vs.
-78.09%) [@ravensbergen2023exploratory].
-Some studies found that a car is viewed as the easier way to incorporate unpaid mobility of care into one’s daily travel as it provides convenience and mitigates concerns about traffic and stranger danger [@maciejewskaHaveChildrenThus2019, @carverParentalChauffeursWhat2013].
-Others, however, have found that women are more likely to complete household-serving travel duties by alternative modes than men due to their lower car access [@RN24;@RN25].
-It seems that car use may be preferred for mobility of care trips, but planning mobility of care for the car may leave those who do not have access to a car, many of them low-income women, with difficult mobility of care trips.
-Studies also observe disparities regarding gender mobility and type of activities, depending on the area of residence; women living in the city center, where more transport systems and activities are available, use more sustainable travel modes than the ones living in the outskirts, which tend to rely more on cars to carry out daily activities [@marquet2014walking; @mejia2021mobility; @montero2023role; @montero2023applying].
-Outer areas are not as dense as in the downtown, and are basically car-oriented developments.
-Thus, studies observe that in the outer areas gender equity is confronted to sustainable mobility.
-
-The discussion around mobility of care and travel mode may also be relevant to larger spatial trends, including gendered spatial entrapment.
-Feminist geographers have noted for decades the spatial impact of women’s disproportionate household-serving labour.
-A relationship has been noted whereby many women’s household responsibilities results in job searches that emphasize job opportunities that are near the home and part-time or with flexible hours to allow for care responsibilities.
-This leads to a lower probability of securing higher paid, and often further afield, employment [@hansonGenderMobilityNew2010;@lawWomenTransportNew1999;@RN28;@RN29].
-This trend has been found to be surprisingly resilient, with improvements only taking places slowly [@RN30].
-Evidence has also shown that this spatial entrapment can be more in car dependent contexts, as access to cars impacts employment rates [@RN31], and is lower amongst women, and especially women of colour [@RN32].
-
-Taken together, there appears to be a preference for travel by car for mobility of care trips in many geographical contexts.
-This is problematic, however, in at least two major ways.
-First, cars are less sustainable than other modes of transport.
-Normalizing car use for care trips does not advance the low-carbon transport transition that most cities are aiming for (at least in principle!).
-Since care trips tend to be short, they are also perfect candidate to be completed by bus, bicycle, or by foot.
-That is, if these alternative and sustainable modes are planned with care in mind.
-Second, there is equity concerns with assuming people are accessing care destinations by car.
-Many people, and low-income women in particular, do not have access to a car.
-Some are also experiencing a degree of spatial entrapment.
-Having safe and convenient access to care destinations using alternatives to the private car is especially important for those who must rely on these options regardless of whether they are well planned.
-
+`r word_comment("I need to clean this up and firmly reiterate the goals", " ")`
 Taken together, city planning is moving to a people-centered approach focusing on accessibility to basic needs.
-However, chrono-urbanism stems from conventional methods that overlooks mobility of care perspectives.
-To counter this, this paper presents a 15-minute cities methodological approach to assess mobility of care.
-We develop this approach using a mid-size typical Canadian city and a novel care destination dataset as case study.
+However, the 15-Minute City has faced criticism as it fails to provide sufficient guidance on how to reach the utopic goal of 15-Minute for all.
+We argue, that significance in what destinations matter within the 15-Minute City could be gained by adopting a Mobility of Care prespective.
+We develop this approach using a mid-size typical Canadian city of Hamilton and a novel care destination dataset as case study.
 The objectives of the paper are as follows:
@@
 We identify to what degree are neighbourhoods caring and/or complete.
-To do so, we make use of self organizing maps (SOM), an unsupervised machine learning data-dimensionality reduction technique to identify caring and complete neighbourhoods.
+To do so, we make use of machine learning techniques namely self organizing maps (SOM) as a data-dimensionality reduction technique and decision trips for the purpose of classification.
 2)	We then apply the same SOM methodology to determine profiles of population groups to understand who enjoys these neighbourhoods and who does not.
@@
 
-In Canada, women have become more educated and have an achieved an increased participation in the labour market [@government_of_canada_gender_2023].
-In fact, the educational level of women has surpassed most men [@government_of_canada_women_2023].
-Reports observe that the gender pay gap has narrowed but disparities persist [@drolet_intersectional_2023;@pay_equity_office_gender_2023].
-According to the statistics, in Ontario between 1998 to 2022 the gender gap narrowed but still women make 87 cents on average for every dollar made by a man [@pay_equity_office_gender_2023].
-All reports observe that it affects women differently, women with disabilities, immigrants, indigenous or racialized suffer especially from the gender pay gap.
-Similarly to other countries worldwide, the occupational segregation remains and women are segregated into lower paid job classes and there is a dominance of women in minimum wage jobs [@drolet_intersectional_2023].
-Furthermore, women are underemployed, and make up most part-time work.
-In this respect, the lack of childcare alternatives enforces this obligatory part-time work [@equal_pay_coalition_myths_2022].
-All researchers and organisations agree that an intersectional approach is of utmost importance to better plan policies to enforce gender equality.
-
 This study focuses on Hamilton, Ontario, a mid-sized city on the shore of Lake Ontario.
 Hamilton has a heterogenous land-use, with a populated and dense urban core, surrounded by suburban development, which is itself surrounded by rural communities.
-The Niagara Escarpment runs through Hamilton, and results in a city with two key elevations: the dense lower city that contains the downtown core and the elevated suburban development referred to as ‘the Mountain’ (ADD THIS TO FIGURES).
+The Niagara Escarpment runs through Hamilton, and results in a city with two key elevations: a more dense lower city that contains the downtown core and the elevated suburban development referred to as ‘the Mountain’.
 In this work, we analyse the residential parcel centroids, approximately ~140,000 locations.
@@
 
-Hamilton also exhibits spatial disparities in social and economic indicators @fig-Fig2.
-The densely populated inner city is characterised by lower average incomes, and a higher prevalence of households living under the low-income cuttoff thresholds (LICO).
+Hamilton also exhibits spatial disparities in social and economic indicators; their spatial distribution is visualised in @fig-Fig2.
+The densely populated inner city is characterised by lower average incomes, and a higher prevalence of households living under the low-income cutoff thresholds (LICO).
 The suburban parts of the city tend to have a greater proportion of children and a lower proportion of one parent households.
@@
 
-## Estimated travel times
+## Care destination dataset
 
-Here we will describe the estimated 15-minute walk travel times using {r5r}, from all residential parcel points to all care destinations assuming default walking speed (3.6km/hr) and OSM road network for Hamilton.
-`r word_comment("TO ADD: add summary statistics", " ")`
-
-## Care destination dataset
-A spatial dataset of care destinations for Hamilton was compiled; sources include provincial and municipal open data portals [@governmentofontarioOntarioDataCatalogue2023; @cityofhamiltonOpenData2023], Data Axle, a consumer dataset compiled of businesses and companies within Canada [@axledataConsumerData2023] and manual retrieved through Google Maps.
-Each destination was categorized based on the specific type of care being accessed following the mobility of care research by @sanchezdemadariagaMeasuringMobilitiesCare2019.
-Initial categories include child, elder,  grocery, health, and errand -centric destinations.
-For the purpose of this study, child- and elder- centric are combined to represent a 'dependent-centric' category.
+A spatial dataset of care destinations for Hamilton was carefully compiled.
+This includes 14 types of destinations that were placed by authors in ( _work removed for double-blind review_ ) into four categories: dependent-centric (e.g., the destinations for child- and elder-centric escorting trips), grocery-centric, health-centric, and errand-centric.
+Notably, these categories were generated following the travel purpose categories created in the mobility of care research by @sanchezdemadariagaMeasuringMobilitiesCare2019.
 Category sources of data and preparation notes are detailed in @tbl-Tbl1.
-Their spatial distribution and sub-categories are visualised in @fig-Fig3.
+The spatial distribution of destination type are visualised in @fig-Fig3 by their category.
 
@@
 
+## Travel time to care estimations
+
+Summarized in the literature review, empirical travel behaviour to care-oriented destinations is often uncounted and thus travel time is unavailable.
+Hence in this work, travel time from the residential parcel locations and care destination locations is approximated for walking assuming an average speed (3.6 km/hr) using the 'travel_time_matrix()' function from the {r5r} package [@pereiraR5rRapidRealistic2021].
+The inputs into the function are: the 143,882 locations of residential parcels as origins, the 2,225 care locations as destinations, and a OpenStreetMap road network including walking infrastructure [@geofabrikOntarioCanadaOpen2023].
+In theme with what destinations can be reached within 15-Minutes, a maximum walking travel time of 15 minutes is selected, and an origin-destination travel time matrix of the shortest travel time from origin to destination is calculated.
+The resulting matrix contains 2,014,502 rows, representing walking travel times from each parcel to reachable care destinations within 15 minutes.
+
 # Methods
 
-## Accessibility and entropy
-A 15-minute normative walk catchment is used to represent the 15-minute neighbourhood concept.
-This has been done in previous literature such as @vale2023accessibility, @graellsCityCitiesMeasuring2021 and @gaglioneUrbanAccessibility15minute2022.
-In this work, the cumulative opportunity score is calculated for each parcel $i$ and aggregated by care category $c$ and takes the following mathematical form:
+In the following sections, the methods taken to calculate the accessibility to each of the 14 destination types from each of the 143,882 residential parcel locations are detailed.
+Then, the entropy measures used to calculate the diversity of accessibility to each of the care categories is described.
+Subsequently, the accessibility and diversity of accessibility values associated with each residential parcel location are then fed into a self-organizing map (SOM) data-reduction algorithm.
+How the resulting output is classified into 9 classifications using decision-trees that represent the amount of "caring access" and "completeness" of this access are then described.
+This methodology presents a data-driven approach to examine what neighbourhoods in a city have the potential to provide 15-minute caring access, at what level of intensity and completeness, as well as in who may be residing in areas that benefit the most.
+
+## Accessibility: the cumulative opportunity measure
+
+To capture the quantity of access to each type of destinations, a cumulative opportunity accessibility score $S^t_i$ is calculated.
+14 scores for each type of care destination $t$ is calculated for every parcel $i$.
+The calculation takes the following mathematical form:
+
 $$
-S^c_i=\sum_{j}O_j^{c}\cdot f(c_{ij})
+S^t_i=\sum_{j}O_j^{t}\cdot f(c_{ij})
 $$
@@
 - $i$ is a set of parcel point origin locations.
-- $j$ is a set of destination locations.
-- $c$ is a set of care categories.
-- $O^c_j$ is a number of opportunities of category $c$ at $j$.
+- $j$ is a set of care destination locations of type $t$.
+- $O^t_j$ is a number of opportunities of category type $t$ at $j$.
 - $c_{ij}$ is the travel cost between $i$ and $j$.
 - $f(\cdot)$ is an impedance function of $c_{ij}$; within the cumulative opportunity approach, it is a binary function that takes the value of 1 if $c_{ij}$ is less than a selected value [@handyMeasuringAccessibilityExploration1997].
-- $S_{i}^c$ is the unconstrained accessibility, the sum of weighted opportunities reachable within $f(\cdot)$, at each $i$ for each $c$.
+- $S_{i}^t$ is the cumulative opportunity accessibility score, the sum of weighted opportunities reachable within $f(\cdot)$, at each $i$ for each $t$.
 
-*Entropy* measure has been widely used to characterise the diversity of land-use mix in early work by @frankLinkingObjectivelyMeasured2005 based on the species diversity Shannon-Wiener index that expresses relative evenness throughout a sequence (this measure among other species diversity indices of the time are reviewed in @whittaker1972evolution).
-It has also been used in social sciences to comprehend the occurrence of social events in a structured manner specially with regard to life trajectories [@ritschardSequenceAnalysisRelated2018].
+## Diversity in opportunity accessibility: the entropy measure
 
-In the realm of urban studies and planning literature the malleable concept of entropy has been used to characterise the diversity in land-use mix [@ewingTravelBuiltEnvironment2010a], especially in the context of active modes [@luUrbanDensityDiversity2017; @mavoaIdentifyingAppropriateLanduse2018] and suburban sprawl [@randallGISBasedLand2015].
-As well, it has been used to understand mobility behaviour [@mcbride2020exploration; @montero2023applying; @montero2023role].
-Work that has used diversity indices alongside accessibility analysis are scarce, though examples of its use as parameters within accessibility scores in the case of employment opportunities [@chengMeasuringUrbanJob2013; @daiIncorporatingJobDiversity2018] and more recently to describe the diversity in transit facilities [@yinIncorporatingFacilityDiversity2024] are present.
-To the authors knowledge, the use of an entropy measure to reflect the diversity in care destination amenity types and in context within a 15-minute neighbourhood, has yet to be done.
+To represent the diversity of care destination accessibility, the entropy measure is used.
+A value is calculated for each parcel that ranges between 0 to 1, where 1 represents total evenness in the number of care opportunities in each category that can be reached.
 
-As such, the entropy measure is used to reflect the diversity of care destinations categories.
-The value ranges from 0 to 1, with 1 representing total evenness in the number of care opportunities in each category that can be reached.
-$i$ is a parcel, $c$ is the category of care destinations, $n_c$ is the count of care destination categories (in this work 4), $S_i^c$ is the spatial accessibility of care destinations within a 15-minute walk for a parcel, and $D_i$ is the diversity.
-As $D_i$ represents evenness in categories, a parcel $i$ can be assigned a diversity of 1 if it has $S_i^c$=1 of each category or if it has $S_i^c$=100 of each category.
-Contrariwise, it can be assigned a low score if it has low but different scores for each category as well as if it has high but different scores for each category.
-
+The mathematical formulation of takes the following form:
 $$
-D_{i} = \frac{-\sum_{c} S_i^{c} / \sum_{c}S^{c}_{i} \times ln(S_i^{c} / \sum_{c}S^{c}_{i} )}{ln(n_c)}
+D_{i} = \frac{-\sum_{t} [S_i^{t} / \sum_{t}S^{t}_{i} \times ln(S_i^{t} / \sum_{t}S^{t}_{i} )]}{ln(n_t)}
 $$
 
-## Self-organizing maps
+\noindent Where:
 
-Self organizing map (SOM) is used to cluster the types of neighbourhoods by their magnitude in 15-minute walking accessibility to caring opportunities and diversity in these opportunities and profile these clusters based on their socio-economic characteristics.
-SOM is a type of artificial neural network (ANN) trained using unsupervised learning that produces a two-dimensional representation of the input space.
-It is a dimensionality reduction method as it applies competitive learning by way of a neighbourhood function thereby preserving the topological properties of the input space and producing combinations of the input space that we intuitively profile.
-It has been used widely in XX but scarcely in transportation related research [@victorianoTimeSpaceMoney2020]
+- $i$ is a set of parcel point origin locations.
+- $t$ is a set of care destination types (e.g., school, grocery, park, etc.)
+- $n_t$ is the count of care destination types $t$. In this work, this value is 14.
+- $S_{i}^t$ is the cumulative opportunity accessibility score, the sum of weighted opportunities reachable within a 15-minute walk from $i$.
+- $D_i$ is the diversity score.
+As $D_i$ represents evenness in type of categories, so if a parcel has a access score of $S_i^t$=0.5 for all types of destinations it will be assigned $D_i$ = 1 in the same was as if it had a score of $S_i^1$=10 for each type of destination.
+Contrariwise, a parcel can be assigned a low $D_i$ score if it has low but different accessibility scores for each category as well as if it has high but different scores for each category.
 
-The SOM procedure is broken into 2 steps: first, five superclusters of the four $S_i^c$ scores and the diversity score $D_i$ are identified.
-The number of superclusters strike a balance between variance explained and supercluster interpretation in what we define as ranging from "not caring" to "very caring" and "not complete" to "complete".
-Second, the superclusters are profiled using a variety of socio-economic and demographic indicators that the mobility of care literature finds significant (@fig-Fig2).
+## Machine learning classification: self-organizing maps and decision trees
 
-## Decision trees
+In this work, we use two machine learning techniques.
+First, SOM is an unsupervised technique implemented to reduce the data dimensional and create interpretable clusters related to the intensity and completeness of caring access.
+Then, a decision tree is deployed to further examine the cluster membership of parcels and another decision tree is run to characterise the socio-economic profile of who resides in neighbourhoods associated with certain (in)complete and (un)caring accessibility clusters.
+Together, this combined approach leverages the unsupervised data-driven classification power of the SOM with the interpretablity of decision trees.
+The procedure deployed in this work closes resembles the work of @victorianoTimeSpaceMoney2020, however instead of each observation representing an individual's daily mobility behaviour (with associated variables) each observation is a parcel location with calculated care accessibility and diversity of type of care accessibility scores.
 
-Decision trees area a supervised learning technique, which can use the results of the clustering algorithm as labels, and the input variables as features.
+For the SOM step, the SOM algorithm implementation in the function 'trainSOM()' from {SOMbrero} R package is used [@VillaVialaneix2017].
+The input variables include the 143,882 parcels, each as individual observations along with 15 variables: the 14 calculated accessibility scores $S_i^$, normalized to the min-max range score within each $t$, and one diversity value $D_i$.
+Otherwise, defaults for all other parameters are assumed, relying on the data-driven heuristics embedded in the 'trainSOM()' function.
+Consequently, a 100 node (10 by 10 gride) SOM structure using euclidean distance and square topology is produced.
+Simply put the SOM algorithm proceeds as follows.
+A 2D grid of nodes is created (in our case 100 nodes) as specified by the analyst, where each node will represent a point in the reduced-dimensional space.
+Upon initialization, each node is assigned a random weight vector of the same dimension as the input data (in our case, 15).
+From the input data, a random observation with its associated weight vector (i.e., one parcel point with 15 variables) is selected and the Euclidean distance between its weight vector and all nodes in the grid is calculated using Euclidean distance.
+The node with the smallest distance (i.e., the smallest dissimilarity) is crowned the 'best matching unit' as it is the node that best represents the input observation.
+After this best matching unit is identified, its own weight and its neighbouring nodes are updated to become more like the input observation.
+The process of finding best matching units and updating their weights is repeated for every observation, multiple times, until the results converge.
+As mentioned, this competitive learning process produces a 100 node SOM structure where each observation (parcel) is assigned to 1 node with an associated dissimilarity index.
+The SOM output is typically examined through a dissimilarity dendogram and an associated dissimilarity variance explained plot to select an appropriately representative 'superclusters' [@VillaVialaneix2017; @victorianoTimeSpaceMoney2020].
+In this work, we refer to these superclusters as simply clusters.
+
+For the decision tree step, the cluster-classified parcels identified in the SOM step are used as _labels_ and socio-economic and demographic indicators available from the most recent Canadian Census related to the mobility of care literature are used as _features_.
+This step is used to profile the clusters in a data-drive way to explore who may reside in what cluster.
+To estimate the decision tree, the 'rpart()' function in the {rpart} package assuming default parameters for classification splitting along with each value being weighted by the population present in the associated DA [@Therneau2023].
+As a summary of the decision tree algorithm, it is a supervised learning technique that begins by splitting a subset of the input data into branches based on a selected feature with the lowest impurity score (i.e., the lowest amount of mixing of label membership in a branch).
+This process is recursively repeated for each subset of data for the next best features.
+Ultimately, the data is classified into distinct classes with class membership explained by traversing the branches defined by features that characterise the partitions in the decision tree.
+Noteably, the absence of features from the decision tree does not necessarily mean they are not relevant for classification, they are just less relevant than other features.
+Put another way, when considering features that are highly correlated, such as income level and LICO, not all relevant variables may be present in the tree [@victorianoTimeSpaceMoney2020].
 
@@
 
-Accessibility results per care category are presented in @fig-Fig4.
-The concentration of opportunities for all category types is concentrated in Hamilton Central - with the highest concentration being in the lower city and pockets of the high concentration on the mountain `r word_comment("(TO ADD: The escarpment line)", " ")`.
-Grocery-centric opportunities are most concentrated, followed by dependent-centric and health-centric.
-Errand-centric is the most sprawled.
+## Overview of access to care destinations 
+How many care destinations can be reached within a 15-minute walk is summed for each of the four care categories defined in @bl-tbl1.
+The median parcel value for each DA is visualised in @fig-Fig4.
+Furthermore, the median parcel diversity of care destinations accessible is presented in @fig-Fig5.
+Together, @fig-Fig4 and @fig-Fig5 represent the 15 variables that served as inputs into the SOM algorithm.
 
@@
 #| fig-cap: "The number of care destinations that can be reached within 15 min walk per care category for a median parcel in each DA.
-Scales are in quartiles.
-These are four of five inputs into the superclusters.
+The values are a summary of the 14 acessibility scores that served as inputs into the SOM.
 Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
@@
 
-The diversity of care destinations accessible is presented in @fig-Fig5.
-Here, it appears that neighbourhoods with a diversity of care destinations accessible are concentrated in the downtown core.
-Of note, some suburban areas also exhibit a high diversity of care access, these areas _have the potential_ to be retrofitted/ designed for access care (ADD MORE).
-
 ```{r plot-figure5, out.width=600, out.height=600}
 #| label: fig-Fig5
-#| fig-cap: "The diversity measure (div) based on the proportion of care category spatial accessibility (figure 4).
-Scale is in tertiles.
-The fifth input into the superclusters.
+#| fig-cap: "The diversity measure based on the proportion of care category spatial accessibility (figure 4).
+These values are a summary of the 15th input into the SOM.
 Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
@@
 
-Superclusters based on accessibility per category and diversity (the two above figures).
-5 clusters are identified, see their dendrogram (right) and grid map (left) in @fig-Fig6 `r word_comment("change colours of SC to match the following plots", ":")`.
+## Identification of (in)completely (un)caring neighbourhoods
 
-```{r plot-figure6-create, eval=FALSE}
-plot1 <- ggdraw() + draw_image(glue::glue(here::here(), "/figures/access_sc_plot_2.png"))
-plot2 <- ggdraw() + draw_image(glue::glue(here::here(), "/figures/access_sc_plot_1.png"))
-plot3 <- plot_grid(plot1, plot2)
+Notably, in @fig-Fig4, the concentration of opportunities for all category types is centered in the downtown of Hamilton, with the highest concentration being in the lower city (near the lake shore) and pockets of the high concentration further from the shoreline.
+Grocery-centric destinations appear to be the most concentrated, followed by health-centric and dependent-centric caring destinations.
+Errand-centric destinations are the most sprawled.
+In many ways, @fig-Fig4 mirrors the spatial distribution of care destinations (@fig-Fig3) as the range of 15-minutes of travel is small.
+In @fig-Fig5, areas that have high care accessibility tend to have high diversity as well, though there are exceptions in pockets of the city outside the downtown core that have lower but equally low levels of accessibility for all care categories.
 
-ggsave(glue::glue(here::here(), "/figures/access_sc_plot_combo.png"), dpi=300)
+Based on the SOM classification methodology discussed and the 15 variables represented in @fig-Fig4 and @fig-Fig5, 7 clusters are identified.
+As diagnostics for the selection of the number of clusters, the dissimilarity-index-based dendrogram with the proportion of parcels represented in each cluster alongside the variance explained plot is visualised in @fig-Fig6.
+```{r plot-figure6, out.width=800, out.height=800}
+#| label: fig-Fig6
+#| fig-cap: "The 7 resulting clusters from the SOM the 10 by 10 grid output represented in a dendrogram (left) and the variance unexplained (right)."
+#| 
+knitr::include_graphics(glue::glue(here::here(), "/figures/all_access_sc_plot_small.png")) 
 ```
 
-```{r plot-figure6, out.width=800, out.height=800}
-#| label: fig-Fig6
-#| fig-cap: "The 5 resulting superclusters from the SOM output (left) and the associated dendrogram structure (right)."
-#| 
-knitr::include_graphics(glue::glue(here::here(), "/figures/access_sc_plot_combo.png")) 
+To characterise the clusters, a table providing summary statistics of each cluster is detailed in @tbl-tbl2:
+```{r creating-table2}
+summary_df_ALL <- access_15_wide |> select(c(starts_with("total_"), div_allcats)) |>
+  group_by() |> 
+  summarise(across(everything(), list(median = median, sd = sd))) %>% 
+ mutate_if(is.numeric, round, digits =1) |>
+  mutate_all(as.character)  
+summary_df_ALL <- cbind(data.frame(scluster = "TOTAL"),summary_df_ALL)
+
+summary_df <- access_15_wide |> select(c(scluster, starts_with("total_"), div_allcats)) |> 
+  mutate(scluster = case_when(scluster == 1 ~ "A",
+                              scluster == 2 ~ "B",
+                              scluster == 3 ~ "D",
+                              scluster == 4 ~ "F",
+                              scluster == 5 ~ "E",
+                              scluster == 6 ~ "C",
+                              scluster == 7 ~ "G")) |>
+  group_by(scluster) |> 
+  summarise(across(everything(), list(median = median, sd = sd))) %>% 
+ mutate_if(is.numeric, round, digits =1) |>
+  mutate_all(as.character)
+
+merge_every_two_columns <- function(df, num_rows) {
+  # Loop through every second column starting from column 2
+  for (col in seq(2, ncol(df), by = 2)) {
+    if (col < ncol(df)) {  # Ensure the column has a next one to merge with
+      for (i in 1:num_rows) {
+        df[i, col] <- paste0(df[i, col], " (", df[i, col + 1], ")")
+      }
+    }
+  }
+  return(df)
+}
+
+summary_df <- merge_every_two_columns(summary_df, 7) |>  select(-contains("sd"))
+summary_df_ALL <- merge_every_two_columns(summary_df_ALL, 1)  |>  select(-contains("sd"))
+summary_df <- rbind(summary_df, summary_df_ALL)
+
+summary_df <- summary_df |> 
+  pivot_longer(!scluster) |>
+  pivot_wider(names_from =scluster )
+
+summary_df_ALL <- summary_df_ALL |> 
+  pivot_longer(!scluster) |>
+  pivot_wider(names_from =scluster )
+
+summary_df$name <- c("GROCERY TOTALS", "Convenience Store", "Grocer",
+                     "DEPENDENT TOTALS", "Community or Rec. centre", "Daycare or EarlyON","LTC or retirment home","Park","School","Senior centre",
+                     "HEALTH TOTALS","Dentist","Hospital or clinic","Pharmacy",
+                     "ERRAND TOTALS","Bank or ATM","Library", "Post office", 
+                     "Category diversity")
 ```
 
-The constitution of the superclusters is shown in the decision tree diagram in @fig-Fig7.
-Decision trees area a supervised learning technique, which can use the results of the clustering algorithm as labels, and the input variables as features.
-The clustering algorithm performed well as class membership is relatively homogeneous as the majority of bar plots consist of a supermajority of one type of supercluster class membership, i.e., supercluster 5 are parcels that have an Errand-centric access of less than 0.5 opportunities (bottom quartile), a diversity score of less than 0.277 (bottom tertile), and any Dependent-centric access score of greater than 0.5 (and can take on any value of the remaining input variables).
+```{r creating-table2-pt2}
+#| tbl-cap: "Median and (standard deviation) of each SOM classified cluster by input variable and additional summary variables (indicated by ALL CAPITAL LETERS)."
+#| label: tbl-Tbl2
+#| 
+summary_df %>%
+  gt() %>%
+  tab_header(
+    title = " ",
+  ) %>%
+  cols_label(
+    name = " "
+  ) %>%
+  tab_style(
+    style = cell_borders(
+      sides = c("top"),
+      color = "black",
+      weight = px(1.5),
+      style = "solid"
+    ),
+    locations = cells_body(rows = c(1,4,11,15))
+  ) |>
+  tab_options(
+    table.font.size = 12,
+    table.width = pct(80)
+  )
+```
+
+Notably, from A to F, the clusters are characterized by caring access and diversity of caring access (completeness). From the descriptive statistics of the clusters, interesting trends emerge: ...
+
+Not caring - not complete (C7)
+Somewhat caring - not complete (C4)
+Somewhat caring - complete (C5)
+Caring - somewhat complete (C3)
+Caring - very complete (C6)
+Very caring - complete (C2)
+Very caring - very complete (C1)
+
+parcels that have an Errand-centric access of less than 0.5 opportunities (bottom quartile), a diversity score of less than 0.277 (bottom tertile), and any Dependent-centric access score of greater than 0.5 (and can take on any value of the remaining input variables).
 This technique provides a data-driven approach to supercluster formation which we then attribute meaning to.
@@
 ```
+
+## Profiling of who resides in (in)completely (un)caring neighbourhoods
 To enhance the meaning of the superclusters beyond a spatial conceptualisation of "caring" and "complete", was _who?_ resides there.
@@
 
+<!--
+
+Instead of prescribing an urban form design principle, such as "all local amenities should be within a 15-minute walking distance", it instead examines an empirical example to determine which areas in the city have the _potential_ to be 15-minute neighbourhood based on the existing spatial accessibility offered by the urban environment and walking transport network.
+To this end, this data-driven methodology introduces a way to identify neighbourhoods that have potential, almost have potential, and are far from containing this potential to support future context-specific qualitative work.
+
+-->
+
 # References

### 2024-09-27 paper/rejected-text.qmd
Still working on the draft...

--- 2024-05-16 Manuscript.qmd
@@
 ---
-title: "Towards caring 15-minute neighbourhoods"
-author:
-  - name: Anastasia Soukhov 
-    email: soukhoa@mcmaster.ca
-    affiliations:
-        - name: McMaster University
-          department: School of Earth, Environment & Society
-          city: Hamilton, Canada
-    attributes:
-        corresponding: true
-  - name: Léa Ravensbergen
-    email: ravensbl@mcmaster.ca
-    affiliations:
-        - name: McMaster University
-          department: School of Earth, Environment & Society
-          city: Hamilton, Canada
-  - name: Lucía Mejía Dorantes
-    email: lucia.mejia.dorantes@gmail.com
-    affiliations:
-        - name: Consultant
-          city: Karlsruhe, Germany
-  - name: Antonio Páez
-    email: paezha@mcmaster.ca
-    affiliations:
-        - name: McMaster University
-          department: School of Earth, Environment & Society
-          city: Hamilton, Canada
-format: 
-  docx
-  # elsevier-pdf:
-  #   keep-tex: true
-  #   include-in-header:
-  #     text: |
-  #       \usepackage{graphicx}
-  #       \usepackage{unicode-math}
-  #       \usepackage{hyperref}
-  #       \def\tightlist{}
-  #       \usepackage{setspace}
-  #       \doublespacing
-  #       \usepackage{lineno}
-  #       \linenumbers
-  #   journal:
-  #     name: 7th International Conference on Women and Gender in Transportation (WGiT)
-  #     formatting: preprint
-  #     model: 3p
-  #     cite-style: authoryear
-keywords: ["Accessibility", "Mobility of Care", "Gender", "15-Minute City", "Care Complete Neighbourhoods"]
-bibliography: bibliography.bib
+title: "rejected_text"
+format: html
 ---
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE,
-                      message = FALSE,
-                      warning = FALSE,
-                      cache=FALSE)
-```
 
-```{r}
-#function for creating comments
-word_comment <- function(comment, highlight = "") {
-  if (isTRUE(knitr:::pandoc_to() == "docx")) {
-    paste0('[', comment, ']{.comment-start id="0" author="Anastasia"',
-           'date="2024-05-16T15:00:00Z"}', highlight, '[]{.comment-end id="0"}')
-  }
-}
-```
+The "15-Minute City" has been recently adopted by leaders as a way to promote human-scaled cities.
+As defined in @moreno_introducing_2021, the 15-Minute City is a urban planning model based in chrono-urbanism, a theory that emphasizes the positive impact on quality of urban life when urban space becomes multi-rythmic [related to chronotopy @mulicekUrbanRhythmsChronotopic2015].
+Multi-rhythmic urban space is in opposition to industrial Fordism and single-use zoning [@mulicekUrbanRhythmsChronotopic2015; @morenoVilleQuartHeure2016]; the segregation of urban time and space for individual uses and mobility.
+The 15-Minute City thus one that accommodates the ability to reach essential destinations such as shops, work, and healthcare within a walkable radius.
+It would allow individuals to reclaim time spent on car mobility and give way to sustainable modes; prompting urban spaces that are responsive to human needs and environmental sensibilities [@Allam2022].
+The 15-Minute City generated much interest in planning practice and public discourse, especially after Paris mayor Anne Hidalgo proposed a 15-minute city plan in her 2020 re-election campaign.
+Though the term 15-Minute City is relatively new, neighbourhood planning is not (e.g., Perry's Neighbourhood Unit Concept (NUC) from the 1930s [@kissfazekasCircleParadigms15minute2022], 15-minute neighbourhoods in China (X)).
+A critique of neighbourhood planning approaches have been their top-down nature [@talenSocialSciencePlanned2017].
+The 15-Minute City esposes to be different, however top-down planning questions of how to equitable select travel-time thresholds (15 minute, or something else?), by what sustainable mode, to what destinations, and for whom continue to plague proximity-based urban planners.
 
-```{r library-setup}
-library(biscale)
-library(cowplot)
-library(sf)
-library(here)
-library(dplyr)
-library(tmap)
-library(mapview)
-library(tmaptools)
-library(pander)
-library(flextable)
-library(ftExtra)
-library(scales)
-```
+While much care labour takes place in the private and feminised sphere, completing care activities requires one to be mobile.
+In other words: caring for oneself and others requires trips to, amongst other places, grocery stores, day cares, elder care, and health care destinations.
 
-```{r}
-load("data/care_dest.rda")
-```
+@moreno_introducing_2021 considers four dimensions on how the 15-minute neighbourhood framework: density (in terms of people per km$^2$); diversity (including mixed-land use and diversity of people); the temporal and spatial proximity to essential services (residential areas, work, commercial areas, education centres, health facilities and other basic institutions) and digitalisation.
 
-```{r reformat care-dest specific categories}
-care_dest <- care_dest |> 
-  mutate(Care_Category_Specific = case_when(
-    Care_Category_Specific  == "Daycare_EarlyON" ~ "Daycare or EarlyON",
-    Care_Category_Specific == "Community_Centre" ~ "Community or Recreation Centre",
-    Care_Category_Specific == "Daycare_Centre" ~ "Daycare or EarlyON",
-    Care_Category_Specific == "Recreation_Centre" ~ "Community or Recreation Centre",
-    Care_Category_Specific == "School" ~ "School",
-    Care_Category_Specific == "Long-Term_Care_Home" ~ "Long-Term Care or Retirement Home",
-    Care_Category_Specific == "Retirement_Home" ~ "Long-Term Care or Retirement Home",
-    Care_Category_Specific == "Senior_Centre" ~ "Senior Centre",
-    Care_Category_Specific == "Seniors_Active_Living_Centre" ~ "Senior Centre",
-    Care_Category_Specific == "Post_Office" ~ "Post Office",
-    Care_Category_Specific == "Library" ~ "Library",
-    Care_Category_Specific == "Banks" ~ "Bank or ATM",
-    Care_Category_Specific == "Grocer_ConvenienceStore" ~ "Convenience Store",
-    Care_Category_Specific == "Clinics" ~ "Hospital or Clinic",
-    Care_Category_Specific == "Grocer_Retail" ~ "Grocer",
-    Care_Category_Specific == "Dentist" ~ "Dentist",
-    Care_Category_Specific == "Hosptial" ~ "Hospital or Clinic",
-    Care_Category_Specific == "Hospital" ~ "Hospital or Clinic",
-    Care_Category_Specific == "Pharmacy" ~ "Pharmacy",
-    Care_Category_Specific == "Park" ~ "Park"),
-  Care_Category = case_when(
-    Care_Category == "Child_Centric" ~ "Dependent-centric",
-    Care_Category == "Elder_Centric" ~ "Dependent-centric",
-    Care_Category == "Errand_Centric" ~ "Errand-centric",
-    Care_Category == "Grocery_Centric" ~ "Grocery-centric",
-    Care_Category == "Health_Centric" ~ "Health-centric"))
+<!-- These critiques were supported by case studies of planned neighbourhoods as...
+lead to social segregation of classes, destination functions, and automobile use.
+From it's onset, it was an efficient way to develop housing after the Second World War.
+The emphasis of proximity to some 'daily essentials' but disconnection from specialised activites (e.g., work) paved the way for car-based subrbia.
+by capital of cities.
+Can the neighbourhood form might be seen as a vehicle for capitalist exploitation, gentrification and displacement???
+The place of neighbourhood planning is actively being debated: can it's social crtiques be addressed in the new form as a "15-Minute City"?
+This study seeks to contribute to this decades old debate.
+-->
 
-care_dest <- care_dest |> rename("care_ID" = "ID") |> mutate(care_ID = as.character(care_ID)) 
-# care_dest|> st_drop_geometry() |> group_by(Care_Category) |> summarise(count = n())
-```
+Before discussing 'what outcome', we first ask 'what amenity' as the focus of this study.
+Often, accessibility analysis studies have focused on a single amenity type (X), often employment destinations and often focused on a scale beyond the neighbourhood.
+However, due to 15-Minute City's emphasis on proximate-level service provision, recent academic work on the topic has employed accessibility analysis to account for multiple 'local' amenities types, but seldom included employment.
+We argue that the inclusion of employment opportunities into the 15-Minute City conceptualisation is critical to it's definition.
+Why?
+We echo critics made by planners and social scientistics in decades past: neighbourhoods that did not mix work with service and culture within the neighbourhood, paved the way for car-based suburbia: cars allowed for the connection of neighbourhoods with their 'lower-order' amenities to areas with 'high-order' destinations like work (X).
+The lower- and higher- order framing of these destinations supported the continued segregation of urban life across many dimensions (X).
 
-```{r re-weighting points so all destinations are considered spatially equal in access counts, include=FALSE}
-Dependent_Centric <- 0.25*2225/1265
-Errand_Centric <- 0.25*2225/158
-Grocery_Centric <- 0.25*2225/381
-Health_Centric <- 0.25*2225/421
+Combing all amenity types into the 15-Minute City concept is also a matter of gender equality.
+There is a branch of research that encompases this critique: Mobility of Care.
+This framework encourages the explicit consideration of destination types that serve unpaid labour (care destinations) into transport planning (as opposed to only considering destinations with paid work).
+Not considering these care destinations have travel time (and hence quality of life) implications on those who do care activites, commonly women and especially disadvantaged women.
+The consideration of _what_ amenity type is a gender issue.
 
-#adds up to 2225 opportunities  
-0.25*2225/1265*(1265) +  0.25*2225/158*(158) + 0.25*2225/381*(381) + 0.25*2225/421*(421)
+From this position, this study aims to holistically investigate the ease of interacting with proximate amenities (accessibility) from the perspective of the 15-Minute City and the Mobility of Care framework.
+As mentioned, accessibility analysis have typically focused on employment (X) and recent 15-Minute City flavoured analysis focus on 'lower-order' amentities.
+We focus on a case study of the mid-sized City of Hamilton, Canada, and aim to classify areas by what degree they may be considered a '15-Minute Neighbourhood" based on the relative quantity of amenity access and diversity of access.
+We offer a novel methodology of holistically considering all proximate-amenities, offer empirical findings, and contribute theoretically to the ongoing debate on the definition of a 15-Minute City by foregrounding care destinations.
 
-#adding the weighted care destination counts -- reweighted such that the 4 categories each make up 25% of the 2225 points.
-care_dest <- care_dest |>
-          mutate(Dest_count_weighted = case_when(Care_Category == "Dependent-centric" ~ Dependent_Centric,
-                                                 Care_Category == "Errand-centric" ~ Errand_Centric,
-                                                 Care_Category == "Grocery-centric" ~ Grocery_Centric,
-                                                 Care_Category == "Health-centric" ~ Health_Centric))
+The objectives of the paper are as follows:
 
-care_dest$Dest_count_weighted |> sum()
-```
+1)	To examine the potential accessibility to, and diversity of, care destinations in the case study.
+We identify to what degree are neighbourhoods caring and/or complete.
+To do so, we make use of self organizing maps (SOM), an unsupervised machine learning data-dimensionality reduction technique to identify caring and complete neighbourhoods.
+2)	We then apply the same SOM methodology to determine profiles of population groups to understand who enjoys these neighbourhoods and who does not.
 
-# Abstract
+In this way, destinations outside of the home involved in caring activities has important implications for spatial and transportation planning; for all people but especially for all women and those from lower income households who complete the most mobility of care trips [@ravensbergen2023exploratory].
 
-The 15-Minute City is a normative conceptualisation gaining ground in urban planning: it frames neighbourhoods as responsive to human needs and environmental sensibilities, where most daily necessities can be reached within a 15-minute walk or bike ride [@Allam2022].
-As a related tool, accessibility measures (the ease of reaching opportunities) are increasingly important amongst transport planners aiming to foster just and sustainable cities [@vale2023accessibility].
-Both the 15-Minute City and accessibility measures are flexible enough to consider all destination types holistically however, gendered examinations have been lacking in the literature.
-For instance, accessibility analyses have historically focused on employment-centric and discretionary travel, types of travel more frequent for working-age and higher-income men.
-
-To counter this masculinist bias, this study investigates a way to gender-mainstream the 15-Minute City through a care lens [e.g., @lawWomenTransportNew1999; @cresswell_gendered_2008; @Oestergaard1992; @levy1991towards; @levy2013travel] supported by the Mobility of Care framework (coined by @sanchezdemadariagaMobilityCareIntroducing2013).
-Mobility of Care emphasizes the importance of travel to unpaid work (care trips) in contrast to the better-studied travel to employment and leisure.
-While all three trip types (work, care, and leisure) are essential, care trips are often relatively shorter-distance, proximate to residential/work/school, and comprise approximately 30% of adults’ daily trips [@sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory; @mejia2021mobility]; fitting well within the 15-Minute City conceptualisation.
-
-Our study provides an empirical example that maps the 15-Minute City onto the Mobility of Care framework.
-Specifically, it identifies which areas in Hamilton, Canada are 'caring 15-minute neighbourhoods'.
-To do so, a database of care destinations is created using secondary data.
-In this database, care destinations include all places associated with sustaining household tasks needed for the reproduction of life namely: shopping (e.g., groceries), errands (e.g., libraries), health (e.g., dentist), and caring for dependents (e.g., schools).
-This database is used to estimate the number and mix of care-destinations that can be reached within a 15-minute walk- and cycling- sheds from census centroids.
-Typologies are generated to illustrate which neighbourhoods can and to what degree facilitate 15-minute access to care.
-
-Results indicate only a few neighbourhoods outside of the downtown core are 'care-complete', i.e., contain a sufficiently high mix of care destinations from all categories and sub-categories.
-However, some neighbourhoods are almost 'care-complete' and provide 15-minute access to some care categories.
-Our study frames these neighbourhoods on the continuum of 'caring' and in need of further intervention.
-The quantitative investigation conducted provides a high-level picture of what neighbourhoods (and their underlying land-use) are connected to transport infrastructure that can support reaching care-destinations.
-
-Taken together, this study provides a theoretical bridge to connect 15-Minute Cities, accessibility analysis and Mobility of Care framework for the purpose of informing policy choice aimed to encourage just and sustainable mobility.
-
-# Introduction
-
-Caring is one of the most fundamental activities in societies, which has allowed humanity to survive and evolve.
-It involves activities targeted to meet physical, psychological, and emotional needs of others [@ilo_care_2018].
-Yet, it is one of the worst devalued, underpaid, and unequal activities worldwide.
-Traditionally, caring activities have remained on women’s shoulders.
-Even though the society has evolved, and nowadays women attain higher educational levels and tend to work out of home more than ever before, the bulk of care remains (at any stage of her lifecourse) heavily on her shoulders.
-The Second Job or the Second Shift remains in all industrial societies, mostly women’s work (Hayden, 1982 (?Citation)) [@hochschild_second_2012].
-
-This unequal share of caring responsibilities leads to gendered differences in among others, career development, differences in contract types and decision-making positions, unbalanced share of women in certain professions, gender pay gap, time poverty, as recognised by different international organisations (European Institute for Gender Equality, 2016 (?Citation)) [@ilo_care_2018].
-In fact, already the first feminists in the United States observed that domestic labor was the most relevant basic cause of women’s inequality [@gordon_womans_1976] (Hayden, 1982 (?Citation)).
-
-According to the [@ilo_care_2018], women and girls perform more than three-quarters of the total amount of unpaid care work and two-thirds of care workers are women.
-The caring economy is part of the women-dominated industries, it is characterised by the evaluative discrimination, which is only one of the dimensions of the gender pay gap [@mejiadorantes_discussing_2019].
-Especially in the case of domestic workers, care workers are most of the times women and migrants’ workers, with poor working conditions, and many times associated with informal jobs which imply lack of labour rights, no social and health protection, susceptible of exploitation, one of the poorest working conditions across the care economy as the ILO observes [@ilo_care_2018].
-Care workers becomes disrespected, leading to a devaluation of their activity, and consequently remains as one of the less well-paid occupations a person can undertake.
-
-If we turn our attention to the urban planning perspective, Caring is a major human activity that may take form in all urban settings, in all transport modes, all times a day and year, and yet it is not described as neither origin nor destination, and until recently, it was inexistent.
-
-Researchers from various fields have highlighted that western modern industrial societies have imposed a physical separation of the private sphere from the public one, splitting production and reproduction, public and private spaces or simply work and home, concentrating on the economic outcome of being mobile [@cresswell_gendered_2008; @lawWomenTransportNew1999; @levy2013travel; @little_gender_1994; @tronto_toward_1990].
-The man with legal rights, paid labour public sphere and woman in the private sphere, dependent, with family duties, personal relations and unpaid labour [@tronto_toward_1990].
-In this line, @cresswell_gendered_2008 described how masculinity was related to mobile and active, whereas femininity to static.
-As an example of this dichotomy, transport planning historically framed non-work/school trips as 'non-mandatory' activities, downgrading these activities in comparison to work.
-
-Women have consistently been found to complete more mobility of care trips than men.
-In one study, mobility of care comprised 32% of women’s daily trips compared to 28% of men’s.
-While this gap was significant, it was found to be far greater in lower income households where women complete 20% more care trips than men [@ravensbergen2023exploratory].
-This gendered division in mobility of care is perhaps unsurprising as women tend to complete a disproportionate amount of household-serving labour, such as caring for children and other dependents, preparing meals, cleaning, and running errands, when compared to men [@RN4].
-This gender gap varies geographically [@RN5].
-In Canada, adult women complete 1.9 hours, on average, more of unpaid care than men every day [@RN4].
-This labour is essential, and a large component of daily life.
-
-From this research motivation, Sánchez de Madariaga coined the term ‘mobility of care’ in 2013 to refer to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands [@sanchezdemadariagaMobilityCareIntroducing2013].
-While decades of research had examined these unique household-serving trips, her seminal work was the first to consider all these trips as one category.
-In doing so, she showed how mobility of care is a significant proportion of daily travel.
-Indeed, in her original work, and in the studies that have followed, it appears that approximately 30% of adults’ trips are for care purposes [@sanchezdemadariagaMobilityCareIntroducing2013; @sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory].
+ In fact, early feminists in the United States had already observed that domestic labor was the single most relevant cause of women’s inequality in society [@gordon_womans_1976; @haydenGrandDomesticRevolution1982].
+....
 Sánchez de Madariaga not only shows how important these mobility of care trips are, but also highlights the ways in which “mobility of care is systematically under-represented in any analysis of urban transport” (p.
@@
 Transport surveys and tools do not directly capture mobility of care, which re-enforces the idea that these trips are not a significant part of daily mobility.
+....
+In this way, destinations outside of the home involved in caring activities has important implications for spatial and transportation planning; for all people but especially for all women and those from lower income households who complete the most mobility of care trips [@ravensbergen2023exploratory].
 
-Cities have simplified individuals as a universal and neutral entity.
-As a result, public spaces are differently experienced depending on our multiple identities.
-Indeed, asymmetric in the distribution of time of citizens is deeply anchored with the public and private spaces organization within cities.
-In this light, we focus on the potential access to destinations of care through an empirical case study of the mid-sized city of Hamilton, Canada.
-We acknowledge that caring is composed by actors and activities but aim to spotlight the role that the urban environment plays on our daily lives, on easing our participation and on exercising our rights to the city as individuals with multiple identities.
-Our work hopes to contribute to a fruitful discussion that may trigger urban and transport policy interventions that increase gender-equity in caring activities among all actors.
+Care also has important implications for spatial and transportation planning.
+Indeed, researchers from various fields have highlighted that western modern industrial societies have imposed a physical separation of the private sphere from the public one, splitting production and reproduction, public and private spaces or simply work and home, concentrating on the economic outcome of being mobile [@cresswell_gendered_2008; @lawWomenTransportNew1999; @levy2013travel; @little_gender_1994; @tronto_toward_1990].
+Men tend to be associated with legal rights, paid labour, and the public sphere while woman are associated with the private sphere, dependents, family duties, personal relations and unpaid labour [@tronto_toward_1990].
+In this line, @cresswell_gendered_2008 described how masculinity has been constructed as mobile and active, whereas femininity is static.
+As an example of this dichotomy, transport planning has historically framed non-work/school trips as 'non-mandatory' activities, downgrading these activities in comparison to mandatory trips like work and school (XXX).
+In more recent years, planners have begun to resist the spatial separation of the public and private sphere to promote for sustainable cities.
+Proximity planning has become more normalized, a popular example is the 15-minute neighbourhood i.e., neighbourhoods where residents can reach all necessary destinations within a 15-minute active transport trip.
+The 15-minute neighbourhood concept, however, has yet to foreground care.
 
-<!-- We experience our urban environment depending on our personal characteristics, such as socioeconomic level, education, profession, household responsibilities, physical or mental disabilities, and very importantly, by gender.
-Furthermore, our living environment is also influenced by historical, economic, social, cultural, political, religion and environmental factors.
-All these issues have an influence on the way we live and experience the city, on how we participate in the community and on our social relations in the public and private space, on how we invest in our socio-economic development, and finally on our well-being.
--->
-
-<!-- Cities have simplified individuals as a universal and neutral entity.
-As a result, public spaces are differently experienced depending on our multiple identities.
-Indeed, the distribution of time of citizens is deeply anchored with the public and private spaces organization within cities.
-In fact, the asymmetries in the distribution of time of individuals are rooted on city spaces, distances to reach them, transport alternatives and different perceived fears, emotions and constraints, along with the roles typically associated with men and women, together with the gendered division of labour [@villagran_urban_2020].
-Furthermore, these authors emphasize that it is a global reality that women invest a greater amount of time to care and to unpaid work and that it is socioeconomically stratified.
--->
-<!-- Many studies worldwide have documented how gender influences mobility patterns.
-Men commute longer distances and present direct travel patterns, whereas women present chained trips to accomplish on their way many different activities (for example, dropping children off at schools or daycare, run errands or accompany family members to medical services).
-For these reasons it has been documented that women tend to have shorter but more complicated transport patterns.
-They also tend to rely more on public transport whereas men use more private transport [@cresswell_gendered_2008; @levy2013travel; @mejia2021mobility; @the_world_bank_handbook_2020].
-Sánchez de Madariaga uses the concept “mobility of care” to recognize, measure, make visible, value and account for all the travel associated to caring and unpaid tasks for the household [@sanchezdemadariagaMobilityCareIntroducing2013].
--->
-
-## Feminist 15-minute neighbourhoods
-
-Within the backdrop of urban planning and 'mobility of care', we connect our work to the 15-minute neighbourhood concept.
-
-The spatial structure of cities and regions has been described in different ways and its advantages and disadvantages has been discussed with regard to a wide variety of issues, for example, regarding health, the environment, economic, social and equity impacts.
+<!-- The spatial structure of cities and regions has been described in different ways and its advantages and disadvantages has been discussed with regard to a wide variety of issues, for example, regarding health, the environment, economic, social and equity impacts.
 After the traditional monocentric vision of the city with a core Centre Business District (CBD), other alternatives appeared, especially due to the expansion of cities caused by the automobile and motorized transport.
@@
 This concept may have different interpretations depending on the spatial and geographical scales [@hall_polycentricity_2009].
+-->
 
-In the last decades, the need for more sustainable, healthier, and liveable cities has received more attention.
-Planners and policy makers have proposed different alternatives where residents reach “essential activities” within walkable or cycling distances, reducing the need for using motorised transportation modes.
-The 15-minutes cities, 20-minutes or x-minutes cities, have been long under the spotlight.
-This approach has also received the name of Chrono-urbanism or the “cities of proximity” [@logan_xminute_2022; @moreno_introducing_2021].
+Within the backdrop of urban planning and ‘mobility of care’, we connect our work to the 15-minute neighbourhood concept.
 
-In the international context, there are different examples that highlight the importance of good access to different services and amenities within walkable distances.
-Probably one of the first cities to adopt the proximity approach was Portland.
-The Portland Plan [@city_of_portland_20minute_2010] was adopted by the City Council on April 25, 2012 aiming at fostering an inclusive city development for the following 30 years.
-The plan is based on the following principles: prosperity, education, health, and equity.
-The progress report states that equity should be achieved, an Office of Equity and Human Rights was established focusing on equitable service delivery to all residents.
-Nonetheless the progress report only measures progress on topics related to racial equity or people with disabilities [@portland_government_portland_2017].
+Both the 15-minute neighborhood and accessibility measures are flexible enough to consider all destination types holistically, however, gendered examinations have been lacking in the literature.
+For instance, accessibility analyses have historically been employment-centric even when adopting a gendered perspective [@paez_jobs_2013].
+Some 15-minute city analyses focus on care destinations, but they seem to not put care at the forefront, or to consider but one care destination (e.g., [@hosford15minuteCityReach2022] on grocery stores).
+Though few 15-minute neighbourhood analyses are explicitly feminist, their potential to build feminist communities have been noted outside of academia.
+For instance, Women in Urbanism have noted how women may benefit more than men from this 15-minute approach to planning.
+For instance, women tend to walk and use transit more than men, as well as to make more frequent, but shorter trips -and often as part of a chain (i.e., with stops along the way).
+These trip characteristics are all well served in 15-minute frameworks.
+15-minute neighbourhoods also promote walkability, which results in more “eyes on the streets” and its associated safety, an important enabler of mobility for women and girls.
+Finally, because women tend to earn less money than men and cars are a large financial burden, planning paradigms that reduce car dependency, such as 15 minute cities, may on average benefit women more than men [@Bertolo2023].
 
-In Paris, the adoption of the 15-Minute city put into the spotlight this approach.
-In Paris, they targeted six social functions that had to be easily accessible at any part of the city: housing, work, health care, groceries, education and leisure [@ville_de_paris_paris_2022].
-In 2021, Ottawa also adopted a 15-Minutes approach in the new City of Ottawa Official Plan [@ville_dottawa_quartier_2021].
-
-More recently, @moreno_introducing_2021 considers four dimensions on how the 15-minute city framework: density (in terms of people per sq.km); diversity (including mixed-land use and diversity of people; the temporal and spatial proximity to essential services (residential areas, work, commercial areas, education centres, health facilities and other basic institutions) and digitalisation.
-
-It can be observed that the feminist perspective of the cities of proximity is still underestimated.
-Only few examples exist that specifically target this issue [@gil_sola_choose_2022; @macintyre_her_2022].
-
-Even if we are on the right direction to facilitate caring activities, develop human-centred urban areas, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities and gender.
-Similarly, it is necessary to provide alternatives to measure progress on this important issue.
-On the other hand, an intersectional approach to understand how different personal identities interact in our daily day activities in the cities is necessary.
-There is still room for improvement.
-
-## Mobility of care and mode-constraints
-
-Given that almost one third of daily trips are for care purposes [@sanchezdemadariagaMobilityCareIntroducing2013], it is important to plan for mobility of care.
-There are also equity considerations to consider: as discussed, women, and especially low-income women, complete this travel more than men.
-Planning for mobility of care may also warrant different considerations because mobility of care trip characteristics differ from other trips, notably those of the commute to work.
-For instance, many of these trips are made through trip-chains, i.e., when a non-work-related stop is made during the commute.
-Women have been found to do so more than men, which is understood as being due to their need to balance unpaid-household serving care and paid employment (e.g., picking up groceries or children on the way home from work) [@ravensbergen2023exploratory; @scheinerWomenComplexDaily2015].
-
-Further, compared to the trip to work, mobility of care trips are more frequently completed by car and by foot, and less frequently done by transit or bicycle [@ravensbergen2023exploratory].
-Evidence for why this modal difference is observed exists.
-For instance, it is possible that walking is more common for care than for work trips because care trips tend to be short [@ravensbergen2023exploratory].
-In other words, if care destinations are closer to the home, it is more likely that these destinations are walkable.
-Ample literature examines the benefits and barriers to walking care destinations, such as schools (e.g., [@RN7;@RN8;@RN9]) and grocery stores (e.g., [@RN10;@RN11]).
-Little of this work uses the mobility of care framework.
-Instead, it tends to consider singular care destinations in research focusing on walkability.
+-- 
+# possible discussion
 
@@
 
-# Data and Methods
+## about employment.
 
-## Case study context
-In Canada, women have become more educated and have an achieved an increased participation in the labour market [@government_of_canada_gender_2023].
+Though not considered, employment is a promiment destination type not considered within this work.
+However, as mentioned, care activites are often trip-chained.
+therefore, the availability of care destinations around work also matter.
+This is for future investigation.
+
+In Canada, women have become more educated and achieved increased participation in the labour market [@government_of_canada_gender_2023].
 In fact, the educational level of women has surpassed most men [@government_of_canada_women_2023].
 Reports observe that the gender pay gap has narrowed but disparities persist [@drolet_intersectional_2023;@pay_equity_office_gender_2023].
-According to the statistics, in Ontario between 1998 to 2022 the gender gap narrowed but still women make 87 cents on average for every dollar made by a man [@pay_equity_office_gender_2023].
-All reports observe that it affects women differently, women with disabilities, immigrants, indigenous or racialized suffer especially from the gender pay gap.
+For instance, in the province of Ontario women still make 87 cents on average for every dollar made by a man and by all reports women of inter-sectional marginalized identies are more impacted [@pay_equity_office_gender_2023].
 Similarly to other countries worldwide, the occupational segregation remains and women are segregated into lower paid job classes and there is a dominance of women in minimum wage jobs [@drolet_intersectional_2023].
@@
 
-- Next: describe hamilton's land-use context; communities which are urban, rural, suburban.
-Also describe the residential parcels summary statistics, ~140,000 points, aggregated at the level of Canadian Census DA.
-See population descriptive plots in @fig-Fig1:
-```{r plot-figure1, out.width=600, out.height=600}
-#| label: fig-Fig1
-#| fig-cap: "The number of residential parcels per DA in 2020 (top) and the rate of population per dwelling per DA (middle) and prevalence of Low-income cut-offs after-tax (LICO-AT) per DA (bottom) retrieved from the 2021 Canadian Census.
-All scales in quartiles.
-Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics("data/pop_descriptive_plots.tiff")
+# possibel conclusions
+
+Cities have simplified individuals as a universal and neutral entity.
+As a result, public spaces are differently experienced depending on our multiple identities.
+Indeed, asymmetric in the distribution of time of citizens is deeply anchored with the public and private spaces organization within cities.
+In this light, we bring together the literatures on 15-minute cities and mobility of care by focusing on the potential access to destinations of care through an empirical case study of the mid-sized city of Hamilton, Canada.
+We acknowledge that caring is composed by actors and activities but aim to spotlight the role that the urban environment plays on our daily lives, on easing our participation and on exercising our rights to the city as individuals with multiple identities.
+Our work hopes to contribute to a fruitful discussion that may trigger urban and transport policy interventions that increase gender-equity in caring activities among all actors.
+
+```{r plot-figure6-create, eval=FALSE}
+plot1 <- ggdraw() + draw_image(glue::glue(here::here(), "/figures/access_sc_plot_2.png"))
+plot2 <- ggdraw() + draw_image(glue::glue(here::here(), "/figures/access_sc_plot_1.png"))
+plot3 <- plot_grid(plot1, plot2)
+
+ggsave(glue::glue(here::here(), "/figures/access_sc_plot_combo.png"), dpi=300)
 ```
 
-## Estimated travel times
-- Next: describe the estimated 15-minute walk travel times using {r5r}, from all residential parcel points to all care destinations assuming default walking speed and OSM road network for Hamilton.
-Provide summary statistics.
-
-## Care destination dataset
-A spatial dataset of care destinations for Hamilton was compiled; sources include provincial and municipal open data portals [@governmentofontarioOntarioDataCatalogue2023; @cityofhamiltonOpenData2023], Data Axle, a consumer dataset compiled of businesses and companies within Canada [@axledataConsumerData2023] and manual retrieved through Google Maps.
-Each destination was categorized based on the specific type of care being accessed following the mobility of care research by Sanchez de Madariaga & Zucchini (2019) [@sanchezdemadariagaMeasuringMobilitiesCare2019].
-Initial categories include child, elder,  grocery, health, and errand -centric destinations.
-For the purpose of this study, child- and elder- centric are combined to represent a 'dependent-centric' category.
-Category sources of data and preparation notes are detailed in @tbl-Tbl1.
-Their spatial distribution and sub-categories are visualised in @fig-Fig2.
-
-```{r creating-table-content}
-dat <- data.frame(matrix(c("Child-centric",
-                           "[@opendatahamiltonEducationalInstitutions2022; @governmentofontarioOntarioDataCatalogue2023; @cityofhamiltonChildCareRegistry2023; @opendatahamiltonParksDatabase2022; @opendatahamiltonRecreationCommunityCentres2022]",
-                           "Schools, daycares, and community centres, recreation centres, and parks: 1,190 locations are included.
-After manual review, all locations that typically do not serve children were removed including: Post-Secondary, Adult-Learning Centres, Group Homes, and Foster Care Centres.
-Further, through examination some Section 23 institutions defined as _“centres for children who cannot attend school to meet the needs of care or treatment, and rehabilitation”_ [@governmentofontarioministryofeducationFundingEducationCommunity2023], were kept due to their innate connection to care.",
-                           "Elder-centric",
-                           "[@opendatahamiltonRecreationCommunityCentres2022; @ontarioministryofhealthgeohubMinistryHealthService2023]",
-                           "Senior centres, long-term care homes, and retirement homes: 75 destinations are identified.",
-                           "Grocery-centric",
-                           "[@axledataConsumerData2023]",
-                           "Grocery stores, namely a place a household could buy groceries ranging from convenience stores to large retail stores: 381 destinations are identified.
-Data is filtered by Company Name, Suite Number, Address, City, Province, Phone Number and Postal Code.
-The type was then identified e.g., grocers specialty foods, grocers retail, grocer health food, grocer wholesale, grocer curbside, grocer delicatessen wholesale, grocer convenience.
-Data was crossreferenced to ensure all included locations were operational and legitimate grocery stores.",
-                           "Health-centric",
-                           "[@ontarioministryofhealthgeohubMinistryHealthService2023; @hamiltonniagarahaldimandbranthealthlineWalkInMedicalClinics2023]",
-                           "Hospitals, pharmacies, clinics, and dentist offices: 421 destinations are identified.
-Hospitals and pharmacies were retrieved while clinics and dentistry clincs were manually scraped from a healthcare services database and checked via Google Maps to remove non-operational locations and confirm dentistry-orientation.
-",
-                           "Errand-centric",
-                           "Hamilton libraries [@opendatahamiltonLibraries2022], post office locations [@axledataConsumerData2023;@canadapostPostOfficeLocator2023], and datasets of all national bank chains [@bankofmontrealBankMontrealBranch2023; @thehongkongandshanghaibankingcorporationlimitedhsbcHSBCBranchABM2023; @nationalbankNationalBankBranches2023; @royalbankofcanadaRBCBranchABM2023; @scotiabankScotiabankABMBranch2023; @thetorontodominionbankTDBankBranch2023].",
-                                                      "Libraries, post offices, and banks: 158 destinations are identified.
-Post offices are retrieved from a mix of databases, and duplicates are removed.
-Banks are also derived from Data Axle and then cross-referenced to ensure data quality with a Bank Locator website for all national banking firms."
-                           ), ncol = 3, byrow = T))
+```{r plot-figure6, out.width=800, out.height=800}
+#| label: fig-Fig6
+#| fig-cap: "The 5 resulting superclusters from the SOM output (left) and the associated dendrogram structure (right)."
+#| 
+knitr::include_graphics(glue::glue(here::here(), "/figures/access_sc_plot_combo.png")) 
 ```
-
-```{r creating-table}
-#| tbl-cap: "Details on the preparation and data sources of care destinations."
-#| label: tbl-Tbl1
-set_flextable_defaults(font.size= 10,
-                       font.family = NULL,
-                       text.align = 'justify',
-                       table.layout = 'fixed')
-
-flextable(dat) %>%
-  ftExtra::colformat_md() %>%
-  compose(
-    i = 1, j = c(1,2,3), part = "header", value = as_paragraph(c("Care category", "Sources", "Data preparation notes"))) %>%
-  width(j = 1, width = 1.7, unit = "cm") %>%
-  width(j = 2, width = 4.3, unit = "cm") %>%
-  width(j = 3, width = 11, unit = "cm")
-```
-
-```{r plot-figure2, out.width=600, out.height=600}
-#| label: fig-Fig2
-#| fig-cap: "The locations of care destinations in Hamilton separated by the author-generated categories of: dependent-, errand-, grocery- and health- centric care categories.
-Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics("data/care_categories_plot.tiff")
-```
-
-For the purpose of this analysis and in absence of city-wide household preferences for care destinations, all locations are re-weighted to make each of the four categories conceptually equivalent i.e., each location within each category $c$ is `r word_comment("Thoughts on this", "weighted to equal")` $W_j^c = \frac{1}{C} \frac{\sum_{c} O_j^c}{O_j^c}$ where $C$ is the total number of categories and $O_j^c$ is a location at destination $j$ within a category $c$.
-The total sum of locations is preserved after weighting such that $\sum_{c} \sum_{j} O_j^c = \sum_{c} \sum_{j} W_j^c * O_j^c$.
-Conceptually, this simplistic re-weight makes destinations in categories with fewer locations more important and destinations in categories with more locations less important.
-The underlying assumption is the status-quo of locations per category is sufficient, e.g., many daycares (smaller capacities) and few hospitals (larger capacities) sufficiently serves  the populations' care needs.
-Re-weighting in this way ensures that populations who are near _assumed_ smaller capacity destinations such as daycares are not assigned a disproportionately high accessibility score and vice versa for populations near destinations with larger service capacities like hospitals.
-In absence of empirical data for amenity weight calibration, this methodological assumption is a limitation.
-
-- Dependent-centric (`r sum(care_dest$Care_Category == "Dependent-centric")` destinations each at `r Dependent_Centric`), 
-- Errand-centric (`r sum(care_dest$Care_Category == "Errand-centric")` destinations each at `r Errand_Centric`),
-- Grocery-centric (`r sum(care_dest$Care_Category == "Grocery-centric")` destinations each at `r Grocery_Centric`) and,
-- Health- centric (`r sum(care_dest$Care_Category == "Health-centric")` destinations each at `r Health_Centric`) .
-
-## Methods
-Cumulative opportunity method; 15-minute normative catchment for walk mode (and cycling, travel time matrix already calculated but to be added later) is used to represent the 15-minute neighbourhood concept.
-This has been done in previous literature such as @vale2023accessibility, @graellsCityCitiesMeasuring2021 and @gaglioneUrbanAccessibility15minute2022.
-
-In this work, the cumulative opportunity score is calculated for each parcel and aggregated by care category at the DA level.
-The Q3 number of opportunities accessibility in a 15 minute walk per care category in retained for each DA.
-So in this analysis, $i$ is both the parcel and DA spatial unit.
-To the author's this is one of the first 15-minute neighbourhood analyse with a focus on care destinations.
-
-**Cumulative opportunity accessibility** takes the following form:
-$$
-S^c_i=\sum_{j}W_j^{c}\cdot f(c_{ij})
-$$
-\noindent Where:
-
-- $i$ is a set of parcel point origin locations.
-- $j$ is a set of destination locations.
-- $c$ is a set of care categories.
-- $W^c_j$ is a number of weighted opportunities ($O_j^c$) of category $c$ at $j$.
-- $c_{ij}$ is the travel cost between $i$ and $j$.
-- $f(\cdot)$ is an impedance function of $c_{ij}$; within the cumulative opportunity approach, it is a binary function that takes the value of 1 if $c_{ij}$ is less than a selected value [@handyMeasuringAccessibilityExploration1997].
-- $S_{i}^c$ is the unconstrained accessibility, the sum of weighted opportunities reachable within $f(\cdot)$, at each $i$ for each $c$.
-
-Next, the Q3 (median + standard deviation) of all $\sum_{c}S_i^c$ values for each DA are taken and represented as $S^{c}_{DA}$.
-$S^{c}_{DA}$ is the upper quartile number of weighted care destinations that can be potentially interacted with within a 15 minute walk by parcels within a DA: an optimistic perspective on how many opportunities can be potentially reached `r word_comment("Thoughts on this choice instead of the median?
-I do this because the parcels are not population weighted, so 1 parcel = 1 travel time.
-There can be obviously multiple households in 1 parcel, so using Q3 makes some sense.", ".")`
-
-*Entropy* measure has been widely used to characterise the diversity of land-use mix in early work by @frankLinkingObjectivelyMeasured2005 based on the species diversity Shannon-Wiener index that expresses relative evenness throughout a sequence (this measure among other species diversity indices of the time are reviewed in @whittaker1972evolution).
-In the realm of urban studies and planning literature, it has since been widely used to characterise the diversity in land-use mix [@ewingTravelBuiltEnvironment2010a], especially in the context of active modes [@luUrbanDensityDiversity2017; @mavoaIdentifyingAppropriateLanduse2018] and suburban sprawl [@randallGISBasedLand2015].
-Work that has used diversity indices alongside accessibility analysis are scarce, though examples of its use as parameters within accessibility scores in the case of employment opportunities [@chengMeasuringUrbanJob2013; @daiIncorporatingJobDiversity2018] and more recently to describe the diversity in transit facilities [@yinIncorporatingFacilityDiversity2024] are present.
-To the authors knowledge, the use of an entropy measure to reflect the diversity in care destination amenity types and in context within a 15-minute neighbourhood, has yet to be done.
-
-In our work, each destination is a specific type $st$ (i.e., `r care_dest$Care_Category_Specific |> unique()`) as well as a care category $c$.
-When $S_i^c$ is aggregated for each category at the $DA$ it reflects how all opportunities from all four care categories can be reached.
-The four care categories have meaning to the mobility of care conceptualisation.
-However, it doesn't explicitly reflect the specific type of destination.
-In our work, the diversity of the specific type of destination is of interest, we calculate $S^{st}_i$ for each parcel $i$ for each $st$ type following the equation for $S^{st}_i$.
-This upper quartile is aggregated at the $DA$ to come up with a $S^{st}_{DA}$ `r word_comment("Thoughts on this choice?
-I think we need to consider both the category type and the specific type in this paper.
-We justify the categories so that's fair but I still think there's value in presenting the specific type  some how - this is what I think this diversity index is good for so I've presented that as an option.
-Note though, it'd be much more simple to just focus on calculating the $D^{st}_{DA}$ using the $S_i^c$ and not introducing the $S_i^{st}$ calculation like I've done..
-so I've made the analysis more complicated.
-Also destination weighting is based on the four categories - so that is distorting the diversity in certain ways.",".")`
-
-As such, the entropy measure is used to reflect the diversity of specific type of care destinations.
-The value ranges from 0 to 1, with 1 representing total evenness in the number of (weighted) specific type $st$ care opportunities can be reached at the aggregated spatial unit $DA$:
-
-$$
-D^{st}_{DA} = \frac{-\sum_{c} S_i^{st} / S^{st}_{DA} \times ln(S_i^{st} / S^{st}_{DA} )}{ln(ST)}
-$$
-\noindent Where:
-
-- $i$ is a set of parcel point origin locations.
-- $st$ is a set of specific type care categories. In our work they are `r care_dest$Care_Category_Specific |> unique()`.
-- $DA$ is a set of DA spatial units (to aggregate parcel point data).
-- $S_{i}^{st}$ is the unconstrained accessibility, the sum of weighted opportunities reachable within $f(\cdot)$, at each $i$ for each $st$.
-- $S^{st}_{DA}$ is the aggregated unconstrained accessibility at each $DA$. In our work, it is the upper quartile of $S_i^{st}$ values in a $DA$. 
-- $ST$ is the total sum of categories. In our work, this is equal to 14. 
-- $D^{st}_{DA}$ is the entropy index.
-
-# Results
-
-```{r plot-figure3, out.width=600, out.height=600}
-#| label: fig-Fig3
-#| fig-cap: "The cumulative opportunity measure.
-The number of care destinations that can be reached, per DA, within 15 min walk.
-Visualised as 'caring' (based on the opportunity reachable number; Caring is approximately the median) and if complete (if all care categories can be reached; this is an arbitrary selection).
-Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics("data/all_cats_copp_weighted_plot.tiff")
-```
-
-```{r plot-figure4, out.width=600, out.height=600}
-#| label: fig-Fig4
-#| fig-cap: "The diversity measure based on the _specific type_ of destination cumulative opportunity DA measure.
-Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics("data/all_cats_entropy_copp_weighted_plot.tiff")
-```
-
-```{r plot-figure5, out.width=600, out.height=600}
-#| label: fig-Fig5
-#| fig-cap: "A bivariate map of the cumulative opportuninities map (aggregated per category at DA - fig3) and the diversity measure (fig 4).
-Amenity diverse = diviersity index above the region median.
-Caring/Complete = accessibility value above the region median.
-The Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics("data/all_cats_bivariate_entropy_copp_plot.tiff")
-```
-
-BONUS!
-Per capita spatial availability.
-This is not discussed in the methods.
-The spatial availability calculation assumes population is 'population density' and weighted opportunity count.
-With this map we can start thinking about per capita availability of opportunities.
-```{r plot-figure6, out.width=600, out.height=600}
-#| label: fig-Fig6
-#| fig-cap: "Per capita spatial availability of care opportunities.
-Uses weighted care opportunities, population density as competition, and 15 minute binary cut-off.
-Aggregates parcel level spatial availability at the DA level (by summation) and divides by the DA population.
-Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics("data/all_cats_vvi_plot.tiff")
-```
-
-`r word_comment("If you're interested in how the unweighted results look like for the same maps see the appendix.
-I've also calculated the cycling travel matrices as well as the 'no dependent' i.e., the same analysis but now only considering 3 care categories: health, errand, and grocery.
-For the time being though let's focus on this.", " ")`
-
-# References
-
-::: {#refs}
-:::
-
-\pagebreak
-
-# Appendix
-
-```{r plot-figure7, out.width=600, out.height=600}
-#| label: fig-Fig7
-#| fig-cap: "UNWEIGHTED -- The cumulative opportunity measure.
-The number of care destinations that can be reached, per DA, within 15 min walk.
-Visualised as 'caring' (based on the opportunity reachable number; Caring is approximately the median) and if complete (if all care categories can be reached; this is an arbitrary selection).
-Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics("data/all_cats_copp_plot.tiff")
-```
-
-```{r plot-figure8, out.width=600, out.height=600}
-#| label: fig-Fig8
-#| fig-cap: "UNWEIGHTED -- The diversity measure based on the _specific type_ of destination cumulative opportunity DA measure.
-Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics("data/all_cats_entropy_copp_plot.tiff")
-```

### 2024-09-27 data/04-bycategoriy-clustering.qmd
small edits and knitted the half done draft

@@
 
+Not caring - not complete (C7)
+Somewhat caring - not complete (C4)
+Somewhat caring - complete (C5)
+Caring - somewhat complete (C3)
+Caring - very complete (C6)
+Very caring - complete (C2)
+Very caring - very complete (C1)
+
 ```{r}
@@
                                max_label_val == SC4 ~ "c. Very caring - mostly complete (SC4)",
-                               max_label_val == SC5 ~ "j. not caring - somewhat complete (SC5)",
-                               max_label_val == SC6 ~ "e. Somewhat caring - mostly complete (SC6)",
+                               max_label_val == SC5 ~ "j. Somewhat caring - somewhat complete (SC5)",
+                               max_label_val == SC6 ~ "e. Somewhat caring - not complete (SC6)",
                                max_label_val == SC7 ~ "f. not caring - not complete (SC7)"))

### 2024-09-27 paper/Manuscript.qmd
small edits and knitted the half done draft

@@
 
-`r word_comment(Maybe Antonio to add? I need to fill this out a bit", " ")`
+`r word_comment("I need to fill this out. A small classification methods paragraph", ".")`
 In addition, a useful tool has been classification algorithms in the identification of spatial typologies and in profiling who may reside there.
@@
 regionlization techniques such as spatially constrained multivariate clustering [@burkeGeospatialAnalysisFramework2022] Use of tree diagrams!
-
-SOM is a type of artificial neural network (ANN) trained using unsupervised learning that produces a two-dimensional representation of the input space.
+....SOM is a type of artificial neural network (ANN) trained using unsupervised learning that produces a two-dimensional representation of the input space.
 It applies competitive learning by way of a neighbourhood function thereby preserving the topological properties of the input space and producing combinations of the input space that we can then intuitively profile.
@@
 The SOM output is typically examined through a dissimilarity dendogram and an associated dissimilarity variance explained plot to select an appropriately representative 'superclusters' [@VillaVialaneix2017; @victorianoTimeSpaceMoney2020].
-In this work, we refer to these superclusters as simply clusters.
 
-For the decision tree step, the cluster-classified parcels identified in the SOM step are used as _labels_ and socio-economic and demographic indicators available from the most recent Canadian Census related to the mobility of care literature are used as _features_.
-This step is used to profile the clusters in a data-drive way to explore who may reside in what cluster.
+For the decision tree step, the supercluster-classified parcels identified in the SOM step are used as _labels_ and socio-economic and demographic indicators available from the most recent Canadian Census related to the mobility of care literature are used as _features_.
+This step is used to profile the superclusters in a data-drive way to explore who may reside in what supercluster.
 To estimate the decision tree, the 'rpart()' function in the {rpart} package assuming default parameters for classification splitting along with each value being weighted by the population present in the associated DA [@Therneau2023].
@@
 
-# Results
+<!-- # Results -->
 
-## Overview of access to care destinations 
-How many care destinations can be reached within a 15-minute walk is summed for each of the four care categories defined in @bl-tbl1.
+<!-- ## Overview of access to care destinations  -->
+<!-- How many care destinations can be reached within a 15-minute walk is summed for each of the four care categories defined in @bl-tbl1.
 The median parcel value for each DA is visualised in @fig-Fig4.
@@
 Together, @fig-Fig4 and @fig-Fig5 represent the 15 variables that served as inputs into the SOM algorithm.
+-->
 
-```{r plot-figure4, out.width=600, out.height=600}
-#| label: fig-Fig4
-#| fig-cap: "The number of care destinations that can be reached within 15 min walk per care category for a median parcel in each DA.
+<!-- ```{r plot-figure4, out.width=600, out.height=600} -->
+<!-- #| label: fig-Fig4 -->
+<!-- #| fig-cap: "The number of care destinations that can be reached within 15 min walk per care category for a median parcel in each DA.
 The values are a summary of the 14 acessibility scores that served as inputs into the SOM.
-Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics(glue::glue(here::here(), "/figures/percats_copp_plot.png"))
-```
+Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]." -->
+<!-- knitr::include_graphics(glue::glue(here::here(), "/figures/percats_copp_plot.png")) -->
+<!-- ``` -->
 
-```{r plot-figure5, out.width=600, out.height=600}
-#| label: fig-Fig5
-#| fig-cap: "The diversity measure based on the proportion of care category spatial accessibility (figure 4).
+<!-- ```{r plot-figure5, out.width=600, out.height=600} -->
+<!-- #| label: fig-Fig5 -->
+<!-- #| fig-cap: "The diversity measure based on the proportion of care category spatial accessibility (figure 4).
 These values are a summary of the 15th input into the SOM.
-Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_entropy_copp_plot.png")) 
-```
+Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]." -->
+<!-- knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_entropy_copp_plot.png"))  -->
+<!-- ``` -->
 
-## Identification of (in)completely (un)caring neighbourhoods
+<!-- ## Identification of (in)completely (un)caring neighbourhoods -->
 
-Notably, in @fig-Fig4, the concentration of opportunities for all category types is centered in the downtown of Hamilton, with the highest concentration being in the lower city (near the lake shore) and pockets of the high concentration further from the shoreline.
+<!-- Notably, in @fig-Fig4, the concentration of opportunities for all category types is centered in the downtown of Hamilton, with the highest concentration being in the lower city (near the lake shore) and pockets of the high concentration further from the shoreline.
 Grocery-centric destinations appear to be the most concentrated, followed by health-centric and dependent-centric caring destinations.
@@
 In @fig-Fig5, areas that have high care accessibility tend to have high diversity as well, though there are exceptions in pockets of the city outside the downtown core that have lower but equally low levels of accessibility for all care categories.
+-->
 
-Based on the SOM classification methodology discussed and the 15 variables represented in @fig-Fig4 and @fig-Fig5, 7 clusters are identified.
-As diagnostics for the selection of the number of clusters, the dissimilarity-index-based dendrogram with the proportion of parcels represented in each cluster alongside the variance explained plot is visualised in @fig-Fig6.
-```{r plot-figure6, out.width=800, out.height=800}
-#| label: fig-Fig6
-#| fig-cap: "The 7 resulting clusters from the SOM the 10 by 10 grid output represented in a dendrogram (left) and the variance unexplained (right)."
-#| 
-knitr::include_graphics(glue::glue(here::here(), "/figures/all_access_sc_plot_small.png")) 
-```
+<!-- Based on the SOM classification methodology discussed and the 15 variables represented in @fig-Fig4 and @fig-Fig5, 7 superclusters are identified.
+As diagnostics for the selection of the number of superclusters, the dissimilarity-index-based dendrogram with the proportion of parcels represented in each supercluster alongside the variance explained plot is visualised in @fig-Fig6.
+-->
+<!-- ```{r plot-figure6, out.width=800, out.height=800} -->
+<!-- #| label: fig-Fig6 -->
+<!-- #| fig-cap: "The 7 resulting superclusters from the SOM the 10 by 10 grid output represented in a dendrogram (left) and the variance unexplained (right)." -->
+<!-- #|  -->
+<!-- knitr::include_graphics(glue::glue(here::here(), "/figures/all_access_sc_plot_small.png"))  -->
+<!-- ``` -->
 
-To characterise the clusters, a table providing summary statistics of each cluster is detailed in @tbl-tbl2:
-```{r creating-table2}
-summary_df_ALL <- access_15_wide |> select(c(starts_with("total_"), div_allcats)) |>
-  group_by() |> 
-  summarise(across(everything(), list(median = median, sd = sd))) %>% 
- mutate_if(is.numeric, round, digits =1) |>
-  mutate_all(as.character)  
-summary_df_ALL <- cbind(data.frame(scluster = "TOTAL"),summary_df_ALL)
+<!-- To characterise the superclusters, a table providing summary statistics of each supercluster is detailed in @tbl-tbl2: -->
+<!-- ```{r creating-table2} -->
+<!-- summary_df_ALL <- access_15_wide |> select(c(starts_with("total_"), div_allcats)) |> -->
+<!--   group_by() |>  -->
+<!--   summarise(across(everything(), list(median = median, sd = sd))) %>%  -->
+<!--  mutate_if(is.numeric, round, digits =1) |> -->
+<!--   mutate_all(as.character)   -->
+<!-- summary_df_ALL <- cbind(data.frame(scluster = "TOTAL"),summary_df_ALL) -->
 
-summary_df <- access_15_wide |> select(c(scluster, starts_with("total_"), div_allcats)) |> 
-  mutate(scluster = case_when(scluster == 1 ~ "A",
-                              scluster == 2 ~ "B",
-                              scluster == 3 ~ "D",
-                              scluster == 4 ~ "F",
-                              scluster == 5 ~ "E",
-                              scluster == 6 ~ "C",
-                              scluster == 7 ~ "G")) |>
-  group_by(scluster) |> 
-  summarise(across(everything(), list(median = median, sd = sd))) %>% 
- mutate_if(is.numeric, round, digits =1) |>
-  mutate_all(as.character)
+<!-- summary_df <- access_15_wide |> select(c(scluster, starts_with("total_"), div_allcats)) |>  -->
+<!--   mutate(scluster = case_when(scluster == 1 ~ "A", -->
+<!--                               scluster == 2 ~ "B", -->
+<!--                               scluster == 3 ~ "D", -->
+<!--                               scluster == 4 ~ "F", -->
+<!--                               scluster == 5 ~ "E", -->
+<!--                               scluster == 6 ~ "C", -->
+<!--                               scluster == 7 ~ "G")) |> -->
+<!--   group_by(scluster) |>  -->
+<!--   summarise(across(everything(), list(median = median, sd = sd))) %>%  -->
+<!--  mutate_if(is.numeric, round, digits =1) |> -->
+<!--   mutate_all(as.character)  -->
 
-merge_every_two_columns <- function(df, num_rows) {
-  # Loop through every second column starting from column 2
-  for (col in seq(2, ncol(df), by = 2)) {
-    if (col < ncol(df)) {  # Ensure the column has a next one to merge with
-      for (i in 1:num_rows) {
-        df[i, col] <- paste0(df[i, col], " (", df[i, col + 1], ")")
-      }
-    }
-  }
-  return(df)
-}
+<!-- merge_every_two_columns <- function(df, num_rows) { -->
+<!--   # Loop through every second column starting from column 2 -->
+<!--   for (col in seq(2, ncol(df), by = 2)) { -->
+<!--     if (col < ncol(df)) {  # Ensure the column has a next one to merge with -->
+<!--       for (i in 1:num_rows) { -->
+<!--         df[i, col] <- paste0(df[i, col], " (", df[i, col + 1], ")") -->
+<!--       } -->
+<!--     } -->
+<!--   } -->
+<!--   return(df) -->
+<!-- } -->
 
-summary_df <- merge_every_two_columns(summary_df, 7) |>  select(-contains("sd"))
-summary_df_ALL <- merge_every_two_columns(summary_df_ALL, 1)  |>  select(-contains("sd"))
-summary_df <- rbind(summary_df, summary_df_ALL)
+<!-- summary_df <- merge_every_two_columns(summary_df, 7) |>  select(-contains("sd")) -->
+<!-- summary_df_ALL <- merge_every_two_columns(summary_df_ALL, 1)  |>  select(-contains("sd")) -->
+<!-- summary_df <- rbind(summary_df, summary_df_ALL) -->
 
-summary_df <- summary_df |> 
-  pivot_longer(!scluster) |>
-  pivot_wider(names_from =scluster )
+<!-- summary_df <- summary_df |>  -->
+<!--   pivot_longer(!scluster) |> -->
+<!--   pivot_wider(names_from =scluster ) -->
 
-summary_df_ALL <- summary_df_ALL |> 
-  pivot_longer(!scluster) |>
-  pivot_wider(names_from =scluster )
+<!-- summary_df_ALL <- summary_df_ALL |>  -->
+<!--   pivot_longer(!scluster) |> -->
+<!--   pivot_wider(names_from =scluster ) -->
 
-summary_df$name <- c("GROCERY TOTALS", "Convenience Store", "Grocer",
-                     "DEPENDENT TOTALS", "Community or Rec. centre", "Daycare or EarlyON","LTC or retirment home","Park","School","Senior centre",
-                     "HEALTH TOTALS","Dentist","Hospital or clinic","Pharmacy",
-                     "ERRAND TOTALS","Bank or ATM","Library", "Post office", 
-                     "Category diversity")
-```
+<!-- summary_df$name <- c("GROCERY TOTALS", "Convenience Store", "Grocer", -->
+<!--                      "DEPENDENT TOTALS", "Community or Rec. centre", "Daycare or EarlyON","LTC or retirment home","Park","School","Senior centre", -->
+<!--                      "HEALTH TOTALS","Dentist","Hospital or clinic","Pharmacy", -->
+<!--                      "ERRAND TOTALS","Bank or ATM","Library", "Post office",  -->
+<!--                      "Category diversity") -->
+<!-- ``` -->
 
-```{r creating-table2-pt2}
-#| tbl-cap: "Median and (standard deviation) of each SOM classified cluster by input variable and additional summary variables (indicated by ALL CAPITAL LETERS)."
-#| label: tbl-Tbl2
-#| 
-summary_df %>%
-  gt() %>%
-  tab_header(
-    title = " ",
-  ) %>%
-  cols_label(
-    name = " "
-  ) %>%
-  tab_style(
-    style = cell_borders(
-      sides = c("top"),
-      color = "black",
-      weight = px(1.5),
-      style = "solid"
-    ),
-    locations = cells_body(rows = c(1,4,11,15))
-  ) |>
-  tab_options(
-    table.font.size = 12,
-    table.width = pct(80)
-  )
-```
+<!-- ```{r creating-table2-pt2} -->
+<!-- #| tbl-cap: "Median and (standard deviation) of each SOM classified cluster by input variable and additional summary variables (indicated by ALL CAPITAL LETERS)." -->
+<!-- #| label: tbl-Tbl2 -->
+<!-- #|  -->
+<!-- summary_df %>% -->
+<!--   gt() %>% -->
+<!--   tab_header( -->
+<!--     title = " ", -->
+<!--   ) %>% -->
+<!--   cols_label( -->
+<!--     name = " " -->
+<!--   ) %>% -->
+<!--   tab_style( -->
+<!--     style = cell_borders( -->
+<!--       sides = c("top"), -->
+<!--       color = "black", -->
+<!--       weight = px(1.5), -->
+<!--       style = "solid" -->
+<!--     ), -->
+<!--     locations = cells_body(rows = c(1,4,11,15)) -->
+<!--   ) |> -->
+<!--   tab_options( -->
+<!--     table.font.size = 12, -->
+<!--     table.width = pct(80) -->
+<!--   ) -->
+<!-- ``` -->
 
-Notably, from A to F, the clusters are characterized by caring access and diversity of caring access (completeness). From the descriptive statistics of the clusters, interesting trends emerge: ...
+<!-- Notably, from A to G, the superclusters are characterized by caring access and diversity of caring access (completeness).
+From the descriptive statistics of the superclusters, interesting trends emerge: ...
+parcels that have an Errand-centric access of less than 0.5 opportunities (bottom quartile), a diversity score of less than 0.277 (bottom tertile), and any Dependent-centric access score of greater than 0.5 (and can take on any value of the remaining input variables).
+This technique provides a data-driven approach to supercluster formation which we then attribute meaning to through the derivation of the following labels:...
+In sum, we intuitively derive classification labels: -->
 
-Not caring - not complete (C7)
-Somewhat caring - not complete (C4)
-Somewhat caring - complete (C5)
-Caring - somewhat complete (C3)
-Caring - very complete (C6)
-Very caring - complete (C2)
-Very caring - very complete (C1)
+<!-- G: Not caring - not complete (C7) -->
+<!-- F: Somewhat caring - not complete (C4) -->
+<!-- E: Somewhat caring - complete (C5) -->
+<!-- D: Caring - somewhat complete (C3) -->
+<!-- C: Caring - very complete (C6) -->
+<!-- B: Very caring - complete (C2) -->
+<!-- A: Very caring - very complete (C1) -->
 
-parcels that have an Errand-centric access of less than 0.5 opportunities (bottom quartile), a diversity score of less than 0.277 (bottom tertile), and any Dependent-centric access score of greater than 0.5 (and can take on any value of the remaining input variables).
-This technique provides a data-driven approach to supercluster formation which we then attribute meaning to.
-
-```{r plot-figure7, out.width=600, out.height=300}
-#| label: fig-Fig7
-#| fig-cap: "The decision tree profiles of the SOM output."
-knitr::include_graphics(glue::glue(here::here(), "/figures/decision_tree_profiles_sc.png"))
-```
-
-The dendrogram demonstrating the supercluster class membership of each parcel in @fig-Fig7 is summarized, with our interpretation, as follows:
-- Very caring (mean category accessibility between 4 to 18.6 and total accessibility median of 46) - very complete (diversity index of mean 0.9 and median 0.90) (SCluster 2)
-- Caring (mean category accessibility between 2.1 to 9 and total accessibility median of 18) -  complete (div index of mean 0.81 and median 0.82) (SCluster 1)
-- Somewhat caring (mean category accessibility between 0.2 to 6 and total accessibility median of 13) - complete div index of mean 0.71 and median 0.74) (SCluster 4)
-- Somewhat caring (mean category accessibility between 0.2 to 6 and total accessibility median 7) - not complete (div index of mean 0.36 and median 0.40) (SCluster 3)
-- Not caring (mean category accessibility between 0 to 4.7 and total accessibility median of 3) - not complete (div index of mean 0.02 and median 0) (SCluster 5)
-
-To further visualize supercluster membership, we spatially demonstrate which supercluster best describes each parcel in @fig-Fig8 (top plot): the median parcel's supercluster membership per DA is tabulated and the majority membership is represented.
-Visually, the super clusters closely follow the quartile split of the accessibility to care destinations (all categories combined) and tertile split of the diversity scores (visualised in @fig-Fig5) `r word_comment("AN IDEA DISCUSSED: Overall, in @fig-Fig8 there are many pockets of good/great care accessibility+diversity of opportunities Hamilton.
+<!-- To further visualize supercluster membership, we spatially demonstrate which superclusters best describes each parcel in @fig-Fig8: <!-- Visually, the super clusters closely follow the quartile split of the accessibility to care destinations (all categories combined) and tertile split of the diversity scores (visualised in @fig-Fig5).
+-->  -->
+<!-- <!-- `r word_comment("AN IDEA DISCUSSED: Overall, in @fig-Fig8 there are many pockets of good/great care accessibility+diversity of opportunities Hamilton.
 These are objectively caring neighbourhoods (BY MEASURE).
@@
 the limitations of objective measures ", ".")`.
+--> -->
+<!-- ```{r plot-figure8, out.width=600, out.height=600} -->
+<!-- #| label: fig-Fig8 -->
+<!-- #| fig-cap: "The median parcel supercluster membership." -->
+<!-- knitr::include_graphics(glue::glue(here::here(), "/figures/SCLab_and_total_copp_plot.png")) -->
+<!-- ``` -->
 
-FIX THIS FIGURE'S LEGEND!!!!!!!
-```{r plot-figure8, out.width=600, out.height=600}
-#| label: fig-Fig8
-#| fig-cap: "The median parcel supercluster membership (top).
-The parcel median total accessibility (bottom).
-The parcel m Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics(glue::glue(here::here(), "/figures/SCLab_and_total_copp_plot.png"))
-```
-
-## Profiling of who resides in (in)completely (un)caring neighbourhoods
-To enhance the meaning of the superclusters beyond a spatial conceptualisation of "caring" and "complete", was _who?_ resides there.
+<!-- ## Profiling of who resides in (in)completely (un)caring neighbourhoods  -->
+<!-- To enhance the meaning of the superclusters beyond a spatial conceptualisation of "caring" and "complete", was _who?_ resides there.
 We implement another decision tree in which the population-weighted socio-demographic characteristics of the parcels (retrieved from the 2021 Canadian Census) are input variables and the resulting output are 'socio-domographic superclusters' of the caring/complete superclusters.
 The input variables included: % households below median household after-tax income, LICO prevalence, average number of children per household, % not in the labour force, Gini index on adjusted household after-tax income, % visible minority, % population aged 0 to 14, % single parent household, % who walk to work (relative to bike, care/truck/van, public transit and 'other'), % of owner household in core housing need (i.e., inadequate housing structure or paying higher than 30% income on housing), % of owner households with a mortgage, % of tenant households in subsidized housing, % of tenant households in core housing need, % no certification, diploma or degree (relative to population with high school, college, or university certificate, diploma or degree).
+-->
 
-Of all the variables, those that proved to be significant in profiling the five superclusters were only four variables: % households below median household after-tax income (bottom income), % population aged 0 to 14, % single parent household, and LICO.
+<!-- Of all the variables, those that proved to be significant in profiling the five superclusters were only four variables: % households below median household after-tax income (bottom income), % population aged 0 to 14, % single parent household, and LICO.
 See @fig-Fig9 that provides the splits.
@@
 However, we can interpret the socio-demographic superclusters (five that result) as labels to tell a narrative of who may reside in what caring/complete supercluster.
+-->
 
-```{r plot-figure9, out.width=600, out.height=300}
-#| label: fig-Fig9
-#| fig-cap: "The dendrogram of the supercluster demographic profiles.."
-knitr::include_graphics(glue::glue(here::here(), "/figures/access_sc_profiles.png"))
-```
-(NOTE: for reference, see the old bivariate map in the appendix for comparison)
+<!-- ```{r plot-figure9, out.width=600, out.height=300} -->
+<!-- #| label: fig-Fig9 -->
+<!-- #| fig-cap: "The dendrogram of the supercluster demographic profiles.." -->
+<!-- knitr::include_graphics(glue::glue(here::here(), "/figures/access_sc_profiles.png")) -->
+<!-- ``` -->
+<!-- (NOTE: for reference, see the old bivariate map in the appendix for comparison) -->
 
-<!--EXPLAIN THE FOLLOWING PARAGRAPH BETTER-->
-In @fig-Fig9 from right-to-left, parcels in DAs with 45% of the population below the median or lower (richer) are often classified with poor accessibility: either "Somewhat caring and not complete" SC3 or "Not caring and not complete" SC5.
+<!-- <!--EXPLAIN THE FOLLOWING PARAGRAPH BETTER--> -->
+<!-- In @fig-Fig9 from right-to-left, parcels in DAs with 45% of the population below the median or lower (richer) are often classified with poor accessibility: either "Somewhat caring and not complete" SC3 or "Not caring and not complete" SC5.
 Parcels within DAs with a higher proportion of single parent households (13% and above) are most often SC3: "Somewhat caring, not complete" and below are most often SC5: "Not caring, not complete".
 For parcels with higher than 45% of bottom-income (poorer), those with a high proportion of children (age 0 to 14) are also likely to be within the "Somewhat caring and not complete" SC3 - just like parcels in richer and more-single-parent DA counterparts.
+-->
 
-These findings related to potential child-care access are notable: single parents provide more care and hence may be more likely to choose, if they are able, more accessible neighbourhoods than the rest of the population in the same household income grouping.
+<!-- These findings related to potential child-care access are notable: single parents provide more care and hence may be more likely to choose, if they are able, more accessible neighbourhoods than the rest of the population in the same household income grouping.
 In a Canadian and U.S.
@@
 Most lone-families are headed by female-lone parent.
-(AS TO ADD IN HAMILTON SPECIFIC NUMBERS)
+(AS TO ADD IN HAMILTON SPECIFIC NUMBERS) -->
 
-Interestingly, parcels in DAs with lower than 13% children population demonstrate a better accessibility to care destination story.
-<!--PERHAPS A MISMATCH ON WHO NEEDS / WHO GETS CARE ACCESSIBLE DESTINATIONS-->
+<!-- Interestingly, parcels in DAs with lower than 13% children population demonstrate a better accessibility to care destination story.
+<!--PERHAPS A MISMATCH ON WHO NEEDS / WHO GETS CARE ACCESSIBLE DESTINATIONS-->  -->
 
-Parcels within DAs that constitute more than 45% population below median household income (poorer) - especially above 9.8% LICO (low income cut-off prevalence - likely to be the poorest) are often within "Very caring and very complete" SC1.
+<!-- Parcels within DAs that constitute more than 45% population below median household income (poorer) - especially above 9.8% LICO (low income cut-off prevalence - likely to be the poorest) are often within "Very caring and very complete" SC1.
 <!-- -maybe people who can’t afford a car choose to live here?
+-->  -->
+
+<!-- Parcels located in DAs with the same lower median income and child population but have _lower_ than 9.8% LICO (not the poorest) also have good accessibility, as they are most often in "Caring and complete" SC1 as well as being in most often in the "Somewhat caring and complete" SC4 relative to other parcels.
 -->
 
-Parcels located in DAs with the same lower median income and child population but have _lower_ than 9.8% LICO (not the poorest) also have good accessibility, as they are most often in "Caring and complete" SC1 as well as being in most often in the "Somewhat caring and complete" SC4 relative to other parcels.
+<!-- Our findings illustrate differences in patterns of 15-minute walk accessibility to care destinations most prominently profiled by income, low-income prevalence, children and single-parent status.
+-->
 
-Our findings illustrate differences in patterns of 15-minute walk accessibility to care destinations most prominently profiled by income, low-income prevalence, children and single-parent status.
-
- `r word_comment("IDEAS DISCUSSED: regarding @fig-Fig9 - 1) We know that highest accessibility & complete is downtown.
+<!--  `r word_comment("IDEAS DISCUSSED: regarding @fig-Fig9 - 1) We know that highest accessibility & complete is downtown.
 But people there have fewer children + lowest income are downtown.
 On the horizon as rent increases,  the most precarious may be pushed out and replaced by people who can pay higher rents.
-", " ")`
+", " ")` -->
 
-  `r word_comment("IDEAS DISCUSSED: regarding @fig-Fig9 - 2) The lowest accessibility & most incomplete are suburban -- they are likely higher income.
+<!--   `r word_comment("IDEAS DISCUSSED: regarding @fig-Fig9 - 2) The lowest accessibility & most incomplete are suburban -- they are likely higher income.
 Interesting to note: one-parent families are more likely to be time poor -- and leaving in low access/incomplete suburbs puts more at risk.
@@
 ", ":")`.
+-->
 
-  <!--Yes: low access and more incomplete = suburban (Figure 8) AND higher income (Figure 9).
+<!--   <!--Yes: low access and more incomplete = suburban (Figure 8) AND higher income (Figure 9).
 Are higher income folks living in low access and low caring places, but don’t care because they can afford to drive to get to these places?
-This has implications for sustainability and car dependence
+This has implications for sustainability and car dependence  -->
 
-  And also: poorer one parent families with children are also in these not caring/ complete places.
+<!--   And also: poorer one parent families with children are also in these not caring/ complete places.
 They are who may feel it: as they likely are more time poor and have lower access to cars.
-This could impact their wellbeing / access to care (which is essential) -->
+This could impact their wellbeing / access to care (which is essential) --> -->
 
- `r word_comment("IDEAS DISCUSSED: regarding @fig-Fig9 - 3) Pockets of the suburbs have high access to care. The suburbs do something -- they have care amenities. ", ":")`.
+<!--  `r word_comment("IDEAS DISCUSSED: regarding @fig-Fig9 - 3) Pockets of the suburbs have high access to care. The suburbs do something -- they have care amenities. ", ":")`.  -->
 
-<!--Potentially suburban have greater access to care compared to jobs.
+<!-- <!--Potentially suburban have greater access to care compared to jobs.
 I mean people live there for a reason I guess!
@@
 Or are these examples of suburban that have been designed to be walkable (which should be emulated)?
--->
+--> -->
 
-# Discussion
+<!-- # Discussion -->
 
-There seems to be a spatial mismatch where dwelling availability and affordability play an important role? Are there social housing programs for lower income households?
+<!-- There seems to be a spatial mismatch where dwelling availability and affordability play an important role? Are there social housing programs for lower income households? -->
 
-Even if we are on the right direction to develop human-centred urban areas by facilitating caring activities, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities, given the multiple impacts it has on everyday mobility, the unequal distribution of tasks among the population and its related socioeconomic consequences.
+<!-- Even if we are on the right direction to develop human-centred urban areas by facilitating caring activities, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities, given the multiple impacts it has on everyday mobility, the unequal distribution of tasks among the population and its related socioeconomic consequences.
 Therefore, it is necessary to provide alternatives to measure progress on this important issue.
@@
 On the other hand, an intersectional approach to understand how different personal identities interact in our daily day activities in the cities is necessary.
+-->
 
-The process of gentrification and displacement are trends also occurring in Hamilton.
+<!-- The process of gentrification and displacement are trends also occurring in Hamilton.
 On the horizon of rising rents, there is a risk of displacement among more vulnerable households, who may be replaced by individuals with higher incomes who can afford different transport modes, with more flexible schedules and are more likely to use ICT arrangements (including telework, e-shopping, and other online services).
 While, lower income households may be particularly vulnerable to locating in areas with worse accessibility to care alternatives, which could have negative consequences for these households and for the society as a whole.
-Even if we are on the right direction to develop human-centred urban areas by facilitating caring activities, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities, given the multiple impacts it has on everyday mobility, the unequal distribution of tasks among the population and its related socioeconomic consequences.
-Therefore, it is necessary to provide alternatives to measure progress on this important issue. 
-This work leaves many doors open for further research. For example, 
-•	the role that gender plays in these households. 
-•	Job arrangements per type of household.
-Some studies observe that working from home arrangements are more likely available for higher education attainment households (likely to have higher income), and families with children, but also living in neighbourhoods with good accessibility,
-•	Finally, an intersectional approach to understand how our personal identities interact in our access to essential activities and caring services.
-City planning should respond to these issues to promote equity.
+-->
+<!-- Even if we are on the right direction to develop human-centred urban areas by facilitating caring activities, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities, given the multiple impacts it has on everyday mobility, the unequal distribution of tasks among the population and its related socioeconomic consequences.
+-->
+<!-- Therefore, it is necessary to provide alternatives to measure progress on this important issue.  -->
+<!-- This work leaves many doors open for further research. For example,  -->
+<!-- •	the role that gender plays in these households.  -->
+<!-- •	Job arrangements per type of household.
+Some studies observe that working from home arrangements are more likely available for higher education attainment households (likely to have higher income), and families with children, but also living in neighbourhoods with good accessibility, -->
+<!-- •	Finally, an intersectional approach to understand how our personal identities interact in our access to essential activities and caring services. -->
+<!-- City planning should respond to these issues to promote equity.
 On the other hand, an intersectional approach to understand how different personal identities interact in our daily day activities in the cities is necessary.
+-->
 
-LUCIA'S:
-As in other case studies, essential opportunities and services tend to concentrate on Hamilton central.
+<!-- LUCIA'S: -->
+<!-- As in other case studies, essential opportunities and services tend to concentrate on Hamilton central.
 Therefore, it is more likely to carry out daily activities, including caring ones, in this area using sustainable modes such as walking or biking.
+-->
 
-However, this study highlights that there seems to be a spatial mismatch between dwelling availability, family profile and eventually (housing and mobility) affordability.
+<!-- However, this study highlights that there seems to be a spatial mismatch between dwelling availability, family profile and eventually (housing and mobility) affordability.
 At neighbourhood level the concentration of opportunities for all category types is in Hamilton Central, but households with children may not live there.
+-->
 
-Richer families may concentrate in either neighborhoods classified with poor accessibility: either “Somewhat caring and not complete” or “Not caring and not complete.
+<!-- Richer families may concentrate in either neighborhoods classified with poor accessibility: either “Somewhat caring and not complete” or “Not caring and not complete.
 However, these families are likely to have different mobility alternatives available.
@@
 e-shopping, service deliveries, taxis and MaaS services, working from home arrangements,,…).
+-->
 
-Our results show a different story for low-income households.
+<!-- Our results show a different story for low-income households.
 If they can afford it, these households will prefer to live in very caring and very complete neighbourhoods, which will ease their accessibility to daily activities and essential services and therefore, avoiding social exclusion.
@@
 Furthermore, school closures or any other type of relocation of essential services puts more pressure on their shoulders.
+-->
 
-Measuring caring neighbourhoods is of utmost importance to promote equity in city planning.
+<!-- Measuring caring neighbourhoods is of utmost importance to promote equity in city planning.
 Our results are of use for social housing programs for lower income households.
@@
 Otherwise, lower income household risk of social exclusion, especially when accessibility to essential services is limited.
+-->
 
-Similarly, this tool has great potential to support city planning, by analyzing the population groups benefited or harmed by relocation, closure or opening of different types of essential services.
+<!-- Similarly, this tool has great potential to support city planning, by analyzing the population groups benefited or harmed by relocation, closure or opening of different types of essential services.
+-->
 
-The process of gentrification and displacement are trends also occurring in Hamilton.
+<!-- The process of gentrification and displacement are trends also occurring in Hamilton.
 On the horizon of rising rents, there is a risk of displacement among more vulnerable households, who may be replaced by individuals with higher incomes who can afford different transport modes, with more flexible schedules and are more likely to use ICT arrangements (including telework, e-shopping, and other online services).
 While, lower income households may be particularly vulnerable to locating in areas with worse accessibility to care alternatives, which could have negative consequences for these households and for the society as a whole.
+-->
 
-Even if we are on the right direction to develop human-centred urban areas by facilitating caring activities, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities, given the multiple impacts it has on everyday mobility, the unequal distribution of tasks among the population and its related socioeconomic consequences.
+<!-- Even if we are on the right direction to develop human-centred urban areas by facilitating caring activities, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities, given the multiple impacts it has on everyday mobility, the unequal distribution of tasks among the population and its related socioeconomic consequences.
+-->
 
-ANATASIA's
-- harkens back to what role a planned neighbourhood has. Is it about serivice provision (the what), social homogenity (the who), local power and democracy? []
+<!-- ANATASIA's -->
+<!-- - harkens back to what role a planned neighbourhood has. Is it about serivice provision (the what), social homogenity (the who), local power and democracy? [] -->
 
-# Conclusion
+<!-- # Conclusion -->
 
-This study makes three types of contributions: empirical, methodological and theoretical.
+<!-- This study makes three types of contributions: empirical, methodological and theoretical.
 At the empirical level, areas of the mid-sized City of Hamilton have been typified by their degree of '15-Minute Caring Neighbourhood' potential.
@@
 We find...
+-->
 
-Theoretically, our work puts forth an explicitly caring 15-Minute Neighbourhod conceptualisation, bridging the Mobility of Care and the "15-Minute City" concept.
+<!-- Theoretically, our work puts forth an explicitly caring 15-Minute Neighbourhod conceptualisation, bridging the Mobility of Care and the "15-Minute City" concept.  -->
 
-<!--
+<!-- <!--  -->
 
-Instead of prescribing an urban form design principle, such as "all local amenities should be within a 15-minute walking distance", it instead examines an empirical example to determine which areas in the city have the _potential_ to be 15-minute neighbourhood based on the existing spatial accessibility offered by the urban environment and walking transport network.
+<!-- Instead of prescribing an urban form design principle, such as "all local amenities should be within a 15-minute walking distance", it instead examines an empirical example to determine which areas in the city have the _potential_ to be 15-minute neighbourhood based on the existing spatial accessibility offered by the urban environment and walking transport network.
 To this end, this data-driven methodology introduces a way to identify neighbourhoods that have potential, almost have potential, and are far from containing this potential to support future context-specific qualitative work.
+-->
 
--->
+<!-- --> -->
 
@@
 
-# Appendix
+<!-- # Appendix -->
 
-```{r plot-figure10, out.width=600, out.height=600}
-#| label: fig-Fig10
-#| fig-cap: "A bivariate map of the cumulative opportuninities map (aggregated per category at DA) and the diversity measure.
+<!-- ```{r plot-figure10, out.width=600, out.height=600} -->
+<!-- #| label: fig-Fig10 -->
+<!-- #| fig-cap: "A bivariate map of the cumulative opportuninities map (aggregated per category at DA) and the diversity measure.
 Amenity diverse = diviersity index above the region median.
 Caring/Complete = accessibility value above the region median.
-The Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
-knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_bivariate_entropy_copp_plot.png"))
-```
+The Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]." -->
+<!-- knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_bivariate_entropy_copp_plot.png")) -->
+<!-- ``` -->

### 2024-09-30 paper/Manuscript.qmd
draft of Abstract through Methods

@@
 
-The "15-Minute City" has been recently adopted by leaders as a way to promote human-scaled cities (e.g., Paris mayor Anne Hidalgo proposed a 15-minute city plan in her 2020 re-election campaign, 15-minute walkable neighbourhood guidance in Shanghai, China [@weng15minuteWalkableNeighborhoods2019]), As defined in @moreno_introducing_2021, the 15-Minute City is a urban planning model based in chrono-urbanism, a theory that emphasizes the positive impact on quality of urban life when urban space becomes multi-rythmic, in opposition to industrial Fordism, single-use zoning, and the segregation of urban time and space for individual uses and mobility [@mulicekUrbanRhythmsChronotopic2015; @morenoVilleQuartHeure2016].
-The 15-Minute City is one that accommodates the ability to reach essential destinations within a walkable, bikeable or transit supportive radius.
-It would allow individuals to reclaim time spent on car mobility and give way to sustainable modes; prompting urban spaces that are responsive to human needs and environmental sensibilities [@Allam2022].
+The "15-Minute City" has been recently adopted by leaders as a way to promote human-scaled cities [@teixeiraClassifying15minuteCities2024; @teixeiraClassifying15minuteCities2024].
+As introduced in @moreno_introducing_2021, the 15-Minute City is a urban planning model based in chrono-urbanism, a theory that emphasizes the positive impact on quality of urban life when urban space becomes multi-rythmic, in opposition to the segregation of urban time and space for individual uses and mobility [@mulicekUrbanRhythmsChronotopic2015; @morenoVilleQuartHeure2016].
+It is in opposition to the forces that supported the formation of single-use neighbourhoods such as industrial Fordism and single-use zoning [@mulicekUrbanRhythmsChronotopic2015; @morenoVilleQuartHeure2016].
+The "15-Minute City" is a utopic ideal, it is a city where one that accommodates the ability for all to reach important destinations within a walkable, bikeable or transit supportive radius.
+This form of neighbourhood planning would allow individuals to reclaim time spent on car mobility, giving way to sustainable modes and prompting urban spaces that are responsive to human needs and environmental sensibilities [@Allam2022].
 
@@
 Social scientists also emphasized that top-down neighbourhood planning approaches overestimated the built environment's influence on social life [@talenSocialSciencePlanned2017].
-Neighbourhood planning approaches such as the 15-Minute City concept are being re-embraced, but from bottom-up perspective.
-However how to equitably prescribe urban form that 'enables' contact with social opportunities (instead of 'engendering it' from the top-down) is still under debate.
-For instance, questions such as _what_ are destinations that matter, by what mode and travel time threshold, and _for whom_ remain under discussion among proximity-based planners [@silvaProximitycentredAccessibilityConceptual2023; @silvaRegionalAccessibilityUndermining2022].
+Neighbourhood planning approaches such as the 15-Minute City concept are being re-embraced by some, but from bottom-up perspective.
+How to equitably prescribe urban form that 'enables' contact with social opportunities (instead of 'engendering it' from the top-down) is still under debate.
+For instance, questions such as _what_ are destinations that matter, by what mode and travel time threshold, and _for whom_ remain under discussion among proximity-based planners who frequently use accessibility-based tools [@silvaProximitycentredAccessibilityConceptual2023; @silvaRegionalAccessibilityUndermining2022; @guzmanProximityEnoughCritical2024].
 
-To add clarity to these questions surrounding the 15-Minute City concept, in this work we aim to build a theoretical bridge to "Mobility of Care" as introduced in @sanchezdemadariagaMobilityCareIntroducing2013.
-Mobility of Care is especially relevant to the 15-Minute City, as it is a purpose of travel that is necessary for all and often occurs at the local level.
-Namely, Mobility of Care refers to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands.
+To clarify questions surrounding the 15-Minute City concept of what destinations matter and for whom, this work builds a theoretical bridge to "Mobility of Care".
+The Mobility of Care term was coined in @sanchezdemadariagaMobilityCareIntroducing2013; it refers to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands.
 While decades of research have examined types of household-serving trips (e.g., shopping trips, escorting to school), @sanchezdemadariagaMobilityCareIntroducing2013 was the first to consider them all as one category.
 In doing so, Mobility of Care is shown to be a significant proportion of daily travel, characterizing the trip purpose of approximately 30% of adults’ trips [@sanchezdemadariagaMobilityCareIntroducing2013; @sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory].
-Notably, care trips take on tremendous gendered and socio-economic importance, as women, and especially those from lower income households, complete the majority of care trips [@ravensbergen2023exploratory].
-Despite its importance for all and in advancing gender equality, the examination of mobility of care and access to care destinations have been under-examined relative to employment destinations.
-This is especially pertinent in accessibility research that has largely focused on employment () and retail ().
+We argue Mobility of Care is especially relevant to the 15-Minute City, as care trips are necessary for _all_ people and they most often occur at the local level.
+Moreover, care trips take on tremendous gendered and socio-economic importance, as women, and especially those from lower income households, complete the majority of care trips [@ravensbergen2023exploratory].
+Despite care travel's importance in advancing gender equality and equity broadly, the examination of mobility of care and access to care destinations have been under-examined relative to employment destinations.
+This is especially pertinent in accessibility research that has largely focused on employment points of interest e.g., [@farberOntarioLineSocioeconomic2019; @duarteInfluenceJobAccessibility2023; @ryanAccessibilitySpaceTime2023].
 
-We aim to theoretically bridge Mobility of Care with the 15-Minute City concept to clarify questions surrounding _what destinations within a 15 minute walkable radius matter_.
-We hypothesis that _care_ destinations matter, they matter for all people but especially for women and lower-income women preforming care trips.
-In this vein, this work pairs this hypothesis with an exploratory clustering methodology to identify which neighbourhoods have the potential to support 15-minute access to care destinations in an empirical case study of Hamilton, Ontario, typical mid-sized Canadian city.
-More concretely, the research objectives are two-fold: after a review of the 15-Minute City concept and associated measurement tools, this work first introduces the Mobility of Care conceptualisation to re-imagine local amenities importance within the 15-Minute City concept, and second to apply a data-driven classification and clustering machine learning methodology to support the bottom-up examination of _15-minute caring neighbourhoods_  through an empirical examination of Hamilton.
+The objectives of this research are two-fold.
+First, we aim to theoretically bridge Mobility of Care with the 15-Minute City concept to to re-imagine what local amenities matter and define an associated accessibility and diversity of accessibility measure.
+Second, taking a bottom-up data-driven approach, we apply a classification and clustering machine learning methodologies to support the bottom-up identification of _15-minute caring neighbourhoods_  through an empirical examination of the mid-cited city of Hamilton, Canada.
+These neighbourhoods are then interpreted on the continuum of having (and not having) the region-relative land-use and transportation infrastructure to support _completely caring_ access.
+Who currently resides in which neighbourhood is examined, and ways to further expand the care accessibility story are discussed.
 
@@
 
-In the last decades, the need for more sustainable, healthier, and liveable cities have received more attention.
-Planners and decision-makers have proposed alternatives to mobility-based neighbourhood planning to one centered on proximity to urban functions [@pozoukidou15MinuteCityDecomposing2021].
+In the last decades, the need for more sustainable, healthier, and liveable cities have received signficant attention.
+Planners and decision-makers have proposed alternatives to mobility-based neighbourhood planning principles to ones that center proximity to urban functions [@pozoukidou15MinuteCityDecomposing2021].
 In this context, the 15-Minute City is now in the public spotlight [@logan_xminute_2022; @moreno_introducing_2021].
 In @moreno_introducing_2021 where the 15-Minute City concept was introduced, four universal dimensions are outlined: density (in terms of people per km$^2$), diversity (including mixed-land use and diversity of people), the temporal and spatial proximity to essential services, <!-- (residential areas, work, commercial areas, education centres, health facilities and other basic institutions)--> and digitalisation.
-The framework presented has been critised within academic and planning circles [@guzmanProximityEnoughCritical2024].
-Namely, critiques have centered on the concept's inadequecy in addressing pre-existing structural forces and individual charctertes that drive inequalities in who does and who could benefit more from such a paradigm [@dimarino15minuteCityConcept2023; @willberg15minuteCityAll2023].
-A 15-Minute City for all people is a utopic ideal, however how will existing inequalities in how people move through and benefit from the city be addressed.
-_Who is in the most need of the 15-Minute City before we all get there?_.
-Without directed answers to these context-sensitive questions makes this increasingly popular concept susceptible to becoming a place-branding slogan [@pozoukidou15MinuteCityDecomposing2021].
+However, the framework presented has been criticized within academic and planning circles [@guzmanProximityEnoughCritical2024].
+Critiques highlight the concept falls short in addressing the  pre-existing structural forces and individual characteristics that drive inequalities, influencing who benefits or could benefit the most from such a paradigm [@dimarino15minuteCityConcept2023; @willberg15minuteCityAll2023].
+A 15-Minute City for all people is an aspirational ideal, but how will existing mobility and accessibility inequalities be addressed.
+Without directed and context-specific solutions, this increasingly popular concept risks becoming a place-branding slogan [@pozoukidou15MinuteCityDecomposing2021].
 
-Reflecting this flexible framework; cities that have adopted 15-Minute City concepts have done so using a diverse range of definitions and tools along with similar indistinct universal approaches.
+`r word_comment("Lucia, do I correctly reflect the plans you reviewed?", " ")`
+Reflecting the flexible and aspirational presentation of the 15-Minute City; cities that have adopted the concept have done so using a diverse range of definitions and tools along with indistinct universal approaches.
 An early adopter was Portland, US in the Portland Plan [@city_of_portland_20minute_2010] of April 2012.
-It aims to foster an inclusive city development based on prosperity, education, health, and equity for a 30 year horizon.
-It promotes neighbourhood self-suffiency and connectivity to city centers and centres of employment.
-The progress report describes a high-level focus on equitable service delivery to all residents with equity concerning topics related to racial equity and people with disabilities [@portland_government_portland_2017]; similarly adopting an "all populations" approach.
-Later, other cities also adopted proximity-based goals into their plans, taking similar neutral entity approaches.
-For instance, in the plan that later catapulted the 15-Minute City into public discourse: Paris mayor Anne Hidalgo proposed a 15-minute city plan in her 2020 re-election campaign that foregrounded six social functions of the city that should be easily accessible from any location, namely: housing, work, health care, groceries, education and leisure [@ville_de_paris_paris_2022].
+It aims to foster an inclusive city development based on prosperity, education, health, and equity over a 30 year horizon.
+It promotes neighbourhood self-sufficiency and connectivity to city centers and centres of employment.
+The progress report describes a high-level focus on equitable service delivery to all residents with equity concerning topics related to racial equity and people with disabilities [@portland_government_portland_2017], similarly taking on an "all populations" approach.
+Later, other cities also incorporated proximity-based goals into their plans, adopting similar neutral entity approaches.
+For instance, in the plan that later catapulted the 15-Minute City into public discourse: Paris mayor Anne Hidalgo proposed a 15-minute city plan in her 2020 re-election campaign that foregrounded six social functions that should be easily accessible from any location, namely: housing, work, health care, groceries, education and leisure [@ville_de_paris_paris_2022].
 These ideas inspired language in the agendas of other cities in the Western world such as Ottawa, in the Canadian context, who also adopted a 15-Minutes approach in their recent Official Plan [@ville_dottawa_quartier_2021].
-From the review of worldwide practice literature in @teixeiraClassifying15minuteCities2024, overall the 15-Minute City concept is in early phases of implementation around the world and the diverse range of definitions, strategies, and instruments present a significant knowledge and implementation gap: the 15-Minute City seems nice but _how do we get there and whom will benefit_?
+From the review of worldwide practice literature in @teixeiraClassifying15minuteCities2024, the 15-Minute City concept is in early phases of implementation around the world and the diverse range of definitions, strategies, and instruments present a significant knowledge and implementation gap: the 15-Minute City is aspirational but _how do we get there and whom will benefit_?
 
@@
 The NUC is a socio-spatial normative scheme widely adopted by government agencies (and the real-estate community) in the Western world after the Second World War [@talenSocialSciencePlanned2017; @solow1969concept].
-Pairing well with the objectives of planning agencies at the time, Perry's NUC would allow for efficient mass-building of cellular units that prioritized the perceived functional needs of women and children: each unit providing proximate access to an elementary school and sufficent supporting community facilities, shopping, parks and housing [@talenSocialSciencePlanned2017; @brodyNeighbourhoodUnitConcept2013].
+Pairing well with the objectives of planning agencies at the time, Perry's NUC would allow for efficient mass-building of cellular units that prioritized the perceived functional needs of women and children: each unit providing proximate access to an elementary school and supporting community facilities, shopping, parks and housing [@talenSocialSciencePlanned2017; @brodyNeighbourhoodUnitConcept2013].
 The NUC primarily prioritized local service provision, though Perry had confidence in good design's contribution to 'neighbourhood spirit' [@hall2014cities].
-From the NUC's conception until the end of the 1960s, planner's attempts to prescribe social meaning to the neighbourhood's physical form had been criticized to near extinction [@talenSocialSciencePlanned2017].
+From the NUC's conception until the end of the 1960s, planners' aspirational attempts to prescribe social meaning to the neighbourhood's physical form had been criticized to near extinction [@talenSocialSciencePlanned2017].
 By social scientists, the overestimation on the impact of built environment on social life was critiqued.
-By other planners, the inability to agree on the specificity of the neighbourhood (i.e., population sized, and the type, quality, and quantity of amenity) as well as how neighbourhood units are connected to other neighbourhoods and the rest of the region.
-In response to these criticisms, neighbourhood planning proponents have redefined deterministic terminology: prescribed physical forms as 'enabling' rather than engendering, and the provision of social contact 'opportunities' instead of determined social behaviour.
-The re-defined bottom-up approach to community and neighbourhood planning has been adopted by New Urbanist from which the 15-Minute City stems [@kissfazekasCircleParadigms15minute2022].
-But with _what bottom-up processes?_; small scale experiments are being tested () with preliminary small-scacle results.
-These processes are only in early stages so critiques of social scientists and planners from decades past still persist.
+By other planners, the inability to agree on the specificity of the neighbourhood (i.e., population size and the type, quality, and quantity of amenity) as well as how neighbourhood units are connected to other neighbourhoods and the rest of the region.
+In response to these criticisms, neighbourhood planning proponents redefined deterministic terminology; their prescriptive physical form will 'enable' social contact with opportunities rather than 'engendering' it.
 
-_How can physical form be planned to enable an improved quality of life, and for whom?_  has also occupied urban planning research in recent years, exploring what are the impacts of built environment on residence's outcomes.
+The re-defined bottom-up approach to community and neighbourhood planning has been adopted by New Urbanists, from which the 15-Minute City stems [@kissfazekasCircleParadigms15minute2022]; however, the effectiveness of these bottom-up processes remain to be seen due to the contemporary nature of the concept.
+Though, related research can be examined.
+In recent years, the question of _how can physical form be planned to enable an improved quality of life, and for whom with what outcomes_ has occupied the urban planning research agenda.
 For instance, the examination of low-income households access to transportation and their likelihood of gaining employment [@blumenbergDriveWorkRelationship2017; @bastiaanssenDoesBetterJob2022] or the relationship between children's access to public transit and participation in after-school activities [@palmRolePublicTransit2020].
-Relatedly, a new wave of researchers and practitioners focused on local and context-specific relationships in the proximity to destinations have also emerged [@silvaProximitycentredAccessibilityConceptual2023; @silvaRegionalAccessibilityUndermining2022].
-As reviewed  in the city plans that have adopted 15-Minute City approaches, the NUC's critisms, and recent proximity-based examinations, the question of _how to enable improved quality of life through urban built environment_ has shown to be highly context-senstive with the need for further investigation.
+A new wave of researchers and practitioners focused on local and context-specific relationships with the proximity to destinations have also emerged [@silvaProximitycentredAccessibilityConceptual2023; @silvaRegionalAccessibilityUndermining2022].
+As reviewed  in the city plans that have adopted 15-Minute City approaches, the NUC's criticisms, and recent urban planning research, the question of _how to enable improved quality of life through urban built environment_ has shown to be highly context-sensitive, with the need for further investigation.
 
@@
 In examining how to enable improved quality of life through urban built environment, an increasingly popular tool have been accessibility measures.
-They are a way to quantify the ease of reaching destinations from a given point in space and have been used to examine urban areas through just and sustainable city planning agendas [@vale2023accessibility; @handyAccessibilityIdeaWhose2020].
+These measures are a way to quantify the ease of reaching destinations from a given point in space and have been used to examine urban areas through just and sustainable city planning agendas [@vale2023accessibility; @handyAccessibilityIdeaWhose2020].
 As the 15-Minute City is an amenity-provision neighbourhood planning concept which aims to enable urban environment that improves life quality, its analysis lends well to the use of accessibility tools [@guzmanProximityEnoughCritical2024].
@@
 This form of accessibility quantifies how many destinations can be reached from a point in space within a given travel time threshold, pairing well with examinations that have a normative travel time threshold, like the 15-Minute City.
-Furthermore, the cumulative opportunity measure is widely appreciated for its intuitive computation and is popular among practitioners [@handyAccessibilityIdeaWhose2020; @handyMeasuringAccessibilityExploration1997; @chengInvestigatingWalkingAccessibility2019].
-However, reflecting the diversity of the accessibility literature, accessibility measures other than the cumulative opportunity have also been applied [@guzmanProximityEnoughCritical2024].
+Furthermore, the cumulative opportunity measure is widely appreciated for its intuitive computation and popularity among practitioners [@handyAccessibilityIdeaWhose2020; @handyMeasuringAccessibilityExploration1997; @chengInvestigatingWalkingAccessibility2019].
+However, accessibility measures other than the cumulative opportunity have also been applied, reflecting the diversity of measures in the literature [@guzmanProximityEnoughCritical2024].
 
 Another method which compliments the assessment of the 15-Minute City as it relates to the diversity of opportunity types, is entropy.
-The entropy measure has been widely used to characterise the diversity of land-use mix in early work by @frankLinkingObjectivelyMeasured2005 based on the species diversity Shannon-Wiener index that expresses relative evenness throughout a sequence (this measure among other species diversity indices of the time are reviewed in @whittaker1972evolution).
-It has also been used in social sciences to comprehend the occurrence of social events in a structured manner specially with regard to life trajectories [@ritschardSequenceAnalysisRelated2018].In the realm of urban studies and planning literature the malleable concept of entropy has been used to characterise the diversity in land-use mix [@ewingTravelBuiltEnvironment2010a], especially in the context of active modes [@luUrbanDensityDiversity2017; @mavoaIdentifyingAppropriateLanduse2018] and suburban sprawl [@randallGISBasedLand2015].
+The entropy measure has been widely used to characterise the diversity of land-use mix in early work by @frankLinkingObjectivelyMeasured2005 based on the species diversity Shannon-Wiener index that expresses relative evenness throughout a sequence [see review in @whittaker1972evolution].
+It has also been used in social sciences to comprehend the occurrence of social events in a structured manner specially with regard to life trajectories [@ritschardSequenceAnalysisRelated2018].
+In the realm of urban studies and planning literature, the malleable concept of entropy has been used to characterise the diversity in land-use mix [@ewingTravelBuiltEnvironment2010a], especially in the context of active modes [@luUrbanDensityDiversity2017; @mavoaIdentifyingAppropriateLanduse2018] and suburban sprawl [@randallGISBasedLand2015].
 As well, it has been used to understand mobility behaviour [@mcbride2020exploration; @montero2023applying; @montero2023role].
@@
 To the authors knowledge, the use of an entropy measure to reflect the diversity in care destination amenity types and in context within a 15-minute neighbourhood, has yet to be done.
-However, entropy measures thematically compliments the 15-Minute City.
+However, entropy measures theoretically compliment the 15-Minute City concept.
 
-`r word_comment("I need to fill this out. A small classification methods paragraph", ".")`
-In addition, a useful tool has been classification algorithms in the identification of spatial typologies and in profiling who may reside there.
-Those too have K means...
-Hierarchical...
-and SOM [@delmelleDifferentiatingPathwaysNeighborhood2017]...
-regionlization techniques such as spatially constrained multivariate clustering [@burkeGeospatialAnalysisFramework2022] Use of tree diagrams!
-....SOM is a type of artificial neural network (ANN) trained using unsupervised learning that produces a two-dimensional representation of the input space.
-It applies competitive learning by way of a neighbourhood function thereby preserving the topological properties of the input space and producing combinations of the input space that we can then intuitively profile.
-It has been used widely in XX but scarcely in transportation related research [@victorianoTimeSpaceMoney2020]
+`r word_comment("I am only vaguely familar with this literature as of these months - how does this sound?", " ")`
+In addition, classification algorithms have been useful tools in the identification of mobility and spatial typologies in relation to transportation planning.
+Machine learning approaches have often been deployed.
+For instance, @burkeGeospatialAnalysisFramework2022 uses spatially constrained multivariate clustering, a regionalisation technique, to develop urban form typologies related to the 15-Minute City.
+The k-means algorithm is another popular method; @ganUnderstandingUrbanMobility2020 applies it to analyse travel behaviour and classify metro stations based on mobility patterns.
+Combing classification techniques is also common; for instance, @delmelleDifferentiatingPathwaysNeighborhood2017 uses self-organizing maps (SOM) algorithm to group US neighbourhoods by minimising dissimilarity in attributes before applying k-means for further classification.
+Similarly, @victorianoTimeSpaceMoney2020 uses SOM to classify individuals' travel patterns according to the dissimilarity in mobility attributes and uses the decision tree algorithm to partition the data into interpretable classifications.
 
-All-in-all, within the context of examining the 15-Minute City there has been a wide range of tools applied in the literature that appear appriorpiate.
-However, the question of _15-minute cities for whom_ requires further examination.
-In this way, this work connects the 15-Minute City to another normative conceptualisation: Mobility of Care.
+All-in-all, within the context of examining the 15-Minute City, a variety of tools have been applied in the literature.
+However, the question of _15-minute cities for whom_ sill demands further examination.
+This work aims to bridge the concept of the 15-Minute City with another normative framework: Mobility of Care.
 
@@
 
-As the 15-Minute City is in early stages of conceptualisation; an alternative way to conceive of _what_ urban functions matter could be through a feminist lens.
-Instead of focusing on _all_ destinations, it may be useful to examine destinations involved in caring activities.
-Caring activities meet the physical, psychological, and emotional needs of others, and are one of the most essential and fundamental activities in society [@ilo_care_2018].
+Rather than focusing on _all_ destinations, it may valuable to examine those related to caring activities.
+This re-framing offers a feminist perspective on  _what_ urban functions matter, connecting well with the 15-Minute City concept.
+Caring activities, which meet the physical, psychological, and emotional needs of others, are among the most essential and fundamental activities in society [@ilo_care_2018].
 Yet, they are the most unequal, undervalued, and even devalued activities worldwide.
-Traditionally, caring activities have been borne on women’s shoulders [@haydenGrandDomesticRevolution1982; @hochschild_second_2012].
+Conventionally, caring activities have been borne on women’s shoulders [@haydenGrandDomesticRevolution1982; @hochschild_second_2012].
 According to @ilo_care_2018, women and girls perform more than three-quarters of the total amount of unpaid care, a gender gap that varies geographically [@RN5].
@@
 From this research motivation, Sánchez de Madariaga coined the term Mobility of Care in 2013 to refer to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands [@sanchezdemadariagaMobilityCareIntroducing2013].
-While decades of research had examined these unique household-serving trips, her seminal work was the first to consider all these trips as one category and demonstrate how mobility of care is a significant proportion of daily travel.
+While an undercurrent of research had examined these unique household-serving trips over the decades, her seminal work was the first to consider all these trips as one category and demonstrate how mobility of care is a significant proportion of daily travel.
 
-The adoption of the Mobility of Care concept within the 15-Minute City also explicitly incorporates equity considerations.
-Perhaps unsurprisingly given the gendered division of care work discussed above, women have consistently been found to complete more mobility of care trips than men.
+The Mobility of Care concept also explicitly integrates inter-sectional equity considerations.
+In considering Mobility of Care within the 15-Minute City one could address a common criticism leveled at the 15-Minute City [@guzmanProximityEnoughCritical2024].
+Perhaps unsurprisingly given the gendered division of care work discussed, women have consistently been found to complete more mobility of care trips than men.
 In one study, mobility of care comprised 32% of women’s daily trips compared to 28% of men’s.
-While this gap was significant, it was found to be far greater in lower income households where women complete 20% more care trips than men [@ravensbergen2023exploratory].
+While this gap is significant, it was found to be far greater in lower income households where women complete 20% more care trips than men [@ravensbergen2023exploratory].
 Sánchez de Madariaga not only shows how important these mobility of care trips are, but also highlights the ways in which “mobility of care is systematically under-represented in any analysis of urban transport” (p.
 37).
-Transport surveys and tools do not directly capture mobility of care, which re-enforces the idea that these trips are not a significant part of daily mobility.
+As a product of the masculinst bias in planning, transport surveys and tools do not directly capture mobility of care, which re-enforces the idea that these trips are not a significant part of daily mobility.
 In this respect, the feminist perspective of the cities of proximity is still underestimated with only few examples on the topic [@gil_sola_choose_2022; @macintyre_her_2022].
-Given the impact of caring activities on everyday life, the chrono-urbanism should specifically target the analysis of caring activities.
 
-Mobility of Care also explicitly prioritizes shorter-distance and local trips, those that could be done by sustainable modes if sufficently supported, making the perspective highly compatible with the 15-Minute City paradigm.Compared to the trip to work, mobility of care trips are more frequently completed by foot, and less frequently done by transit or bicycle [@ravensbergen2023exploratory].
-Further, evidence suggests that walking is more common for care than for work trips because care trips tend to be short [@ravensbergen2023exploratory]; if care destinations are closer to the home, it is more likely that these destinations are walkable.
-Ample literature examines the benefits and barriers to walking care destinations, such as schools (e.g., [@RN7;@RN8;@RN9]) and grocery stores (e.g., [@RN10;@RN11]).
-However, little of this work uses the mobility of care framework.
-Instead, they tend to consider singular care destinations in research focusing on walkability.Furthermore, mobility of care trip characteristics differ from other trips, notably those of the commute to work.
-For instance, many of these trips are made through trip-chains, i.e., when a non-work-related stop is made during the commute.
-Women have been found to do so more than men, which is understood as being due to their need to balance unpaid-household serving care and paid employment (e.g., picking up groceries or children on the way home from work) [@ravensbergen2023exploratory; @scheinerWomenComplexDaily2015].
+Pairing well with the focus on shorter trips and the potential use of sustainable modes within the 15-Minute City, Mobility of Care trips are more local, shorter-distance and trip-chained.
+Compared to the trip to work, mobility of care trips are more frequently completed by foot, and less frequently by transit or bicycle [@ravensbergen2023exploratory].
+However, little work to date examines walking to care destinations through the Mobility of Care framework.
+Though there is ample literature that examines walking to care destinations, such as schools (e.g., [@RN7;@RN8;@RN9]) and grocery stores (e.g., [@RN10;@RN11]), they tend to consider singular care destinations in research focusing on walkability.
+
+The objectives of this research are two-fold.
+First, we aim to theoretically bridge Mobility of Care with the 15-Minute City concept to to re-imagine what local amenities matter and define an associated accessibility and diversity of accessibility measure.
+Second, taking a bottom-up data-driven approach, we apply a classification and clustering machine learning methodologies to support the bottom-up identification of _15-minute caring neighbourhoods_  through an empirical examination of the mid-cited city of Hamilton, Canada.
+These neighbourhoods are then interpreted on the continuum of having (and not having) the region-relative land-use and transportation infrastructure to support _completely caring_ access.
+Who currently resides in which neighbourhood is examined, and ways to further expand the care accessibility story are discussed.
 
 `r word_comment("I need to clean this up and firmly reiterate the goals", " ")`
-Taken together, city planning is moving to a people-centered approach focusing on accessibility to basic needs.
-However, the 15-Minute City has faced criticism as it fails to provide sufficient guidance on how to reach the utopic goal of 15-Minute for all.
-We argue, that significance in what destinations matter within the 15-Minute City could be gained by adopting a Mobility of Care prespective.
-We develop this approach using a mid-size typical Canadian city of Hamilton and a novel care destination dataset as case study.
-The objectives of the paper are as follows:
+Due to it's popularity, the 15-Minute Concept is moving city planning towards a more people-centered approach focusing on accessibility to basic needs.
+While an aspirational ideal, the concept has faced criticism as a inadequately providing a way forward to address structural inequalities.
+In this context, we bridge the 15-Minute City with the Mobility of Care concept, to add significance to what destinations and associated trips matter for all, but especially along gendered and socio-economic lines.
+We apply the "Caring 15-Minute City" concept to measure what neighourhoods in the mid-size Canadian city of Hamilton have the potential to be "caring 15 minute neighbourhoods".
+Making use of the tools reviewed and applied in the transportation research, the objectives of the paper are completed as follows:
 
-1)	To examine the potential accessibility to, and diversity of, care destinations in the case study.
-We identify to what degree are neighbourhoods caring and/or complete.
-To do so, we make use of machine learning techniques namely self organizing maps (SOM) as a data-dimensionality reduction technique and decision trips for the purpose of classification.
-2)	We then apply the same SOM methodology to determine profiles of population groups to understand who enjoys these neighbourhoods and who does not.
+1)	The cumulative accessibility measure and the entropy measure are used to quantify both the level of access and the diversity of access to care destinations within a 15-minute walk.
+2) Based on these accessibility and diversity outputs, "Completely Caring 15-Minute Neighbourhood" typologies are generated using the SOM algorithm.  
+3)	A decision tree is then applied to these typologies and socio-economic variables of populations who reside in these neighbourhoods.
+This is done to delineate data-driven profiles on which groups have the potential to benefit from what Completely Caring 15-Minute Neighbourhood.
 
@@
 
-This study focuses on Hamilton, Ontario, a mid-sized city on the shore of Lake Ontario.
+This work focuses on Hamilton, Ontario, a mid-sized city on the shore of Lake Ontario.
 Hamilton has a heterogenous land-use, with a populated and dense urban core, surrounded by suburban development, which is itself surrounded by rural communities.
 The Niagara Escarpment runs through Hamilton, and results in a city with two key elevations: a more dense lower city that contains the downtown core and the elevated suburban development referred to as ‘the Mountain’.
-In this work, we analyse the residential parcel centroids, approximately ~140,000 locations.
+In this work, we analyse the residential parcel centroids, 143,882 locations.
 We aggregated the points at the level of Canadian Census Dissemination Area (DA) along with the population and population per parcel plots in @fig-Fig1.
@@
 The densely populated inner city is characterised by lower average incomes, and a higher prevalence of households living under the low-income cutoff thresholds (LICO).
-The suburban parts of the city tend to have a greater proportion of children and a lower proportion of one parent households.
+Noteably, the suburban areas of the city tend to have a greater proportion of children and a lower proportion of one parent households.
 
@@
 A spatial dataset of care destinations for Hamilton was carefully compiled.
-This includes 14 types of destinations that were placed by authors in ( _work removed for double-blind review_ ) into four categories: dependent-centric (e.g., the destinations for child- and elder-centric escorting trips), grocery-centric, health-centric, and errand-centric.
+The dataset includes 14 types of destinations that were placed by authors in ( _work removed for double-blind review_ ) into four categories: dependent-centric (e.g., the destinations for child- and elder-centric escorting trips), grocery-centric, health-centric, and errand-centric.
 Notably, these categories were generated following the travel purpose categories created in the mobility of care research by @sanchezdemadariagaMeasuringMobilitiesCare2019.
@@
 Then, the entropy measures used to calculate the diversity of accessibility to each of the care categories is described.
-Subsequently, the accessibility and diversity of accessibility values associated with each residential parcel location are then fed into a self-organizing map (SOM) data-reduction algorithm.
-How the resulting output is classified into 9 classifications using decision-trees that represent the amount of "caring access" and "completeness" of this access are then described.
-This methodology presents a data-driven approach to examine what neighbourhoods in a city have the potential to provide 15-minute caring access, at what level of intensity and completeness, as well as in who may be residing in areas that benefit the most.
+Subsequently, the accessibility and diversity of accessibility values associated with each residential parcel location are fed into a self-organizing map (SOM) algorithm.
+The resulting output is then classified into 7 superclusters the amount of "caring access" and "completeness" of this access.
+A Decision Tree, using the superclusters as features and socio-economic variables for the populations who reside in the city as lavels, is generated to identify patterns on who may benefit (or be excluded from) the potential of Completely Caring 15-Minute Neighbourhoods.
+
+In sum, this methodology presents a data-driven approach to examine what neighbourhoods in a city have the potential to provide 15-minute caring access, at what level of intensity and completeness, as well as in who may be residing in areas that benefit the most.
 
@@
 First, SOM is an unsupervised technique implemented to reduce the data dimensional and create interpretable clusters related to the intensity and completeness of caring access.
-Then, a decision tree is deployed to further examine the cluster membership of parcels and another decision tree is run to characterise the socio-economic profile of who resides in neighbourhoods associated with certain (in)complete and (un)caring accessibility clusters.
-Together, this combined approach leverages the unsupervised data-driven classification power of the SOM with the interpretablity of decision trees.
-The procedure deployed in this work closes resembles the work of @victorianoTimeSpaceMoney2020, however instead of each observation representing an individual's daily mobility behaviour (with associated variables) each observation is a parcel location with calculated care accessibility and diversity of type of care accessibility scores.
+This is done by imputing each parcel as an observation with its associated accessibility and diversity attributes, and the data being rearranged onto a two-dimensional space based on its minimizing dissimilarity in its neighbourhood.
+Then, a decision tree is run to characterise the socio-economic profile of who resides in neighbourhoods associated with certain (in)complete and (un)caring accessibility clusters.
+Together, this combined approach leverages the unsupervised data-driven classification power of the SOM with the interpretation of decision trees.
+The procedure used in this work is similar to the approache used in @victorianoTimeSpaceMoney2020.
+However, instead of each observation representing an individual's daily mobility behaviour (with associated variables), here each observation represents a parcel location with calculated care accessibility and diversity of type of care accessibility scores.
 
-For the SOM step, the SOM algorithm implementation in the function 'trainSOM()' from {SOMbrero} R package is used [@VillaVialaneix2017].
+For the SOM step, the SOM implementation in the function 'trainSOM()' from {SOMbrero} R package is used [@VillaVialaneix2017].
 The input variables include the 143,882 parcels, each as individual observations along with 15 variables: the 14 calculated accessibility scores $S_i^$, normalized to the min-max range score within each $t$, and one diversity value $D_i$.
 Otherwise, defaults for all other parameters are assumed, relying on the data-driven heuristics embedded in the 'trainSOM()' function.
-Consequently, a 100 node (10 by 10 gride) SOM structure using euclidean distance and square topology is produced.
-Simply put the SOM algorithm proceeds as follows.
-A 2D grid of nodes is created (in our case 100 nodes) as specified by the analyst, where each node will represent a point in the reduced-dimensional space.
+Consequently, a 100 node (10 by 10 grid) SOM structure using euclidean distance and square topology is produced.
+Simply put, the SOM algorithm proceeds as follows: a 2D grid of nodes is created (in our case 100 nodes) as specified by the analyst, where each node will represent a point in the reduced-dimensional space.
 Upon initialization, each node is assigned a random weight vector of the same dimension as the input data (in our case, 15).
-From the input data, a random observation with its associated weight vector (i.e., one parcel point with 15 variables) is selected and the Euclidean distance between its weight vector and all nodes in the grid is calculated using Euclidean distance.
-The node with the smallest distance (i.e., the smallest dissimilarity) is crowned the 'best matching unit' as it is the node that best represents the input observation.
+From the input data, a random observation with its associated weight vector (i.e., one parcel point with 15 variables) is selected and the Euclidean distance between its weight vector and all nodes in the grid is calculated.
+The node with the smallest distance (i.e., the smallest dissimilarity) is labelled the 'best matching unit' as it is the node that best represents the input observation.
 After this best matching unit is identified, its own weight and its neighbouring nodes are updated to become more like the input observation.
@@
 For the decision tree step, the supercluster-classified parcels identified in the SOM step are used as _labels_ and socio-economic and demographic indicators available from the most recent Canadian Census related to the mobility of care literature are used as _features_.
-This step is used to profile the superclusters in a data-drive way to explore who may reside in what supercluster.
-To estimate the decision tree, the 'rpart()' function in the {rpart} package assuming default parameters for classification splitting along with each value being weighted by the population present in the associated DA [@Therneau2023].
-As a summary of the decision tree algorithm, it is a supervised learning technique that begins by splitting a subset of the input data into branches based on a selected feature with the lowest impurity score (i.e., the lowest amount of mixing of label membership in a branch).
-This process is recursively repeated for each subset of data for the next best features.
-Ultimately, the data is classified into distinct classes with class membership explained by traversing the branches defined by features that characterise the partitions in the decision tree.
-Noteably, the absence of features from the decision tree does not necessarily mean they are not relevant for classification, they are just less relevant than other features.
+This step is used to profile the superclusters so we may explore who reside in what supercluster in a data-driven way.
+To estimate the decision tree, the 'rpart()' function in the {rpart} R package is used [@Therneau2023]; default parameters for classification splitting along with each value being weighted by the population present in the associated DA is assumed.
+To summarize the decision tree algorithm, it is a supervised learning technique that begins by splitting a subset of the input data into branches based on a selected feature with the lowest impurity score (i.e., the least mixing of label membership within a branch).
+This process is recursively repeated for each data subset, selecting the next best features.
+Ultimately, the data is classified into distinct classes, with class membership explained by traversing the branches defined by the features that characterise the partitions within the decision tree.
+Notably, the absence of features from the decision tree does not necessarily imply they are irrelevant for classification, rather, they are just less relevant than other features.
 Put another way, when considering features that are highly correlated, such as income level and LICO, not all relevant variables may be present in the tree [@victorianoTimeSpaceMoney2020].

### 2024-09-30 paper/rejected-text.qmd
draft of Abstract through Methods

@@
 ---
+
+## rejected from intro
+
+(e.g., Paris mayor Anne Hidalgo proposed plan in her 2020 re-election campaign [@teixeiraClassifying15minuteCities2024], neighbourhood guidance in Shanghai, China [@weng15minuteWalkableNeighborhoods2019])
 
@@
 Finally, because women tend to earn less money than men and cars are a large financial burden, planning paradigms that reduce car dependency, such as 15 minute cities, may on average benefit women more than men [@Bertolo2023].
+
+Furthermore, mobility of care trip characteristics differ from other trips, notably those of the commute to work.
+For instance, many of these trips are made through trip-chains, i.e., when a non-work-related stop is made during the commute.
+Women have been found to do so more than men, which is understood as being due to their need to balance unpaid-household serving care and paid employment (e.g., picking up groceries or children on the way home from work) [@ravensbergen2023exploratory; @scheinerWomenComplexDaily2015].
 

### 2024-10-04 data/04-bycategoriy-clustering.qmd
Updates including Lucias comments

--- 2024-09-27 data/04-bycategoriy-clustering.qmd
@@
 ```{r}
-access_15_wide |> select(total_tot_copp_isc_groc,total_tot_copp_isc_dep,total_tot_copp_isc_health,total_tot_copp_isc_err,div_allcats) |> summary()
-access_15_wide |> group_by(scluster) |> summarise(total_tot_copp_isc_groc = median(total_tot_copp_isc_groc),
-                                                  total_tot_copp_isc_dep = median(total_tot_copp_isc_dep),
-                                                  total_tot_copp_isc_health = median(total_tot_copp_isc_health),
-                                                  total_tot_copp_isc_err = median(total_tot_copp_isc_err),
-                                                  div_allcats = median(div_allcats),
-                                                  pc = n()/143882)
+# norm_tot_copp_isc_norm_groc + norm_Convience + norm_Grocer + #div_allcats +
+#                                               norm_tot_copp_isc_norm_dep + norm_Rec + norm_Daycare + norm_LTC + norm_Park + norm_School + norm_Senior + #div_dep +
+#                                               norm_tot_copp_isc_norm_health + norm_Pharmacy + norm_Dentist + norm_Hos_clinc + #div_health +
+#                                               norm_tot_copp_isc_norm_err + norm_PostOff + norm_Library+ norm_Bank_ATM + div_allcats
+
+#total_tot_copp_isc_dep,total_tot_copp_isc_health,total_tot_copp_isc_err,div_allcats
+#'total_Community or Recreation Centre', "total_Daycare or EarlyON","total_Long_Term Care or Retirement Home","total_Park","total_School",'total_Senior Centre'
+
+access_15_wide |> 
+  select(c(
+    scluster,total_tot_copp_isc_groc,total_tot_copp_isc_dep,total_tot_copp_isc_health,total_tot_copp_isc_err,div_allcats
+    #scluster,norm_tot_copp_isc_norm_groc, norm_tot_copp_isc_norm_dep,norm_tot_copp_isc_norm_health,norm_tot_copp_isc_norm_err,div_allcats
+    )) |> 
+  mutate(tot = rowSums(across(c(total_tot_copp_isc_dep,total_tot_copp_isc_health,total_tot_copp_isc_err))))|>
+  group_by(scluster) |>
+  summarise_all(list(median))|>
+    mutate(lab = case_when(scluster == 1 ~ "A+",
+                               scluster == 2 ~ "A",
+                               scluster == 3 ~ "A-",
+                               scluster == 4 ~ "D",
+                               scluster == 5 ~ "B",
+                               scluster == 6 ~ "B+",
+                               scluster == 7 ~ "F")) |> arrange(tot)
+
+  # total_tot_copp_isc_groc = mean(total_tot_copp_isc_groc),
+  #                                                 total_tot_copp_isc_dep = mean(total_tot_copp_isc_dep),
+  #                                                 total_tot_copp_isc_health = mean(total_tot_copp_isc_health),
+  #                                                 total_tot_copp_isc_err = mean(total_tot_copp_isc_err),
+  #                                                 div_allcats = mean(div_allcats),
+                                                  #pc = n()/143882) |>
+# access_15_wide |> 
+#   select(c(scluster,norm_tot_copp_isc_norm_groc, norm_tot_copp_isc_norm_dep,norm_tot_copp_isc_norm_health,norm_tot_copp_isc_norm_err,div_allcats)) |> summary()
 ```
@@
 The low diversities - low accessibility
-C7 - the WORST - below Q3 in everything
-
-Mixed diversities - Higher accessibility for some stuff
-C4 - above median dependent-centric - but low diversity and low everything else (THESE AREAS HAVE POTENTIAL! the need more GROC, HEALTH, and ERRs) 
-C5 - like medium - about median everything with slightly lower HIGH diversity
-C3 - exceptionally high dependent-centric, and high everything else but low diversity
-
-High diversities and high accessibility
-C6 - medium - about median everything with the highest diversity
-C2 - better than medium - slightly above median everything with high diversity
-C1 - the best - above Q3 everything
+
+Lower accessibility (under median) with mixed diversity
+F - C7 - the absolute worst, low in all but with some dep, making it very low diversity - "NOT GOOD - need to add it all"
+D - C4 - has nothing except dep, making it very low diversity - "NOT GOOD - need to add err, groc, health"
+
+B - C5 - LIKE C6 but slightly lower everything. Overall good balance so good diversity - "ALL AROUND GOOD" -- firmly average. 
+B+ - C6 - about median dep, above median groc and health and Q3 errand. Overall great balance so high diversity - "ALL AROUND GREAT"
+
+High accessibilities (high caring) -- all at or above median
+A- - C3 - about median groc, health, errand and Q3 dep-centric, so lower diversity score - "ALL AROUND GREAT - with really high children-centric dests" 
+A - C2 - above Q3 everything and high diversity - however all is slightly lower than A - "ALL AROUND EXCELLENT"
+A+ - C1 - the best - above Q3 everything with high diversity- "ALL AROUND SUPER EXCELLENT"
 
 THE LABELS:
-Not caring - not complete (C7)
-Somewhat caring - not complete (C4)
-Somewhat caring - complete (C5)
-Caring - somewhat complete (C3)
-Caring - very complete (C6)
-Very caring - complete (C2)
-Very caring - very complete (C1)
+
+F (Nothing much: Not caring & not complete) (C7)
+D (Daycares, parks and schools, not much else: Somewhat caring & not complete) (C4)
+
+B (Average: Caring & complete) (C5)
+B+ (Above average: Caring & very complete) (C6)
+
+A- (Excellent, but dep- heavy: Very caring & not complete (C3)
+A (Excellent: Very caring & very complete) (C2)
+A+ (Exceptional: very caring & very complete) (C1)
 
@@
          Total_copp_health = Hos_clinc+Dentist,
-         Total_copp_err = PostOff +Library+Bank_ATM + Pharmacy)
+         Total_copp_err = PostOff +Library+Bank_ATM + Pharmacy) |>
+  rename("GROCERY TOTALS" = "total_tot_copp_isc_groc",
+             "DEPENDENT TOTALS"= total_tot_copp_isc_dep,
+             "HEALTH TOTALS"= total_tot_copp_isc_health,
+             "ERRAND TOTALS"= total_tot_copp_isc_err, 
+             "CATEGORY DIVERSITY"=div_allcats)
 ```
@@
 #                                               PostOff +Library+Bank_ATM + Pharmacy + div_err))
+
 classmod.som <- rpart(data = df,
-                      formula = scluster ~ (norm_tot_copp_isc_norm_groc + norm_Convience + norm_Grocer + #div_allcats +
-                                              norm_tot_copp_isc_norm_dep + norm_Rec + norm_Daycare + norm_LTC + norm_Park + norm_School + norm_Senior + #div_dep +
-                                              norm_tot_copp_isc_norm_health + norm_Pharmacy + norm_Dentist + norm_Hos_clinc + #div_health +
-                                              norm_tot_copp_isc_norm_err + norm_PostOff + norm_Library+ norm_Bank_ATM + div_allcats#+ div_err
+                      formula = scluster ~ (`GROCERY TOTALS` +
+                                            #norm_tot_copp_isc_norm_groc + 
+                                              #norm_Convience + norm_Grocer + #div_allcats +
+                                              `DEPENDENT TOTALS` +
+                                              #norm_tot_copp_isc_norm_dep + 
+                                              #norm_Rec + norm_Daycare + norm_LTC + norm_Park + norm_School + norm_Senior + #div_dep +
+                                              `HEALTH TOTALS` +
+                                              #norm_tot_copp_isc_norm_health + 
+                                              #norm_Pharmacy + norm_Dentist + norm_Hos_clinc + #div_health +
+                                              `ERRAND TOTALS` + 
+                                              #norm_tot_copp_isc_norm_err + 
+                                              #norm_PostOff + norm_Library+ norm_Bank_ATM + 
+                                              `CATEGORY DIVERSITY`#+ div_err
                                             ))
@@
 # df$SC9 <- predicted_labels$`9`
-
-```
-
-Trying ggparty to visualise, but first convert the rpart object to a party object:
+```
+
+Variable importance:
+```{r}
+var_importance <- data.frame(imp = classmod.som$variable.importance)
+var_importance <- var_importance %>% 
+  tibble::rownames_to_column() %>% 
+  dplyr::rename("variable" = rowname) %>% 
+  dplyr::arrange(imp) %>%
+  dplyr::mutate(variable = forcats::fct_inorder(variable))
+ggplot2::ggplot(var_importance) +
+  geom_col(aes(x = variable, y = imp),
+           col = "black", show.legend = F) +
+  coord_flip() +
+  scale_fill_grey() +
+  theme_bw()
+
+```
+
+<!-- Trying ggparty to visualise, but first convert the rpart object to a party object: -->
+<!-- ```{r} -->
+<!-- classmod.sompar <- as.party(classmod.som) -->
+
+<!-- #Autoplot as ggplot: -->
+<!-- autoplot(classmod.sompar) -->
+<!-- ``` -->
+
+<!-- Frequency of parcels in the super clusters: -->
+<!-- ```{r} -->
+<!-- freq_sc <- ggplot(access_15_wide |>  -->
+<!--          mutate(scluster = factor(scluster))) + -->
+<!--   geom_bar(aes(x = "",  -->
+<!--                group = scluster,  -->
+<!--                fill = scluster)) + -->
+<!--   scale_fill_discrete(name = "Supercluster") #+ -->
+<!--   #theme_void() + -->
+<!--   #guides(fill = guide_legend(nrow = 1)) + -->
+<!--   #theme(legend.position = "bottom") -->
+
+<!-- #with new names and colours -->
+<!-- freq_sc_renamed <- ggplot(access_15_wide |>  -->
+<!--          mutate(scluster = factor(scluster))) + -->
+<!--   geom_bar(aes(x = "",  -->
+<!--                group = scluster,  -->
+<!--                fill = scluster)) + -->
+<!--   scale_fill_manual(name = "Neighbourhood typologies", -->
+<!--                     values = c( "#1A9850","#91CF60", "#D9EF8B", "#FFFFBF","#FEE08B","#FC8D59","#D73027"), -->
+<!--                     labels = c("F (Nothing much: Not caring & not complete) (C7)", -->
+<!-- "D (Daycares, parks and schools, not much else: Somewhat caring & not complete) (C4)", -->
+<!-- "B (Average: Caring & complete) (C5)", -->
+<!-- "B+ (Above average: Caring & very complete) (C6)", -->
+<!-- "A- (Excellent, but dep- heavy: Very caring & not complete (C3)", -->
+<!-- "A (Excellent: Very caring & very complete) (C2)", -->
+<!-- "A+ (Exceptional: very caring & very complete) (C1)")) -->
+
+<!-- ``` -->
+
+<!-- Extract the legend: -->
+<!-- ```{r} -->
+<!-- legd <- cowplot::get_legend(freq_sc) -->
+
+<!-- legd -->
+<!-- ``` -->
+
+Make a prettier plot:
 ```{r}
@@
 
-#Autoplot as ggplot:
-autoplot(classmod.sompar)
-```
-
-Frequency of parcels in the super clusters:
-```{r}
-freq_sc <- ggplot(access_15_wide |> 
-         mutate(scluster = factor(scluster))) +
-  geom_bar(aes(x = "", 
-               group = scluster, 
-               fill = scluster)) +
-  scale_fill_discrete(name = "Supercluster") #+
-  #theme_void() +
-  #guides(fill = guide_legend(nrow = 1)) +
-  #theme(legend.position = "bottom")
-
-#with new names and colours
-freq_sc_renamed <- ggplot(access_15_wide |> 
-         mutate(scluster = factor(scluster))) +
-  geom_bar(aes(x = "", 
-               group = scluster, 
-               fill = scluster)) +
-  scale_fill_manual(name = "Neighbourhood typologies",
-                    values = c( "#1A9850","#91CF60", "#D9EF8B", "#FFFFBF","#FEE08B","#FC8D59","#D73027"),
-                    labels = c("Not caring - not complete (C2)",
-                               "Somewhat caring - not complete (C1)",
-                               "Somewhat caring - complete (C4)",
-                               "Caring - somewhat complete (C3)",
-                               "Caring - very complete (C6)",
-                               "Very caring - complete (C5)",
-                               "Very caring - very complete (C7)"))
-
-```
-
-Extract the legend:
-```{r}
-legd <- cowplot::get_legend(freq_sc)
-
-legd
-```
-
-Make a prettier plot:
-```{r}
 party_plot <- ggparty(classmod.sompar) +
   geom_edge() +
-  geom_edge_label(size = 1.6) +
-  geom_node_label(aes(label = splitvar),
+  geom_edge_label(size = 2.5) +
+  geom_node_label(aes(label = splitvar ),
                   id = "inner",
@@
                                ylab(""),
-                               theme_void()
-                               #theme(axis.text = element_blank())
-                               ))
+                               theme_void(),
+                               scale_fill_manual(values = c("#1A9850","#91CF60", "#D9EF8B", "#F46D43", "#FDAE61", "#FEE08B","#D73027")),
+                              theme(axis.text = element_blank())))
 party_plot
@@
 ```{r}
-png(file= file.path(here("figures/decision_tree_profiles_sc.png")),width= 2404, height= 1600, units="px",  res=300)
-plot_grid(party_plot, legd, nrow = 1, rel_widths = c(1, 0.2))
+png(file= file.path(here("figures/access_sc_profiles_usingSCVARS.png")),width= 3000, height= 1600, units="px",  res=300)
+plot_grid(party_plot, legd_renamed, nrow = 1, rel_widths = c(2, 1.5))
 dev.off()
@@
 OKAY! SO LET"S MAKE SOME VISUAL PLOTS:
+
+"F (Nothing much: Not caring & not complete) (C7)",
+"D (Daycares, parks and schools, not much else: Somewhat caring & not complete) (C4)",
+"B (Average: Caring & complete) (C5)",
+"B+ (Above average: Caring & very complete) (C6)",
+"A- (Excellent, but dep- heavy: Very caring & not complete (C3)",
+"A (Excellent: Very caring & very complete) (C2)",
+"A+ (Exceptional: very caring & very complete) (C1)"
 
@@
     Total_copp_err = median(Total_copp_err)) |>
-  mutate(max_label = case_when(max_label_val == SC1 ~ "a. Very caring - very complete (SC1)",
-                               max_label_val == SC2 ~ "b. Very caring - mostly complete (SC2)",
-                               max_label_val == SC3 ~ "d. Somewhat caring - mostly complete (SC3)",
-                               max_label_val == SC4 ~ "c. Very caring - mostly complete (SC4)",
-                               max_label_val == SC5 ~ "j. not caring - somewhat complete (SC5)",
-                               max_label_val == SC6 ~ "e. Somewhat caring - mostly complete (SC6)",
-                               max_label_val == SC7 ~ "f. not caring - not complete (SC7)"))
-
-                    labels = c("Not caring - not complete (C2)",
-                               "Somewhat caring - not complete (C1)",
-                               "Somewhat caring - complete (C4)",
-                               "Caring - somewhat complete (C3)",
-                               "Caring - very complete (C6)",
-                               "Very caring - complete (C5)",
-                               "Very caring - very complete (C7)"))
+  mutate(max_label = case_when(max_label_val == SC1 ~ "a. (Exceptional: very caring & very complete) (C1)",
+                               max_label_val == SC2 ~ "b. (Excellent: Very caring & very complete) (C2)",
+                               max_label_val == SC3 ~ "c. (Excellent, but dep- heavy: Very caring & not complete (C3)",
+                               max_label_val == SC4 ~ "f. (Daycares, parks and schools, not much else: Somewhat caring & not complete) (C4)",
+                               max_label_val == SC5 ~ "e. (Average: Caring & complete) (C5)",
+                               max_label_val == SC6 ~ "d. (Above average: Caring & very complete) (C6)",
+                               max_label_val == SC7 ~ "g. (Nothing much: Not caring & not complete) (C7)"))
 
@@
 df_DA$parcel_count |> sum() #a check
+
+save(df_DA, file = "df_DA.RData")
 ```
@@
 hydro_p_LakeOntario_cropped<-st_crop(hydro_p_LakeOntario$geometry, Community_Boundaries$geometry)
+
+Escpartment_Boundary <- st_read("data-raw/boundaries/Escarpment.shp", quiet = TRUE) |> st_transform(Escpartment_Boundary, crs = 4326) |> st_cast("LINESTRING")
+Escpartment_Boundary_HAM <- Escpartment_Boundary |> st_intersection(City_Boundary) 
 ```
@@
 ```{r, eval=FALSE}
-total_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
-  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
-  tm_shape(df_DA) +
-  tm_polygons("Total_copp_groc",
-              border.alpha = 0.2,
-              #style = "quantile", n=4,
-              palette = "Purples",
-              title = "Total access to care destinations",) +
-  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
-  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
-  tm_compass(position = c("left", "top"), size=1.0)+
-  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4,
-            bg.color = "grey",
-            legend.position = c("right","top"))
-
-total_copp_plot
-```
-
-```{r, eval=FALSE}
-SClabel_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+# RColorBrewer::brewer.pal(n = 8, name = "RdYlGn")
+# "#D73027" "#F46D43" "#FDAE61" "#FEE08B" "#D9EF8B" "#A6D96A" "#66BD63" "#1A9850"
+
+SClabel_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
   tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
@@
               border.alpha = 0.2,
-              # palette = c("#44ce1b","#bbdb44", "#f7e379", "#f2a134","#e51f1f"),
-              # labels = (c("Very caring - very complete",
-              #             "Caring - Complete",
-              # "Somewhat caring - complete",
-              # "Somewhat caring - somewhat complete", 
-              # "Somewhat caring - somewhat complete")),
+              palette = c("#1A9850","#91CF60", "#D9EF8B", "#FEE08B","#FDAE61","#F46D43","#D73027"),
+              labels = c("(A+) Exceptional: very caring & very complete", #C1
+                         "(A ) Excellent: very caring & very complete ", #(C2)
+                         "(A-) Dep- heavy, otherwise excellent: very caring & not complete", #(C3)
+                         "(B+) Above average: caring & very complete", # (C6)
+                         "(B ) Average: caring & complete", #(C5)
+                         "(B- ) Dep- heavy, otherwise below average: somewhat caring & not complete", # (C4)
+                         "(D ) Below average: not caring & not complete"), # (C7)
               title = "Neighbourhood typology") +
   tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
-  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
-  tm_compass(position = c("left", "top"), size=1.0)+
+  tm_shape(Escpartment_Boundary) + tm_lines(col = "grey30", lty = "dashed", lwd=1.5, legend.show=TRUE)+
+  tm_scale_bar(position = c("left", "top"), breaks = c(0, 1, 5, 10)) +
+  tm_compass(position = c("right", "top"), size=1.0)+
   tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4,
             bg.color = "grey",
-            legend.position = c("right","top"))
-SClabel_copp_plot
-
-SCLab_and_total_copp_plot <- tmap_arrange(SClabel_copp_plot, total_copp_plot, nrow=2)
-```
-
-```{r, eval=FALSE}
-SCLab_and_total_copp_plot
-```
-
-```{r, eval=FALSE}
-tmap_save(SCLab_and_total_copp_plot, 
-          file= file.path(here("figures/SCLab_and_total_copp_plot.png")),
+            legend.text.size=0.8,
+            legend.title.size = 1.2,
+            legend.width = 2,
+            legend.position = c("left","bottom"))
+SClabel_plot
+
+tmap_save(SClabel_plot, 
+          file= file.path(here("figures/SClabel_plot.png")),
           dpi=300)
-
-tmap_save(SClabel_copp_plot, 
-          file= file.path(here("figures/SClabel_copp_plot.png")),
-          dpi=300)
-```
+```
+
+F (Nothing much: Not caring & not complete) (C7)
+D (Daycares, parks and schools, not much else: Somewhat caring & not complete) (C4)
+
+B (Average: Caring & complete) (C5)
+B+ (Above average: Caring & very complete) (C6)
+
+A- (Excellent, but dep- heavy: Very caring & not complete (C3)
+A (Excellent: Very caring & very complete) (C2)
+A+ (Exceptional: very caring & very complete) (C1)
 
@@
 Select data for classification:
-```{r}
-
-df <- access_15_wide |> left_join(access_15_aggregated, by = "from_id") |> 
+
+```{r}
+# #min-max normalization
+# rescale <- function(x, na.rm = TRUE) {
+#   # Remove NA values if na.rm is TRUE
+#   if (na.rm) {
+#     min_x <- min(x, na.rm = TRUE)
+#     max_x <- max(x, na.rm = TRUE)
+#   } else {
+#     min_x <- min(x)
+#     max_x <- max(x)
+#   }
+#   
+#   # Avoid division by zero (min == max) and ensure no Inf values
+#   if (min_x == max_x) {
+#     warning("All non-NA values are the same. Returning zeros.")
+#     return(rep(0, length(x)))
+#   }
+#   
+#   # Apply Min-Max normalization, keeping NAs where they were
+#   rescaled_x <- (x - min_x) / (max_x - min_x)
+#   
+#   # Replace potential Inf/-Inf values with NA, though this should not happen
+#   rescaled_x[is.infinite(rescaled_x)] <- NA
+#   
+#   return(rescaled_x)
+# }
+
+# RESCALING DOESN"T HAVE AN IMPACT!
+df <- access_15_wide |> 
+  left_join(access_15 |> dplyr::select(c("from_id","GeoUID")), by = "from_id") |>  
+  left_join(HAM_census_21, by=c("GeoUID")) |>
   transmute(scluster = factor(scluster),
-            Income_bottom = `v_CA21_1103: In bottom half of the distribution`/(`v_CA21_1103: In bottom half of the distribution` + `v_CA21_1121: In top half of the distribution`),
-            LICO = `v_CA21_1085: Prevalence of low income based on the Low-income cut-offs, after tax (LICO-AT) (%)`,
+            Income_bottom_n = `v_CA21_1103: In bottom half of the distribution`,
+            Income_top_n = `v_CA21_1121: In top half of the distribution`,
+            Income_bottom = Income_bottom_n / (Income_bottom_n + Income_top_n),
+            Income = `v_CA21_907: Median after-tax income of household in 2020 ($)` ,
+            LICO = `v_CA21_1085: Prevalence of low income based on the Low-income cut-offs, after tax (LICO-AT) (%)` ,
             Avg_numb_of_kids = `v_CA21_498: Average number of children in census families with children`,
             Not_in_labforce = `v_CA21_6504: Not in the labour force`/(`v_CA21_6498: Employed` + `v_CA21_6501: Unemployed` + `v_CA21_6504: Not in the labour force`),
-            Income_gini = `v_CA21_1142: Gini index on adjusted household after-tax income`,
+            Unemployed = (`v_CA21_6501: Unemployed`)/(`v_CA21_6498: Employed` + `v_CA21_6501: Unemployed` + `v_CA21_6504: Not in the labour force`),
+            Income_gini = `v_CA21_1142: Gini index on adjusted household after-tax income` ,
             Vis_minority = `v_CA21_4875: Total visible minority population`/ (`v_CA21_4914: Not a visible minority`+ `v_CA21_4875: Total visible minority population`),
-            Age_0to14 =(`v_CA21_11: 0 to 14 years`) / (`v_CA21_11: 0 to 14 years` + `v_CA21_68: 15 to 64 years` + `v_CA21_251: 65 years and over`),
+            Age_0to14 = (`v_CA21_11: 0 to 14 years`) / (`v_CA21_11: 0 to 14 years` + `v_CA21_68: 15 to 64 years` + `v_CA21_251: 65 years and over`) ,
+            Age_65above = (`v_CA21_251: 65 years and over`) / (`v_CA21_11: 0 to 14 years` + `v_CA21_68: 15 to 64 years` + `v_CA21_251: 65 years and over`) ,
             One_parent_families = `v_CA21_507: Total one-parent families` / (`v_CA21_507: Total one-parent families` + `v_CA21_500: Total couple families`),
-            walk_towork =  `v_CA21_7647: Walked` / (`v_CA21_7647: Walked` + `v_CA21_7635: Car, truck or van` + `v_CA21_7644: Public transit` + `v_CA21_7650: Bicycle` + `v_CA21_7653: Other method` ),
-            Owner_hld_CHN = `v_CA21_4308: % in core housing need (57)`,
-            Hld_ownership = `v_CA21_4306: % of owner households with a mortgage (58)`,
-            Tenant_subsi_hld = `v_CA21_4314: % of tenant households in subsidized housing (61)`,
-            Tenant_hld_CHH =`v_CA21_4316: % in core housing need (57)`,
+            walk_towork =  `v_CA21_7647: Walked`/ (`v_CA21_7647: Walked` + `v_CA21_7635: Car, truck or van` + `v_CA21_7644: Public transit` + `v_CA21_7650: Bicycle` + `v_CA21_7653: Other method` ),
+            bike_towork =  `v_CA21_7650: Bicycle` / (`v_CA21_7647: Walked` + `v_CA21_7635: Car, truck or van` + `v_CA21_7644: Public transit` + `v_CA21_7650: Bicycle` + `v_CA21_7653: Other method` ),
+            pt_towork = `v_CA21_7644: Public transit`/ (`v_CA21_7647: Walked` + `v_CA21_7635: Car, truck or van` + `v_CA21_7644: Public transit` + `v_CA21_7650: Bicycle` + `v_CA21_7653: Other method` ),
+            car_towork =  `v_CA21_7635: Car, truck or van`/ (`v_CA21_7647: Walked` + `v_CA21_7635: Car, truck or van` + `v_CA21_7644: Public transit` + `v_CA21_7650: Bicycle` + `v_CA21_7653: Other method` ),
+            Owner_hld_CHN = `v_CA21_4308: % in core housing need (57)` ,
+            Hld_ownership = `v_CA21_4306: % of owner households with a mortgage (58)` ,
+            Hld_sheltercosts = `v_CA21_4307: % of owner households spending 30% or more of its income on shelter costs (55)` ,
+            Tenant_sheltercosts = `v_CA21_4315: % of tenant households spending 30% or more of its income on shelter costs (55)` ,
+            Tenant_subsi_hld = `v_CA21_4314: % of tenant households in subsidized housing (61)` ,
+            Tenant_hld_CHH =`v_CA21_4316: % in core housing need (57)` ,
             Population,
@@
 
-# examining correlations
-cor(df |> select(-c("scluster")), use = "complete.obs")
-
-#Train a decision tree using the data (renamed!):
+#df |> summary()
+
+# # examining correlations
+#df_cor <- cov.wt(df |> select(-c("scluster")) |> na.omit(), wt =  na.omit(df$Population), cor=TRUE) |> as.data.frame()
+```
+```{r}
+#Train a decision tree using the data:
 classdemo.som <- rpart(data = df,
-                      formula = scluster ~ (Income_bottom + Age_0to14 + One_parent_families + Vis_minority + Not_in_labforce+ Income_gini+walk_towork+Owner_hld_CHN+Hld_ownership+Tenant_subsi_hld+Tenant_hld_CHH+LICO),
+                      formula = scluster ~ (One_parent_families + Avg_numb_of_kids + # Age_0to14+
+                                              Not_in_labforce+Vis_minority+Unemployed +
+                                              Owner_hld_CHN+Tenant_subsi_hld+Tenant_hld_CHH+
+                                              Hld_sheltercosts+ Tenant_sheltercosts+
+                                              LICO+Low_or_no_deg+Income_gini+#Income+
+                                              Income_bottom+
+                                              walk_towork #+car_towork
+                                            ),
+                                            #+  +
+                                            #+bike_towork+pt_towork+Hld_ownership
+                                            #++  Income +Age_65above+++Income_bottom+)
                       weights = Population
@@
 # rpart.plot(classdemo.som, cex = 0.5, type = 0, varlen = 8)
-```
+
+#rpart.plot(classdemo.som, main = "")
+```
+Low income, LICO, single parent, and no degree -- highly correlated.
 
@@
                                theme_void(),
-                               scale_fill_manual(values = c("#bbdb44","#44ce1b",  "#f2a134", "#f7e379","#e51f1f")),
+                               scale_fill_manual(values = c("#1A9850","#91CF60", "#D9EF8B", "#F46D43", "#FDAE61", "#FEE08B","#D73027")),
                               theme(axis.text = element_blank())))
 
+```
+```{r}
+party_plot
+
+#creating a legend
+freq_sc_renamed <- ggplot(access_15_wide |> 
+         mutate(scluster = factor(scluster))) +
+  geom_bar(aes(x = "", 
+               group = scluster, 
+               fill = scluster)) +
+  scale_fill_manual(name = "Neighbourhood typologies",
+                    values = c("#1A9850","#91CF60", "#D9EF8B", "#FEE08B","#FDAE61","#F46D43","#D73027"),
+              labels = c("(A+) Exceptional: very caring & very complete", #C1
+                         "(A ) Excellent: very caring & very complete ", #(C2)
+                         "(A-) Dep- heavy, otherwise excellent: very caring & not complete", #(C3)
+                         "(B+) Above average: caring & very complete", # (C6)
+                         "(B ) Average: caring & complete", #(C5)
+                         "(B- ) Dep- heavy, otherwise below average: somewhat caring & not complete", # (C4)
+                         "(D ) Below average: not caring & not complete") # (C7)
+  ) 
+freq_sc_renamed
+
 legd_renamed <- cowplot::get_legend(freq_sc_renamed)
-party_plot
+
+```
+Variable importance:
+```{r}
+var_importance <- data.frame(imp = classdemo.som$variable.importance)
+var_importance <- var_importance %>% 
+  tibble::rownames_to_column() %>% 
+  dplyr::rename("variable" = rowname) %>% 
+  dplyr::arrange(imp) %>%
+  dplyr::mutate(variable = forcats::fct_inorder(variable))
+ggplot2::ggplot(var_importance) +
+  geom_col(aes(x = variable, y = imp),
+           col = "black", show.legend = F) +
+  coord_flip() +
+  scale_fill_grey() +
+  theme_bw()
+
 ```
@@
 #plot and save
-# NOTE: edited this in paint.. be careful saving over
-png(file= file.path(here("figures/access_sc_profiles_renamed.png")),width= 2600, height= 1600, units="px",  res=300)
-plot_grid(party_plot, legd_renamed, nrow = 1, rel_widths = c(1, 0.3))
+# NOTE: combined and edited this in paint, becareful when saving over
+png(file= file.path(here("figures/access_sc_profiles_justincome.png")),width= 2600, height= 1600, units="px",  res=300)
+plot_grid(party_plot)
 dev.off()
@@
 
-This is interesting!
-So for parcels with less than 45% of bottom-income people (richer) -- parcels with higher proportion of 1 parent households (13% and above) are more likely to have bad care/incomplete access (c3) than _really_ bad care/very incomplete access (c5).
-
-For parcels with higher than 45% of bottom-income (poorer), those with high a high proportion of children (age 0 to 14) are likely to have bad care/incomplete acces (c3).
-But those with a lower proportion of children - especially 9.8% LICO and above are going to have _super high_ care/complete (c1) or under 9.8% LICO will have _high_ care/complete (c2)
-For less than 63%, it's a different story. If between 45% and 63%, they are described by cluster 1 (caring and complete).
+```{r, eval=FALSE}
+png(file= file.path(here("figures/access_sc_profiles_legend.png")),width= 2600, height= 1600, units="px",  res=300)
+plot_grid(legd_renamed)
+dev.off()
+```

### 2024-10-04 paper/Manuscript.qmd
Updates including Lucias comments

@@
 library(gt)
+#install.packages("gtExtras")
+library(gtExtras)
 ```
@@
 load(glue::glue(here::here(), "/data/access_15_wide.RData"))
+load(glue::glue(here::here(), "/data/df_DA.RData"))
 ```
@@
 
+```{r creating boxplot, eval=FALSE}
+summary_boxplots <- access_15_wide |> select(c(scluster, starts_with("norm_"), div_allcats)) |>
+  mutate(scluster = case_when(scluster == 1 ~ "A+",
+                              scluster == 2 ~ "A",
+                              scluster == 3 ~ "A-",
+                              scluster == 4 ~ "B-",
+                              scluster == 5 ~ "B",
+                              scluster == 6 ~ "B+",
+                              scluster == 7 ~ "D"),
+         scluster = factor(scluster, ordered = TRUE,
+                           levels = c("A+", "A", "A-","B+","B","B-","D")))
+
+# summary_boxplots_Aplus <- summary_boxplots |> filter(scluster == "A+") |> select(-c("scluster"))
+# summary_boxplots_A <- summary_boxplots |> filter(scluster == "A") |> select(-c("scluster"))
+# summary_boxplots_Aminus <- summary_boxplots |> filter(scluster == "A-") |> select(-c("scluster"))
+# summary_boxplots_Bplus <- summary_boxplots |> filter(scluster == "B+") |> select(-c("scluster"))
+# summary_boxplots_B <- summary_boxplots |> filter(scluster == "B") |> select(-c("scluster"))
+# summary_boxplots_Bminus <- summary_boxplots |> filter(scluster == "B-") |> select(-c("scluster"))
+# summary_boxplots_D <- summary_boxplots |> filter(scluster == "D") |> select(-c("scluster"))
+
+summary_boxplots_tidy <- summary_boxplots |>
+  select(c("scluster","div_allcats",
+           "norm_tot_copp_isc_norm_dep", "norm_tot_copp_isc_norm_err",
+           "norm_tot_copp_isc_norm_groc", "norm_tot_copp_isc_norm_health")) |>
+  rename("1"="div_allcats",
+         "4"="norm_tot_copp_isc_norm_dep", 
+         "2"="norm_tot_copp_isc_norm_err",
+         "5"="norm_tot_copp_isc_norm_groc", 
+         "3"="norm_tot_copp_isc_norm_health") |>
+  pivot_longer(!scluster, names_to="variables", values_to = "normalized")
+```
+
+```{r, eval=FALSE}
+box_plot <- ggplot(summary_boxplots_tidy, 
+       aes(x=variables, y=normalized, fill=scluster)) +
+  geom_boxplot(outlier.shape = NA,
+               linewidth = 0.1) +
+  scale_fill_manual(values=c("A+" = "#1A9850",
+                               "A" = "#91CF60", 
+                               "A-" = "#D9EF8B", 
+                               "B+" = "#FEE08B",
+                               "B" = "#FDAE61",
+                               "B-" = "#F46D43",
+                               "D" = "#D73027"),
+                    name = " ") +
+    scale_x_discrete(labels = c("1" = "Category Totals",
+                                "2" = "ERRAND TOTALS",
+                                "3" = "HEALTH TOTALS",
+                                "4" = "DEPENDENT TOTALS",
+                                "5" = "GROCERY TOTALS")) +
+  coord_flip()+
+  theme_bw() +
+  theme(axis.title.y=element_blank()) +
+  ylab("Min-max normalized values") 
+box_plot
+ggsave(file= file.path(here("figures/box-plot.png")))
+```
+
+```{r creating-table2-pt1}
+summary_df_ALL <- access_15_wide |> select(c(starts_with("total_"), div_allcats)) |>
+  group_by() |>
+  summarise(across(everything(), list(mean = mean, sd = sd))) %>%
+ mutate_if(is.numeric, round, digits =1) |>
+  mutate_all(as.character)
+summary_df_ALL <- cbind(data.frame(scluster = "TOTAL"),summary_df_ALL)
+
+summary_df <- access_15_wide |> select(c(scluster, starts_with("total_"), div_allcats)) |>
+  mutate(scluster = case_when(scluster == 1 ~ "A+",
+                              scluster == 2 ~ "A",
+                              scluster == 3 ~ "A-",
+                              scluster == 4 ~ "B-",
+                              scluster == 5 ~ "B",
+                              scluster == 6 ~ "B+",
+                              scluster == 7 ~ "D")) |>
+  group_by(scluster) |>
+  summarise(across(everything(), list(mean = mean, sd = sd))) %>%
+ mutate_if(is.numeric, round, digits =1) |>
+  mutate_all(as.character)
+
+merge_every_two_columns <- function(df, num_rows) {
+  # Loop through every second column starting from column 2
+  for (col in seq(2, ncol(df), by = 2)) {
+    if (col < ncol(df)) {  # Ensure the column has a next one to merge with
+      for (i in 1:num_rows) {
+        df[i, col] <- paste0(df[i, col], " (", df[i, col + 1], ")")
+      }
+    }
+  }
+  return(df)
+}
+
+summary_df <- merge_every_two_columns(summary_df, 7) |>  select(-contains("sd"))
+summary_df_ALL <- merge_every_two_columns(summary_df_ALL, 1)  |>  select(-contains("sd"))
+summary_df <- rbind(summary_df, summary_df_ALL)
+
+summary_df <- summary_df |>
+  pivot_longer(!scluster) |>
+  pivot_wider(names_from =scluster )
+
+summary_df_ALL <- summary_df_ALL |>
+  pivot_longer(!scluster) |>
+  pivot_wider(names_from =scluster )
+
+summary_df$name <- c("GROCERY TOTALS", "Convenience Store", "Grocer",
+                     "DEPENDENT TOTALS", "Community or Rec. centre", "Daycare or EarlyON","LTC or retirment home","Park","School","Senior centre",
+                     "HEALTH TOTALS","Dentist","Hospital or clinic","Pharmacy",
+                     "ERRAND TOTALS","Bank or ATM","Library", "Post office",
+                     "Category diversity")
+
+summary_df <- summary_df[,c(1,3,2,4,6,5,7,8,9)]
+```
+
+```{r, echo=FALSE}
+df_DA_nogeom <- df_DA |> st_drop_geometry()
+df_DA_nogeom <- df_DA_nogeom[,30:81]
+
+df_DA_nogeom <- df_DA_nogeom |> mutate(
+  Income = `v_CA21_907: Median after-tax income of household in 2020 ($)`,
+  Avg_numb_of_kids = `v_CA21_498: Average number of children in census families with children`,
+  Not_in_labforce = `v_CA21_6504: Not in the labour force`/(`v_CA21_6498: Employed` + `v_CA21_6501: Unemployed` + `v_CA21_6504: Not in the labour force`),
+            Unemployed = (`v_CA21_6501: Unemployed`)/(`v_CA21_6498: Employed` + `v_CA21_6501: Unemployed` + `v_CA21_6504: Not in the labour force`),
+            Income_gini = `v_CA21_1142: Gini index on adjusted household after-tax income` ,
+            Vis_minority = `v_CA21_4875: Total visible minority population`/ (`v_CA21_4914: Not a visible minority`+ `v_CA21_4875: Total visible minority population`),
+            walk_towork =  `v_CA21_7647: Walked`/ (`v_CA21_7647: Walked` + `v_CA21_7635: Car, truck or van` + `v_CA21_7644: Public transit` + `v_CA21_7650: Bicycle` + `v_CA21_7653: Other method` ),
+            bike_towork =  `v_CA21_7650: Bicycle` / (`v_CA21_7647: Walked` + `v_CA21_7635: Car, truck or van` + `v_CA21_7644: Public transit` + `v_CA21_7650: Bicycle` + `v_CA21_7653: Other method` ),
+            pt_towork = `v_CA21_7644: Public transit`/ (`v_CA21_7647: Walked` + `v_CA21_7635: Car, truck or van` + `v_CA21_7644: Public transit` + `v_CA21_7650: Bicycle` + `v_CA21_7653: Other method` ),
+            car_towork =  `v_CA21_7635: Car, truck or van`/ (`v_CA21_7647: Walked` + `v_CA21_7635: Car, truck or van` + `v_CA21_7644: Public transit` + `v_CA21_7650: Bicycle` + `v_CA21_7653: Other method` ),
+            Owner_hld_CHN = `v_CA21_4308: % in core housing need (57)` ,
+            Hld_ownership = `v_CA21_4306: % of owner households with a mortgage (58)` ,
+            Hld_sheltercosts = `v_CA21_4307: % of owner households spending 30% or more of its income on shelter costs (55)` ,
+            Tenant_sheltercosts = `v_CA21_4315: % of tenant households spending 30% or more of its income on shelter costs (55)` ,
+            Tenant_subsi_hld = `v_CA21_4314: % of tenant households in subsidized housing (61)` ,
+            Tenant_hld_CHH =`v_CA21_4316: % in core housing need (57)` ,
+            Low_or_no_deg = (`v_CA21_5868: No certificate, diploma or degree` + `v_CA21_5871: High (secondary) school diploma or equivalency certificate`) / (`v_CA21_5868: No certificate, diploma or degree` + `v_CA21_5871: High (secondary) school diploma or equivalency certificate` + `v_CA21_5874: Postsecondary certificate, diploma or degree`))
+```
+```{r, echo=FALSE}
+cors_df_DA_nogeom <- cor(df_DA_nogeom, use = "complete.obs") |> as.data.frame()
+#cors_df_DA_nogeom[cors_df_DA_nogeom <0.7 | cors_df_DA_nogeom == 1] <- ""
+
+One_parent_family_inspect <- cors_df_DA_nogeom |> select(One_parent_families)
+LICO_inspect <- cors_df_DA_nogeom |> select(LICO)
+Income_inspect <- cors_df_DA_nogeom |> select(Income)
+```
+
 # Abstract
@@
 Inequalities in the relationship between the identified caring typologies and social-economic indicators of who currently reside in the neighbourhoods were then explored.
-Our results suggest the majority of 'caring and complete' neighbourhoods are within the urban core where a higher proportion of lower-income and single-parent households reside.
+Our results suggest the majority of 'caring and complete' neighbourhoods are within the urban core where a higher proportion of lower-income households reside.
+Contrawise, neighbourhoods that are 'not caring and not complete' are located in peripheral regions that tend to have higher household incomes.
 Further, certain neighbourhoods in suburban and rural communities are 'caring and somewhat complete', showing the potential to become 'complete' through possible urban policy intervention.
-Taken together, this study offers a theoretical bridge to connect Mobility of Care and the 15-minute neighbourhoods concept, a data-driven classification and clustering methodology to assess neighbourhoods on a 'caring and complete' continuum, and discussion of relevant gender mainstreaming policy interventions.
+Implications of neighbourhoods that have a high potential to be improved, and socio-demographic characteristics associated with income such as single-parent households and economic poverty are discussed.
+Taken together, this study offers a theoretical bridge to connect Mobility of Care and the 15-minute neighbourhoods concept, a data-driven classification and clustering methodology to assess neighbourhoods on a 'caring and complete' continuum, and discussion of implications relevant to gender mainstreaming within transportation planning.
 
@@
 
-The "15-Minute City" has been recently adopted by leaders as a way to promote human-scaled cities [@teixeiraClassifying15minuteCities2024; @teixeiraClassifying15minuteCities2024].
+The "15-Minute City" has been recently adopted by leaders as a way to promote human-scaled cities [@teixeiraClassifying15minuteCities2024].
 As introduced in @moreno_introducing_2021, the 15-Minute City is a urban planning model based in chrono-urbanism, a theory that emphasizes the positive impact on quality of urban life when urban space becomes multi-rythmic, in opposition to the segregation of urban time and space for individual uses and mobility [@mulicekUrbanRhythmsChronotopic2015; @morenoVilleQuartHeure2016].
 It is in opposition to the forces that supported the formation of single-use neighbourhoods such as industrial Fordism and single-use zoning [@mulicekUrbanRhythmsChronotopic2015; @morenoVilleQuartHeure2016].
-The "15-Minute City" is a utopic ideal, it is a city where one that accommodates the ability for all to reach important destinations within a walkable, bikeable or transit supportive radius.
+The "15-Minute City" refers to a city where relevant destinations within a walkable (or bikeable or transit supportive) radius are reachable by all.
 This form of neighbourhood planning would allow individuals to reclaim time spent on car mobility, giving way to sustainable modes and prompting urban spaces that are responsive to human needs and environmental sensibilities [@Allam2022].
@@
 As recent examples, planned neighbourhoods dominated post-WWII built urban form: the Neighbourhood Unit Concept in the western world [@brodyNeighbourhoodUnitConcept2013] and a parallel concept of the “mikrorayons” neighbourhoods in the Soviet Union  [@kissfazekasCircleParadigms15minute2022].
-These planning forms have also been extensively critiqued: as summarized in @talenSocialSciencePlanned2017, urban planners struggled to determine suitably universal design aspects such as neighborhood size, the type and quality of amenities, and local and regional connectivity.
-Social scientists also emphasized that top-down neighbourhood planning approaches overestimated the built environment's influence on social life [@talenSocialSciencePlanned2017].
-Neighbourhood planning approaches such as the 15-Minute City concept are being re-embraced by some, but from bottom-up perspective.
+Along different dimensions, these planning forms have been extensively critiqued [@talenSocialSciencePlanned2017].
+In this respect, neighbourhood planning approaches such as the 15-Minute City concept are being re-embraced by some, but from bottom-up perspective using co-creation tools and meaningful resident participation to achieve equitable and justice neighbourhood forms [@mahmoudCocreationPathwayUrban2021].
 How to equitably prescribe urban form that 'enables' contact with social opportunities (instead of 'engendering it' from the top-down) is still under debate.
 For instance, questions such as _what_ are destinations that matter, by what mode and travel time threshold, and _for whom_ remain under discussion among proximity-based planners who frequently use accessibility-based tools [@silvaProximitycentredAccessibilityConceptual2023; @silvaRegionalAccessibilityUndermining2022; @guzmanProximityEnoughCritical2024].
+
+From another perspective, feminist geographers, urban planners and other research groups agree that urban structure influences travel, but city planning fails short when planning for a neutral identity.
+In contrast, cities ought to be planned according to the multiple identities of the inhabitants [@vacchelli_towards_2018; @urbandevelopmentviennaGenderMainstreamingUrban2013].
+Researchers have demonstrated how mobility patterns differ according to gender [@lawWomenTransportNew1999; @cresswell_gendered_2008; @levy2013travel; @little_gender_1994; @tronto_toward_1990].
+Gender and social norms and cultural factors play a role in how, where, with whom, what for, and how far we move.
+Women have higher activities related to care [@ilo_care_2018], no matter at what life stage they are [@garciaromanGenderDifferencesTime2022].
+For all these reasons the spatial organization of activities has a different impact according to gender, the 15-Minute City concept cannot be gender blind.
 
@@
 
-In the last decades, the need for more sustainable, healthier, and liveable cities have received signficant attention.
+In the last decades, the need for more sustainable, healthier, and livable cities have received significant attention.
 Planners and decision-makers have proposed alternatives to mobility-based neighbourhood planning principles to ones that center proximity to urban functions [@pozoukidou15MinuteCityDecomposing2021].
@@
 
-`r word_comment("Lucia, do I correctly reflect the plans you reviewed?", " ")`
 Reflecting the flexible and aspirational presentation of the 15-Minute City; cities that have adopted the concept have done so using a diverse range of definitions and tools along with indistinct universal approaches.
-An early adopter was Portland, US in the Portland Plan [@city_of_portland_20minute_2010] of April 2012.
-It aims to foster an inclusive city development based on prosperity, education, health, and equity over a 30 year horizon.
-It promotes neighbourhood self-sufficiency and connectivity to city centers and centres of employment.
+A trail blazer was Portland, US in the Portland Plan [@city_of_portland_20minute_2010] of April 2012.
+Adopted before the '15-Minute City' concept of today, it aims to foster an inclusive city development based on prosperity, education, health, and equity over a 30 year horizon.
+The plan aims to promote neighbourhood self-sufficiency and connectivity to city centers and centres of employment.
 The progress report describes a high-level focus on equitable service delivery to all residents with equity concerning topics related to racial equity and people with disabilities [@portland_government_portland_2017], similarly taking on an "all populations" approach.
@@
 The k-means algorithm is another popular method; @ganUnderstandingUrbanMobility2020 applies it to analyse travel behaviour and classify metro stations based on mobility patterns.
-Combing classification techniques is also common; for instance, @delmelleDifferentiatingPathwaysNeighborhood2017 uses self-organizing maps (SOM) algorithm to group US neighbourhoods by minimising dissimilarity in attributes before applying k-means for further classification.
+Similarly, [@montero2023role; @montero2023applying] cluster travel behavior profile typologies representing different population segments and place of residence in Barcelona Metropolitan Area.
+Combing classification techniques is also common; for instance, @delmelleDifferentiatingPathwaysNeighborhood2017 uses self-organizing maps (SOM) algorithm to group US neighbourhoods by minimizing dissimilarity in attributes before applying k-means for further classification.
 Similarly, @victorianoTimeSpaceMoney2020 uses SOM to classify individuals' travel patterns according to the dissimilarity in mobility attributes and uses the decision tree algorithm to partition the data into interpretable classifications.
@@
 Conventionally, caring activities have been borne on women’s shoulders [@haydenGrandDomesticRevolution1982; @hochschild_second_2012].
-According to @ilo_care_2018, women and girls perform more than three-quarters of the total amount of unpaid care, a gender gap that varies geographically [@RN5].
+According to @ilo_care_2018, women and girls perform more than three-quarters of the total amount of unpaid care worldwide, a gender gap that varies geographically [@RN5].
 This unequal share of caring responsibilities leads to multifaceted gendered differences: in career development, profession selection, contract type, pay gap, and time poverty, as recognized by various international organisations [@EIGE2016; @ilo_care_2018].
@@
 37).
-As a product of the masculinst bias in planning, transport surveys and tools do not directly capture mobility of care, which re-enforces the idea that these trips are not a significant part of daily mobility.
+As a product of the masculinst bias in planning, transport surveys and tools most of the times do not directly capture mobility of care, which re-enforces the idea that these trips are not a significant part of daily mobility.
 In this respect, the feminist perspective of the cities of proximity is still underestimated with only few examples on the topic [@gil_sola_choose_2022; @macintyre_her_2022].
@@
 The objectives of this research are two-fold. 
-First, we aim to theoretically bridge Mobility of Care with the 15-Minute City concept to to re-imagine what local amenities matter and define an associated accessibility and diversity of accessibility measure.
-Second, taking a bottom-up data-driven approach, we apply a classification and clustering machine learning methodologies to support the bottom-up identification of _15-minute caring neighbourhoods_  through an empirical examination of the mid-cited city of Hamilton, Canada.
+- First, we aim to theoretically bridge Mobility of Care with the 15-Minute City concept to re-imagine what local amenities matter and define an associated accessibility and diversity of accessibility measure.
+- Second, taking a bottom-up data-driven approach, we apply a classification and clustering machine learning methodologies to support the bottom-up identification of _15-minute caring neighbourhoods_  through an empirical examination of the mid-cited city of Hamilton, Canada.
 These neighbourhoods are then interpreted on the continuum of having (and not having) the region-relative land-use and transportation infrastructure to support _completely caring_ access.
@@
 
-`r word_comment("I need to clean this up and firmly reiterate the goals", " ")`
-Due to it's popularity, the 15-Minute Concept is moving city planning towards a more people-centered approach focusing on accessibility to basic needs.
-While an aspirational ideal, the concept has faced criticism as a inadequately providing a way forward to address structural inequalities.
-In this context, we bridge the 15-Minute City with the Mobility of Care concept, to add significance to what destinations and associated trips matter for all, but especially along gendered and socio-economic lines.
+<!-- In this context, we bridge the 15-Minute City with the Mobility of Care concept, to add significance to what destinations and associated trips matter for all, but especially along gendered and socio-economic lines.
 We apply the "Caring 15-Minute City" concept to measure what neighourhoods in the mid-size Canadian city of Hamilton have the potential to be "caring 15 minute neighbourhoods".
-Making use of the tools reviewed and applied in the transportation research, the objectives of the paper are completed as follows:
+Making use of the tools reviewed and applied in the transportation research, the objectives of the paper are completed as follows: -->
 
-1)	The cumulative accessibility measure and the entropy measure are used to quantify both the level of access and the diversity of access to care destinations within a 15-minute walk.
-2) Based on these accessibility and diversity outputs, "Completely Caring 15-Minute Neighbourhood" typologies are generated using the SOM algorithm.  
-3)	A decision tree is then applied to these typologies and socio-economic variables of populations who reside in these neighbourhoods.
+<!-- 1)	The cumulative accessibility measure and the entropy measure are used to quantify both the level of access and the diversity of access to care destinations within a 15-minute walk. -->
+<!-- 2) Based on these accessibility and diversity outputs, "Completely Caring 15-Minute Neighbourhood" typologies are generated using the SOM algorithm.   -->
+<!-- 3)	A decision tree is then applied to these typologies and socio-economic variables of populations who reside in these neighbourhoods.
 This is done to delineate data-driven profiles on which groups have the potential to benefit from what Completely Caring 15-Minute Neighbourhood.
+-->
 
@@
 
-The mathematical formulation of takes the following form:
+The mathematical formulation takes the following form:
 $$
@@
 - $D_i$ is the diversity score.
-As $D_i$ represents evenness in type of categories, so if a parcel has a access score of $S_i^t$=0.5 for all types of destinations it will be assigned $D_i$ = 1 in the same was as if it had a score of $S_i^1$=10 for each type of destination.
+As $D_i$ represents evenness in type of categories, so if a parcel has a access score of $S_i^t$ =0.5 for all types of destinations it will be assigned $D_i$ = 1 in the same was as if it had a score of $S_i^t$ =10 for each type of destination.
 Contrariwise, a parcel can be assigned a low $D_i$ score if it has low but different accessibility scores for each category as well as if it has high but different scores for each category.
@@
 For the SOM step, the SOM implementation in the function 'trainSOM()' from {SOMbrero} R package is used [@VillaVialaneix2017].
-The input variables include the 143,882 parcels, each as individual observations along with 15 variables: the 14 calculated accessibility scores $S_i^$, normalized to the min-max range score within each $t$, and one diversity value $D_i$.
+The input variables include the 143,882 parcels, each as individual observations along with 15 variables: the 14 calculated accessibility scores $S_i^t$, normalized to the min-max range score within each $t$, and one diversity value $D_i$.
 Otherwise, defaults for all other parameters are assumed, relying on the data-driven heuristics embedded in the 'trainSOM()' function.
@@
 
-For the decision tree step, the supercluster-classified parcels identified in the SOM step are used as _labels_ and socio-economic and demographic indicators available from the most recent Canadian Census related to the mobility of care literature are used as _features_.
+For the decision tree step, the supercluster-classified parcels identified in the SOM step are used as _labels_ and socio-economic and demographic indicators available from the 2021 Canadian Census [@governmentofcanadaCensusPopulation2023] related to the mobility of care literature are used as _features_.
 This step is used to profile the superclusters so we may explore who reside in what supercluster in a data-driven way.
@@
 
-<!-- # Results -->
+# Results
 
-<!-- ## Overview of access to care destinations  -->
-<!-- How many care destinations can be reached within a 15-minute walk is summed for each of the four care categories defined in @bl-tbl1.
+## Overview of access to care destinations
+How many care destinations can be reached within a 15-minute walk is summed for each of the four care categories defined in @bl-tbl1.
 The median parcel value for each DA is visualised in @fig-Fig4.
@@
 Together, @fig-Fig4 and @fig-Fig5 represent the 15 variables that served as inputs into the SOM algorithm.
+
+```{r plot-figure4, out.width=600, out.height=600}
+#| label: fig-Fig4
+#| fig-cap: "The number of care destinations that can be reached within 15 min walk per care category for a median parcel in each DA.
+The values are a summary of the 14 acessibility scores that served as inputs into the SOM.
+Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
+knitr::include_graphics(glue::glue(here::here(), "/figures/percats_copp_plot.png"))
+```
+
+```{r plot-figure5, out.width=600, out.height=600}
+#| label: fig-Fig5
+#| fig-cap: "The diversity measure based on the proportion of care category spatial accessibility (figure 4).
+These values are a summary of the 15th input into the SOM.
+Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
+knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_entropy_copp_plot.png"))
+```
+
+## Identification of (in)completely (un)caring neighbourhoods
+<!--
+
+A+ is now A+
+A is now A
+A- is now A-
+
+B+ is now B+
+B is now B
+D is now B-
+
+F is now D
+
 -->
 
-<!-- ```{r plot-figure4, out.width=600, out.height=600} -->
-<!-- #| label: fig-Fig4 -->
-<!-- #| fig-cap: "The number of care destinations that can be reached within 15 min walk per care category for a median parcel in each DA.
-The values are a summary of the 14 acessibility scores that served as inputs into the SOM.
-Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]." -->
-<!-- knitr::include_graphics(glue::glue(here::here(), "/figures/percats_copp_plot.png")) -->
-<!-- ``` -->
-
-<!-- ```{r plot-figure5, out.width=600, out.height=600} -->
-<!-- #| label: fig-Fig5 -->
-<!-- #| fig-cap: "The diversity measure based on the proportion of care category spatial accessibility (figure 4).
-These values are a summary of the 15th input into the SOM.
-Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]." -->
-<!-- knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_entropy_copp_plot.png"))  -->
-<!-- ``` -->
-
-<!-- ## Identification of (in)completely (un)caring neighbourhoods -->
-
-<!-- Notably, in @fig-Fig4, the concentration of opportunities for all category types is centered in the downtown of Hamilton, with the highest concentration being in the lower city (near the lake shore) and pockets of the high concentration further from the shoreline.
+Notably, in @fig-Fig4, the concentration of opportunities for all category types is centered in the downtown of Hamilton, with the highest concentration being in the lower city (near the lake shore) and pockets of the high concentration further from the shoreline.
 Grocery-centric destinations appear to be the most concentrated, followed by health-centric and dependent-centric caring destinations.
@@
 In @fig-Fig5, areas that have high care accessibility tend to have high diversity as well, though there are exceptions in pockets of the city outside the downtown core that have lower but equally low levels of accessibility for all care categories.
+
+Based on the SOM classification methodology discussed and the 15 variables represented in @fig-Fig4 and @fig-Fig5, 7 superclusters are identified from the 10 by 10 grid of SOM nodes.
+As diagnostics for the selection of the number of superclusters, the dissimilarity-index-based dendrogram with the proportion of parcels represented in each supercluster alongside the variance explained plot is visualised in @fig-Fig6.
+
+```{r plot-figure6, out.width=800, out.height=800}
+#| label: fig-Fig6
+#| fig-cap: "The 7 resulting superclusters from the SOM the 10 by 10 grid output represented in a dendrogram (left) and the proportion of variance unexplained by cluster (right)."
+#|
+knitr::include_graphics(glue::glue(here::here(), "/figures/all_access_sc_plot_small.png"))
+```
+As seen in @fig-Fig6, the 7 superclusters are assigned labels from A to F, corresponding to their 'grade' of quantity of accessibility and diversity in accessibility per care category.
+What differentiates the dissimilarities in the clusters the most, i.e., nodes of grade A/A+ to everything else, is their high levels of care accessibility and often high diversity in care categories calculated for each parcel.
+A/A+ grades represent only 24% of the parcels, with the other grades being assigned lower accessibility and diversity of accessibility scores.
+To characterise the superclusters, a table providing summary statistics of each supercluster is detailed in @tbl-tbl2 and boxplots in @fig-Fig7.
+
+```{r creating-table2-pt2}
+#| tbl-cap: "Mean and (standard deviation) of each SOM classified cluster by input variable and additional summary variables (indicated by ALL CAPITAL LETERS)."
+#| label: tbl-Tbl2
+#|
+summary_df %>%
+  gt() %>%
+  tab_header(
+    title = " ",
+  ) %>%
+  cols_label(
+    name = " "
+  ) %>%
+  gt_highlight_cols(
+    2:3, fill = "green", alpha = 0.3
+    ) %>%
+  gt_highlight_cols(
+    4, fill = "green", alpha = 0.2
+    ) %>%
+  gt_highlight_cols(
+    5:6, fill = "orange", alpha = 0.3
+    ) %>%
+    gt_highlight_cols(
+    7, fill = "orange", alpha = 0.2
+    ) %>%
+  gt_highlight_cols(
+    8, fill = "red", alpha = 0.3
+    ) %>%
+  tab_style(
+    style = cell_borders(
+      sides = c("top"),
+      color = "black",
+      weight = px(1.5),
+      style = "solid"
+    ),
+    locations = cells_body(rows = c(1,4,11,15,19))
+  ) |>
+  tab_options(
+    table.font.size = 12,
+    table.width = pct(80)
+  )
+```
+
+```{r plot-figure7, out.width=600, out.height=300}
+#| label: fig-Fig7
+#| fig-cap: "A boxplot demonstrating summary variables and category diversity that define the superclusters."
+knitr::include_graphics(glue::glue(here::here(), "/figures/box-plot.png"))
+```
+
+The superclusters are characterized by caring access and diversity of caring access (completeness).
+This technique provides a data-driven approach to supercluster formation which we then attribute meaning to through the derivation of the following labels.
+From the descriptive statistics of the superclusters, interesting trends emerge:
+
+- A+, A, and A- all have exceptionally high or excellent caring accessibility for all destination types; all destination type and diversity scores are top quantile or above.
+Interestingly, A- is like A but with much higher dependent centric destinations, particularly parks, schools, and daycares, with high scores for everything else.
+This disbalance lowers its overall completeness score but is a supercluster that demonstrates potential in being retrofitted to provide A+ level completely caring access.
+Together, these grades represent 28% of all parcels.
+- B+, B and B- superclusters present about average completely caring access.
+These superclusters, particuarly B, can serve as the benchmark for what 'average' completely caring access in Hamilton currently looks like.
+B+ access and diversity scores are above B, while B- has low values for everything with exception to above average number of dependent-centric destinations, particularly parks, daycares and schools.
+These three clusters represent 34% of the parcels.
+- D superclusters demonstrate the lows scores all around, presenting plenty of room for improvement in complete and caring 15-minute access. This cluster represents the 38% of all parcels.
+
+<!-- D (Nothing much: Not caring & not complete) (C7) -->
+
+<!-- B- (Below average, but dep- heavy: Somewhat caring & not complete) (C4) -->
+<!-- B (Average: Caring & complete) (C5) -->
+<!-- B+ (Above average: Caring & very complete) (C6) -->
+
+<!-- A- (Excellent, but dep- heavy: Very caring & not complete (C3) -->
+<!-- A (Excellent: Very caring & very complete) (C2) -->
+<!-- A+ (Exceptional: very caring & very complete) (C1) -->
+
+To meaningfully visualise where the superclusters are located, we visualise the representative supercluster label in each DA in @fig-Fig8.
+This is done by grouping parcels by their DA, calculating the median supercluster label for each label, and selecting the label which is most dominate within that DA.
+For reference, the median number of parcels in each DA is `r df_DA$parcel_count |> median() |> round()`.
+Further, supercluster label membership within a DA is typically pure, with the median membership being `r df_DA$max_label_val |> median() |> scales::percent()` of a single cluster.
+
+`r word_comment("What do you think about these grades as labels?
+Maybe there's another way we can frame it.
+A+ and A are excellent, B+ and B are average, D are bad.
+However A- and B- show promise: they have some caring access (child-centric) and can be improved!
+A- can be more easily improved than B-.", " ")`
+```{r plot-figure8, out.width=600, out.height=600}
+#| label: fig-Fig8
+#| fig-cap: "The maximum median parcel supercluster membership per DA. Escarpment is visualised as a grey dashed line."
+knitr::include_graphics(glue::glue(here::here(), "/figures/SClabel_plot.png"))
+```
+The resulting superclusters visualised in @fig-Fig8 are a combination of the cumulative opportunity and diversity of access measures trasnlated into meaningful typologies through the SOM methodology.
+It can be observed that the excellent and exceptional neighbourhoods, which are very caring and very complete 15-minute neighbourhoods, are located within the center of the city and most closer to the shoreline in the 'lower' city (below the escarpment).
+The average and below average neighbourhoods are often proximate to the excellent and exceptional neighbourhoods and are more prevalent in the 'mountain' part of the center of Hamilton.
+The escarpment is a physical barrier, with few pedestrian-accessible access points to traverse; hence the typologies describing neighbouring DAs separated by the escarpment are often different.
+
+As accessibility is the measure of _potential_ interaction, e.g., how many care destinations one could reach within a 15-minute walk, these typologies are useful in identifying which areas of the city are providing regionally high levels of complete caring access, which areas are regionally average (oranges), and which areas are regionally below average.
+Furthermore, neighbourhoods which provide disbalanced access can be further investigated.
+For instance, neighbourhoods that provide A- provide otherwise high levels of 15-minute walking access to all types of destinations, but particularly to Community and recreational centres, daycares, parks, and schools.
+These A- neighbourhoods are suitable for all, but may be particularly desirable for households that have children, needed proximity to care destinations.
+Furthermore, B- neighbourhoods are like A- neighbourhoods as they provide high access to children-centric destinations like schools, parks, and daycares, however are low in all other destination types.
+These neighbourhoods can be understood as currently being conducive to completing children-centric care trips, and having the potential to become complete if additional destination types were located within their neighbourhoods.
+
+## Profiling of who resides in (in)completely (un)caring neighbourhoods
+
+To enhance the meaning of the superclusters beyond a descriptive and spatial conceptualisation of "caring" and "complete",  _who?_ resides in what neighbourhood currently can also be investigated.
+We run a decision tree and include the population-weighted socio-demographic characteristics of the parcels (retrieved from the 2021 Canadian Census) as input variables and the resulting output are 'socio-demographic superclusters' of the superclusters.
+The input variables included: median household income, % below the median household income, % LICO prevalence, average number of children per household, % population aged 0 to 14, % not in the labour force, % not employed, Gini index on adjusted household after-tax income, % visible minority, % single parent household, % who walk to work (relative to bike, care/truck/van, public transit and 'other'), % of owner household in core housing need (i.e., inadequate housing structure or paying higher than 30% income on housing), % of tenant households in subsidized housing, % of tenant households in core housing need, % no certification or with only a highschool diploma.
+
+Of all the included input variables, the one that proves most meaningful in partitioning the superclusters data into subsets was only one variable: median household income of the DA.
+@fig-Fig9 provides the splits that were significant in partitioning the data in the decision tree algorithim.
+While the algorithim is unable to homogenously use each supercluster, we still use it to report a narrative of who may reside in what caring/complete supercluster.
+Particularly, the decision tree demonstrates a more pure supercluster membership for only two of the three terminal decision tree nodes: supercluster A+ (exceptional completely caring access) and D (not caring and complete access).
+For reference, the city-wide mean household income is `r df_DA_nogeom$Income |> mean(na.rm=T)|> round() |> prettyNum(big.mark = ",") ` (SD: `r df_DA_nogeom$Income |> sd(na.rm=T)|> round() |> prettyNum(big.mark = ",")` and median: `r quantile(df_DA_nogeom$Income,0.5, na.rm=T) |> unname()  |> prettyNum(big.mark = ",")`).
+
+<!-- - % single parent household (median: `r df_DA$One_parent_families |> median(na.rm=T) |> scales::percent()` (`r df_DA$One_parent_families |> sd(na.rm=T) |> scales::percent()`) ), and  -->
+<!-- - % prevalence of LICO (median: `r df_DA$LICO |> median(na.rm=T) |> round() |> paste0("%")` (`r df_DA$LICO |> sd(na.rm=T) |> round() |> paste0("%")`) ) -->
+
+```{r plot-figure9, out.width=600, out.height=300}
+#| label: fig-Fig9
+#| fig-cap: "The decision tree demonstrating socio-ecnomic profiles of superclusters."
+knitr::include_graphics(glue::glue(here::here(), "/figures/access_sc_profiles_justincome.png"))
+```
+Describing @fig-Fig9, the most valuable variable for splitting the data was the median household income in each DA.
+The terminal nodes represented as bar columns depict the percentage composition by supercluster.
+For the column on the right, the majority of the parcels are D superclusters with the second most supercluster membership being B- parcels, described by having greater than 91,500 CAD household income, above the city-wide mean.
+The left column is dominated by A+ along with A and A- (in the largest quantities relative to other bars), though proportions of all other superclusters are also present.
+This column is defined by parcels present in DAs with lower incomes, e.g., 68,750 CAD.
+The middle columns's classification is mixed: with all types of superclusters present in DAs with about average households incomes (between 68,750 CAD and 91,500 CAD).
+
+Though median income was the most useful in partitioning the parcels by their supercluster labels, other variables are still important but were not as helpful in partitioning the data.
+Namely, the following variables were also important, in order of importance and correlation coefficient with the median household income variable in brackets: the % below the median household income (`r Income_inspect[c("Income_bottom"),] |> round(digits=2)`), % single parent household (`r Income_inspect[c("One_parent_families"),] |> round(digits=2)`), % no certification or with only a highschool diploma (`r Income_inspect[c("Low_or_no_deg"),] |> round(digits=2)`), the average number of children per household (`r Income_inspect[c("Avg_numb_of_kids"),] |> round(digits=2)`), % not in the labour force (`r Income_inspect[c("Not_in_labforce"),] |> round(digits=2)`), % LICO prevalence (`r Income_inspect[c("LICO"),] |> round(digits=2)`), and  % who walk to work (`r Income_inspect[c("walk_towork"),] |> round(digits=2)`).
+Of note, when median household income is removed from the decision tree, a similar structure is created, except with above average Single-Parent family and LICO prevalence splitting the data.
+It is notable that median household income is highly or moderately correlated with many of these important variables, and discussing these correlation is useful for interpretation alongside the tree diagram.
+
+For instance, single-parent families are most positively correlated with the proportion of households being in the bottom income distribution (`r Income_inspect[c("Income_bottom"),] |> round(digits=2)`) and low or new degree (`r One_parent_family_inspect[c("Low_or_no"),] |> round(digits=2)`), and most negatively with the median after-tax household income (`r One_parent_family_inspect[c("v_CA21_907: Median after-tax income of household in 2020 ($)"),] |> round(digits=2)`), reflecting nationwide trends [@governmentofcanadaDailyMainHighlights2024; @governmentofcanadaPrevalenceLowIncome2024].
+LICO is also an indicator of low-income, and though not all single-family households are below the LICO (in fact, single-family households are not highly correlated with LICO at the DA level in Hamilton), DAs where higher concentrations of single-family and where high LICO prevalence persist tend to have A+ access.
+These findings are notable as currently, the highest grade (A+) serves a disadvantaged group that is also often disadvantaged along different intersecting socio-economic characteristics.
+Also interestingly, DAs with lower single-parent households (who also tend to have higher incomes) are in DAs with a higher concentration of F parcels.
+These considerations are important when considering the equity of caring 15-minute policy interventions.
+
+<!-- Describing @fig-Fig9, the most valuable variable for splitting the data was the proportion of single family households and % of LICO prevalence in the DA.
+The the terminal nodes represented as bar columns depict the percentage composition by supercluster.
+For the bar on the right, the majority of the parcels are F superclusters.
+Following the tree diagram, these F superclusters are described by having less than 15.8% single parent families (this is below the median of `r df_DA$One_parent_families |> median(na.rm=T) |> scales::percent()`).
+The bar on the left is mostly composed of parcels that are labeled A+.
+These A+ parcels are located in DAs greater than 15.8% single parent families and % LICO prevalence of 6.5% and greater (this is above the median of `r df_DA$LICO |> median(na.rm=T) |> round() |> paste0("%")`).
+In sum: residences in parcels with exceptionally complete and caring access tend to be lower-income and single-parent families, while residences in parcels with not caring are not complete access tend to have a lower presence of single parent households.
+Households with a higher proportion of single-family households (15.8% and greater) where LICO is lower have a mixed label membership (bar column in @fig-Fig9).
 -->
 
-<!-- Based on the SOM classification methodology discussed and the 15 variables represented in @fig-Fig4 and @fig-Fig5, 7 superclusters are identified.
-As diagnostics for the selection of the number of superclusters, the dissimilarity-index-based dendrogram with the proportion of parcels represented in each supercluster alongside the variance explained plot is visualised in @fig-Fig6.
--->
-<!-- ```{r plot-figure6, out.width=800, out.height=800} -->
-<!-- #| label: fig-Fig6 -->
-<!-- #| fig-cap: "The 7 resulting superclusters from the SOM the 10 by 10 grid output represented in a dendrogram (left) and the variance unexplained (right)." -->
-<!-- #|  -->
-<!-- knitr::include_graphics(glue::glue(here::here(), "/figures/all_access_sc_plot_small.png"))  -->
-<!-- ``` -->
+# Discussion
 
-<!-- To characterise the superclusters, a table providing summary statistics of each supercluster is detailed in @tbl-tbl2: -->
-<!-- ```{r creating-table2} -->
-<!-- summary_df_ALL <- access_15_wide |> select(c(starts_with("total_"), div_allcats)) |> -->
-<!--   group_by() |>  -->
-<!--   summarise(across(everything(), list(median = median, sd = sd))) %>%  -->
-<!--  mutate_if(is.numeric, round, digits =1) |> -->
-<!--   mutate_all(as.character)   -->
-<!-- summary_df_ALL <- cbind(data.frame(scluster = "TOTAL"),summary_df_ALL) -->
+The descriptive statistics associated with each supercluster, their spatial distribution, and the socio-economic profiles that characterise them, are revealing points of discussions and further examined.
 
-<!-- summary_df <- access_15_wide |> select(c(scluster, starts_with("total_"), div_allcats)) |>  -->
-<!--   mutate(scluster = case_when(scluster == 1 ~ "A", -->
-<!--                               scluster == 2 ~ "B", -->
-<!--                               scluster == 3 ~ "D", -->
-<!--                               scluster == 4 ~ "F", -->
-<!--                               scluster == 5 ~ "E", -->
-<!--                               scluster == 6 ~ "C", -->
-<!--                               scluster == 7 ~ "G")) |> -->
-<!--   group_by(scluster) |>  -->
-<!--   summarise(across(everything(), list(median = median, sd = sd))) %>%  -->
-<!--  mutate_if(is.numeric, round, digits =1) |> -->
-<!--   mutate_all(as.character)  -->
+Hamilton is a city is not a caring 15-Minute City for all, but it is for some.
+There are evident spatial inequalities in areas that are, are not, and are somewhere between a caring and complete 15-minute neighbourhood (from labels A+ and A being excellent, B being about average, D being well below average, and A- and B- demonstrating potential in their grade group).
+The downtown core of the city has the most _caring 15-minute neighbourhoods_, but many pockets of the suburbs do as well.
+Though this contrasts what may be seen when considering employment opportunities [as in the work of @elgeneidyNonstopEquityAssessing2016, the paper's findings align with the neighbourhood unit planning literature reviewed.
+As many North American Suburbs, Hamilton followed certain characteristic design patterns that were initially designed to be proximate to amenities, coinciding with the conceptualisation of _caring 15-minute cities_ [@aliNeighbourhoodPlanningHamilton1991].
 
-<!-- merge_every_two_columns <- function(df, num_rows) { -->
-<!--   # Loop through every second column starting from column 2 -->
-<!--   for (col in seq(2, ncol(df), by = 2)) { -->
-<!--     if (col < ncol(df)) {  # Ensure the column has a next one to merge with -->
-<!--       for (i in 1:num_rows) { -->
-<!--         df[i, col] <- paste0(df[i, col], " (", df[i, col + 1], ")") -->
-<!--       } -->
-<!--     } -->
-<!--   } -->
-<!--   return(df) -->
-<!-- } -->
+Who in currently resides in Hamilton's caring 15-Minute neighbhorhoods is also currently an optimistic story.
+Parcels that provide A+ completely caring access tend to be in DAs that are economically disadvantaged.
+Economic disadvantage tends to intersect with other identities such as gender [@lightmanMeasuringEconomicExclusion2018].
+And as reviewed in this work, all women and especially those from lower income households tend to complete the most mobility of care trips [@ravensbergen2023exploratory].
+Furthermore, lower-income households tend to also be single-parent households.
+Broadly, single-parent households are also more likely to be time-disadvantaged [@nieuwenhuisSingleparentFamiliesWork2018], and tasked with a higher proportion of care duties [@craigTimeCareComparison2004].
+In this way, the most economically disadvantaged groups having A+ 15-minute care accessibility is an optimistic finding.
+However, Hamilton is experiencing gentrification [@ellisyoungWeReJust2018]; rents are rising along the future light rail transit corridor and throughout the city [@vandermerweSpilloverGentrificationMidsized2021; @mayersLightTransitDocumenting2023].
+The larger and high-rent city of of Toronto is spilling gentrification into Hamilton's downtown core, (re)producing neighbourhoods based on Toronto's middle class [@mayersLightTransitDocumenting2023].
+In this way, the lower income residents of Hamilton's A+ neighbourhoods are more likely to be displayed and forced to re-located; and if they want to stay in Hamilton, there are not many areas that provide the same exceptional level of access as those they are currently located in.
+This is a matter of wellbeing and justice.
 
-<!-- summary_df <- merge_every_two_columns(summary_df, 7) |>  select(-contains("sd")) -->
-<!-- summary_df_ALL <- merge_every_two_columns(summary_df_ALL, 1)  |>  select(-contains("sd")) -->
-<!-- summary_df <- rbind(summary_df, summary_df_ALL) -->
+But what is the most equitable way forward in working to increase caring 15-Minute neighbourhoods?
+The work's findings indicate that parcels that provide very low 15-minute completely caring access (F) are households that tend to be higher income.
+These areas tend to be more rural, single-use zoned and car-dependent.
+Further, of the parcels that provide high child-centrics but low otherwise, A- parcels (better access) tend to be in DAs with lower-income households more so than B- parcels (lower acess than A-).
+From the perspective of ameliorating land-use to support the equitable distribution of 15-minute caring neighbourhoods which areas be targeted?
+Who should benefit, and how?
+If the initiative is targeted to specific neighbourhoods: sustainability linked to car dependency and equity are in-tension.
+This harkens to what role a planned neighbourhood, and bottom-up planning approaches that should be adopted to achieve equitable 15-minute cities.
+In this way, the methodology presented shows potential to support city planning, by demonstrating where neighbourhods that provide walkable access to opportunities that matter to all (care destinations) are located as city-wide relative typologies and who currently resides there, as a stepping stone for further investigation.
 
-<!-- summary_df <- summary_df |>  -->
-<!--   pivot_longer(!scluster) |> -->
-<!--   pivot_wider(names_from =scluster ) -->
+As all research work, the results should also be interpreted along with methodological assumptions, namely, spatial accessibility yields a measure of potential interaction with all reachable destinations from an origin.
+These destinations however may not be relevant to people at an origin, e.g., the may be underutilized (walkable access to 2 schools is unneeded, as a child only needs to attend 1 school), the trip physically undesirable (the walk may be along an arterial with high traffic speeds, making the trip unlikely to ever happen by foot), or the average walking speed assumed may not reflect the walking speeds of all populations [@willberg15minuteCityAll2023].
+All-in-all, people who reside in these neighbourhoods may also disagree with the neighbourhood's label, the labels are region-relative (e.g., high accessibility in Hamilton may be subjectively insufficient for some) and they do not consider subjective perceptions that influence accessibility (e.g., though a neighbourhood has many opportunities, it may not feel safe to access them).
+Furthermore, accessibility is calculated from the point of residential parcels.
+Care trips are not necessarily completed from home, in-fact, they are often completed in complex trip-chains [@scheinerWomenComplexDaily2015].
+In this way, the results flatten the dynamic patterns of care trips.
 
-<!-- summary_df_ALL <- summary_df_ALL |>  -->
-<!--   pivot_longer(!scluster) |> -->
-<!--   pivot_wider(names_from =scluster ) -->
+# Conclusion
 
-<!-- summary_df$name <- c("GROCERY TOTALS", "Convenience Store", "Grocer", -->
-<!--                      "DEPENDENT TOTALS", "Community or Rec. centre", "Daycare or EarlyON","LTC or retirment home","Park","School","Senior centre", -->
-<!--                      "HEALTH TOTALS","Dentist","Hospital or clinic","Pharmacy", -->
-<!--                      "ERRAND TOTALS","Bank or ATM","Library", "Post office",  -->
-<!--                      "Category diversity") -->
-<!-- ``` -->
-
-<!-- ```{r creating-table2-pt2} -->
-<!-- #| tbl-cap: "Median and (standard deviation) of each SOM classified cluster by input variable and additional summary variables (indicated by ALL CAPITAL LETERS)." -->
-<!-- #| label: tbl-Tbl2 -->
-<!-- #|  -->
-<!-- summary_df %>% -->
-<!--   gt() %>% -->
-<!--   tab_header( -->
-<!--     title = " ", -->
-<!--   ) %>% -->
-<!--   cols_label( -->
-<!--     name = " " -->
-<!--   ) %>% -->
-<!--   tab_style( -->
-<!--     style = cell_borders( -->
-<!--       sides = c("top"), -->
-<!--       color = "black", -->
-<!--       weight = px(1.5), -->
-<!--       style = "solid" -->
-<!--     ), -->
-<!--     locations = cells_body(rows = c(1,4,11,15)) -->
-<!--   ) |> -->
-<!--   tab_options( -->
-<!--     table.font.size = 12, -->
-<!--     table.width = pct(80) -->
-<!--   ) -->
-<!-- ``` -->
-
-<!-- Notably, from A to G, the superclusters are characterized by caring access and diversity of caring access (completeness).
-From the descriptive statistics of the superclusters, interesting trends emerge: ...
-parcels that have an Errand-centric access of less than 0.5 opportunities (bottom quartile), a diversity score of less than 0.277 (bottom tertile), and any Dependent-centric access score of greater than 0.5 (and can take on any value of the remaining input variables).
-This technique provides a data-driven approach to supercluster formation which we then attribute meaning to through the derivation of the following labels:...
-In sum, we intuitively derive classification labels: -->
-
-<!-- G: Not caring - not complete (C7) -->
-<!-- F: Somewhat caring - not complete (C4) -->
-<!-- E: Somewhat caring - complete (C5) -->
-<!-- D: Caring - somewhat complete (C3) -->
-<!-- C: Caring - very complete (C6) -->
-<!-- B: Very caring - complete (C2) -->
-<!-- A: Very caring - very complete (C1) -->
-
-<!-- To further visualize supercluster membership, we spatially demonstrate which superclusters best describes each parcel in @fig-Fig8: <!-- Visually, the super clusters closely follow the quartile split of the accessibility to care destinations (all categories combined) and tertile split of the diversity scores (visualised in @fig-Fig5).
--->  -->
-<!-- <!-- `r word_comment("AN IDEA DISCUSSED: Overall, in @fig-Fig8 there are many pockets of good/great care accessibility+diversity of opportunities Hamilton.
-These are objectively caring neighbourhoods (BY MEASURE).
-However, people may subjectively disagree...
-Maybe they don't feel safe, that is not captured.
-the limitations of objective measures ", ".")`.
---> -->
-<!-- ```{r plot-figure8, out.width=600, out.height=600} -->
-<!-- #| label: fig-Fig8 -->
-<!-- #| fig-cap: "The median parcel supercluster membership." -->
-<!-- knitr::include_graphics(glue::glue(here::here(), "/figures/SCLab_and_total_copp_plot.png")) -->
-<!-- ``` -->
-
-<!-- ## Profiling of who resides in (in)completely (un)caring neighbourhoods  -->
-<!-- To enhance the meaning of the superclusters beyond a spatial conceptualisation of "caring" and "complete", was _who?_ resides there.
-We implement another decision tree in which the population-weighted socio-demographic characteristics of the parcels (retrieved from the 2021 Canadian Census) are input variables and the resulting output are 'socio-domographic superclusters' of the caring/complete superclusters.
-The input variables included: % households below median household after-tax income, LICO prevalence, average number of children per household, % not in the labour force, Gini index on adjusted household after-tax income, % visible minority, % population aged 0 to 14, % single parent household, % who walk to work (relative to bike, care/truck/van, public transit and 'other'), % of owner household in core housing need (i.e., inadequate housing structure or paying higher than 30% income on housing), % of owner households with a mortgage, % of tenant households in subsidized housing, % of tenant households in core housing need, % no certification, diploma or degree (relative to population with high school, college, or university certificate, diploma or degree).
--->
-
-<!-- Of all the variables, those that proved to be significant in profiling the five superclusters were only four variables: % households below median household after-tax income (bottom income), % population aged 0 to 14, % single parent household, and LICO.
-See @fig-Fig9 that provides the splits.
-Unfortunately, the supercluster membership is not as homogeneous as the caring/complete supercluster dendrogram (@fig-Fig7).
-However, we can interpret the socio-demographic superclusters (five that result) as labels to tell a narrative of who may reside in what caring/complete supercluster.
--->
-
-<!-- ```{r plot-figure9, out.width=600, out.height=300} -->
-<!-- #| label: fig-Fig9 -->
-<!-- #| fig-cap: "The dendrogram of the supercluster demographic profiles.." -->
-<!-- knitr::include_graphics(glue::glue(here::here(), "/figures/access_sc_profiles.png")) -->
-<!-- ``` -->
-<!-- (NOTE: for reference, see the old bivariate map in the appendix for comparison) -->
-
-<!-- <!--EXPLAIN THE FOLLOWING PARAGRAPH BETTER--> -->
-<!-- In @fig-Fig9 from right-to-left, parcels in DAs with 45% of the population below the median or lower (richer) are often classified with poor accessibility: either "Somewhat caring and not complete" SC3 or "Not caring and not complete" SC5.
-Parcels within DAs with a higher proportion of single parent households (13% and above) are most often SC3: "Somewhat caring, not complete" and below are most often SC5: "Not caring, not complete".
-For parcels with higher than 45% of bottom-income (poorer), those with a high proportion of children (age 0 to 14) are also likely to be within the "Somewhat caring and not complete" SC3 - just like parcels in richer and more-single-parent DA counterparts.
--->
-
-<!-- These findings related to potential child-care access are notable: single parents provide more care and hence may be more likely to choose, if they are able, more accessible neighbourhoods than the rest of the population in the same household income grouping.
-In a Canadian and U.S.
-context, households with children may locate to catchments with 'better' schools, often more suburban with lower potential access scores overall.
-Furthermore, statistics show that in the last years, the number of single parents families in Canada has increased.
-In 2022, there were about 1.84 million single parents compared to about 1.56 million in 2010 and 1,3 million in 2001 (https://www.statista.com/statistics/443342/single-parent-families-in-canada/ & https://www.justice.gc.ca/eng/rp-pr/jr/jr12/p5a.html#ftn4).
-Furthermore, Ontario is the Canadian province where most of the lone parent families live (749,769 families, https://www.statista.com/statistics/446667/lone-parent-families-in-canada-by-province/).
-Most lone-families are headed by female-lone parent.
-(AS TO ADD IN HAMILTON SPECIFIC NUMBERS) -->
-
-<!-- Interestingly, parcels in DAs with lower than 13% children population demonstrate a better accessibility to care destination story.
-<!--PERHAPS A MISMATCH ON WHO NEEDS / WHO GETS CARE ACCESSIBLE DESTINATIONS-->  -->
-
-<!-- Parcels within DAs that constitute more than 45% population below median household income (poorer) - especially above 9.8% LICO (low income cut-off prevalence - likely to be the poorest) are often within "Very caring and very complete" SC1.
-<!-- -maybe people who can’t afford a car choose to live here?
--->  -->
-
-<!-- Parcels located in DAs with the same lower median income and child population but have _lower_ than 9.8% LICO (not the poorest) also have good accessibility, as they are most often in "Caring and complete" SC1 as well as being in most often in the "Somewhat caring and complete" SC4 relative to other parcels.
--->
-
-<!-- Our findings illustrate differences in patterns of 15-minute walk accessibility to care destinations most prominently profiled by income, low-income prevalence, children and single-parent status.
--->
-
-<!--  `r word_comment("IDEAS DISCUSSED: regarding @fig-Fig9 - 1) We know that highest accessibility & complete is downtown.
-But people there have fewer children + lowest income are downtown.
-On the horizon as rent increases,  the most precarious may be pushed out and replaced by people who can pay higher rents.
-", " ")` -->
-
-<!--   `r word_comment("IDEAS DISCUSSED: regarding @fig-Fig9 - 2) The lowest accessibility & most incomplete are suburban -- they are likely higher income.
-Interesting to note: one-parent families are more likely to be time poor -- and leaving in low access/incomplete suburbs puts more at risk.
-We know a think or 2 about single parents and jobs -- access to jobs was high where single parents were not.
-(Catherine, farber Paez paper).
-", ":")`.
--->
-
-<!--   <!--Yes: low access and more incomplete = suburban (Figure 8) AND higher income (Figure 9).
-Are higher income folks living in low access and low caring places, but don’t care because they can afford to drive to get to these places?
-This has implications for sustainability and car dependence  -->
-
-<!--   And also: poorer one parent families with children are also in these not caring/ complete places.
-They are who may feel it: as they likely are more time poor and have lower access to cars.
-This could impact their wellbeing / access to care (which is essential) --> -->
-
-<!--  `r word_comment("IDEAS DISCUSSED: regarding @fig-Fig9 - 3) Pockets of the suburbs have high access to care. The suburbs do something -- they have care amenities. ", ":")`.  -->
-
-<!-- <!--Potentially suburban have greater access to care compared to jobs.
-I mean people live there for a reason I guess!
-And they are often seen as ‘child friendly’.
-If these suburban areas have high access to care by foot, are they going there by foot?
-What can be done to get them to walk to care destinations?
-Or are these examples of suburban that have been designed to be walkable (which should be emulated)?
---> -->
-
-<!-- # Discussion -->
-
-<!-- There seems to be a spatial mismatch where dwelling availability and affordability play an important role? Are there social housing programs for lower income households? -->
-
-<!-- Even if we are on the right direction to develop human-centred urban areas by facilitating caring activities, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities, given the multiple impacts it has on everyday mobility, the unequal distribution of tasks among the population and its related socioeconomic consequences.
-Therefore, it is necessary to provide alternatives to measure progress on this important issue.
-City planning should respond to these issues to promote equity.
-On the other hand, an intersectional approach to understand how different personal identities interact in our daily day activities in the cities is necessary.
--->
-
-<!-- The process of gentrification and displacement are trends also occurring in Hamilton.
-On the horizon of rising rents, there is a risk of displacement among more vulnerable households, who may be replaced by individuals with higher incomes who can afford different transport modes, with more flexible schedules and are more likely to use ICT arrangements (including telework, e-shopping, and other online services).
-While, lower income households may be particularly vulnerable to locating in areas with worse accessibility to care alternatives, which could have negative consequences for these households and for the society as a whole.
--->
-<!-- Even if we are on the right direction to develop human-centred urban areas by facilitating caring activities, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities, given the multiple impacts it has on everyday mobility, the unequal distribution of tasks among the population and its related socioeconomic consequences.
--->
-<!-- Therefore, it is necessary to provide alternatives to measure progress on this important issue.  -->
-<!-- This work leaves many doors open for further research. For example,  -->
-<!-- •	the role that gender plays in these households.  -->
-<!-- •	Job arrangements per type of household.
-Some studies observe that working from home arrangements are more likely available for higher education attainment households (likely to have higher income), and families with children, but also living in neighbourhoods with good accessibility, -->
-<!-- •	Finally, an intersectional approach to understand how our personal identities interact in our access to essential activities and caring services. -->
-<!-- City planning should respond to these issues to promote equity.
-On the other hand, an intersectional approach to understand how different personal identities interact in our daily day activities in the cities is necessary.
--->
-
-<!-- LUCIA'S: -->
-<!-- As in other case studies, essential opportunities and services tend to concentrate on Hamilton central.
-Therefore, it is more likely to carry out daily activities, including caring ones, in this area using sustainable modes such as walking or biking.
--->
-
-<!-- However, this study highlights that there seems to be a spatial mismatch between dwelling availability, family profile and eventually (housing and mobility) affordability.
-At neighbourhood level the concentration of opportunities for all category types is in Hamilton Central, but households with children may not live there.
--->
-
-<!-- Richer families may concentrate in either neighborhoods classified with poor accessibility: either “Somewhat caring and not complete” or “Not caring and not complete.
-However, these families are likely to have different mobility alternatives available.
-The literature shows that higher income households have a lower mobility share compared to lower income households.
-Furthermore, they are more likely to be assisted by different service arrangements (i.e.
-e-shopping, service deliveries, taxis and MaaS services, working from home arrangements,,…).
--->
-
-<!-- Our results show a different story for low-income households.
-If they can afford it, these households will prefer to live in very caring and very complete neighbourhoods, which will ease their accessibility to daily activities and essential services and therefore, avoiding social exclusion.
-In this respect, one-parent families are more likely to be time poor.
-For them, living in low access/incomplete suburbs increases their chancesof  falling into social exclusion.
-Furthermore, school closures or any other type of relocation of essential services puts more pressure on their shoulders.
--->
-
-<!-- Measuring caring neighbourhoods is of utmost importance to promote equity in city planning.
-Our results are of use for social housing programs for lower income households.
-Housing is a human right which may not be disentangle from mobility needs.
-Otherwise, lower income household risk of social exclusion, especially when accessibility to essential services is limited.
--->
-
-<!-- Similarly, this tool has great potential to support city planning, by analyzing the population groups benefited or harmed by relocation, closure or opening of different types of essential services.
--->
-
-<!-- The process of gentrification and displacement are trends also occurring in Hamilton.
-On the horizon of rising rents, there is a risk of displacement among more vulnerable households, who may be replaced by individuals with higher incomes who can afford different transport modes, with more flexible schedules and are more likely to use ICT arrangements (including telework, e-shopping, and other online services).
-While, lower income households may be particularly vulnerable to locating in areas with worse accessibility to care alternatives, which could have negative consequences for these households and for the society as a whole.
--->
-
-<!-- Even if we are on the right direction to develop human-centred urban areas by facilitating caring activities, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities, given the multiple impacts it has on everyday mobility, the unequal distribution of tasks among the population and its related socioeconomic consequences.
--->
-
-<!-- ANATASIA's -->
-<!-- - harkens back to what role a planned neighbourhood has. Is it about serivice provision (the what), social homogenity (the who), local power and democracy? [] -->
-
-<!-- # Conclusion -->
-
-<!-- This study makes three types of contributions: empirical, methodological and theoretical.
+This study makes three types of contributions: empirical, methodological and theoretical.
 At the empirical level, areas of the mid-sized City of Hamilton have been typified by their degree of '15-Minute Caring Neighbourhood' potential.
@@
 These values were then clustered using a novel machine learning approach, SOM, to generate meaningful typologies for discussion and further comparision with socio-economic characteristic of the area.
-We find...
--->
+We find A+ and A 15-minute caring neighbourhoods are located within the downtown core and in certain suburban pockets of the city, while the peripheral regions provide F level caring access.
+A- and B- areas are also identified as neighbourhoods that already support a high amount of children-centric destination access, and could be improved to provide better complete care access.
+Theoretically, our work puts forth an explicitly caring 15-Minute Neighbourhod conceptualisation, bridging the Mobility of Care and the "15-Minute City" concept.
+We discuss how measuring caring neighbourhoods is a way equity can be implicated considered within city planning.
 
-<!-- Theoretically, our work puts forth an explicitly caring 15-Minute Neighbourhod conceptualisation, bridging the Mobility of Care and the "15-Minute City" concept.  -->
-
-<!-- <!--  -->
-
-<!-- Instead of prescribing an urban form design principle, such as "all local amenities should be within a 15-minute walking distance", it instead examines an empirical example to determine which areas in the city have the _potential_ to be 15-minute neighbourhood based on the existing spatial accessibility offered by the urban environment and walking transport network.
+Our work is of relevance to research and practitioners planning equitable and sustainable cities.
+Instead of prescribing an urban form design principle, such as "all local amenities should be within a 15-minute walking distance", it instead examines an empirical example to determine which areas in the city have the _potential_ to be 15-minute neighbourhood based on the existing spatial accessibility offered by the urban environment and walking transport network.
 To this end, this data-driven methodology introduces a way to identify neighbourhoods that have potential, almost have potential, and are far from containing this potential to support future context-specific qualitative work.
--->
-
-<!-- --> -->
 
@@
 
-<!-- `r word_comment("If you're interested in how the unweighted results look like for the same maps see the appendix.
-I've also calculated the cycling travel matrices as well as the 'no dependent' i.e., the same analysis but now only considering 3 care categories: health, errand, and grocery.
-For the time being though let's focus on this.", " ")` -->
+<!-- As a tool to visualise the variables that describe the superclusters, a decision tree is fit with the DEPENDENT TOTALS, ERRAND TOTALS, GROCERY TOTALS, HEALTH TOTALS, and CATEGORY DIVERSITY scores.
+Aligning with the descriptive summaries of the superclusters provided, describing the tree-diagram from left to right (roughly from the most caring & complete, to the least), the first three terminal nodes (represented by bar columns that visualise the proportion of supercluster membership at that node) are mostly A+ and A superclusters, and are defined by their high dependent-centric access (about median and above), and high diversity score.
+A+ is differentiated from A by its exceptionally high health-centric totals while A most just has excellent scores.
+Columns four, six and seven are dominated by the presence of B+ and B parcels.
+On average the B+ parcels provide access to an average level of dependent-centric destinations, so the B+ parcels with greater than 6.5 dependent-centric destinations are found in the more left-most columns, and those below this split on the right-most.
+The far right-most column is dominated by the presence of F parcels, defined by lower DEPENDENT TOTALS and LOW DIVERSITY.
+It is interesting to note the 5th column, which is defined by D parcels but also the most amount of A- parcels than any other column.
+A- and D parcels are both high in dependent-centric destinations and low in diversity (though D is much lower than A-).
+These two parcels again, as shown in @fig-Fig7, can be thought of as parcels in neighbourhoods containing a high amount of some caring destinations, but lacking others, indicating high efficiency for improvement potential.
+-->
 
-<!-- # Appendix -->
-
-<!-- ```{r plot-figure10, out.width=600, out.height=600} -->
-<!-- #| label: fig-Fig10 -->
-<!-- #| fig-cap: "A bivariate map of the cumulative opportuninities map (aggregated per category at DA) and the diversity measure.
-Amenity diverse = diviersity index above the region median.
-Caring/Complete = accessibility value above the region median.
-The Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]." -->
-<!-- knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_bivariate_entropy_copp_plot.png")) -->
+<!-- ```{r plot-figure7, out.width=600, out.height=300} -->
+<!-- #| label: fig-Fig7 -->
+<!-- #| fig-cap: "The decision tree demonstrating variables that define the superclusters." -->
+<!-- knitr::include_graphics(glue::glue(here::here(), "/figures/access_sc_profiles_usingSCVARS.png")) -->
 <!-- ``` -->

### 2024-10-04 paper/rejected-text.qmd
Updates including Lucias comments

@@
 
+Second.5.
+the discusison of self-sorting: single parents provide more care and hence may be more likely to choose, if they are able, more accessible neighbourhoods than the rest of the population in the same household income grouping.
+In a Canadian and U.S.
+context, households with children may locate to catchments with 'better' schools, often more suburban with lower potential access scores overall.
+Furthermore, statistics show that in the last years, the number of single parents families in Canada has increased.
+In 2022, there were about 1.84 million single parents compared to about 1.56 million in 2010 and 1,3 million in 2001 (https://www.statista.com/statistics/443342/single-parent-families-in-canada/ & https://www.justice.gc.ca/eng/rp-pr/jr/jr12/p5a.html#ftn4).
+Furthermore, Ontario is the Canadian province where most of the lone parent families live (749,769 families, https://www.statista.com/statistics/446667/lone-parent-families-in-canada-by-province/).
+Most lone-families are headed by female-lone parent.
+(AS TO ADD IN HAMILTON SPECIFIC NUMBERS)
+
 In car dependent cities, cycling commute rates tend to be low.
@@
 
+Even if we are on the right direction to develop human-centred urban areas by facilitating caring activities, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities, given the multiple impacts it has on everyday mobility, the unequal distribution of tasks among the population and its related socioeconomic consequences.
+
+The process of gentrification and displacement are trends also occurring in Hamilton.
+On the horizon of rising rents, there is a risk of displacement among more vulnerable households, who may be replaced by individuals with higher incomes who can afford different transport modes, with more flexible schedules and are more likely to use ICT arrangements (including telework, e-shopping, and other online services).
+While, lower income households may be particularly vulnerable to locating in areas with worse accessibility to care alternatives, which could have negative consequences for these households and for the society as a whole.
+
+Even if we are on the right direction to develop human-centred urban areas by facilitating caring activities, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities, given the multiple impacts it has on everyday mobility, the unequal distribution of tasks among the population and its related socioeconomic consequences.
+Therefore, it is necessary to provide alternatives to measure progress on this important issue.
+City planning should respond to these issues to promote equity.
+On the other hand, an intersectional approach to understand how different personal identities interact in our daily day activities in the cities is necessary.
+
+Even if we are on the right direction to develop human-centred urban areas by facilitating caring activities, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities, given the multiple impacts it has on everyday mobility, the unequal distribution of tasks among the population and its related socioeconomic consequences.
+Therefore, it is necessary to provide alternatives to measure progress on this important issue.
+This work leaves many doors open for further research. For example,
+•	the role that gender plays in these households.
+•	Job arrangements per type of household.
+Some studies observe that working from home arrangements are more likely available for higher education attainment households (likely to have higher income), and families with children, but also living in neighbourhoods with good accessibility,
+•	Finally, an intersectional approach to understand how our personal identities interact in our access to essential activities and caring services.
+City planning should respond to these issues to promote equity.
+On the other hand, an intersectional approach to understand how different personal identities interact in our daily day activities in the cities is necessary.
+
+As in other case studies, essential opportunities and services tend to concentrate on Hamilton central.
+Therefore, it is more likely to carry out daily activities, including caring ones, in this area using sustainable modes such as walking or biking.
+
+However, this study highlights that there seems to be a spatial mismatch between dwelling availability, family profile and eventually (housing and mobility) affordability.
+At the neighbourhood level, the concentration of opportunities for all category types is in Hamilton Central, but households with children may not live there.
+
+Richer families may concentrate in either neighborhoods classified with poor accessibility: either “Somewhat caring and not complete” or “Not caring and not complete.
+However, these families are likely to have different mobility alternatives available.
+The literature shows that higher income households have a lower mobility share compared to lower income households.
+Furthermore, they are more likely to be assisted by different service arrangements (i.e.
+e-shopping, service deliveries, taxis and MaaS services, working from home arrangements,,…).
+
+Our results show a different story for low-income households.
+If they can afford it, these households will prefer to live in very caring and very complete neighbourhoods, which will ease their accessibility to daily activities and essential services and therefore, avoiding social exclusion.
+In this respect, one-parent families are more likely to be time poor.
+For them, living in low access/incomplete suburbs increases their chancesof  falling into social exclusion.
+Furthermore, school closures or any other type of relocation of essential services puts more pressure on their shoulders.
+
 ## about employment.

### 2024-10-04 paper/Manuscript.qmd
Proofreading with Anastasia

@@
 library(sf)
+library(tidyr)
 library(tmap)

### 2024-10-08 data/04-bycategoriy-clustering.qmd
Draft for full team review

@@
 classdemo.som <- rpart(data = df,
-                      formula = scluster ~ (One_parent_families + Avg_numb_of_kids + # Age_0to14+
+                      formula = scluster ~ (One_parent_families + Avg_numb_of_kids + Age_0to14+
                                               Not_in_labforce+Vis_minority+Unemployed +
@@
                                               Hld_sheltercosts+ Tenant_sheltercosts+
-                                              LICO+Low_or_no_deg+Income_gini+#Income+
+                                              LICO+Low_or_no_deg+Income_gini+ Income+
                                               Income_bottom+

### 2024-10-08 paper/Manuscript.qmd
Draft for full team review

--- 2024-10-04 paper/Manuscript.qmd
@@
 library(gtExtras)
+library(tidyr)
 ```
@@
 
+```{r population in each bar column for the Decision Tree, eval=FALSE}
+df_DA |> filter(`v_CA21_907: Median after-tax income of household in 2020 ($)` <= 68750) |>  select("Population") |> st_drop_geometry()  |> sum(na.rm=T) / df_DA |>  select("Population") |> st_drop_geometry()  |> sum(na.rm=T)
+
+df_DA |> filter(`v_CA21_907: Median after-tax income of household in 2020 ($)` > 68750 & `v_CA21_907: Median after-tax income of household in 2020 ($)` <  91500) |>  select("Population") |> st_drop_geometry()  |> sum(na.rm=T) / df_DA |>  select("Population") |> st_drop_geometry()  |> sum(na.rm=T)
+
+df_DA |> filter(`v_CA21_907: Median after-tax income of household in 2020 ($)` >=  91500) |>  select("Population") |> st_drop_geometry()  |> sum(na.rm=T) / df_DA |>  select("Population") |> st_drop_geometry()  |> sum(na.rm=T)
+```
+
 # Abstract
@@
 
-`r word_comment("I am only vaguely familar with this literature as of these months - how does this sound?", " ")`
 In addition, classification algorithms have been useful tools in the identification of mobility and spatial typologies in relation to transportation planning.
@@
 
+`r word_comment("After reading the results and discussion, do you think fig 2 plots are helpful? Maybe they should be removed?", " ")`
 ```{r plot-figure2, out.width=600, out.height=600}
@@
 The spatial distribution of destination type are visualised in @fig-Fig3 by their category.
+
+`r word_comment("Any ideas on how to condense this table...? Maybe we can cite the forthcoming JTG article, what do you think?", " ")`
 
@@
 Consequently, a 100 node (10 by 10 grid) SOM structure using euclidean distance and square topology is produced.
-Simply put, the SOM algorithm proceeds as follows: a 2D grid of nodes is created (in our case 100 nodes) as specified by the analyst, where each node will represent a point in the reduced-dimensional space.
+Simply put, the SOM algorithm proceeds as follows: a 2D grid of nodes is created as specified by the analyst, where each node will represent a point in the reduced-dimensional space.
 Upon initialization, each node is assigned a random weight vector of the same dimension as the input data (in our case, 15).
@@
 ## Overview of access to care destinations
-How many care destinations can be reached within a 15-minute walk is summed for each of the four care categories defined in @bl-tbl1.
-The median parcel value for each DA is visualised in @fig-Fig4.
-Furthermore, the median parcel diversity of care destinations accessible is presented in @fig-Fig5.
+How many care destinations can be reached within a 15-minute walk is summed for each of the four care categories defined in @tbl-Tbl1.
+The median parcel value for each DA is visualised in @fig-Fig4 and the median parcel diversity of care destinations accessible is presented in @fig-Fig5.
 Together, @fig-Fig4 and @fig-Fig5 represent the 15 variables that served as inputs into the SOM algorithm.
@@
 ## Identification of (in)completely (un)caring neighbourhoods
-<!--
 
-A+ is now A+
-A is now A
-A- is now A-
-
-B+ is now B+
-B is now B
-D is now B-
-
-F is now D
-
--->
-
-Notably, in @fig-Fig4, the concentration of opportunities for all category types is centered in the downtown of Hamilton, with the highest concentration being in the lower city (near the lake shore) and pockets of the high concentration further from the shoreline.
+In @fig-Fig4, the concentration of opportunities for all category types is centered in the downtown of Hamilton, with the highest concentration being in the lower city (near the lake shore) and pockets of the high concentration further from the shoreline.
 Grocery-centric destinations appear to be the most concentrated, followed by health-centric and dependent-centric caring destinations.
@@
 
-Based on the SOM classification methodology discussed and the 15 variables represented in @fig-Fig4 and @fig-Fig5, 7 superclusters are identified from the 10 by 10 grid of SOM nodes.
+Based on the SOM methodology discussed, 7 superclusters are identified from the 10 by 10 grid of SOM nodes.
 As diagnostics for the selection of the number of superclusters, the dissimilarity-index-based dendrogram with the proportion of parcels represented in each supercluster alongside the variance explained plot is visualised in @fig-Fig6.
+Labels representing grades A to D qualifying the 7 superclusters are then assigned.
+These grades as also reflected in @fig-Fig6 and represent the quantity of accessibility and diversity in accessibility per care category.
+Higher grades (A+ and A) corresponding to the highest accessibility and diversity, while lowest grades (D) representing lowest accessibility and diversity scores.
 
@@
 ```
-As seen in @fig-Fig6, the 7 superclusters are assigned labels from A to F, corresponding to their 'grade' of quantity of accessibility and diversity in accessibility per care category.
-What differentiates the dissimilarities in the clusters the most, i.e., nodes of grade A/A+ to everything else, is their high levels of care accessibility and often high diversity in care categories calculated for each parcel.
-A/A+ grades represent only 24% of the parcels, with the other grades being assigned lower accessibility and diversity of accessibility scores.
-To characterise the superclusters, a table providing summary statistics of each supercluster is detailed in @tbl-tbl2 and boxplots in @fig-Fig7.
+
+The grade for each supercluster was carefully assigned by the authors by referring to descriptive statistics and their interesting trends.
+These statistics are summarised in @tbl-Tbl2 and in the boxplots in @fig-Fig7.
+
+- A+ and A have exceptionally high caring accessibility for all destination types and high diversity scores in the top quantile or above. Together, these grades represent 24% of all parcels.
+- A- has high caring accessibility, but low diversity.
+Interestingly, A- is like A but with much higher dependent-centric destinations, particularly parks, schools, and daycares, with only moderately high scores for all  other destination types.
+This disbalance lowers its overall diversity score.
+However, it can be characterized as a supercluster that demonstrates potential in being retrofitted to provide A+ level completely caring access.
+This grade  represent 4% of parcels.
+- B+ and B superclusters present about average completely caring access.
+These superclusters can serve as the benchmark for what 'average' completely caring access in Hamilton currently looks like and they represent 15% of parcels.
+- B- provides above average dependent-centric destination access, particularly parks, daycares and schools, but below average access to other destination types and hence has low diversity scores.
+B- is like A- as it demonstrates complete caring 15-minute potential if retrofitted.
+As these parcels demonstrate caring access to some destinations, they may have the potential to be retrofitted to support complete access to all caring destination types.
+This grade represents 19% of parcels.
+- D superclusters demonstrate the lowest scores all-around, representing room for land-use improvement that addressed complete and caring 15-minute access.
+This supercluster characterizes the most number of parcels, representing 38% of parcels in the city.
 
 ```{r creating-table2-pt2}
-#| tbl-cap: "Mean and (standard deviation) of each SOM classified cluster by input variable and additional summary variables (indicated by ALL CAPITAL LETERS)."
+#| tbl-cap: "Mean and (standard deviation) of each SOM classified cluster by input variable and additional summary variables.
+Variables included in the SOM algorthim are in regular case, while additional summary variables are indicated by ALL CAPITAL LETERS."
 #| label: tbl-Tbl2
 #|
-summary_df %>%
-  gt() %>%
-  tab_header(
-    title = " ",
-  ) %>%
-  cols_label(
-    name = " "
-  ) %>%
-  gt_highlight_cols(
-    2:3, fill = "green", alpha = 0.3
-    ) %>%
-  gt_highlight_cols(
-    4, fill = "green", alpha = 0.2
-    ) %>%
-  gt_highlight_cols(
-    5:6, fill = "orange", alpha = 0.3
-    ) %>%
-    gt_highlight_cols(
-    7, fill = "orange", alpha = 0.2
-    ) %>%
-  gt_highlight_cols(
-    8, fill = "red", alpha = 0.3
-    ) %>%
-  tab_style(
-    style = cell_borders(
-      sides = c("top"),
-      color = "black",
-      weight = px(1.5),
-      style = "solid"
-    ),
-    locations = cells_body(rows = c(1,4,11,15,19))
-  ) |>
-  tab_options(
-    table.font.size = 12,
-    table.width = pct(80)
-  )
+# summary_df %>%
+#   gt() %>%
+#   tab_header(
+#     title = " ",
+#   ) %>%
+#   cols_label(
+#     name = " "
+#   ) %>%
+#   gt_highlight_cols(
+#     2:3, fill = "green", alpha = 0.3
+#     ) %>%
+#   gt_highlight_cols(
+#     4, fill = "green", alpha = 0.2
+#     ) %>%
+#   gt_highlight_cols(
+#     5:6, fill = "orange", alpha = 0.3
+#     ) %>%
+#     gt_highlight_cols(
+#     7, fill = "orange", alpha = 0.2
+#     ) %>%
+#   gt_highlight_cols(
+#     8, fill = "red", alpha = 0.3
+#     ) %>%
+#   tab_style(
+#     style = cell_borders(
+#       sides = c("top"),
+#       color = "black",
+#       weight = px(1.5),
+#       style = "solid"
+#     ),
+#     locations = cells_body(rows = c(1,4,11,15,19))
+#   ) |>
+#   tab_options(
+#     table.font.size = 10,
+#     table.layout = 'fixed',
+#     table.align = 'center',
+#     table.width = pct(100)
+#  )
+
+set_flextable_defaults(font.size= 8,
+                       font.family = NULL,
+                       text.align = 'justify',
+                       table.layout = 'fixed')
+
+summary_df <- summary_df |> rename(" " = "name")
+
+flextable(summary_df) %>%
+ ftExtra::colformat_md() %>%
+  flextable::width(j = 1, width = 3.06, unit = "cm") %>%
+  flextable::width(j = 2:9, width = 1.53, unit = "cm") %>%
+  bg(j= 2, bg = "#48B376") |>
+  bg(j= 3, bg = "#91CF60") |>
+  bg(j= 4, bg = "#D9EF8B") |>
+  bg(j= 5, bg = "#FFFFBF") |>
+  bg(j= 6, bg = "#FEE08B") |>
+  bg(j= 7, bg = "#FDAE61") |> 
+  bg(j= 8, bg = "#D75C56") |>
+  hline(i = c(3,10,14,18))
+
+# RColorBrewer::brewer.pal(7, "RdYlGn")
+# "#D73027" "#FC8D59" "#FEE08B" "#FFFFBF" "#D9EF8B" "#91CF60" "#1A9850"
 ```
@@
 ```
-
-The superclusters are characterized by caring access and diversity of caring access (completeness).
-This technique provides a data-driven approach to supercluster formation which we then attribute meaning to through the derivation of the following labels.
-From the descriptive statistics of the superclusters, interesting trends emerge:
-
-- A+, A, and A- all have exceptionally high or excellent caring accessibility for all destination types; all destination type and diversity scores are top quantile or above.
-Interestingly, A- is like A but with much higher dependent centric destinations, particularly parks, schools, and daycares, with high scores for everything else.
-This disbalance lowers its overall completeness score but is a supercluster that demonstrates potential in being retrofitted to provide A+ level completely caring access.
-Together, these grades represent 28% of all parcels.
-- B+, B and B- superclusters present about average completely caring access.
-These superclusters, particuarly B, can serve as the benchmark for what 'average' completely caring access in Hamilton currently looks like.
-B+ access and diversity scores are above B, while B- has low values for everything with exception to above average number of dependent-centric destinations, particularly parks, daycares and schools.
-These three clusters represent 34% of the parcels.
-- D superclusters demonstrate the lows scores all around, presenting plenty of room for improvement in complete and caring 15-minute access. This cluster represents the 38% of all parcels.
 
@@
 
-To meaningfully visualise where the superclusters are located, we visualise the representative supercluster label in each DA in @fig-Fig8.
-This is done by grouping parcels by their DA, calculating the median supercluster label for each label, and selecting the label which is most dominate within that DA.
-For reference, the median number of parcels in each DA is `r df_DA$parcel_count |> median() |> round()`.
-Further, supercluster label membership within a DA is typically pure, with the median membership being `r df_DA$max_label_val |> median() |> scales::percent()` of a single cluster.
+To spatially demonstrate where the superclusters are located, representative supercluster labels for each DA in visualised in @fig-Fig8.
+This visualisation is created by grouping parcels by their DA, calculating the median supercluster label for each label, and selecting the label that is most dominant within that DA.
+For reference, the median number of parcels in each DA is `r df_DA$parcel_count |> median() |> round()` and supercluster label membership within a DA is typically pure, with the median membership being `r df_DA$max_label_val |> median() |> scales::percent()` of a single cluster.
 
@@
 ```
-The resulting superclusters visualised in @fig-Fig8 are a combination of the cumulative opportunity and diversity of access measures trasnlated into meaningful typologies through the SOM methodology.
-It can be observed that the excellent and exceptional neighbourhoods, which are very caring and very complete 15-minute neighbourhoods, are located within the center of the city and most closer to the shoreline in the 'lower' city (below the escarpment).
-The average and below average neighbourhoods are often proximate to the excellent and exceptional neighbourhoods and are more prevalent in the 'mountain' part of the center of Hamilton.
+The resulting superclusters visualised in @fig-Fig8 are a combination of the cumulative opportunity and diversity of access measures translated into meaningful typologies through the SOM methodology.
+Accessibility is the measure of _potential_ interaction, e.g., how many care destinations one could reach within a 15-minute walk, these typologies are useful in identifying which areas of the city are providing regionally high levels of complete caring access, which areas are regionally average, and which areas are regionally below average.
+Furthermore, neighbourhoods which provide disbalanced access can be further investigated.
+
+In @fig-Fig8, it can be observed that the excellent and exceptional neighbourhoods (A, A+), which are very caring and very complete 15-minute neighbourhoods, are located within the center of the city and closest to the shoreline in the Lower City.
+The above and below average neighbourhoods (B+ B) are often proximate to the excellent and exceptional neighbourhoods and are more prevalent in the Mountain part of the center of Hamilton.
 The escarpment is a physical barrier, with few pedestrian-accessible access points to traverse; hence the typologies describing neighbouring DAs separated by the escarpment are often different.
+Below average (D) neighbourhoods are located in peripheral areas outside the center of Hamilton, in areas where urban form is characterised by lower density residential housing and auto-dependent mobility.
 
-As accessibility is the measure of _potential_ interaction, e.g., how many care destinations one could reach within a 15-minute walk, these typologies are useful in identifying which areas of the city are providing regionally high levels of complete caring access, which areas are regionally average (oranges), and which areas are regionally below average.
-Furthermore, neighbourhoods which provide disbalanced access can be further investigated.
-For instance, neighbourhoods that provide A- provide otherwise high levels of 15-minute walking access to all types of destinations, but particularly to Community and recreational centres, daycares, parks, and schools.
-These A- neighbourhoods are suitable for all, but may be particularly desirable for households that have children, needed proximity to care destinations.
-Furthermore, B- neighbourhoods are like A- neighbourhoods as they provide high access to children-centric destinations like schools, parks, and daycares, however are low in all other destination types.
-These neighbourhoods can be understood as currently being conducive to completing children-centric care trips, and having the potential to become complete if additional destination types were located within their neighbourhoods.
+Furthermore in @fig-Fig8, A- and B- grades stand out as offering high access for children-centric destinations for their grade-group but below group-average access in other destination types.
+These neighbourhoods may be more suitable for populations who prioritize walkable access to destinations like schools, parks and daycares, and find access to other types of caring destinations less important.
+These neighbourhoods also stand out as demonstrating potential to be retrofitted to provide more complete access if additional destination types were located within their neighbourhoods.
 
@@
 
-To enhance the meaning of the superclusters beyond a descriptive and spatial conceptualisation of "caring" and "complete",  _who?_ resides in what neighbourhood currently can also be investigated.
-We run a decision tree and include the population-weighted socio-demographic characteristics of the parcels (retrieved from the 2021 Canadian Census) as input variables and the resulting output are 'socio-demographic superclusters' of the superclusters.
-The input variables included: median household income, % below the median household income, % LICO prevalence, average number of children per household, % population aged 0 to 14, % not in the labour force, % not employed, Gini index on adjusted household after-tax income, % visible minority, % single parent household, % who walk to work (relative to bike, care/truck/van, public transit and 'other'), % of owner household in core housing need (i.e., inadequate housing structure or paying higher than 30% income on housing), % of tenant households in subsidized housing, % of tenant households in core housing need, % no certification or with only a highschool diploma.
+To enhance the meaning of the superclusters beyond a descriptive and spatial conceptualisation of "caring" and "complete",  _who?_ resides in what neighbourhood is investigated through the Decision Tree results.
+The input features of the Decision Tree are the supercluster labels and the feature variables are various population-weighted socio-demographic characteristics of the 2021 Canadian Census, namely: median household income, % below the median household income, % LICO prevalence, average number of children per household, % population aged 0 to 14, % not in the labour force, % not employed, Gini index on adjusted household after-tax income, % visible minority, % single parent household, % who walk to work (relative to bike, care/truck/van, public transit and 'other'), % of owner household in core housing need (i.e., inadequate housing structure or paying higher than 30% income on housing), % of tenant households in subsidized housing, % of tenant households in core housing need, % no certification or with only a highschool diploma.
 
-Of all the included input variables, the one that proves most meaningful in partitioning the superclusters data into subsets was only one variable: median household income of the DA.
-@fig-Fig9 provides the splits that were significant in partitioning the data in the decision tree algorithim.
-While the algorithim is unable to homogenously use each supercluster, we still use it to report a narrative of who may reside in what caring/complete supercluster.
-Particularly, the decision tree demonstrates a more pure supercluster membership for only two of the three terminal decision tree nodes: supercluster A+ (exceptional completely caring access) and D (not caring and complete access).
+Of all the included input variables, _median household income_ proved to be the most meaningful in partitioning the superclusters data.
+@fig-Fig9 provides the Decision Tree with the significant splits in median household income and the proportion composition of supercluster along each branch for three terminal Decision Tree nodes.
+While the algorithm is unable to homogenously use each supercluster, @fig-Fig9 is helpful to report a narrative of who may reside in what caring/complete supercluster.
+Particularly, the Decision Tree demonstrates a more pure supercluster membership for only two of the three terminal decision tree nodes: supercluster A+ (exceptional completely caring access) and D (not caring and complete access).
 For reference, the city-wide mean household income is `r df_DA_nogeom$Income |> mean(na.rm=T)|> round() |> prettyNum(big.mark = ",") ` (SD: `r df_DA_nogeom$Income |> sd(na.rm=T)|> round() |> prettyNum(big.mark = ",")` and median: `r quantile(df_DA_nogeom$Income,0.5, na.rm=T) |> unname()  |> prettyNum(big.mark = ",")`).
+The notable splits for income are $\gt$ \$91,500 $\lt$ \$68,750, or between \$68,750 and \$91,500, roughly split by lower income, middle income, and higher income median household brackets.
 
@@
 #| label: fig-Fig9
-#| fig-cap: "The decision tree demonstrating socio-ecnomic profiles of superclusters."
+#| fig-cap: "The decision tree demonstrating median household income splits and superclusters composition within each branch."
 knitr::include_graphics(glue::glue(here::here(), "/figures/access_sc_profiles_justincome.png"))
 ```
-Describing @fig-Fig9, the most valuable variable for splitting the data was the median household income in each DA.
-The terminal nodes represented as bar columns depict the percentage composition by supercluster.
-For the column on the right, the majority of the parcels are D superclusters with the second most supercluster membership being B- parcels, described by having greater than 91,500 CAD household income, above the city-wide mean.
-The left column is dominated by A+ along with A and A- (in the largest quantities relative to other bars), though proportions of all other superclusters are also present.
-This column is defined by parcels present in DAs with lower incomes, e.g., 68,750 CAD.
-The middle columns's classification is mixed: with all types of superclusters present in DAs with about average households incomes (between 68,750 CAD and 91,500 CAD).
+Describing @fig-Fig9, the bar column on the right represents DAs with higher household income.
+In these DAs, the majority of parcels are labeled as D superclusters, i.e not caring or complete 15-minute neighbourhoods.
+Within this higher-income-representing bar column, the second highest proportion of supercluster membership are B- parcels.
+B- parcels are low in most amenity types except a few child-centric destinations and show promise in being more easily improved than the D parcels.
+The higher income column represents 39% of the population, the largest proportion of any of the three columns.
+The left column in @fig-Fig9 correpnds to the lowest income, i.e., below 68,750 CAD median household income, and accounts for 32% the population.
+It is dominated by A+ parcels along with A and A- parcels, in the largest quantities relative to other bar columns, however, proportions of all other superclusters are also present.
+The middle column is defined by parcels with a mix of supercluster classifications and represents DAs with middle households incomes, between 68,750 CAD and 91,500 CAD representing 28% of the population.
 
-Though median income was the most useful in partitioning the parcels by their supercluster labels, other variables are still important but were not as helpful in partitioning the data.
-Namely, the following variables were also important, in order of importance and correlation coefficient with the median household income variable in brackets: the % below the median household income (`r Income_inspect[c("Income_bottom"),] |> round(digits=2)`), % single parent household (`r Income_inspect[c("One_parent_families"),] |> round(digits=2)`), % no certification or with only a highschool diploma (`r Income_inspect[c("Low_or_no_deg"),] |> round(digits=2)`), the average number of children per household (`r Income_inspect[c("Avg_numb_of_kids"),] |> round(digits=2)`), % not in the labour force (`r Income_inspect[c("Not_in_labforce"),] |> round(digits=2)`), % LICO prevalence (`r Income_inspect[c("LICO"),] |> round(digits=2)`), and  % who walk to work (`r Income_inspect[c("walk_towork"),] |> round(digits=2)`).
-Of note, when median household income is removed from the decision tree, a similar structure is created, except with above average Single-Parent family and LICO prevalence splitting the data.
-It is notable that median household income is highly or moderately correlated with many of these important variables, and discussing these correlation is useful for interpretation alongside the tree diagram.
+Though median income was the most useful in partitioning the parcels by their supercluster labels in @fig-Fig9, other variables are still important.
+Namely, the following variables were also important (in order of importance and correlation coefficient with the median household income variable in brackets): the % below the median household income (`r Income_inspect[c("Income_bottom"),] |> round(digits=2)`), % single parent household (`r Income_inspect[c("One_parent_families"),] |> round(digits=2)`), % no certification or with only a highschool diploma (`r Income_inspect[c("Low_or_no_deg"),] |> round(digits=2)`), the average number of children per household (`r Income_inspect[c("Avg_numb_of_kids"),] |> round(digits=2)`), % not in the labour force (`r Income_inspect[c("Not_in_labforce"),] |> round(digits=2)`), % LICO prevalence (`r Income_inspect[c("LICO"),] |> round(digits=2)`), and % who walk to work (`r Income_inspect[c("walk_towork"),] |> round(digits=2)`).
+<!-- Of note, when median household income is removed from the decision tree, a similar structure is created, except with above average Single-Parent family and LICO prevalence splitting the data.-->
 
+It is notable that median household income is highly or moderately correlated with many of these important variables, and discussing these correlation is useful for interpretation alongside the Decision Tree diagram.
 For instance, single-parent families are most positively correlated with the proportion of households being in the bottom income distribution (`r Income_inspect[c("Income_bottom"),] |> round(digits=2)`) and low or new degree (`r One_parent_family_inspect[c("Low_or_no"),] |> round(digits=2)`), and most negatively with the median after-tax household income (`r One_parent_family_inspect[c("v_CA21_907: Median after-tax income of household in 2020 ($)"),] |> round(digits=2)`), reflecting nationwide trends [@governmentofcanadaDailyMainHighlights2024; @governmentofcanadaPrevalenceLowIncome2024].
-LICO is also an indicator of low-income, and though not all single-family households are below the LICO (in fact, single-family households are not highly correlated with LICO at the DA level in Hamilton), DAs where higher concentrations of single-family and where high LICO prevalence persist tend to have A+ access.
-These findings are notable as currently, the highest grade (A+) serves a disadvantaged group that is also often disadvantaged along different intersecting socio-economic characteristics.
-Also interestingly, DAs with lower single-parent households (who also tend to have higher incomes) are in DAs with a higher concentration of F parcels.
-These considerations are important when considering the equity of caring 15-minute policy interventions.
-
-<!-- Describing @fig-Fig9, the most valuable variable for splitting the data was the proportion of single family households and % of LICO prevalence in the DA.
-The the terminal nodes represented as bar columns depict the percentage composition by supercluster.
-For the bar on the right, the majority of the parcels are F superclusters.
-Following the tree diagram, these F superclusters are described by having less than 15.8% single parent families (this is below the median of `r df_DA$One_parent_families |> median(na.rm=T) |> scales::percent()`).
-The bar on the left is mostly composed of parcels that are labeled A+.
-These A+ parcels are located in DAs greater than 15.8% single parent families and % LICO prevalence of 6.5% and greater (this is above the median of `r df_DA$LICO |> median(na.rm=T) |> round() |> paste0("%")`).
-In sum: residences in parcels with exceptionally complete and caring access tend to be lower-income and single-parent families, while residences in parcels with not caring are not complete access tend to have a lower presence of single parent households.
-Households with a higher proportion of single-family households (15.8% and greater) where LICO is lower have a mixed label membership (bar column in @fig-Fig9).
--->
+Though not all single-family households are below the LICO (single-family households are not highly correlated with LICO at the DA-level in Hamilton), DAs with higher concentrations of single-family households and LICO prevalence tend to be A+ completely caring 15-minute neighbourhood.
+Conversely, DAs with lower single-parent households are in DAs with a higher concentration of F parcels.
+These findings are notable from an equity perspective and in considering caring 15-minute policy interventions, especially since economic disadvantaged often intersects with other socio-demographic characteristics.
 
@@
 
-The descriptive statistics associated with each supercluster, their spatial distribution, and the socio-economic profiles that characterise them, are revealing points of discussions and further examined.
+`r word_comment("Ideas on this section (idea for an extra paragraph or additional points to each paragraph to make them stronger, with citations) would greatly be appreciated!", " ")`
 
-Hamilton is a city is not a caring 15-Minute City for all, but it is for some.
+Geographically, Hamilton is city that offers very completely caring 15-Minute neighbourhoods for some, but not everyone.
 There are evident spatial inequalities in areas that are, are not, and are somewhere between a caring and complete 15-minute neighbourhood (from labels A+ and A being excellent, B being about average, D being well below average, and A- and B- demonstrating potential in their grade group).
 The downtown core of the city has the most _caring 15-minute neighbourhoods_, but many pockets of the suburbs do as well.
-Though this contrasts what may be seen when considering employment opportunities [as in the work of @elgeneidyNonstopEquityAssessing2016, the paper's findings align with the neighbourhood unit planning literature reviewed.
+Though this contrasts what may be seen when considering employment opportunities [as in the work of @elgeneidyNonstopEquityAssessing2016], the paper's findings align with the neighbourhood unit planning literature reviewed.
 As many North American Suburbs, Hamilton followed certain characteristic design patterns that were initially designed to be proximate to amenities, coinciding with the conceptualisation of _caring 15-minute cities_ [@aliNeighbourhoodPlanningHamilton1991].
 
-Who in currently resides in Hamilton's caring 15-Minute neighbhorhoods is also currently an optimistic story.
+Who currently resides in Hamilton's completely caring 15-Minute neighbhorhoods is also an optimistic story.
 Parcels that provide A+ completely caring access tend to be in DAs that are economically disadvantaged.
@@
 Furthermore, lower-income households tend to also be single-parent households.
-Broadly, single-parent households are also more likely to be time-disadvantaged [@nieuwenhuisSingleparentFamiliesWork2018], and tasked with a higher proportion of care duties [@craigTimeCareComparison2004].
-In this way, the most economically disadvantaged groups having A+ 15-minute care accessibility is an optimistic finding.
+Broadly, single-parent households are more likely to be time-disadvantaged [@nieuwenhuisSingleparentFamiliesWork2018], and tasked with a higher proportion of care duties [@craigTimeCareComparison2004].
+In this way, the most economically disadvantaged groups having A+ complete and caring access is an optimistic finding.
 However, Hamilton is experiencing gentrification [@ellisyoungWeReJust2018]; rents are rising along the future light rail transit corridor and throughout the city [@vandermerweSpilloverGentrificationMidsized2021; @mayersLightTransitDocumenting2023].
-The larger and high-rent city of of Toronto is spilling gentrification into Hamilton's downtown core, (re)producing neighbourhoods based on Toronto's middle class [@mayersLightTransitDocumenting2023].
-In this way, the lower income residents of Hamilton's A+ neighbourhoods are more likely to be displayed and forced to re-located; and if they want to stay in Hamilton, there are not many areas that provide the same exceptional level of access as those they are currently located in.
-This is a matter of wellbeing and justice.
+Toronto, Hamilton's larger and higher-rent neighbouring city, is spilling gentrification into Hamilton's downtown core, (re)producing neighbourhoods based on Toronto's middle class identities [@mayersLightTransitDocumenting2023].
+In these ways, the lower income residents of Hamilton's A+ neighbourhoods are more likely to be displaced, which is matter of wellbeing and justice.
+There are few low-rent choices that provide the same exceptional level of access as the downtown core, hence lower household income residents that currently reside in A+ neighhourhoods would likely need to relocate to a different city.
 
-But what is the most equitable way forward in working to increase caring 15-Minute neighbourhoods?
-The work's findings indicate that parcels that provide very low 15-minute completely caring access (F) are households that tend to be higher income.
-These areas tend to be more rural, single-use zoned and car-dependent.
-Further, of the parcels that provide high child-centrics but low otherwise, A- parcels (better access) tend to be in DAs with lower-income households more so than B- parcels (lower acess than A-).
+`r word_comment("How does this paragraph sound..? I like it but I feel like it doesn't have a solid point to make... yet.", " ")`
+In discussing policy interventions that equitably increase completely caring 15-Minute neighbourhoods in Hamilton, this work's presents a methodology to create city-wide relative typologies and investigate who currently resides in what neighbourhood, as a stepping stone for further investigation.
+Neighbourhoods with the lowest grades (D) and with the highest potential in being improved (i.e., high accessibiltiy for certain types of destinations but not all) are neighbourhoods with B- and A- grades.
+However, the work's findings demonstrate a mismatch in household income and completely caring access.
+Is it equitable to focus policy on ameliorating neighbourhoods that are already higher-rent though they tend to be more rural, single-use zoned and car-dependent (parcels with D grades)?
+Further, of the parcels that provide high child-centrics but low otherwise, A- parcels (better access) tend to be in DAs with lower-income households more so than B- parcels (lower access than A-).
 From the perspective of ameliorating land-use to support the equitable distribution of 15-minute caring neighbourhoods which areas be targeted?
 Who should benefit, and how?
-If the initiative is targeted to specific neighbourhoods: sustainability linked to car dependency and equity are in-tension.
+If the policy initiative is targeted to specific neighbourhoods: sustainability linked to car dependency and equity are in-tension.
 This harkens to what role a planned neighbourhood, and bottom-up planning approaches that should be adopted to achieve equitable 15-minute cities.
-In this way, the methodology presented shows potential to support city planning, by demonstrating where neighbourhods that provide walkable access to opportunities that matter to all (care destinations) are located as city-wide relative typologies and who currently resides there, as a stepping stone for further investigation.
 
-As all research work, the results should also be interpreted along with methodological assumptions, namely, spatial accessibility yields a measure of potential interaction with all reachable destinations from an origin.
+As all research work, the results should also be interpreted along with methodological assumptions.
+This work measures spatial accessibility which is a measure of potential interaction with all reachable destinations from an origin.
 These destinations however may not be relevant to people at an origin, e.g., the may be underutilized (walkable access to 2 schools is unneeded, as a child only needs to attend 1 school), the trip physically undesirable (the walk may be along an arterial with high traffic speeds, making the trip unlikely to ever happen by foot), or the average walking speed assumed may not reflect the walking speeds of all populations [@willberg15minuteCityAll2023].
-All-in-all, people who reside in these neighbourhoods may also disagree with the neighbourhood's label, the labels are region-relative (e.g., high accessibility in Hamilton may be subjectively insufficient for some) and they do not consider subjective perceptions that influence accessibility (e.g., though a neighbourhood has many opportunities, it may not feel safe to access them).
+Further, people who reside in these neighbourhoods may also disagree with the neighbourhood's label.
+The grade labels are region-relative (e.g., high accessibility in Hamilton may be subjectively insufficient for some) and they do not consider subjective perceptions that influence accessibility (e.g., though a neighbourhood has many opportunities, it may not feel safe to access them).
 Furthermore, accessibility is calculated from the point of residential parcels.
@@
 In this way, the results flatten the dynamic patterns of care trips.
+These methodological assumptions should all be taken together when interpreting the results.
+In this way, the methodologyand findings presented identifies spatial and socio-economic trends that should be further investigated.
 
@@
 
-This study makes three types of contributions: empirical, methodological and theoretical.
+`r word_comment("Thoughts on this section? Strong ending? anything more to add?", " ")`
+This study makes three types of contributions to the transportation and city planning literature: empirical, methodological and theoretical.
 At the empirical level, areas of the mid-sized City of Hamilton have been typified by their degree of '15-Minute Caring Neighbourhood' potential.
 Methodologically, we applied the longstanding accessibility measure of cumulative opportunities and entropy to classify spatial areas based on how many destinations could be reached in a 15-minute walk and the diversity of destination type.
-These values were then clustered using a novel machine learning approach, SOM, to generate meaningful typologies for discussion and further comparision with socio-economic characteristic of the area.
-We find A+ and A 15-minute caring neighbourhoods are located within the downtown core and in certain suburban pockets of the city, while the peripheral regions provide F level caring access.
+These values were then clustered using a novel machine learning approach, SOM, to generate meaningful typologies for discussion and further comparison with socio-economic composition of the area.
+We find A+ and A 15-minute caring neighbourhoods are located within the downtown core and in certain suburban pockets of the city, while the peripheral regions provide D level caring access.
 A- and B- areas are also identified as neighbourhoods that already support a high amount of children-centric destination access, and could be improved to provide better complete care access.
-Theoretically, our work puts forth an explicitly caring 15-Minute Neighbourhod conceptualisation, bridging the Mobility of Care and the "15-Minute City" concept.
-We discuss how measuring caring neighbourhoods is a way equity can be implicated considered within city planning.
+Theoretically, our work puts forth an explicitly caring 15-Minute Neighbourhod conceptualisation, bridging the Mobility of Care and the 15-Minute City concepts.
+We discuss how measuring caring neighbourhoods can be explicitly considered within city planning.
 

### 2024-10-08 paper/rejected-text.qmd
Draft for full team review

@@
 ```
+
+<!-- Describing @fig-Fig9, the most valuable variable for splitting the data was the proportion of single family households and % of LICO prevalence in the DA.
+The the terminal nodes represented as bar columns depict the percentage composition by supercluster.
+For the bar on the right, the majority of the parcels are F superclusters.
+Following the tree diagram, these F superclusters are described by having less than 15.8% single parent families (this is below the median of `r df_DA$One_parent_families |> median(na.rm=T) |> scales::percent()`).
+The bar on the left is mostly composed of parcels that are labeled A+.
+These A+ parcels are located in DAs greater than 15.8% single parent families and % LICO prevalence of 6.5% and greater (this is above the median of `r df_DA$LICO |> median(na.rm=T) |> round() |> paste0("%")`).
+In sum: residences in parcels with exceptionally complete and caring access tend to be lower-income and single-parent families, while residences in parcels with not caring are not complete access tend to have a lower presence of single parent households.
+Households with a higher proportion of single-family households (15.8% and greater) where LICO is lower have a mixed label membership (bar column in @fig-Fig9).
+-->

### 2024-10-10 editorial correspondance/Highlights.qmd
Draft to submit to journal. And some repo clean-up.

@@
 
-- 15-minute neighbourhoods (FMN) have (re)-emerged as a popular concept in urban-planning city-planning; emphasizing the access to urban amenities by foot or bicycle.
-However, historically, neighbourhood planning processes have been top-down; FMC is linked to the Neighbourhood Unit Concept of the 1920s, largely critized for supporting segregated neighbourhoods, leaving a void in neighbourhood planning concepts that were filled by auto-centric suburban sprawl.
-Regardless, fundamental concepts within FMC remain popular: people identify with neighbourhoods, see cities through that lense.
-These ideas have in part re-emerged in the popularity of 15-minute neighbourhoods.
+- The 15-Minute City have emerged in popular urban planning and public discourse, emphasizing urban access to amenities by foot.
+However, critiques of the concept highlight how it falls short in addressing the pre-existing structural forces that drive inequalities.
 
-- But how do we uplift the positive, popular, aspects of FMN without opening it up to the critics of the NUC?
-In this work, we turn to feminist geographies and the Mobility of Care conceptualisation and apply it to FMC.
-We argue, conventional planning practice is masculinist.
-This can threaten the FMC concept by shaping how 'important' amenities are conceptualised thus favouring work- and leisure- centered amenities in cities that are increasingly capital-driven.
-Though FMN is fundamentally based in bottom-up collaborative planning processes, we exist in the context of planning pracice which is largely masculinist.
+-	The Mobility of Care framework is introduced as a way to clarify this critique of the 15-Minute City from the perspective of amenity focus.
+It argues that conventional, masculinist planning focuses on capital-driven amenities (work and leisure), neglecting essential "Care Destinations" like trips to drop-off children at daycare, grocery shopping, errands, attending medical appointment.
+These are pertinent to all people: but especially for women, and women who are lower-income.
 
-- FMN are for people - people do unpaid labour, those people are often women and intersectionally women-of-colour and low-income.
-However, this work should be for all people.
-Cities should be designed with this conceptualisation in mind.
-This is to combate the growing capitalisation of cities.
-
-- Our paper offers this nuanced understanding of FNC -- calling for them to center Care Destinations as important amenity types with the explicit conceptualisation that their destination type matters especially for women (low-income and of colour) who interact with them most, but also for all people.
-Their explicit positioning also can indirectly combat finacialisation of cities and co-opting of the FMC concept to one that centers capital-generating destination types like work and prohibitive leisure.
-Not to say those destination types should or cannot co-exist with Care destinations -- they should -- but markets are already incentivizing them.
-The market does not speak for non-economically geberating destination types, so we must explicitly call for them and support them within the FNC framework.
-- 
-- 
--
+- Through the empirical examination of Hamilton, Canada, accessibility to care destinations and the diversity of this accessibility is calculated for all residential parcels.
+Machine learning techniques are then used to classify areas of the city by their extent of _caring 15-minute neighbourhoods_ potential and create residential profiles along socio-economic dimensions.

### 2024-10-10 paper/Manuscript.qmd
Draft to submit to journal. And some repo clean-up.

@@
 ---
-title: "Identifying 15-minute neighbourhoods with complete caring potential"
+title: "Caring 15-minute neighbourhoods"
 author:
@@
 format: 
-  docx
-  # elsevier-pdf:
-  #   keep-tex: true
-  #   include-in-header:
-  #     text: |
-  #       \usepackage{graphicx}
-  #       \usepackage{unicode-math}
-  #       \usepackage{hyperref}
-  #       \def\tightlist{}
-  #       \usepackage{setspace}
-  #       \doublespacing
-  #       \usepackage{lineno}
-  #       \linenumbers
-  #   journal:
-  #     name: 7th International Conference on Women and Gender in Transportation (WGiT)
-  #     formatting: preprint
-  #     model: 3p
-  #     cite-style: authoryear
+  # docx
+  elsevier-pdf:
+    keep-tex: true
+    include-in-header:
+      text: |
+        \usepackage{graphicx}
+        \usepackage{unicode-math}
+        \usepackage{hyperref}
+        \def\tightlist{}
+        \usepackage{setspace}
+        \doublespacing
+        \usepackage{lineno}
+        \linenumbers
+    journal:
+      #name: "Transportation Research Part A: Policy and Practice"
+      formatting: preprint
+      model: 3p
+      cite-style: authoryear
 keywords: ["Accessibility", "Mobility of Care", "Gender", "15-minute neighbourhood", "Care Complete Neighbourhoods"]
@@
 summary_df$name <- c("GROCERY TOTALS", "Convenience Store", "Grocer",
-                     "DEPENDENT TOTALS", "Community or Rec. centre", "Daycare or EarlyON","LTC or retirment home","Park","School","Senior centre",
+                     "DEP. TOTALS", "Comm. or Rec. centre", "Daycare or EarlyON","LTC or retirment home","Park","School","Senior centre",
                      "HEALTH TOTALS","Dentist","Hospital or clinic","Pharmacy",
@@
 
-The "15-Minute Neighbourhood" concept has been adopted by leaders around the world as a way to promote human-scale cities.
-The concept emphasizes a transport and land-use mix that supports short-trips to daily necessities by foot or bicycle from ones' residence.
-This study adopts the 15-minute neighbourhood concept through the lens of "Mobility of Care", a framework that foregrounds travel to care destinations, travel done predominately by women, in contrast to the better-studied travel to employment and leisure.
-While the 15-minute neighbourhood concept is flexible enough to consider all destination types, gendered examinations are relatively lacking in the literature and no research to date has focused explicitly on care destinations.
-In this study, we address these gaps by identifying which areas in Hamilton, Canada have the potential to be ‘caring 15-minute neighbourhoods'.
-To do so, a database of care destinations was compiled and used to estimate both the number (cumulative opportunities indicator) and diversity of mix (entropy measure) of destinations reachable within a 15-minute walk from all residential parcels.
-Parcel-level results were classified into 'caring' (quantity of reachable opportunities) and 'complete' (diversity of care destinations) neighbourhood typologies of varying degrees using machine learning techniques (Self-Organizing Maps (SOM) and hierarchical clustering).
-Inequalities in the relationship between the identified caring typologies and social-economic indicators of who currently reside in the neighbourhoods were then explored.
-Our results suggest the majority of 'caring and complete' neighbourhoods are within the urban core where a higher proportion of lower-income households reside.
-Contrawise, neighbourhoods that are 'not caring and not complete' are located in peripheral regions that tend to have higher household incomes.
-Further, certain neighbourhoods in suburban and rural communities are 'caring and somewhat complete', showing the potential to become 'complete' through possible urban policy intervention.
-Implications of neighbourhoods that have a high potential to be improved, and socio-demographic characteristics associated with income such as single-parent households and economic poverty are discussed.
-Taken together, this study offers a theoretical bridge to connect Mobility of Care and the 15-minute neighbourhoods concept, a data-driven classification and clustering methodology to assess neighbourhoods on a 'caring and complete' continuum, and discussion of implications relevant to gender mainstreaming within transportation planning.
+The "15-Minute City" concept has been embraced by global leaders to promote human-scale neighbourhoods with transport and land-use designs that support short trips to daily necessities.
+This paper bridges the 15-Minute City to "Mobility of Care",  a framework that foregrounds travel to care destinations, travel done predominately by women.
+This focus contrasts the more commonly studied travel to employment and leisure destinations.
+While the 15-Minute City concept is flexible enough to consider all destination types, gendered examinations are relatively lacking in the literature, and little research to date has focused explicitly on care destinations.
+This gap is addressed in this paper by identifying which areas in the city of Hamilton, Canada are ‘caring 15-minute neighbourhoods’.
+To do so, a database of care destinations was compiled to estimate the number (using the cumulative opportunities accessibility measure) and diversity of mix (using the entropy measure) of care destinations within a 15-minute walk from residential parcels.
+Using data-driven machine learning techniques (Self-Organizing Maps and Decision Trees), neighborhoods were classified into 'caring 15-minute neighbourhood' typologies that are examined across residential socio-economic profiles.
+Our results suggest that the majority of caring 15-minute neighbourhoods are in the urban core, where more lower-income households reside.
+In contrast, not caring 15-minute neighbourhoods in higher-income peripheral areas.
+Areas that make good candidates for urban policy intervention are identified, and the implications of enhancing 15-minute walkable caring access are discussed in relation to gender mainstreaming in transportation planning.
 
@@
 How to equitably prescribe urban form that 'enables' contact with social opportunities (instead of 'engendering it' from the top-down) is still under debate.
-For instance, questions such as _what_ are destinations that matter, by what mode and travel time threshold, and _for whom_ remain under discussion among proximity-based planners who frequently use accessibility-based tools [@silvaProximitycentredAccessibilityConceptual2023; @silvaRegionalAccessibilityUndermining2022; @guzmanProximityEnoughCritical2024].
+Questions remain, including: what destinations really matter?, what mode and travel time threshold is most appropriate?, and who benefits?
+For instance, these dimensions are under discussion among proximity-based planners who use accessibility-based tools [@silvaProximitycentredAccessibilityConceptual2023; @silvaRegionalAccessibilityUndermining2022; @guzmanProximityEnoughCritical2024].
 
-From another perspective, feminist geographers, urban planners and other research groups agree that urban structure influences travel, but city planning fails short when planning for a neutral identity.
+From another perspective, feminist geographers, urban planners and other researchers agree that urban structure influences travel, but city planning falls short in planning for a neutral identity.
 In contrast, cities ought to be planned according to the multiple identities of the inhabitants [@vacchelli_towards_2018; @urbandevelopmentviennaGenderMainstreamingUrban2013].
 Researchers have demonstrated how mobility patterns differ according to gender [@lawWomenTransportNew1999; @cresswell_gendered_2008; @levy2013travel; @little_gender_1994; @tronto_toward_1990].
-Gender and social norms and cultural factors play a role in how, where, with whom, what for, and how far we move.
+Gendered social norms and cultural factors play a role in how, where, with whom, what for, and how far we move.
 Women have higher activities related to care [@ilo_care_2018], no matter at what life stage they are [@garciaromanGenderDifferencesTime2022].
-For all these reasons the spatial organization of activities has a different impact according to gender, the 15-Minute City concept cannot be gender blind.
+For all these reasons the spatial organization of activities has a different impact according to gender, the 15-Minute City concept should not be gender blind.
 
-To clarify questions surrounding the 15-Minute City concept of what destinations matter and for whom, this work builds a theoretical bridge to "Mobility of Care".
+This work builds a theoretical bridge between the 15-minute city and mobility of care concepts to answer questions about what destinations matter and who benefits from proximity planning.
 The Mobility of Care term was coined in @sanchezdemadariagaMobilityCareIntroducing2013; it refers to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands.
@@
 Moreover, care trips take on tremendous gendered and socio-economic importance, as women, and especially those from lower income households, complete the majority of care trips [@ravensbergen2023exploratory].
-Despite care travel's importance in advancing gender equality and equity broadly, the examination of mobility of care and access to care destinations have been under-examined relative to employment destinations.
+Despite care travel's importance in advancing gender equality -and equity more broadly-, mobility of care and access to care destinations have been under-examined relative to employment destinations.
 This is especially pertinent in accessibility research that has largely focused on employment points of interest e.g., [@farberOntarioLineSocioeconomic2019; @duarteInfluenceJobAccessibility2023; @ryanAccessibilitySpaceTime2023].
@@
 The objectives of this research are two-fold.
-First, we aim to theoretically bridge Mobility of Care with the 15-Minute City concept to to re-imagine what local amenities matter and define an associated accessibility and diversity of accessibility measure.
-Second, taking a bottom-up data-driven approach, we apply a classification and clustering machine learning methodologies to support the bottom-up identification of _15-minute caring neighbourhoods_ through an empirical examination of the mid-cited city of Hamilton, Canada.
-These neighbourhoods are then interpreted on the continuum of having (and not having) the region-relative land-use and transportation infrastructure to support _completely caring_ access.
-Who currently resides in which neighbourhood is examined, and ways to further expand the care accessibility story are discussed.
+First, we aim to theoretically bridge Mobility of Care with the 15-Minute City concept to re-imagine what local amenities matter and to define an associated accessibility and diversity of accessibility measures.
+Second, we take a data-driven approach to classify areas into 'caring 15-minute neighbourhoods' typologies and examine associated residential profiles in an empirical examination of the city of Hamilton, Canada.
 
@@
 In this context, the 15-Minute City is now in the public spotlight [@logan_xminute_2022; @moreno_introducing_2021].
-In @moreno_introducing_2021 where the 15-Minute City concept was introduced, four universal dimensions are outlined: density (in terms of people per km$^2$), diversity (including mixed-land use and diversity of people), the temporal and spatial proximity to essential services, <!-- (residential areas, work, commercial areas, education centres, health facilities and other basic institutions)--> and digitalisation.
+The 15-Minute City was introduced in @moreno_introducing_2021 along with four dimensions: density (in terms of people per km$^2$), diversity (including mixed-land use and diversity of people), the temporal and spatial proximity to essential services, and digitalisation.
 However, the framework presented has been criticized within academic and planning circles [@guzmanProximityEnoughCritical2024].
-Critiques highlight the concept falls short in addressing the  pre-existing structural forces and individual characteristics that drive inequalities, influencing who benefits or could benefit the most from such a paradigm [@dimarino15minuteCityConcept2023; @willberg15minuteCityAll2023].
+Critiques highlight how the concept falls short in addressing the pre-existing structural forces and individual characteristics that drive inequalities, influencing who benefits or could benefit the most from such a paradigm [@dimarino15minuteCityConcept2023; @willberg15minuteCityAll2023].
 A 15-Minute City for all people is an aspirational ideal, but how will existing mobility and accessibility inequalities be addressed.
-Without directed and context-specific solutions, this increasingly popular concept risks becoming a place-branding slogan [@pozoukidou15MinuteCityDecomposing2021].
+Without directed and context-specific solutions, this popular concept risks becoming an empty city branding exercise.
+[@pozoukidou15MinuteCityDecomposing2021; @gowerPlanningInnovationCity2022].
 
@@
 The progress report describes a high-level focus on equitable service delivery to all residents with equity concerning topics related to racial equity and people with disabilities [@portland_government_portland_2017], similarly taking on an "all populations" approach.
-Later, other cities also incorporated proximity-based goals into their plans, adopting similar neutral entity approaches.
-For instance, in the plan that later catapulted the 15-Minute City into public discourse: Paris mayor Anne Hidalgo proposed a 15-minute city plan in her 2020 re-election campaign that foregrounded six social functions that should be easily accessible from any location, namely: housing, work, health care, groceries, education and leisure [@ville_de_paris_paris_2022].
-These ideas inspired language in the agendas of other cities in the Western world such as Ottawa, in the Canadian context, who also adopted a 15-Minutes approach in their recent Official Plan [@ville_dottawa_quartier_2021].
-From the review of worldwide practice literature in @teixeiraClassifying15minuteCities2024, the 15-Minute City concept is in early phases of implementation around the world and the diverse range of definitions, strategies, and instruments present a significant knowledge and implementation gap: the 15-Minute City is aspirational but _how do we get there and whom will benefit_?
+Later, other cities adopted proximity-based goals using similar neutral approaches.
+For example, the 15-Minute City plan announced during the Paris mayor Anne Hidalgo’s 2020 re-election campaign, which helped popularize the concept in today's public discourse, emphasized six key social functions that should be easily accessible from any location [@ville_de_paris_paris_2022].
+These locations included: housing, work, health care, groceries, education and leisure.
+The 15-Minute City concept inspired language in the agendas of other cities in the Western world such as Ottawa, in the Canadian context, who also adopted a 15-Minutes approach in their recent Official Plan [@ville_dottawa_quartier_2021].
+@teixeiraClassifying15minuteCities2024's global review finds that the 15-Minute City concept is in early phases of implementation around the world and the diverse range of definitions, strategies, and instruments present a significant knowledge and implementation gap.
+In other words: the 15-Minute City is aspirational but how do we get there and who will benefit?
 
@@
 
-The re-defined bottom-up approach to community and neighbourhood planning has been adopted by New Urbanists, from which the 15-Minute City stems [@kissfazekasCircleParadigms15minute2022]; however, the effectiveness of these bottom-up processes remain to be seen due to the contemporary nature of the concept.
+The redefined bottom-up approach to community and neighbourhood planning has been adopted by New Urbanists, from which the 15-Minute City stems [@kissfazekasCircleParadigms15minute2022]; however, the effectiveness of these bottom-up processes remain to be seen due to the contemporary nature of the concept.
 Though, related research can be examined.
-In recent years, the question of _how can physical form be planned to enable an improved quality of life, and for whom with what outcomes_ has occupied the urban planning research agenda.
+In recent years, the question of _how can physical form be planned to enable an improved quality of life, for whom, and with what outcomes_ has occupied the urban planning research agenda.
 For instance, the examination of low-income households access to transportation and their likelihood of gaining employment [@blumenbergDriveWorkRelationship2017; @bastiaanssenDoesBetterJob2022] or the relationship between children's access to public transit and participation in after-school activities [@palmRolePublicTransit2020].
@@
 
-In examining how to enable improved quality of life through urban built environment, an increasingly popular tool have been accessibility measures.
+In examining how to enable improved quality of life through urban built environment, accessibility measures have become an increasingly popular tool.
 These measures are a way to quantify the ease of reaching destinations from a given point in space and have been used to examine urban areas through just and sustainable city planning agendas [@vale2023accessibility; @handyAccessibilityIdeaWhose2020].
-As the 15-Minute City is an amenity-provision neighbourhood planning concept which aims to enable urban environment that improves life quality, its analysis lends well to the use of accessibility tools [@guzmanProximityEnoughCritical2024].
-Foreseeably, recent works have applied accessibility measures to investigate the 15-Minute City across various geographic scopes.
+The 15-Minute City is an amenity-provision neighbourhood planning concept aimed at enabling the creation of urban environments that enhance life quality, making it well-suited for analysis using accessibility methodologies [@guzmanProximityEnoughCritical2024].
+Recent works have applied accessibility measures to investigate the 15-Minute City across different geographic scopes.
 For instance, in Naples, Italy [@gaglioneUrbanAccessibility15minute2022], Barcelona, Spain [@graellsCityCitiesMeasuring2021], Vancouver, Canada [@hosford15minuteCityReach2022], and in urban areas across Europe [@vale2023accessibility].
-The "cumulative opportunity" measure has been applied to many 15-Minute City accessibility examinations.
-This form of accessibility quantifies how many destinations can be reached from a point in space within a given travel time threshold, pairing well with examinations that have a normative travel time threshold, like the 15-Minute City.
+The "cumulative opportunity" measure has been applied in many 15-Minute City examinations.
+This measure quantifies how many destinations can be reached from a point in space within a given travel time threshold, pairing well with examinations that have a normative travel time threshold, like the 15-Minute City.
 Furthermore, the cumulative opportunity measure is widely appreciated for its intuitive computation and popularity among practitioners [@handyAccessibilityIdeaWhose2020; @handyMeasuringAccessibilityExploration1997; @chengInvestigatingWalkingAccessibility2019].
@@
 
-Another method which compliments the assessment of the 15-Minute City as it relates to the diversity of opportunity types, is entropy.
-The entropy measure has been widely used to characterise the diversity of land-use mix in early work by @frankLinkingObjectivelyMeasured2005 based on the species diversity Shannon-Wiener index that expresses relative evenness throughout a sequence [see review in @whittaker1972evolution].
-It has also been used in social sciences to comprehend the occurrence of social events in a structured manner specially with regard to life trajectories [@ritschardSequenceAnalysisRelated2018].
-In the realm of urban studies and planning literature, the malleable concept of entropy has been used to characterise the diversity in land-use mix [@ewingTravelBuiltEnvironment2010a], especially in the context of active modes [@luUrbanDensityDiversity2017; @mavoaIdentifyingAppropriateLanduse2018] and suburban sprawl [@randallGISBasedLand2015].
-As well, it has been used to understand mobility behaviour [@mcbride2020exploration; @montero2023applying; @montero2023role].
-Work that has used diversity indices alongside accessibility analysis are scarce, though examples of its use as parameters within accessibility scores in the case of employment opportunities [@chengMeasuringUrbanJob2013; @daiIncorporatingJobDiversity2018] and more recently to describe the diversity in transit facilities [@yinIncorporatingFacilityDiversity2024] are present.
-To the authors knowledge, the use of an entropy measure to reflect the diversity in care destination amenity types and in context within a 15-minute neighbourhood, has yet to be done.
-However, entropy measures theoretically compliment the 15-Minute City concept.
+Another method that compliments the assessment of the 15-Minute City regarding the diversity of opportunity types is entropy.
+The entropy measure, based on the Shannon-Wiener index of species diversity, expresses relative evenness within a sequence [@whittaker1972evolution].
+In urban studies and planning literature, the flexible concept of entropy has been used to characterize land-use mix diversity [@frankLinkingObjectivelyMeasured2005; @ewingTravelBuiltEnvironment2010a], particularly in the context of non-work trips [@cerveroTravelDemand3Ds1997], walking as a mode [@luUrbanDensityDiversity2017; @mavoaIdentifyingAppropriateLanduse2018], and suburban sprawl [@randallGISBasedLand2015] as relevant to this work.
+It has also been used to understand mobility behaviour [@mcbride2020exploration; @montero2023applying; @montero2023role].
+However, work incorporating entropy indices alongside accessibility analysis are not exceptionally common.
+There are examples of their use as parameters within accessibility scores for employment opportunities [@chengMeasuringUrbanJob2013; @daiIncorporatingJobDiversity2018] and, more recently, to describe the diversity in transit facilities [@yinIncorporatingFacilityDiversity2024].
+Within the conceptualisation of the 15-Minute City, diversity of land-use is a key feature.
+In this way, the entropy measure theoretically complements the 15-Minute City.
 
-In addition, classification algorithms have been useful tools in the identification of mobility and spatial typologies in relation to transportation planning.
-Machine learning approaches have often been deployed.
-For instance, @burkeGeospatialAnalysisFramework2022 uses spatially constrained multivariate clustering, a regionalisation technique, to develop urban form typologies related to the 15-Minute City.
-The k-means algorithm is another popular method; @ganUnderstandingUrbanMobility2020 applies it to analyse travel behaviour and classify metro stations based on mobility patterns.
-Similarly, [@montero2023role; @montero2023applying] cluster travel behavior profile typologies representing different population segments and place of residence in Barcelona Metropolitan Area.
-Combing classification techniques is also common; for instance, @delmelleDifferentiatingPathwaysNeighborhood2017 uses self-organizing maps (SOM) algorithm to group US neighbourhoods by minimizing dissimilarity in attributes before applying k-means for further classification.
-Similarly, @victorianoTimeSpaceMoney2020 uses SOM to classify individuals' travel patterns according to the dissimilarity in mobility attributes and uses the decision tree algorithm to partition the data into interpretable classifications.
-
-All-in-all, within the context of examining the 15-Minute City, a variety of tools have been applied in the literature.
-However, the question of _15-minute cities for whom_ sill demands further examination.
-This work aims to bridge the concept of the 15-Minute City with another normative framework: Mobility of Care.
+Classification algorithms also show promise as a tool in the assessment of the 15-Minute City, as they have been useful in identifying mobility and spatial typologies within the transportation planning literature.
+Often, these algorithms take the form of machine learning approaches: for instance, @burkeGeospatialAnalysisFramework2022 uses spatially constrained multivariate clustering, to develop urban form typologies related to the 15-Minute City.
+The k-means algorithm is another popular method: @ganUnderstandingUrbanMobility2020 applies it to analyse travel behaviour and classify metro stations based on mobility patterns.
+Similarly, [@montero2023role; @montero2023applying] cluster travel behavior profile typologies representing different population segments and place of residence in the Barcelona Metropolitan Area.
+Combing classification techniques is also common: for instance, @delmelleDifferentiatingPathwaysNeighborhood2017 uses self-organizing maps (SOM) algorithm to group U.S.
+neighbourhoods by minimizing dissimilarity in attributes before applying k-means for further classification.
+Similarly, @victorianoTimeSpaceMoney2020 uses SOM to classify individuals' travel patterns according to the dissimilarity in mobility attributes and uses the Decision Tree algorithm to partition the data into interpretable classifications.
 
@@
 Rather than focusing on _all_ destinations, it may valuable to examine those related to caring activities.
-This re-framing offers a feminist perspective on  _what_ urban functions matter, connecting well with the 15-Minute City concept.
-Caring activities, which meet the physical, psychological, and emotional needs of others, are among the most essential and fundamental activities in society [@ilo_care_2018].
+This re-framing offers a feminist perspective on which urban functions matter, connecting well with the 15-Minute City concept.
+Caring activities, which fulfill the physical, psychological, and emotional needs of others, are among the most essential and fundamental activities in society [@ilo_care_2018].
 Yet, they are the most unequal, undervalued, and even devalued activities worldwide.
@@
 
-The Mobility of Care concept also explicitly integrates inter-sectional equity considerations.
-In considering Mobility of Care within the 15-Minute City one could address a common criticism leveled at the 15-Minute City [@guzmanProximityEnoughCritical2024].
+The Mobility of Care concept also explicitly integrates inter-sectional equity considerations, a common criticism leveled at the 15-Minute City [@guzmanProximityEnoughCritical2024].
 Perhaps unsurprisingly given the gendered division of care work discussed, women have consistently been found to complete more mobility of care trips than men.
@@
 37).
-As a product of the masculinst bias in planning, transport surveys and tools most of the times do not directly capture mobility of care, which re-enforces the idea that these trips are not a significant part of daily mobility.
+As a product of the masculinst bias in planning, transport surveys and tools often do not directly capture mobility of care, which re-enforces the idea that these trips are not a significant part of daily mobility.
 In this respect, the feminist perspective of the cities of proximity is still underestimated with only few examples on the topic [@gil_sola_choose_2022; @macintyre_her_2022].
@@
 
-The objectives of this research are two-fold. 
-- First, we aim to theoretically bridge Mobility of Care with the 15-Minute City concept to re-imagine what local amenities matter and define an associated accessibility and diversity of accessibility measure.
-- Second, taking a bottom-up data-driven approach, we apply a classification and clustering machine learning methodologies to support the bottom-up identification of _15-minute caring neighbourhoods_  through an empirical examination of the mid-cited city of Hamilton, Canada.
-These neighbourhoods are then interpreted on the continuum of having (and not having) the region-relative land-use and transportation infrastructure to support _completely caring_ access.
-Who currently resides in which neighbourhood is examined, and ways to further expand the care accessibility story are discussed.
+In summary, we reviewed the neighborhood planning framework from which the 15-Minute City concept stems and is situated within.
+We then reviewed tools common and useful in investigating the 15-Minute City.
+Finally, we introduced the Mobility of Care concept, theoretically linking it to the 15-Minute City to highlight the importance of care destinations for all but especially along gendered and socio-economic lines.
+The objectives of this work are two-fold, achieved through an empirical examination of Hamilton, Canada, a mid-sized city:
 
-<!-- In this context, we bridge the 15-Minute City with the Mobility of Care concept, to add significance to what destinations and associated trips matter for all, but especially along gendered and socio-economic lines.
-We apply the "Caring 15-Minute City" concept to measure what neighourhoods in the mid-size Canadian city of Hamilton have the potential to be "caring 15 minute neighbourhoods".
-Making use of the tools reviewed and applied in the transportation research, the objectives of the paper are completed as follows: -->
+- First, by theoretically bridging Mobility of Care with the 15-Minute City concept, we re-imagine what local amenities matter and calculate associated accessibility and entropy measure for care destinations.
+- Second, using a bottom-up data-driven approach, we apply machine learning methodologies to classification areas into _15-minute caring neighbourhoods_ and investigation their associated residential profiles along socio-economic lines.
+
+<!--  the objectives of the paper are completed as follows: -->
 
@@
 <!-- 2) Based on these accessibility and diversity outputs, "Completely Caring 15-Minute Neighbourhood" typologies are generated using the SOM algorithm.   -->
-<!-- 3)	A decision tree is then applied to these typologies and socio-economic variables of populations who reside in these neighbourhoods.
+<!-- 3)	A Decision Tree is then applied to these typologies and socio-economic variables of populations who reside in these neighbourhoods.
 This is done to delineate data-driven profiles on which groups have the potential to benefit from what Completely Caring 15-Minute Neighbourhood.
@@
 The densely populated inner city is characterised by lower average incomes, and a higher prevalence of households living under the low-income cutoff thresholds (LICO).
-Noteably, the suburban areas of the city tend to have a greater proportion of children and a lower proportion of one parent households.
+Notably, the suburban areas of the city tend to have a greater proportion of children and a lower proportion of one parent households.
 
-`r word_comment("After reading the results and discussion, do you think fig 2 plots are helpful? Maybe they should be removed?", " ")`
 ```{r plot-figure2, out.width=600, out.height=600}
@@
 
+\newpage
+
 ## Care destination dataset
 
-A spatial dataset of care destinations for Hamilton was carefully compiled.
-The dataset includes 14 types of destinations that were placed by authors in ( _work removed for double-blind review_ ) into four categories: dependent-centric (e.g., the destinations for child- and elder-centric escorting trips), grocery-centric, health-centric, and errand-centric.
+A spatial dataset of care destinations for Hamilton was compiled as detailed in ( _work removed for double-blind review_ ).
+The dataset includes 14 types of care destinations that were classified into four categories: dependent-centric (e.g., the destinations for child- and elder-centric escorting trips), grocery-centric, health-centric, and errand-centric.
 Notably, these categories were generated following the travel purpose categories created in the mobility of care research by @sanchezdemadariagaMeasuringMobilitiesCare2019.
-Category sources of data and preparation notes are detailed in @tbl-Tbl1.
+Care categories, sources of data, and preparation notes are detailed in @tbl-Tbl1.
 The spatial distribution of destination type are visualised in @fig-Fig3 by their category.
-
-`r word_comment("Any ideas on how to condense this table...? Maybe we can cite the forthcoming JTG article, what do you think?", " ")`
 
@@
 dat <- data.frame(matrix(c("Depedent-centric",
+                           "Schools, daycares, and community centres, recreation centres, parks, senior centres, long-term care homes, and retirement homes: 1,265 locations are included.",
                            "[@opendatahamiltonEducationalInstitutions2022; @governmentofontarioOntarioDataCatalogue2023; @cityofhamiltonChildCareRegistry2023; @opendatahamiltonParksDatabase2022; @opendatahamiltonRecreationCommunityCentres2022; @ontarioministryofhealthgeohubMinistryHealthService2023]",
-                           "Schools, daycares, and community centres, recreation centres, parks, senior centres, long-term care homes, and retirement homes: 1,265 locations are included.
-After manual review, all locations that typically do not serve children were removed including: Post-Secondary, Adult-Learning Centres, Group Homes, and Foster Care Centres.
-Further, through examination some Section 23 institutions defined as _“centres for children who cannot attend school to meet the needs of care or treatment, and rehabilitation”_ [@governmentofontarioministryofeducationFundingEducationCommunity2023], were kept due to their innate connection to care.",
                            "Grocery-centric",
+                           "Convenience stores and grocery stores (e.g., large retailers as well as speciality food grocers, health food grocers): 381 destinations are included.",
                            "[@axledataConsumerData2023]",
-                           "Grocery stores, namely a place a household could buy groceries ranging from convenience stores to large retail stores: 381 destinations are identified.
-Data is filtered by Company Name, Suite Number, Address, City, Province, Phone Number and Postal Code.
-The type was then identified e.g., grocers specialty foods, grocers retail, grocer health food, grocer wholesale, grocer curbside, grocer delicatessen wholesale, grocer convenience.
-Data was cross-referenced to ensure all included locations were operational and legitimate grocery stores.",
                            "Health-centric",
+                           "Hospitals, pharmacies, clinics, and dentist offices: 421 destinations are identified.",
                            "[@ontarioministryofhealthgeohubMinistryHealthService2023; @hamiltonniagarahaldimandbranthealthlineWalkInMedicalClinics2023]",
-                           "Hospitals, pharmacies, clinics, and dentist offices: 421 destinations are identified.
-Hospitals and pharmacies were retrieved while clinics and dentistry clinics were manually scraped from a healthcare services database and checked via Google Maps to remove non-operational locations and confirm dentistry-orientation.
-",
                            "Errand-centric",
-                           "Hamilton libraries [@opendatahamiltonLibraries2022], post office locations [@axledataConsumerData2023;@canadapostPostOfficeLocator2023], and datasets of all national bank chains [@bankofmontrealBankMontrealBranch2023; @thehongkongandshanghaibankingcorporationlimitedhsbcHSBCBranchABM2023; @nationalbankNationalBankBranches2023; @royalbankofcanadaRBCBranchABM2023; @scotiabankScotiabankABMBranch2023; @thetorontodominionbankTDBankBranch2023].",
-                                                      "Libraries, post offices, and banks: 158 destinations are identified.
-Post offices are retrieved from a mix of databases, and duplicates are removed.
-Banks are also derived from Data Axle and then cross-referenced to ensure data quality with a Bank Locator website for all national banking firms."
+                           "Libraries, post offices, and banks: 158 destinations are identified.",
+                           "[@opendatahamiltonLibraries2022; @axledataConsumerData2023; @canadapostPostOfficeLocator2023; @bankofmontrealBankMontrealBranch2023; @thehongkongandshanghaibankingcorporationlimitedhsbcHSBCBranchABM2023; @nationalbankNationalBankBranches2023; @royalbankofcanadaRBCBranchABM2023; @scotiabankScotiabankABMBranch2023; @thetorontodominionbankTDBankBranch2023]."
                            ), ncol = 3, byrow = T))
@@
 
-```{r creating-table}
-#| tbl-cap: "Details on the preparation and data sources of care destinations."
+```{r creating-table, ft.arraystretch = 1}
+#| tbl-cap: "Description of care destinations categories, notes on data preparation and associated data sources."
 #| label: tbl-Tbl1
-set_flextable_defaults(font.size= 10,
+set_flextable_defaults(font.size= 8,
                        font.family = NULL,
@@
   compose(
-    i = 1, j = c(1,2,3), part = "header", value = as_paragraph(c("Care category", "Sources", "Data preparation notes"))) %>%
+    i = 1, j = c(1,2,3), part = "header", value = as_paragraph(c("Care category", "Care destination types", "Sources"))) %>%
   width(j = 1, width = 1.7, unit = "cm") %>%
-  width(j = 2, width = 4.3, unit = "cm") %>%
-  width(j = 3, width = 11, unit = "cm")
+  width(j = 2, width = 10, unit = "cm") %>%
+  width(j = 3, width = 5.3, unit = "cm")|>
+  flextable::line_spacing(space = 0,
+                          part = "body")
 ```
@@
 
-## Travel time to care estimations
+\newpage
 
-Summarized in the literature review, empirical travel behaviour to care-oriented destinations is often uncounted and thus travel time is unavailable.
-Hence in this work, travel time from the residential parcel locations and care destination locations is approximated for walking assuming an average speed (3.6 km/hr) using the 'travel_time_matrix()' function from the {r5r} package [@pereiraR5rRapidRealistic2021].
-The inputs into the function are: the 143,882 locations of residential parcels as origins, the 2,225 care locations as destinations, and a OpenStreetMap road network including walking infrastructure [@geofabrikOntarioCanadaOpen2023].
-In theme with what destinations can be reached within 15-Minutes, a maximum walking travel time of 15 minutes is selected, and an origin-destination travel time matrix of the shortest travel time from origin to destination is calculated.
+## Travel time to care destinations estimations
+
+Travel behaviour to care-oriented destinations is often uncounted and hence comprehensive travel times to all destination types in @tbl-Tbl1 is unavailable.
+To overcome this gap, travel time from residential parcel locations in Hamilton and care destinations in @fig-Fig3 are approximated assuming travel by foot at an average speed (3.6 km/hr).
+This estimation is done using the 'travel_time_matrix()' function from the {r5r} package [@pereiraR5rRapidRealistic2021] using R version 4.3.2 [@coreR2023].
+The inputs into the function are: the locations of 143,882 residential parcels as origins, the 2,225 care locations as destinations, and a OpenStreetMap road network including walking infrastructure [@geofabrikOntarioCanadaOpen2023].
+In line with the 15-Minute City, a maximum walking travel time of 15 minutes is specified and an origin-destination travel time matrix of the shortest travel time from origin to destination is calculated.
 The resulting matrix contains 2,014,502 rows, representing walking travel times from each parcel to reachable care destinations within 15 minutes.
@@
 
-In the following sections, the methods taken to calculate the accessibility to each of the 14 destination types from each of the 143,882 residential parcel locations are detailed.
-Then, the entropy measures used to calculate the diversity of accessibility to each of the care categories is described.
-Subsequently, the accessibility and diversity of accessibility values associated with each residential parcel location are fed into a self-organizing map (SOM) algorithm.
-The resulting output is then classified into 7 superclusters the amount of "caring access" and "completeness" of this access.
-A Decision Tree, using the superclusters as features and socio-economic variables for the populations who reside in the city as lavels, is generated to identify patterns on who may benefit (or be excluded from) the potential of Completely Caring 15-Minute Neighbourhoods.
-
-In sum, this methodology presents a data-driven approach to examine what neighbourhoods in a city have the potential to provide 15-minute caring access, at what level of intensity and completeness, as well as in who may be residing in areas that benefit the most.
+The following sub-sections detail the methods to classify Hamilton into spatial degrees of _caring 15-minute neighbourhood_.
+First, accessibility to each of the 14 destination types from each of the 143,882 residential parcel locations is described.
+Second, the entropy measures used to calculate the diversity of accessibility to each of the care categories is detailed.
+Third, we describe how accessibility and diversity values for each parcel are input into a Self-Organizing Map (SOM) algorithm, and how the resulting clusters are analyzed using a Decision Tree algorithm to narrate residential profiles based on socio-economic variables.
+In sum, this methodology presents a data-driven approach to examine what neighbourhoods in a city have the potential to provide 15-minute caring access, at what level of intensity and completeness, and the socio-economic characteristics of those who are most benefited or burdened.
 
@@
 To capture the quantity of access to each type of destinations, a cumulative opportunity accessibility score $S^t_i$ is calculated.
-14 scores for each type of care destination $t$ is calculated for every parcel $i$.
+Scores for each of the 14 care destination types $t$ is calculated for every parcel $i$.
 The calculation takes the following mathematical form:
@@
 
-To represent the diversity of care destination accessibility, the entropy measure is used.
-A value is calculated for each parcel that ranges between 0 to 1, where 1 represents total evenness in the number of care opportunities in each category that can be reached.
+The entropy measure, as defined in @cerveroTravelDemand3Ds1997, is used to represent the diversity of care destination accessibility.
+For each parcel, a value between 0 to 1 is calculated, where 1 represents total evenness in the number of care opportunities in each category that can be reached.
+The mathematical formulation takes the following form:
 
-The mathematical formulation takes the following form:
 $$
@@
 - $D_i$ is the diversity score.
-As $D_i$ represents evenness in type of categories, so if a parcel has a access score of $S_i^t$ =0.5 for all types of destinations it will be assigned $D_i$ = 1 in the same was as if it had a score of $S_i^t$ =10 for each type of destination.
-Contrariwise, a parcel can be assigned a low $D_i$ score if it has low but different accessibility scores for each category as well as if it has high but different scores for each category.
 
-## Machine learning classification: self-organizing maps and decision trees
+Notably, $D_i$ represents evenness in the type of care categories a parcel can access.
+For example, if a parcel has an access score of $S_i^t = 0.5$ for all destination types, it will receive a diversity score of $D_i = 1$, just as if it had an access score of $S_i^t = 10$ for each destination.
+Conversely, a parcel may be assigned a low $D_i$ score if its accessibility scores differ across categories, regardless of whether those scores are low or high.
 
-In this work, we use two machine learning techniques.
-First, SOM is an unsupervised technique implemented to reduce the data dimensional and create interpretable clusters related to the intensity and completeness of caring access.
+## Machine learning classification: SOM and Decision Trees
+
+We use two machine learning techniques in this work.
+First, SOM is an unsupervised technique implemented to reduce the data dimensionality and create interpretable clusters related to the intensity and completeness of caring access.
 This is done by imputing each parcel as an observation with its associated accessibility and diversity attributes, and the data being rearranged onto a two-dimensional space based on its minimizing dissimilarity in its neighbourhood.
-Then, a decision tree is run to characterise the socio-economic profile of who resides in neighbourhoods associated with certain (in)complete and (un)caring accessibility clusters.
-Together, this combined approach leverages the unsupervised data-driven classification power of the SOM with the interpretation of decision trees.
-The procedure used in this work is similar to the approache used in @victorianoTimeSpaceMoney2020.
-However, instead of each observation representing an individual's daily mobility behaviour (with associated variables), here each observation represents a parcel location with calculated care accessibility and diversity of type of care accessibility scores.
+An appropriate number of superclusters are selected and assigned labels based on the quantity and diversity of care access provided, i.e., the degree by each the parcel is located in a _caring 15-minute neighbourhood_.
+Then, a Decision Tree is run to characterise the socio-economic profiles of who resides in neighbourhoods associated with the superclusters.
+Together, this combined approach leverages the unsupervised data-driven classification power of the SOM with the interpretation of Decision Trees.
+The procedure used in this work is similar to that used in @victorianoTimeSpaceMoney2020.
+In this work, rather than each observation representing an individual's daily mobility behaviour (with associated variables), each observation represents a parcel location with calculated care accessibility and accessibility diversity scores.
 
-For the SOM step, the SOM implementation in the function 'trainSOM()' from {SOMbrero} R package is used [@VillaVialaneix2017].
-The input variables include the 143,882 parcels, each as individual observations along with 15 variables: the 14 calculated accessibility scores $S_i^t$, normalized to the min-max range score within each $t$, and one diversity value $D_i$.
+For the SOM step, the function 'trainSOM()' from {SOMbrero} R package is used [@VillaVialaneix2017].
+The input variables include the 143,882 parcels, each as individual observation along with 15 variables: the 14 calculated accessibility scores $S_i^t$, normalized to the min-max range score within each $t$, and one diversity value $D_i$.
 Otherwise, defaults for all other parameters are assumed, relying on the data-driven heuristics embedded in the 'trainSOM()' function.
 Consequently, a 100 node (10 by 10 grid) SOM structure using euclidean distance and square topology is produced.
-Simply put, the SOM algorithm proceeds as follows: a 2D grid of nodes is created as specified by the analyst, where each node will represent a point in the reduced-dimensional space.
+Simply put, the SOM algorithm proceeds as follows: a 2D grid of nodes is created as specified by the analyst, where a node represents a point in the reduced-dimensional space.
 Upon initialization, each node is assigned a random weight vector of the same dimension as the input data (in our case, 15).
-From the input data, a random observation with its associated weight vector (i.e., one parcel point with 15 variables) is selected and the Euclidean distance between its weight vector and all nodes in the grid is calculated.
+From the input data, a random observation with its associated weight vector (i.e., one parcel point with 15 variables) is selected and the Euclidean distance between its weight vector and all nodes in the grid are calculated.
 The node with the smallest distance (i.e., the smallest dissimilarity) is labelled the 'best matching unit' as it is the node that best represents the input observation.
@@
 
-For the decision tree step, the supercluster-classified parcels identified in the SOM step are used as _labels_ and socio-economic and demographic indicators available from the 2021 Canadian Census [@governmentofcanadaCensusPopulation2023] related to the mobility of care literature are used as _features_.
-This step is used to profile the superclusters so we may explore who reside in what supercluster in a data-driven way.
-To estimate the decision tree, the 'rpart()' function in the {rpart} R package is used [@Therneau2023]; default parameters for classification splitting along with each value being weighted by the population present in the associated DA is assumed.
-To summarize the decision tree algorithm, it is a supervised learning technique that begins by splitting a subset of the input data into branches based on a selected feature with the lowest impurity score (i.e., the least mixing of label membership within a branch).
-This process is recursively repeated for each data subset, selecting the next best features.
-Ultimately, the data is classified into distinct classes, with class membership explained by traversing the branches defined by the features that characterise the partitions within the decision tree.
-Notably, the absence of features from the decision tree does not necessarily imply they are irrelevant for classification, rather, they are just less relevant than other features.
+For the Decision Tree step, the supercluster-classified parcels identified in the SOM step are used as _labels_ and socio-economic and demographic indicators related to the Mobility of Care literature are used as _features_.
+Features are retrieved from the 2021 Canadian Census [@governmentofcanadaCensusPopulation2023].
+This step creates residential profiles of the superclusters, allowing us to explore the characteristics of residents within the superclusters in a data-driven way.
+To estimate the Decision Tree, the 'rpart()' function in the {rpart} R package is used [@Therneau2023]; default parameters for classification splitting along with each value being weighted by the population present in the associated DA is assumed.
+The Decision Tree algorithm is a supervised learning technique that begins by splitting a subset of the input data into branches based on a selected feature with the lowest impurity score (i.e., the least mixing of label membership within a branch).
+This process is recursively repeated for each data subset, selecting the next best feature.
+Ultimately, the data is classified into distinct classes, with class membership explained by traversing the branches defined by the features that characterise the partitions within the Decision Tree.
+Notably, the absence of features from the Decision Tree does not necessarily imply they are irrelevant for classification, rather, they are less relevant than other features.
 Put another way, when considering features that are highly correlated, such as income level and LICO, not all relevant variables may be present in the tree [@victorianoTimeSpaceMoney2020].
@@
 The median parcel value for each DA is visualised in @fig-Fig4 and the median parcel diversity of care destinations accessible is presented in @fig-Fig5.
-Together, @fig-Fig4 and @fig-Fig5 represent the 15 variables that served as inputs into the SOM algorithm.
+Together, @fig-Fig4 and @fig-Fig5 represent summaries of 15 variables that served as inputs into the SOM algorithm.
 
@@
 #| label: fig-Fig4
-#| fig-cap: "The number of care destinations that can be reached within 15 min walk per care category for a median parcel in each DA.
+#| fig-cap: "The number of care destinations that can be reached within a 15-min walk per care category for a median parcel in each DA.
 The values are a summary of the 14 acessibility scores that served as inputs into the SOM.
@@
 
-```{r plot-figure5, out.width=600, out.height=600}
+```{r plot-figure5, out.width=450, out.height=450}
 #| label: fig-Fig5
-#| fig-cap: "The diversity measure based on the proportion of care category spatial accessibility (figure 4).
-These values are a summary of the 15th input into the SOM.
+#| fig-cap: "The diversity measure based on the proportion of care category accessibility in Figure 4.
+These values are also a summary of the 15th input into the SOM.
 Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
@@
 
-## Identification of (in)completely (un)caring neighbourhoods
+\newpage
+
+## Identification of completely caring 15-minute neighbourhood typologies
 
@@
 In many ways, @fig-Fig4 mirrors the spatial distribution of care destinations (@fig-Fig3) as the range of 15-minutes of travel is small.
-In @fig-Fig5, areas that have high care accessibility tend to have high diversity as well, though there are exceptions in pockets of the city outside the downtown core that have lower but equally low levels of accessibility for all care categories.
+In @fig-Fig5, areas that have high care accessibility tend to have high diversity as well, though there are exceptions in pockets of the city outside the downtown core that have higher diversity but low levels of accessibility for all care categories.
+Similarly, there are areas with low diversity within the downtown core that have only moderate or high accessibility to certain care destination types.
 
@@
 
-```{r plot-figure6, out.width=800, out.height=800}
+```{r plot-figure6, out.width = '100%', out.height = '100%'}
 #| label: fig-Fig6
 #| fig-cap: "The 7 resulting superclusters from the SOM the 10 by 10 grid output represented in a dendrogram (left) and the proportion of variance unexplained by cluster (right)."
-#|
 knitr::include_graphics(glue::glue(here::here(), "/figures/all_access_sc_plot_small.png"))
@@
 
-The grade for each supercluster was carefully assigned by the authors by referring to descriptive statistics and their interesting trends.
+\newpage
+
+The grade for each supercluster was carefully assigned by the authors by referring to descriptive statistics and their notable trends.
 These statistics are summarised in @tbl-Tbl2 and in the boxplots in @fig-Fig7.
 
-- A+ and A have exceptionally high caring accessibility for all destination types and high diversity scores in the top quantile or above. Together, these grades represent 24% of all parcels.
+- A+ and A have exceptionally high caring accessibility for all destination types and high diversity scores in the top quantile or above.
+Together, these grades represent 24% of all parcels in the city.
 - A- has high caring accessibility, but low diversity.
@@
 This disbalance lowers its overall diversity score.
-However, it can be characterized as a supercluster that demonstrates potential in being retrofitted to provide A+ level completely caring access.
-This grade  represent 4% of parcels.
+However, it can be characterized as a supercluster that demonstrates potential in being retrofitted to provide A+ or A level of completely caring access.
+4% of all parcels are represented by the A- grade.
 - B+ and B superclusters present about average completely caring access.
-These superclusters can serve as the benchmark for what 'average' completely caring access in Hamilton currently looks like and they represent 15% of parcels.
+These superclusters can serve as the benchmark for what 'average' completely caring access in Hamilton currently looks like.
+These grades represent 15% of all parcels.
 - B- provides above average dependent-centric destination access, particularly parks, daycares and schools, but below average access to other destination types and hence has low diversity scores.
@@
 As these parcels demonstrate caring access to some destinations, they may have the potential to be retrofitted to support complete access to all caring destination types.
-This grade represents 19% of parcels.
-- D superclusters demonstrate the lowest scores all-around, representing room for land-use improvement that addressed complete and caring 15-minute access.
-This supercluster characterizes the most number of parcels, representing 38% of parcels in the city.
+This grade represents 19% of all parcels.
+- D superclusters demonstrate the lowest scores all-around, representing room for land-use improvement that addresses complete and caring 15-minute access.
+This supercluster characterizes the largest number of parcels, representing 38% of parcels in the city.
 
-```{r creating-table2-pt2}
+```{r creating-table2-pt2, ft.arraystretch = 1}
 #| tbl-cap: "Mean and (standard deviation) of each SOM classified cluster by input variable and additional summary variables.
@@
   bg(j= 8, bg = "#D75C56") |>
-  hline(i = c(3,10,14,18))
+  hline(i = c(3,10,14,18))  |>
+  flextable::line_spacing(space = 0,
+                          part = "body")
 
@@
 
-```{r plot-figure7, out.width=600, out.height=300}
+```{r plot-figure7, out.width=800, out.height=300}
 #| label: fig-Fig7
@@
 
-<!-- D (Nothing much: Not caring & not complete) (C7) -->
+\newpage
 
-<!-- B- (Below average, but dep- heavy: Somewhat caring & not complete) (C4) -->
-<!-- B (Average: Caring & complete) (C5) -->
-<!-- B+ (Above average: Caring & very complete) (C6) -->
-
-<!-- A- (Excellent, but dep- heavy: Very caring & not complete (C3) -->
-<!-- A (Excellent: Very caring & very complete) (C2) -->
-<!-- A+ (Exceptional: very caring & very complete) (C1) -->
-
-To spatially demonstrate where the superclusters are located, representative supercluster labels for each DA in visualised in @fig-Fig8.
+To spatially demonstrate where the superclusters are located, representative supercluster labels for each DA are visualised in @fig-Fig8.
 This visualisation is created by grouping parcels by their DA, calculating the median supercluster label for each label, and selecting the label that is most dominant within that DA.
@@
 
-`r word_comment("What do you think about these grades as labels?
-Maybe there's another way we can frame it.
-A+ and A are excellent, B+ and B are average, D are bad.
-However A- and B- show promise: they have some caring access (child-centric) and can be improved!
-A- can be more easily improved than B-.", " ")`
-```{r plot-figure8, out.width=600, out.height=600}
+```{r plot-figure8, out.width=450, out.height=450}
 #| label: fig-Fig8
-#| fig-cap: "The maximum median parcel supercluster membership per DA. Escarpment is visualised as a grey dashed line."
+#| fig-cap: "The maximum median parcel supercluster membership per DA.
+Escarpment is visualised as a grey dashed line.
+Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
 knitr::include_graphics(glue::glue(here::here(), "/figures/SClabel_plot.png"))
 ```
+
 The resulting superclusters visualised in @fig-Fig8 are a combination of the cumulative opportunity and diversity of access measures translated into meaningful typologies through the SOM methodology.
-Accessibility is the measure of _potential_ interaction, e.g., how many care destinations one could reach within a 15-minute walk, these typologies are useful in identifying which areas of the city are providing regionally high levels of complete caring access, which areas are regionally average, and which areas are regionally below average.
+To reiterate, accessibility is the measure of _potential_ interaction, e.g., how many care destinations one could reach within a 15-minute walk, these typologies are useful in identifying which areas of the city are providing regionally high levels of complete caring access, which areas are regionally average, and which areas are regionally below average.
 Furthermore, neighbourhoods which provide disbalanced access can be further investigated.
 
-In @fig-Fig8, it can be observed that the excellent and exceptional neighbourhoods (A, A+), which are very caring and very complete 15-minute neighbourhoods, are located within the center of the city and closest to the shoreline in the Lower City.
-The above and below average neighbourhoods (B+ B) are often proximate to the excellent and exceptional neighbourhoods and are more prevalent in the Mountain part of the center of Hamilton.
+In @fig-Fig8, it can be observed that the excellent and exceptional neighbourhoods (A, A+), which are very caring and very complete 15-minute neighbourhoods, are located within the center of the city and closest to the shoreline in the downtown core of Hamilton.
+The above and below average neighbourhoods (B+ B) are often proximate to the excellent and exceptional neighbourhoods and are more prevalent in south of the escarpment within the center Hamilton.
 The escarpment is a physical barrier, with few pedestrian-accessible access points to traverse; hence the typologies describing neighbouring DAs separated by the escarpment are often different.
@@
 
-Furthermore in @fig-Fig8, A- and B- grades stand out as offering high access for children-centric destinations for their grade-group but below group-average access in other destination types.
+Furthermore in @fig-Fig8, A- and B- grades stand out as offering high access for children-centric destinations for their grade-group but below grade-group average access in other destination types.
 These neighbourhoods may be more suitable for populations who prioritize walkable access to destinations like schools, parks and daycares, and find access to other types of caring destinations less important.
@@
 
-## Profiling of who resides in (in)completely (un)caring neighbourhoods
+## Profiles of who does and does not reside in caring 15-minute neighbourhoods
 
@@
 While the algorithm is unable to homogenously use each supercluster, @fig-Fig9 is helpful to report a narrative of who may reside in what caring/complete supercluster.
-Particularly, the Decision Tree demonstrates a more pure supercluster membership for only two of the three terminal decision tree nodes: supercluster A+ (exceptional completely caring access) and D (not caring and complete access).
+Particularly, the Decision Tree demonstrates a more pure supercluster membership for only two of the three terminal Decision Tree nodes: supercluster A+ (exceptional completely caring access) and D (not caring and complete access).
 For reference, the city-wide mean household income is `r df_DA_nogeom$Income |> mean(na.rm=T)|> round() |> prettyNum(big.mark = ",") ` (SD: `r df_DA_nogeom$Income |> sd(na.rm=T)|> round() |> prettyNum(big.mark = ",")` and median: `r quantile(df_DA_nogeom$Income,0.5, na.rm=T) |> unname()  |> prettyNum(big.mark = ",")`).
-The notable splits for income are $\gt$ \$91,500 $\lt$ \$68,750, or between \$68,750 and \$91,500, roughly split by lower income, middle income, and higher income median household brackets.
+The notable splits for income are >\$91,500, >\$68,750, or between \$68,750 and \$91,500; roughly split by lower income, middle income, and higher income median household brackets.
 
@@
 
-```{r plot-figure9, out.width=600, out.height=300}
+```{r plot-figure9, out.width=600, out.height=450}
 #| label: fig-Fig9
-#| fig-cap: "The decision tree demonstrating median household income splits and superclusters composition within each branch."
+#| fig-cap: "The Decision Tree demonstrating median household income splits and superclusters composition within each branch."
 knitr::include_graphics(glue::glue(here::here(), "/figures/access_sc_profiles_justincome.png"))
 ```
+
 Describing @fig-Fig9, the bar column on the right represents DAs with higher household income.
@@
 The higher income column represents 39% of the population, the largest proportion of any of the three columns.
-The left column in @fig-Fig9 correpnds to the lowest income, i.e., below 68,750 CAD median household income, and accounts for 32% the population.
+The left column in @fig-Fig9 corresponds to the lowest income, i.e., below 68,750 CAD median household income, and accounts for 32% the population.
 It is dominated by A+ parcels along with A and A- parcels, in the largest quantities relative to other bar columns, however, proportions of all other superclusters are also present.
@@
 Though median income was the most useful in partitioning the parcels by their supercluster labels in @fig-Fig9, other variables are still important.
-Namely, the following variables were also important (in order of importance and correlation coefficient with the median household income variable in brackets): the % below the median household income (`r Income_inspect[c("Income_bottom"),] |> round(digits=2)`), % single parent household (`r Income_inspect[c("One_parent_families"),] |> round(digits=2)`), % no certification or with only a highschool diploma (`r Income_inspect[c("Low_or_no_deg"),] |> round(digits=2)`), the average number of children per household (`r Income_inspect[c("Avg_numb_of_kids"),] |> round(digits=2)`), % not in the labour force (`r Income_inspect[c("Not_in_labforce"),] |> round(digits=2)`), % LICO prevalence (`r Income_inspect[c("LICO"),] |> round(digits=2)`), and % who walk to work (`r Income_inspect[c("walk_towork"),] |> round(digits=2)`).
-<!-- Of note, when median household income is removed from the decision tree, a similar structure is created, except with above average Single-Parent family and LICO prevalence splitting the data.-->
+Namely, the following variables in order of importance and correlation coefficient with the median household income variable in brackets, are listed: the % below the median household income (`r Income_inspect[c("Income_bottom"),] |> round(digits=2)`), % single parent household (`r Income_inspect[c("One_parent_families"),] |> round(digits=2)`), % no certification or with only a highschool diploma (`r Income_inspect[c("Low_or_no_deg"),] |> round(digits=2)`), the average number of children per household (`r Income_inspect[c("Avg_numb_of_kids"),] |> round(digits=2)`), % not in the labour force (`r Income_inspect[c("Not_in_labforce"),] |> round(digits=2)`), % LICO prevalence (`r Income_inspect[c("LICO"),] |> round(digits=2)`), and % who walk to work (`r Income_inspect[c("walk_towork"),] |> round(digits=2)`).
+<!-- Of note, when median household income is removed from the Decision Tree, a similar structure is created, except with above average Single-Parent family and LICO prevalence splitting the data.-->
 
-It is notable that median household income is highly or moderately correlated with many of these important variables, and discussing these correlation is useful for interpretation alongside the Decision Tree diagram.
-For instance, single-parent families are most positively correlated with the proportion of households being in the bottom income distribution (`r Income_inspect[c("Income_bottom"),] |> round(digits=2)`) and low or new degree (`r One_parent_family_inspect[c("Low_or_no"),] |> round(digits=2)`), and most negatively with the median after-tax household income (`r One_parent_family_inspect[c("v_CA21_907: Median after-tax income of household in 2020 ($)"),] |> round(digits=2)`), reflecting nationwide trends [@governmentofcanadaDailyMainHighlights2024; @governmentofcanadaPrevalenceLowIncome2024].
-Though not all single-family households are below the LICO (single-family households are not highly correlated with LICO at the DA-level in Hamilton), DAs with higher concentrations of single-family households and LICO prevalence tend to be A+ completely caring 15-minute neighbourhood.
-Conversely, DAs with lower single-parent households are in DAs with a higher concentration of F parcels.
-These findings are notable from an equity perspective and in considering caring 15-minute policy interventions, especially since economic disadvantaged often intersects with other socio-demographic characteristics.
+It is notable that median household income is highly or moderately correlated with many of these important variables, and discussing these correlations is useful for interpretation alongside the Decision Tree diagram.
+For instance, single-parent families are most positively correlated with the proportion of households being in the bottom income distribution (`r Income_inspect[c("Income_bottom"),] |> round(digits=2)`) and lower or no diploma (`r One_parent_family_inspect[c("Low_or_no"),] |> round(digits=2)`), and most negatively with the median after-tax household income (`r One_parent_family_inspect[c("v_CA21_907: Median after-tax income of household in 2020 ($)"),] |> round(digits=2)`), reflecting nationwide trends [@governmentofcanadaDailyMainHighlights2024; @governmentofcanadaPrevalenceLowIncome2024].
+Though not all single-family households are below the LICO (i.e., single-family households are not highly correlated with LICO at the DA-level in Hamilton), DAs with higher concentrations of single-family households and LICO prevalence tend to have A+ access.
+Conversely, DAs with lower single-parent households are in DAs with a higher concentration of parcels with D access.
+These findings are notable from an equity perspective since economic disadvantaged often intersects with other socio-demographic characteristics.
 
@@
 
-`r word_comment("Ideas on this section (idea for an extra paragraph or additional points to each paragraph to make them stronger, with citations) would greatly be appreciated!", " ")`
+Geographically, Hamilton is city that offers very caring - and completely caring - 15-Minute neighbourhoods for some, but not all.
+There are evident spatial inequalities, with areas ranging from excellent (A+ and A), average (B+ and B), to well below average (D).
+Some areas, like those labelled A- and B-, show potential for improvement within their grade groups.
+While the downtown core has the highest concentration of _caring 15-minute neighborhoods_, certain areas outside the core do as well.
+This finding somewhat contrasts employment accessibility studies, such as @elgeneidyNonstopEquityAssessing2016's assessment of public transit access to employment in the Greater Toronto Area (including Hamilton), where access is heavily concentrated in the downtown core, more so than appears to be present in this work.
+Though our paper uses a different methodology, both studies highlight distinct patterns: indeed, A+, A or A- care access is concentrated in the downtown core but it is also present in certain pockets outside of the center of Hamilton, leading to interesting conjectures.
+For instance, some of the A-grade neighborhoods in Hamilton likely follow land-use principles that emphasized residential proximity to amenities, similar to the NUC reviewed.
+The potential overlap of the NUC with caring 15-minute neighborhoods warrants further investigation.
 
-Geographically, Hamilton is city that offers very completely caring 15-Minute neighbourhoods for some, but not everyone.
-There are evident spatial inequalities in areas that are, are not, and are somewhere between a caring and complete 15-minute neighbourhood (from labels A+ and A being excellent, B being about average, D being well below average, and A- and B- demonstrating potential in their grade group).
-The downtown core of the city has the most _caring 15-minute neighbourhoods_, but many pockets of the suburbs do as well.
-Though this contrasts what may be seen when considering employment opportunities [as in the work of @elgeneidyNonstopEquityAssessing2016], the paper's findings align with the neighbourhood unit planning literature reviewed.
-As many North American Suburbs, Hamilton followed certain characteristic design patterns that were initially designed to be proximate to amenities, coinciding with the conceptualisation of _caring 15-minute cities_ [@aliNeighbourhoodPlanningHamilton1991].
-
-Who currently resides in Hamilton's completely caring 15-Minute neighbhorhoods is also an optimistic story.
+Who currently resides in Hamilton's completely caring 15-Minute neighbhorhoods is also demonstrated to be a somewhat optimistic story.
 Parcels that provide A+ completely caring access tend to be in DAs that are economically disadvantaged.
 Economic disadvantage tends to intersect with other identities such as gender [@lightmanMeasuringEconomicExclusion2018].
-And as reviewed in this work, all women and especially those from lower income households tend to complete the most mobility of care trips [@ravensbergen2023exploratory].
+And as reviewed in this work, all women and especially those from lower income households tend to complete most mobility of care trips [@ravensbergen2023exploratory].
 Furthermore, lower-income households tend to also be single-parent households.
@@
 In these ways, the lower income residents of Hamilton's A+ neighbourhoods are more likely to be displaced, which is matter of wellbeing and justice.
-There are few low-rent choices that provide the same exceptional level of access as the downtown core, hence lower household income residents that currently reside in A+ neighhourhoods would likely need to relocate to a different city.
+There are now few low-rent choices that provide the same exceptional level of access as the downtown core, hence lower household income residents that currently reside in A+ neighhourhoods will likely displaced in the coming years if current trends continue.
 
-`r word_comment("How does this paragraph sound..? I like it but I feel like it doesn't have a solid point to make... yet.", " ")`
 In discussing policy interventions that equitably increase completely caring 15-Minute neighbourhoods in Hamilton, this work's presents a methodology to create city-wide relative typologies and investigate who currently resides in what neighbourhood, as a stepping stone for further investigation.
 Neighbourhoods with the lowest grades (D) and with the highest potential in being improved (i.e., high accessibiltiy for certain types of destinations but not all) are neighbourhoods with B- and A- grades.
-However, the work's findings demonstrate a mismatch in household income and completely caring access.
-Is it equitable to focus policy on ameliorating neighbourhoods that are already higher-rent though they tend to be more rural, single-use zoned and car-dependent (parcels with D grades)?
+However, the work's findings demonstrate a mismatch in household income and completely caring access: higher income areas are often the ones targeted for improvement.
+This raises important questions from the land-use policy perspective that need to be further analysed.
+For instance, is it equitable to focus policy on ameliorating neighbourhoods that are already higher-rent though they tend to be more rural, single-use zoned and car-dependent (parcels with D grades)?
 Further, of the parcels that provide high child-centrics but low otherwise, A- parcels (better access) tend to be in DAs with lower-income households more so than B- parcels (lower access than A-).
@@
 Who should benefit, and how?
-If the policy initiative is targeted to specific neighbourhoods: sustainability linked to car dependency and equity are in-tension.
-This harkens to what role a planned neighbourhood, and bottom-up planning approaches that should be adopted to achieve equitable 15-minute cities.
+If the policy initiative is targeted to specific neighbourhoods: sustainability linked to car dependency and equity are in tension.
+This harkens to what role a planned neighbourhood, and bottom-up planning approaches that include the evaluation of travel behaviour by socio-economic and demographic profile should be adopted to achieve equitable 15-minute cities.
 
-As all research work, the results should also be interpreted along with methodological assumptions.
+As is the case for all research, the results should also be interpreted along with methodological assumptions.
 This work measures spatial accessibility which is a measure of potential interaction with all reachable destinations from an origin.
-These destinations however may not be relevant to people at an origin, e.g., the may be underutilized (walkable access to 2 schools is unneeded, as a child only needs to attend 1 school), the trip physically undesirable (the walk may be along an arterial with high traffic speeds, making the trip unlikely to ever happen by foot), or the average walking speed assumed may not reflect the walking speeds of all populations [@willberg15minuteCityAll2023].
+These destinations, however, may not be relevant to people at an origin, e.g., they may be underutilized (walkable access to 2 schools may be underused, as children often only attend 1 school), the trip physically undesirable (the walk may be along an arterial with high traffic speeds, making the trip unlikely to ever happen by foot), or the average walking speed assumed may not reflect the walking speeds of all populations [@willberg15minuteCityAll2023].
 Further, people who reside in these neighbourhoods may also disagree with the neighbourhood's label.
@@
 
-`r word_comment("Thoughts on this section? Strong ending? anything more to add?", " ")`
 This study makes three types of contributions to the transportation and city planning literature: empirical, methodological and theoretical.
@@
 
-Our work is of relevance to research and practitioners planning equitable and sustainable cities.
+Our work is of relevance to researchers and practitioners planning equitable and sustainable cities.
 Instead of prescribing an urban form design principle, such as "all local amenities should be within a 15-minute walking distance", it instead examines an empirical example to determine which areas in the city have the _potential_ to be 15-minute neighbourhood based on the existing spatial accessibility offered by the urban environment and walking transport network.
@@
 :::
-
-<!-- As a tool to visualise the variables that describe the superclusters, a decision tree is fit with the DEPENDENT TOTALS, ERRAND TOTALS, GROCERY TOTALS, HEALTH TOTALS, and CATEGORY DIVERSITY scores.
-Aligning with the descriptive summaries of the superclusters provided, describing the tree-diagram from left to right (roughly from the most caring & complete, to the least), the first three terminal nodes (represented by bar columns that visualise the proportion of supercluster membership at that node) are mostly A+ and A superclusters, and are defined by their high dependent-centric access (about median and above), and high diversity score.
-A+ is differentiated from A by its exceptionally high health-centric totals while A most just has excellent scores.
-Columns four, six and seven are dominated by the presence of B+ and B parcels.
-On average the B+ parcels provide access to an average level of dependent-centric destinations, so the B+ parcels with greater than 6.5 dependent-centric destinations are found in the more left-most columns, and those below this split on the right-most.
-The far right-most column is dominated by the presence of F parcels, defined by lower DEPENDENT TOTALS and LOW DIVERSITY.
-It is interesting to note the 5th column, which is defined by D parcels but also the most amount of A- parcels than any other column.
-A- and D parcels are both high in dependent-centric destinations and low in diversity (though D is much lower than A-).
-These two parcels again, as shown in @fig-Fig7, can be thought of as parcels in neighbourhoods containing a high amount of some caring destinations, but lacking others, indicating high efficiency for improvement potential.
--->
-
-<!-- ```{r plot-figure7, out.width=600, out.height=300} -->
-<!-- #| label: fig-Fig7 -->
-<!-- #| fig-cap: "The decision tree demonstrating variables that define the superclusters." -->
-<!-- knitr::include_graphics(glue::glue(here::here(), "/figures/access_sc_profiles_usingSCVARS.png")) -->
-<!-- ``` -->

### 2024-10-10 paper/Manuscript.tex
Draft to submit to journal. And some repo clean-up.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  authoryear,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{fontspec}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{hhline}
\newlength\Oldarrayrulewidth
\newlength\Oldtabcolsep
\usepackage{longtable}
\usepackage{array}
\usepackage{hyperref}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{unicode-math}
\usepackage{hyperref}
\def\tightlist{}
\usepackage{setspace}
\doublespacing
\usepackage{lineno}
\linenumbers
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Caring 15-minute neighbourhoods},
  pdfauthor={Anastasia Soukhov; Léa Ravensbergen; Lucía Mejía Dorantes; Antonio Páez},
  pdfkeywords={Accessibility, Mobility of Care, Gender, 15-minute
neighbourhood, Care Complete Neighbourhoods},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Caring 15-minute neighbourhoods}
\author[1]{Anastasia Soukhov%
\corref{cor1}%
}
 \ead{soukhoa@mcmaster.ca} 
\author[1]{Léa Ravensbergen%
%
}
 \ead{ravensbl@mcmaster.ca} 
\author[2]{Lucía Mejía Dorantes%
%
}
 \ead{lucia.mejia.dorantes@gmail.com} 
\author[1]{Antonio Páez%
%
}
 \ead{paezha@mcmaster.ca}

\affiliation[1]{organization={McMaster University, School of Earth,
Environment \& Society},city={Hamilton, Canada},postcodesep={}}
\affiliation[2]{organization={Consultant},city={Karlsruhe,
Germany},postcodesep={}}

\cortext[cor1]{Corresponding author}

\begin{keyword}
    Accessibility \sep Mobility of Care \sep Gender \sep 15-minute
neighbourhood \sep 
    Care Complete Neighbourhoods
\end{keyword}
\end{frontmatter}

\section{Abstract}\label{abstract}

The ``15-Minute City'' concept has been embraced by global leaders to
promote human-scale neighbourhoods with transport and land-use designs
that support short trips to daily necessities. This paper bridges the
15-Minute City to ``Mobility of Care'', a framework that foregrounds
travel to care destinations, travel done predominately by women. This
focus contrasts the more commonly studied travel to employment and
leisure destinations. While the 15-Minute City concept is flexible
enough to consider all destination types, gendered examinations are
relatively lacking in the literature, and little research to date has
focused explicitly on care destinations. This gap is addressed in this
paper by identifying which areas in the city of Hamilton, Canada are
`caring 15-minute neighbourhoods'. To do so, a database of care
destinations was compiled to estimate the number (using the cumulative
opportunities accessibility measure) and diversity of mix (using the
entropy measure) of care destinations within a 15-minute walk from
residential parcels. Using data-driven machine learning techniques
(Self-Organizing Maps and Decision Trees), neighborhoods were classified
into `caring 15-minute neighbourhood' typologies that are examined
across residential socio-economic profiles. Our results suggest that the
majority of caring 15-minute neighbourhoods are in the urban core, where
more lower-income households reside. In contrast, not caring 15-minute
neighbourhoods in higher-income peripheral areas. Areas that make good
candidates for urban policy intervention are identified, and the
implications of enhancing 15-minute walkable caring access are discussed
in relation to gender mainstreaming in transportation planning.

Keywords: Accessibility; 15-Minute Neighborhoods; Chrono-Urbanism;
Cities of Proximity; Mobility of Care; Inequality; Gender Gap;

\section{Introduction}\label{introduction}

The ``15-Minute City'' has been recently adopted by leaders as a way to
promote human-scaled cities
\citep{teixeiraClassifying15minuteCities2024}. As introduced in
\citet{moreno_introducing_2021}, the 15-Minute City is a urban planning
model based in chrono-urbanism, a theory that emphasizes the positive
impact on quality of urban life when urban space becomes multi-rythmic,
in opposition to the segregation of urban time and space for individual
uses and mobility
\citep{mulicekUrbanRhythmsChronotopic2015, morenoVilleQuartHeure2016}.
It is in opposition to the forces that supported the formation of
single-use neighbourhoods such as industrial Fordism and single-use
zoning
\citep{mulicekUrbanRhythmsChronotopic2015, morenoVilleQuartHeure2016}.
The ``15-Minute City'' refers to a city where relevant destinations
within a walkable (or bikeable or transit supportive) radius are
reachable by all. This form of neighbourhood planning would allow
individuals to reclaim time spent on car mobility, giving way to
sustainable modes and prompting urban spaces that are responsive to
human needs and environmental sensibilities \citep{Allam2022}.

Though the term 15-Minute City is relatively new, neighbourhood planning
practice aiming to foster community and local travel to amenities is
not. As recent examples, planned neighbourhoods dominated post-WWII
built urban form: the Neighbourhood Unit Concept in the western world
\citep{brodyNeighbourhoodUnitConcept2013} and a parallel concept of the
``mikrorayons'' neighbourhoods in the Soviet Union
\citep{kissfazekasCircleParadigms15minute2022}. Along different
dimensions, these planning forms have been extensively critiqued
\citep{talenSocialSciencePlanned2017}. In this respect, neighbourhood
planning approaches such as the 15-Minute City concept are being
re-embraced by some, but from bottom-up perspective using co-creation
tools and meaningful resident participation to achieve equitable and
justice neighbourhood forms \citep{mahmoudCocreationPathwayUrban2021}.
How to equitably prescribe urban form that `enables' contact with social
opportunities (instead of `engendering it' from the top-down) is still
under debate. Questions remain, including: what destinations really
matter?, what mode and travel time threshold is most appropriate?, and
who benefits? For instance, these dimensions are under discussion among
proximity-based planners who use accessibility-based tools
\citep{silvaProximitycentredAccessibilityConceptual2023, silvaRegionalAccessibilityUndermining2022, guzmanProximityEnoughCritical2024}.

From another perspective, feminist geographers, urban planners and other
researchers agree that urban structure influences travel, but city
planning falls short in planning for a neutral identity. In contrast,
cities ought to be planned according to the multiple identities of the
inhabitants
\citep{vacchelli_towards_2018, urbandevelopmentviennaGenderMainstreamingUrban2013}.
Researchers have demonstrated how mobility patterns differ according to
gender
\citep{lawWomenTransportNew1999, cresswell_gendered_2008, levy2013travel, little_gender_1994, tronto_toward_1990}.
Gendered social norms and cultural factors play a role in how, where,
with whom, what for, and how far we move. Women have higher activities
related to care \citep{ilo_care_2018}, no matter at what life stage they
are \citep{garciaromanGenderDifferencesTime2022}. For all these reasons
the spatial organization of activities has a different impact according
to gender, the 15-Minute City concept should not be gender blind.

This work builds a theoretical bridge between the 15-minute city and
mobility of care concepts to answer questions about what destinations
matter and who benefits from proximity planning. The Mobility of Care
term was coined in
\citet{sanchezdemadariagaMobilityCareIntroducing2013}; it refers to all
travel required to sustain the needs of a household, such as grocery
shopping, escorting children, travelling to health appointments, and
running errands. While decades of research have examined types of
household-serving trips (e.g., shopping trips, escorting to school),
\citet{sanchezdemadariagaMobilityCareIntroducing2013} was the first to
consider them all as one category. In doing so, Mobility of Care is
shown to be a significant proportion of daily travel, characterizing the
trip purpose of approximately 30\% of adults' trips
\citep{sanchezdemadariagaMobilityCareIntroducing2013, sanchezdemadariagaMeasuringMobilitiesCare2019, ravensbergen2023exploratory}.
We argue Mobility of Care is especially relevant to the 15-Minute City,
as care trips are necessary for \emph{all} people and they most often
occur at the local level. Moreover, care trips take on tremendous
gendered and socio-economic importance, as women, and especially those
from lower income households, complete the majority of care trips
\citep{ravensbergen2023exploratory}. Despite care travel's importance in
advancing gender equality -and equity more broadly-, mobility of care
and access to care destinations have been under-examined relative to
employment destinations. This is especially pertinent in accessibility
research that has largely focused on employment points of interest e.g.,
\citep{farberOntarioLineSocioeconomic2019, duarteInfluenceJobAccessibility2023, ryanAccessibilitySpaceTime2023}.

The objectives of this research are two-fold. First, we aim to
theoretically bridge Mobility of Care with the 15-Minute City concept to
re-imagine what local amenities matter and to define an associated
accessibility and diversity of accessibility measures. Second, we take a
data-driven approach to classify areas into `caring 15-minute
neighbourhoods' typologies and examine associated residential profiles
in an empirical examination of the city of Hamilton, Canada.

\section{Review of neighbourhood planning
literature}\label{review-of-neighbourhood-planning-literature}

\subsection{From the 15-Minute City to the
NUC}\label{from-the-15-minute-city-to-the-nuc}

In the last decades, the need for more sustainable, healthier, and
livable cities have received significant attention. Planners and
decision-makers have proposed alternatives to mobility-based
neighbourhood planning principles to ones that center proximity to urban
functions \citep{pozoukidou15MinuteCityDecomposing2021}. In this
context, the 15-Minute City is now in the public spotlight
\citep{logan_xminute_2022, moreno_introducing_2021}. The 15-Minute City
was introduced in \citet{moreno_introducing_2021} along with four
dimensions: density (in terms of people per km\(^2\)), diversity
(including mixed-land use and diversity of people), the temporal and
spatial proximity to essential services, and digitalisation. However,
the framework presented has been criticized within academic and planning
circles \citep{guzmanProximityEnoughCritical2024}. Critiques highlight
how the concept falls short in addressing the pre-existing structural
forces and individual characteristics that drive inequalities,
influencing who benefits or could benefit the most from such a paradigm
\citep{dimarino15minuteCityConcept2023, willberg15minuteCityAll2023}. A
15-Minute City for all people is an aspirational ideal, but how will
existing mobility and accessibility inequalities be addressed. Without
directed and context-specific solutions, this popular concept risks
becoming an empty city branding exercise.
\citep{pozoukidou15MinuteCityDecomposing2021, gowerPlanningInnovationCity2022}.

Reflecting the flexible and aspirational presentation of the 15-Minute
City; cities that have adopted the concept have done so using a diverse
range of definitions and tools along with indistinct universal
approaches. A trail blazer was Portland, US in the Portland Plan
\citep{city_of_portland_20minute_2010} of April 2012. Adopted before the
`15-Minute City' concept of today, it aims to foster an inclusive city
development based on prosperity, education, health, and equity over a 30
year horizon. The plan aims to promote neighbourhood self-sufficiency
and connectivity to city centers and centres of employment. The progress
report describes a high-level focus on equitable service delivery to all
residents with equity concerning topics related to racial equity and
people with disabilities \citep{portland_government_portland_2017},
similarly taking on an ``all populations'' approach. Later, other cities
adopted proximity-based goals using similar neutral approaches. For
example, the 15-Minute City plan announced during the Paris mayor Anne
Hidalgo's 2020 re-election campaign, which helped popularize the concept
in today's public discourse, emphasized six key social functions that
should be easily accessible from any location
\citep{ville_de_paris_paris_2022}. These locations included: housing,
work, health care, groceries, education and leisure. The 15-Minute City
concept inspired language in the agendas of other cities in the Western
world such as Ottawa, in the Canadian context, who also adopted a
15-Minutes approach in their recent Official Plan
\citep{ville_dottawa_quartier_2021}.
\citet{teixeiraClassifying15minuteCities2024}'s global review finds that
the 15-Minute City concept is in early phases of implementation around
the world and the diverse range of definitions, strategies, and
instruments present a significant knowledge and implementation gap. In
other words: the 15-Minute City is aspirational but how do we get there
and who will benefit?

The past can shed light on cautionary tales. While the 15-Minute City
concept is new, neighbourhood planning to improve society outcomes is
not: in fact the literature has drawn parallels from the 15-Minute City
to Clarence A. Perry's Neighbourhood Unit Concept (NUC) from the 1930s
\citep{kissfazekasCircleParadigms15minute2022}. The NUC is a
socio-spatial normative scheme widely adopted by government agencies
(and the real-estate community) in the Western world after the Second
World War \citep{talenSocialSciencePlanned2017, solow1969concept}.
Pairing well with the objectives of planning agencies at the time,
Perry's NUC would allow for efficient mass-building of cellular units
that prioritized the perceived functional needs of women and children:
each unit providing proximate access to an elementary school and
supporting community facilities, shopping, parks and housing
\citep{talenSocialSciencePlanned2017, brodyNeighbourhoodUnitConcept2013}.
The NUC primarily prioritized local service provision, though Perry had
confidence in good design's contribution to `neighbourhood spirit'
\citep{hall2014cities}. From the NUC's conception until the end of the
1960s, planners' aspirational attempts to prescribe social meaning to
the neighbourhood's physical form had been criticized to near extinction
\citep{talenSocialSciencePlanned2017}. By social scientists, the
overestimation on the impact of built environment on social life was
critiqued. By other planners, the inability to agree on the specificity
of the neighbourhood (i.e., population size and the type, quality, and
quantity of amenity) as well as how neighbourhood units are connected to
other neighbourhoods and the rest of the region. In response to these
criticisms, neighbourhood planning proponents redefined deterministic
terminology; their prescriptive physical form will `enable' social
contact with opportunities rather than `engendering' it.

The redefined bottom-up approach to community and neighbourhood planning
has been adopted by New Urbanists, from which the 15-Minute City stems
\citep{kissfazekasCircleParadigms15minute2022}; however, the
effectiveness of these bottom-up processes remain to be seen due to the
contemporary nature of the concept. Though, related research can be
examined. In recent years, the question of \emph{how can physical form
be planned to enable an improved quality of life, for whom, and with
what outcomes} has occupied the urban planning research agenda. For
instance, the examination of low-income households access to
transportation and their likelihood of gaining employment
\citep{blumenbergDriveWorkRelationship2017, bastiaanssenDoesBetterJob2022}
or the relationship between children's access to public transit and
participation in after-school activities
\citep{palmRolePublicTransit2020}. A new wave of researchers and
practitioners focused on local and context-specific relationships with
the proximity to destinations have also emerged
\citep{silvaProximitycentredAccessibilityConceptual2023, silvaRegionalAccessibilityUndermining2022}.
As reviewed in the city plans that have adopted 15-Minute City
approaches, the NUC's criticisms, and recent urban planning research,
the question of \emph{how to enable improved quality of life through
urban built environment} has shown to be highly context-sensitive, with
the need for further investigation.

\subsection{Tools: accessibility methods, diversity measures and
typology-classification}\label{tools-accessibility-methods-diversity-measures-and-typology-classification}

In examining how to enable improved quality of life through urban built
environment, accessibility measures have become an increasingly popular
tool. These measures are a way to quantify the ease of reaching
destinations from a given point in space and have been used to examine
urban areas through just and sustainable city planning agendas
\citep{vale2023accessibility, handyAccessibilityIdeaWhose2020}. The
15-Minute City is an amenity-provision neighbourhood planning concept
aimed at enabling the creation of urban environments that enhance life
quality, making it well-suited for analysis using accessibility
methodologies \citep{guzmanProximityEnoughCritical2024}. Recent works
have applied accessibility measures to investigate the 15-Minute City
across different geographic scopes. For instance, in Naples, Italy
\citep{gaglioneUrbanAccessibility15minute2022}, Barcelona, Spain
\citep{graellsCityCitiesMeasuring2021}, Vancouver, Canada
\citep{hosford15minuteCityReach2022}, and in urban areas across Europe
\citep{vale2023accessibility}. The ``cumulative opportunity'' measure
has been applied in many 15-Minute City examinations. This measure
quantifies how many destinations can be reached from a point in space
within a given travel time threshold, pairing well with examinations
that have a normative travel time threshold, like the 15-Minute City.
Furthermore, the cumulative opportunity measure is widely appreciated
for its intuitive computation and popularity among practitioners
\citep{handyAccessibilityIdeaWhose2020, handyMeasuringAccessibilityExploration1997, chengInvestigatingWalkingAccessibility2019}.
However, accessibility measures other than the cumulative opportunity
have also been applied, reflecting the diversity of measures in the
literature \citep{guzmanProximityEnoughCritical2024}.

Another method that compliments the assessment of the 15-Minute City
regarding the diversity of opportunity types is entropy. The entropy
measure, based on the Shannon-Wiener index of species diversity,
expresses relative evenness within a sequence
\citep{whittaker1972evolution}. In urban studies and planning
literature, the flexible concept of entropy has been used to
characterize land-use mix diversity
\citep{frankLinkingObjectivelyMeasured2005, ewingTravelBuiltEnvironment2010a},
particularly in the context of non-work trips
\citep{cerveroTravelDemand3Ds1997}, walking as a mode
\citep{luUrbanDensityDiversity2017, mavoaIdentifyingAppropriateLanduse2018},
and suburban sprawl \citep{randallGISBasedLand2015} as relevant to this
work. It has also been used to understand mobility behaviour
\citep{mcbride2020exploration, montero2023applying, montero2023role}.
However, work incorporating entropy indices alongside accessibility
analysis are not exceptionally common. There are examples of their use
as parameters within accessibility scores for employment opportunities
\citep{chengMeasuringUrbanJob2013, daiIncorporatingJobDiversity2018}
and, more recently, to describe the diversity in transit facilities
\citep{yinIncorporatingFacilityDiversity2024}. Within the
conceptualisation of the 15-Minute City, diversity of land-use is a key
feature. In this way, the entropy measure theoretically complements the
15-Minute City.

Classification algorithms also show promise as a tool in the assessment
of the 15-Minute City, as they have been useful in identifying mobility
and spatial typologies within the transportation planning literature.
Often, these algorithms take the form of machine learning approaches:
for instance, \citet{burkeGeospatialAnalysisFramework2022} uses
spatially constrained multivariate clustering, to develop urban form
typologies related to the 15-Minute City. The k-means algorithm is
another popular method: \citet{ganUnderstandingUrbanMobility2020}
applies it to analyse travel behaviour and classify metro stations based
on mobility patterns. Similarly,
\citep{montero2023role, montero2023applying} cluster travel behavior
profile typologies representing different population segments and place
of residence in the Barcelona Metropolitan Area. Combing classification
techniques is also common: for instance,
\citet{delmelleDifferentiatingPathwaysNeighborhood2017} uses
self-organizing maps (SOM) algorithm to group U.S. neighbourhoods by
minimizing dissimilarity in attributes before applying k-means for
further classification. Similarly, \citet{victorianoTimeSpaceMoney2020}
uses SOM to classify individuals' travel patterns according to the
dissimilarity in mobility attributes and uses the Decision Tree
algorithm to partition the data into interpretable classifications.

\subsection{Mobility of Care and feminist 15-minute
neighbourhoods}\label{mobility-of-care-and-feminist-15-minute-neighbourhoods}

Rather than focusing on \emph{all} destinations, it may valuable to
examine those related to caring activities. This re-framing offers a
feminist perspective on which urban functions matter, connecting well
with the 15-Minute City concept. Caring activities, which fulfill the
physical, psychological, and emotional needs of others, are among the
most essential and fundamental activities in society
\citep{ilo_care_2018}. Yet, they are the most unequal, undervalued, and
even devalued activities worldwide. Conventionally, caring activities
have been borne on women's shoulders
\citep{haydenGrandDomesticRevolution1982, hochschild_second_2012}.
According to \citet{ilo_care_2018}, women and girls perform more than
three-quarters of the total amount of unpaid care worldwide, a gender
gap that varies geographically \citep{RN5}. This unequal share of caring
responsibilities leads to multifaceted gendered differences: in career
development, profession selection, contract type, pay gap, and time
poverty, as recognized by various international organisations
\citep{EIGE2016, ilo_care_2018}. In terms of spatial and transportation
planning, almost one third of daily trips are for care purposes
\citep{sanchezdemadariagaMobilityCareIntroducing2013, sanchezdemadariagaMeasuringMobilitiesCare2019, ravensbergen2023exploratory}.
From this research motivation, Sánchez de Madariaga coined the term
Mobility of Care in 2013 to refer to all travel required to sustain the
needs of a household, such as grocery shopping, escorting children,
travelling to health appointments, and running errands
\citep{sanchezdemadariagaMobilityCareIntroducing2013}. While an
undercurrent of research had examined these unique household-serving
trips over the decades, her seminal work was the first to consider all
these trips as one category and demonstrate how mobility of care is a
significant proportion of daily travel.

The Mobility of Care concept also explicitly integrates inter-sectional
equity considerations, a common criticism leveled at the 15-Minute City
\citep{guzmanProximityEnoughCritical2024}. Perhaps unsurprisingly given
the gendered division of care work discussed, women have consistently
been found to complete more mobility of care trips than men. In one
study, mobility of care comprised 32\% of women's daily trips compared
to 28\% of men's. While this gap is significant, it was found to be far
greater in lower income households where women complete 20\% more care
trips than men \citep{ravensbergen2023exploratory}. Sánchez de Madariaga
not only shows how important these mobility of care trips are, but also
highlights the ways in which ``mobility of care is systematically
under-represented in any analysis of urban transport'' (p.~37). As a
product of the masculinst bias in planning, transport surveys and tools
often do not directly capture mobility of care, which re-enforces the
idea that these trips are not a significant part of daily mobility. In
this respect, the feminist perspective of the cities of proximity is
still underestimated with only few examples on the topic
\citep{gil_sola_choose_2022, macintyre_her_2022}.

Pairing well with the focus on shorter trips and the potential use of
sustainable modes within the 15-Minute City, Mobility of Care trips are
more local, shorter-distance and trip-chained. Compared to the trip to
work, mobility of care trips are more frequently completed by foot, and
less frequently by transit or bicycle
\citep{ravensbergen2023exploratory}. However, little work to date
examines walking to care destinations through the Mobility of Care
framework. Though there is ample literature that examines walking to
care destinations, such as schools (e.g., \citep{RN7, RN8, RN9}) and
grocery stores (e.g., \citep{RN10, RN11}), they tend to consider
singular care destinations in research focusing on walkability.

In summary, we reviewed the neighborhood planning framework from which
the 15-Minute City concept stems and is situated within. We then
reviewed tools common and useful in investigating the 15-Minute City.
Finally, we introduced the Mobility of Care concept, theoretically
linking it to the 15-Minute City to highlight the importance of care
destinations for all but especially along gendered and socio-economic
lines. The objectives of this work are two-fold, achieved through an
empirical examination of Hamilton, Canada, a mid-sized city:

\begin{itemize}
\tightlist
\item
  First, by theoretically bridging Mobility of Care with the 15-Minute
  City concept, we re-imagine what local amenities matter and calculate
  associated accessibility and entropy measure for care destinations.
\item
  Second, using a bottom-up data-driven approach, we apply machine
  learning methodologies to classification areas into \emph{15-minute
  caring neighbourhoods} and investigation their associated residential
  profiles along socio-economic lines.
\end{itemize}

\section{Data}\label{data}

\subsection{Case study context}\label{case-study-context}

This work focuses on Hamilton, Ontario, a mid-sized city on the shore of
Lake Ontario. Hamilton has a heterogenous land-use, with a populated and
dense urban core, surrounded by suburban development, which is itself
surrounded by rural communities. The Niagara Escarpment runs through
Hamilton, and results in a city with two key elevations: a more dense
lower city that contains the downtown core and the elevated suburban
development referred to as `the Mountain'. In this work, we analyse the
residential parcel centroids, 143,882 locations. We aggregated the
points at the level of Canadian Census Dissemination Area (DA) along
with the population and population per parcel plots in
Figure~\ref{fig-Fig1}.

\begin{figure}

\centering{

\includegraphics[width=6.25in,height=6.25in]{../figures/pop_descriptive_plots_1.png}

}

\caption{\label{fig-Fig1}The number of residential parcels per DA in
2020 (top), the population (middle) retrieved from the 2021 Canadian
Census, and the rate of population per parcel per DA (bottom). All
scales in quartiles. Basemap shapefiles are sourced from the Open Data
Hamilton Portal \citep{opendatahamiltonCityBoundary2023} and the USGS
\citep{greatlakesUSGS2010}.}

\end{figure}%

Hamilton also exhibits spatial disparities in social and economic
indicators; their spatial distribution is visualised in
Figure~\ref{fig-Fig2}. The densely populated inner city is characterised
by lower average incomes, and a higher prevalence of households living
under the low-income cutoff thresholds (LICO). Notably, the suburban
areas of the city tend to have a greater proportion of children and a
lower proportion of one parent households.

\begin{figure}

\centering{

\includegraphics[width=6.25in,height=6.25in]{../figures/pop_descriptive_plots_2.png}

}

\caption{\label{fig-Fig2}Socio-economic and demographic variables that
characterise accessibility to care destinations retrieved from the 2021
Canadian Census. All scales in quartiles. Basemap shapefiles are sourced
from the Open Data Hamilton Portal
\citep{opendatahamiltonCityBoundary2023} and the USGS
\citep{greatlakesUSGS2010}.}

\end{figure}%

\newpage

\subsection{Care destination dataset}\label{care-destination-dataset}

A spatial dataset of care destinations for Hamilton was compiled as
detailed in ( \emph{work removed for double-blind review} ). The dataset
includes 14 types of care destinations that were classified into four
categories: dependent-centric (e.g., the destinations for child- and
elder-centric escorting trips), grocery-centric, health-centric, and
errand-centric. Notably, these categories were generated following the
travel purpose categories created in the mobility of care research by
\citet{sanchezdemadariagaMeasuringMobilitiesCare2019}. Care categories,
sources of data, and preparation notes are detailed in
Table~\ref{tbl-Tbl1}. The spatial distribution of destination type are
visualised in Figure~\ref{fig-Fig3} by their category.

\global\setlength{\Oldarrayrulewidth}{\arrayrulewidth}

\global\setlength{\Oldtabcolsep}{\tabcolsep}

\setlength{\tabcolsep}{2pt}

\renewcommand*{\arraystretch}{1}

\providecommand{\ascline}[3]{\noalign{\global\arrayrulewidth #1}\arrayrulecolor[HTML]{#2}\cline{#3}}

\begin{longtable}[c]{|p{0.67in}|p{3.94in}|p{2.09in}}

\caption{\label{tbl-Tbl1}Description of care destinations categories,
notes on data preparation and associated data sources.}

\tabularnewline

\ascline{1.5pt}{666666}{1-3}

\multicolumn{1}{>{\raggedright}m{\dimexpr 0.67in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{Care\ category}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 3.94in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{Care\ destination\ types}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 2.09in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{Sources}}} \\

\ascline{1.5pt}{666666}{1-3}\endfirsthead

\ascline{1.5pt}{666666}{1-3}

\multicolumn{1}{>{\raggedright}m{\dimexpr 0.67in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{Care\ category}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 3.94in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{Care\ destination\ types}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 2.09in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{Sources}}} \\

\ascline{1.5pt}{666666}{1-3}\endhead

\multicolumn{1}{>{\raggedright}m{\dimexpr 0.67in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Depedent-centric}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 3.94in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Schools,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{daycares,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{and}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{community}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{centres,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{recreation}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{centres,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{parks,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{senior}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{centres,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{long-term}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{care}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{homes,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{and}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{retirement}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{homes:}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1,265}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{locations}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{are}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{included.}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 2.09in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(Hamilton}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2022a,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2022c,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2022d;}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Ontario}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023;}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Ontario GeoHub}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023)}}} \\

\multicolumn{1}{>{\raggedright}m{\dimexpr 0.67in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Grocery-centric}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 3.94in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Convenience}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{stores}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{and}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{grocery}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{stores}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(e.g.,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{large}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{retailers}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{as}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{well}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{as}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{speciality}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{food}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{grocers,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{health}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{food}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{grocers):}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{381}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{destinations}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{are}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{included.}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 2.09in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(Axle Data}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023)}}} \\

\multicolumn{1}{>{\raggedright}m{\dimexpr 0.67in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Health-centric}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 3.94in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Hospitals,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{pharmacies,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{clinics,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{and}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{dentist}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{offices:}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{421}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{destinations}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{are}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{identified.}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 2.09in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(Ontario GeoHub}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023;}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{HNHB Healthline}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023)}}} \\

\multicolumn{1}{>{\raggedright}m{\dimexpr 0.67in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Errand-centric}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 3.94in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Libraries,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{post}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{offices,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{and}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{banks:}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{158}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{destinations}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{are}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{identified.}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 2.09in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(Hamilton}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2022b;}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Axle Data}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023;}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Canada Post}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023;}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{BMO}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023;}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{HSBC}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023;}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{National Bank}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023;}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{RBC}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023;}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Scotiabank}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023;}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{TD Bank}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023)}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{.}}} \\

\ascline{1.5pt}{666666}{1-3}

\end{longtable}

\arrayrulecolor[HTML]{000000}

\global\setlength{\arrayrulewidth}{\Oldarrayrulewidth}

\global\setlength{\tabcolsep}{\Oldtabcolsep}

\renewcommand*{\arraystretch}{1}

\begin{figure}

\centering{

\includegraphics[width=6.25in,height=6.25in]{../figures/care_categories_plot.png}

}

\caption{\label{fig-Fig3}The locations of care destinations in Hamilton
separated by the author-generated categories of: dependent-, errand-,
grocery- and health- centric care categories. Basemap shapefiles are
sourced from the Open Data Hamilton Portal
\citep{opendatahamiltonCityBoundary2023} and the USGS
\citep{greatlakesUSGS2010}.}

\end{figure}%

\newpage

\subsection{Travel time to care destinations
estimations}\label{travel-time-to-care-destinations-estimations}

Travel behaviour to care-oriented destinations is often uncounted and
hence comprehensive travel times to all destination types in
Table~\ref{tbl-Tbl1} is unavailable. To overcome this gap, travel time
from residential parcel locations in Hamilton and care destinations in
Figure~\ref{fig-Fig3} are approximated assuming travel by foot at an
average speed (3.6 km/hr). This estimation is done using the
`travel\_time\_matrix()' function from the \{r5r\} package
\citep{pereiraR5rRapidRealistic2021} using R version 4.3.2
\citep{coreR2023}. The inputs into the function are: the locations of
143,882 residential parcels as origins, the 2,225 care locations as
destinations, and a OpenStreetMap road network including walking
infrastructure \citep{geofabrikOntarioCanadaOpen2023}. In line with the
15-Minute City, a maximum walking travel time of 15 minutes is specified
and an origin-destination travel time matrix of the shortest travel time
from origin to destination is calculated. The resulting matrix contains
2,014,502 rows, representing walking travel times from each parcel to
reachable care destinations within 15 minutes.

\section{Methods}\label{methods}

The following sub-sections detail the methods to classify Hamilton into
spatial degrees of \emph{caring 15-minute neighbourhood}. First,
accessibility to each of the 14 destination types from each of the
143,882 residential parcel locations is described. Second, the entropy
measures used to calculate the diversity of accessibility to each of the
care categories is detailed. Third, we describe how accessibility and
diversity values for each parcel are input into a Self-Organizing Map
(SOM) algorithm, and how the resulting clusters are analyzed using a
Decision Tree algorithm to narrate residential profiles based on
socio-economic variables. In sum, this methodology presents a
data-driven approach to examine what neighbourhoods in a city have the
potential to provide 15-minute caring access, at what level of intensity
and completeness, and the socio-economic characteristics of those who
are most benefited or burdened.

\subsection{Accessibility: the cumulative opportunity
measure}\label{accessibility-the-cumulative-opportunity-measure}

To capture the quantity of access to each type of destinations, a
cumulative opportunity accessibility score \(S^t_i\) is calculated.
Scores for each of the 14 care destination types \(t\) is calculated for
every parcel \(i\). The calculation takes the following mathematical
form:

\[
S^t_i=\sum_{j}O_j^{t}\cdot f(c_{ij})
\] \noindent Where:

\begin{itemize}
\tightlist
\item
  \(i\) is a set of parcel point origin locations.
\item
  \(j\) is a set of care destination locations of type \(t\).
\item
  \(O^t_j\) is a number of opportunities of category type \(t\) at
  \(j\).
\item
  \(c_{ij}\) is the travel cost between \(i\) and \(j\).
\item
  \(f(\cdot)\) is an impedance function of \(c_{ij}\); within the
  cumulative opportunity approach, it is a binary function that takes
  the value of 1 if \(c_{ij}\) is less than a selected value
  \citep{handyMeasuringAccessibilityExploration1997}.
\item
  \(S_{i}^t\) is the cumulative opportunity accessibility score, the sum
  of weighted opportunities reachable within \(f(\cdot)\), at each \(i\)
  for each \(t\).
\end{itemize}

\subsection{Diversity in opportunity accessibility: the entropy
measure}\label{diversity-in-opportunity-accessibility-the-entropy-measure}

The entropy measure, as defined in \citet{cerveroTravelDemand3Ds1997},
is used to represent the diversity of care destination accessibility.
For each parcel, a value between 0 to 1 is calculated, where 1
represents total evenness in the number of care opportunities in each
category that can be reached. The mathematical formulation takes the
following form:

\[
D_{i} = \frac{-\sum_{t} [S_i^{t} / \sum_{t}S^{t}_{i} \times ln(S_i^{t} / \sum_{t}S^{t}_{i} )]}{ln(n_t)}
\]

\noindent Where:

\begin{itemize}
\tightlist
\item
  \(i\) is a set of parcel point origin locations.
\item
  \(t\) is a set of care destination types (e.g., school, grocery, park,
  etc.)
\item
  \(n_t\) is the count of care destination types \(t\). In this work,
  this value is 14.
\item
  \(S_{i}^t\) is the cumulative opportunity accessibility score, the sum
  of weighted opportunities reachable within a 15-minute walk from
  \(i\).
\item
  \(D_i\) is the diversity score.
\end{itemize}

Notably, \(D_i\) represents evenness in the type of care categories a
parcel can access. For example, if a parcel has an access score of
\(S_i^t = 0.5\) for all destination types, it will receive a diversity
score of \(D_i = 1\), just as if it had an access score of
\(S_i^t = 10\) for each destination. Conversely, a parcel may be
assigned a low \(D_i\) score if its accessibility scores differ across
categories, regardless of whether those scores are low or high.

\subsection{Machine learning classification: SOM and Decision
Trees}\label{machine-learning-classification-som-and-decision-trees}

We use two machine learning techniques in this work. First, SOM is an
unsupervised technique implemented to reduce the data dimensionality and
create interpretable clusters related to the intensity and completeness
of caring access. This is done by imputing each parcel as an observation
with its associated accessibility and diversity attributes, and the data
being rearranged onto a two-dimensional space based on its minimizing
dissimilarity in its neighbourhood. An appropriate number of
superclusters are selected and assigned labels based on the quantity and
diversity of care access provided, i.e., the degree by each the parcel
is located in a \emph{caring 15-minute neighbourhood}. Then, a Decision
Tree is run to characterise the socio-economic profiles of who resides
in neighbourhoods associated with the superclusters. Together, this
combined approach leverages the unsupervised data-driven classification
power of the SOM with the interpretation of Decision Trees. The
procedure used in this work is similar to that used in
\citet{victorianoTimeSpaceMoney2020}. In this work, rather than each
observation representing an individual's daily mobility behaviour (with
associated variables), each observation represents a parcel location
with calculated care accessibility and accessibility diversity scores.

For the SOM step, the function `trainSOM()' from \{SOMbrero\} R package
is used \citep{VillaVialaneix2017}. The input variables include the
143,882 parcels, each as individual observation along with 15 variables:
the 14 calculated accessibility scores \(S_i^t\), normalized to the
min-max range score within each \(t\), and one diversity value \(D_i\).
Otherwise, defaults for all other parameters are assumed, relying on the
data-driven heuristics embedded in the `trainSOM()' function.
Consequently, a 100 node (10 by 10 grid) SOM structure using euclidean
distance and square topology is produced. Simply put, the SOM algorithm
proceeds as follows: a 2D grid of nodes is created as specified by the
analyst, where a node represents a point in the reduced-dimensional
space. Upon initialization, each node is assigned a random weight vector
of the same dimension as the input data (in our case, 15). From the
input data, a random observation with its associated weight vector
(i.e., one parcel point with 15 variables) is selected and the Euclidean
distance between its weight vector and all nodes in the grid are
calculated. The node with the smallest distance (i.e., the smallest
dissimilarity) is labelled the `best matching unit' as it is the node
that best represents the input observation. After this best matching
unit is identified, its own weight and its neighbouring nodes are
updated to become more like the input observation. The process of
finding best matching units and updating their weights is repeated for
every observation, multiple times, until the results converge. As
mentioned, this competitive learning process produces a 100 node SOM
structure where each observation (parcel) is assigned to 1 node with an
associated dissimilarity index. The SOM output is typically examined
through a dissimilarity dendogram and an associated dissimilarity
variance explained plot to select an appropriately representative
`superclusters'
\citep{VillaVialaneix2017, victorianoTimeSpaceMoney2020}.

For the Decision Tree step, the supercluster-classified parcels
identified in the SOM step are used as \emph{labels} and socio-economic
and demographic indicators related to the Mobility of Care literature
are used as \emph{features}. Features are retrieved from the 2021
Canadian Census \citep{governmentofcanadaCensusPopulation2023}. This
step creates residential profiles of the superclusters, allowing us to
explore the characteristics of residents within the superclusters in a
data-driven way. To estimate the Decision Tree, the `rpart()' function
in the \{rpart\} R package is used \citep{Therneau2023}; default
parameters for classification splitting along with each value being
weighted by the population present in the associated DA is assumed. The
Decision Tree algorithm is a supervised learning technique that begins
by splitting a subset of the input data into branches based on a
selected feature with the lowest impurity score (i.e., the least mixing
of label membership within a branch). This process is recursively
repeated for each data subset, selecting the next best feature.
Ultimately, the data is classified into distinct classes, with class
membership explained by traversing the branches defined by the features
that characterise the partitions within the Decision Tree. Notably, the
absence of features from the Decision Tree does not necessarily imply
they are irrelevant for classification, rather, they are less relevant
than other features. Put another way, when considering features that are
highly correlated, such as income level and LICO, not all relevant
variables may be present in the tree
\citep{victorianoTimeSpaceMoney2020}.

\section{Results}\label{results}

\subsection{Overview of access to care
destinations}\label{overview-of-access-to-care-destinations}

How many care destinations can be reached within a 15-minute walk is
summed for each of the four care categories defined in
Table~\ref{tbl-Tbl1}. The median parcel value for each DA is visualised
in Figure~\ref{fig-Fig4} and the median parcel diversity of care
destinations accessible is presented in Figure~\ref{fig-Fig5}. Together,
Figure~\ref{fig-Fig4} and Figure~\ref{fig-Fig5} represent summaries of
15 variables that served as inputs into the SOM algorithm.

\begin{figure}

\centering{

\includegraphics[width=6.25in,height=6.25in]{../figures/percats_copp_plot.png}

}

\caption{\label{fig-Fig4}The number of care destinations that can be
reached within a 15-min walk per care category for a median parcel in
each DA. The values are a summary of the 14 acessibility scores that
served as inputs into the SOM. Basemap shapefiles are retrieved from the
2021 Canadian census \citep{governmentofcanadaCensusPopulation2023}, the
Open Data Hamilton Portal \citep{opendatahamiltonCityBoundary2023} and
the USGS \citep{greatlakesUSGS2010}.}

\end{figure}%

\begin{figure}

\centering{

\includegraphics[width=4.6875in,height=4.6875in]{../figures/all_cats_entropy_copp_plot.png}

}

\caption{\label{fig-Fig5}The diversity measure based on the proportion
of care category accessibility in Figure 4. These values are also a
summary of the 15th input into the SOM. Basemap shapefiles are retrieved
from the 2021 Canadian census
\citep{governmentofcanadaCensusPopulation2023}, the Open Data Hamilton
Portal \citep{opendatahamiltonCityBoundary2023} and the USGS
\citep{greatlakesUSGS2010}.}

\end{figure}%

\newpage

\subsection{Identification of completely caring 15-minute neighbourhood
typologies}\label{identification-of-completely-caring-15-minute-neighbourhood-typologies}

In Figure~\ref{fig-Fig4}, the concentration of opportunities for all
category types is centered in the downtown of Hamilton, with the highest
concentration being in the lower city (near the lake shore) and pockets
of the high concentration further from the shoreline. Grocery-centric
destinations appear to be the most concentrated, followed by
health-centric and dependent-centric caring destinations. Errand-centric
destinations are the most sprawled. In many ways, Figure~\ref{fig-Fig4}
mirrors the spatial distribution of care destinations
(Figure~\ref{fig-Fig3}) as the range of 15-minutes of travel is small.
In Figure~\ref{fig-Fig5}, areas that have high care accessibility tend
to have high diversity as well, though there are exceptions in pockets
of the city outside the downtown core that have higher diversity but low
levels of accessibility for all care categories. Similarly, there are
areas with low diversity within the downtown core that have only
moderate or high accessibility to certain care destination types.

Based on the SOM methodology discussed, 7 superclusters are identified
from the 10 by 10 grid of SOM nodes. As diagnostics for the selection of
the number of superclusters, the dissimilarity-index-based dendrogram
with the proportion of parcels represented in each supercluster
alongside the variance explained plot is visualised in
Figure~\ref{fig-Fig6}. Labels representing grades A to D qualifying the
7 superclusters are then assigned. These grades as also reflected in
Figure~\ref{fig-Fig6} and represent the quantity of accessibility and
diversity in accessibility per care category. Higher grades (A+ and A)
corresponding to the highest accessibility and diversity, while lowest
grades (D) representing lowest accessibility and diversity scores.

\begin{figure}

\centering{

\includegraphics[width=1\textwidth,height=1\textheight]{../figures/all_access_sc_plot_small.png}

}

\caption{\label{fig-Fig6}The 7 resulting superclusters from the SOM the
10 by 10 grid output represented in a dendrogram (left) and the
proportion of variance unexplained by cluster (right).}

\end{figure}%

\newpage

The grade for each supercluster was carefully assigned by the authors by
referring to descriptive statistics and their notable trends. These
statistics are summarised in Table~\ref{tbl-Tbl2} and in the boxplots in
Figure~\ref{fig-Fig7}.

\begin{itemize}
\tightlist
\item
  A+ and A have exceptionally high caring accessibility for all
  destination types and high diversity scores in the top quantile or
  above. Together, these grades represent 24\% of all parcels in the
  city.
\item
  A- has high caring accessibility, but low diversity. Interestingly, A-
  is like A but with much higher dependent-centric destinations,
  particularly parks, schools, and daycares, with only moderately high
  scores for all other destination types. This disbalance lowers its
  overall diversity score. However, it can be characterized as a
  supercluster that demonstrates potential in being retrofitted to
  provide A+ or A level of completely caring access. 4\% of all parcels
  are represented by the A- grade.
\item
  B+ and B superclusters present about average completely caring access.
  These superclusters can serve as the benchmark for what `average'
  completely caring access in Hamilton currently looks like. These
  grades represent 15\% of all parcels.
\item
  B- provides above average dependent-centric destination access,
  particularly parks, daycares and schools, but below average access to
  other destination types and hence has low diversity scores. B- is like
  A- as it demonstrates complete caring 15-minute potential if
  retrofitted. As these parcels demonstrate caring access to some
  destinations, they may have the potential to be retrofitted to support
  complete access to all caring destination types. This grade represents
  19\% of all parcels.
\item
  D superclusters demonstrate the lowest scores all-around, representing
  room for land-use improvement that addresses complete and caring
  15-minute access. This supercluster characterizes the largest number
  of parcels, representing 38\% of parcels in the city.
\end{itemize}

\global\setlength{\Oldarrayrulewidth}{\arrayrulewidth}

\global\setlength{\Oldtabcolsep}{\tabcolsep}

\setlength{\tabcolsep}{2pt}

\renewcommand*{\arraystretch}{1}

\providecommand{\ascline}[3]{\noalign{\global\arrayrulewidth #1}\arrayrulecolor[HTML]{#2}\cline{#3}}

\begin{longtable}[c]{|p{1.20in}|p{0.60in}|p{0.60in}|p{0.60in}|p{0.60in}|p{0.60in}|p{0.60in}|p{0.60in}|p{0.60in}}

\caption{\label{tbl-Tbl2}Mean and (standard deviation) of each SOM
classified cluster by input variable and additional summary variables.
Variables included in the SOM algorthim are in regular case, while
additional summary variables are indicated by ALL CAPITAL LETERS.}

\tabularnewline

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}-}

\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{\ }}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{A+}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{A}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{A-}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{B+}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{B}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{B-}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{D}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{TOTAL}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}-}\endfirsthead

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}-}

\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{\ }}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{A+}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{A}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{A-}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{B+}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{B}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{B-}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{D}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{TOTAL}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}-}\endhead

\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{GROCERY}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{TOTALS}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{12.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(5.7)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{7}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(3.1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{4.8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.7}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.2)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3.6}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(5.1)}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Convenience}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Store}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{4.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.9)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.9)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(3.4)}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Grocer}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{4.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.8)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.6}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.7}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.9)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.4)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2)}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{DEP.}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{TOTALS}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{17.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(5.8)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{11.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(4.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{13.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{6.7}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{5.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{9.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.1)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{8.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(6.2)}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Comm.}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{or}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Rec.}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{centre}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.7}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.9)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.8)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.3)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.8)}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Daycare}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{or}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{EarlyON}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{4.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.6}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.6}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.8)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.2)}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{LTC}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{or}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{retirment}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{home}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.8)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.7)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.4)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.8)}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Park}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{6.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{4.8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{5.6}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.7}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.3)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3.6}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.6)}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{School}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.8)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.7}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.7)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.9)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.5)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.3)}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Senior}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{centre}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.9)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.7)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.2)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.5)}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{HEALTH}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{TOTALS}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{11.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(6.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{5.8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.8)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3.6}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.8)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.8)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.9)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.6}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(4.7)}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Dentist}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{4.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(3.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.7)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.4)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.2)}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Hospital}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{or}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{clinic}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.6}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.7)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.7)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.2)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.5)}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Pharmacy}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{6.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(3.7)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.7}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.8)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.7)}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{ERRAND}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{TOTALS}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.4)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.8)}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Bank}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{or}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{ATM}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.8)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.7}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.7)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.2)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.7}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.2)}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Library}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.2)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.4)}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Post}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{office}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.8)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.7}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.7)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.8)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.6}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.7)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.2)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.7)}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1pt}-}

\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Category}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{diversity}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.6}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.4)}}} \\

\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}

\hhline{>{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}->{\arrayrulecolor[HTML]{666666}\global\arrayrulewidth=1.5pt}-}

\end{longtable}

\arrayrulecolor[HTML]{000000}

\global\setlength{\arrayrulewidth}{\Oldarrayrulewidth}

\global\setlength{\tabcolsep}{\Oldtabcolsep}

\renewcommand*{\arraystretch}{1}

\begin{figure}

\centering{

\includegraphics[width=8.33333in,height=3.125in]{../figures/box-plot.png}

}

\caption{\label{fig-Fig7}A boxplot demonstrating summary variables and
category diversity that define the superclusters.}

\end{figure}%

\newpage

To spatially demonstrate where the superclusters are located,
representative supercluster labels for each DA are visualised in
Figure~\ref{fig-Fig8}. This visualisation is created by grouping parcels
by their DA, calculating the median supercluster label for each label,
and selecting the label that is most dominant within that DA. For
reference, the median number of parcels in each DA is 150 and
supercluster label membership within a DA is typically pure, with the
median membership being 80\% of a single cluster.

\begin{figure}

\centering{

\includegraphics[width=4.6875in,height=4.6875in]{../figures/SClabel_plot.png}

}

\caption{\label{fig-Fig8}The maximum median parcel supercluster
membership per DA. Escarpment is visualised as a grey dashed line.
Basemap shapefiles are retrieved from the 2021 Canadian census
\citep{governmentofcanadaCensusPopulation2023}, the Open Data Hamilton
Portal \citep{opendatahamiltonCityBoundary2023} and the USGS
\citep{greatlakesUSGS2010}.}

\end{figure}%

The resulting superclusters visualised in Figure~\ref{fig-Fig8} are a
combination of the cumulative opportunity and diversity of access
measures translated into meaningful typologies through the SOM
methodology. To reiterate, accessibility is the measure of
\emph{potential} interaction, e.g., how many care destinations one could
reach within a 15-minute walk, these typologies are useful in
identifying which areas of the city are providing regionally high levels
of complete caring access, which areas are regionally average, and which
areas are regionally below average. Furthermore, neighbourhoods which
provide disbalanced access can be further investigated.

In Figure~\ref{fig-Fig8}, it can be observed that the excellent and
exceptional neighbourhoods (A, A+), which are very caring and very
complete 15-minute neighbourhoods, are located within the center of the
city and closest to the shoreline in the downtown core of Hamilton. The
above and below average neighbourhoods (B+ B) are often proximate to the
excellent and exceptional neighbourhoods and are more prevalent in south
of the escarpment within the center Hamilton. The escarpment is a
physical barrier, with few pedestrian-accessible access points to
traverse; hence the typologies describing neighbouring DAs separated by
the escarpment are often different. Below average (D) neighbourhoods are
located in peripheral areas outside the center of Hamilton, in areas
where urban form is characterised by lower density residential housing
and auto-dependent mobility.

Furthermore in Figure~\ref{fig-Fig8}, A- and B- grades stand out as
offering high access for children-centric destinations for their
grade-group but below grade-group average access in other destination
types. These neighbourhoods may be more suitable for populations who
prioritize walkable access to destinations like schools, parks and
daycares, and find access to other types of caring destinations less
important. These neighbourhoods also stand out as demonstrating
potential to be retrofitted to provide more complete access if
additional destination types were located within their neighbourhoods.

\subsection{Profiles of who does and does not reside in caring 15-minute
neighbourhoods}\label{profiles-of-who-does-and-does-not-reside-in-caring-15-minute-neighbourhoods}

To enhance the meaning of the superclusters beyond a descriptive and
spatial conceptualisation of ``caring'' and ``complete'', \emph{who?}
resides in what neighbourhood is investigated through the Decision Tree
results. The input features of the Decision Tree are the supercluster
labels and the feature variables are various population-weighted
socio-demographic characteristics of the 2021 Canadian Census, namely:
median household income, \% below the median household income, \% LICO
prevalence, average number of children per household, \% population aged
0 to 14, \% not in the labour force, \% not employed, Gini index on
adjusted household after-tax income, \% visible minority, \% single
parent household, \% who walk to work (relative to bike, care/truck/van,
public transit and `other'), \% of owner household in core housing need
(i.e., inadequate housing structure or paying higher than 30\% income on
housing), \% of tenant households in subsidized housing, \% of tenant
households in core housing need, \% no certification or with only a
highschool diploma.

Of all the included input variables, \emph{median household income}
proved to be the most meaningful in partitioning the superclusters data.
Figure~\ref{fig-Fig9} provides the Decision Tree with the significant
splits in median household income and the proportion composition of
supercluster along each branch for three terminal Decision Tree nodes.
While the algorithm is unable to homogenously use each supercluster,
Figure~\ref{fig-Fig9} is helpful to report a narrative of who may reside
in what caring/complete supercluster. Particularly, the Decision Tree
demonstrates a more pure supercluster membership for only two of the
three terminal Decision Tree nodes: supercluster A+ (exceptional
completely caring access) and D (not caring and complete access). For
reference, the city-wide mean household income is 81,316 (SD: 25,239 and
median: 80,000). The notable splits for income are \textgreater\$91,500,
\textgreater\$68,750, or between \$68,750 and \$91,500; roughly split by
lower income, middle income, and higher income median household
brackets.

\begin{figure}

\centering{

\includegraphics[width=6.25in,height=4.6875in]{../figures/access_sc_profiles_justincome.png}

}

\caption{\label{fig-Fig9}The Decision Tree demonstrating median
household income splits and superclusters composition within each
branch.}

\end{figure}%

Describing Figure~\ref{fig-Fig9}, the bar column on the right represents
DAs with higher household income. In these DAs, the majority of parcels
are labeled as D superclusters, i.e not caring or complete 15-minute
neighbourhoods. Within this higher-income-representing bar column, the
second highest proportion of supercluster membership are B- parcels. B-
parcels are low in most amenity types except a few child-centric
destinations and show promise in being more easily improved than the D
parcels. The higher income column represents 39\% of the population, the
largest proportion of any of the three columns. The left column in
Figure~\ref{fig-Fig9} corresponds to the lowest income, i.e., below
68,750 CAD median household income, and accounts for 32\% the
population. It is dominated by A+ parcels along with A and A- parcels,
in the largest quantities relative to other bar columns, however,
proportions of all other superclusters are also present. The middle
column is defined by parcels with a mix of supercluster classifications
and represents DAs with middle households incomes, between 68,750 CAD
and 91,500 CAD representing 28\% of the population.

Though median income was the most useful in partitioning the parcels by
their supercluster labels in Figure~\ref{fig-Fig9}, other variables are
still important. Namely, the following variables in order of importance
and correlation coefficient with the median household income variable in
brackets, are listed: the \% below the median household income (-0.89),
\% single parent household (-0.57), \% no certification or with only a
highschool diploma (-0.49), the average number of children per household
(0.38), \% not in the labour force (-0.41), \% LICO prevalence (-0.67),
and \% who walk to work (-0.37).

It is notable that median household income is highly or moderately
correlated with many of these important variables, and discussing these
correlations is useful for interpretation alongside the Decision Tree
diagram. For instance, single-parent families are most positively
correlated with the proportion of households being in the bottom income
distribution (-0.89) and lower or no diploma (0.6), and most negatively
with the median after-tax household income (-0.57), reflecting
nationwide trends
\citep{governmentofcanadaDailyMainHighlights2024, governmentofcanadaPrevalenceLowIncome2024}.
Though not all single-family households are below the LICO (i.e.,
single-family households are not highly correlated with LICO at the
DA-level in Hamilton), DAs with higher concentrations of single-family
households and LICO prevalence tend to have A+ access. Conversely, DAs
with lower single-parent households are in DAs with a higher
concentration of parcels with D access. These findings are notable from
an equity perspective since economic disadvantaged often intersects with
other socio-demographic characteristics.

\section{Discussion}\label{discussion}

Geographically, Hamilton is city that offers very caring - and
completely caring - 15-Minute neighbourhoods for some, but not all.
There are evident spatial inequalities, with areas ranging from
excellent (A+ and A), average (B+ and B), to well below average (D).
Some areas, like those labelled A- and B-, show potential for
improvement within their grade groups. While the downtown core has the
highest concentration of \emph{caring 15-minute neighborhoods}, certain
areas outside the core do as well. This finding somewhat contrasts
employment accessibility studies, such as
\citet{elgeneidyNonstopEquityAssessing2016}'s assessment of public
transit access to employment in the Greater Toronto Area (including
Hamilton), where access is heavily concentrated in the downtown core,
more so than appears to be present in this work. Though our paper uses a
different methodology, both studies highlight distinct patterns: indeed,
A+, A or A- care access is concentrated in the downtown core but it is
also present in certain pockets outside of the center of Hamilton,
leading to interesting conjectures. For instance, some of the A-grade
neighborhoods in Hamilton likely follow land-use principles that
emphasized residential proximity to amenities, similar to the NUC
reviewed. The potential overlap of the NUC with caring 15-minute
neighborhoods warrants further investigation.

Who currently resides in Hamilton's completely caring 15-Minute
neighbhorhoods is also demonstrated to be a somewhat optimistic story.
Parcels that provide A+ completely caring access tend to be in DAs that
are economically disadvantaged. Economic disadvantage tends to intersect
with other identities such as gender
\citep{lightmanMeasuringEconomicExclusion2018}. And as reviewed in this
work, all women and especially those from lower income households tend
to complete most mobility of care trips
\citep{ravensbergen2023exploratory}. Furthermore, lower-income
households tend to also be single-parent households. Broadly,
single-parent households are more likely to be time-disadvantaged
\citep{nieuwenhuisSingleparentFamiliesWork2018}, and tasked with a
higher proportion of care duties \citep{craigTimeCareComparison2004}. In
this way, the most economically disadvantaged groups having A+ complete
and caring access is an optimistic finding. However, Hamilton is
experiencing gentrification \citep{ellisyoungWeReJust2018}; rents are
rising along the future light rail transit corridor and throughout the
city
\citep{vandermerweSpilloverGentrificationMidsized2021, mayersLightTransitDocumenting2023}.
Toronto, Hamilton's larger and higher-rent neighbouring city, is
spilling gentrification into Hamilton's downtown core, (re)producing
neighbourhoods based on Toronto's middle class identities
\citep{mayersLightTransitDocumenting2023}. In these ways, the lower
income residents of Hamilton's A+ neighbourhoods are more likely to be
displaced, which is matter of wellbeing and justice. There are now few
low-rent choices that provide the same exceptional level of access as
the downtown core, hence lower household income residents that currently
reside in A+ neighhourhoods will likely displaced in the coming years if
current trends continue.

In discussing policy interventions that equitably increase completely
caring 15-Minute neighbourhoods in Hamilton, this work's presents a
methodology to create city-wide relative typologies and investigate who
currently resides in what neighbourhood, as a stepping stone for further
investigation. Neighbourhoods with the lowest grades (D) and with the
highest potential in being improved (i.e., high accessibiltiy for
certain types of destinations but not all) are neighbourhoods with B-
and A- grades. However, the work's findings demonstrate a mismatch in
household income and completely caring access: higher income areas are
often the ones targeted for improvement. This raises important questions
from the land-use policy perspective that need to be further analysed.
For instance, is it equitable to focus policy on ameliorating
neighbourhoods that are already higher-rent though they tend to be more
rural, single-use zoned and car-dependent (parcels with D grades)?
Further, of the parcels that provide high child-centrics but low
otherwise, A- parcels (better access) tend to be in DAs with
lower-income households more so than B- parcels (lower access than A-).
From the perspective of ameliorating land-use to support the equitable
distribution of 15-minute caring neighbourhoods which areas be targeted?
Who should benefit, and how? If the policy initiative is targeted to
specific neighbourhoods: sustainability linked to car dependency and
equity are in tension. This harkens to what role a planned
neighbourhood, and bottom-up planning approaches that include the
evaluation of travel behaviour by socio-economic and demographic profile
should be adopted to achieve equitable 15-minute cities.

As is the case for all research, the results should also be interpreted
along with methodological assumptions. This work measures spatial
accessibility which is a measure of potential interaction with all
reachable destinations from an origin. These destinations, however, may
not be relevant to people at an origin, e.g., they may be underutilized
(walkable access to 2 schools may be underused, as children often only
attend 1 school), the trip physically undesirable (the walk may be along
an arterial with high traffic speeds, making the trip unlikely to ever
happen by foot), or the average walking speed assumed may not reflect
the walking speeds of all populations
\citep{willberg15minuteCityAll2023}. Further, people who reside in these
neighbourhoods may also disagree with the neighbourhood's label. The
grade labels are region-relative (e.g., high accessibility in Hamilton
may be subjectively insufficient for some) and they do not consider
subjective perceptions that influence accessibility (e.g., though a
neighbourhood has many opportunities, it may not feel safe to access
them). Furthermore, accessibility is calculated from the point of
residential parcels. Care trips are not necessarily completed from home,
in-fact, they are often completed in complex trip-chains
\citep{scheinerWomenComplexDaily2015}. In this way, the results flatten
the dynamic patterns of care trips. These methodological assumptions
should all be taken together when interpreting the results. In this way,
the methodologyand findings presented identifies spatial and
socio-economic trends that should be further investigated.

\section{Conclusion}\label{conclusion}

This study makes three types of contributions to the transportation and
city planning literature: empirical, methodological and theoretical. At
the empirical level, areas of the mid-sized City of Hamilton have been
typified by their degree of `15-Minute Caring Neighbourhood' potential.
Methodologically, we applied the longstanding accessibility measure of
cumulative opportunities and entropy to classify spatial areas based on
how many destinations could be reached in a 15-minute walk and the
diversity of destination type. These values were then clustered using a
novel machine learning approach, SOM, to generate meaningful typologies
for discussion and further comparison with socio-economic composition of
the area. We find A+ and A 15-minute caring neighbourhoods are located
within the downtown core and in certain suburban pockets of the city,
while the peripheral regions provide D level caring access. A- and B-
areas are also identified as neighbourhoods that already support a high
amount of children-centric destination access, and could be improved to
provide better complete care access. Theoretically, our work puts forth
an explicitly caring 15-Minute Neighbourhod conceptualisation, bridging
the Mobility of Care and the 15-Minute City concepts. We discuss how
measuring caring neighbourhoods can be explicitly considered within city
planning.

Our work is of relevance to researchers and practitioners planning
equitable and sustainable cities. Instead of prescribing an urban form
design principle, such as ``all local amenities should be within a
15-minute walking distance'', it instead examines an empirical example
to determine which areas in the city have the \emph{potential} to be
15-minute neighbourhood based on the existing spatial accessibility
offered by the urban environment and walking transport network. To this
end, this data-driven methodology introduces a way to identify
neighbourhoods that have potential, almost have potential, and are far
from containing this potential to support future context-specific
qualitative work.

\section{References}\label{references}

\renewcommand{\bibsection}{}
\bibliography{bibliography.bib}

\end{document}

### 2024-10-10 paper/rejected-text.qmd
Draft to submit to journal. And some repo clean-up.

@@
 -->
+
+As related to Mobility of Care, care trips and hence care destinations are of key focus.
+While using entropy to measure the diversity of care destinations within 15-minute neighborhoods has yet to be explored, it theoretically complements the 15-Minute City concept.
+
+The entropy measure has been widely used to characterise the diversity of land-use mix in early work by [@frankLinkingObjectivelyMeasured2005] based on the species diversity Shannon-Wiener index that expresses relative evenness throughout a sequence (see review in @whittaker1972evolution).
+It has also been used in social sciences to comprehend the occurrence of social events in a structured manner specially with regard to life trajectories [@ritschardSequenceAnalysisRelated2018]
+
+A variety of tools have been applied in the literature to examine the 15-Minute City.
+However, the question of _what destinations matter_  still demands further examination.
+The following sub-section bridges the concept of the 15-Minute City with another normative framework: Mobility of Care.
 
@@
 -->
+
+<!-- As a tool to visualise the variables that describe the superclusters, a Decision Tree is fit with the DEPENDENT TOTALS, ERRAND TOTALS, GROCERY TOTALS, HEALTH TOTALS, and CATEGORY DIVERSITY scores.
+Aligning with the descriptive summaries of the superclusters provided, describing the tree-diagram from left to right (roughly from the most caring & complete, to the least), the first three terminal nodes (represented by bar columns that visualise the proportion of supercluster membership at that node) are mostly A+ and A superclusters, and are defined by their high dependent-centric access (about median and above), and high diversity score.
+A+ is differentiated from A by its exceptionally high health-centric totals while A most just has excellent scores.
+Columns four, six and seven are dominated by the presence of B+ and B parcels.
+On average the B+ parcels provide access to an average level of dependent-centric destinations, so the B+ parcels with greater than 6.5 dependent-centric destinations are found in the more left-most columns, and those below this split on the right-most.
+The far right-most column is dominated by the presence of F parcels, defined by lower DEPENDENT TOTALS and LOW DIVERSITY.
+It is interesting to note the 5th column, which is defined by D parcels but also the most amount of A- parcels than any other column.
+A- and D parcels are both high in dependent-centric destinations and low in diversity (though D is much lower than A-).
+These two parcels again, as shown in @fig-Fig7, can be thought of as parcels in neighbourhoods containing a high amount of some caring destinations, but lacking others, indicating high efficiency for improvement potential.
+-->
+
+<!-- ```{r plot-figure7, out.width=600, out.height=300} -->
+<!-- #| label: fig-Fig7 -->
+<!-- #| fig-cap: "The Decision Tree demonstrating variables that define the superclusters." -->
+<!-- knitr::include_graphics(glue::glue(here::here(), "/figures/access_sc_profiles_usingSCVARS.png")) -->
+<!-- ``` -->
+
+<!-- D (Nothing much: Not caring & not complete) (C7) -->
+
+<!-- B- (Below average, but dep- heavy: Somewhat caring & not complete) (C4) -->
+<!-- B (Average: Caring & complete) (C5) -->
+<!-- B+ (Above average: Caring & very complete) (C6) -->
+
+<!-- A- (Excellent, but dep- heavy: Very caring & not complete (C3) -->
+<!-- A (Excellent: Very caring & very complete) (C2) -->
+<!-- A+ (Exceptional: very caring & very complete) (C1) -->
+
+OLD RANDOM HIGHLIGHTS:
+
+- The 15-Minut City have re-emerged as a popular concept in urban-planning city-planning; emphasizing the access to urban amenities by foot or bicycle.
+However, historically, neighbourhood planning processes have been top-down; FMC is linked to the Neighbourhood Unit Concept of the 1920s, largely critized for supporting segregated neighbourhoods, leaving a void in neighbourhood planning concepts that were filled by auto-centric suburban sprawl.
+Regardless, fundamental concepts within FMC remain popular: people identify with neighbourhoods, see cities through that lense.
+These ideas have in part re-emerged in the popularity of 15-minute neighbourhoods.
+
+- But how do we uplift the positive, popular, aspects of FMN without opening it up to the critics of the NUC?
+In this work, we turn to feminist geographies and the Mobility of Care conceptualisation and apply it to FMC.
+We argue, conventional planning practice is masculinist.
+This can threaten the FMC concept by shaping how 'important' amenities are conceptualised thus favouring work- and leisure- centered amenities in cities that are increasingly capital-driven.
+Though FMN is fundamentally based in bottom-up collaborative planning processes, we exist in the context of planning pracice which is largely masculinist.
+
+- FMN are for people - people do unpaid labour, those people are often women and intersectionally women-of-colour and low-income.
+However, this work should be for all people.
+Cities should be designed with this conceptualisation in mind.
+This is to combate the growing capitalisation of cities.
+
+- Our paper offers this nuanced understanding of FNC -- calling for them to center Care Destinations as important amenity types with the explicit conceptualisation that their destination type matters especially for women (low-income and of colour) who interact with them most, but also for all people.
+Their explicit positioning also can indirectly combat finacialisation of cities and co-opting of the FMC concept to one that centers capital-generating destination types like work and prohibitive leisure.
+Not to say those destination types should or cannot co-exist with Care destinations -- they should -- but markets are already incentivizing them.
+The market does not speak for non-economically geberating destination types, so we must explicitly call for them and support them within the FNC framework.
+- 
+-

### 2024-10-10 paper/Manuscript.qmd
Edits before submission

@@
 
-Though the term 15-Minute City is relatively new, neighbourhood planning practice aiming to foster community and local travel to amenities is not.
+The term 15-Minute City is relatively new but the concept is not, as it is simply a normative cumulative-opportunities accessibility measure [@paez_measuring_2012].
+Its newfound fame under the moniker of chronourbanism comes to complement long-standing efforts to foster community and local travel to amenities in neighbourhood planning practice.
 As recent examples, planned neighbourhoods dominated post-WWII built urban form: the Neighbourhood Unit Concept in the western world [@brodyNeighbourhoodUnitConcept2013] and a parallel concept of the “mikrorayons” neighbourhoods in the Soviet Union  [@kissfazekasCircleParadigms15minute2022].
 Along different dimensions, these planning forms have been extensively critiqued [@talenSocialSciencePlanned2017].
-In this respect, neighbourhood planning approaches such as the 15-Minute City concept are being re-embraced by some, but from bottom-up perspective using co-creation tools and meaningful resident participation to achieve equitable and justice neighbourhood forms [@mahmoudCocreationPathwayUrban2021].
-How to equitably prescribe urban form that 'enables' contact with social opportunities (instead of 'engendering it' from the top-down) is still under debate.
-Questions remain, including: what destinations really matter?, what mode and travel time threshold is most appropriate?, and who benefits?
-For instance, these dimensions are under discussion among proximity-based planners who use accessibility-based tools [@silvaProximitycentredAccessibilityConceptual2023; @silvaRegionalAccessibilityUndermining2022; @guzmanProximityEnoughCritical2024].
+In this respect, neighbourhood planning approaches such as the 15-Minute City concept offer an opportunity for bottom-up approaches that leverage co-creation tools and meaningful resident participation to achieve equitable and just neighbourhood forms [@mahmoudCocreationPathwayUrban2021].
+How to prescribe equitable urban forms that 'enable' contact with social opportunities (instead of 'engendering it' from the top-down) is still a matter for debate.
+Questions remain, including: what destinations really matter?
+What mode and travel time threshold is most appropriate?
+And who benefits?
+To be certain, these dimensions are under discussion among proximity-based planners who use accessibility-based tools [@silvaProximitycentredAccessibilityConceptual2023; @silvaRegionalAccessibilityUndermining2022; @guzmanProximityEnoughCritical2024].
 
@@
 In contrast, cities ought to be planned according to the multiple identities of the inhabitants [@vacchelli_towards_2018; @urbandevelopmentviennaGenderMainstreamingUrban2013].
-Researchers have demonstrated how mobility patterns differ according to gender [@lawWomenTransportNew1999; @cresswell_gendered_2008; @levy2013travel; @little_gender_1994; @tronto_toward_1990].
-Gendered social norms and cultural factors play a role in how, where, with whom, what for, and how far we move.
-Women have higher activities related to care [@ilo_care_2018], no matter at what life stage they are [@garciaromanGenderDifferencesTime2022].
-For all these reasons the spatial organization of activities has a different impact according to gender, the 15-Minute City concept should not be gender blind.
+Researchers have demonstrated how mobility patterns differ according to gender [@lawWomenTransportNew1999; @cresswell_gendered_2008; @levy2013travel; @little_gender_1994; @tronto_toward_1990; @neto_gender_2015].
+The causes for this are varied, and related to gendered social norms and cultural factors that play a role in how, where, with whom, what for, and how far we move.
+Of note is that women tend to travel more frequently and for shorter distance [@roorda_triap_2010; @morency_distance_2011] and have more activities related to care [@ilo_care_2018] no matter at what life stage they are [@garciaromanGenderDifferencesTime2022].
+Combined, these findings indicate that the spatial organization of activities has a different impact according to gender, and therefore the 15-Minute City concept should not be gender blind.
 
-This work builds a theoretical bridge between the 15-minute city and mobility of care concepts to answer questions about what destinations matter and who benefits from proximity planning.
+This work builds a theoretical bridge between the 15-Minute City and mobility of care concepts to answer questions about what destinations matter and who benefits from proximity planning.
 The Mobility of Care term was coined in @sanchezdemadariagaMobilityCareIntroducing2013; it refers to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands.
 While decades of research have examined types of household-serving trips (e.g., shopping trips, escorting to school), @sanchezdemadariagaMobilityCareIntroducing2013 was the first to consider them all as one category.
-In doing so, Mobility of Care is shown to be a significant proportion of daily travel, characterizing the trip purpose of approximately 30% of adults’ trips [@sanchezdemadariagaMobilityCareIntroducing2013; @sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory].
-We argue Mobility of Care is especially relevant to the 15-Minute City, as care trips are necessary for _all_ people and they most often occur at the local level.
-Moreover, care trips take on tremendous gendered and socio-economic importance, as women, and especially those from lower income households, complete the majority of care trips [@ravensbergen2023exploratory].
+In doing so, care-related mobility can be shown to comprise a significant proportion of daily travel, characterizing approximately 30% of adults’ daily trips [@sanchezdemadariagaMobilityCareIntroducing2013; @sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory].
+We argue that Mobility of Care is especially relevant to the 15-Minute City, as care trips are necessary for _all_ people and they most often occur at the local level.
+Moreover, care trips are of particular gendered and socio-economic import, as women, and especially those in lower income households, complete the majority of care trips [@ravensbergen2023exploratory].
 Despite care travel's importance in advancing gender equality -and equity more broadly-, mobility of care and access to care destinations have been under-examined relative to employment destinations.
@@
 The objectives of this research are two-fold.
-First, we aim to theoretically bridge Mobility of Care with the 15-Minute City concept to re-imagine what local amenities matter and to define an associated accessibility and diversity of accessibility measures.
-Second, we take a data-driven approach to classify areas into 'caring 15-minute neighbourhoods' typologies and examine associated residential profiles in an empirical examination of the city of Hamilton, Canada.
+First, we aim to theoretically bridge Mobility of Care with the 15-Minute City concept to re-imagine what local amenities matter from the perspective of care and to define an associated accessibility and diversity of accessibility measures.
+And secondly, we take a data-driven approach to classify areas into 'caring 15-minute neighbourhoods' typologies and examine associated residential profiles in an empirical examination of the city of Hamilton, Canada.
 
@@
 
-In the last decades, the need for more sustainable, healthier, and livable cities have received significant attention.
-Planners and decision-makers have proposed alternatives to mobility-based neighbourhood planning principles to ones that center proximity to urban functions [@pozoukidou15MinuteCityDecomposing2021].
+In the last decades, the need for more sustainable, healthier, and livable cities has become a prominent concern.
+Planners and decision-makers have proposed a shift to neighbourhood planning principles centered on proximity to urban functions [@pozoukidou15MinuteCityDecomposing2021].
 In this context, the 15-Minute City is now in the public spotlight [@logan_xminute_2022; @moreno_introducing_2021].
-The 15-Minute City was introduced in @moreno_introducing_2021 along with four dimensions: density (in terms of people per km$^2$), diversity (including mixed-land use and diversity of people), the temporal and spatial proximity to essential services, and digitalisation.
-However, the framework presented has been criticized within academic and planning circles [@guzmanProximityEnoughCritical2024].
-Critiques highlight how the concept falls short in addressing the pre-existing structural forces and individual characteristics that drive inequalities, influencing who benefits or could benefit the most from such a paradigm [@dimarino15minuteCityConcept2023; @willberg15minuteCityAll2023].
-A 15-Minute City for all people is an aspirational ideal, but how will existing mobility and accessibility inequalities be addressed.
+The 15-Minute City was introduced in @moreno_introducing_2021 along with four dimensions: density (in terms of people per km$^2$), diversity (including mixed-land use and diversity of people), the temporal and spatial proximity to essential services, and digitalisation [q.v.
+@cerveroTravelDemand3Ds1997].
+However, the framework presented has been criticized within academic and planning circles [e.g., @guzmanProximityEnoughCritical2024; @mouratidisTimeChallenge15minute2024] among other things for shortfalls in terms of addressing pre-existing structural forces and individual characteristics that drive inequalities and influence who benefits or could benefit the most from such an approach [@dimarino15minuteCityConcept2023; @willberg15minuteCityAll2023].
+A 15-Minute City for all people is an aspirational goal, but does not fully confront existing mobility and accessibility inequalities in need of redress.
 Without directed and context-specific solutions, this popular concept risks becoming an empty city branding exercise.
@@
 
-Reflecting the flexible and aspirational presentation of the 15-Minute City; cities that have adopted the concept have done so using a diverse range of definitions and tools along with indistinct universal approaches.
-A trail blazer was Portland, US in the Portland Plan [@city_of_portland_20minute_2010] of April 2012.
-Adopted before the '15-Minute City' concept of today, it aims to foster an inclusive city development based on prosperity, education, health, and equity over a 30 year horizon.
-The plan aims to promote neighbourhood self-sufficiency and connectivity to city centers and centres of employment.
+Reflective of the flexible and aspirational presentation of the 15-Minute City, the concept has been adopted by cities using a diverse range of definitions and tools along with indistinct universal approaches.
+A trail blazer in this respect was Portland (US) in the Portland Plan [@city_of_portland_20minute_2010] of April 2012.
+Adopted before the '15-Minute City' concept of today, this plan aimed to foster inclusive city development based on prosperity, education, health, and equity over a 30 year horizon.
+Central to the plan was the promotion of neighbourhood self-sufficiency and connectivity to city centers and centres of employment.
 The progress report describes a high-level focus on equitable service delivery to all residents with equity concerning topics related to racial equity and people with disabilities [@portland_government_portland_2017], similarly taking on an "all populations" approach.
-Later, other cities adopted proximity-based goals using similar neutral approaches.
-For example, the 15-Minute City plan announced during the Paris mayor Anne Hidalgo’s 2020 re-election campaign, which helped popularize the concept in today's public discourse, emphasized six key social functions that should be easily accessible from any location [@ville_de_paris_paris_2022].
+Subsequently, other cities adopted proximity-based goals using similarly neutral approaches.
+For example, the 15-Minute City plan announced during the re-election campaign of Paris mayor Anne Hidalgo in 2020 emphasized six key social functions that should be easily accessible from any location [@ville_de_paris_paris_2022].
 These locations included: housing, work, health care, groceries, education and leisure.
@@
 
-The past can shed light on cautionary tales.
-While the 15-Minute City concept is new, neighbourhood planning to improve society outcomes is not: in fact the literature has drawn parallels from the 15-Minute City to Clarence A.
+The past can provide cautionary tales.
+While the 15-Minute City brand of planning is new, neighbourhood planning to improve society outcomes is not: in fact the literature has drawn parallels from the 15-Minute City to Clarence A.
 Perry's Neighbourhood Unit Concept (NUC) from the 1930s [@kissfazekasCircleParadigms15minute2022].
@@
 The NUC primarily prioritized local service provision, though Perry had confidence in good design's contribution to 'neighbourhood spirit' [@hall2014cities].
-From the NUC's conception until the end of the 1960s, planners' aspirational attempts to prescribe social meaning to the neighbourhood's physical form had been criticized to near extinction [@talenSocialSciencePlanned2017].
-By social scientists, the overestimation on the impact of built environment on social life was critiqued.
-By other planners, the inability to agree on the specificity of the neighbourhood (i.e., population size and the type, quality, and quantity of amenity) as well as how neighbourhood units are connected to other neighbourhoods and the rest of the region.
-In response to these criticisms, neighbourhood planning proponents redefined deterministic terminology; their prescriptive physical form will 'enable' social contact with opportunities rather than 'engendering' it.
+By the end of the 1960s, planners' aspirational attempts to prescribe social meaning to the neighbourhood's physical form had been criticized to near extinction [@talenSocialSciencePlanned2017].
+A critique by social scientists was an overestimation of the impact of built environment on social life.
+Planners, on the other hand, could not reach consensus on the specificity of the neighbourhood (i.e., population size and the type, quality, and quantity of amenity) or how neighbourhood units connect between them and the rest of the region.
+In response to these criticisms, neighbourhood planning proponents redefined their deterministic terminology; their prescriptive physical form would 'enable' social contact with opportunities rather than 'engendering' it.
 
-The redefined bottom-up approach to community and neighbourhood planning has been adopted by New Urbanists, from which the 15-Minute City stems [@kissfazekasCircleParadigms15minute2022]; however, the effectiveness of these bottom-up processes remain to be seen due to the contemporary nature of the concept.
+The redefined bottom-up approach to community and neighbourhood planning was adopted  in the 1980s by American New Urbanists [@trudeau_new_2013], from which the 15-Minute City eventually stemmed [@kissfazekasCircleParadigms15minute2022].
+However, the bottom-up effectiveness of these ideas remains to be seen due to the contemporary nature of the concept.
 Though, related research can be examined.
@@
 A new wave of researchers and practitioners focused on local and context-specific relationships with the proximity to destinations have also emerged [@silvaProximitycentredAccessibilityConceptual2023; @silvaRegionalAccessibilityUndermining2022].
-As reviewed in the city plans that have adopted 15-Minute City approaches, the NUC's criticisms, and recent urban planning research, the question of _how to enable improved quality of life through urban built environment_ has shown to be highly context-sensitive, with the need for further investigation.
+As reviewed in the city plans that have adopted 15-Minute City approaches, the NUC's criticisms, and recent urban planning research, the question of _how to enable improved quality of life through urban built environment_ is highly context-sensitive, prompting the need for further investigation.
 
@@
 The "cumulative opportunity" measure has been applied in many 15-Minute City examinations.
-This measure quantifies how many destinations can be reached from a point in space within a given travel time threshold, pairing well with examinations that have a normative travel time threshold, like the 15-Minute City.
+This measure quantifies how many destinations can be reached from a point in space within a given travel time threshold , pairing well with normative examinations [see @paez_measuring_2012] that use travel time thresholds like x-minute cities [@logan_xminute_2022.
 Furthermore, the cumulative opportunity measure is widely appreciated for its intuitive computation and popularity among practitioners [@handyAccessibilityIdeaWhose2020; @handyMeasuringAccessibilityExploration1997; @chengInvestigatingWalkingAccessibility2019].
@@
 
-Another method that compliments the assessment of the 15-Minute City regarding the diversity of opportunity types is entropy.
+Another concept that complements the assessment of the 15-Minute City regarding the diversity of opportunity types is entropy.
 The entropy measure, based on the Shannon-Wiener index of species diversity, expresses relative evenness within a sequence [@whittaker1972evolution].
-In urban studies and planning literature, the flexible concept of entropy has been used to characterize land-use mix diversity [@frankLinkingObjectivelyMeasured2005; @ewingTravelBuiltEnvironment2010a], particularly in the context of non-work trips [@cerveroTravelDemand3Ds1997], walking as a mode [@luUrbanDensityDiversity2017; @mavoaIdentifyingAppropriateLanduse2018], and suburban sprawl [@randallGISBasedLand2015] as relevant to this work.
-It has also been used to understand mobility behaviour [@mcbride2020exploration; @montero2023applying; @montero2023role].
-However, work incorporating entropy indices alongside accessibility analysis are not exceptionally common.
+In urban studies and planning literature, the concept of entropy has been used to characterize land-use mix diversity [@frankLinkingObjectivelyMeasured2005; @ewingTravelBuiltEnvironment2010a; @moniruzzaman_accessibility_2012] including to understand mobility behaviour [@mcbride2020exploration; @montero2023applying; @montero2023role], in the context of non-work trips [@cerveroTravelDemand3Ds1997], walking [@luUrbanDensityDiversity2017; @mavoaIdentifyingAppropriateLanduse2018], and suburban sprawl [@randallGISBasedLand2015].
+However, entropy indices are rare in accessibility analysis.
 There are examples of their use as parameters within accessibility scores for employment opportunities [@chengMeasuringUrbanJob2013; @daiIncorporatingJobDiversity2018] and, more recently, to describe the diversity in transit facilities [@yinIncorporatingFacilityDiversity2024].
-Within the conceptualisation of the 15-Minute City, diversity of land-use is a key feature.
+Given how 15-Minute Cities aim to provide access to numerous amenities, diversity of land-use is a key feature.
 In this way, the entropy measure theoretically complements the 15-Minute City.
@@
 Classification algorithms also show promise as a tool in the assessment of the 15-Minute City, as they have been useful in identifying mobility and spatial typologies within the transportation planning literature.
-Often, these algorithms take the form of machine learning approaches: for instance, @burkeGeospatialAnalysisFramework2022 uses spatially constrained multivariate clustering, to develop urban form typologies related to the 15-Minute City.
-The k-means algorithm is another popular method: @ganUnderstandingUrbanMobility2020 applies it to analyse travel behaviour and classify metro stations based on mobility patterns.
-Similarly, [@montero2023role; @montero2023applying] cluster travel behavior profile typologies representing different population segments and place of residence in the Barcelona Metropolitan Area.
-Combing classification techniques is also common: for instance, @delmelleDifferentiatingPathwaysNeighborhood2017 uses self-organizing maps (SOM) algorithm to group U.S.
-neighbourhoods by minimizing dissimilarity in attributes before applying k-means for further classification.
-Similarly, @victorianoTimeSpaceMoney2020 uses SOM to classify individuals' travel patterns according to the dissimilarity in mobility attributes and uses the Decision Tree algorithm to partition the data into interpretable classifications.
+Often, these algorithms take the form of machine learning approaches: in transportation, the use of Self-Organizing Maps (SOM) was pioneered by Delmelle [-@delmelle_spatial_2012] and has been used to group U.S.
+neighbourhoods by minimizing dissimilarity in attributes [@delmelleDifferentiatingPathwaysNeighborhood2017] and to classify individuals' travel patterns according to the dissimilarity in mobility attributes and uses the Decision Tree algorithm to partition the data into interpretable classifications [@victorianoTimeSpaceMoney2020].
+Other approaches include the use of spatially constrained multivariate clustering, to develop urban form typologies related to the 15-Minute City [@burkeGeospatialAnalysisFramework2022] and the use of k-means to analyse travel behaviour and classify metro stations based on mobility patterns[ @ganUnderstandingUrbanMobility2020].
 
@@
 
-Rather than focusing on _all_ destinations, it may valuable to examine those related to caring activities.
-This re-framing offers a feminist perspective on which urban functions matter, connecting well with the 15-Minute City concept.
+Rather than focusing on _all_ destinations, it is valuable to examine those related to caring activities.
+This framing offers a feminist perspective on urban functions that matter from a care perspective, and connects well with the 15-Minute City concept.
 Caring activities, which fulfill the physical, psychological, and emotional needs of others, are among the most essential and fundamental activities in society [@ilo_care_2018].
@@
 Though there is ample literature that examines walking to care destinations, such as schools (e.g., [@RN7;@RN8;@RN9]) and grocery stores (e.g., [@RN10;@RN11]), they tend to consider singular care destinations in research focusing on walkability.
-
 In summary, we reviewed the neighborhood planning framework from which the 15-Minute City concept stems and is situated within.

### 2024-10-10 paper/Manuscript.tex
Edits before submission

@@
 % Make \paragraph and \subparagraph free-standing
+\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}{
+    \@ifstar
+      \xxxParagraphStar
+      \xxxParagraphNoStar
+  }
+  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
+  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}{
+    \@ifstar
+      \xxxSubParagraphStar
+      \xxxSubParagraphNoStar
+  }
+  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
+  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
+\makeatother
 
@@
 \makeatother
+
 \ifLuaTeX
@@
 
-Though the term 15-Minute City is relatively new, neighbourhood planning
-practice aiming to foster community and local travel to amenities is
-not. As recent examples, planned neighbourhoods dominated post-WWII
+The term 15-Minute City is relatively new but the concept is not, as it
+is simply a normative cumulative-opportunities accessibility measure
+\citep{paez_measuring_2012}. Its newfound fame under the moniker of
+chronourbanism comes to complement long-standing efforts to foster
+community and local travel to amenities in neighbourhood planning
+practice. As recent examples, planned neighbourhoods dominated post-WWII
 built urban form: the Neighbourhood Unit Concept in the western world
@@
 \citep{talenSocialSciencePlanned2017}. In this respect, neighbourhood
-planning approaches such as the 15-Minute City concept are being
-re-embraced by some, but from bottom-up perspective using co-creation
-tools and meaningful resident participation to achieve equitable and
-justice neighbourhood forms \citep{mahmoudCocreationPathwayUrban2021}.
-How to equitably prescribe urban form that `enables' contact with social
-opportunities (instead of `engendering it' from the top-down) is still
-under debate. Questions remain, including: what destinations really
-matter?, what mode and travel time threshold is most appropriate?, and
-who benefits? For instance, these dimensions are under discussion among
+planning approaches such as the 15-Minute City concept offer an
+opportunity for bottom-up approaches that leverage co-creation tools and
+meaningful resident participation to achieve equitable and just
+neighbourhood forms \citep{mahmoudCocreationPathwayUrban2021}. How to
+prescribe equitable urban forms that `enable' contact with social
+opportunities (instead of `engendering it' from the top-down) is still a
+matter for debate. Questions remain, including: what destinations really
+matter? What mode and travel time threshold is most appropriate? And who
+benefits? To be certain, these dimensions are under discussion among
 proximity-based planners who use accessibility-based tools
@@
 gender
-\citep{lawWomenTransportNew1999, cresswell_gendered_2008, levy2013travel, little_gender_1994, tronto_toward_1990}.
-Gendered social norms and cultural factors play a role in how, where,
-with whom, what for, and how far we move. Women have higher activities
-related to care \citep{ilo_care_2018}, no matter at what life stage they
-are \citep{garciaromanGenderDifferencesTime2022}. For all these reasons
-the spatial organization of activities has a different impact according
-to gender, the 15-Minute City concept should not be gender blind.
-
-This work builds a theoretical bridge between the 15-minute city and
+\citep{lawWomenTransportNew1999, cresswell_gendered_2008, levy2013travel, little_gender_1994, tronto_toward_1990, neto_gender_2015}.
+The causes for this are varied, and related to gendered social norms and
+cultural factors that play a role in how, where, with whom, what for,
+and how far we move. Of note is that women tend to travel more
+frequently and for shorter distance
+\citep{roorda_triap_2010, morency_distance_2011} and have more
+activities related to care \citep{ilo_care_2018} no matter at what life
+stage they are \citep{garciaromanGenderDifferencesTime2022}. Combined,
+these findings indicate that the spatial organization of activities has
+a different impact according to gender, and therefore the 15-Minute City
+concept should not be gender blind.
+
+This work builds a theoretical bridge between the 15-Minute City and
 mobility of care concepts to answer questions about what destinations
@@
 \citet{sanchezdemadariagaMobilityCareIntroducing2013} was the first to
-consider them all as one category. In doing so, Mobility of Care is
-shown to be a significant proportion of daily travel, characterizing the
-trip purpose of approximately 30\% of adults' trips
+consider them all as one category. In doing so, care-related mobility
+can be shown to comprise a significant proportion of daily travel,
+characterizing approximately 30\% of adults' daily trips
 \citep{sanchezdemadariagaMobilityCareIntroducing2013, sanchezdemadariagaMeasuringMobilitiesCare2019, ravensbergen2023exploratory}.
-We argue Mobility of Care is especially relevant to the 15-Minute City,
-as care trips are necessary for \emph{all} people and they most often
-occur at the local level. Moreover, care trips take on tremendous
-gendered and socio-economic importance, as women, and especially those
-from lower income households, complete the majority of care trips
+We argue that Mobility of Care is especially relevant to the 15-Minute
+City, as care trips are necessary for \emph{all} people and they most
+often occur at the local level. Moreover, care trips are of particular
+gendered and socio-economic import, as women, and especially those in
+lower income households, complete the majority of care trips
 \citep{ravensbergen2023exploratory}. Despite care travel's importance in
@@
 theoretically bridge Mobility of Care with the 15-Minute City concept to
-re-imagine what local amenities matter and to define an associated
-accessibility and diversity of accessibility measures. Second, we take a
-data-driven approach to classify areas into `caring 15-minute
-neighbourhoods' typologies and examine associated residential profiles
-in an empirical examination of the city of Hamilton, Canada.
+re-imagine what local amenities matter from the perspective of care and
+to define an associated accessibility and diversity of accessibility
+measures. And secondly, we take a data-driven approach to classify areas
+into `caring 15-minute neighbourhoods' typologies and examine associated
+residential profiles in an empirical examination of the city of
+Hamilton, Canada.
 
@@
 In the last decades, the need for more sustainable, healthier, and
-livable cities have received significant attention. Planners and
-decision-makers have proposed alternatives to mobility-based
-neighbourhood planning principles to ones that center proximity to urban
-functions \citep{pozoukidou15MinuteCityDecomposing2021}. In this
-context, the 15-Minute City is now in the public spotlight
+livable cities has become a prominent concern. Planners and
+decision-makers have proposed a shift to neighbourhood planning
+principles centered on proximity to urban functions
+\citep{pozoukidou15MinuteCityDecomposing2021}. In this context, the
+15-Minute City is now in the public spotlight
 \citep{logan_xminute_2022, moreno_introducing_2021}. The 15-Minute City
@@
 (including mixed-land use and diversity of people), the temporal and
-spatial proximity to essential services, and digitalisation. However,
-the framework presented has been criticized within academic and planning
-circles \citep{guzmanProximityEnoughCritical2024}. Critiques highlight
-how the concept falls short in addressing the pre-existing structural
-forces and individual characteristics that drive inequalities,
-influencing who benefits or could benefit the most from such a paradigm
+spatial proximity to essential services, and digitalisation
+\citep[q.v.][]{cerveroTravelDemand3Ds1997}. However, the framework
+presented has been criticized within academic and planning circles
+\citep[e.g.,][]{guzmanProximityEnoughCritical2024, mouratidisTimeChallenge15minute2024}
+among other things for shortfalls in terms of addressing pre-existing
+structural forces and individual characteristics that drive inequalities
+and influence who benefits or could benefit the most from such an
+approach
 \citep{dimarino15minuteCityConcept2023, willberg15minuteCityAll2023}. A
-15-Minute City for all people is an aspirational ideal, but how will
-existing mobility and accessibility inequalities be addressed. Without
-directed and context-specific solutions, this popular concept risks
-becoming an empty city branding exercise.
+15-Minute City for all people is an aspirational goal, but does not
+fully confront existing mobility and accessibility inequalities in need
+of redress. Without directed and context-specific solutions, this
+popular concept risks becoming an empty city branding exercise.
 \citep{pozoukidou15MinuteCityDecomposing2021, gowerPlanningInnovationCity2022}.
 
-Reflecting the flexible and aspirational presentation of the 15-Minute
-City; cities that have adopted the concept have done so using a diverse
+Reflective of the flexible and aspirational presentation of the
+15-Minute City, the concept has been adopted by cities using a diverse
 range of definitions and tools along with indistinct universal
-approaches. A trail blazer was Portland, US in the Portland Plan
-\citep{city_of_portland_20minute_2010} of April 2012. Adopted before the
-`15-Minute City' concept of today, it aims to foster an inclusive city
-development based on prosperity, education, health, and equity over a 30
-year horizon. The plan aims to promote neighbourhood self-sufficiency
-and connectivity to city centers and centres of employment. The progress
-report describes a high-level focus on equitable service delivery to all
-residents with equity concerning topics related to racial equity and
-people with disabilities \citep{portland_government_portland_2017},
-similarly taking on an ``all populations'' approach. Later, other cities
-adopted proximity-based goals using similar neutral approaches. For
-example, the 15-Minute City plan announced during the Paris mayor Anne
-Hidalgo's 2020 re-election campaign, which helped popularize the concept
-in today's public discourse, emphasized six key social functions that
-should be easily accessible from any location
+approaches. A trail blazer in this respect was Portland (US) in the
+Portland Plan \citep{city_of_portland_20minute_2010} of April 2012.
+Adopted before the `15-Minute City' concept of today, this plan aimed to
+foster inclusive city development based on prosperity, education,
+health, and equity over a 30 year horizon. Central to the plan was the
+promotion of neighbourhood self-sufficiency and connectivity to city
+centers and centres of employment. The progress report describes a
+high-level focus on equitable service delivery to all residents with
+equity concerning topics related to racial equity and people with
+disabilities \citep{portland_government_portland_2017}, similarly taking
+on an ``all populations'' approach. Subsequently, other cities adopted
+proximity-based goals using similarly neutral approaches. For example,
+the 15-Minute City plan announced during the re-election campaign of
+Paris mayor Anne Hidalgo in 2020 emphasized six key social functions
+that should be easily accessible from any location
 \citep{ville_de_paris_paris_2022}. These locations included: housing,
@@
 
-The past can shed light on cautionary tales. While the 15-Minute City
-concept is new, neighbourhood planning to improve society outcomes is
+The past can provide cautionary tales. While the 15-Minute City brand of
+planning is new, neighbourhood planning to improve society outcomes is
 not: in fact the literature has drawn parallels from the 15-Minute City
@@
 confidence in good design's contribution to `neighbourhood spirit'
-\citep{hall2014cities}. From the NUC's conception until the end of the
-1960s, planners' aspirational attempts to prescribe social meaning to
-the neighbourhood's physical form had been criticized to near extinction
-\citep{talenSocialSciencePlanned2017}. By social scientists, the
-overestimation on the impact of built environment on social life was
-critiqued. By other planners, the inability to agree on the specificity
-of the neighbourhood (i.e., population size and the type, quality, and
-quantity of amenity) as well as how neighbourhood units are connected to
-other neighbourhoods and the rest of the region. In response to these
-criticisms, neighbourhood planning proponents redefined deterministic
-terminology; their prescriptive physical form will `enable' social
-contact with opportunities rather than `engendering' it.
+\citep{hall2014cities}. By the end of the 1960s, planners' aspirational
+attempts to prescribe social meaning to the neighbourhood's physical
+form had been criticized to near extinction
+\citep{talenSocialSciencePlanned2017}. A critique by social scientists
+was an overestimation of the impact of built environment on social life.
+Planners, on the other hand, could not reach consensus on the
+specificity of the neighbourhood (i.e., population size and the type,
+quality, and quantity of amenity) or how neighbourhood units connect
+between them and the rest of the region. In response to these
+criticisms, neighbourhood planning proponents redefined their
+deterministic terminology; their prescriptive physical form would
+`enable' social contact with opportunities rather than `engendering' it.
 
 The redefined bottom-up approach to community and neighbourhood planning
-has been adopted by New Urbanists, from which the 15-Minute City stems
-\citep{kissfazekasCircleParadigms15minute2022}; however, the
-effectiveness of these bottom-up processes remain to be seen due to the
+was adopted in the 1980s by American New Urbanists
+\citep{trudeau_new_2013}, from which the 15-Minute City eventually
+stemmed \citep{kissfazekasCircleParadigms15minute2022}. However, the
+bottom-up effectiveness of these ideas remains to be seen due to the
 contemporary nature of the concept. Though, related research can be
@@
 the question of \emph{how to enable improved quality of life through
-urban built environment} has shown to be highly context-sensitive, with
-the need for further investigation.
+urban built environment} is highly context-sensitive, prompting the need
+for further investigation.
 
@@
 quantifies how many destinations can be reached from a point in space
-within a given travel time threshold, pairing well with examinations
-that have a normative travel time threshold, like the 15-Minute City.
+within a given travel time threshold , pairing well with normative
+examinations \citep[see][]{paez_measuring_2012} that use travel time
+thresholds like x-minute cities {[}\citet{logan_xminute_2022}.
 Furthermore, the cumulative opportunity measure is widely appreciated
@@
 
-Another method that compliments the assessment of the 15-Minute City
+Another concept that complements the assessment of the 15-Minute City
 regarding the diversity of opportunity types is entropy. The entropy
@@
 \citep{whittaker1972evolution}. In urban studies and planning
-literature, the flexible concept of entropy has been used to
-characterize land-use mix diversity
-\citep{frankLinkingObjectivelyMeasured2005, ewingTravelBuiltEnvironment2010a},
-particularly in the context of non-work trips
-\citep{cerveroTravelDemand3Ds1997}, walking as a mode
+literature, the concept of entropy has been used to characterize
+land-use mix diversity
+\citep{frankLinkingObjectivelyMeasured2005, ewingTravelBuiltEnvironment2010a, moniruzzaman_accessibility_2012}
+including to understand mobility behaviour
+\citep{mcbride2020exploration, montero2023applying, montero2023role}, in
+the context of non-work trips \citep{cerveroTravelDemand3Ds1997},
+walking
 \citep{luUrbanDensityDiversity2017, mavoaIdentifyingAppropriateLanduse2018},
-and suburban sprawl \citep{randallGISBasedLand2015} as relevant to this
-work. It has also been used to understand mobility behaviour
-\citep{mcbride2020exploration, montero2023applying, montero2023role}.
-However, work incorporating entropy indices alongside accessibility
-analysis are not exceptionally common. There are examples of their use
-as parameters within accessibility scores for employment opportunities
+and suburban sprawl \citep{randallGISBasedLand2015}. However, entropy
+indices are rare in accessibility analysis. There are examples of their
+use as parameters within accessibility scores for employment
+opportunities
 \citep{chengMeasuringUrbanJob2013, daiIncorporatingJobDiversity2018}
 and, more recently, to describe the diversity in transit facilities
-\citep{yinIncorporatingFacilityDiversity2024}. Within the
-conceptualisation of the 15-Minute City, diversity of land-use is a key
-feature. In this way, the entropy measure theoretically complements the
-15-Minute City.
+\citep{yinIncorporatingFacilityDiversity2024}. Given how 15-Minute
+Cities aim to provide access to numerous amenities, diversity of
+land-use is a key feature. In this way, the entropy measure
+theoretically complements the 15-Minute City.
 
@@
 and spatial typologies within the transportation planning literature.
-Often, these algorithms take the form of machine learning approaches:
-for instance, \citet{burkeGeospatialAnalysisFramework2022} uses
-spatially constrained multivariate clustering, to develop urban form
-typologies related to the 15-Minute City. The k-means algorithm is
-another popular method: \citet{ganUnderstandingUrbanMobility2020}
-applies it to analyse travel behaviour and classify metro stations based
-on mobility patterns. Similarly,
-\citep{montero2023role, montero2023applying} cluster travel behavior
-profile typologies representing different population segments and place
-of residence in the Barcelona Metropolitan Area. Combing classification
-techniques is also common: for instance,
-\citet{delmelleDifferentiatingPathwaysNeighborhood2017} uses
-self-organizing maps (SOM) algorithm to group U.S. neighbourhoods by
-minimizing dissimilarity in attributes before applying k-means for
-further classification. Similarly, \citet{victorianoTimeSpaceMoney2020}
-uses SOM to classify individuals' travel patterns according to the
-dissimilarity in mobility attributes and uses the Decision Tree
-algorithm to partition the data into interpretable classifications.
+Often, these algorithms take the form of machine learning approaches: in
+transportation, the use of Self-Organizing Maps (SOM) was pioneered by
+Delmelle \citeyearpar{delmelle_spatial_2012} and has been used to group
+U.S. neighbourhoods by minimizing dissimilarity in attributes
+\citep{delmelleDifferentiatingPathwaysNeighborhood2017} and to classify
+individuals' travel patterns according to the dissimilarity in mobility
+attributes and uses the Decision Tree algorithm to partition the data
+into interpretable classifications \citep{victorianoTimeSpaceMoney2020}.
+Other approaches include the use of spatially constrained multivariate
+clustering, to develop urban form typologies related to the 15-Minute
+City \citep{burkeGeospatialAnalysisFramework2022} and the use of k-means
+to analyse travel behaviour and classify metro stations based on
+mobility patterns\citep{ganUnderstandingUrbanMobility2020}.
 
@@
 
-Rather than focusing on \emph{all} destinations, it may valuable to
-examine those related to caring activities. This re-framing offers a
-feminist perspective on which urban functions matter, connecting well
-with the 15-Minute City concept. Caring activities, which fulfill the
-physical, psychological, and emotional needs of others, are among the
-most essential and fundamental activities in society
-\citep{ilo_care_2018}. Yet, they are the most unequal, undervalued, and
-even devalued activities worldwide. Conventionally, caring activities
-have been borne on women's shoulders
+Rather than focusing on \emph{all} destinations, it is valuable to
+examine those related to caring activities. This framing offers a
+feminist perspective on urban functions that matter from a care
+perspective, and connects well with the 15-Minute City concept. Caring
+activities, which fulfill the physical, psychological, and emotional
+needs of others, are among the most essential and fundamental activities
+in society \citep{ilo_care_2018}. Yet, they are the most unequal,
+undervalued, and even devalued activities worldwide. Conventionally,
+caring activities have been borne on women's shoulders
 \citep{haydenGrandDomesticRevolution1982, hochschild_second_2012}.
@@
 grocery stores (e.g., \citep{RN10, RN11}), they tend to consider
-singular care destinations in research focusing on walkability.
-
-In summary, we reviewed the neighborhood planning framework from which
-the 15-Minute City concept stems and is situated within. We then
-reviewed tools common and useful in investigating the 15-Minute City.
-Finally, we introduced the Mobility of Care concept, theoretically
-linking it to the 15-Minute City to highlight the importance of care
-destinations for all but especially along gendered and socio-economic
-lines. The objectives of this work are two-fold, achieved through an
-empirical examination of Hamilton, Canada, a mid-sized city:
+singular care destinations in research focusing on walkability. In
+summary, we reviewed the neighborhood planning framework from which the
+15-Minute City concept stems and is situated within. We then reviewed
+tools common and useful in investigating the 15-Minute City. Finally, we
+introduced the Mobility of Care concept, theoretically linking it to the
+15-Minute City to highlight the importance of care destinations for all
+but especially along gendered and socio-economic lines. The objectives
+of this work are two-fold, achieved through an empirical examination of
+Hamilton, Canada, a mid-sized city:
 

### 2024-10-10 paper/Manuscript.qmd
Submitted to journal with editorial correspondance. With lock file update.

@@
 ---
-title: "Caring 15-minute neighbourhoods"
+title: "Towards completely caring 15-minute neighbourhoods"
 author:
@@
 This gap is addressed in this paper by identifying which areas in the city of Hamilton, Canada are ‘caring 15-minute neighbourhoods’.
-To do so, a database of care destinations was compiled to estimate the number (using the cumulative opportunities accessibility measure) and diversity of mix (using the entropy measure) of care destinations within a 15-minute walk from residential parcels.
+To do so, a database of care destinations was compiled to estimate the number (using the cumulative opportunity accessibility measure) and diversity of mix (using the entropy measure) of care destinations within a 15-minute walk from residential parcels.
 Using data-driven machine learning techniques (Self-Organizing Maps and Decision Trees), neighborhoods were classified into 'caring 15-minute neighbourhood' typologies that are examined across residential socio-economic profiles.
@@
 
-Keywords: Accessibility; 15-Minute Neighborhoods; Chrono-Urbanism; Cities of Proximity; Mobility of Care; Inequality; Gender Gap;
+Keywords: 15-Minute City; Chrono-Urbanism; Accessibility; Mobility of Care; Inequality; Gender Gap
+<!-- Keywords: 15-Minute City; 15-Minute Neighbourhoods; Chrono-Urbanism; Accessibility; Mobility of Care; Cities of Proximity; Inequality; Gender Gap -->
 
@@
 
-The term 15-Minute City is relatively new but the concept is not, as it is simply a normative cumulative-opportunities accessibility measure [@paez_measuring_2012].
-Its newfound fame under the moniker of chronourbanism comes to complement long-standing efforts to foster community and local travel to amenities in neighbourhood planning practice.
+The term 15-Minute City is relatively new, but the concept is not.
+Its newfound fame under the moniker of chrono-urbanism comes to complement long-standing efforts to foster community and local travel to amenities in neighbourhood planning practice.
 As recent examples, planned neighbourhoods dominated post-WWII built urban form: the Neighbourhood Unit Concept in the western world [@brodyNeighbourhoodUnitConcept2013] and a parallel concept of the “mikrorayons” neighbourhoods in the Soviet Union  [@kissfazekasCircleParadigms15minute2022].
@@
 To be certain, these dimensions are under discussion among proximity-based planners who use accessibility-based tools [@silvaProximitycentredAccessibilityConceptual2023; @silvaRegionalAccessibilityUndermining2022; @guzmanProximityEnoughCritical2024].
+In this way, the 15-Minute City could be understood to be a normative cumulative opportunity accessibility measure [@paez_measuring_2012].
 
@@
 The causes for this are varied, and related to gendered social norms and cultural factors that play a role in how, where, with whom, what for, and how far we move.
-Of note is that women tend to travel more frequently and for shorter distance [@roorda_triap_2010; @morency_distance_2011] and have more activities related to care [@ilo_care_2018] no matter at what life stage they are [@garciaromanGenderDifferencesTime2022].
+Women tend to travel shorter distances and more frequently [@roorda_triap_2010; @morency_distance_2011], and conduct more activities related to care no matter their stage of life [@ilo_care_2018; @garciaromanGenderDifferencesTime2022].
 Combined, these findings indicate that the spatial organization of activities has a different impact according to gender, and therefore the 15-Minute City concept should not be gender blind.
 
-This work builds a theoretical bridge between the 15-Minute City and mobility of care concepts to answer questions about what destinations matter and who benefits from proximity planning.
+This work builds a theoretical bridge between the 15-Minute City and Mobility of Care concepts to answer questions about what destinations matter and who benefits from proximity planning.
 The Mobility of Care term was coined in @sanchezdemadariagaMobilityCareIntroducing2013; it refers to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands.
@@
 We argue that Mobility of Care is especially relevant to the 15-Minute City, as care trips are necessary for _all_ people and they most often occur at the local level.
-Moreover, care trips are of particular gendered and socio-economic import, as women, and especially those in lower income households, complete the majority of care trips [@ravensbergen2023exploratory].
-Despite care travel's importance in advancing gender equality -and equity more broadly-, mobility of care and access to care destinations have been under-examined relative to employment destinations.
+Moreover, care trips are of particular gendered and socio-economic importance, as women, and especially those within lower income households, complete the majority of care trips [@ravensbergen2023exploratory].
+Despite care travel's importance in advancing gender equality -and equity more broadly-, Mobility of Care and access to care destinations have been under-examined relative to employment destinations.
 This is especially pertinent in accessibility research that has largely focused on employment points of interest e.g., [@farberOntarioLineSocioeconomic2019; @duarteInfluenceJobAccessibility2023; @ryanAccessibilitySpaceTime2023].
@@
 In this context, the 15-Minute City is now in the public spotlight [@logan_xminute_2022; @moreno_introducing_2021].
-The 15-Minute City was introduced in @moreno_introducing_2021 along with four dimensions: density (in terms of people per km$^2$), diversity (including mixed-land use and diversity of people), the temporal and spatial proximity to essential services, and digitalisation [q.v.
-@cerveroTravelDemand3Ds1997].
+The 15-Minute City was introduced in @moreno_introducing_2021 along with four dimensions: density (in terms of people per km$^2$), diversity (including mixed-land use and diversity of people), the temporal and spatial proximity to essential services, and digitalisation [related concepts in @cerveroTravelDemand3Ds1997].
 However, the framework presented has been criticized within academic and planning circles [e.g., @guzmanProximityEnoughCritical2024; @mouratidisTimeChallenge15minute2024] among other things for shortfalls in terms of addressing pre-existing structural forces and individual characteristics that drive inequalities and influence who benefits or could benefit the most from such an approach [@dimarino15minuteCityConcept2023; @willberg15minuteCityAll2023].
@@
 Reflective of the flexible and aspirational presentation of the 15-Minute City, the concept has been adopted by cities using a diverse range of definitions and tools along with indistinct universal approaches.
-A trail blazer in this respect was Portland (US) in the Portland Plan [@city_of_portland_20minute_2010] of April 2012.
+A trail blazer in this respect was Portland (U.S.) in the Portland Plan [@city_of_portland_20minute_2010] of April 2012.
 Adopted before the '15-Minute City' concept of today, this plan aimed to foster inclusive city development based on prosperity, education, health, and equity over a 30 year horizon.
@@
 The "cumulative opportunity" measure has been applied in many 15-Minute City examinations.
-This measure quantifies how many destinations can be reached from a point in space within a given travel time threshold , pairing well with normative examinations [see @paez_measuring_2012] that use travel time thresholds like x-minute cities [@logan_xminute_2022.
+This measure quantifies how many destinations can be reached from a point in space within a given travel time threshold, pairing well with normative examinations [see @paez_measuring_2012] that use travel time thresholds like x-minute cities [@logan_xminute_2022].
 Furthermore, the cumulative opportunity measure is widely appreciated for its intuitive computation and popularity among practitioners [@handyAccessibilityIdeaWhose2020; @handyMeasuringAccessibilityExploration1997; @chengInvestigatingWalkingAccessibility2019].
@@
 Classification algorithms also show promise as a tool in the assessment of the 15-Minute City, as they have been useful in identifying mobility and spatial typologies within the transportation planning literature.
-Often, these algorithms take the form of machine learning approaches: in transportation, the use of Self-Organizing Maps (SOM) was pioneered by Delmelle [-@delmelle_spatial_2012] and has been used to group U.S.
-neighbourhoods by minimizing dissimilarity in attributes [@delmelleDifferentiatingPathwaysNeighborhood2017] and to classify individuals' travel patterns according to the dissimilarity in mobility attributes and uses the Decision Tree algorithm to partition the data into interpretable classifications [@victorianoTimeSpaceMoney2020].
-Other approaches include the use of spatially constrained multivariate clustering, to develop urban form typologies related to the 15-Minute City [@burkeGeospatialAnalysisFramework2022] and the use of k-means to analyse travel behaviour and classify metro stations based on mobility patterns[ @ganUnderstandingUrbanMobility2020].
+Often, these algorithms take the form of machine learning approaches: in transportation, the use of Self-Organizing Maps (SOM) was pioneered by Delmelle [-@delmelle_spatial_2012].
+It has been used to group U.S.
+neighbourhoods by minimizing dissimilarity in attributes [@delmelleDifferentiatingPathwaysNeighborhood2017].
+SOM has also been used to classify individuals' travel patterns according to the dissimilarity in mobility attributes along with the Decision Tree algorithm to partition the data into interpretable classifications [@victorianoTimeSpaceMoney2020].
+Other approaches include the use of spatially constrained multivariate clustering, to develop urban form typologies related to the 15-Minute City [@burkeGeospatialAnalysisFramework2022] and the use of k-means to analyse travel behaviour and classify metro stations based on mobility patterns [@ganUnderstandingUrbanMobility2020].
 
@@
 From this research motivation, Sánchez de Madariaga coined the term Mobility of Care in 2013 to refer to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands [@sanchezdemadariagaMobilityCareIntroducing2013].
-While an undercurrent of research had examined these unique household-serving trips over the decades, her seminal work was the first to consider all these trips as one category and demonstrate how mobility of care is a significant proportion of daily travel.
+While an undercurrent of research had examined these unique household-serving trips over the decades, her seminal work was the first to consider all these trips as one category and demonstrate how Mobility of Care is a significant proportion of daily travel.
 
 The Mobility of Care concept also explicitly integrates inter-sectional equity considerations, a common criticism leveled at the 15-Minute City [@guzmanProximityEnoughCritical2024].
-Perhaps unsurprisingly given the gendered division of care work discussed, women have consistently been found to complete more mobility of care trips than men.
-In one study, mobility of care comprised 32% of women’s daily trips compared to 28% of men’s.
+Perhaps unsurprisingly given the gendered division of care work discussed, women have consistently been found to complete more Mobility of Care trips than men.
+In one study, Mobility of Care comprised 32% of women’s daily trips compared to 28% of men’s.
 While this gap is significant, it was found to be far greater in lower income households where women complete 20% more care trips than men [@ravensbergen2023exploratory].
-Sánchez de Madariaga not only shows how important these mobility of care trips are, but also highlights the ways in which “mobility of care is systematically under-represented in any analysis of urban transport” (p.
+Sánchez de Madariaga not only shows how important these Mobility of Care trips are, but also highlights the ways in which “Mobility of Care is systematically under-represented in any analysis of urban transport” (p.
 37).
-As a product of the masculinst bias in planning, transport surveys and tools often do not directly capture mobility of care, which re-enforces the idea that these trips are not a significant part of daily mobility.
+As a product of the masculinst bias in planning, transport surveys and tools often do not directly capture Mobility of Care, which re-enforces the idea that these trips are not a significant part of daily mobility.
 In this respect, the feminist perspective of the cities of proximity is still underestimated with only few examples on the topic [@gil_sola_choose_2022; @macintyre_her_2022].
@@
 Pairing well with the focus on shorter trips and the potential use of sustainable modes within the 15-Minute City, Mobility of Care trips are more local, shorter-distance and trip-chained.
-Compared to the trip to work, mobility of care trips are more frequently completed by foot, and less frequently by transit or bicycle [@ravensbergen2023exploratory].
+Compared to the trip to work, Mobility of Care trips are more frequently completed by foot, and less frequently by transit or bicycle [@ravensbergen2023exploratory].
 However, little work to date examines walking to care destinations through the Mobility of Care framework.
@@
 
-A spatial dataset of care destinations for Hamilton was compiled as detailed in ( _work removed for double-blind review_ ).
+A spatial dataset of care destinations for Hamilton was compiled as detailed in the forthcoming work of @soukhovforthcoming2024.
 The dataset includes 14 types of care destinations that were classified into four categories: dependent-centric (e.g., the destinations for child- and elder-centric escorting trips), grocery-centric, health-centric, and errand-centric.
-Notably, these categories were generated following the travel purpose categories created in the mobility of care research by @sanchezdemadariagaMeasuringMobilitiesCare2019.
-Care categories, sources of data, and preparation notes are detailed in @tbl-Tbl1.
+Notably, these categories were generated following the travel purpose categories created in the Mobility of Care research by @sanchezdemadariagaMeasuringMobilitiesCare2019.
+Care categories, sources of data, and descriptive notes are detailed in @tbl-Tbl1.
 The spatial distribution of destination type are visualised in @fig-Fig3 by their category.
@@
 Economic disadvantage tends to intersect with other identities such as gender [@lightmanMeasuringEconomicExclusion2018].
-And as reviewed in this work, all women and especially those from lower income households tend to complete most mobility of care trips [@ravensbergen2023exploratory].
+And as reviewed in this work, all women and especially those from lower income households tend to complete most Mobility of Care trips [@ravensbergen2023exploratory].
 Furthermore, lower-income households tend to also be single-parent households.

### 2024-10-10 paper/Manuscript.tex
Submitted to journal with editorial correspondance. With lock file update.

@@
 % Make \paragraph and \subparagraph free-standing
-\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}{
-    \@ifstar
-      \xxxParagraphStar
-      \xxxParagraphNoStar
-  }
-  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
-  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}{
-    \@ifstar
-      \xxxSubParagraphStar
-      \xxxSubParagraphNoStar
-  }
-  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
-  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
-\makeatother
 
@@
 \makeatother
-
 \ifLuaTeX
@@
 \hypersetup{
-  pdftitle={Caring 15-minute neighbourhoods},
+  pdftitle={Towards completely caring 15-minute neighbourhoods},
   pdfauthor={Anastasia Soukhov; Léa Ravensbergen; Lucía Mejía Dorantes; Antonio Páez},
@@
 \begin{frontmatter}
-\title{Caring 15-minute neighbourhoods}
+\title{Towards completely caring 15-minute neighbourhoods}
 \author[1]{Anastasia Soukhov%
@@
 destinations was compiled to estimate the number (using the cumulative
-opportunities accessibility measure) and diversity of mix (using the
+opportunity accessibility measure) and diversity of mix (using the
 entropy measure) of care destinations within a 15-minute walk from
@@
 
-Keywords: Accessibility; 15-Minute Neighborhoods; Chrono-Urbanism;
-Cities of Proximity; Mobility of Care; Inequality; Gender Gap;
+Keywords: 15-Minute City; 15-Minute Neighbourhoods; Chrono-Urbanism;
+Accessibility; Mobility of Care; Cities of Proximity; Inequality; Gender
+Gap
 
@@
 
-The term 15-Minute City is relatively new but the concept is not, as it
-is simply a normative cumulative-opportunities accessibility measure
-\citep{paez_measuring_2012}. Its newfound fame under the moniker of
-chronourbanism comes to complement long-standing efforts to foster
-community and local travel to amenities in neighbourhood planning
-practice. As recent examples, planned neighbourhoods dominated post-WWII
-built urban form: the Neighbourhood Unit Concept in the western world
+The term 15-Minute City is relatively new, but the concept is not. Its
+newfound fame under the moniker of chrono-urbanism comes to complement
+long-standing efforts to foster community and local travel to amenities
+in neighbourhood planning practice. As recent examples, planned
+neighbourhoods dominated post-WWII built urban form: the Neighbourhood
+Unit Concept in the western world
 \citep{brodyNeighbourhoodUnitConcept2013} and a parallel concept of the
@@
 \citep{silvaProximitycentredAccessibilityConceptual2023, silvaRegionalAccessibilityUndermining2022, guzmanProximityEnoughCritical2024}.
+In this way, the 15-Minute City could be understood to be a normative
+cumulative opportunity accessibility measure
+\citep{paez_measuring_2012}.
 
@@
 cultural factors that play a role in how, where, with whom, what for,
-and how far we move. Of note is that women tend to travel more
-frequently and for shorter distance
-\citep{roorda_triap_2010, morency_distance_2011} and have more
-activities related to care \citep{ilo_care_2018} no matter at what life
-stage they are \citep{garciaromanGenderDifferencesTime2022}. Combined,
+and how far we move. Women tend to travel shorter distances and more
+frequently \citep{roorda_triap_2010, morency_distance_2011}, and conduct
+more activities related to care no matter their stage of life
+\citep{ilo_care_2018, garciaromanGenderDifferencesTime2022}. Combined,
 these findings indicate that the spatial organization of activities has
@@
 This work builds a theoretical bridge between the 15-Minute City and
-mobility of care concepts to answer questions about what destinations
+Mobility of Care concepts to answer questions about what destinations
 matter and who benefits from proximity planning. The Mobility of Care
@@
 often occur at the local level. Moreover, care trips are of particular
-gendered and socio-economic import, as women, and especially those in
-lower income households, complete the majority of care trips
+gendered and socio-economic importance, as women, and especially those
+within lower income households, complete the majority of care trips
 \citep{ravensbergen2023exploratory}. Despite care travel's importance in
-advancing gender equality -and equity more broadly-, mobility of care
+advancing gender equality -and equity more broadly-, Mobility of Care
 and access to care destinations have been under-examined relative to
@@
 spatial proximity to essential services, and digitalisation
-\citep[q.v.][]{cerveroTravelDemand3Ds1997}. However, the framework
-presented has been criticized within academic and planning circles
+\citep[related concepts in][]{cerveroTravelDemand3Ds1997}. However, the
+framework presented has been criticized within academic and planning
+circles
 \citep[e.g.,][]{guzmanProximityEnoughCritical2024, mouratidisTimeChallenge15minute2024}
@@
 range of definitions and tools along with indistinct universal
-approaches. A trail blazer in this respect was Portland (US) in the
+approaches. A trail blazer in this respect was Portland (U.S.) in the
 Portland Plan \citep{city_of_portland_20minute_2010} of April 2012.
@@
 quantifies how many destinations can be reached from a point in space
-within a given travel time threshold , pairing well with normative
+within a given travel time threshold, pairing well with normative
 examinations \citep[see][]{paez_measuring_2012} that use travel time
-thresholds like x-minute cities {[}\citet{logan_xminute_2022}.
-Furthermore, the cumulative opportunity measure is widely appreciated
-for its intuitive computation and popularity among practitioners
+thresholds like x-minute cities \citep{logan_xminute_2022}. Furthermore,
+the cumulative opportunity measure is widely appreciated for its
+intuitive computation and popularity among practitioners
 \citep{handyAccessibilityIdeaWhose2020, handyMeasuringAccessibilityExploration1997, chengInvestigatingWalkingAccessibility2019}.
@@
 transportation, the use of Self-Organizing Maps (SOM) was pioneered by
-Delmelle \citeyearpar{delmelle_spatial_2012} and has been used to group
+Delmelle \citeyearpar{delmelle_spatial_2012}. It has been used to group
 U.S. neighbourhoods by minimizing dissimilarity in attributes
-\citep{delmelleDifferentiatingPathwaysNeighborhood2017} and to classify
-individuals' travel patterns according to the dissimilarity in mobility
-attributes and uses the Decision Tree algorithm to partition the data
-into interpretable classifications \citep{victorianoTimeSpaceMoney2020}.
-Other approaches include the use of spatially constrained multivariate
-clustering, to develop urban form typologies related to the 15-Minute
-City \citep{burkeGeospatialAnalysisFramework2022} and the use of k-means
-to analyse travel behaviour and classify metro stations based on
-mobility patterns\citep{ganUnderstandingUrbanMobility2020}.
+\citep{delmelleDifferentiatingPathwaysNeighborhood2017}. SOM has also
+been used to classify individuals' travel patterns according to the
+dissimilarity in mobility attributes along with the Decision Tree
+algorithm to partition the data into interpretable classifications
+\citep{victorianoTimeSpaceMoney2020}. Other approaches include the use
+of spatially constrained multivariate clustering, to develop urban form
+typologies related to the 15-Minute City
+\citep{burkeGeospatialAnalysisFramework2022} and the use of k-means to
+analyse travel behaviour and classify metro stations based on mobility
+patterns \citep{ganUnderstandingUrbanMobility2020}.
 
@@
 trips over the decades, her seminal work was the first to consider all
-these trips as one category and demonstrate how mobility of care is a
+these trips as one category and demonstrate how Mobility of Care is a
 significant proportion of daily travel.
@@
 the gendered division of care work discussed, women have consistently
-been found to complete more mobility of care trips than men. In one
-study, mobility of care comprised 32\% of women's daily trips compared
+been found to complete more Mobility of Care trips than men. In one
+study, Mobility of Care comprised 32\% of women's daily trips compared
 to 28\% of men's. While this gap is significant, it was found to be far
@@
 trips than men \citep{ravensbergen2023exploratory}. Sánchez de Madariaga
-not only shows how important these mobility of care trips are, but also
-highlights the ways in which ``mobility of care is systematically
+not only shows how important these Mobility of Care trips are, but also
+highlights the ways in which ``Mobility of Care is systematically
 under-represented in any analysis of urban transport'' (p.~37). As a
 product of the masculinst bias in planning, transport surveys and tools
-often do not directly capture mobility of care, which re-enforces the
+often do not directly capture Mobility of Care, which re-enforces the
 idea that these trips are not a significant part of daily mobility. In
@@
 more local, shorter-distance and trip-chained. Compared to the trip to
-work, mobility of care trips are more frequently completed by foot, and
+work, Mobility of Care trips are more frequently completed by foot, and
 less frequently by transit or bicycle
@@
 A spatial dataset of care destinations for Hamilton was compiled as
-detailed in ( \emph{work removed for double-blind review} ). The dataset
-includes 14 types of care destinations that were classified into four
-categories: dependent-centric (e.g., the destinations for child- and
-elder-centric escorting trips), grocery-centric, health-centric, and
+detailed in the forthcoming work of \citet{soukhovforthcoming2024}. The
+dataset includes 14 types of care destinations that were classified into
+four categories: dependent-centric (e.g., the destinations for child-
+and elder-centric escorting trips), grocery-centric, health-centric, and
 errand-centric. Notably, these categories were generated following the
-travel purpose categories created in the mobility of care research by
+travel purpose categories created in the Mobility of Care research by
 \citet{sanchezdemadariagaMeasuringMobilitiesCare2019}. Care categories,
-sources of data, and preparation notes are detailed in
+sources of data, and descriptive notes are detailed in
 Table~\ref{tbl-Tbl1}. The spatial distribution of destination type are
@@
 work, all women and especially those from lower income households tend
-to complete most mobility of care trips
+to complete most Mobility of Care trips
 \citep{ravensbergen2023exploratory}. Furthermore, lower-income

### 2025-01-30 data/02-access-and-prelim-plots.qmd
R1 submitted.

@@
 ```{r}
-load("data-raw/parcels/R_PARCELS_CENTS_2020.rda") #add the census id to the parcels, then the ttm
-
+load("data-raw/parcels/R_PARCELS_2020.rda") #add the census id to the parcels, then the ttm
+R_PARCELS_CENTS_2020 <- st_centroid(R_PARCELS_2020)
 #first join it to census.
@@
 care_categories_plot <- 
+  tm_shape(Community_Boundaries, bbox=Community_Boundaries) + tm_polygons(col = "white", border.alpha=0)+
   tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
   tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+ 
-  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
-  tm_shape(Community_Boundaries, bbox=Community_Boundaries) + tm_polygons(col = "white", border.col = "black", lwd=1)+
-  tm_shape(care_dest) + tm_dots("Care_Category_Specific", palette = RColorBrewer::brewer.pal(n = 4, name = 'Dark2'), size=0.025, title = "")+
+    tm_shape(Community_Boundaries, bbox=Community_Boundaries) + tm_polygons(col = NA, alpha=0, border.col = "black", lwd=1)+
+  tm_shape(care_dest) + tm_symbols(shape="Care_Category_Specific", size=0.03, border.lwd = 0, shapes = c(21,22,23,24,25,15), col = "#E6AB02", title.shape = "")+
+  # tm_shape(care_dest) + tm_dots("Care_Category_Specific", palette = RColorBrewer::brewer.pal(n = 4, name = 'Dark2'),
+  #                              size=0.025, title = "")+
   tm_facets("Care_Category", nrow = 2, free.scales= TRUE)+ 
@@
   tm_compass(position = c("left", "top"), size=0.9) +
-  tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 1, legend.bg.color = "white", legend.bg.alpha = 0.4,
+  tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), scale = 1.1, legend.bg.color = "white", legend.bg.alpha = 0.65,
             panel.labels = c("1,265 Dependent-centric destinations", 
@@
             panel.label.size = 1.1) 
-
+#care_categories_plot
 tmap_save(care_categories_plot,
           file= file.path(here("figures/care_categories_plot.png")),
-          dpi=300)
+          dpi=600)
 ```

### 2025-01-30 data/03-clustering-and-profiling.qmd
R1 submitted.

@@
               title = "% hld below AT median income",
-              legend.is.portrait = TRUE) +
+              legend.is.portrait = TRUE, 
+              colorNA = "white",
+              textNA= "Unavailable") +
   tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
@@
               title = "% population aged 0 to 14",
-              legend.is.portrait = TRUE) +
+              legend.is.portrait = TRUE, 
+              colorNA = "white",
+              textNA= "Unavailable") +
   tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
@@
               title = "% one parent hlds",
-              legend.is.portrait = TRUE) +
+              legend.is.portrait = TRUE, 
+              colorNA = "white",
+              textNA= "Unavailable") +
   tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
@@
               title = "Prevalance of household LICO-AT",
-              legend.is.portrait = TRUE) +
+              legend.is.portrait = TRUE, 
+              colorNA = "white",
+              textNA= "Unavailable") +
   tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
@@
 ```{r, eval=FALSE}
-dep_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+dep_copp_plot <- 
   tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
-  tm_shape(access_15_DA) +
+  tm_shape(access_15_DA |> mutate(med_dep_copp = ifelse(med_dep_copp == 0, NA, med_dep_copp))) +
   tm_polygons("med_dep_copp",
@@
               palette = "Oranges",
-              title = "Dependent-centric",) +
+              title = "Dependent-centric",
+              colorNA = "white",
+              textNA = 0,
+              border.col = NULL) +
+  tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
   tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
@@
   tm_compass(position = c("left", "top"), size=1.0)+
-  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4,
+  tm_layout(legend.bg.color = "white", legend.bg.alpha = 0.4, scale=0.8,
             bg.color = "grey",
@@
 
-health_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+health_copp_plot <- 
   tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
-  tm_shape(access_15_DA) +
+  tm_shape(access_15_DA |> mutate(med_health_copp = ifelse(med_health_copp == 0, NA, med_health_copp))) +
   tm_polygons("med_health_copp",
@@
               palette = "Purples",
-              title = "Health-centric",) +
+              title = "Health-centric",
+              colorNA = "white",
+              textNA = 0,
+              border.col = NULL) +
+  tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
   tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
@@
 
-groc_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+groc_copp_plot <-
   tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
-  tm_shape(access_15_DA) +
+  tm_shape(access_15_DA |> mutate(med_groc_copp = ifelse(med_groc_copp == 0, NA, med_groc_copp))) +
   tm_polygons("med_groc_copp",
@@
               palette = "Greens",
-              title = "Grocery-centric",) +
+              title = "Grocery-centric",
+              colorNA = "white",
+              textNA = 0,
+              border.col = NULL) +
+  tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
   tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
@@
 
-err_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+err_copp_plot <- 
   tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
-  tm_shape(access_15_DA) +
+  tm_shape(access_15_DA |> mutate(med_err_copp = ifelse(med_err_copp == 0, NA, med_err_copp))) +
   tm_polygons("med_err_copp",
@@
               palette = "Reds",
-              title = "Errand-centric",) +
+              title = "Errand-centric",
+              colorNA = "white",
+              textNA = 0,
+              border.col = NULL) +
+  tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
   tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
@@
 ```{r, eval=FALSE}
-all_cats_entropy_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
+all_cats_entropy_copp_plot <- 
   tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
-  tm_shape(access_15_DA) +
+  tm_shape(access_15_DA |> mutate(med_div = ifelse(med_div == 0, NA, med_div))) +
   tm_polygons("med_div",
               border.alpha = 0.2,
-              style = "quantile", n=3, #are the quantiles, but are also similar to the clustered profiles
-              label = c("[0 to 0.49) Not complete", "[0.49 to 0.85) Complete", "[0.85 to 1] Very complete"),
+              style = "quantile", n=4, #are the quantiles, but are also similar to the clustered profiles
+              #label = c("[0 to 0.49) Not complete", "[0.49 to 0.85) Complete", "[0.85 to 1] Very complete"),
               palette = "Greys",
-              title = "Diversity of care destinations within a 15-minute walk for a median parcel",
-              legend.is.portrait = TRUE) +
+              title = "Diversity of care destinations within \na 15-minute walk for a median parcel",
+              legend.is.portrait = TRUE,
+              colorNA = "white",
+              textNA = 0,
+              border.col = NULL) +
   tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
+  tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
   tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
@@
             panel.label.color = "black",
-            legend.outside = TRUE,
-            legend.position = c("center", "bottom"),
-            legend.outside.position = "bottom",
-            legend.outside.size = 0.2,
-            legend.width = 12)
+            #legend.outside = TRUE,
+            legend.position = c("right", "top"))
+            #legend.outside.position = "bottom",
+            #legend.outside.size = 0.2,
+            # legend.width = 12)
 

### 2025-01-30 paper/Manuscript.qmd
R1 submitted.

@@
         \linenumbers
+        \floatplacement{figure}{H}
     journal:
@@
       cite-style: authoryear
-keywords: ["Accessibility", "Mobility of Care", "Gender", "15-minute neighbourhood", "Care Complete Neighbourhoods"]
+keywords: ["15-Minute City", "Chrono-Urbanism","Accessibility", "Mobility of Care", "Inequality", "Gender Gap"]
 bibliography: bibliography.bib
@@
 
+```{r to calc perc urbanised parcels in each scluster}
+load(glue::glue(here::here(), "/data/data-raw/parcels/R_PARCELS_2020.rda"))
+R_PARCELS_2020 <- R_PARCELS_2020 |> select(c("ID","URBAN_RURA")) |> st_drop_geometry()
+R_PARCELS_2020$ID <- as.character(R_PARCELS_2020$ID)
+
+access_15_urban_summary <- access_15_wide |> 
+  select(c("from_id","scluster")) |>
+  left_join(R_PARCELS_2020,by=c("from_id"= "ID"))
+
+access_15_urban_summary <- access_15_urban_summary |> group_by(scluster,URBAN_RURA) |> 
+  summarise(count_urb_parcels = n(), 
+            .groups = "drop")
+```
+
 # Abstract
@@
 In contrast, not caring 15-minute neighbourhoods in higher-income peripheral areas.
-Areas that make good candidates for urban policy intervention are identified, and the implications of enhancing 15-minute walkable caring access are discussed in relation to gender mainstreaming in transportation planning.
+Areas that make good candidates for urban policy intervention are identified and the implications of enhancing 15-minute walkable caring access are discussed in relation to gender mainstreaming in transportation planning and limitations of this work.
 
@@
 
-From another perspective, feminist geographers, urban planners and other researchers agree that urban structure influences travel, but city planning falls short in planning for a neutral identity.
+An increasing number of human geographers, planners and other researchers agree that urban structure influences travel, but city planning falls short in planning for a neutral identity.
 In contrast, cities ought to be planned according to the multiple identities of the inhabitants [@vacchelli_towards_2018; @urbandevelopmentviennaGenderMainstreamingUrban2013].
-Researchers have demonstrated how mobility patterns differ according to gender [@lawWomenTransportNew1999; @cresswell_gendered_2008; @levy2013travel; @little_gender_1994; @tronto_toward_1990; @neto_gender_2015].
+Researchers have demonstrated how mobility patterns differ according to gender [@lawWomenTransportNew1999; @cresswell_gendered_2008; @levy2013travel; @little_gender_1994; @tronto_toward_1990].
 The causes for this are varied, and related to gendered social norms and cultural factors that play a role in how, where, with whom, what for, and how far we move.
@@
 Moreover, care trips are of particular gendered and socio-economic importance, as women, and especially those within lower income households, complete the majority of care trips [@ravensbergen2023exploratory].
-Despite care travel's importance in advancing gender equality -and equity more broadly-, Mobility of Care and access to care destinations have been under-examined relative to employment destinations.
+Despite care travel's importance in advancing gender equality -and equity^[The United Nations defines "gender equality" as equal rights, responsibilities, and opportunities for all genders, regardless of gender at birth [@unwomenHandbookGenderMainstreaming2022]; they prefer "gender equality," as "equity" relates to the concept of fairness that are usually based on traditions, customs, or cultural beliefs which can perpetuate disadvantage.
+However, "equity," in the broader context of social justice, is widely accepted.
+As defined in the UNICEF Glossary of DEI terms, equity is the "process of being fair to all individuals and groups, by addressing present and historical inequality in order to work towards equality in outcomes" and may involve temporary special measures to address systemic discrimination faced by marginalized groups [@unicefAnnexDIVERSITYEQUITY2024].] more broadly-, Mobility of Care and access to care destinations have been under-examined relative to employment destinations.
 This is especially pertinent in accessibility research that has largely focused on employment points of interest e.g., [@farberOntarioLineSocioeconomic2019; @duarteInfluenceJobAccessibility2023; @ryanAccessibilitySpaceTime2023].
@@
 
-# Data
+# Case study and data
 
@@
 The Niagara Escarpment runs through Hamilton, and results in a city with two key elevations: a more dense lower city that contains the downtown core and the elevated suburban development referred to as ‘the Mountain’.
-In this work, we analyse the residential parcel centroids, 143,882 locations.
+In this work, we analyse the residential parcel centroids, 143,882 locations in the city [@Teranet_2021].
 We aggregated the points at the level of Canadian Census Dissemination Area (DA) along with the population and population per parcel plots in @fig-Fig1.
+The DA is the most spatially granual level of publicly available Canadian census data.
+DAs are created by Statistics Canada for the purpose of the census and each represent between 400 to 700 people (Q1:442 and Q3:664 from @fig-Fig1).
 
@@
 ```
-
-\newpage
 
@@
     i = 1, j = c(1,2,3), part = "header", value = as_paragraph(c("Care category", "Care destination types", "Sources"))) %>%
-  width(j = 1, width = 1.7, unit = "cm") %>%
-  width(j = 2, width = 10, unit = "cm") %>%
-  width(j = 3, width = 5.3, unit = "cm")|>
+  width(j = 1, width = 3, unit = "cm") %>%
+  width(j = 2, width = 9, unit = "cm") %>%
+  width(j = 3, width = 5, unit = "cm")|>
   flextable::line_spacing(space = 0,
                           part = "body")
+
 ```
@@
 
-## Accessibility: the cumulative opportunity measure
+## The cumulative opportunity measure of accessibility
 
-To capture the quantity of access to each type of destinations, a cumulative opportunity accessibility score $S^t_i$ is calculated.
+To capture the quantity of spatial access to each type of destinations, a cumulative opportunity accessibility score $S^t_i$ is calculated.
 Scores for each of the 14 care destination types $t$ is calculated for every parcel $i$.
-The calculation takes the following mathematical form:
+The calculation takes the mathematical form in @eq-access-cumulative:
 
@@
 S^t_i=\sum_{j}O_j^{t}\cdot f(c_{ij})
-$$
+$${#eq-access-cumulative}
 \noindent Where:
@@
 - $f(\cdot)$ is an impedance function of $c_{ij}$; within the cumulative opportunity approach, it is a binary function that takes the value of 1 if $c_{ij}$ is less than a selected value [@handyMeasuringAccessibilityExploration1997].
-- $S_{i}^t$ is the cumulative opportunity accessibility score, the sum of weighted opportunities reachable within $f(\cdot)$, at each $i$ for each $t$.
+- $S_{i}^t$ is the cumulative opportunity accessibility score, the sum of opportunities reachable within $f(\cdot)$, at each $i$ for each $t$.
 
@@
 D_{i} = \frac{-\sum_{t} [S_i^{t} / \sum_{t}S^{t}_{i} \times ln(S_i^{t} / \sum_{t}S^{t}_{i} )]}{ln(n_t)}
-$$
+$${#eq-entropy-measure}
 
@@
 - $n_t$ is the count of care destination types $t$. In this work, this value is 14.
-- $S_{i}^t$ is the cumulative opportunity accessibility score, the sum of weighted opportunities reachable within a 15-minute walk from $i$.
+- $S_{i}^t$ is the cumulative opportunity accessibility score, the sum of opportunities reachable within a 15-minute walk from $i$.
 - $D_i$ is the diversity score.
@@
 ## Overview of access to care destinations
-How many care destinations can be reached within a 15-minute walk is summed for each of the four care categories defined in @tbl-Tbl1.
+The number of care destinations that can be reached within a 15-minute walk is summarised for each of the four care categories (as defined in @tbl-Tbl1).
 The median parcel value for each DA is visualised in @fig-Fig4 and the median parcel diversity of care destinations accessible is presented in @fig-Fig5.
-Together, @fig-Fig4 and @fig-Fig5 represent summaries of 15 variables that served as inputs into the SOM algorithm.
+Together, @fig-Fig4 and @fig-Fig5 represent summaries of 15 variables were inputs into the SOM algorithm.
 
@@
 #| fig-cap: "The number of care destinations that can be reached within a 15-min walk per care category for a median parcel in each DA.
-The values are a summary of the 14 acessibility scores that served as inputs into the SOM.
-Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
+The values are a summary of the 14 acessibility scores that are inputs into the SOM.
+Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010].
+Scale is in quartiles."
 knitr::include_graphics(glue::glue(here::here(), "/figures/percats_copp_plot.png"))
@@
 These values are also a summary of the 15th input into the SOM.
-Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
+Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010].
+Scale is in quartiles."
 knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_entropy_copp_plot.png"))
@@
 
-\newpage
-
 ## Identification of completely caring 15-minute neighbourhood typologies
 
-In @fig-Fig4, the concentration of opportunities for all category types is centered in the downtown of Hamilton, with the highest concentration being in the lower city (near the lake shore) and pockets of the high concentration further from the shoreline.
-Grocery-centric destinations appear to be the most concentrated, followed by health-centric and dependent-centric caring destinations.
-Errand-centric destinations are the most sprawled.
-In many ways, @fig-Fig4 mirrors the spatial distribution of care destinations (@fig-Fig3) as the range of 15-minutes of travel is small.
-In @fig-Fig5, areas that have high care accessibility tend to have high diversity as well, though there are exceptions in pockets of the city outside the downtown core that have higher diversity but low levels of accessibility for all care categories.
+In @fig-Fig4, for all care categories, access to opportunities is concentrated in downtown Hamilton (Hamilton Central), with the highest concentration (top quantiles) near the lake shore of Hamilton Central.
+Grocery-centric destinations follow this trend most prominently, followed by health-centric and dependent-centric caring destinations.
+Access to errand-centric destinations appears to most spatially homogeneous of the four care destination types, though the top quantiles are still concentrated in Hamilton Central.
+As the 15-minute walking accessibility to destinations is a spatially weighted summary measure of the count of destinations, @fig-Fig4 mirrors the spatial distribution in @fig-Fig3.
+In @fig-Fig5, areas that have high diversity are broadly located in downtown Hamilton as well, however, there are exceptions.
+Namely, there are pockets of high diversity outside the downtown core that that score low levels of accessibility for all care categories.
 Similarly, there are areas with low diversity within the downtown core that have only moderate or high accessibility to certain care destination types.
 
-Based on the SOM methodology discussed, 7 superclusters are identified from the 10 by 10 grid of SOM nodes.
-As diagnostics for the selection of the number of superclusters, the dissimilarity-index-based dendrogram with the proportion of parcels represented in each supercluster alongside the variance explained plot is visualised in @fig-Fig6.
-Labels representing grades A to D qualifying the 7 superclusters are then assigned.
-These grades as also reflected in @fig-Fig6 and represent the quantity of accessibility and diversity in accessibility per care category.
-Higher grades (A+ and A) corresponding to the highest accessibility and diversity, while lowest grades (D) representing lowest accessibility and diversity scores.
+Based on the SOM methodology discussed and using inputs summarised in @fig-Fig4 and @fig-Fig5, seven superclusters are identified as shown in the dissimilarity-index-based dendrogram (left plot of @fig-Fig6).
+Seven superclusters are selected to represent the data, as they explain a high amount of the variance (over 90%; right plot in @fig-Fig6).
+Each supercluster can be intuitively labelled based on the quantity of accessibility and diversity in accessibility per care category offered by a parcel, as mapped onto the SOM input grid.
+Labels representing grades A through D are assigned to qualify the seven superclusters.
+Higher grades (A+ and A) corresponding to the highest accessibility and diversity, while lowest grades (D) represent the lowest accessibility and diversity scores.
 
@@
 #| label: fig-Fig6
-#| fig-cap: "The 7 resulting superclusters from the SOM the 10 by 10 grid output represented in a dendrogram (left) and the proportion of variance unexplained by cluster (right)."
+#| fig-cap: "The seven resulting superclusters from the SOM output are represented in the dendrogram (left).
+As a diagnostic, the proportion of variance unexplained by the number superclusters is displayed on the right, with red lines indicating seven superclusters."
 knitr::include_graphics(glue::glue(here::here(), "/figures/all_access_sc_plot_small.png"))
@@
 
-\newpage
+More specifically, each supercluster was labelled a grade by the authors by referring to the supercluster's descriptive statistics, as summarised in @tbl-Tbl2 and the boxplots in @fig-Fig7.
+Additionally, the grades for each supercluster are further explained in context of the proportion of urban, suburban and rural parcels as classified by the City of Hamilton within the parcel dataset [@Teranet_2021].
+Notably, the majority of parcels in Hamilton are considered urban (`r ((access_15_urban_summary |> filter(URBAN_RURA == "U") |> select(count_urb_parcels) |> sum()) / (access_15_urban_summary  |> select(count_urb_parcels) |> sum() )) |> scales::percent(0.1)`), while suburban and rural parcels constitute the remaining `r ((access_15_urban_summary |> filter(URBAN_RURA == "S") |> select(count_urb_parcels) |> sum()) / (access_15_urban_summary  |> select(count_urb_parcels) |> sum() )) |> scales::percent(0.1)` and `r ((access_15_urban_summary |> filter(URBAN_RURA == "R") |> select(count_urb_parcels) |> sum()) / (access_15_urban_summary  |> select(count_urb_parcels) |> sum() )) |> scales::percent(0.1)`, respectively:
 
-The grade for each supercluster was carefully assigned by the authors by referring to descriptive statistics and their notable trends.
-These statistics are summarised in @tbl-Tbl2 and in the boxplots in @fig-Fig7.
+<!--"(A+) Exceptional: very caring & very complete", #C1
+"(A ) Excellent: very caring & very complete ", #(C2)
+"(A-) Dep- heavy, otherwise excellent: very caring & not complete", #(C3)
 
-- A+ and A have exceptionally high caring accessibility for all destination types and high diversity scores in the top quantile or above.
-Together, these grades represent 24% of all parcels in the city.
+"(B+) Above average: caring & very complete", # (C6)
+"(B ) Average: caring & complete", #(C5)
+"(B- ) Dep- heavy, otherwise below average: somewhat caring & not complete", # (C4)
+
+"(D ) Below average: not caring & not complete") # (C7)-->
+
+- A+ and A represent superclusters with exceptionally high caring accessibility for all destination types and high diversity scores in the top quantile or above.
+Together, these grades represent 24% of all parcels in the city and `r ((access_15_urban_summary |> filter((scluster == 1 | scluster == 2) & (URBAN_RURA == "U")) |> select(count_urb_parcels) |> sum()) / (access_15_urban_summary |> filter((scluster == 1 | scluster == 2)) |> select(count_urb_parcels) |> sum() )) |> scales::percent()` are located in urban neighbourhoods.
+Based on the spatial access and diversity of care destination access offered by these neighbourhoods, they could be called "very caring" and "very complete".
 - A- has high caring accessibility, but low diversity.
+These neighbourhoods could be classified as "very caring" but and "not complete".
 Interestingly, A- is like A but with much higher dependent-centric destinations, particularly parks, schools, and daycares, with only moderately high scores for all other destination types.
-This disbalance lowers its overall diversity score.
-However, it can be characterized as a supercluster that demonstrates potential in being retrofitted to provide A+ or A level of completely caring access.
-4% of all parcels are represented by the A- grade.
-- B+ and B superclusters present about average completely caring access.
+This disbalance contributes to reducing its overall diversity score.
+However, it could be characterized as a supercluster that demonstrates potential in being retrofitted to provide A+ or A level of completely caring access.
+4% of all parcels are represented by the A- grade and, similar to parcels in A+ and A superclusters, `r ((access_15_urban_summary |> filter((scluster == 3) & (URBAN_RURA == "U")) |> select(count_urb_parcels) |> sum()) / (access_15_urban_summary |> filter((scluster == 3)) |> select(count_urb_parcels) |> sum() )) |> scales::percent()` are located in urban neighbourhoods.
+- B+ and B superclusters present about average "completely caring" access.
 These superclusters can serve as the benchmark for what 'average' completely caring access in Hamilton currently looks like.
-These grades represent 15% of all parcels.
+These grades represent 15% of all parcels and like parcels in A+, A and A- superclusters, `r ((access_15_urban_summary |> filter((scluster == 6 | scluster == 5) & (URBAN_RURA == "U")) |> select(count_urb_parcels) |> sum()) / (access_15_urban_summary |> filter((scluster == 6 | scluster == 5)) |> select(count_urb_parcels) |> sum() )) |> scales::percent()` are located in urban neighbourhoods.
 - B- provides above average dependent-centric destination access, particularly parks, daycares and schools, but below average access to other destination types and hence has low diversity scores.
+B- neighbourhoods could be labelled "somewhat caring" but "not complete".
 B- is like A- as it demonstrates complete caring 15-minute potential if retrofitted.
@@
 This grade represents 19% of all parcels.
+While the majority of parcels are still classified as urban neighbourhoods (`r ((access_15_urban_summary |> filter((scluster == 4) & (URBAN_RURA == "U")) |> select(count_urb_parcels) |> sum()) / (access_15_urban_summary |> filter((scluster == 4)) |> select(count_urb_parcels) |> sum() )) |> scales::percent()`), this supercluster represents `r ( (access_15_urban_summary |> filter((scluster == 4) & (URBAN_RURA == "S")) |> select(count_urb_parcels) |> sum()) / (access_15_urban_summary |> filter(URBAN_RURA == "S") |> select(count_urb_parcels) |> sum()) ) |> scales::percent(0.1)` of the suburban parcels and `r ( (access_15_urban_summary |> filter((scluster == 4) & (URBAN_RURA == "R")) |> select(count_urb_parcels) |> sum()) / (access_15_urban_summary |> filter(URBAN_RURA == "R") |> select(count_urb_parcels) |> sum()) ) |> scales::percent(0.1)` rural parcels in the city.
+These suburban and rural parcels are in proximity to urban parcels, as elaborated in the next subsection.
 - D superclusters demonstrate the lowest scores all-around, representing room for land-use improvement that addresses complete and caring 15-minute access.
-This supercluster characterizes the largest number of parcels, representing 38% of parcels in the city.
+D neighbourhoods could be classified as "not caring" and "not complete" due to the low amount of access to all care destinations offered and the care category diversity of this access.
+Notably, this supercluster characterizes the largest number of parcels, representing 38% of parcels in the city.
+Only `r ((access_15_urban_summary |> filter((scluster == 7) & (URBAN_RURA == "U")) |> select(count_urb_parcels) |> sum()) / (access_15_urban_summary |> filter((scluster == 7)) |> select(count_urb_parcels) |> sum() )) |> scales::percent()` of these parcels are 'urban', and this supercluster represents the vast majority of suburban and rural parcels in the city (`r ( (access_15_urban_summary |> filter((scluster == 7) & (URBAN_RURA == "S")) |> select(count_urb_parcels) |> sum()) / (access_15_urban_summary |> filter(URBAN_RURA == "S") |> select(count_urb_parcels) |> sum()) ) |> scales::percent(0.1)` and `r ( (access_15_urban_summary |> filter((scluster == 7) & (URBAN_RURA == "R")) |> select(count_urb_parcels) |> sum()) / (access_15_urban_summary |> filter(URBAN_RURA == "R") |> select(count_urb_parcels) |> sum()) ) |> scales::percent(0.1)`).
 
@@
 
-\newpage
+To spatially demonstrate where the superclusters are located, representative supercluster grades for each DA are visualised in @fig-Fig8.
+This visualisation is created by grouping parcels by their DA and selecting the grade that is most dominant within that DA.
+For reference, the median number of parcels in each DA is `r df_DA$parcel_count |> median() |> round()` and supercluster grade membership within a DA is typically pure, with the median membership being `r df_DA$max_label_val |> median() |> scales::percent()` of a single cluster.
 
-To spatially demonstrate where the superclusters are located, representative supercluster labels for each DA are visualised in @fig-Fig8.
-This visualisation is created by grouping parcels by their DA, calculating the median supercluster label for each label, and selecting the label that is most dominant within that DA.
-For reference, the median number of parcels in each DA is `r df_DA$parcel_count |> median() |> round()` and supercluster label membership within a DA is typically pure, with the median membership being `r df_DA$max_label_val |> median() |> scales::percent()` of a single cluster.
-
-```{r plot-figure8, out.width=450, out.height=450}
+```{r plot-figure8, out.width=800, out.height=450}
 #| label: fig-Fig8
@@
 
-The resulting superclusters visualised in @fig-Fig8 are a combination of the cumulative opportunity and diversity of access measures translated into meaningful typologies through the SOM methodology.
-To reiterate, accessibility is the measure of _potential_ interaction, e.g., how many care destinations one could reach within a 15-minute walk, these typologies are useful in identifying which areas of the city are providing regionally high levels of complete caring access, which areas are regionally average, and which areas are regionally below average.
-Furthermore, neighbourhoods which provide disbalanced access can be further investigated.
+Recall, the superclusters in @fig-Fig8 are a combination of the cumulative opportunity and diversity of access measures translated into meaningful typologies through the SOM methodology.
+To reiterate, accessibility is the measure of _potential_ interaction, e.g., how many care destinations one could reach within a 15-minute walk, these typologies are useful in identifying which areas of the city are providing city-wide high levels of complete caring access, which areas are city-wide average, and which areas are city-wide below average.
 
 In @fig-Fig8, it can be observed that the excellent and exceptional neighbourhoods (A, A+), which are very caring and very complete 15-minute neighbourhoods, are located within the center of the city and closest to the shoreline in the downtown core of Hamilton.
-The above and below average neighbourhoods (B+ B) are often proximate to the excellent and exceptional neighbourhoods and are more prevalent in south of the escarpment within the center Hamilton.
+The above and below average neighbourhoods (B+, B) are often proximate to the excellent and exceptional neighbourhoods and are more prevalent in south of the escarpment within the center Hamilton.
 The escarpment is a physical barrier, with few pedestrian-accessible access points to traverse; hence the typologies describing neighbouring DAs separated by the escarpment are often different.
-Below average (D) neighbourhoods are located in peripheral areas outside the center of Hamilton, in areas where urban form is characterised by lower density residential housing and auto-dependent mobility.
+Below average (D) neighbourhoods are located in peripheral areas outside the center of Hamilton, in areas where urban form is characterised by lower density residential housing and auto-dependent mobility (majority rural and suburban).
+D neighbourhoods also tend to have the lowest diversity scores, with either no access or access to only one type of care destination category.
 
-Furthermore in @fig-Fig8, A- and B- grades stand out as offering high access for children-centric destinations for their grade-group but below grade-group average access in other destination types.
-These neighbourhoods may be more suitable for populations who prioritize walkable access to destinations like schools, parks and daycares, and find access to other types of caring destinations less important.
+Furthermore in @fig-Fig8, A- and B- grades stand out as offering high access for children-centric destinations for their grade-group but below grade-group average access in other destination types (e.g., low or lowest diversity scores).
+These neighbourhoods may be more suitable for populations who prioritize walkable access to children-centric destinations like schools, parks and daycares, and find access to other types of caring destinations less important.
 These neighbourhoods also stand out as demonstrating potential to be retrofitted to provide more complete access if additional destination types were located within their neighbourhoods.
@@
 For reference, the city-wide mean household income is `r df_DA_nogeom$Income |> mean(na.rm=T)|> round() |> prettyNum(big.mark = ",") ` (SD: `r df_DA_nogeom$Income |> sd(na.rm=T)|> round() |> prettyNum(big.mark = ",")` and median: `r quantile(df_DA_nogeom$Income,0.5, na.rm=T) |> unname()  |> prettyNum(big.mark = ",")`).
-The notable splits for income are >\$91,500, >\$68,750, or between \$68,750 and \$91,500; roughly split by lower income, middle income, and higher income median household brackets.
+The notable splits for income are >\$91,500, <\$68,750, or between \$68,750 and \$91,500; roughly split by higher income, lower income, and middle income median household brackets.
 
@@
 
-Describing @fig-Fig9, the bar column on the right represents DAs with higher household income.
+Describing @fig-Fig9, the bar column on the right represents DAs with higher household income (>\$91,500 median household income).
 In these DAs, the majority of parcels are labeled as D superclusters, i.e not caring or complete 15-minute neighbourhoods.
@@
 B- parcels are low in most amenity types except a few child-centric destinations and show promise in being more easily improved than the D parcels.
-The higher income column represents 39% of the population, the largest proportion of any of the three columns.
-The left column in @fig-Fig9 corresponds to the lowest income, i.e., below 68,750 CAD median household income, and accounts for 32% the population.
-It is dominated by A+ parcels along with A and A- parcels, in the largest quantities relative to other bar columns, however, proportions of all other superclusters are also present.
-The middle column is defined by parcels with a mix of supercluster classifications and represents DAs with middle households incomes, between 68,750 CAD and 91,500 CAD representing 28% of the population.
+This higher income column represents 39% of the population, the largest proportion of any of the three columns.
+The left column in @fig-Fig9 corresponds to the lowest income (<\$68,750 median household income), and accounts for 32% the population.
+It is dominated by A+ parcels along with A and A- parcels, in the largest quantities relative to other bar columns.
+However, proportions of all other superclusters are also present.
+The middle column is defined by parcels with a mix of supercluster classifications and represents DAs with middle households incomes (between $68,750 and 91,500) representing 28% of the population.
 
-Though median income was the most useful in partitioning the parcels by their supercluster labels in @fig-Fig9, other variables are still important.
-Namely, the following variables in order of importance and correlation coefficient with the median household income variable in brackets, are listed: the % below the median household income (`r Income_inspect[c("Income_bottom"),] |> round(digits=2)`), % single parent household (`r Income_inspect[c("One_parent_families"),] |> round(digits=2)`), % no certification or with only a highschool diploma (`r Income_inspect[c("Low_or_no_deg"),] |> round(digits=2)`), the average number of children per household (`r Income_inspect[c("Avg_numb_of_kids"),] |> round(digits=2)`), % not in the labour force (`r Income_inspect[c("Not_in_labforce"),] |> round(digits=2)`), % LICO prevalence (`r Income_inspect[c("LICO"),] |> round(digits=2)`), and % who walk to work (`r Income_inspect[c("walk_towork"),] |> round(digits=2)`).
+Though median income was the most useful in partitioning the parcels by their supercluster labels in @fig-Fig9, other variables could also be important in partitioning the data.
+Namely, the following variables in order of importance and their correlation coefficient with the median household income variable in brackets, are listed: the % below the median household income (`r Income_inspect[c("Income_bottom"),] |> round(digits=2)`), % single parent household (`r Income_inspect[c("One_parent_families"),] |> round(digits=2)`), % no certification or with only a highschool diploma (`r Income_inspect[c("Low_or_no_deg"),] |> round(digits=2)`), the average number of children per household (`r Income_inspect[c("Avg_numb_of_kids"),] |> round(digits=2)`), % not in the labour force (`r Income_inspect[c("Not_in_labforce"),] |> round(digits=2)`), % LICO prevalence (`r Income_inspect[c("LICO"),] |> round(digits=2)`), and % who walk to work (`r Income_inspect[c("walk_towork"),] |> round(digits=2)`).
 <!-- Of note, when median household income is removed from the Decision Tree, a similar structure is created, except with above average Single-Parent family and LICO prevalence splitting the data.-->
@@
 For instance, single-parent families are most positively correlated with the proportion of households being in the bottom income distribution (`r Income_inspect[c("Income_bottom"),] |> round(digits=2)`) and lower or no diploma (`r One_parent_family_inspect[c("Low_or_no"),] |> round(digits=2)`), and most negatively with the median after-tax household income (`r One_parent_family_inspect[c("v_CA21_907: Median after-tax income of household in 2020 ($)"),] |> round(digits=2)`), reflecting nationwide trends [@governmentofcanadaDailyMainHighlights2024; @governmentofcanadaPrevalenceLowIncome2024].
-Though not all single-family households are below the LICO (i.e., single-family households are not highly correlated with LICO at the DA-level in Hamilton), DAs with higher concentrations of single-family households and LICO prevalence tend to have A+ access.
-Conversely, DAs with lower single-parent households are in DAs with a higher concentration of parcels with D access.
-These findings are notable from an equity perspective since economic disadvantaged often intersects with other socio-demographic characteristics.
+Though not all single-family households are below the LICO (i.e., single-family households are not highly correlated with LICO at the DA-level in Hamilton), DAs with higher concentrations of single-family households and LICO prevalence tend to have A+ very completely caring access.
+Conversely, DAs with lower single-parent households are in DAs with a higher concentration of parcels with D not complete or caring access.
+These findings are notable from an equity perspective since economic disadvantaged often intersects with other socio-demographic characteristics [@ravensbergen_exploratory_2023; @lightmanMeasuringEconomicExclusion2018].
 
@@
 
-Geographically, Hamilton is city that offers very caring - and completely caring - 15-Minute neighbourhoods for some, but not all.
+Spatially, Hamilton is a city that offers some completely caring 15-Minute neighbourhoods for some, but not all.
 There are evident spatial inequalities, with areas ranging from excellent (A+ and A), average (B+ and B), to well below average (D).
@@
 Though our paper uses a different methodology, both studies highlight distinct patterns: indeed, A+, A or A- care access is concentrated in the downtown core but it is also present in certain pockets outside of the center of Hamilton, leading to interesting conjectures.
-For instance, some of the A-grade neighborhoods in Hamilton likely follow land-use principles that emphasized residential proximity to amenities, similar to the NUC reviewed.
+For instance, some of the A- neighborhoods likely follow land-use principles that emphasized residential proximity to amenities, similar to the NUC reviewed.
 The potential overlap of the NUC with caring 15-minute neighborhoods warrants further investigation.
@@
 There are now few low-rent choices that provide the same exceptional level of access as the downtown core, hence lower household income residents that currently reside in A+ neighhourhoods will likely displaced in the coming years if current trends continue.
+These questions of gentrification, displacement and changing 15-Minute neighbourhoods in Hamilton are topical subjects that warrant further investigation.
 
-In discussing policy interventions that equitably increase completely caring 15-Minute neighbourhoods in Hamilton, this work's presents a methodology to create city-wide relative typologies and investigate who currently resides in what neighbourhood, as a stepping stone for further investigation.
-Neighbourhoods with the lowest grades (D) and with the highest potential in being improved (i.e., high accessibiltiy for certain types of destinations but not all) are neighbourhoods with B- and A- grades.
-However, the work's findings demonstrate a mismatch in household income and completely caring access: higher income areas are often the ones targeted for improvement.
-This raises important questions from the land-use policy perspective that need to be further analysed.
+In discussing policy interventions that equitably increase completely caring 15-Minute neighbourhoods in Hamilton, this work presents a methodology to create city-wide relative typologies and investigate who currently resides in what neighbourhood, as a stepping stone for further investigation.
+Neighbourhoods with the lowest grades (D) and with the highest potential in being improved (i.e., high accessibility for certain types of destinations but not all) are neighbourhoods with B- and A- grades.
+However, our work demonstrates that higher income households tend to reside in these neighbourhoods.
+This raises important questions for land-use policy that needs further analysis.
 For instance, is it equitable to focus policy on ameliorating neighbourhoods that are already higher-rent though they tend to be more rural, single-use zoned and car-dependent (parcels with D grades)?
 Further, of the parcels that provide high child-centrics but low otherwise, A- parcels (better access) tend to be in DAs with lower-income households more so than B- parcels (lower access than A-).
-From the perspective of ameliorating land-use to support the equitable distribution of 15-minute caring neighbourhoods which areas be targeted?
-Who should benefit, and how?
+From the perspective of ameliorating land-use to support the equitable distribution of 15-minute caring neighbourhoods, which areas should be targeted?
+Who should potentially benefit, and how?
 If the policy initiative is targeted to specific neighbourhoods: sustainability linked to car dependency and equity are in tension.
-This harkens to what role a planned neighbourhood, and bottom-up planning approaches that include the evaluation of travel behaviour by socio-economic and demographic profile should be adopted to achieve equitable 15-minute cities.
+This harkens to what role a planned neighbourhood, and bottom-up planning approaches that include the evaluation of travel behaviour by socio-economic and demographic profiles along with qualitative perspectives, should play in planning for _equitable_ 15-minute cities.
 
@@
 This work measures spatial accessibility which is a measure of potential interaction with all reachable destinations from an origin.
-These destinations, however, may not be relevant to people at an origin, e.g., they may be underutilized (walkable access to 2 schools may be underused, as children often only attend 1 school), the trip physically undesirable (the walk may be along an arterial with high traffic speeds, making the trip unlikely to ever happen by foot), or the average walking speed assumed may not reflect the walking speeds of all populations [@willberg15minuteCityAll2023].
-Further, people who reside in these neighbourhoods may also disagree with the neighbourhood's label.
-The grade labels are region-relative (e.g., high accessibility in Hamilton may be subjectively insufficient for some) and they do not consider subjective perceptions that influence accessibility (e.g., though a neighbourhood has many opportunities, it may not feel safe to access them).
+These destinations, however, may not be relevant to people at an origin, e.g., they may be underutilized such as a parcel with a single-child household having 15-minute walkable access to two schools, as a child only needs to attend one school not two.
+As other examples: the trip may be physically undesirable e.g., the walk may be along an arterial with high traffic speeds, making the trip unlikely to ever happen by foot, or the average walking speed assumed may not reflect the walking speeds of all populations [@willberg15minuteCityAll2023].
+Furthermore, the SOM methodology only incorporates aspects of spatial accessibility, and people who reside in these neighbourhoods may disagree with the neighbourhood's completely caring access grade.
+The grade labels are region-relative (e.g., high accessibility in Hamilton may be subjectively insufficient for some) and they do not consider subjective perceptions that influence accessibility (e.g., though a neighbourhood has many opportunities, residences may not feel safe to access them).
 Furthermore, accessibility is calculated from the point of residential parcels.
-Care trips are not necessarily completed from home, in-fact, they are often completed in complex trip-chains [@scheinerWomenComplexDaily2015].
+Care trips are not necessarily completed from home, in fact, they are often completed in complex trip-chains [@scheinerWomenComplexDaily2015].
+Also different care destination types can be more or else important to different people, indicating an aspect of competition and destination quality could be considered within the accessibility measure itself.
 In this way, the results flatten the dynamic patterns of care trips.
 These methodological assumptions should all be taken together when interpreting the results.
-In this way, the methodologyand findings presented identifies spatial and socio-economic trends that should be further investigated.
+In this way, the methodology and findings presented identifies spatial and socio-economic variations that should be further investigated.
+They can be examined through additional quantitative investigation but also through qualitative and mixed methods that can capture trends at the individual and personal level.
 
@@
 
-This study makes three types of contributions to the transportation and city planning literature: empirical, methodological and theoretical.
+This work makes three types of contributions to the transportation and city planning literature: empirical, methodological and theoretical.
 At the empirical level, areas of the mid-sized City of Hamilton have been typified by their degree of '15-Minute Caring Neighbourhood' potential.
@@
 These values were then clustered using a novel machine learning approach, SOM, to generate meaningful typologies for discussion and further comparison with socio-economic composition of the area.
-We find A+ and A 15-minute caring neighbourhoods are located within the downtown core and in certain suburban pockets of the city, while the peripheral regions provide D level caring access.
+We find A+ and A (very completely caring 15-minute neighbourhoods) are located within the downtown core and in certain suburban pockets of the city, while the peripheral regions provide D level caring access.
 A- and B- areas are also identified as neighbourhoods that already support a high amount of children-centric destination access, and could be improved to provide better complete care access.
+Our work also demonstrates that residents of lower care access neighbourhoods tend to be wealthier than higher care access neighbourhoods highlighting the potential tensions in implementing sustainable land-use policy.
+We also detail limitations associated with our data and methodology and hence our results throughout the manuscript.
 Theoretically, our work puts forth an explicitly caring 15-Minute Neighbourhod conceptualisation, bridging the Mobility of Care and the 15-Minute City concepts.
-We discuss how measuring caring neighbourhoods can be explicitly considered within city planning.
+We discuss how measuring caring neighbourhoods can be explicitly considered within city planning, along with trends that warrant further investigation in future studies.
 
-Our work is of relevance to researchers and practitioners planning equitable and sustainable cities.
-Instead of prescribing an urban form design principle, such as "all local amenities should be within a 15-minute walking distance", it instead examines an empirical example to determine which areas in the city have the _potential_ to be 15-minute neighbourhood based on the existing spatial accessibility offered by the urban environment and walking transport network.
+This work is of relevance to researchers and practitioners planning equitable and sustainable cities.
+Instead of prescribing an urban form design principle, such as "all local amenities should be within a 15-minute walking distance", this work instead examines an empirical example to determine which areas in the city have the _potential_ to be 15-minute neighbourhood based on the existing spatial accessibility offered by the urban environment and walking transport network.
 To this end, this data-driven methodology introduces a way to identify neighbourhoods that have potential, almost have potential, and are far from containing this potential to support future context-specific qualitative work.

### 2025-01-30 paper/Manuscript.tex
R1 submitted.

@@
 \linenumbers
+\floatplacement{figure}{H}
 \makeatletter
@@
   pdfauthor={Anastasia Soukhov; Léa Ravensbergen; Lucía Mejía Dorantes; Antonio Páez},
-  pdfkeywords={Accessibility, Mobility of Care, Gender, 15-minute
-neighbourhood, Care Complete Neighbourhoods},
+  pdfkeywords={15-Minute City, Chrono-Urbanism, Accessibility, Mobility
+of Care, Inequality, Gender Gap},
   colorlinks=true,
@@
 \begin{keyword}
-    Accessibility \sep Mobility of Care \sep Gender \sep 15-minute
-neighbourhood \sep 
-    Care Complete Neighbourhoods
+    15-Minute City \sep Chrono-Urbanism \sep Accessibility \sep Mobility
+of Care \sep Inequality \sep 
+    Gender Gap
 \end{keyword}
@@
 neighbourhoods in higher-income peripheral areas. Areas that make good
-candidates for urban policy intervention are identified, and the
+candidates for urban policy intervention are identified and the
 implications of enhancing 15-minute walkable caring access are discussed
-in relation to gender mainstreaming in transportation planning.
-
-Keywords: 15-Minute City; 15-Minute Neighbourhoods; Chrono-Urbanism;
-Accessibility; Mobility of Care; Cities of Proximity; Inequality; Gender
-Gap
+in relation to gender mainstreaming in transportation planning and
+limitations of this work.
+
+Keywords: 15-Minute City; Chrono-Urbanism; Accessibility; Mobility of
+Care; Inequality; Gender Gap
 
@@
 
-From another perspective, feminist geographers, urban planners and other
+An increasing number of human geographers, planners and other
 researchers agree that urban structure influences travel, but city
@@
 gender
-\citep{lawWomenTransportNew1999, cresswell_gendered_2008, levy2013travel, little_gender_1994, tronto_toward_1990, neto_gender_2015}.
+\citep{lawWomenTransportNew1999, cresswell_gendered_2008, levy2013travel, little_gender_1994, tronto_toward_1990}.
 The causes for this are varied, and related to gendered social norms and
@@
 \citep{ravensbergen2023exploratory}. Despite care travel's importance in
-advancing gender equality -and equity more broadly-, Mobility of Care
-and access to care destinations have been under-examined relative to
-employment destinations. This is especially pertinent in accessibility
-research that has largely focused on employment points of interest e.g.,
+advancing gender equality -and equity\footnote{The United Nations
+  defines ``gender equality'' as equal rights, responsibilities, and
+  opportunities for all genders, regardless of gender at birth
+  \citep{unwomenHandbookGenderMainstreaming2022}; they prefer ``gender
+  equality,'' as ``equity'' relates to the concept of fairness that are
+  usually based on traditions, customs, or cultural beliefs which can
+  perpetuate disadvantage. However, ``equity,'' in the broader context
+  of social justice, is widely accepted. As defined in the UNICEF
+  Glossary of DEI terms, equity is the ``process of being fair to all
+  individuals and groups, by addressing present and historical
+  inequality in order to work towards equality in outcomes'' and may
+  involve temporary special measures to address systemic discrimination
+  faced by marginalized groups \citep{unicefAnnexDIVERSITYEQUITY2024}.}
+more broadly-, Mobility of Care and access to care destinations have
+been under-examined relative to employment destinations. This is
+especially pertinent in accessibility research that has largely focused
+on employment points of interest e.g.,
 \citep{farberOntarioLineSocioeconomic2019, duarteInfluenceJobAccessibility2023, ryanAccessibilitySpaceTime2023}.
@@
 
-\section{Data}\label{data}
+\section{Case study and data}\label{case-study-and-data}
 
@@
 development referred to as `the Mountain'. In this work, we analyse the
-residential parcel centroids, 143,882 locations. We aggregated the
-points at the level of Canadian Census Dissemination Area (DA) along
-with the population and population per parcel plots in
-Figure~\ref{fig-Fig1}.
+residential parcel centroids, 143,882 locations in the city
+\citep{Teranet_2021}. We aggregated the points at the level of Canadian
+Census Dissemination Area (DA) along with the population and population
+per parcel plots in Figure~\ref{fig-Fig1}. The DA is the most spatially
+granual level of publicly available Canadian census data. DAs are
+created by Statistics Canada for the purpose of the census and each
+represent between 400 to 700 people (Q1:442 and Q3:664 from
+Figure~\ref{fig-Fig1}).
 
@@
 \end{figure}%
-
-\newpage
 
@@
 
-\begin{longtable}[c]{|p{0.67in}|p{3.94in}|p{2.09in}}
+\begin{longtable}[c]{|p{1.18in}|p{3.54in}|p{1.97in}}
 
@@
 
-\multicolumn{1}{>{\raggedright}m{\dimexpr 0.67in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{Care\ category}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 3.94in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{Care\ destination\ types}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 2.09in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{Sources}}} \\
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{\global\setmainfont{Arial}{Care\ category}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 3.54in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{\global\setmainfont{Arial}{Care\ destination\ types}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 1.97in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{\global\setmainfont{Arial}{Sources}}}} \\
 
@@
 
-\multicolumn{1}{>{\raggedright}m{\dimexpr 0.67in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Depedent-centric}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 3.94in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Schools,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{daycares,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{and}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{community}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{centres,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{recreation}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{centres,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{parks,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{senior}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{centres,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{long-term}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{care}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{homes,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{and}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{retirement}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{homes:}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1,265}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{locations}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{are}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{included.}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 2.09in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(Hamilton}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2022a,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2022c,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2022d;}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Ontario}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023;}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Ontario GeoHub}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023)}}} \\
-
-\multicolumn{1}{>{\raggedright}m{\dimexpr 0.67in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Grocery-centric}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 3.94in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Convenience}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{stores}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{and}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{grocery}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{stores}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(e.g.,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{large}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{retailers}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{as}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{well}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{as}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{speciality}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{food}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{grocers,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{health}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{food}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{grocers):}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{381}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{destinations}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{are}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{included.}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 2.09in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(Axle Data}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023)}}} \\
-
-\multicolumn{1}{>{\raggedright}m{\dimexpr 0.67in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Health-centric}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 3.94in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Hospitals,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{pharmacies,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{clinics,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{and}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{dentist}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{offices:}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{421}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{destinations}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{are}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{identified.}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 2.09in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(Ontario GeoHub}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023;}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{HNHB Healthline}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023)}}} \\
-
-\multicolumn{1}{>{\raggedright}m{\dimexpr 0.67in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Errand-centric}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 3.94in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Libraries,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{post}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{offices,}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{and}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{banks:}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{158}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{destinations}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{are}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{identified.}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 2.09in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(Hamilton}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2022b;}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Axle Data}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023;}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Canada Post}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023;}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{BMO}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023;}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{HSBC}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023;}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{National Bank}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023;}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{RBC}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023;}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Scotiabank}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023;}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{TD Bank}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2023)}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{.}}} \\
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Depedent-centric}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 3.54in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Schools,}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{daycares,}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{and}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{community}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{centres,}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{recreation}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{centres,}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{parks,}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{senior}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{centres,}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{long-term}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{care}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{homes,}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{and}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{retirement}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{homes:}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1,265}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{locations}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{are}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{included.}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 1.97in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(Hamilton}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2022a,}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2023,}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2022c,}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2022d;}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Ontario}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2023;}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Ontario GeoHub}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2023)}}}} \\
+
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Grocery-centric}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 3.54in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Convenience}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{stores}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{and}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{grocery}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{stores}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(e.g.,}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{large}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{retailers}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{as}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{well}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{as}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{speciality}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{food}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{grocers,}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{health}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{food}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{grocers):}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{381}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{destinations}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{are}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{included.}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 1.97in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(Axle Data}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2023)}}}} \\
+
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Health-centric}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 3.54in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Hospitals,}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{pharmacies,}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{clinics,}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{and}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{dentist}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{offices:}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{421}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{destinations}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{are}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{identified.}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 1.97in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(Ontario GeoHub}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2023;}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{HNHB Healthline}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2023)}}}} \\
+
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.18in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Errand-centric}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 3.54in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Libraries,}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{post}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{offices,}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{and}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{banks:}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{158}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{destinations}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{are}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{identified.}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 1.97in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(Hamilton}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2022b;}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Axle Data}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2023;}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Canada Post}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2023;}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{BMO}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2023;}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{HSBC}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2023;}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{National Bank}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2023;}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{RBC}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2023;}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Scotiabank}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2023;}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{TD Bank}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2023)}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{.}}}} \\
 
@@
 
-\subsection{Accessibility: the cumulative opportunity
-measure}\label{accessibility-the-cumulative-opportunity-measure}
-
-To capture the quantity of access to each type of destinations, a
-cumulative opportunity accessibility score \(S^t_i\) is calculated.
+\subsection{The cumulative opportunity measure of
+accessibility}\label{the-cumulative-opportunity-measure-of-accessibility}
+
+To capture the quantity of spatial access to each type of destinations,
+a cumulative opportunity accessibility score \(S^t_i\) is calculated.
 Scores for each of the 14 care destination types \(t\) is calculated for
-every parcel \(i\). The calculation takes the following mathematical
-form:
-
-\[
+every parcel \(i\). The calculation takes the mathematical form in
+Equation~\ref{eq-access-cumulative}:
+
+\begin{equation}\phantomsection\label{eq-access-cumulative}{
 S^t_i=\sum_{j}O_j^{t}\cdot f(c_{ij})
-\] \noindent Where:
+}\end{equation} \noindent Where:
 
@@
   \(S_{i}^t\) is the cumulative opportunity accessibility score, the sum
-  of weighted opportunities reachable within \(f(\cdot)\), at each \(i\)
-  for each \(t\).
+  of opportunities reachable within \(f(\cdot)\), at each \(i\) for each
+  \(t\).
 \end{itemize}
@@
 
-\[
+\begin{equation}\phantomsection\label{eq-entropy-measure}{
 D_{i} = \frac{-\sum_{t} [S_i^{t} / \sum_{t}S^{t}_{i} \times ln(S_i^{t} / \sum_{t}S^{t}_{i} )]}{ln(n_t)}
-\]
+}\end{equation}
 
@@
   \(S_{i}^t\) is the cumulative opportunity accessibility score, the sum
-  of weighted opportunities reachable within a 15-minute walk from
-  \(i\).
+  of opportunities reachable within a 15-minute walk from \(i\).
 \item
@@
 
-How many care destinations can be reached within a 15-minute walk is
-summed for each of the four care categories defined in
-Table~\ref{tbl-Tbl1}. The median parcel value for each DA is visualised
+The number of care destinations that can be reached within a 15-minute
+walk is summarised for each of the four care categories (as defined in
+Table~\ref{tbl-Tbl1}). The median parcel value for each DA is visualised
 in Figure~\ref{fig-Fig4} and the median parcel diversity of care
@@
 Figure~\ref{fig-Fig4} and Figure~\ref{fig-Fig5} represent summaries of
-15 variables that served as inputs into the SOM algorithm.
+15 variables were inputs into the SOM algorithm.
 
@@
 reached within a 15-min walk per care category for a median parcel in
-each DA. The values are a summary of the 14 acessibility scores that
-served as inputs into the SOM. Basemap shapefiles are retrieved from the
-2021 Canadian census \citep{governmentofcanadaCensusPopulation2023}, the
-Open Data Hamilton Portal \citep{opendatahamiltonCityBoundary2023} and
-the USGS \citep{greatlakesUSGS2010}.}
+each DA. The values are a summary of the 14 acessibility scores that are
+inputs into the SOM. Basemap shapefiles are retrieved from the 2021
+Canadian census \citep{governmentofcanadaCensusPopulation2023}, the Open
+Data Hamilton Portal \citep{opendatahamiltonCityBoundary2023} and the
+USGS \citep{greatlakesUSGS2010}. Scale is in quartiles.}
 
@@
 Portal \citep{opendatahamiltonCityBoundary2023} and the USGS
-\citep{greatlakesUSGS2010}.}
+\citep{greatlakesUSGS2010}. Scale is in quartiles.}
 
 \end{figure}%
-
-\newpage
 
@@
 
-In Figure~\ref{fig-Fig4}, the concentration of opportunities for all
-category types is centered in the downtown of Hamilton, with the highest
-concentration being in the lower city (near the lake shore) and pockets
-of the high concentration further from the shoreline. Grocery-centric
-destinations appear to be the most concentrated, followed by
-health-centric and dependent-centric caring destinations. Errand-centric
-destinations are the most sprawled. In many ways, Figure~\ref{fig-Fig4}
-mirrors the spatial distribution of care destinations
-(Figure~\ref{fig-Fig3}) as the range of 15-minutes of travel is small.
-In Figure~\ref{fig-Fig5}, areas that have high care accessibility tend
-to have high diversity as well, though there are exceptions in pockets
-of the city outside the downtown core that have higher diversity but low
-levels of accessibility for all care categories. Similarly, there are
-areas with low diversity within the downtown core that have only
-moderate or high accessibility to certain care destination types.
-
-Based on the SOM methodology discussed, 7 superclusters are identified
-from the 10 by 10 grid of SOM nodes. As diagnostics for the selection of
-the number of superclusters, the dissimilarity-index-based dendrogram
-with the proportion of parcels represented in each supercluster
-alongside the variance explained plot is visualised in
-Figure~\ref{fig-Fig6}. Labels representing grades A to D qualifying the
-7 superclusters are then assigned. These grades as also reflected in
-Figure~\ref{fig-Fig6} and represent the quantity of accessibility and
-diversity in accessibility per care category. Higher grades (A+ and A)
-corresponding to the highest accessibility and diversity, while lowest
-grades (D) representing lowest accessibility and diversity scores.
+In Figure~\ref{fig-Fig4}, for all care categories, access to
+opportunities is concentrated in downtown Hamilton (Hamilton Central),
+with the highest concentration (top quantiles) near the lake shore of
+Hamilton Central. Grocery-centric destinations follow this trend most
+prominently, followed by health-centric and dependent-centric caring
+destinations. Access to errand-centric destinations appears to most
+spatially homogeneous of the four care destination types, though the top
+quantiles are still concentrated in Hamilton Central. As the 15-minute
+walking accessibility to destinations is a spatially weighted summary
+measure of the count of destinations, Figure~\ref{fig-Fig4} mirrors the
+spatial distribution in Figure~\ref{fig-Fig3}. In Figure~\ref{fig-Fig5},
+areas that have high diversity are broadly located in downtown Hamilton
+as well, however, there are exceptions. Namely, there are pockets of
+high diversity outside the downtown core that that score low levels of
+accessibility for all care categories. Similarly, there are areas with
+low diversity within the downtown core that have only moderate or high
+accessibility to certain care destination types.
+
+Based on the SOM methodology discussed and using inputs summarised in
+Figure~\ref{fig-Fig4} and Figure~\ref{fig-Fig5}, seven superclusters are
+identified as shown in the dissimilarity-index-based dendrogram (left
+plot of Figure~\ref{fig-Fig6}). Seven superclusters are selected to
+represent the data, as they explain a high amount of the variance (over
+90\%; right plot in Figure~\ref{fig-Fig6}). Each supercluster can be
+intuitively labelled based on the quantity of accessibility and
+diversity in accessibility per care category offered by a parcel, as
+mapped onto the SOM input grid. Labels representing grades A through D
+are assigned to qualify the seven superclusters. Higher grades (A+ and
+A) corresponding to the highest accessibility and diversity, while
+lowest grades (D) represent the lowest accessibility and diversity
+scores.
 
@@
 
-\caption{\label{fig-Fig6}The 7 resulting superclusters from the SOM the
-10 by 10 grid output represented in a dendrogram (left) and the
-proportion of variance unexplained by cluster (right).}
+\caption{\label{fig-Fig6}The seven resulting superclusters from the SOM
+output are represented in the dendrogram (left). As a diagnostic, the
+proportion of variance unexplained by the number superclusters is
+displayed on the right, with red lines indicating seven superclusters.}
 
@@
 
-\newpage
-
-The grade for each supercluster was carefully assigned by the authors by
-referring to descriptive statistics and their notable trends. These
-statistics are summarised in Table~\ref{tbl-Tbl2} and in the boxplots in
-Figure~\ref{fig-Fig7}.
+More specifically, each supercluster was labelled a grade by the authors
+by referring to the supercluster's descriptive statistics, as summarised
+in Table~\ref{tbl-Tbl2} and the boxplots in Figure~\ref{fig-Fig7}.
+Additionally, the grades for each supercluster are further explained in
+context of the proportion of urban, suburban and rural parcels as
+classified by the City of Hamilton within the parcel dataset
+\citep{Teranet_2021}. Notably, the majority of parcels in Hamilton are
+considered urban (93.4\%), while suburban and rural parcels constitute
+the remaining 2.3\% and 4.2\%, respectively:
 
@@
 \item
-  A+ and A have exceptionally high caring accessibility for all
-  destination types and high diversity scores in the top quantile or
-  above. Together, these grades represent 24\% of all parcels in the
-  city.
-\item
-  A- has high caring accessibility, but low diversity. Interestingly, A-
-  is like A but with much higher dependent-centric destinations,
-  particularly parks, schools, and daycares, with only moderately high
-  scores for all other destination types. This disbalance lowers its
-  overall diversity score. However, it can be characterized as a
-  supercluster that demonstrates potential in being retrofitted to
-  provide A+ or A level of completely caring access. 4\% of all parcels
-  are represented by the A- grade.
-\item
-  B+ and B superclusters present about average completely caring access.
-  These superclusters can serve as the benchmark for what `average'
-  completely caring access in Hamilton currently looks like. These
-  grades represent 15\% of all parcels.
+  A+ and A represent superclusters with exceptionally high caring
+  accessibility for all destination types and high diversity scores in
+  the top quantile or above. Together, these grades represent 24\% of
+  all parcels in the city and 100\% are located in urban neighbourhoods.
+  Based on the spatial access and diversity of care destination access
+  offered by these neighbourhoods, they could be called ``very caring''
+  and ``very complete''.
+\item
+  A- has high caring accessibility, but low diversity. These
+  neighbourhoods could be classified as ``very caring'' but and ``not
+  complete''. Interestingly, A- is like A but with much higher
+  dependent-centric destinations, particularly parks, schools, and
+  daycares, with only moderately high scores for all other destination
+  types. This disbalance contributes to reducing its overall diversity
+  score. However, it could be characterized as a supercluster that
+  demonstrates potential in being retrofitted to provide A+ or A level
+  of completely caring access. 4\% of all parcels are represented by the
+  A- grade and, similar to parcels in A+ and A superclusters, 100\% are
+  located in urban neighbourhoods.
+\item
+  B+ and B superclusters present about average ``completely caring''
+  access. These superclusters can serve as the benchmark for what
+  `average' completely caring access in Hamilton currently looks like.
+  These grades represent 15\% of all parcels and like parcels in A+, A
+  and A- superclusters, 100\% are located in urban neighbourhoods.
 \item
@@
   particularly parks, daycares and schools, but below average access to
-  other destination types and hence has low diversity scores. B- is like
-  A- as it demonstrates complete caring 15-minute potential if
-  retrofitted. As these parcels demonstrate caring access to some
-  destinations, they may have the potential to be retrofitted to support
-  complete access to all caring destination types. This grade represents
-  19\% of all parcels.
+  other destination types and hence has low diversity scores. B-
+  neighbourhoods could be labelled ``somewhat caring'' but ``not
+  complete''. B- is like A- as it demonstrates complete caring 15-minute
+  potential if retrofitted. As these parcels demonstrate caring access
+  to some destinations, they may have the potential to be retrofitted to
+  support complete access to all caring destination types. This grade
+  represents 19\% of all parcels. While the majority of parcels are
+  still classified as urban neighbourhoods (99\%), this supercluster
+  represents 5.9\% of the suburban parcels and 0.9\% rural parcels in
+  the city. These suburban and rural parcels are in proximity to urban
+  parcels, as elaborated in the next subsection.
 \item
@@
   room for land-use improvement that addresses complete and caring
-  15-minute access. This supercluster characterizes the largest number
-  of parcels, representing 38\% of parcels in the city.
+  15-minute access. D neighbourhoods could be classified as ``not
+  caring'' and ``not complete'' due to the low amount of access to all
+  care destinations offered and the care category diversity of this
+  access. Notably, this supercluster characterizes the largest number of
+  parcels, representing 38\% of parcels in the city. Only 83\% of these
+  parcels are `urban', and this supercluster represents the vast
+  majority of suburban and rural parcels in the city (94.1\% and
+  99.1\%).
 \end{itemize}
@@
 
-\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{\ }}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{A+}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{A}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{A-}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{B+}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{B}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{B-}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{D}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{TOTAL}}} \\
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{\global\setmainfont{Arial}{\ }}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{\global\setmainfont{Arial}{A+}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{\global\setmainfont{Arial}{A}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{\global\setmainfont{Arial}{A-}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{\global\setmainfont{Arial}{B+}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{\global\setmainfont{Arial}{B}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{\global\setmainfont{Arial}{B-}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{\global\setmainfont{Arial}{D}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{8}\selectfont{\global\setmainfont{Arial}{TOTAL}}}} \\
 
@@
 
-\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{GROCERY}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{TOTALS}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{12.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(5.7)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{7}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(3.1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{4.8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.7}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.2)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3.6}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(5.1)}}} \\
-
-\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
-
-\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Convenience}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Store}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{4.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.9)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.9)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(3.4)}}} \\
-
-\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
-
-\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Grocer}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{4.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.8)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.6}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.7}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.9)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.4)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2)}}} \\
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{GROCERY}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{TOTALS}}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{12.2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(5.7)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{7}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(3.1)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2.8}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2.6)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{4.8}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2.5)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2.7}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2.2)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.6)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.5}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.2)}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{3.6}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(5.1)}}}} \\
+
+\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
+
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Convenience}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Store}}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{8}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(4)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{4.5}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2.6)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2.1)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{3}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.9)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1.8}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.6)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.8}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.2)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.4}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.9)}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2.4}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(3.4)}}}} \\
+
+\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
+
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Grocer}}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{4.2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2.8)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2.6}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.5)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.7}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.9)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1.9}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.2)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.9}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.6)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.4)}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1.2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2)}}}} \\
 
@@
 
-\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{DEP.}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{TOTALS}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{17.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(5.8)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{11.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(4.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{13.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{6.7}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{5.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{9.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.1)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{8.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(6.2)}}} \\
-
-\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
-
-\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Comm.}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{or}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Rec.}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{centre}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.7}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.9)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.8)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.3)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.8)}}} \\
-
-\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
-
-\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Daycare}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{or}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{EarlyON}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{4.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.6}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.6}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.8)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.2)}}} \\
-
-\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
-
-\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{LTC}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{or}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{retirment}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{home}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.8)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.7)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.4)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.8)}}} \\
-
-\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
-
-\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Park}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{6.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{4.8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{5.6}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.7}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.3)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3.6}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.6)}}} \\
-
-\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
-
-\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{School}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.8)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.7}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.7)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.9)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.5)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.3)}}} \\
-
-\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
-
-\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Senior}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{centre}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.9)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.7)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.2)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.5)}}} \\
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{DEP.}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{TOTALS}}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{17.5}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(5.8)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{11.5}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(4.3)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{13.4}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(3)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{6.7}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2.4)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{5.3}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2.4)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{9.5}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2.4)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2.9}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2.1)}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{8.1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(6.2)}}}} \\
+
+\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
+
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Comm.}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{or}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Rec.}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{centre}}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1.3}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.7}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.9)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.8)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.4)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.4)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.4}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.6)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.3)}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.4}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.8)}}}} \\
+
+\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
+
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Daycare}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{or}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{EarlyON}}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{4.9}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2.2)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{3.3}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2.1)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{3.9}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2.3)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1.8}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.4)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1.6}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.5)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{3}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.5)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.6}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.8)}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2.2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2.2)}}}} \\
+
+\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
+
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{LTC}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{or}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{retirment}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{home}}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1.1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.3)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.5}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.8)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.4}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.7)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.3}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.6)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.5)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.3}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.6)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.4)}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.4}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.8)}}}} \\
+
+\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
+
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Park}}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{6.9}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(3)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{4.8}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2.3)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{5.6}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2.2)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{3.2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.6)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2.5}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.4)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{3.9}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.6)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1.7}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.3)}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{3.6}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2.6)}}}} \\
+
+\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
+
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{School}}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2.8}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.5)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2.1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.1)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1.9}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.2)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1.2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.8)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.7}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.7)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1.8}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.9)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.4}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.5)}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1.3}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.3)}}}} \\
+
+\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
+
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Senior}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{centre}}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.5}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.9)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.3}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.5)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.5}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.7)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.2)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.4)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.3)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.2)}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.5)}}}} \\
 
@@
 
-\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{HEALTH}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{TOTALS}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{11.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(6.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{5.8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.8)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3.6}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.8)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.8)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.9)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.6}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(4.7)}}} \\
-
-\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
-
-\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Dentist}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{4.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(3.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.7)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.4)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.2)}}} \\
-
-\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
-
-\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Hospital}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{or}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{clinic}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.6}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.7)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.7)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.2)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.5)}}} \\
-
-\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
-
-\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Pharmacy}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{6.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(3.7)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.7}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.8)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2.7)}}} \\
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{HEALTH}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{TOTALS}}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{11.1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(6.3)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{5.8}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2.8)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{3.6}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.8)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{3.9}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.8)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2.9}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.9)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1.5}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.4)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.6}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1)}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{3.4}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(4.7)}}}} \\
+
+\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
+
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Dentist}}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{4.1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(3.4)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2.2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.8}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.1)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1.4}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.4)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.9}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.1)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.3}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.7)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.4)}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1.1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2.2)}}}} \\
+
+\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
+
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Hospital}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{or}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{clinic}}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.6}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.7)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.3}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.5)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.3}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.6)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.5)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.3}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.7)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.3)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.2)}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.5)}}}} \\
+
+\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
+
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Pharmacy}}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{6.4}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(3.7)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{3.3}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.5)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2.5}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.3)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2.3}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.2)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1.7}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.2)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1.2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.2)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.4}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.8)}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2.1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2.7)}}}} \\
 
@@
 
-\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{ERRAND}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{TOTALS}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{3.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.6)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.4)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.8)}}} \\
-
-\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
-
-\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Bank}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{or}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{ATM}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{2.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.8)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.7}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.7)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.2)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.7}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(1.2)}}} \\
-
-\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
-
-\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Library}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.2)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.4)}}} \\
-
-\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
-
-\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Post}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{office}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{1.2}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.8)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.7}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.7)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.5}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.8)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.6}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.7)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.5)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.1}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.2)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.4}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.7)}}} \\
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{ERRAND}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{TOTALS}}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{3.9}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(2)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2.5}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.4)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1.1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2.2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.6)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.3}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.6)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.4)}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1.2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.8)}}}} \\
+
+\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
+
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Bank}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{or}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{ATM}}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{2.2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.8)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1.5}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.2)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.4}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.5)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1.5}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.1)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.7}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.7)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.3)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.2)}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.7}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(1.2)}}}} \\
+
+\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
+
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Library}}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.5}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.5)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.3}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.5)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.4)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.2)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.2)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.3)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.2)}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.4)}}}} \\
+
+\noalign{\global\arrayrulewidth 0pt}\arrayrulecolor[HTML]{000000}
+
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Post}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{office}}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{1.2}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.8)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.7}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.7)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.5}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.8)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.6}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.7)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.3}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.5)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.1}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.3)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.2)}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.4}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.7)}}}} \\
 
@@
 
-\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{Category}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{diversity}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.8}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.2)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.3)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.9}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.6}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.4)}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.1)}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0)}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{0.3}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\ }}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{(0.4)}}} \\
+\multicolumn{1}{>{\raggedright}m{\dimexpr 1.2in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{Category}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{diversity}}}} & \multicolumn{1}{>{\cellcolor[HTML]{48B376}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.9}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.1)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{91CF60}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.8}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.2)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D9EF8B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.3}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.3)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FFFFBF}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.9}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.1)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FEE08B}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.6}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.4)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{FDAE61}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.1)}}}} & \multicolumn{1}{>{\cellcolor[HTML]{D75C56}\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0)}}}} & \multicolumn{1}{>{\raggedright}m{\dimexpr 0.6in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{0.3}}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{\ }}}\textcolor[HTML]{000000}{\fontsize{8}{0}\selectfont{\global\setmainfont{Arial}{(0.4)}}}} \\
 
@@
 
-\newpage
-
 To spatially demonstrate where the superclusters are located,
-representative supercluster labels for each DA are visualised in
+representative supercluster grades for each DA are visualised in
 Figure~\ref{fig-Fig8}. This visualisation is created by grouping parcels
-by their DA, calculating the median supercluster label for each label,
-and selecting the label that is most dominant within that DA. For
-reference, the median number of parcels in each DA is 150 and
-supercluster label membership within a DA is typically pure, with the
+by their DA and selecting the grade that is most dominant within that
+DA. For reference, the median number of parcels in each DA is 150 and
+supercluster grade membership within a DA is typically pure, with the
 median membership being 80\% of a single cluster.
@@
 
-\includegraphics[width=4.6875in,height=4.6875in]{../figures/SClabel_plot.png}
+\includegraphics[width=8.33333in,height=4.6875in]{../figures/SClabel_plot.png}
 
@@
 
-The resulting superclusters visualised in Figure~\ref{fig-Fig8} are a
-combination of the cumulative opportunity and diversity of access
-measures translated into meaningful typologies through the SOM
-methodology. To reiterate, accessibility is the measure of
-\emph{potential} interaction, e.g., how many care destinations one could
-reach within a 15-minute walk, these typologies are useful in
-identifying which areas of the city are providing regionally high levels
-of complete caring access, which areas are regionally average, and which
-areas are regionally below average. Furthermore, neighbourhoods which
-provide disbalanced access can be further investigated.
+Recall, the superclusters in Figure~\ref{fig-Fig8} are a combination of
+the cumulative opportunity and diversity of access measures translated
+into meaningful typologies through the SOM methodology. To reiterate,
+accessibility is the measure of \emph{potential} interaction, e.g., how
+many care destinations one could reach within a 15-minute walk, these
+typologies are useful in identifying which areas of the city are
+providing city-wide high levels of complete caring access, which areas
+are city-wide average, and which areas are city-wide below average.
 
@@
 city and closest to the shoreline in the downtown core of Hamilton. The
-above and below average neighbourhoods (B+ B) are often proximate to the
-excellent and exceptional neighbourhoods and are more prevalent in south
-of the escarpment within the center Hamilton. The escarpment is a
+above and below average neighbourhoods (B+, B) are often proximate to
+the excellent and exceptional neighbourhoods and are more prevalent in
+south of the escarpment within the center Hamilton. The escarpment is a
 physical barrier, with few pedestrian-accessible access points to
@@
 where urban form is characterised by lower density residential housing
-and auto-dependent mobility.
+and auto-dependent mobility (majority rural and suburban). D
+neighbourhoods also tend to have the lowest diversity scores, with
+either no access or access to only one type of care destination
+category.
 
@@
 grade-group but below grade-group average access in other destination
-types. These neighbourhoods may be more suitable for populations who
-prioritize walkable access to destinations like schools, parks and
-daycares, and find access to other types of caring destinations less
-important. These neighbourhoods also stand out as demonstrating
-potential to be retrofitted to provide more complete access if
-additional destination types were located within their neighbourhoods.
+types (e.g., low or lowest diversity scores). These neighbourhoods may
+be more suitable for populations who prioritize walkable access to
+children-centric destinations like schools, parks and daycares, and find
+access to other types of caring destinations less important. These
+neighbourhoods also stand out as demonstrating potential to be
+retrofitted to provide more complete access if additional destination
+types were located within their neighbourhoods.
 
@@
 median: 80,000). The notable splits for income are \textgreater\$91,500,
-\textgreater\$68,750, or between \$68,750 and \$91,500; roughly split by
-lower income, middle income, and higher income median household
+\textless\$68,750, or between \$68,750 and \$91,500; roughly split by
+higher income, lower income, and middle income median household
 brackets.
@@
 Describing Figure~\ref{fig-Fig9}, the bar column on the right represents
-DAs with higher household income. In these DAs, the majority of parcels
-are labeled as D superclusters, i.e not caring or complete 15-minute
-neighbourhoods. Within this higher-income-representing bar column, the
-second highest proportion of supercluster membership are B- parcels. B-
-parcels are low in most amenity types except a few child-centric
-destinations and show promise in being more easily improved than the D
-parcels. The higher income column represents 39\% of the population, the
-largest proportion of any of the three columns. The left column in
-Figure~\ref{fig-Fig9} corresponds to the lowest income, i.e., below
-68,750 CAD median household income, and accounts for 32\% the
-population. It is dominated by A+ parcels along with A and A- parcels,
-in the largest quantities relative to other bar columns, however,
-proportions of all other superclusters are also present. The middle
-column is defined by parcels with a mix of supercluster classifications
-and represents DAs with middle households incomes, between 68,750 CAD
-and 91,500 CAD representing 28\% of the population.
+DAs with higher household income (\textgreater\$91,500 median household
+income). In these DAs, the majority of parcels are labeled as D
+superclusters, i.e not caring or complete 15-minute neighbourhoods.
+Within this higher-income-representing bar column, the second highest
+proportion of supercluster membership are B- parcels. B- parcels are low
+in most amenity types except a few child-centric destinations and show
+promise in being more easily improved than the D parcels. This higher
+income column represents 39\% of the population, the largest proportion
+of any of the three columns. The left column in Figure~\ref{fig-Fig9}
+corresponds to the lowest income (\textless\$68,750 median household
+income), and accounts for 32\% the population. It is dominated by A+
+parcels along with A and A- parcels, in the largest quantities relative
+to other bar columns. However, proportions of all other superclusters
+are also present. The middle column is defined by parcels with a mix of
+supercluster classifications and represents DAs with middle households
+incomes (between \$68,750 and 91,500) representing 28\% of the
+population.
 
 Though median income was the most useful in partitioning the parcels by
-their supercluster labels in Figure~\ref{fig-Fig9}, other variables are
-still important. Namely, the following variables in order of importance
-and correlation coefficient with the median household income variable in
-brackets, are listed: the \% below the median household income (-0.89),
-\% single parent household (-0.57), \% no certification or with only a
-highschool diploma (-0.49), the average number of children per household
-(0.38), \% not in the labour force (-0.41), \% LICO prevalence (-0.67),
-and \% who walk to work (-0.37).
+their supercluster labels in Figure~\ref{fig-Fig9}, other variables
+could also be important in partitioning the data. Namely, the following
+variables in order of importance and their correlation coefficient with
+the median household income variable in brackets, are listed: the \%
+below the median household income (-0.89), \% single parent household
+(-0.57), \% no certification or with only a highschool diploma (-0.49),
+the average number of children per household (0.38), \% not in the
+labour force (-0.41), \% LICO prevalence (-0.67), and \% who walk to
+work (-0.37).
 
@@
 DA-level in Hamilton), DAs with higher concentrations of single-family
-households and LICO prevalence tend to have A+ access. Conversely, DAs
-with lower single-parent households are in DAs with a higher
-concentration of parcels with D access. These findings are notable from
-an equity perspective since economic disadvantaged often intersects with
-other socio-demographic characteristics.
+households and LICO prevalence tend to have A+ very completely caring
+access. Conversely, DAs with lower single-parent households are in DAs
+with a higher concentration of parcels with D not complete or caring
+access. These findings are notable from an equity perspective since
+economic disadvantaged often intersects with other socio-demographic
+characteristics
+\citep{ravensbergen_exploratory_2023, lightmanMeasuringEconomicExclusion2018}.
 
@@
 
-Geographically, Hamilton is city that offers very caring - and
-completely caring - 15-Minute neighbourhoods for some, but not all.
-There are evident spatial inequalities, with areas ranging from
-excellent (A+ and A), average (B+ and B), to well below average (D).
-Some areas, like those labelled A- and B-, show potential for
-improvement within their grade groups. While the downtown core has the
-highest concentration of \emph{caring 15-minute neighborhoods}, certain
-areas outside the core do as well. This finding somewhat contrasts
-employment accessibility studies, such as
-\citet{elgeneidyNonstopEquityAssessing2016}'s assessment of public
-transit access to employment in the Greater Toronto Area (including
-Hamilton), where access is heavily concentrated in the downtown core,
-more so than appears to be present in this work. Though our paper uses a
-different methodology, both studies highlight distinct patterns: indeed,
-A+, A or A- care access is concentrated in the downtown core but it is
-also present in certain pockets outside of the center of Hamilton,
-leading to interesting conjectures. For instance, some of the A-grade
-neighborhoods in Hamilton likely follow land-use principles that
+Spatially, Hamilton is a city that offers some completely caring
+15-Minute neighbourhoods for some, but not all. There are evident
+spatial inequalities, with areas ranging from excellent (A+ and A),
+average (B+ and B), to well below average (D). Some areas, like those
+labelled A- and B-, show potential for improvement within their grade
+groups. While the downtown core has the highest concentration of
+\emph{caring 15-minute neighborhoods}, certain areas outside the core do
+as well. This finding somewhat contrasts employment accessibility
+studies, such as \citet{elgeneidyNonstopEquityAssessing2016}'s
+assessment of public transit access to employment in the Greater Toronto
+Area (including Hamilton), where access is heavily concentrated in the
+downtown core, more so than appears to be present in this work. Though
+our paper uses a different methodology, both studies highlight distinct
+patterns: indeed, A+, A or A- care access is concentrated in the
+downtown core but it is also present in certain pockets outside of the
+center of Hamilton, leading to interesting conjectures. For instance,
+some of the A- neighborhoods likely follow land-use principles that
 emphasized residential proximity to amenities, similar to the NUC
@@
 reside in A+ neighhourhoods will likely displaced in the coming years if
-current trends continue.
+current trends continue. These questions of gentrification, displacement
+and changing 15-Minute neighbourhoods in Hamilton are topical subjects
+that warrant further investigation.
 
 In discussing policy interventions that equitably increase completely
-caring 15-Minute neighbourhoods in Hamilton, this work's presents a
+caring 15-Minute neighbourhoods in Hamilton, this work presents a
 methodology to create city-wide relative typologies and investigate who
@@
 investigation. Neighbourhoods with the lowest grades (D) and with the
-highest potential in being improved (i.e., high accessibiltiy for
+highest potential in being improved (i.e., high accessibility for
 certain types of destinations but not all) are neighbourhoods with B-
-and A- grades. However, the work's findings demonstrate a mismatch in
-household income and completely caring access: higher income areas are
-often the ones targeted for improvement. This raises important questions
-from the land-use policy perspective that need to be further analysed.
-For instance, is it equitable to focus policy on ameliorating
-neighbourhoods that are already higher-rent though they tend to be more
-rural, single-use zoned and car-dependent (parcels with D grades)?
-Further, of the parcels that provide high child-centrics but low
-otherwise, A- parcels (better access) tend to be in DAs with
-lower-income households more so than B- parcels (lower access than A-).
-From the perspective of ameliorating land-use to support the equitable
-distribution of 15-minute caring neighbourhoods which areas be targeted?
-Who should benefit, and how? If the policy initiative is targeted to
-specific neighbourhoods: sustainability linked to car dependency and
-equity are in tension. This harkens to what role a planned
-neighbourhood, and bottom-up planning approaches that include the
-evaluation of travel behaviour by socio-economic and demographic profile
-should be adopted to achieve equitable 15-minute cities.
+and A- grades. However, our work demonstrates that higher income
+households tend to reside in these neighbourhoods. This raises important
+questions for land-use policy that needs further analysis. For instance,
+is it equitable to focus policy on ameliorating neighbourhoods that are
+already higher-rent though they tend to be more rural, single-use zoned
+and car-dependent (parcels with D grades)? Further, of the parcels that
+provide high child-centrics but low otherwise, A- parcels (better
+access) tend to be in DAs with lower-income households more so than B-
+parcels (lower access than A-). From the perspective of ameliorating
+land-use to support the equitable distribution of 15-minute caring
+neighbourhoods, which areas should be targeted? Who should potentially
+benefit, and how? If the policy initiative is targeted to specific
+neighbourhoods: sustainability linked to car dependency and equity are
+in tension. This harkens to what role a planned neighbourhood, and
+bottom-up planning approaches that include the evaluation of travel
+behaviour by socio-economic and demographic profiles along with
+qualitative perspectives, should play in planning for \emph{equitable}
+15-minute cities.
 
@@
 not be relevant to people at an origin, e.g., they may be underutilized
-(walkable access to 2 schools may be underused, as children often only
-attend 1 school), the trip physically undesirable (the walk may be along
-an arterial with high traffic speeds, making the trip unlikely to ever
-happen by foot), or the average walking speed assumed may not reflect
-the walking speeds of all populations
-\citep{willberg15minuteCityAll2023}. Further, people who reside in these
-neighbourhoods may also disagree with the neighbourhood's label. The
-grade labels are region-relative (e.g., high accessibility in Hamilton
-may be subjectively insufficient for some) and they do not consider
-subjective perceptions that influence accessibility (e.g., though a
-neighbourhood has many opportunities, it may not feel safe to access
-them). Furthermore, accessibility is calculated from the point of
-residential parcels. Care trips are not necessarily completed from home,
-in-fact, they are often completed in complex trip-chains
-\citep{scheinerWomenComplexDaily2015}. In this way, the results flatten
-the dynamic patterns of care trips. These methodological assumptions
-should all be taken together when interpreting the results. In this way,
-the methodologyand findings presented identifies spatial and
-socio-economic trends that should be further investigated.
+such as a parcel with a single-child household having 15-minute walkable
+access to two schools, as a child only needs to attend one school not
+two. As other examples: the trip may be physically undesirable e.g., the
+walk may be along an arterial with high traffic speeds, making the trip
+unlikely to ever happen by foot, or the average walking speed assumed
+may not reflect the walking speeds of all populations
+\citep{willberg15minuteCityAll2023}. Furthermore, the SOM methodology
+only incorporates aspects of spatial accessibility, and people who
+reside in these neighbourhoods may disagree with the neighbourhood's
+completely caring access grade. The grade labels are region-relative
+(e.g., high accessibility in Hamilton may be subjectively insufficient
+for some) and they do not consider subjective perceptions that influence
+accessibility (e.g., though a neighbourhood has many opportunities,
+residences may not feel safe to access them). Furthermore, accessibility
+is calculated from the point of residential parcels. Care trips are not
+necessarily completed from home, in fact, they are often completed in
+complex trip-chains \citep{scheinerWomenComplexDaily2015}. Also
+different care destination types can be more or else important to
+different people, indicating an aspect of competition and destination
+quality could be considered within the accessibility measure itself. In
+this way, the results flatten the dynamic patterns of care trips. These
+methodological assumptions should all be taken together when
+interpreting the results. In this way, the methodology and findings
+presented identifies spatial and socio-economic variations that should
+be further investigated. They can be examined through additional
+quantitative investigation but also through qualitative and mixed
+methods that can capture trends at the individual and personal level.
 
@@
 
-This study makes three types of contributions to the transportation and
+This work makes three types of contributions to the transportation and
 city planning literature: empirical, methodological and theoretical. At
@@
 for discussion and further comparison with socio-economic composition of
-the area. We find A+ and A 15-minute caring neighbourhoods are located
-within the downtown core and in certain suburban pockets of the city,
-while the peripheral regions provide D level caring access. A- and B-
-areas are also identified as neighbourhoods that already support a high
-amount of children-centric destination access, and could be improved to
-provide better complete care access. Theoretically, our work puts forth
-an explicitly caring 15-Minute Neighbourhod conceptualisation, bridging
-the Mobility of Care and the 15-Minute City concepts. We discuss how
-measuring caring neighbourhoods can be explicitly considered within city
-planning.
-
-Our work is of relevance to researchers and practitioners planning
+the area. We find A+ and A (very completely caring 15-minute
+neighbourhoods) are located within the downtown core and in certain
+suburban pockets of the city, while the peripheral regions provide D
+level caring access. A- and B- areas are also identified as
+neighbourhoods that already support a high amount of children-centric
+destination access, and could be improved to provide better complete
+care access. Our work also demonstrates that residents of lower care
+access neighbourhoods tend to be wealthier than higher care access
+neighbourhoods highlighting the potential tensions in implementing
+sustainable land-use policy. We also detail limitations associated with
+our data and methodology and hence our results throughout the
+manuscript. Theoretically, our work puts forth an explicitly caring
+15-Minute Neighbourhod conceptualisation, bridging the Mobility of Care
+and the 15-Minute City concepts. We discuss how measuring caring
+neighbourhoods can be explicitly considered within city planning, along
+with trends that warrant further investigation in future studies.
+
+This work is of relevance to researchers and practitioners planning
 equitable and sustainable cities. Instead of prescribing an urban form
 design principle, such as ``all local amenities should be within a
-15-minute walking distance'', it instead examines an empirical example
-to determine which areas in the city have the \emph{potential} to be
-15-minute neighbourhood based on the existing spatial accessibility
-offered by the urban environment and walking transport network. To this
-end, this data-driven methodology introduces a way to identify
-neighbourhoods that have potential, almost have potential, and are far
-from containing this potential to support future context-specific
-qualitative work.
+15-minute walking distance'', this work instead examines an empirical
+example to determine which areas in the city have the \emph{potential}
+to be 15-minute neighbourhood based on the existing spatial
+accessibility offered by the urban environment and walking transport
+network. To this end, this data-driven methodology introduces a way to
+identify neighbourhoods that have potential, almost have potential, and
+are far from containing this potential to support future
+context-specific qualitative work.
 
